### 2021-09-02 Code/Draft_ruddy_duck_productivity.Rmd
figures prep.

---
lang: en-EN
bibliography: ../../../../Biblio/JabRef/Library_OFB/Library_OFB.bib
csl: ../../../../Biblio/JabRef/Library_OFB/Styles/ecology.csl
title: "Productivity estimation in waterfowl using a non-invasive method"
author: "Adrien"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    latex_engine: pdflatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      out.width = "70%", 
                      fig.align = "center") # , fig.pos = 'H' pour tenir la position

Sys.setlocale("LC_ALL", "English")

```

\newpage

# Abstract {-}

The response of a waterfowl population to a harvest pressure depends on its capacity to renew. The recruitment, i.e. the number of young adults reproducing for the first time, is a key to describe the renewal of a population. The productivity, i.e. the number of recruits per breeder, is even more informative because it is independant of the population size and allows a comparison over time. The proportion of young adults in a waterfowl population is often estimated from game-hunting samples. However, this proportion, which is a main step to estimate the productivity, is not accessible on the years without harvest, or with a low harvest rate. The age-structure in the harvest samples, also called the hunting bag, does not necessary reflect the underlying age-structure of the population. It is often skewed towards juveniles and might lead to an overestimation of the productivity. In waterfowl, the adult males displays often brighter colors than juveniles and females. This dichromatism can be easily characterized and monitored from count surveys. Such information can be used to estimate the proportion of young adults, and consequently the productivity. In using two populations of ruddy duck, this study develops a bayesian method to estimate the productivity from count surveys. To judge the accuracy of this approach, the results are compared to productivity estimates from samples. Since the adult survival is commonly estimated in the litterature and can be estimated from the counting and the sampling methods, the consistency of the two approaches is discussed.

# Introduction

breeding/reproduction success good but not enough because variable survival of juveniles

Introduction:

productivity often defined as Jt/At
but more realistic to defined as Jt/At-1

The carrying capacity of a habitat determines the maximum number of individuals of a species. When a population has occupied a habitat for a long while, its size varies around the carrying capacity because the lack of space and ressources induce a lower reproduction success and a higher youth mortality. The productivity of such population is thus dampened by density dependant effects. When considering a population colonising a new habitat, its size increases close to its maximum growth rate because the realized productivity approaches its maximum potential. Such populations are thus a good model to estimate this maximum productivity rate. This is the case of the alien species colonizing new territories. These species are often introduced by a human release and its presence causes some se species are often arrived because of undesired released

In Europe, 
a population following its maximum growth rate is a perfect study model

productivity is more closely related to recruitment, which defined with the adult survival the population growth rate and consequently the evolution of the population size and the maximum sustainable catch rate
we defined productivity as the number of recruits per their breeders

This parameter is complex because it is composed of two sub-parameters, the reproduction success, and the youth survival.

Nichols -> il faut compter de la façon dont on demande

idée: estimer l'age moyen de la pop grâce la vulnerabilité et le sexe ratio chez les adultes

# Materials & methods

develop the counting method :

2 ways to check the method :

sampling method?
Ok but few data as poorly harvested

realistic survival? 
if > 1, estimated productivity is too low to support the maximum growth rate
in the range of similar species in litterature -> satisfying

blabla

# Results

blabla

# Discussion

blabla

# References {-}

<div id="refs"></div>

### 2021-09-02 Code/Model_productivity.R

@@ -140,7 +140,7 @@ frag %>%
   scale_color_manual(values = c_pop, guide = "none") +
   scale_y_percent +
   scale_x_date_own(5e-2) +
-  guides(size = guide_legend(title = "Total counts")) +
+  guides(size = guide_legend(title = "Samples")) +
   labs(y = "Male proportion")
 
 read_rds("../Ruddy_duck_data/Data/Ruddy_duck_data.rds") %>% pluck(2) %>% 
@@ -665,3 +665,4 @@ JuvOut4 %>%
   scale_color_manual(values = c_met, guide = "none") +
 labs(title = "Adult survival rate",
        y = NULL, x = NULL)
+

### 2021-09-02 Code/Draft_ruddy_duck_productivity.Rmd
Figure preparation

@@
 title: "Productivity estimation in waterfowl using a non-invasive method"
-author: "Adrien"
+author: 
+- Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr - contribution in data analyses & writing]
+- Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk - contribution in data collection]
+- Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr - contribution in data collection]
+- Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr - contribution in writing]
+
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
                       message = FALSE, 
-                      out.width = "70%", 
+                      out.width = "100%", 
                       fig.align = "center") # , fig.pos = 'H' pour tenir la position
@@
 The recruitment, i.e.
-the number of young adults reproducing for the first time, is a key to describe the renewal of a population.
+the number of young adults reproducing for the first time, is a key indicator to describe the renewal of a population, and therefore an essential tool in population management.
 The productivity, i.e.
-the number of recruits per breeder, is even more informative because it is independant of the population size and allows a comparison over time.
+the number of recruits per breeder, is even more informative because it is independent of the population size and allows a comparison over time and between species.
 The proportion of young adults in a waterfowl population is often estimated from game-hunting samples.
-However, this proportion, which is a main step to estimate the productivity, is not accessible on the years without harvest, or with a low harvest rate.
-The age-structure in the harvest samples, also called the hunting bag, does not necessary reflect the underlying age-structure of the population.
-It is often skewed towards juveniles and might lead to an overestimation of the productivity.
-In waterfowl, the adult males displays often brighter colors than juveniles and females.
-This dichromatism can be easily characterized and monitored from count surveys.
-Such information can be used to estimate the proportion of young adults, and consequently the productivity.
-In using two populations of ruddy duck, this study develops a bayesian method to estimate the productivity from count surveys.
+However, this proportion, which is a main step in estimating productivity, is not accessible in years without harvest, or with a low harvest rate.
+Moreover, the age-structure in the harvest samples, commonly called hunting bag, does not necessarily reflect the underlying age-structure of the population.
+It is often skewed towards juveniles and can lead to an overestimation of the productivity.
+In waterfowl, adult males usually display brighter colours than juveniles and females.
+This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
+Such information can be used to infer the proportion of young adults, and consequently the productivity.
+In using two populations of ruddy duck, this study develops a Bayesian method to estimate the productivity from count data.
 To judge the accuracy of this approach, the results are compared to productivity estimates from samples.
-Since the adult survival is commonly estimated in the litterature and can be estimated from the counting and the sampling methods, the consistency of the two approaches is discussed.
+The adult survival can be estimated from both the counting and the sampling methods and is commonly found in the literature.
+The consistency of the two approaches is thus discussed by comparing the survival estimates with the literature values.
 
 # Introduction
+
+## The use of productivity in population management
+
+The management of an exploited population generally aims at...
+
+The productivity of a population is the average number of recruits produced per breeder.
+If one considers a species reaching the sexual maturity as early as its first year, all the population is mature just before the reproduction season.
+The productivity is thus the ratio of the number of young adults at the reproduction season the year *n* on the number of all adults at the reproduction season the year *n - 1*.
+
+The productivity is closely related to recruitment. When combined with the adult survival, the recruitment determines the population growth rate and consequently the evolution of the population size.
+
+This parameter is complex because it is composed of two sub-parameters: the reproduction success and the youth survival.
+
+## The variability of the productivity
+
+A waterfowl species that is released in a favourable habitat will naturally expand (malthus).
+The growth of such a population can be explained by its productivity parameter that is higher than its adult mortality.
+This expansion is limited by the carrying capacity of the habitat (ref).
+When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity without because the individuals compete for space or/and food (ref).
+This competition commonly induces a lower reproduction success or/and a higher youth mortality (ref).
+The productivity then decreases because it is dampened by density-dependent effects (holt).
+For an endemic species evolving in stable environmental conditions, the productivity and the adult mortality are balanced, which explains the stabilization of the number of individuals.
+
+Hunted waterfowl species can experience high harvest rate.
+Such high harvest pressure increases adult and juvenile mortality (ref).
+Paradoxically, this high juvenile mortality does not induce a decrease of productivity (ref).
+Indeed, one expects that a population with a high adult mortality combined to a low productivity would rapidly extinct.
+However, hunting waterfowl has a long history, and most of these species have persisted over time.
+The reason is that high harvest pressure reduces the competition for space and food.
+The density-dependent effects affecting the productivity in case of high population density are dampened by the harvest mortality.
+The exploitation of a population thus increases adult mortality and productivity.
+These two parameters can reach a new equilibrium if the population is exploited.
+The size of a newly exploited population in stable environmental conditions should thus reach a new equilibrium, which is expected to be lower than the size without exploitation.
+
+Productivity is thus a varying parameter that reaches its highest values when no density-dependent effects occurs.
+Exploring the gradient of productivity, and defining the maximum productivity, and consequently the response of a population to a harvest pressure...
+
+## Is the variability of the productivity huge or not? What part of the productivity gradient should we know for population management?
+
+## What is the ideal biological model to explore the gradient of the productivity?
+
+Hunting waterfowl has a long history, and few species have observed long-term moratorium
+Alien species colonizing new territories and then heavily controled are a good model to overcome density-dependent effects and explore the gradient productivity.
+These species are often introduced during unvoluntary release events.
+and its presence causes some se species are often arrived because of undesired released
+
+In Europe, 
+a population following its maximum growth rate is a perfect study model
+
+## What are the limitations of the sampling method to estimate the productivity?
+
+- We cannot explore the productivity when there is no harvest, however, it is crucial to explore what happens when there is no harvest in a scenario where we suggest a moratorium to let the population recover.
+
+- Biais because hunting like any predation action is selective.
+Often towards the weakest individuals.
+So overstimation of productivity, and potentially maximum sustainable harvest rate overestimated and consequently depletion of the managed population.
+
+## Why counting data is a potential tool to estimate the productivity?
+
+Long time series, easy access, no need of samples
+
+## What is the methodology of this study?
+
+Nichols -> il faut compter de la façon dont on demande
+
+idée: estimer l'age moyen de la pop grâce la vulnerabilité et le sexe ratio chez les adultes
 
@@
 but more realistic to defined as Jt/At-1
-
-The carrying capacity of a habitat determines the maximum number of individuals of a species.
-When a population has occupied a habitat for a long while, its size varies around the carrying capacity because the lack of space and ressources induce a lower reproduction success and a higher youth mortality.
-The productivity of such population is thus dampened by density dependant effects.
-When considering a population colonising a new habitat, its size increases close to its maximum growth rate because the realized productivity approaches its maximum potential.
-Such populations are thus a good model to estimate this maximum productivity rate.
-This is the case of the alien species colonizing new territories.
-These species are often introduced by a human release and its presence causes some se species are often arrived because of undesired released
-
-In Europe, 
-a population following its maximum growth rate is a perfect study model
-
-productivity is more closely related to recruitment, which defined with the adult survival the population growth rate and consequently the evolution of the population size and the maximum sustainable catch rate
-we defined productivity as the number of recruits per their breeders
-
-This parameter is complex because it is composed of two sub-parameters, the reproduction success, and the youth survival.
-
-Nichols -> il faut compter de la façon dont on demande
-
-idée: estimer l'age moyen de la pop grâce la vulnerabilité et le sexe ratio chez les adultes
 
@@
 
+Nichols -> il faut compter de la façon dont on demande
+
+# Aknowledgements {-}
+
+This work was partly funded by the LIFE Oxyura project (LIFE17 NAT/FR/000942) through the LIFE program.
+This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the Office Français de la Biodiversité, and Jean-Marc Gillier from the Société Nationale pour la Protection de la Nature.
+The authors aknowledge all the contributors of the data collection in the UK and in France, especially: **UK names**, Vincent Fontaine, Denis Lacourpaille, Jules Joly, Justin Potier, Valentin Boniface, Alexis Laroche, and Médéric Lortion, **other names**.
+
 # References {-}

### 2021-09-02 Code/Model_productivity.R

@@ -16,7 +16,7 @@ require(tls) # total least squares regression
 
 tibble(id = 1:20) %>% 
   ggplot(aes(x = 1, y = id, color = factor(id))) +
-  geom_point(size = 10) + scale_y_continuous(aes(breaks = id)) -> p ; p
+  geom_point(size = 10) + scale_y_continuous(aes(breaks = id)) -> p; p
 
 ggplot_build(p)$data
 
@@ -50,7 +50,7 @@ raw_plot <- function(para,
   JuvOut4 %>%
     filter(par == as_name(para)) %>% 
     ggplot(aes(x = year, y = `50%`, color = method, fill = method)) +
-    facet_wrap(~ pop, nrow = 1, scales = "free_x") +
+    facet_wrap(~ sub_pop, nrow = 1, scales = "free_x") +
     geom_line(linetype = "dashed", alpha = .5) +
     geom_point() +
     geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5) +
@@ -194,7 +194,6 @@ frag %>%
 
 # Lambda max formatting ------------------------------------------------------------
 
-
 # estimatation of lambda_max to test productivity reliability
 # when ruddy ducks start to be permanently seen?
 counts_1 %>% 
@@ -302,7 +301,7 @@ counts_1 %>%
   scale_shape_manual(values = c(16, 16, 16), guide = "none") +
   scale_alpha_manual(values = c(0.5, 1, 1), guide = "none") +
   scale_color_manual(values = c_pop, guide = "none") +
-  labs(y = "Population size") -> raw ; raw
+  labs(y = "Population size") -> raw_pop; raw_pop
 
 # Combined dataset ------------------------------------------------------------------
 
@@ -585,33 +584,25 @@ bind_rows(JuvOut3, survival_avg) -> JuvOut4
 
 # Output plot -----------------------------------------------------------------------
 
-raw_plot(p_juv, title = "Juvenile proportion", percent = TRUE)
-raw_plot(productivity, title = "Productivity : recruits per breeder") + 
-  scale_y_continuous(breaks = 0:9 / 10)
-raw_plot(recruits, title = "Recruit number") + 
-  facet_wrap(~ pop, nrow = 1, scales = "free")
-
+raw_plot(p_juv, title = "Proportion of recruits in the population", 
+         percent = TRUE) -> raw_prop; raw_prop
[…]

### 2021-09-06 Code/Draft_ruddy_duck_productivity.Rmd
Save

@@
 - Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr - contribution in data analyses & writing]
+- Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr - contribution in writing]
 - Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk - contribution in data collection]
 - Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr - contribution in data collection]
-- Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr - contribution in writing]
+- Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthieu.guillemain@ofb.gouv.fr - contribution in writing]
 
@@
     latex_engine: pdflatex
+    toc: false
 ---
@@
 This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
-Such information can be used to infer the proportion of young adults, and consequently the productivity.
-In using two populations of ruddy duck, this study develops a Bayesian method to estimate the productivity from count data.
+Such information can be used to infer the proportion of recruits, and consequently the productivity.
+In using two populations of ruddy duck, *Oxyura oxyura*, this study develops a Bayesian method to estimate the productivity from count data.
 To judge the accuracy of this approach, the results are compared to productivity estimates from samples.
@@
 
-## The use of productivity in population management
+## The use of productivity in population management {-}
 
@@
 If one considers a species reaching the sexual maturity as early as its first year, all the population is mature just before the reproduction season.
-The productivity is thus the ratio of the number of young adults at the reproduction season the year *n* on the number of all adults at the reproduction season the year *n - 1*.
+The productivity is thus the ratio of the number of young adults at the reproduction season the year $t$ on the number of all adults at the reproduction season the year $t - 1$.
 
-The productivity is closely related to recruitment. When combined with the adult survival, the recruitment determines the population growth rate and consequently the evolution of the population size.
+When combined to the breeding population size, the productivity determines the recruitment.
+
+When combined with the adult survival, the productivity determines the population growth rate and consequently the evolution of the population size.
 
@@
 
-## The variability of the productivity
+## The variability of the productivity {-}
 
@@
 Such high harvest pressure increases adult and juvenile mortality (ref).
-Paradoxically, this high juvenile mortality does not induce a decrease of productivity (ref).
+Paradoxically, this high juvenile mortality does not always induce a decrease of productivity (ref).
 Indeed, one expects that a population with a high adult mortality combined to a low productivity would rapidly extinct.
@@
 The exploitation of a population thus increases adult mortality and productivity.
-These two parameters can reach a new equilibrium if the population is exploited.
+These two parameters can reach a new equilibrium if the population is exploited to a constant harvest rate.
 The size of a newly exploited population in stable environmental conditions should thus reach a new equilibrium, which is expected to be lower than the size without exploitation.
@@
 Productivity is thus a varying parameter that reaches its highest values when no density-dependent effects occurs.
-Exploring the gradient of productivity, and defining the maximum productivity, and consequently the response of a population to a harvest pressure...
+Exploring the gradient of productivity, and defining the maximum productivity to predict the response of a population to different harvest rates is a challenge in population management.
 
-## Is the variability of the productivity huge or not? What part of the productivity gradient should we know for population management?
+## What part of the productivity gradient should we know for population management? {-}
 
-## What is the ideal biological model to explore the gradient of the productivity?
+## What is the ideal biological model to explore the gradient of the productivity? {-}
 
-Hunting waterfowl has a long history, and few species have observed long-term moratorium
+Hunting waterfowl has a long history, and few species have observed long-term moratorium, so few opportunities to explore the response of the productivity to a gradient of pressure. 
 Alien species colonizing new territories and then heavily controled are a good model to overcome density-dependent effects and explore the gradient productivity.
 These species are often introduced during unvoluntary release events.
-and its presence causes some se species are often arrived because of undesired released
+Since its presence might impact the balance of the colonized ecosytem, it happens that  managers take control measures to restrict the growth of these populations or to eradicate them.
 
@@
 
-## What are the limitations of the sampling method to estimate the productivity?
+## What are the limitations of the sampling method to estimate the productivity? {-}
 
@@
 
-## Why counting data is a potential tool to estimate the productivity?
+## Why counting data is a potential tool to estimate the productivity? {-}
 
-Long time series, easy access, no need of samples
+Long time series, easy access, no need of samples.
+Can only provide subestimation of productivity because if some male recruits have already  their male colours, they are considered as adult and not a recruit.
 
-## What is the methodology of this study?
+## What is the methodology of this study? {-}
 
@@
 # Materials & methods
+
+## Few words on the species {-}
+
+The ruddy duck is a species introduced in the United Kingdom in the 50's.
+The first reproduction was observed in the 60's.
+The population rapidely grew until reaching about more than 5000 individuals spread over the entire country.
+A new population set up in France from the 90's, likely because of an arrival of few individuals from the United Kingdom.
+This population set up around a single wintering spot, the Grand Lieu Lake.
+Since this species is considered as a major threat to the white-headed duck because of the risk of hybridization (AEWA report), a European plan of eradication has been concluded (EU report).
+Control measures were taken in both countries from 1999 in the UK and 2004 in France.
+
+## Dichromatism characterization {-}
+
+Pictures of male juvenile in winter
+Pictures of male in winter
+
+## Time series {-}
+
+The French and the UK wintering populations have been monitored since the presence of the first individuals was validated.
+
+## Model {-}
+
+|Name|Class|Description|
+|:--|:----|:----------------------------|
+|$SM$|Data|Total number of adult males sampled|
+|$SF$|Data|Total number of adult females sampled|
+|$PM$|Parameter|Proportion of males in the adult part of population at its equilibrium|
+|$S_{i, t}$|Data|Number of individuals sampled in population $i$ in year $t$|
+|$SJ_{i, t}$|Data|Number of juveniles sampled in population $i$ in year $t$|
+|$CM_{i, t}$|Data|Cumulated number of type-male individuals counted in population $i$ in year $t$|
+|$C_{i, t}$|Data|Cumulated number of individuals counted in population $i$ in year $t$|
+|$Cmax_{i, t}$|Data|Maximum number of individuals counted in population $i$ in year $t$|
+|$PR_{i, t}$|Parameter|Proportion of recruits in population $i$ in year $t$|
+|$R_{i, t}$|Parameter|Number of recruits in population $i$ in year $t$|
+|$P_{i, t}$|Parameter|Number of recruits in population $i$ in year $t$ per breeder in year $t - 1$|
+|$N0_{j}$|Parameter|Intercept of the regression model, the $j$ index refers to a restricted time period for a single population|
+|$SD_{j}$|Parameter|Standard deviation of the regression model, the $j$ index refers to a restricted time period for a single population|
+|$\lambda_{j}$|Parameter|Average annual growth rate, the $j$ index refers to a restricted time period for a single population|
+|$S_{j}$|Parameter|Average annual growth rate, the $j$ index refers to a restricted time period for a single population|
+
+The proportion of males in the adult part of a population at its equilibrium is inferred from the samples.
+
+Indeed, the available French samples were insufficient to get Since the . Age in samples  to 2003 to 2020 for the United Kingdom and
+
+\begin{equation}
+PM \sim {\sf Beta}(SM, SF)
+(\#eq:PM)
+\end{equation}
+
+$PR_{i, t}$ estimation by the counting method
+
+\begin{equation}
+CM_{i, t} \sim {\sf Binom}(PM.(1 - PR_{i, t}), C_{i, t})
+(\#eq:PRC)
+\end{equation}
+
+$PR_{i, t}$ estimation by the sampling method
+
+\begin{equation}
+SJ_{i, t} \sim {\sf Binom}(PR_{i, t}, S_{i, t})
+(\#eq:PRS)
+\end{equation}
+
+From $PR_{i, t}$ estimates, $R_{i, t}$ and $P_{i, t}$ are expressed as:
+
+\begin{equation}
+R_{i, t} = PR_{i, t}.Cmax_{i, t}
+(\#eq:R)
+\end{equation}
+
+\begin{equation}
+P_{i, t} = \frac{PR_{i, t}.Cmax_{i, t}}{Cmax_{i, t - 1}}
+(\#eq:P)
+\end{equation}
+
+The index $j$ refers to a single population $i$ and a restricted time period for $t$.
+The evolution of the population size can be related to the time to infer the population growth rate.
+It is a linear regression on the logarithm scale:
+
+\begin{equation}
+log(Cmax_{i, t}) \sim {\sf Norm}(N0_{j} + log(\lambda_{j}).t, SD_{j})
+(\#eq:L)
+\end{equation}
+
+\begin{equation}
+S_{j} = \lambda_{j} - \overline{P_{i, t}}
+(\#eq:S)
+\end{equation}
+
+## Survival estimation {-}
 
@@
 
+Productivité: 
+Stabilité grosse pop, 
+variabilité petite pop.
+
+Productivity: drop from 2009
+
 blabla

### 2021-09-06 Code/Model_productivity.R

@@ -50,7 +50,7 @@ raw_plot <- function(para,
   JuvOut4 %>%
     filter(par == as_name(para)) %>% 
     ggplot(aes(x = year, y = `50%`, color = method, fill = method)) +
-    facet_wrap(~ sub_pop, nrow = 1, scales = "free_x") +
+    facet_wrap(~ pop, nrow = 1, scales = "free_x") +
     geom_line(linetype = "dashed", alpha = .5) +
     geom_point() +
     geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5) +
@@ -306,8 +306,12 @@ counts_1 %>%
 # Combined dataset ------------------------------------------------------------------
 
 ds_1 %>% 
-  left_join(counts_2 %>% select(year, pop, sub_pop)) -> ds
-
+  left_join(counts_2 %>% select(year, pop, sub_pop)) %>% 
+  mutate(pop_title = case_when(
+    pop == "FR" ~ "FR small pop.",
+    pop == "UK" & year(year) < 1981 ~ "UK small pop.",
+    pop == "UK" & year(year) > 1980 ~ "UK large pop.",
+    TRUE ~ NA_character_) %>% as_factor()) -> ds
 
 # Data formatting function ----------------------------------------------------------
 
@@ -479,11 +483,11 @@ nimbleMCMC(
 
 # Output check ----------------------------------------------------------------------
 
-# vérif autorcorrélation et burning
+# autocorrelation and burning check
 mcmcplot(mcmcout = JuvOut,
          parms = JuvMon[6]) 
 
-# verif convergence + visualisation update des priors
+# convergence check + vizualisation of prior update
 MCMCtrace(object = JuvOut, 
           params = JuvMon[6], 
           iter = nsample_1,
@@ -499,6 +503,7 @@ MCMCplot(JuvOut,
          ylim = c(0, 1.5))
 
 # Output formatting -----------------------------------------------------------------
+
 MCMCsummary(object = JuvOut, 
             Rhat = TRUE, 
             n.eff = TRUE)
@@ -534,8 +539,8 @@ JuvOut2 %>%
                         rep("sampling", dis(spl_tot, pop, id, TRUE)))[id],
     par == "lambda_max" ~ levels(dis(sub_pop, sub_pop, sub_pop))[id],
     TRUE ~ NA_character_)) %>% 
[…]

### 2021-09-07 Code/Draft_ruddy_duck_productivity.Rmd
abstract modif
biblio path

@@
 lang: en-EN
-bibliography: ../../../../Biblio/JabRef/Library_OFB/Library_OFB.bib
-csl: ../../../../Biblio/JabRef/Library_OFB/Styles/ecology.csl
+bibliography: ../../../Dropbox/Biblio/Library_OFB.bib
+csl: ../../../Dropbox/Biblio/Styles/ecology.csl
 title: "Productivity estimation in waterfowl using a non-invasive method"
@@
 This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
-Such information can be used to infer the proportion of recruits, and consequently the productivity.
+Such information can be used to infer the proportion of young adults, and consequently the productivity.
 In using two populations of ruddy duck, *Oxyura oxyura*, this study develops a Bayesian method to estimate the productivity from count data.

### 2021-09-09 Code/Draft_ruddy_duck_productivity.Rmd
light modif

@@
 lang: en-EN
-bibliography: ../../../Dropbox/Biblio/Library_OFB.bib
-csl: ../../../Dropbox/Biblio/Styles/ecology.csl
+bibliography: ../../../../Dropbox/Biblio/Library_OFB.bib
+csl: ../../../../Dropbox/Biblio/Styles/ecology.csl
 title: "Productivity estimation in waterfowl using a non-invasive method"

### 2021-09-09 Code/Model_productivity.R

@@ -12,7 +12,7 @@ require(MCMCvis)
 require(mcmcplots)
 require(tls) # total least squares regression
 
-# Colors selection -------------------------------------------------------------------
+# Colours selection -------------------------------------------------------------------
 
 tibble(id = 1:20) %>% 
   ggplot(aes(x = 1, y = id, color = factor(id))) +
@@ -194,7 +194,7 @@ frag %>%
 
 # Lambda max formatting ------------------------------------------------------------
 
-# estimatation of lambda_max to test productivity reliability
+# estimation of lambda_max to test productivity reliability
 # when ruddy ducks start to be permanently seen?
 counts_1 %>% 
   ggplot(aes(x = year, y = n_pop, color = pop)) + 
@@ -335,7 +335,7 @@ dis <- function(fil = NA, gp = NA, col = NA, size = FALSE) {
 
 JuvCode <- nimbleCode(
   {
-    # conjugate posterior for the mal proportion (adult only)
+    # conjugate posterior for the male proportion (adult only)
     p_mal ~ dbeta(mal_ad + 1, fem_ad + 1)
     
     # parameter estimation from apparent population structure

### 2021-09-10 Code/Model_productivity.R
changes in breeding pop size

@@ -114,7 +114,7 @@ cor_plot <- function(col, title = NULL, limits = NA) {
 
 # Data import -----------------------------------------------------------------------
 
-read_rds("../Ruddy_duck_data/Data/Ruddy_duck_data.rds") %>% pluck(1) %>% 
+read_rds("../Ruddy_duck_data/Output/Ruddy_duck_data.rds") %>% pluck(1) %>% 
   mutate(across(pop, as_factor)) %>% arrange(pop) -> frag
 
 frag %>% 
@@ -143,7 +143,7 @@ frag %>%
   guides(size = guide_legend(title = "Samples")) +
   labs(y = "Male proportion")
 
-read_rds("../Ruddy_duck_data/Data/Ruddy_duck_data.rds") %>% pluck(2) %>% 
+read_rds("../Ruddy_duck_data/Output/Ruddy_duck_data.rds") %>% pluck(2) %>% 
   mutate(across(pop, as_factor)) %>% arrange(pop) -> counts
 
 # Count and sample formatting ------------------------------------------------------
@@ -161,7 +161,8 @@ counts %>%
   mutate(across(c(ad, no_ad), ~ if_else(ad + no_ad > 100, .x, NA_real_))) %>% 
   left_join(
     counts %>% 
-      mutate(year = year + years(1)) %>% 
+      mutate(year = year + years(1),
+             n_breed = count - killed_before_rep) %>% 
       select(year, pop, n_breed = count)) %>% 
   full_join(counts %>% select(year, count, pop)) %>% 
   rename(n_pop = count) %>% 

### 2021-09-10 Code/Model_productivity.R
commit before switch to master branch

@@ -151,25 +151,33 @@ read_rds("../Ruddy_duck_data/Output/Ruddy_duck_data.rds") %>% pluck(2) %>%
 # one keeps only counts from Dec and Jan
 # one keeps only samples when total > 100
 counts %>% 
-  unnest(age) %>% 
   unnest(sex_app) %>% 
   filter(obs_type_mal + obs_type_fem > 0, month(date) %in% c(12, 1)) %>% 
-  group_by(across(c(year:no_ad, pop))) %>% 
+  group_by(across(c(year, pop))) %>% 
   summarize(across(starts_with("obs"), sum)) %>% 
   ungroup() %>% 
-  filter(obs_type_mal + obs_type_fem > 100) %>% 
-  mutate(across(c(ad, no_ad), ~ if_else(ad + no_ad > 100, .x, NA_real_))) %>% 
-  left_join(
-    counts %>% 
-      mutate(year = year + years(1),
-             n_breed = count - killed_before_rep) %>% 
-      select(year, pop, n_breed = count)) %>% 
-  full_join(counts %>% select(year, count, pop)) %>% 
-  rename(n_pop = count) %>% 
+  filter(obs_type_mal + obs_type_fem > 100) -> count_sex_app
+
+counts %>% 
+  unnest(age) %>% 
+  filter(ad + no_ad > 100) %>% 
+  select(year, pop, ad, no_ad) -> sample_recruit
+
+counts %>% 
+  mutate(year = year + years(1),
+         killed_before_rep = killed_before_rep %>% replace_na(0),
+         n_breed = count - killed_before_rep) %>% 
+  select(year, pop, n_breed) -> count_breed
+
+counts %>% 
+  select(year, n_pop = count, pop) %>% 
+  left_join(count_breed) %>% 
+  left_join(count_sex_app) %>% 
+  left_join(sample_recruit) %>% 
   arrange(pop, year) -> counts_1
 
 # male and female proportion in samples
-# one keeps only years with only few unidentified sex
+# one keeps only years with only few unidentified individuals
 # when regulation starts and prevents from estimating r_max?
 
 frag %>% 
@@ -182,7 +190,7 @@ frag %>%
 
 frag %>%  
   filter(age == "ad", sex != "ind") %>% 
-  filter(!(pop == "FR" & year(year) < 2011), !(pop == "UK" & year(year) < 2002)) %>%
[…]

### 2021-09-10 Code/Model_productivity.R
breeding size acounts for removals before the reproduction

@@ -11,6 +11,7 @@ require(nimble)
 require(MCMCvis)
 require(mcmcplots)
 require(tls) # total least squares regression
+library(scales)
 
 # Colours selection -------------------------------------------------------------------
 
@@ -65,7 +66,7 @@ raw_plot <- function(para,
   if (percent == TRUE) {tmp + scale_y_percent} else {tmp}
 }
 
-cor_plot <- function(col, title = NULL, limits = NA) {
+cor_plot <- function(col, title = NULL, limits = NA, breaks = NA) {
   
   col2 <- str_c(col, c("2.5%", "50%", "97.5%"), sep = "_")
   col_s <- str_c(col2, "sampling", sep = "_")
@@ -105,11 +106,17 @@ cor_plot <- function(col, title = NULL, limits = NA) {
     scale_x_continuous(expand = expansion(mult = c(0, 0)), 
                        limits = if (is.na(limits)) {
                          c(0, 1.1 * max(unlist(tmp[-1])))
-                       } else {limits}) +
+                       } else {limits}, 
+                       breaks = if (is.na(breaks)) {
+                         pretty_breaks()
+                       } else {breaks_extended(n = breaks)}) +
     scale_y_continuous(expand = expansion(mult = c(0, 0)), 
                        limits = if (is.na(limits)) {
                          c(0, 1.1 * max(unlist(tmp[-1])))
-                       } else {limits})
+                       } else {limits}, 
+                       breaks = if (is.na(breaks)) {
+                         pretty_breaks()
+                       } else {breaks_extended(n = breaks)})
 }
 
 # Data import -----------------------------------------------------------------------
@@ -201,7 +208,7 @@ frag %>%
          spl_tot = ad + no_ad) %>% 
   rowid_to_column() -> ds_1
 
-# Lambda max formatting ------------------------------------------------------------
+# Lambda formatting ------------------------------------------------------------
 
 # estimation of lambda_max to test productivity reliability
 # when ruddy ducks start to be permanently seen?
@@ -228,7 +235,7 @@ frag %>%
 
 counts_1 %>% 
   left_join(frag %>% 
-              filter(!(age %in% c("no_ad", "ind") & repro == "after_rep")) %>% 
[…]

### 2021-09-15 Code/Draft_ruddy_duck_productivity.Rmd
model running with pop size = f(year)

--- 2021-09-07 Code/Draft_ruddy_duck_productivity.Rmd
@@
 lang: en-EN
-bibliography: ../../../Dropbox/Biblio/Library_OFB.bib
-csl: ../../../Dropbox/Biblio/Styles/ecology.csl
+bibliography: ../../../../Biblio/JabRef/Library_OFB.bib
+csl: ../../../../Biblio/JabRef/Styles/ecology.csl
 title: "Productivity estimation in waterfowl using a non-invasive method"
@@
 The productivity, i.e.
-the number of recruits per breeder, is even more informative because it is independent of the population size and allows a comparison over time and between species.
+the number of recruits produced per breeder, is even more informative because it is independent of the population size and allows a comparison over time and between species.
 The proportion of young adults in a waterfowl population is often estimated from game-hunting samples.
@@
 To judge the accuracy of this approach, the results are compared to productivity estimates from samples.
-The adult survival can be estimated from both the counting and the sampling methods and is commonly found in the literature.
-The consistency of the two approaches is thus discussed by comparing the survival estimates with the literature values.
+The adult survival is commonly assessed by capture mark recapture and published in the literature.
+Since the survival can also be estimated from both the counting and the sampling methods, the consistency of the two approaches is discussed by comparing the survival estimates with the literature values.
 
 # Introduction
+
+## Definitions
+
+One defines the productivity of a breeding population as the average number of recruits produced per each breeder (ref).
+If one considers a species reaching the sexual maturity as early as its first year, all the population is mature just before the reproduction season.
+The productivity is thus the ratio of the number of young adults at the reproduction season in year $t$, i.e.
+the recruits, to the number of all individuals before the reproduction season in year $t - 1$, i.e.
+the recruits and the adults older than one year.
+
+The population size in year $t$ is equal to the number of breeders that survived over year $t - 1$ plus the juveniles produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
+the recruitment in year $t$.
+The growth rate of a population is thus the sum of the productivity and the adult survival rate.
+
+The productivity parameter is complex because it is composed of two sub-parameters: the reproduction success and the juvenile survival.
 
@@
 
-The productivity of a population is the average number of recruits produced per breeder.
-If one considers a species reaching the sexual maturity as early as its first year, all the population is mature just before the reproduction season.
-The productivity is thus the ratio of the number of young adults at the reproduction season the year $t$ on the number of all adults at the reproduction season the year $t - 1$.
-
-When combined to the breeding population size, the productivity determines the recruitment.
-
-When combined with the adult survival, the productivity determines the population growth rate and consequently the evolution of the population size.
-
-This parameter is complex because it is composed of two sub-parameters: the reproduction success and the youth survival.
-
 ## The variability of the productivity {-}
@@
 A waterfowl species that is released in a favourable habitat will naturally expand (malthus).
-The growth of such a population can be explained by its productivity parameter that is higher than its adult mortality.
+The growth of such a population can be explained by its productivity parameter that is higher than its adult mortality rate.
 This expansion is limited by the carrying capacity of the habitat (ref).
-When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity without because the individuals compete for space or/and food (ref).
+When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity because the individuals compete for space or/and food (ref), which prevents the population from expanding.
 This competition commonly induces a lower reproduction success or/and a higher youth mortality (ref).
 The productivity then decreases because it is dampened by density-dependent effects (holt).
-For an endemic species evolving in stable environmental conditions, the productivity and the adult mortality are balanced, which explains the stabilization of the number of individuals.
+For an endemic species evolving in stable environmental conditions, the productivity and the adult mortality rate are balanced, which explains the stabilization of the number of individuals.
 
-Hunted waterfowl species can experience high harvest rate.
-Such high harvest pressure increases adult and juvenile mortality (ref).
-Paradoxically, this high juvenile mortality does not always induce a decrease of productivity (ref).
-Indeed, one expects that a population with a high adult mortality combined to a low productivity would rapidly extinct.
+Hunting waterfowl induces adult and juvenile mortality (ref).
+Paradoxically, this juvenile mortality does not always induce a decrease of productivity (ref).
+Indeed, one expects that a population with a high adult mortality rate combined to a low productivity would rapidly extinct.
 However, hunting waterfowl has a long history, and most of these species have persisted over time.
-The reason is that high harvest pressure reduces the competition for space and food.
+The reason is that harvest pressure reduces the competition for space and food.
 The density-dependent effects affecting the productivity in case of high population density are dampened by the harvest mortality.
-The exploitation of a population thus increases adult mortality and productivity.
-These two parameters can reach a new equilibrium if the population is exploited to a constant harvest rate.
+The exploitation of a population thus increases adult mortality, but also productivity.
+A balance between these two parameters can be reached if the population is exploited to a constant and moderate harvest rate.
 The size of a newly exploited population in stable environmental conditions should thus reach a new equilibrium, which is expected to be lower than the size without exploitation.

### 2021-09-15 Code/Model_productivity.R

@@ -246,10 +246,12 @@ counts_1 %>%
   geom_hline(yintercept = c(.1, .2), linetype = "dashed") +
   geom_point() +
   geom_line() +
-  facet_wrap( ~ pop, nrow = 2) +
-  scale_y_percent +
+  facet_wrap(~ pop, nrow = 2) +
   scale_x_date_own(1e-2) +
-  scale_color_manual(values = c_pop)
+  scale_color_manual(values = c_pop) +
+  scale_y_continuous(labels = scales::percent, 
+                     breaks = scales::pretty_breaks(), 
+                     limits = c(0, 1.1))
 # from 1999 in UK, and from 1996 in FR si 0%
 # from 1999 in UK, and from 2004 in FR si 10%
 # from 2006 in UK, and from 2008 in FR si 20%

### 2021-09-15 Code/Model_productivity.R
New lambda calculation strategy

@@ -211,7 +211,7 @@ frag %>%
 # Lambda formatting ------------------------------------------------------------
 
 # estimation of lambda_max to test productivity reliability
-# when ruddy ducks start to be permanently seen?
+# when ruddy ducks start to be permanently seen and growing?
 counts_1 %>% 
   ggplot(aes(x = year, y = n_pop, color = pop)) + 
   geom_line() +
@@ -221,8 +221,8 @@ counts_1 %>%
   scale_y_log10() +
   scale_x_date_own(1e-2)
 
-# from 1961 in UK, from 1994 in FR
-# growth shift from 1981 in UK
+# from 1972 in UK, from 1994 in FR
+# growth shift from 1981 in UK when pop > 1000
 
 # when exploitation starts and prevents from estimating lambda_max
 frag %>% 
@@ -243,7 +243,7 @@ counts_1 %>%
   mutate(across(shot, replace_na, 0),
          exploitation_rate = (shot / n_pop) %>% replace_na(0)) %>% 
   ggplot(aes(x = year, y = exploitation_rate, color = pop)) +
-  geom_hline(yintercept = c(.1, .2), linetype = "dashed") +
+  geom_hline(yintercept = c(.1), linetype = "dashed") +
   geom_point() +
   geom_line() +
   facet_wrap(~ pop, nrow = 2) +
@@ -253,37 +253,95 @@ counts_1 %>%
                      breaks = scales::pretty_breaks(), 
                      limits = c(0, 1.1))
 # from 1999 in UK, and from 1996 in FR si 0%
-# from 1999 in UK, and from 2004 in FR si 10%
+# from 1999 in UK, and from 2004 in FR si 10% 
+# but remove lambda on year 2000 & 2001 because high exploitation in year 1999 & 2000
 # from 2006 in UK, and from 2008 in FR si 20%
 
+
+
+
 counts_1 %>% 
-  filter(!(pop == "UK" & year(year) > 1998), !(pop == "UK" & year(year) < 1961),
-         !(pop == "FR" & year(year) > 2003), !(pop == "FR" & year(year) < 1994)) %>% 
+  group_by(pop) %>% 
+  nest() %>% 
+  mutate(data = data %>% 
+           map(~ .x %>% mutate(n_pop_prev = c(0, .$n_pop[-length(.$n_pop)])))) %>% 
+  unnest(data) %>% 
[…]

### 2021-09-29 Code/Model_productivity.R
Plots for article

@@ -12,6 +12,10 @@ require(MCMCvis)
 require(mcmcplots)
 require(tls) # total least squares regression
 require(scales)
+require(gridExtra)
+require(cowplot)
+
+theme_set(theme_bw())
 
 # Colours selection -------------------------------------------------------------------
 
@@ -52,10 +56,10 @@ raw_plot <- function(para,
     filter(par == as_name(para)) %>% 
     ggplot(aes(x = year, y = `50%`, color = method, fill = method)) +
     facet_wrap(~ pop, nrow = 1, scales = "free_x") +
-    geom_line(linetype = "dashed", alpha = .5) +
+    # geom_line(linetype = "dashed", alpha = .5) +
     geom_point() +
     geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5) +
-    scale_y_continuous(limits = if (is.na(ylimits)) {c(0, NA)} else {ylimits}) +
+    scale_y_continuous(limits = if (is.na(ylimits)) {c(0, NA)} else {c(0, ylimits)}) +
     scale_x_date(date_breaks = "2 years", 
                  date_minor_breaks = "1 year", 
                  date_labels = "%Y") +
@@ -475,7 +479,7 @@ JuvInit <- list(p_mal = 0.6,
                 prior = rep(.1, 3),
                 p_juv = rep(.1, JuvConst$C + JuvConst$S),
                 log_N_int = rep(0, 6),
-                r_avg = rep(.1, 5),
+                r_avg = rep(.01, 5),
                 N_sd = rep(1, 5))
 
 # node targets
@@ -597,16 +601,22 @@ sel %>%
       rowMeans()
     lambda = as_tibble(JuvOut) %>% pull(unique(x$slope))
     survival = quantile(lambda - prod, c(.025, .5, .975))
+    prod_avg = quantile(prod, c(.025, .5, .975))
+    sur <- tibble(survival, name = names(survival)) %>% 
+      pivot_wider(names_from = name, values_from = survival) %>% 
+      mutate(par = "survival_avg")
+    pro <- tibble(prod_avg, name = names(prod_avg)) %>% 
+      pivot_wider(names_from = name, values_from = prod_avg) %>% 
+      mutate(par = "productivity_avg")
     out = x %>% 
       distinct(pop, gp, method) %>% 
-      bind_cols(tibble(survival, name = names(survival)) %>% 
[…]

### 2021-09-29 Code/Draft_ruddy_duck_productivity.Rmd
final before merging

@@
     toc: false
+header-includes:
+  \usepackage{float}
+
 ---
@@
                       out.width = "100%", 
-                      fig.align = "center") # , fig.pos = 'H' pour tenir la position
+                      fig.align = "center",
+                      fig.pos = 'H') # pour tenir la position
 
@@
 
+```{r Harv, fig.cap = "blabla"}
+knitr::include_graphics("../Output/plot_1.png")
+```
+
 ## Model {-}

### 2021-09-29 Code/Draft_ruddy_duck_productivity.Rmd
end of the day

@@
 - Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthieu.guillemain@ofb.gouv.fr - contribution in writing]
-
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-```{r Harv, fig.cap = "blabla"}
+|Name|Description|
+|:--|:----------------------------|
+|$N_{t}$|Number of individuals in year $t$|
+|$D_{t}$|Number of individuals removed in year $t$|
+|$R_{t}$|Number of recruits in year $t$|
+|$S_{t}$|Adult survival rate, i.e. proportion of breeder in year $t-1$ still alive in year $t$|
+|$P_{t}$|Number of recruits in year $t$ produced per breeder in year $t - 1$|
+
+The proportion of males in the adult part of a population at its equilibrium is inferred from the samples.
+
+Indeed, the available French samples were insufficient to get Since the . Age in samples  to 2003 to 2020 for the United Kingdom and
+
+\begin{align} 
+N_{t} & = N_{t-1} - D_{t-1} + R_{t} \\
+N_{t} & = N_{t-1} - (1 - S_{t}).N_{t-1} + P_{t}.N_{t-1} \notag \\
+N_{t} & = S_{t}.N_{t-1} + P_{t}.N_{t-1} \notag \\
+\frac{N_{t}}{N_{t-1}} & = S_{t} + P_{t} \notag \\
+\lambda_{t} & = S_{t} + P_{t} 
+(\#eq:GR)
+\end{align}
+
+$PR_{i, t}$ estimation by the counting method
+
+```{r MM1, fig.cap = "(ref:MM1cap)"}
 knitr::include_graphics("../Output/plot_1.png")
+```
+
+```{r MM2, fig.cap = "(ref:MM2cap)"}
+knitr::include_graphics("../Output/plot_2.png")
 ```
@@
 |$SD_{j}$|Parameter|Standard deviation of the regression model, the $j$ index refers to a restricted time period for a single population|
-|$\lambda_{j}$|Parameter|Average annual growth rate, the $j$ index refers to a restricted time period for a single population|
-|$S_{j}$|Parameter|Average annual growth rate, the $j$ index refers to a restricted time period for a single population|
+|$\lambda_{j}$|Parameter|Population growth rate (in $year^{-1}$), the $j$ index refers to a restricted time period for a single population|
+|$S_{j}$|Parameter|Adult survival rate (in $year^{-1}$), the $j$ index refers to a restricted time period for a single population|
 
@@
 
+```{r R1, fig.cap = "(ref:R1cap)"}
+knitr::include_graphics("../Output/plot_3.png")
+```
+
+```{r R2, fig.cap = "(ref:R2cap)"}
+knitr::include_graphics("../Output/plot_4.png")
+```
+
+```{r R3, fig.cap = "(ref:R3cap)"}
+knitr::include_graphics("../Output/plot_5.png")
+```
+
+```{r R4, fig.cap = "(ref:R4cap)"}
+knitr::include_graphics("../Output/plot_6.png")
+```
+
 # Discussion
@@
 <div id="refs"></div>
+
+\newpage
+
+# Supplement 1 {-}
+
+```{r S1, fig.cap = "(ref:S1cap)"}
+knitr::include_graphics("../Output/plot_7.png")
+```
+
+Comments?
+
+(ref:MM1cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+
+(ref:MM2cap) Evolution of the size of the two Ruddy duck populations in light of the data time series; on restricted time periods with available data, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+
+(ref:R1cap) (A) Estimates of productivity, i.e.
+average number of recruits produced per breeder; (B) Comparison of the productivity values between the two estimation methods; bars define the 95% confidence intervals
+
+(ref:R2cap) (A) Estimates of the population growth rate without harvest; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates, bars define the 95\\% confidence intervals
+
+(ref:R3cap) (A) Estimates of the population growth rate on time intervals with data availability.
+This aims at producing survival estimates to discuss the reliability of the methods to estimate productivity; (B) Growth rate estimates, bars define the 95% confidence intervals
+
+(ref:R4cap) (A) Average productivity estimates on the time intervals defined for survival analysis
+
+(ref:S1cap) (A1) Estimates of the number of recruits, i.e.
+the young adults reproducing for their first time; (B1) estimates of the proportion of recruits in the population at the reproduction period; (A2) & (B2) comparison of the values between the two estimation methods; bars define the 95% confidence intervals

### 2021-10-01 Code/Model_productivity.R
dataset within the repertory

@@ -125,7 +125,7 @@ cor_plot <- function(col, title = NULL, limits = NA, breaks = NA) {
 
 # Data import -----------------------------------------------------------------------
 
-read_rds("../Ruddy_duck_data/Output/Ruddy_duck_data.rds") %>% pluck(1) %>% 
+read_rds("./Data/Ruddy_duck_data.rds") %>% pluck(1) %>% 
   mutate(across(pop, as_factor)) %>% arrange(pop) -> frag
 
 frag %>% 
@@ -154,7 +154,7 @@ frag %>%
   guides(size = guide_legend(title = "Samples")) +
   labs(y = "Male proportion")
 
-read_rds("../Ruddy_duck_data/Output/Ruddy_duck_data.rds") %>% pluck(2) %>% 
+read_rds("./Data/Ruddy_duck_data.rds") %>% pluck(2) %>% 
   mutate(across(pop, as_factor)) %>% arrange(pop) -> counts
 
 # Count and sample formatting ------------------------------------------------------
@@ -181,7 +181,8 @@ counts %>%
   select(year, pop, n_breed) -> count_breed
 
 counts %>% 
-  select(year, n_pop = count, pop) %>% 
+  mutate(n_pop = count - killed_before_rep) %>% 
+  select(year, n_win = count, n_pop, pop) %>% 
   left_join(count_breed) %>% 
   left_join(count_sex_app) %>% 
   left_join(sample_recruit) %>% 
@@ -246,7 +247,7 @@ counts_1 %>%
               ungroup() %>% 
               filter(shot > 0)) %>%
   mutate(across(shot, replace_na, 0),
-         exploitation_rate = (shot / n_pop) %>% replace_na(0)) %>% 
+         exploitation_rate = (shot / n_win) %>% replace_na(0)) %>% 
   ggplot(aes(x = year, y = exploitation_rate, color = pop)) +
   geom_hline(yintercept = c(.1), linetype = "dashed") +
   geom_point() +
@@ -263,7 +264,7 @@ counts_1 %>%
 #lambda dataset formatting
 
 counts_1 %>% 
-  mutate(plot1 = if_else(pop == "UK" & year(year) >= 2006 & year(year) <= 2010, 
+  mutate(plot1 = if_else(pop == "UK" & year(year) >= 2006 & year(year) <= 2014, 
                          "UK decrease", NA_character_),
          plot2 = if_else(pop == "FR" & year(year) >= 2004 & year(year) <= 2018, 
                          "FR constant", NA_character_),
[…]

### 2021-10-01 Code/Draft_ruddy_duck_productivity.Rmd
figure update

@@
 
-## Few words on the species {-}
+## The studied species
 
@@
 
-## Dichromatism characterization {-}
+## Dichromatism characterization
 
@@
 
-## Time series {-}
+## Data time series
 
 The French and the UK wintering populations have been monitored since the presence of the first individuals was validated.
-
-|Name|Description|
-|:--|:----------------------------|
-|$N_{t}$|Number of individuals in year $t$|
-|$D_{t}$|Number of individuals removed in year $t$|
-|$R_{t}$|Number of recruits in year $t$|
-|$S_{t}$|Adult survival rate, i.e. proportion of breeder in year $t-1$ still alive in year $t$|
-|$P_{t}$|Number of recruits in year $t$ produced per breeder in year $t - 1$|
-
-The proportion of males in the adult part of a population at its equilibrium is inferred from the samples.
-
-Indeed, the available French samples were insufficient to get Since the . Age in samples  to 2003 to 2020 for the United Kingdom and
-
-\begin{align} 
-N_{t} & = N_{t-1} - D_{t-1} + R_{t} \\
-N_{t} & = N_{t-1} - (1 - S_{t}).N_{t-1} + P_{t}.N_{t-1} \notag \\
-N_{t} & = S_{t}.N_{t-1} + P_{t}.N_{t-1} \notag \\
-\frac{N_{t}}{N_{t-1}} & = S_{t} + P_{t} \notag \\
-\lambda_{t} & = S_{t} + P_{t} 
-(\#eq:GR)
-\end{align}
 
@@
 
-```{r MM2, fig.cap = "(ref:MM2cap)"}
-knitr::include_graphics("../Output/plot_2.png")
-```
-
-## Model {-}
+## Model
 
@@
 
-## Survival estimation {-}
+## Result validation using survival analysis
+
+|Name|Description|
+|:--|:----------------------------|
+|$N_{t}$|Number of breeders in year $t$|
+|$D_{t}$|Number of breeders removed in year $t$|
+|$R_{t}$|Number of recruits in year $t$|
+|$S_{t}$|Adult survival rate, i.e. proportion of breeder in year $t-1$ still alive in year $t$|
+|$P_{t}$|Number of recruits in year $t$ produced per breeder in year $t - 1$|
+
+The proportion of males in the adult part of a population at its equilibrium is inferred from the samples.
+
+Indeed, the available French samples were insufficient to get Since the . Age in samples  to 2003 to 2020 for the United Kingdom and
+
+\begin{align} 
+N_{t} & = N_{t-1} - D_{t-1} + R_{t} \\
+N_{t} & = N_{t-1} - (1 - S_{t}).N_{t-1} + P_{t}.N_{t-1} \notag \\
+N_{t} & = S_{t}.N_{t-1} + P_{t}.N_{t-1} \notag \\
+\frac{N_{t}}{N_{t-1}} & = S_{t} + P_{t} \notag \\
+\lambda_{t} & = S_{t} + P_{t} 
+(\#eq:GR)
+\end{align}
 
@@
 
+## Result validation by comparison to maximum productivity
+
+```{r MM2, fig.cap = "(ref:MM2cap)"}
+knitr::include_graphics("../Output/plot_2.png")
+```
+
 # Results
 
-blabla
+## Productivity estimates
 
@@
 ```
+
+## Result validation using survival analysis
 
@@
 ```
+
+## Result validation by comparison to maximum productivity
 
@@
 
-(ref:MM1cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:MM1cap) Evolution of the size of the two Ruddy duck populations in light of the data time series; on restricted time periods with available data, population growth rates can be estimated (slopes in red for the UK and in blue for France)
 
-(ref:MM2cap) Evolution of the size of the two Ruddy duck populations in light of the data time series; on restricted time periods with available data, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:MM2cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, population growth rates can be estimated (slopes in red for the UK and in blue for France)
 
@@
 
-(ref:R2cap) (A) Estimates of the population growth rate without harvest; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates, bars define the 95\\% confidence intervals
-
-(ref:R3cap) (A) Estimates of the population growth rate on time intervals with data availability.
+(ref:R2cap) (A) Estimates of the population growth rate on time intervals with data availability.
 This aims at producing survival estimates to discuss the reliability of the methods to estimate productivity; (B) Growth rate estimates, bars define the 95% confidence intervals
 
-(ref:R4cap) (A) Average productivity estimates on the time intervals defined for survival analysis
+(ref:R3cap) (A) Average productivity estimates on the time intervals defined for survival analysis; (B) Underlying average survival estimates
+
+(ref:R4cap) (A) Estimates of the population growth rate without harvest; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates, bars define the 95% confidence intervals
 

### 2021-10-01 Code/Model_productivity.R

@@ -264,7 +264,7 @@ counts_1 %>%
 #lambda dataset formatting
 
 counts_1 %>% 
-  mutate(plot1 = if_else(pop == "UK" & year(year) >= 2006 & year(year) <= 2014, 
+  mutate(plot1 = if_else(pop == "UK" & year(year) >= 2006 & year(year) <= 2013, 
                          "UK decrease", NA_character_),
          plot2 = if_else(pop == "FR" & year(year) >= 2004 & year(year) <= 2018, 
                          "FR constant", NA_character_),
@@ -761,6 +761,70 @@ JuvOut4 %>%
 
 # Plots for article ------------------------------------------------------------------
 
+# Time series
+JuvOut3 %>%
+  filter(par == "productivity") %>% 
+  group_split(pop, method) %>% 
+  map(~ .x %>% 
+        mutate(diff = c(1, diff(year(year))) - 1, 
+               diff = if_else(diff != 0, 1, 0), 
+               diff = cumsum(diff)) %>% 
+        group_by(pop, method, diff) %>% 
+        summarize(xmin = min(year) - months(6),
+                  xmax = max(year) + months(6)) %>% 
+        ungroup()) %>%
+  bind_rows() %>% 
+  mutate(label = str_c(method, " time series")) %>% 
+  select(pop, label, xmin, xmax) %>% 
+  bind_rows(
+    ds %>% 
+      mutate(
+        rec = case_when(
+          pop == "UK" & year(year) < 1999 ~ 1,
+          pop == "UK" & year(year) >= 1999 ~ 2,
+          pop == "FR" & year(year) < 1999 ~ 3, 
+          pop == "FR" & year(year) %in% c(1999:2000) ~ 4,
+          pop == "FR" & year(year) %in% c(2001:2003) ~ 5,
+          pop == "FR" & year(year) >= 2004 ~ 6),  
+        label = if_else(rec %in% c(1, 3, 5), 
+                        "no harvest", 
+                        "harvest pressure")) %>% 
+      group_by(rec, label, pop) %>% 
+      summarize(xmin = min(year),
+                xmax = max(year) + years(1)) %>% 
+      rowwise() %>% 
+      mutate(xmax = min(xmax, ymd(20210101))) %>% 
+      ungroup() %>% 
+      select(pop, label, xmin, xmax)) %>% 
+  rowwise() %>% 
+  mutate(label = factor(label, 
+                        levels = c("counting time series", 
+                                   "sampling time series",
[…]

### 2021-10-04 Code/Model_productivity.R
figure update

@@ -708,11 +708,12 @@ JuvOut4 %>%
   facet_wrap(~ pop, nrow = 1, scales = "free_x") +
   geom_point() +
   geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`)) +
-  geom_text(aes(x = method, y = .365, label = txt), color = "#de2d26", angle = -90) +
+  geom_text(aes(x = method, y = .365, label = txt), color = "#de2d26", angle = 90) +
   scale_y_continuous(
     limits = c(0, .7),
     breaks = 0:7 / 10) +
   scale_color_manual(values = c_met, guide = "none") +
+  guides(x = guide_axis(angle = 45)) +
   labs(title = "Productivity: number of recruits per breeder",
        y = NULL, x = NULL) -> prod_plot; prod_plot 
 
@@ -726,12 +727,13 @@ JuvOut4 %>%
   facet_wrap(~ pop, nrow = 1, scales = "free_x") +
   geom_point() +
   geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`)) +
-  geom_text(aes(x = method, y = .5, label = txt), color = "#de2d26", angle = -90) +
+  geom_text(aes(x = method, y = .5, label = txt), color = "#de2d26", angle = 90) +
   scale_y_continuous(
     labels = scales::percent_format(accuracy = 1L),
     limits = c(-.04, 1),
     breaks = 0:10 / 10) +
   scale_color_manual(values = c_met, guide = "none") +
+  guides(x = guide_axis(angle = 45)) +
   labs(title = "Annual adult survival rate",
        y = NULL, x = NULL) -> surviv_plot; surviv_plot 
 
@@ -750,14 +752,66 @@ JuvOut4 %>%
   facet_wrap(~ pop, nrow = 1, scales = "free_x") +
   geom_point() +
   geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`)) +
-  geom_text(aes(x = method, y = .5, label = txt), color = "#de2d26", angle = -90) +
+  geom_text(aes(x = method, y = .5, label = txt), color = "#de2d26", angle = 90) +
   scale_y_continuous(
     labels = scales::percent_format(accuracy = 1L),
     limits = c(-.04, 1.5),
     breaks = 0:15 / 10) +
   scale_color_manual(values = c_met, guide = "none") +
+  guides(x = guide_axis(angle = 45)) +
   labs(title = "Annual adult survival rate",
[…]

### 2021-10-05 Code/Draft_ruddy_duck_productivity.Rmd
update

@@
 
+```{r MM2, fig.cap = "(ref:MM2cap)"}
+knitr::include_graphics("../Output/plot_2.png")
+```
+
 ## Result validation by comparison to maximum productivity
 
-```{r MM2, fig.cap = "(ref:MM2cap)"}
-knitr::include_graphics("../Output/plot_2.png")
+```{r MM3, fig.cap = "(ref:MM3cap)"}
+knitr::include_graphics("../Output/plot_3.png")
 ```
@@
 ```{r R1, fig.cap = "(ref:R1cap)"}
-knitr::include_graphics("../Output/plot_3.png")
+knitr::include_graphics("../Output/plot_4.png")
 ```
@@
 ```{r R2, fig.cap = "(ref:R2cap)"}
-knitr::include_graphics("../Output/plot_4.png")
+knitr::include_graphics("../Output/plot_5.png")
 ```
@@
 ```{r R3, fig.cap = "(ref:R3cap)"}
-knitr::include_graphics("../Output/plot_5.png")
+knitr::include_graphics("../Output/plot_6.png")
 ```
@@
 ```{r R4, fig.cap = "(ref:R4cap)"}
-knitr::include_graphics("../Output/plot_6.png")
+knitr::include_graphics("../Output/plot_7.png")
+```
+
+```{r R5, fig.cap = "(ref:R5cap)"}
+knitr::include_graphics("../Output/plot_8.png")
 ```
@@
 ```{r S1, fig.cap = "(ref:S1cap)"}
-knitr::include_graphics("../Output/plot_7.png")
+knitr::include_graphics("../Output/plot_9.png")
 ```
@@
 
+(ref:MM3cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+
 (ref:R1cap) (A) Estimates of productivity, i.e.
@@
 (ref:R2cap) (A) Estimates of the population growth rate on time intervals with data availability.
-This aims at producing survival estimates to discuss the reliability of the methods to estimate productivity; (B) Growth rate estimates, bars define the 95% confidence intervals
+This aims at producing survival estimates to discuss the reliability of the methods to estimate productivity; (B) Growth rate estimates; bars define the 95% confidence intervals
 
-(ref:R3cap) (A) Average productivity estimates on the time intervals defined for survival analysis; (B) Underlying average survival estimates
+(ref:R3cap) (A) Average productivity estimates on the time intervals defined for survival analysis; (B) Average survival estimates deduced from growth rate and productivity estimates on the time intervals; bars define the 95% confidence intervals
 
-(ref:R4cap) (A) Estimates of the population growth rate without harvest; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates, bars define the 95% confidence intervals
+(ref:R4cap) (A) Estimates of the population growth rate without harvest; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates; bars define the 95% confidence intervals
+
+(ref:R5cap) Comparison between a proxy of maximum productivity and the average productivity estimated following *counting* and *sampling* methods; the proxy is estimated during the maximum growth periods by combining the growth rate to the hypothesis of a 100% adult survival (*max growth*); this hypothesis implies that the true maximum of productivity is likely higher; bars define the 95% confidence intervals
 

### 2021-10-05 Code/Model_productivity.R

@@ -859,7 +859,6 @@ JuvOut3 %>%
          ymin = 10^(log10(6.5e3) + 0.1 * min(c(2, (as.numeric(label) - 1)))),
          ymax = 10^(log10(ymin) + 0.1)) -> time_series
 
-
 ds %>% 
   ggplot(aes(x = year, y = n_pop)) + 
   facet_wrap( ~ pop, ncol = 1) +
@@ -1076,4 +1075,3 @@ list(plo, c(1:length(plo)), plo_dim) %>%
                  ..1, 
                  width = ..3[1], 
                  height = ..3[2], dpi = 600))
-

### 2021-10-19 Code/Model_productivity.R
figure for introduction

@@ -128,6 +128,13 @@ cor_plot <- function(col, title = NULL, limits = NA, breaks = NA) {
 read_rds("./Data/Ruddy_duck_data.rds") %>% pluck(1) %>% 
   mutate(across(pop, as_factor)) %>% arrange(pop) -> frag
 
+read_rds("./Data/Ruddy_duck_data.rds") %>% pluck(2) %>% 
+  mutate(across(pop, as_factor)) %>% arrange(pop) -> counts
+
+
+# male proportion -------------------------------------------------------------------
+
+# adult
 frag %>% 
   filter(age == 'ad', sex != "ind") %>% 
   group_by(year, pop, sex) %>% 
@@ -152,10 +159,48 @@ frag %>%
   scale_y_percent +
   scale_x_date_own(5e-2) +
   guides(size = guide_legend(title = "Samples")) +
-  labs(y = "Male proportion")
+  labs(y = "Male proportion adult")
 
-read_rds("./Data/Ruddy_duck_data.rds") %>% pluck(2) %>% 
-  mutate(across(pop, as_factor)) %>% arrange(pop) -> counts
+# juvenile
+frag %>% 
+  filter(age == 'no_ad', sex != "ind") %>% 
+  group_by(year, pop, sex) %>% 
+  summarize(across(shot, sum)) %>% 
+  group_by(year, pop) %>% 
+  mutate(tot = sum(shot)) %>%  
+  filter(sex == "mal", tot > 0) %>% 
+  group_by(pop) %>% 
+  mutate(inf = qbeta(0.025, shot, tot - shot), 
+         male_proportion = qbeta(0.5, shot, tot - shot), 
+         sup = qbeta(0.975, shot, tot - shot),
+         avg_inf = qbeta(0.025, sum(shot), sum(tot - shot)),
+         avg = qbeta(0.5, sum(shot), sum(tot - shot)),
+         avg_sup = qbeta(0.975, sum(shot), sum(tot - shot))) %>% 
+  ggplot(aes(x = year, y = male_proportion, color = pop)) +
+  geom_ribbon(aes(ymin = avg_inf, ymax = avg_sup), alpha = 0.3, color = "gray") +
+  geom_hline(aes(yintercept = avg), linetype = "dashed") +
+  geom_linerange(aes(ymin = inf, ymax = sup), alpha = .5) +
+  geom_point(aes(size = tot), alpha = .8) +
[…]

### 2021-10-20 Code/Draft_ruddy_duck_productivity.Rmd
update

@@
 The productivity, i.e.
-the number of recruits produced per breeder, is even more informative because it is independent of the population size and allows a comparison over time and between species.
-The proportion of young adults in a waterfowl population is often estimated from game-hunting samples.
-However, this proportion, which is a main step in estimating productivity, is not accessible in years without harvest, or with a low harvest rate.
+the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
+The proportion of young adults in a waterfowl population, which is a main step in estimating productivity, is often estimated from game-hunting samples.
+However, this proportion is not accessible in years without harvest, or with a low harvest rate.
 Moreover, the age-structure in the harvest samples, commonly called hunting bag, does not necessarily reflect the underlying age-structure of the population.
 It is often skewed towards juveniles and can lead to an overestimation of the productivity.
-In waterfowl, adult males usually display brighter colours than juveniles and females.
+In waterfowl, adult males usually display brighter colours than immatures and females.
 This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
@@
 In using two populations of ruddy duck, *Oxyura oxyura*, this study develops a Bayesian method to estimate the productivity from count data.
-To judge the accuracy of this approach, the results are compared to productivity estimates from samples.
-The adult survival is commonly assessed by capture mark recapture and published in the literature.
-Since the survival can also be estimated from both the counting and the sampling methods, the consistency of the two approaches is discussed by comparing the survival estimates with the literature values.
+The results are compared to productivity estimates from samples and are discussed by testing their consistency with the population growth rates.
+The results suggest that the counting method is a versatile tool to estimate the productivity if the species is monitored during the appropriate time period.
 
@@
 
-One defines the productivity of a breeding population as the average number of recruits produced per each breeder (ref).
-If one considers a species reaching the sexual maturity as early as its first year, all the population is mature just before the reproduction season.
+One defines the productivity of a breeding population as the average number of recruits produced per each breeder [@Johnson1987].
+If one considers a species reaching the sexual maturity as early as its first year, all individuals are mature at the reproduction season.
 The productivity is thus the ratio of the number of young adults at the reproduction season in year $t$, i.e.
-the recruits, to the number of all individuals before the reproduction season in year $t - 1$, i.e.
-the recruits and the adults older than one year.
+the recruits, to the number of all individuals at the reproduction season in the previous year, $t - 1$.
 
@@
 
-The productivity parameter is complex because it is composed of two sub-parameters: the reproduction success and the juvenile survival.
+The productivity parameter is complex because it is composed of two sub-parameters: the reproduction succes, which is the product of the nesting and the hatching success, and the juvenile survival.
 
@@
 
+Many studies focus on the monitoring of the reproduction success of waterfowl species is not sufficient to manage properly a population to estimate population growth rate changes [@Dzus1998].
+Productivity defined as recruit per breeder is better, but nesting and hatching success are keys to understand the reasons of productivity variability.
+
 ## The variability of the productivity {-}
 
-A waterfowl species that is released in a favourable habitat will naturally expand (malthus).
+A waterfowl species that is released in a favourable habitat will naturally expand  [@Malthus1872].
 The growth of such a population can be explained by its productivity parameter that is higher than its adult mortality rate.
-This expansion is limited by the carrying capacity of the habitat (ref).
-When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity because the individuals compete for space or/and food (ref), which prevents the population from expanding.
-This competition commonly induces a lower reproduction success or/and a higher youth mortality (ref).
-The productivity then decreases because it is dampened by density-dependent effects (holt).
+This expansion is limited by the carrying capacity of the habitat [@Sayre2008].
+When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity because the individuals compete for space or/and food [@Nummi2015], which prevents the population from expanding.
+This competition commonly induces a lower reproduction success or/and a higher youth mortality.
+The productivity then decreases because it is dampened by density-dependent effects.
 For an endemic species evolving in stable environmental conditions, the productivity and the adult mortality rate are balanced, which explains the stabilization of the number of individuals.
@@
 
+For hunted duck species in North America, the mortality rate on first-year immatures compared to adults is higher by a factor ranging from 1.4 to 2 (.
+
 ## Why counting data is a potential tool to estimate the productivity? {-}
@@
 
+```{r MM1, fig.cap = "(ref:MM1cap)"}
+knitr::include_graphics("../Output/flock.png")
+```
+
+```{r MM2, fig.cap = "(ref:MM2cap)"}
+knitr::include_graphics("../Output/plot_1.png")
+```
+
 Pictures of male juvenile in winter
@@
 
-```{r MM1, fig.cap = "(ref:MM1cap)"}
-knitr::include_graphics("../Output/plot_1.png")
+```{r MM3, fig.cap = "(ref:MM3cap)"}
+knitr::include_graphics("../Output/plot_2.png")
 ```
@@
 
+The adult survival is commonly assessed by capture mark recapture and published in the literature.
+Since the survival can also be estimated from both the counting and the sampling methods, the consistency of the two approaches is discussed by comparing the survival estimates with the literature values.
+
 |Name|Description|
@@
 
-```{r MM2, fig.cap = "(ref:MM2cap)"}
-knitr::include_graphics("../Output/plot_2.png")
+```{r MM4, fig.cap = "(ref:MM4cap)"}
+knitr::include_graphics("../Output/plot_3.png")
 ```
@@
 
-```{r MM3, fig.cap = "(ref:MM3cap)"}
-knitr::include_graphics("../Output/plot_3.png")
+```{r MM5, fig.cap = "(ref:MM5cap)"}
+knitr::include_graphics("../Output/plot_4.png")
 ```
@@
 
-## Productivity estimates
+## Productivity variability
 
 ```{r R1, fig.cap = "(ref:R1cap)"}
-knitr::include_graphics("../Output/plot_4.png")
+knitr::include_graphics("../Output/plot_5.png")
 ```
@@
 ```{r R2, fig.cap = "(ref:R2cap)"}
-knitr::include_graphics("../Output/plot_5.png")
+knitr::include_graphics("../Output/plot_6.png")
 ```
@@
 ```{r R3, fig.cap = "(ref:R3cap)"}
-knitr::include_graphics("../Output/plot_6.png")
+knitr::include_graphics("../Output/plot_7.png")
 ```
@@
 ```{r R4, fig.cap = "(ref:R4cap)"}
-knitr::include_graphics("../Output/plot_7.png")
+knitr::include_graphics("../Output/plot_8.png")
 ```
@@
 ```{r R5, fig.cap = "(ref:R5cap)"}
-knitr::include_graphics("../Output/plot_8.png")
+knitr::include_graphics("../Output/plot_9.png")
 ```
@@
 ```{r S1, fig.cap = "(ref:S1cap)"}
-knitr::include_graphics("../Output/plot_9.png")
+knitr::include_graphics("../Output/plot_10.png")
 ```
 
-Comments?
+(ref:MM1cap) Typical ruddy duck flock observation in winter: 4 male-like individuals with a white cheek and a black cap, 10 female-like individuals with a whitish cheek and a dark stripe across it, 3 unidentified individuals © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
 
-(ref:MM1cap) Evolution of the size of the two Ruddy duck populations in light of the data time series; on restricted time periods with available data, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:MM2cap) Apparent male proportion in winter counts vs male proportion in adult samples; for the first category, one point corresponds to the proportion of male-like plumage in a population counted in winter; for the second category, the male proportion is estimated from the adult ruddy ducks sampled over one year
 
-(ref:MM2cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:MM3cap) Evolution of the size of the two Ruddy duck populations in light of the data time series; on restricted time periods with available data, population growth rates can be estimated (slopes in red for the UK and in blue for France)
 
-(ref:MM3cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:MM4cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+
+(ref:MM5cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, population growth rates can be estimated (slopes in red for the UK and in blue for France)
 

### 2021-10-20 Code/Model_productivity.R

@@ -265,7 +265,7 @@ count_sex_app %>%
          `male proportion` = obs_type_mal / tot) %>% 
   select(method, `male proportion`, tot) %>% 
   bind_rows(frag %>%  
-              filter(age == "ad", sex != "ind") %>% 
+              filter(sex != "ind") %>% 
               filter(!(pop == "FR" & year(year) < 2011), !(pop == "UK" & year(year) > 2009)) %>%
               group_by(year, pop) %>% 
               mutate(tot = sum(shot)) %>%  
@@ -275,7 +275,7 @@ count_sex_app %>%
               ungroup() %>% 
               filter(tot > 50) %>%
               mutate(`male proportion` = mal / tot, 
-                     method = "adult samples") %>%
+                     method = "samples") %>%
               select(method, `male proportion`, tot)) %>% 
   mutate(method = as_factor(method)) %>% 
   ggplot(aes(x = method, y = `male proportion`, size = tot)) + 

### 2021-10-22 Code/Draft_ruddy_duck_productivity.Rmd
update

@@
 the number of young adults reproducing for the first time, is a key indicator to describe the renewal of a population, and therefore an essential tool in population management.
-The productivity, i.e.
-the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
+The productivity, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
 The proportion of young adults in a waterfowl population, which is a main step in estimating productivity, is often estimated from game-hunting samples.
@@
 If one considers a species reaching the sexual maturity as early as its first year, all individuals are mature at the reproduction season.
-The productivity is thus the ratio of the number of young adults at the reproduction season in year $t$, i.e.
-the recruits, to the number of all individuals at the reproduction season in the previous year, $t - 1$.
+The productivity is thus the ratio of the number of young adults at the reproduction season of year $t$, i.e.
+the recruits, to the number of all individuals at the reproduction season of the previous year, $t - 1$.
 
-The population size in year $t$ is equal to the number of breeders that survived over year $t - 1$ plus the juveniles produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
-the recruitment in year $t$.
-The growth rate of a population is thus the sum of the productivity and the adult survival rate.
-
-The productivity parameter is complex because it is composed of two sub-parameters: the reproduction succes, which is the product of the nesting and the hatching success, and the juvenile survival.
+The productivity parameter is complex because it is composed of two main sub-parameters: the reproductive success and the juvenile survival.
+Both sub-parameters can be split into other sub-parameters.
+For instance, the reproductive success is the product of the nesting success and the hatching success.
+The juvenile survival is the combination of the post-hatching survival and the fledging juvenile survival.
+According to @Blums2004, the recruitment for diving ducks is not related to the reproductive success, but depends mostly on the juvenile survival, which is driven by weather conditions.
+However, the results of this study does not necessarily reflect a general reality because several sources of variations of the reproductive success are controlled in the experimental plan of this study.
+The productivity as defined in the present study is thus a parameter integrating many sources of variability affecting all life stages from the egg to the young adult breeding for its first time.
 
@@
 
-The management of an exploited population generally aims at...
-
-Many studies focus on the monitoring of the reproduction success of waterfowl species is not sufficient to manage properly a population to estimate population growth rate changes [@Dzus1998].
-Productivity defined as recruit per breeder is better, but nesting and hatching success are keys to understand the reasons of productivity variability.
+The management of an exploited population generally aims at preserving the conservation status of the population, generally by targeting a minimum population size.
+To do so, a first strategy is to test iteratively management rules to reach progressively an optimum.
+Such approach can be time-consuming given the intrinsic variability of the population response to its changing environment.
+It might also suffer from a rejection of the public because of its lack of transparency and its arbitrary process.
+Another approach is to try to predict the growth rate of the population as a function of the harvest rate and other environment inputs.
+Such modelling approach can provide clues on the recovering time of a population upon a variety of managment rules.
+These decision-support tools are then key inputs for the stakeholders to status on the best management rules to implement.
+The growth rate of a population is the sum of the adult survival and the productivity, see Equation \@ref(eq:GR) in Section \@ref(MM).
+The adult survival is generally estimated by Capture Mark Recapture methods.
+If the harvest pressure behaves as an additive mortality to the natural mortality, it is quite straightforward to predict its impact of the adult survival.
+The key to predict the growth rate of a population is then to estimate the value and the variability of the productivity as defined inhere and to understand if this parameter is impacted or not by a harvest pressure.
 
@@
 A waterfowl species that is released in a favourable habitat will naturally expand  [@Malthus1872].
-The growth of such a population can be explained by its productivity parameter that is higher than its adult mortality rate.
+The growth of such a population can be explained with a productivity parameter that is higher than the adult mortality rate.
 This expansion is limited by the carrying capacity of the habitat [@Sayre2008].
 When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity because the individuals compete for space or/and food [@Nummi2015], which prevents the population from expanding.
-This competition commonly induces a lower reproduction success or/and a higher youth mortality.
-The productivity then decreases because it is dampened by density-dependent effects.
+This competition commonly induces a lower reproductive success or/and a higher youth mortality.
+The population is then less productive because the productivity parameter is dampened by density-dependent effects.
 For an endemic species evolving in stable environmental conditions, the productivity and the adult mortality rate are balanced, which explains the stabilization of the number of individuals.
 
-Hunting waterfowl induces adult and juvenile mortality (ref).
-Paradoxically, this juvenile mortality does not always induce a decrease of productivity (ref).
-Indeed, one expects that a population with a high adult mortality rate combined to a low productivity would rapidly extinct.
+Hunting waterfowl induces adult and juvenile mortality [@Bellrose1980].
+Paradoxically, this juvenile mortality does not always induce a decrease of productivity.
+Indeed, one expects that a population with a high adult mortality rate combined to a lower productivity would rapidly extinct.
 However, hunting waterfowl has a long history, and most of these species have persisted over time.
 The reason is that harvest pressure reduces the competition for space and food.
-The density-dependent effects affecting the productivity in case of high population density are dampened by the harvest mortality.
+The density-dependent effects affecting the productivity in case of high population density are reduced by the harvest mortality.
 The exploitation of a population thus increases adult mortality, but also productivity.
@@
 
-Productivity is thus a varying parameter that reaches its highest values when no density-dependent effects occurs.
-Exploring the gradient of productivity, and defining the maximum productivity to predict the response of a population to different harvest rates is a challenge in population management.
+The reproductive success and the juvenile survival varies a lot following weather conditions [@Blums2004]. So the response of the productivity to the harvest rate can be dilluted by a lot.
+
+Productivity is thus a varying parameter that reaches its highest values when no density-dependent effect occurs and when the harvest pressure on juveniles does not exceed a certain limit.
+Defining the maximum productivity and exploring the variability of productivity to predict the response of a population to different harvest rates is a challenge in population management.
 
@@
 
+We need to know what is the expected maximum productivity. What is the limit of the harvest pressure before affecting the productivity. What is the range of productivity variability.
+
+If the maximum productivity is not reached under optimal conditions (maximum population size not reached, and limited harvest pressure), it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the productivity.
+
 ## What is the ideal biological model to explore the gradient of the productivity? {-}
 
-Hunting waterfowl has a long history, and few species have observed long-term moratorium, so few opportunities to explore the response of the productivity to a gradient of pressure. 
-Alien species colonizing new territories and then heavily controled are a good model to overcome density-dependent effects and explore the gradient productivity.
-These species are often introduced during unvoluntary release events.
-Since its presence might impact the balance of the colonized ecosytem, it happens that  managers take control measures to restrict the growth of these populations or to eradicate them.
+Hunting waterfowl has a long history, and few species have observed long-term moratorium.
+So there are few opportunities to explore the response of the productivity to a gradient of pressure in such species.
+Alien species colonizing new territories and then heavily controlled are then a good model to overcome density-dependent effects and explore the gradient productivity.
+These species are often introduced during involuntary release events.
+Since its presence might impact the balance of the colonized ecosystem, it happens that  managers take control measures to restrict the growth of these populations or to eradicate them.
 
-In Europe, 
-a population following its maximum growth rate is a perfect study model
+In Europe, a population following its maximum growth rate is a perfect study model.
 
@@
 
-- Biais because hunting like any predation action is selective.
+- Bias because hunting like any predation action is selective.
 Often towards the weakest individuals.
-So overstimation of productivity, and potentially maximum sustainable harvest rate overestimated and consequently depletion of the managed population.
+So overestimation of productivity, and potentially maximum sustainable harvest rate overestimated and consequently depletion of the managed population.
 
@@
 Long time series, easy access, no need of samples.
-Can only provide subestimation of productivity because if some male recruits have already  their male colours, they are considered as adult and not a recruit.
+Can only provide underestimation of productivity because if some male recruits have already  their male colours, they are considered as adult and not a recruit.
 
@@
 
-breeding/reproduction success good but not enough because variable survival of juveniles
+breeding/reproductive success good but not enough because variable survival of juveniles
 
@@
 
-# Materials & methods
+# Materials & methods {#MM}
 
@@
 |$N_{t}$|Number of breeders in year $t$|
-|$D_{t}$|Number of breeders removed in year $t$|
+|$D_{t}$|Number of breeders dead after reproduction in year $t$|
 |$R_{t}$|Number of recruits in year $t$|
 |$S_{t}$|Adult survival rate, i.e. proportion of breeder in year $t-1$ still alive in year $t$|
-|$P_{t}$|Number of recruits in year $t$ produced per breeder in year $t - 1$|
-
-The proportion of males in the adult part of a population at its equilibrium is inferred from the samples.
-
-Indeed, the available French samples were insufficient to get Since the . Age in samples  to 2003 to 2020 for the United Kingdom and
+|$P_{t}$|Number of recruits in year $t$ produced per breeder in year $t-1$|
+|$\lambda_{t}$|Growth rate of the population between the reproduction period of year $t-1$ and year $t$|
 
 \begin{align} 
-N_{t} & = N_{t-1} - D_{t-1} + R_{t} \\
+N_{t} & = N_{t-1} - D_{t-1} + R_{t} \notag \\
 N_{t} & = N_{t-1} - (1 - S_{t}).N_{t-1} + P_{t}.N_{t-1} \notag \\
@@
 \end{align}
+
+The population size in year $t$ is equal to the number of breeders that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
+the recruitment in year $t$.
+The growth rate of a population is thus the sum of the productivity and the adult survival rate.
+
+The proportion of males in the adult part of a population at its equilibrium is inferred from the samples.
+
+Indeed, the available French samples were insufficient to get Since the . Age in samples  to 2003 to 2020 for the United Kingdom and
 
@@
 
+quite stable when harvest rate is limited (France 2000), but is lower and noisy when harvest rate is huge and the population depleted.
+
 ## Result validation using survival analysis
@@
 
+Au dela du fait que la croissance de la pop est la somme de la survie et de la productivité, 
+intéressant de parler de la variability de la survie vs variabilité de la productivity 
+pour parler de risk management.
+
 Productivity: drop from 2009
@@
 
-(ref:MM1cap) Typical ruddy duck flock observation in winter: 4 male-like individuals with a white cheek and a black cap, 10 female-like individuals with a whitish cheek and a dark stripe across it, 3 unidentified individuals © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
+(ref:MM1cap) Typical observation of a ruddy duck flock in winter: 4 male-like individuals with a white cheek and a black cap, 10 female-like individuals with a whitish cheek and a dark stripe across it, 3 unidentified individuals © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
 

### 2021-10-27 Code/Draft_ruddy_duck_productivity.Rmd
Intro done

@@
 The results are compared to productivity estimates from samples and are discussed by testing their consistency with the population growth rates.
-The results suggest that the counting method is a versatile tool to estimate the productivity if the species is monitored during the appropriate time period.
+The results suggest that the counting method is a versatile tool to estimate the productivity if the species is monitored during the appropriate time window.
 
@@
 However, the results of this study does not necessarily reflect a general reality because several sources of variations of the reproductive success are controlled in the experimental plan of this study.
-The productivity as defined in the present study is thus a parameter integrating many sources of variability affecting all life stages from the egg to the young adult breeding for its first time.
+The productivity as defined in the present study is thus a parameter integrating many sources of variability affecting all life stages from the egg to the young adult trying to breed for its first time.
 
@@
 Such approach can be time-consuming given the intrinsic variability of the population response to its changing environment.
-It might also suffer from a rejection of the public because of its lack of transparency and its arbitrary process.
+It might also suffer from a rejection of the stakeholders because of the lack of transparency of its arbitrary process.
 Another approach is to try to predict the growth rate of the population as a function of the harvest rate and other environment inputs.
 Such modelling approach can provide clues on the recovering time of a population upon a variety of managment rules.
-These decision-support tools are then key inputs for the stakeholders to status on the best management rules to implement.
-The growth rate of a population is the sum of the adult survival and the productivity, see Equation \@ref(eq:GR) in Section \@ref(MM).
+These decision-support tools are then key inputs to status on the best management rules to implement.
+The growth rate of a population can be simply defined as the sum of the adult survival and the productivity, see Equation \@ref(eq:GR) in Section \@ref(MM).
 The adult survival is generally estimated by Capture Mark Recapture methods.
 If the harvest pressure behaves as an additive mortality to the natural mortality, it is quite straightforward to predict its impact of the adult survival.
-The key to predict the growth rate of a population is then to estimate the value and the variability of the productivity as defined inhere and to understand if this parameter is impacted or not by a harvest pressure.
+The key to predict the growth rate of a population is then to estimate the value and the variability of the productivity and to understand if this parameter is impacted or not by a harvest pressure.
 
@@
 A waterfowl species that is released in a favourable habitat will naturally expand  [@Malthus1872].
-The growth of such a population can be explained with a productivity parameter that is higher than the adult mortality rate.
+The growth of such a population is possible because the productivity parameter is higher than the adult mortality rate.
 This expansion is limited by the carrying capacity of the habitat [@Sayre2008].
 When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity because the individuals compete for space or/and food [@Nummi2015], which prevents the population from expanding.
-This competition commonly induces a lower reproductive success or/and a higher youth mortality.
+This competition commonly induces a lower reproductive success or/and a higher juvenile mortality.
 The population is then less productive because the productivity parameter is dampened by density-dependent effects.
-For an endemic species evolving in stable environmental conditions, the productivity and the adult mortality rate are balanced, which explains the stabilization of the number of individuals.
+For an endemic species evolving in stable environmental conditions, a balance between the productivity and the adult mortality rate explains the stabilization of the number of individuals.
 
 Hunting waterfowl induces adult and juvenile mortality [@Bellrose1980].
-Paradoxically, this juvenile mortality does not always induce a decrease of productivity.
-Indeed, one expects that a population with a high adult mortality rate combined to a lower productivity would rapidly extinct.
+Intuitively, one expects that such a hunted population would observe lower adult survival and productivity and then would rapidly extinct.
 However, hunting waterfowl has a long history, and most of these species have persisted over time.
-The reason is that harvest pressure reduces the competition for space and food.
-The density-dependent effects affecting the productivity in case of high population density are reduced by the harvest mortality.
-The exploitation of a population thus increases adult mortality, but also productivity.
-A balance between these two parameters can be reached if the population is exploited to a constant and moderate harvest rate.
-The size of a newly exploited population in stable environmental conditions should thus reach a new equilibrium, which is expected to be lower than the size without exploitation.
+So paradoxically, hunting immature individuals does not always induce a decrease of productivity.
+A reason could be that the harvest pressure reduces the competition for space and food.
+The density-dependent effects affecting the productivity in case of high population density are then reduced by the harvest mortality.
+In conclusion, a limited exploitation of a population is expected to increase both adult mortality and productivity.
+A balance between these two parameters can be theoritically reached if the population is exploited to a constant and moderate harvest rate.
+The size of a newly exploited population in stable environmental conditions should thus reach a different equilibrium, which is expected to be lower than the size without exploitation.
 
-The reproductive success and the juvenile survival varies a lot following weather conditions [@Blums2004]. So the response of the productivity to the harvest rate can be dilluted by a lot.
+The reproductive success and the juvenile survival, the two components of the productivity, are very sensitive to weather conditions [@Blums2004; @Folliot2017].
+The environment conditions are thus an additional source of variability of the productivity.
+The response of the productivity to a harvest pressure is thus not straightforward.
+It is noisy because of its sensitivity to environmental factors.
+It also depends directly on the intensity of harvest on the immature individuals and indirectly on the mitigation of the density-dependent effects.
 
-Productivity is thus a varying parameter that reaches its highest values when no density-dependent effect occurs and when the harvest pressure on juveniles does not exceed a certain limit.
-Defining the maximum productivity and exploring the variability of productivity to predict the response of a population to different harvest rates is a challenge in population management.
+## What part of the productivity gradient is necessary for population management? {-}
 
-## What part of the productivity gradient should we know for population management? {-}
-
-We need to know what is the expected maximum productivity. What is the limit of the harvest pressure before affecting the productivity. What is the range of productivity variability.
-
-If the maximum productivity is not reached under optimal conditions (maximum population size not reached, and limited harvest pressure), it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the productivity.
+In population mangement, a common question for a threaten or an invasive species is to determine the measure that will allow to reach a certain population size in a limited time period.
+For such question, the carrying capacity of the habitat is theoritically not reached.
+So a manager should mostly focus on collecting knowledge on the maximum productivity and its variability by removing the potential density-dependent effects occurring only for the highest densities.
 
@@
 
-Hunting waterfowl has a long history, and few species have observed long-term moratorium.
+Hunting waterfowl has a long history, but few species have observed long-term moratorium.
 So there are few opportunities to explore the response of the productivity to a gradient of pressure in such species.
-Alien species colonizing new territories and then heavily controlled are then a good model to overcome density-dependent effects and explore the gradient productivity.
-These species are often introduced during involuntary release events.
-Since its presence might impact the balance of the colonized ecosystem, it happens that  managers take control measures to restrict the growth of these populations or to eradicate them.
+Alien species are often introduced during involuntary release events.
+They observe commonly a first period of colonisation on new territories without management measures.
+Since its presence might impact the balance of the colonized ecosystem, it happens that managers take strong control measures to restrict the growth of these populations or to eradicate them.
+Such species are then a good biological model to overcome density-dependent effects and explore the effects of a harvest gradient on productivity.
 
-In Europe, a population following its maximum growth rate is a perfect study model.
+## How to estimate productivity? {-}
 
-## What are the limitations of the sampling method to estimate the productivity? {-}
+The productivity parameter of a population can be assessed by monitoring the number of eggs that survives to the age of the first-breeding event.
+To do so, a first step consists in quantifying the reproductive success by estimating nesting and hatching success thanks to surveys in habitats representative of the managed population.
+The survival of the post-hatched chick to its recruitment can be done by capture mark recapture method [@Blums2004].
+Even if such approach provides valuable information on the dynamics of the recruitment and can highlight the bottlenecks, the fieldwork required to estimate both components is too time-consuming to generalize the method for waterfowl population management.
 
-- We cannot explore the productivity when there is no harvest, however, it is crucial to explore what happens when there is no harvest in a scenario where we suggest a moratorium to let the population recover.
+In waterfowl, it is generally impossible to differenciate immatures from adults without examining the birds in hand.
+Aging individuals from samples coming from hunting bags is thus another possible approach.
+Such approach suffers from two caveats.
+First, it is impossible to explore the productivity when there is no harvest whereas it is necessary to predict what would happen if a moratorium seems required to let a population recover, or to calibrate a priori the harvest effort necessary to control an invasive species [@Smith2005].
+Secondly, like any predation action, hunting is selective and provides a biased picture of the population age structure in favor of the immatures [@Bellrose1980].
+The hunting bag analysis might then lead to overestimate the productivity and consequently the maximum sustainable harvest rate, with adverse consequences on population management.
 
-- Bias because hunting like any predation action is selective.
-Often towards the weakest individuals.
-So overestimation of productivity, and potentially maximum sustainable harvest rate overestimated and consequently depletion of the managed population.
+In waterfowl, it is common that adult males display brighter colors than females when  immatures from both sexes display cover-up plumage similar to adult females *ref*.
+Knowing sex ratios, this discrepancy between adult and immature males might be used to estimate the proportion of immature in a waterfowl population just by counting them from distance.
+Such approach is promising because it requires only to count the proportion of birds of the two plumage-kinds to infer an estimate of productivity.
+Unlike the two previous methods, this one is non-invasive and requires a little time investment, which allows to collect long time series and consequently to track productivity variability.
 
-For hunted duck species in North America, the mortality rate on first-year immatures compared to adults is higher by a factor ranging from 1.4 to 2 (.
+## Methodology of this study {-}
 
-## Why counting data is a potential tool to estimate the productivity? {-}
-
-Long time series, easy access, no need of samples.
-Can only provide underestimation of productivity because if some male recruits have already  their male colours, they are considered as adult and not a recruit.
-
-## What is the methodology of this study? {-}
-
-Nichols -> il faut compter de la façon dont on demande
-
-idée: estimer l'age moyen de la pop grâce la vulnerabilité et le sexe ratio chez les adultes
-
-breeding/reproductive success good but not enough because variable survival of juveniles
-
-Introduction:
-
-productivity often defined as Jt/At
-but more realistic to defined as Jt/At-1
+From two populations of ruddy duck, an exotic species introduced in Europe in the 50's, this study applies a method of productivity estimation from count data.
 
 # Materials & methods {#MM}
+
+The best picture stands just before the breeding period, so generally in winter for waterfowl.
 
@@
 
+Hypothesis of time invariant sex ratios in immature and adult: limits
+
+For hunted duck species in North America, the mortality rate on first-year immatures compared to adults is higher by a factor ranging from 1.4 to 2 (Bellrose 1980).
+
 Productivité: 
@@
 Nichols -> il faut compter de la façon dont on demande
+
+Taper des jeunes c'est pas forcément très utile car density dependence
+
+If the maximum productivity is not reached under optimal conditions (maximum population size not reached, and limited harvest pressure), it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the productivity.
 
@@
 
-(ref:MM4cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:MM4cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods with data time series, population growth rates can be estimated (slopes in red for the UK and in blue for France)
 
-(ref:MM5cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:MM5cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)
 

### 2021-11-03 Code/Draft_ruddy_duck_productivity.Rmd
update

@@
 bibliography: ../../../../Biblio/JabRef/Library_OFB.bib
-csl: ../../../../Biblio/JabRef/Styles/ecology.csl
+csl: ../../../../Biblio/JabRef/Styles/APA.csl
+link-citations: true
 title: "Productivity estimation in waterfowl using a non-invasive method"
@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
+abstract: The response of a waterfowl population to a harvest pressure depends on its capacity to renew.
+The recruitment, i.e.
+the number of young adults reproducing for the first time, is a key indicator to describe the renewal of a population, and therefore an essential tool in population management.
+The productivity, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
+The proportion of young adults in a waterfowl population, which is a main step in estimating productivity, is often estimated from game-hunting samples.
+However, this proportion is not accessible in years without harvest, or with a low harvest rate.
+Moreover, the age-structure in the harvest samples, commonly called hunting bag, does not necessarily reflect the underlying age-structure of the population.
+It is often skewed towards juveniles and can lead to an overestimation of the productivity.
+In waterfowl, adult males usually display brighter colours than immatures and females.
+This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
+Such information can be used to infer the proportion of young adults, and consequently the productivity.
+In using two populations of ruddy duck, *Oxyura oxyura*, this study develops a Bayesian method to estimate the productivity from count data.
+The results are compared to productivity estimates from samples and are discussed by testing their consistency with the population growth rates.
+The results suggest that the counting method is a versatile tool to estimate the productivity if the species is monitored during the appropriate time window.
 output:
@@
   \usepackage{float}
-
 ---
@@
 
-# Abstract {-}
-
-The response of a waterfowl population to a harvest pressure depends on its capacity to renew.
-The recruitment, i.e.
-the number of young adults reproducing for the first time, is a key indicator to describe the renewal of a population, and therefore an essential tool in population management.
-The productivity, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
-The proportion of young adults in a waterfowl population, which is a main step in estimating productivity, is often estimated from game-hunting samples.
-However, this proportion is not accessible in years without harvest, or with a low harvest rate.
-Moreover, the age-structure in the harvest samples, commonly called hunting bag, does not necessarily reflect the underlying age-structure of the population.
-It is often skewed towards juveniles and can lead to an overestimation of the productivity.
-In waterfowl, adult males usually display brighter colours than immatures and females.
-This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
-Such information can be used to infer the proportion of young adults, and consequently the productivity.
-In using two populations of ruddy duck, *Oxyura oxyura*, this study develops a Bayesian method to estimate the productivity from count data.
-The results are compared to productivity estimates from samples and are discussed by testing their consistency with the population growth rates.
-The results suggest that the counting method is a versatile tool to estimate the productivity if the species is monitored during the appropriate time window.
-
 # Introduction
 
-## Definitions
+*Definitions*
 
 One defines the productivity of a breeding population as the average number of recruits produced per each breeder [@Johnson1987].
-If one considers a species reaching the sexual maturity as early as its first year, all individuals are mature at the reproduction season.
+This parameter might sometimes be called "annual fertility" [@Koons2014], or "reproductive success".
+If one considers a species reaching its sexual maturity as early as its first year, all individuals are mature at the reproduction season.
 The productivity is thus the ratio of the number of young adults at the reproduction season of year $t$, i.e.
@@
 
-## The use of productivity in population management {-}
+*The use of productivity in population management*
 
-The management of an exploited population generally aims at preserving the conservation status of the population, generally by targeting a minimum population size.
+The management of an exploited population generally aims at preserving its conservation status, generally by targeting a population size **ref required**.
 To do so, a first strategy is to test iteratively management rules to reach progressively an optimum.
 Such approach can be time-consuming given the intrinsic variability of the population response to its changing environment.
-It might also suffer from a rejection of the stakeholders because of the lack of transparency of its arbitrary process.
+It might also suffer from a rejection of the stakeholders because of the lack of transparency of its arbitrary process **ref required** [@Nichols2000]?.
 Another approach is to try to predict the growth rate of the population as a function of the harvest rate and other environment inputs.
-Such modelling approach can provide clues on the recovering time of a population upon a variety of managment rules.
+Such modelling approach can provide clues on the recovering time of a population upon a variety of managment rules **ref required**.
 These decision-support tools are then key inputs to status on the best management rules to implement.
 The growth rate of a population can be simply defined as the sum of the adult survival and the productivity, see Equation \@ref(eq:GR) in Section \@ref(MM).
-The adult survival is generally estimated by Capture Mark Recapture methods.
+The adult survival is generally estimated by capture mark recapture methods **ref required**.
 If the harvest pressure behaves as an additive mortality to the natural mortality, it is quite straightforward to predict its impact of the adult survival.
-The key to predict the growth rate of a population is then to estimate the value and the variability of the productivity and to understand if this parameter is impacted or not by a harvest pressure.
+The key to predict the growth rate of a population is then to estimate the average value and the variability of the productivity and to understand how this parameter is impacted by a harvest pressure.
 
-## The variability of the productivity {-}
+*The variability of the productivity*
 
 A waterfowl species that is released in a favourable habitat will naturally expand  [@Malthus1872].
-The growth of such a population is possible because the productivity parameter is higher than the adult mortality rate.
+The growth of such a population reflects that the productivity parameter is higher than the adult mortality rate.
 This expansion is limited by the carrying capacity of the habitat [@Sayre2008].
@@
 This competition commonly induces a lower reproductive success or/and a higher juvenile mortality.
-The population is then less productive because the productivity parameter is dampened by density-dependent effects.
-For an endemic species evolving in stable environmental conditions, a balance between the productivity and the adult mortality rate explains the stabilization of the number of individuals.
+The population is then less productive because the productivity parameter is affected by density-dependent effects [@Gunnarsson2013].
+For an endemic species evolving in stable environmental conditions, a balance between the productivity and the adult mortality rate explains the stabilization of the number of individuals **ref required**.
 
 Hunting waterfowl induces adult and juvenile mortality [@Bellrose1980].
-Intuitively, one expects that such a hunted population would observe lower adult survival and productivity and then would rapidly extinct.
-However, hunting waterfowl has a long history, and most of these species have persisted over time.
+Intuitively, one expects that such a hunted population would observe lower adult survival and productivity and then would rapidly deplete to its extinction.
+However, hunting waterfowl has a long history, and most of these species have persisted over time **ref required**.
 So paradoxically, hunting immature individuals does not always induce a decrease of productivity.
-A reason could be that the harvest pressure reduces the competition for space and food.
+A reason could be that the harvest pressure reduces the competition for space and food **ref required**.
 The density-dependent effects affecting the productivity in case of high population density are then reduced by the harvest mortality.
 In conclusion, a limited exploitation of a population is expected to increase both adult mortality and productivity.
-A balance between these two parameters can be theoritically reached if the population is exploited to a constant and moderate harvest rate.
-The size of a newly exploited population in stable environmental conditions should thus reach a different equilibrium, which is expected to be lower than the size without exploitation.
+A balance between these two parameters can be theoritically reached if the population is exploited to a constant and moderate harvest rate **ref required**.
+The size of a newly exploited population in stable environmental conditions should thus reach a different equilibrium, which is expected to be lower than the size without exploitation **ref required**.
 
@@
 
-## What part of the productivity gradient is necessary for population management? {-}
+*What part of the productivity gradient is necessary for population management?*
 
-In population mangement, a common question for a threaten or an invasive species is to determine the measure that will allow to reach a certain population size in a limited time period.
+In population mangement, a common question for a threaten or an invasive species is to determine the measure that will allow to reach a targetted population size/density in a limited time period **ref required**.
 For such question, the carrying capacity of the habitat is theoritically not reached.
-So a manager should mostly focus on collecting knowledge on the maximum productivity and its variability by removing the potential density-dependent effects occurring only for the highest densities.
+So a manager should mostly focus on collecting knowledge on the maximum productivity and its variability by removing the potential density-dependent effects occurring only for the highest theoritical densities.
 
-## What is the ideal biological model to explore the gradient of the productivity? {-}
+*What is the ideal biological model to explore the gradient of the productivity?*
 
-Hunting waterfowl has a long history, but few species have observed long-term moratorium.
-So there are few opportunities to explore the response of the productivity to a gradient of pressure in such species.
-Alien species are often introduced during involuntary release events.
+Hunting waterfowl has a long history, but few species have observed long-term moratorium **ref required**.
+So there are few opportunities to explore the response of the productivity to a gradient of pressure for such species.
+Alien species are often introduced during involuntary release events **ref required**.
 They observe commonly a first period of colonisation on new territories without management measures.
-Since its presence might impact the balance of the colonized ecosystem, it happens that managers take strong control measures to restrict the growth of these populations or to eradicate them.
+Since its presence might impact the balance of the colonized ecosystem, it happens that managers take strong control measures to restrict the growth of these populations or to eradicate them **ref required**.
 Such species are then a good biological model to overcome density-dependent effects and explore the effects of a harvest gradient on productivity.
 
-## How to estimate productivity? {-}
+*How to estimate productivity?*
 
 The productivity parameter of a population can be assessed by monitoring the number of eggs that survives to the age of the first-breeding event.
-To do so, a first step consists in quantifying the reproductive success by estimating nesting and hatching success thanks to surveys in habitats representative of the managed population.
+To do so, a first step consists in quantifying the reproductive success by estimating nesting and hatching success thanks to surveys in habitats representative of the managed population **ref required**.
 The survival of the post-hatched chick to its recruitment can be done by capture mark recapture method [@Blums2004].
@@
 
-In waterfowl, it is generally impossible to differenciate immatures from adults without examining the birds in hand.
+In waterfowl, it is generally impossible to differenciate immatures from adults without examining birds in hand.
 Aging individuals from samples coming from hunting bags is thus another possible approach.
@@
 Secondly, like any predation action, hunting is selective and provides a biased picture of the population age structure in favor of the immatures [@Bellrose1980].
-The hunting bag analysis might then lead to overestimate the productivity and consequently the maximum sustainable harvest rate, with adverse consequences on population management.
+The hunting bag analysis might then lead to overestimate the productivity and consequently the maximum sustainable harvest rate, with adverse consequences on population management **ref required**.
 
-In waterfowl, it is common that adult males display brighter colors than females when  immatures from both sexes display cover-up plumage similar to adult females *ref*.
-Knowing sex ratios, this discrepancy between adult and immature males might be used to estimate the proportion of immature in a waterfowl population just by counting them from distance.
-Such approach is promising because it requires only to count the proportion of birds of the two plumage-kinds to infer an estimate of productivity.
+In waterfowl, it is common that adult males display brighter colors than females when  immatures from both sexes display cover-up plumage similar to adult females **ref required**.
+Knowing sex ratios, this discrepancy between adult and immature males might be used to estimate the proportion of immature in a waterfowl population just by counting them from distance **ref required**.
+Such approach is promising because it requires only to count the proportion of birds of the two plumage kinds to infer an estimate of productivity.
 Unlike the two previous methods, this one is non-invasive and requires a little time investment, which allows to collect long time series and consequently to track productivity variability.
 
-## Methodology of this study {-}
+*Methodology of this study*
 
-From two populations of ruddy duck, an exotic species introduced in Europe in the 50's, this study applies a method of productivity estimation from count data.
+From two populations of ruddy duck, an species introduced in Europe in the 40's, this study develops a Bayesian framework to infer productivity from count data.
+A comparison to the results of the common approach using samples from hunting bags aims at checking if the productivity variability is well tracked and if the hypothesis of the productivity overestimation by sampling methods is consistent with the results.
+Using two other approaches using average population growth rates, one testes the consistency of the order of magnitude of the productivity estimates from the counting method.
 
@@
 
-The best picture stands just before the breeding period, so generally in winter for waterfowl.
+## Biological model
 
-## The studied species
+The ruddy duck is a diving duck species introduced in the United Kingdom in the 40's [@GutierrezExposito2020].
+The first reproduction was observed in the 60's.
+The population rapidely grew to reach more than 5000 individuals spread over the entire country.
+A new population set up in France from the 90's, likely because of an arrival of few individuals from the United Kingdom.
+This population set up around a single wintering spot, the Grand Lieu Lake (geographical coordinates: 47.09°N, 1.67°W).
+Since this species is considered as a threat to the white-headed duck because of the risk of genetic introgression [@MunozFuentes2007], an European plan of eradication has been adopted and control measures were taken in both countries [@GutierrezExposito2020].
 
-The ruddy duck is a species introduced in the United Kingdom in the 50's.
-The first reproduction was observed in the 60's.
-The population rapidely grew until reaching about more than 5000 individuals spread over the entire country.
-A new population set up in France from the 90's, likely because of an arrival of few individuals from the United Kingdom.
-This population set up around a single wintering spot, the Grand Lieu Lake.
-Since this species is considered as a major threat to the white-headed duck because of the risk of hybridization (AEWA report), a European plan of eradication has been concluded (EU report).
-Control measures were taken in both countries from 1999 in the UK and 2004 in France.
-
-## Dichromatism characterization
+Just before the reproduction period, i.e.
+in winter, one can distinguish from shore individuals with male-like plumage from female-like individuals (Figure \@ref(fig:MM1)).
+It is noticeable that the apparent proportion of males in the counts from shore is far below the proportion of males in the samples from hunting bags (Figure \@ref(fig:MM2)).
 
@@
 
-Pictures of male juvenile in winter
-Pictures of male in winter
+Since sex identification of samples is very accurate **ref required**, the counting method misses the identification of a significant part of males in the population.
+In many waterfowl species, immature individuals of both sexes display cover-up plumage similar to adult female plumage **ref required**.
+Consequently, the counts in winter allows adult males to be distinguished from all other individuals (adult females, immature females, and immature males).
+This property can be valuable to estimate the recruitment with a few assumptions.
+One assumes in this study that the immature individuals have a equally balanced sex ratio [@Bellrose1961; @Blums1996; @Wood2021]
+
+Ruddy duck populations from the UK and from France follow similar demographic histories (Figure \@ref(fig:MM1)).
+They grew freely during some years in both countries until 1999.
+Limited control measures have been then applied from 1999 to 2005 in the UK and from 1999 to 2018 in France.
+In both countries, these management measures led to stop the expansion of the ruddy ducks.
+From 2005 in the UK and from 2018 in France, higher control effort led to a rapid depletion in both populations.
+
+```{r MM3, fig.cap = "(ref:MM3cap)"}
+knitr::include_graphics("../Output/plot_2.png")
+```
+
+Such biological model is thus suited to study the response of the harvest pressure on productivity.
 
 ## Data time series
+
+The best picture stands just before the breeding period, so generally in winter for waterfowl.
 
@@
 
-```{r MM3, fig.cap = "(ref:MM3cap)"}
-knitr::include_graphics("../Output/plot_2.png")
-```
-
-## Model
+## Modelling
 
@@
 
-## Result validation using survival analysis
+[@RCT2020] R package Nimble [@Valpine2017]
+
+Bayesian framework is efficient to propagate error through the parameters
+
+We generated three chains of length 25,000, discarding the first 12,500 samples as burn-
+in. Convergence of all chains was assessed using the Gelman and Rubin convergence
+diagnostic (R<1.1, Gelman and Rubin 1992). We fit the models using NIMBLE (de
+Valpine et al. 2017) run from R (R Core Team 2019).
+
+## Comparison with the sampling method
+
+# Result validation using survival analysis
 
@@
 
-## Result validation by comparison to maximum productivity
+## Result validation using maximum productivity proxy
 
@@
 # Discussion
+
+productivity the most sensitive parameter to density dependent effect [@Koons2014]
 
@@
 
+one considers harvest as an impact only adult survival, not on age structure, but we should, oversimplistics
+
 blabla
 
-Nichols -> il faut compter de la façon dont on demande
+@Nichols2006 -> il faut compter de la façon dont on demande
 
@@
 
+Condition of application to other species.
+
 # Aknowledgements {-}
@@
 This work was partly funded by the LIFE Oxyura project (LIFE17 NAT/FR/000942) through the LIFE program.
-This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the Office Français de la Biodiversité, and Jean-Marc Gillier from the Société Nationale pour la Protection de la Nature.
+This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the *Office Français de la Biodiversité*, and Jean-Marc Gillier from the *Société Nationale pour la Protection de la Nature*.
 The authors aknowledge all the contributors of the data collection in the UK and in France, especially: **UK names**, Vincent Fontaine, Denis Lacourpaille, Jules Joly, Justin Potier, Valentin Boniface, Alexis Laroche, and Médéric Lortion, **other names**.
@@
 
-# Supplement 1 {-}
+# (APPENDIX) Supplement {-}
+
+# Supplement
 
@@
 
-(ref:MM2cap) Apparent male proportion in winter counts vs male proportion in adult samples; for the first category, one point corresponds to the proportion of male-like plumage in a population counted in winter; for the second category, the male proportion is estimated from the adult ruddy ducks sampled over one year
+(ref:MM2cap) Apparent male proportion in winter counts vs male proportion in samples; for the first category, one point corresponds to the proportion of male-like plumage in a population counted in winter; for the second category, the male proportion is estimated from all ruddy ducks sampled over one year
 
-(ref:MM3cap) Evolution of the size of the two Ruddy duck populations in light of the data time series; on restricted time periods with available data, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:MM3cap) Evolution of the size of the two Ruddy duck populations in light of the data time series;
 

### 2021-11-03 Code/Model_productivity.R

@@ -58,7 +58,7 @@ raw_plot <- function(para,
     facet_wrap(~ pop, nrow = 1, scales = "free_x") +
     # geom_line(linetype = "dashed", alpha = .5) +
     geom_point() +
-    geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5) +
+    geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5, size = 1.2) +
     scale_y_continuous(limits = if (is.na(ylimits)) {c(0, NA)} else {c(0, ylimits)}) +
     scale_x_date(date_breaks = "2 years", 
                  date_minor_breaks = "1 year", 
@@ -102,9 +102,9 @@ cor_plot <- function(col, title = NULL, limits = NA, breaks = NA) {
     geom_abline(slope = 1, linetype = "dashed") +
     geom_point(aes(color = year)) +
     geom_errorbarh(aes(xmin = !!sym(col_s[1]), xmax = !!sym(col_s[3]), color = year), 
-                   height = 0) +
+                   height = 0, alpha = .5, size = 1.2) +
     geom_errorbar(aes(ymin = !!sym(col_c[1]), ymax = !!sym(col_c[3]), color = year), 
-                  width = 0) +
+                  width = 0, alpha = .5, size = 1.2) +
     labs(title = title, 
          y = "counting method", x = "sampling method") +
     scale_x_continuous(expand = expansion(mult = c(0, 0)), 
@@ -777,7 +777,7 @@ JuvOut4 %>%
   ggplot(aes(x = method, y = `50%`, group = pop, color = method)) +
   facet_wrap(~ pop, nrow = 1, scales = "free_x") +
   geom_point() +
-  geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`)) +
+  geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5, size = 1.2) +
   geom_text(aes(x = method, y = .365, label = txt), color = "#de2d26", angle = 90) +
   scale_y_continuous(
     limits = c(0, .7),
@@ -796,7 +796,7 @@ JuvOut4 %>%
   ggplot(aes(x = method, y = `50%`, group = pop, color = method)) +
   facet_wrap(~ pop, nrow = 1, scales = "free_x") +
   geom_point() +
-  geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`)) +
[…]

### 2021-11-16 Code/Draft_ruddy_duck_productivity.Rmd
update

@@
 It is often skewed towards juveniles and can lead to an overestimation of the productivity.
-In waterfowl, adult males usually display brighter colours than immatures and females.
+In waterfowl, adult males usually display brighter colours than females and immatures from both sexes.
 This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
@@
 in winter, one can distinguish from shore individuals with male-like plumage from female-like individuals (Figure \@ref(fig:MM1)).
-It is noticeable that the apparent proportion of males in the counts from shore is far below the proportion of males in the samples from hunting bags (Figure \@ref(fig:MM2)).
+It is noticeable that the apparent proportion of males in the counts from shore is far below the proportion of males in the samples from the hunting bag (Figure \@ref(fig:MM2)).
 
@@
 
-We generated three chains of length 25,000, discarding the first 12,500 samples as burn-
+We generated three chains of length 25000, discarding the first 12500 samples as burn-
 in. Convergence of all chains was assessed using the Gelman and Rubin convergence

### 2021-11-16 Code/Draft_ruddy_duck_productivity.Rmd
update

@@
 In using two populations of ruddy duck, *Oxyura oxyura*, this study develops a Bayesian method to estimate the productivity from count data.
-The results are compared to productivity estimates from samples and are discussed by testing their consistency with the population growth rates.
+The results are compared to productivity estimates from samples and are discussed by testing their consistency to support the observed population growth rates.
 The results suggest that the counting method is a versatile tool to estimate the productivity if the species is monitored during the appropriate time window.
+For the relevant waterfowl species, these results argue for considering a two-category protocol in winter count surveys.
 output:
@@
 One defines the productivity of a breeding population as the average number of recruits produced per each breeder [@Johnson1987].
-This parameter might sometimes be called "annual fertility" [@Koons2014], or "reproductive success".
+This parameter might sometimes be called "annual fertility" [@Koons2014], or "reproductive rate" [@Cooch2014], or "recruitment rate" [@Flint2015].
 If one considers a species reaching its sexual maturity as early as its first year, all individuals are mature at the reproduction season.
@@
 
-The productivity parameter is complex because it is composed of two main sub-parameters: the reproductive success and the juvenile survival.
+The productivity parameter is complex because it is composed of two main sub-parameters: the reproductive success followed by the juvenile survival.
 Both sub-parameters can be split into other sub-parameters.
@@
 However, the results of this study does not necessarily reflect a general reality because several sources of variations of the reproductive success are controlled in the experimental plan of this study.
-The productivity as defined in the present study is thus a parameter integrating many sources of variability affecting all life stages from the egg to the young adult trying to breed for its first time.
+The productivity as defined in the present study is thus a parameter integrating many sources of variability affecting all life stages from the egg to the young adult trying to breed for its first time [@Koons2017].
 
@@
 
-The management of an exploited population generally aims at preserving its conservation status, generally by targeting a population size **ref required**.
-To do so, a first strategy is to test iteratively management rules to reach progressively an optimum.
+Population conservation, harvesting, and control require tools in population management to reach targets such as a population size for example [@Shea1998].
+To do so, a first "blind" strategy is to test iteratively management rules to reach progressively an optimum.
 Such approach can be time-consuming given the intrinsic variability of the population response to its changing environment.
-It might also suffer from a rejection of the stakeholders because of the lack of transparency of its arbitrary process **ref required** [@Nichols2000]?.
+It might also suffer from a rejection of the stakeholders because of the lack of transparency of an arbitrary process [@Williams2013].
 Another approach is to try to predict the growth rate of the population as a function of the harvest rate and other environment inputs.
-Such modelling approach can provide clues on the recovering time of a population upon a variety of managment rules **ref required**.
+Such modelling approach can provide clues on the recovering time of a population upon a variety of managment rules [@Otis2006].
 These decision-support tools are then key inputs to status on the best management rules to implement.
-The growth rate of a population can be simply defined as the sum of the adult survival and the productivity, see Equation \@ref(eq:GR) in Section \@ref(MM).
-The adult survival is generally estimated by capture mark recapture methods **ref required**.
-If the harvest pressure behaves as an additive mortality to the natural mortality, it is quite straightforward to predict its impact of the adult survival.
-The key to predict the growth rate of a population is then to estimate the average value and the variability of the productivity and to understand how this parameter is impacted by a harvest pressure.
+The growth rate of a population depends on the adult survival and the productivity, see Equation \@ref(eq:GR) in Section \@ref(MM).
+The adult survival is generally estimated by mark recapture/recovery methods [@Lebreton2001].
+If the harvest pressure behaves as an additive mortality to the natural mortality for adults [e.g.
+@Iverson2013], it is quite straightforward to predict its impact of the adult survival.
+The key to predict the growth rate of a population is then to estimate the average value and the variability of the productivity and to understand how this parameter is impacted by a harvest pressure [@Anders2005].
 
@@
 The population is then less productive because the productivity parameter is affected by density-dependent effects [@Gunnarsson2013].
-For an endemic species evolving in stable environmental conditions, a balance between the productivity and the adult mortality rate explains the stabilization of the number of individuals **ref required**.
+For an endemic species evolving in stable environmental conditions, a balance between the productivity and the adult mortality rate explains the stabilization of the number of individuals [@Flint2015].
 
@@
 Intuitively, one expects that such a hunted population would observe lower adult survival and productivity and then would rapidly deplete to its extinction.
-However, hunting waterfowl has a long history, and most of these species have persisted over time **ref required**.
+However, hunting waterfowl has a long history, and most of these species have persisted over time [@Cooch2014].
 So paradoxically, hunting immature individuals does not always induce a decrease of productivity.
-A reason could be that the harvest pressure reduces the competition for space and food **ref required**.
-The density-dependent effects affecting the productivity in case of high population density are then reduced by the harvest mortality.
-In conclusion, a limited exploitation of a population is expected to increase both adult mortality and productivity.
-A balance between these two parameters can be theoritically reached if the population is exploited to a constant and moderate harvest rate **ref required**.
-The size of a newly exploited population in stable environmental conditions should thus reach a different equilibrium, which is expected to be lower than the size without exploitation **ref required**.
+The hetereogeneity of quality among immature individuals might explain this by inducing a compensation process if the harvest selects mostly individuals with a low reproductive potential [@Lindberg2013; @Gimenez2017].
+The productivity can even be enhanced because the harvest pressure reduces the densities and potentially the competition for space and food [@Nummi2015; @Peron2012].
+In conclusion, a limited exploitation of a population observing heteregeneity and/or density-dependent effect might increase or at least not affect the productivity.
+A balance between adult mortality and productivity can be theoritically reached if the population is exploited to a constant and moderate harvest rate [@Tsikliras2018].
+The size of a newly exploited population in stable environmental conditions should thus reach a different equilibrium, which is expected to be lower than the size without exploitation.
 
@@
 For such question, the carrying capacity of the habitat is theoritically not reached.
-So a manager should mostly focus on collecting knowledge on the maximum productivity and its variability by removing the potential density-dependent effects occurring only for the highest theoritical densities.
+So a manager should mostly focus on collecting knowledge on the maximum productivity and its variability by removing the potential density-dependent effects occurring only for the highest theoritical densities [@Peron2013].
 
@@
 First, it is impossible to explore the productivity when there is no harvest whereas it is necessary to predict what would happen if a moratorium seems required to let a population recover, or to calibrate a priori the harvest effort necessary to control an invasive species [@Smith2005].
-Secondly, like any predation action, hunting is selective and provides a biased picture of the population age structure in favor of the immatures [@Bellrose1980].
-The hunting bag analysis might then lead to overestimate the productivity and consequently the maximum sustainable harvest rate, with adverse consequences on population management **ref required**.
+Secondly, like any predation action, hunting is selective and provides a biased picture of the population age structure in favor of the immatures [@Bellrose1980; @Fox2014].
+The hunting bag analysis might then lead to overestimate the productivity and consequently the maximum sustainable harvest rate, with adverse consequences on population management ; [@Fox2014].
 

### 2021-11-18 Code/Draft_ruddy_duck_recruitment_rate.Rmd
terminology modif in titles

--- 2021-11-16 Code/Draft_ruddy_duck_productivity.Rmd
@@
 link-citations: true
-title: "Productivity estimation in waterfowl using a non-invasive method"
+title: "Recruitment rate estimation in waterfowl using a non-invasive method"
 author: 
@@
 the number of young adults reproducing for the first time, is a key indicator to describe the renewal of a population, and therefore an essential tool in population management.
-The productivity, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
-The proportion of young adults in a waterfowl population, which is a main step in estimating productivity, is often estimated from game-hunting samples.
+The recruitment rate, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
+The proportion of young adults in a waterfowl population, which is a main step in estimating recruitment rate, is often estimated from game-hunting samples.
 However, this proportion is not accessible in years without harvest, or with a low harvest rate.
 Moreover, the age-structure in the harvest samples, commonly called hunting bag, does not necessarily reflect the underlying age-structure of the population.
-It is often skewed towards juveniles and can lead to an overestimation of the productivity.
+It is often skewed towards juveniles and can lead to an overestimation of the recruitment rate.
 In waterfowl, adult males usually display brighter colours than females and immatures from both sexes.
 This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
-Such information can be used to infer the proportion of young adults, and consequently the productivity.
-In using two populations of ruddy duck, *Oxyura oxyura*, this study develops a Bayesian method to estimate the productivity from count data.
-The results are compared to productivity estimates from samples and are discussed by testing their consistency to support the observed population growth rates.
-The results suggest that the counting method is a versatile tool to estimate the productivity if the species is monitored during the appropriate time window.
+Such information can be used to infer the proportion of young adults, and consequently the recruitment rate.
+In using two populations of ruddy duck, *Oxyura oxyura*, this study develops a Bayesian method to estimate the recruitment rate from count data.
+The results are compared to recruitment rate estimates from samples and are discussed by testing their consistency to support the observed population growth rates.
+The results suggest that the counting method is a versatile tool to estimate the recruitment rate if the species is monitored during the appropriate time window.
 For the relevant waterfowl species, these results argue for considering a two-category protocol in winter count surveys.
@@
   bookdown::pdf_document2:
-    latex_engine: pdflatex
+    latex_engine: xelatex
     toc: false
@@
                       fig.align = "center",
-                      fig.pos = 'H') # pour tenir la position
+                      fig.pos = 'H') # hold position
 
@@
 
-One defines the productivity of a breeding population as the average number of recruits produced per each breeder [@Johnson1987].
-This parameter might sometimes be called "annual fertility" [@Koons2014], or "reproductive rate" [@Cooch2014], or "recruitment rate" [@Flint2015].
+One defines the recruitment rate of a breeding population as the average number of recruits produced per each breeder [@Flint2015].
+This parameter might sometimes be called "annual fertility" [@Koons2014], "reproductive rate" [@Cooch2014], "reproductive success" [@Etterson2011], or "productivity" [@Johnson1987].
 If one considers a species reaching its sexual maturity as early as its first year, all individuals are mature at the reproduction season.
-The productivity is thus the ratio of the number of young adults at the reproduction season of year $t$, i.e.
+The recruitment rate is thus the ratio of the number of young adults at the reproduction season of year $t$, i.e.
 the recruits, to the number of all individuals at the reproduction season of the previous year, $t - 1$.
+The declination for species with age at first reproduction higher than 1 follows the same process [see @Robertson2008].
 
-The productivity parameter is complex because it is composed of two main sub-parameters: the reproductive success followed by the juvenile survival.
+The recruitment rate parameter is complex because it is composed of two main sub-parameters : the fecundity followed by the juvenile survival [@Etterson2011].
 Both sub-parameters can be split into other sub-parameters.
-For instance, the reproductive success is the product of the nesting success and the hatching success.
-The juvenile survival is the combination of the post-hatching survival and the fledging juvenile survival.
-According to @Blums2004, the recruitment for diving ducks is not related to the reproductive success, but depends mostly on the juvenile survival, which is driven by weather conditions.
-However, the results of this study does not necessarily reflect a general reality because several sources of variations of the reproductive success are controlled in the experimental plan of this study.
-The productivity as defined in the present study is thus a parameter integrating many sources of variability affecting all life stages from the egg to the young adult trying to breed for its first time [@Koons2017].
+The fecundity can be defined by the product of the nesting success, the average clutch size, and the hatching success.
+The juvenile survival can be defined as the product of the post-hatching survival and the post-fledging survival.
+The recruitment rate as defined in the present study is thus a parameter integrating multiple life steps from the egg to the young adult attempting to breed for its first time [@Koons2017].
 
-*The use of productivity in population management*
+*The use of recruitment rate in population management*
 
@@
 These decision-support tools are then key inputs to status on the best management rules to implement.
-The growth rate of a population depends on the adult survival and the productivity, see Equation \@ref(eq:GR) in Section \@ref(MM).
+The growth rate of a population depends on the adult survival and the recruitment rate, see Equation \@ref(eq:GR) in Section \@ref(MM).
 The adult survival is generally estimated by mark recapture/recovery methods [@Lebreton2001].
@@
 @Iverson2013], it is quite straightforward to predict its impact of the adult survival.
-The key to predict the growth rate of a population is then to estimate the average value and the variability of the productivity and to understand how this parameter is impacted by a harvest pressure [@Anders2005].
+According to @Blums2004, the recruitment for diving ducks is poorly related to fecundity, but depends mostly on juvenile survival, which is driven by weather conditions.
+Studies on other birds concludes also on the weak correlation between fecundity and recruitment [e.g.
+@Murray2000].
+The difficulty to track juvenile survival [@Schmidt2008] demonstrates the interest to focus on the recruitment rate to predict population growth rate over long time periods.
+The key to predict the growth rate of a population is then to estimate the average value and the variability of the recruitment rate, potentially influenced by the harvest pressure [@Anders2005].
 
-*The variability of the productivity*
+*The variability of the recruitment rate*
 
 A waterfowl species that is released in a favourable habitat will naturally expand  [@Malthus1872].
-The growth of such a population reflects that the productivity parameter is higher than the adult mortality rate.
+The growth of such a population reflects that the recruitment rate parameter is higher than the adult mortality rate.
 This expansion is limited by the carrying capacity of the habitat [@Sayre2008].
 When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity because the individuals compete for space or/and food [@Nummi2015], which prevents the population from expanding.
-This competition commonly induces a lower reproductive success or/and a higher juvenile mortality.
-The population is then less productive because the productivity parameter is affected by density-dependent effects [@Gunnarsson2013].
-For an endemic species evolving in stable environmental conditions, a balance between the productivity and the adult mortality rate explains the stabilization of the number of individuals [@Flint2015].
+This competition commonly induces a lower fecundity or/and a higher juvenile mortality.
+The population renewal is slowed because the recruitment rate parameter is affected by density-dependent effects [@Gunnarsson2013].
+For an endemic species evolving in stable environmental conditions, a balance between the recruitment rate and the adult mortality rate explains the stabilization of the number of individuals [@Flint2015].
 
 Hunting waterfowl induces adult and juvenile mortality [@Bellrose1980].
-Intuitively, one expects that such a hunted population would observe lower adult survival and productivity and then would rapidly deplete to its extinction.
+Intuitively, one expects that such a hunted population would observe lower adult survival and recruitment rate and then would rapidly deplete to its extinction.
 However, hunting waterfowl has a long history, and most of these species have persisted over time [@Cooch2014].
-So paradoxically, hunting immature individuals does not always induce a decrease of productivity.
+So paradoxically, hunting immature individuals does not always induce a decrease of recruitment rate.
 The hetereogeneity of quality among immature individuals might explain this by inducing a compensation process if the harvest selects mostly individuals with a low reproductive potential [@Lindberg2013; @Gimenez2017].
-The productivity can even be enhanced because the harvest pressure reduces the densities and potentially the competition for space and food [@Nummi2015; @Peron2012].
-In conclusion, a limited exploitation of a population observing heteregeneity and/or density-dependent effect might increase or at least not affect the productivity.
-A balance between adult mortality and productivity can be theoritically reached if the population is exploited to a constant and moderate harvest rate [@Tsikliras2018].
+The recruitment rate can even be enhanced because the harvest pressure reduces the densities and potentially the competition for space and food [@Nummi2015; @Peron2012].
+In conclusion, a limited exploitation of a population observing heteregeneity and/or density-dependent effect might increase or at least not affect the recruitment rate.
+A balance between adult mortality and recruitment rate can be theoritically reached if the population is exploited to a constant and moderate harvest rate [@Tsikliras2018].
 The size of a newly exploited population in stable environmental conditions should thus reach a different equilibrium, which is expected to be lower than the size without exploitation.
 
-The reproductive success and the juvenile survival, the two components of the productivity, are very sensitive to weather conditions [@Blums2004; @Folliot2017].
-The environment conditions are thus an additional source of variability of the productivity.
-The response of the productivity to a harvest pressure is thus not straightforward.
+The fecundity and the juvenile survival, the two components of the recruitment rate, are very sensitive to weather conditions [@Blums2004; @Folliot2017].
+The environment conditions are thus an additional source of variability of the recruitment rate.
+The response of the recruitment rate to a harvest pressure is thus not straightforward.
 It is noisy because of its sensitivity to environmental factors.
@@
 
-*What part of the productivity gradient is necessary for population management?*
+*What part of the recruitment rate gradient is necessary for population management?*
 
-In population mangement, a common question for a threaten or an invasive species is to determine the measure that will allow to reach a targetted population size/density in a limited time period **ref required**.
+In population mangement, a common question for a threaten or an invasive species is to determine the measure that will allow to reach a targetted population size/density in a limited time period [@Shea1998].
 For such question, the carrying capacity of the habitat is theoritically not reached.
-So a manager should mostly focus on collecting knowledge on the maximum productivity and its variability by removing the potential density-dependent effects occurring only for the highest theoritical densities [@Peron2013].
+So a manager should mostly focus on collecting knowledge on the maximum recruitment rate and its variability by removing the potential density-dependent effects occurring only for the highest theoritical densities [@Peron2013, @Eraud2021].
 
-*What is the ideal biological model to explore the gradient of the productivity?*
+*What is the ideal biological model to explore the gradient of the recruitment rate?*
 
-Hunting waterfowl has a long history, but few species have observed long-term moratorium **ref required**.
-So there are few opportunities to explore the response of the productivity to a gradient of pressure for such species.
-Alien species are often introduced during involuntary release events **ref required**.
+Hunting waterfowl has a long history, but the biological effects of long-term moratorium have been poorly surveyed [@MartinezAbrain2013].
+So there are few opportunities to explore the response of the recruitment rate to a gradient of pressure for such species.
+Alien vertebrate species are often introduced during  unintentional or intentional release events [@Saul2016].
 They observe commonly a first period of colonisation on new territories without management measures.
-Since its presence might impact the balance of the colonized ecosystem, it happens that managers take strong control measures to restrict the growth of these populations or to eradicate them **ref required**.
-Such species are then a good biological model to overcome density-dependent effects and explore the effects of a harvest gradient on productivity.
+Since its presence might impact the balance of the colonized ecosystem, it happens that managers take strong control measures to restrict the growth of these populations or to eradicate them [@Oficialdegui2020].
+Such species are then a good biological model to overcome density-dependent effects and explore the effects of a harvest gradient on recruitment rate.
 
-*How to estimate productivity?*
+*How to estimate recruitment rate?*
 
-The productivity parameter of a population can be assessed by monitoring the number of eggs that survives to the age of the first-breeding event.
-To do so, a first step consists in quantifying the reproductive success by estimating nesting and hatching success thanks to surveys in habitats representative of the managed population **ref required**.
-The survival of the post-hatched chick to its recruitment can be done by capture mark recapture method [@Blums2004].
-Even if such approach provides valuable information on the dynamics of the recruitment and can highlight the bottlenecks, the fieldwork required to estimate both components is too time-consuming to generalize the method for waterfowl population management.
+The recruitment rate parameter of a population can be assessed first by evaluating the fecundity of the breeding population and then by monitoring the survival from the egg life stage to the first-breeding event.
+To do so, it is necessary to monitor nests in habitats representative of the managed population [@Blums2004].
+The survival the chick to its recruitment can be done by capture mark recapture method quite tricky to set up [@Blums2004; Schmidt2008].
+Even if such approach provides valuable information on the dynamics of the recruitment and can highlight the bottlenecks, the fieldwork required to estimate both components is too time-consuming to generalize such approach for waterfowl population management.
 
@@
 Such approach suffers from two caveats.
-First, it is impossible to explore the productivity when there is no harvest whereas it is necessary to predict what would happen if a moratorium seems required to let a population recover, or to calibrate a priori the harvest effort necessary to control an invasive species [@Smith2005].
+First, it is impossible to assess the recruitment rate when there is no harvest whereas it is necessary to predict what would happen if a moratorium seems required to allow a population to recover, or to calibrate a priori the harvest effort necessary to control an invasive species [@Smith2005].
 Secondly, like any predation action, hunting is selective and provides a biased picture of the population age structure in favor of the immatures [@Bellrose1980; @Fox2014].
-The hunting bag analysis might then lead to overestimate the productivity and consequently the maximum sustainable harvest rate, with adverse consequences on population management ; [@Fox2014].
+The hunting bag analysis might then lead to overestimate the recruitment rate and consequently the maximum sustainable harvest rate, with adverse consequences on population management [@Fox2014].
 
-In waterfowl, it is common that adult males display brighter colors than females when  immatures from both sexes display cover-up plumage similar to adult females **ref required**.
-Knowing sex ratios, this discrepancy between adult and immature males might be used to estimate the proportion of immature in a waterfowl population just by counting them from distance **ref required**.
-Such approach is promising because it requires only to count the proportion of birds of the two plumage kinds to infer an estimate of productivity.
-Unlike the two previous methods, this one is non-invasive and requires a little time investment, which allows to collect long time series and consequently to track productivity variability.
+In waterfowl, it is common that adult males display brighter colors than females when  immatures from both sexes display cover-up plumage similar to adult females [@Johnson1999].
+Knowing sex ratios, this discrepancy between adult and immature males might be used to estimate the proportion of immature in a waterfowl population just by counting them from distance [@Smith2001].
+Such approach is promising because it requires only to count the proportion of birds of the two plumage kinds to infer the recruitment rate.
+Unlike the two previous methods, this one is non-invasive and requires a little time investment, which allows to collect long time series and consequently to track recruitment rate variability.
 
@@
 
-From two populations of ruddy duck, an species introduced in Europe in the 40's, this study develops a Bayesian framework to infer productivity from count data.
-A comparison to the results of the common approach using samples from hunting bags aims at checking if the productivity variability is well tracked and if the hypothesis of the productivity overestimation by sampling methods is consistent with the results.
-Using two other approaches using average population growth rates, one testes the consistency of the order of magnitude of the productivity estimates from the counting method.
+From two populations of ruddy duck, an species introduced in Europe in the 40's, this study develops a Bayesian framework to infer recruitment rate from count data.
+A comparison to the results of the common approach using samples from hunting bags aims at checking if the recruitment rate variability is well tracked and if the hypothesis of the recruitment rate overestimation by sampling methods is consistent with the results.
+Using two other approaches using average population growth rates, one testes the consistency of the order of magnitude of the recruitment rate estimates from the counting method.
 
@@
 A new population set up in France from the 90's, likely because of an arrival of few individuals from the United Kingdom.
-This population set up around a single wintering spot, the Grand Lieu Lake (geographical coordinates: 47.09°N, 1.67°W).
+This population set up around a single wintering spot, the Grand Lieu Lake (47.09°N, 1.67°W).
 Since this species is considered as a threat to the white-headed duck because of the risk of genetic introgression [@MunozFuentes2007], an European plan of eradication has been adopted and control measures were taken in both countries [@GutierrezExposito2020].
@@
 This property can be valuable to estimate the recruitment with a few assumptions.
-One assumes in this study that the immature individuals have a equally balanced sex ratio [@Bellrose1961; @Blums1996; @Wood2021]
+One assumes in this study that the immature individuals have a equally balanced sex ratio [@Bellrose1961; @Blums1996; @Wood2021].
 
 Ruddy duck populations from the UK and from France follow similar demographic histories (Figure \@ref(fig:MM1)).
-They grew freely during some years in both countries until 1999.
+They grew freely in both countries until 1999.
 Limited control measures have been then applied from 1999 to 2005 in the UK and from 1999 to 2018 in France.
@@
 
-Such biological model is thus suited to study the response of the harvest pressure on productivity.
+Such biological model is thus suited to study the response of the harvest pressure on recruitment rate.
 
@@
 the recruitment in year $t$.
-The growth rate of a population is thus the sum of the productivity and the adult survival rate.
+The growth rate of a population is thus the sum of the recruitment rate and the adult survival rate.
 
@@
 realistic survival? 
-if > 1, estimated productivity is too low to support the maximum growth rate
+if > 1, estimated recruitment rate is too low to support the maximum growth rate
 in the range of similar species in litterature -> satisfying
@@
 
-## Result validation using maximum productivity proxy
+## Result validation using maximum recruitment rate proxy
 
@@
 
-## Productivity variability
+## recruitment rate variability
 
@@
 
-productivity the most sensitive parameter to density dependent effect [@Koons2014]
+According to @Blums2004, the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, which is driven by weather conditions.
+Studies on other birds also conclude of the poor correlation between fecundity and recruitment [e.g.
+@Murray2000].
+
+age proportion -> [0, 1]
+recruitment rate [@, + inf] donc plus sensitive
+
+recruitment rate the most sensitive parameter to density dependent effect [@Koons2014]
 
@@
 
-Productivité: 
+recruitment rate: 
 Stabilité grosse pop, 
@@
 
-Au dela du fait que la croissance de la pop est la somme de la survie et de la productivité, 
-intéressant de parler de la variability de la survie vs variabilité de la productivity 
+Au dela du fait que la croissance de la pop est la somme de la survie et de la recruitment rate, 
+intéressant de parler de la variability de la survie vs variabilité de la recruitment rate 
 pour parler de risk management.
 
-Productivity: drop from 2009
+recruitment rate: drop from 2009
 
@@
 
-If the maximum productivity is not reached under optimal conditions (maximum population size not reached, and limited harvest pressure), it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the productivity.
+If the maximum recruitment rate is not reached under optimal conditions (maximum population size not reached, and limited harvest pressure), it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the recruitment rate.
 
@@
 
-(ref:R1cap) (A) Estimates of productivity, i.e.
-average number of recruits produced per breeder; (B) Comparison of the productivity values between the two estimation methods; bars define the 95% confidence intervals
+(ref:R1cap) (A) Estimates of recruitment rate, i.e.
+average number of recruits produced per breeder; (B) Comparison of the recruitment rate values between the two estimation methods; bars define the 95% confidence intervals
 
 (ref:R2cap) (A) Estimates of the population growth rate on time intervals with data availability.
-This aims at producing survival estimates to discuss the reliability of the methods to estimate productivity; (B) Growth rate estimates; bars define the 95% confidence intervals
+This aims at producing survival estimates to discuss the reliability of the methods to estimate recruitment rate; (B) Growth rate estimates; bars define the 95% confidence intervals
 
-(ref:R3cap) (A) Average productivity estimates on the time intervals defined for survival analysis; (B) Average survival estimates deduced from growth rate and productivity estimates on the time intervals; bars define the 95% confidence intervals
+(ref:R3cap) (A) Average recruitment rate estimates on the time intervals defined for survival analysis; (B) Average survival estimates deduced from growth rate and recruitment rate estimates on the time intervals; bars define the 95% confidence intervals
 
@@
 
-(ref:R5cap) Comparison between a proxy of maximum productivity and the average productivity estimated following *counting* and *sampling* methods; the proxy is estimated during the maximum growth periods by combining the growth rate to the hypothesis of a 100% adult survival (*max growth*); this hypothesis implies that the true maximum of productivity is likely higher; bars define the 95% confidence intervals
+(ref:R5cap) Comparison between a proxy of maximum recruitment rate and the average recrutiment rate estimated following *counting* and *sampling* methods; the proxy is estimated during the maximum growth periods by combining the growth rate to the hypothesis of a 100% adult survival (*max growth*); this hypothesis implies that the true maximum of recruitment rate is likely higher; bars define the 95% confidence intervals
 

### 2021-11-24 Code/Draft_ruddy_duck_recruitment_rate.Rmd
result writing

@@
 lang: en-EN
-bibliography: ../../../../Biblio/JabRef/Library_OFB.bib
-csl: ../../../../Biblio/JabRef/Styles/APA.csl
+bibliography: C:/Users/adrien.tableau/Dropbox/JabRef/Library_OFB.bib
+csl: https://www.zotero.org/styles/apa
 link-citations: true
@@
 - Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr - contribution in data collection]
-- Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthieu.guillemain@ofb.gouv.fr - contribution in writing]
+- Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthi(UK).guillemain@ofb.gouv.fr - contribution in writing]
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-abstract: The response of a waterfowl population to a harvest pressure depends on its capacity to renew.
+abstract: "The response of a waterfowl population to a harvest pressure depends on its capacity to renew.
 The recruitment, i.e.
@@
 The recruitment rate, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
-The proportion of young adults in a waterfowl population, which is a main step in estimating recruitment rate, is often estimated from game-hunting samples.
+The proportion of young adult in a waterfowl population, which is a main step in estimating recruitment rate, is often estimated from game-hunting samples.
 However, this proportion is not accessible in years without harvest, or with a low harvest rate.
-Moreover, the age-structure in the harvest samples, commonly called hunting bag, does not necessarily reflect the underlying age-structure of the population.
-It is often skewed towards juveniles and can lead to an overestimation of the recruitment rate.
+Moreover, the age-structure in the harvest samples does not necessarily reflect the underlying age-structure of the population.
+It is often skewed towards immatures and can lead to an overestimation of the recruitment rate.
 In waterfowl, adult males usually display brighter colours than females and immatures from both sexes.
 This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
-Such information can be used to infer the proportion of young adults, and consequently the recruitment rate.
-In using two populations of ruddy duck, *Oxyura oxyura*, this study develops a Bayesian method to estimate the recruitment rate from count data.
-The results are compared to recruitment rate estimates from samples and are discussed by testing their consistency to support the observed population growth rates.
+Such information can be used to infer the proportion of immatures, and consequently the recruitment rate.
+In using two populations of ruddy duck, *Oxyura oxyura*, this study introduces a Bayesian method to estimate the recruitment rate from count data.
+The results are compared to recruitment rate estimates from samples and their consistency to support the observed population growth rates is tested.
 The results suggest that the counting method is a versatile tool to estimate the recruitment rate if the species is monitored during the appropriate time window.
-For the relevant waterfowl species, these results argue for considering a two-category protocol in winter count surveys.
+The results argue for considering a two-category count protocol in winter surveys of waterfowl species when adult males can be identified.
+\\par \\textbf{Keywords:} fecundity - productivity - reproductive success - age ratio - juvenile - duck"
 output:
@@
 Another approach is to try to predict the growth rate of the population as a function of the harvest rate and other environment inputs.
-Such modelling approach can provide clues on the recovering time of a population upon a variety of managment rules [@Otis2006].
+Such modelling approach can provide clues on the recovering time of a population upon a variety of management rules [@Otis2006].
 These decision-support tools are then key inputs to status on the best management rules to implement.
@@
 If the harvest pressure behaves as an additive mortality to the natural mortality for adults [e.g.
-@Iverson2013], it is quite straightforward to predict its impact of the adult survival.
+@Iverson2013], it is quite straightforward to predict its impact on the adult survival.
+
 According to @Blums2004, the recruitment for diving ducks is poorly related to fecundity, but depends mostly on juvenile survival, which is driven by weather conditions.
@@
 @Murray2000].
-The difficulty to track juvenile survival [@Schmidt2008] demonstrates the interest to focus on the recruitment rate to predict population growth rate over long time periods.
+The difficulty to track juvenile survival [@Schmidt2008] demonstrates the interest to focus on the recruitment rate when accessible.
 The key to predict the growth rate of a population is then to estimate the average value and the variability of the recruitment rate, potentially influenced by the harvest pressure [@Anders2005].
@@
 This competition commonly induces a lower fecundity or/and a higher juvenile mortality.
-The population renewal is slowed because the recruitment rate parameter is affected by density-dependent effects [@Gunnarsson2013].
+The population renewal is slowed down because the recruitment rate parameter is affected by density-dependent effects [@Gunnarsson2013].
 For an endemic species evolving in stable environmental conditions, a balance between the recruitment rate and the adult mortality rate explains the stabilization of the number of individuals [@Flint2015].
@@
 So paradoxically, hunting immature individuals does not always induce a decrease of recruitment rate.
-The hetereogeneity of quality among immature individuals might explain this by inducing a compensation process if the harvest selects mostly individuals with a low reproductive potential [@Lindberg2013; @Gimenez2017].
+The heterogeneity of quality among immature individuals might explain this by inducing a compensation process if the harvest selects mostly individuals with a low reproductive potential [@Lindberg2013; @Gimenez2017].
 The recruitment rate can even be enhanced because the harvest pressure reduces the densities and potentially the competition for space and food [@Nummi2015; @Peron2012].
-In conclusion, a limited exploitation of a population observing heteregeneity and/or density-dependent effect might increase or at least not affect the recruitment rate.
-A balance between adult mortality and recruitment rate can be theoritically reached if the population is exploited to a constant and moderate harvest rate [@Tsikliras2018].
+In conclusion, a limited exploitation of a population observing heterogeneity and/or density-dependent effect might increase or at least not affect the recruitment rate.
+Under moderate exploitation, a new balance between adult mortality and recruitment rate can be theoretically reached [@Tsikliras2018].
 The size of a newly exploited population in stable environmental conditions should thus reach a different equilibrium, which is expected to be lower than the size without exploitation.
+This process explains why the exploitation of many waterfowl species is sustainable and has been lasting over a long time.
 
@@
 
-In population mangement, a common question for a threaten or an invasive species is to determine the measure that will allow to reach a targetted population size/density in a limited time period [@Shea1998].
-For such question, the carrying capacity of the habitat is theoritically not reached.
-So a manager should mostly focus on collecting knowledge on the maximum recruitment rate and its variability by removing the potential density-dependent effects occurring only for the highest theoritical densities [@Peron2013, @Eraud2021].
+In population management, a common question for a threaten or an invasive species is to determine the measure that will allow to reach a targeted population size/density in a limited time period [@Shea1998].
+For such question, the carrying capacity of the habitat is theoretically not reached.
+So a manager should mostly focus on collecting knowledge on the maximum recruitment rate and its variability by removing the potential density-dependent effects occurring only for the highest theoretical densities [@Peron2013, @Eraud2021].
 
@@
 So there are few opportunities to explore the response of the recruitment rate to a gradient of pressure for such species.
-Alien vertebrate species are often introduced during  unintentional or intentional release events [@Saul2016].
-They observe commonly a first period of colonisation on new territories without management measures.
+Alien vertebrate species are commonly introduced during release events [@Saul2016].
+They usually observe a first period of colonisation on new territories without management measures.
 Since its presence might impact the balance of the colonized ecosystem, it happens that managers take strong control measures to restrict the growth of these populations or to eradicate them [@Oficialdegui2020].
@@
 To do so, it is necessary to monitor nests in habitats representative of the managed population [@Blums2004].
-The survival the chick to its recruitment can be done by capture mark recapture method quite tricky to set up [@Blums2004; Schmidt2008].
-Even if such approach provides valuable information on the dynamics of the recruitment and can highlight the bottlenecks, the fieldwork required to estimate both components is too time-consuming to generalize such approach for waterfowl population management.
+The survival of a chick to its recruitment can be done by capture mark recapture method quite tricky to set up [@Blums2004; Schmidt2008].
+Even if such approach provides valuable information on the dynamics of the recruitment and can highlight the bottlenecks, the fieldwork required to estimate both components is too time-consuming to generalize such approach for waterfowl population management on a long-term horizon.
 
-In waterfowl, it is generally impossible to differenciate immatures from adults without examining birds in hand.
-Aging individuals from samples coming from hunting bags is thus another possible approach.
+In waterfowl, it is generally impossible to differentiate immatures from adults without examining birds in hand.
+Aging individuals from samples coming from hunting bags is thus another opportunity to estimate the recruit proportion.
 Such approach suffers from two caveats.
 First, it is impossible to assess the recruitment rate when there is no harvest whereas it is necessary to predict what would happen if a moratorium seems required to allow a population to recover, or to calibrate a priori the harvest effort necessary to control an invasive species [@Smith2005].
-Secondly, like any predation action, hunting is selective and provides a biased picture of the population age structure in favor of the immatures [@Bellrose1980; @Fox2014].
-The hunting bag analysis might then lead to overestimate the recruitment rate and consequently the maximum sustainable harvest rate, with adverse consequences on population management [@Fox2014].
+Secondly, like any predation action, hunting is selective and provides a biased picture of the population age structure in favour of the immatures [@Bellrose1980; @Fox2014].
+The hunting bag analysis might then lead to overestimate the recruitment rate and consequently the maximum sustainable harvest rate, with adverse consequences on population conservation [@Fox2014].
 
-In waterfowl, it is common that adult males display brighter colors than females when  immatures from both sexes display cover-up plumage similar to adult females [@Johnson1999].
-Knowing sex ratios, this discrepancy between adult and immature males might be used to estimate the proportion of immature in a waterfowl population just by counting them from distance [@Smith2001].
-Such approach is promising because it requires only to count the proportion of birds of the two plumage kinds to infer the recruitment rate.
-Unlike the two previous methods, this one is non-invasive and requires a little time investment, which allows to collect long time series and consequently to track recruitment rate variability.
+In waterfowl, it is common that adult males display brighter colours than adult females and immatures of both sexes, which display cover-up plumage similar to adult females [@Johnson1999].
+This discrepancy is valuable because it can be observed and quantified from distance.
+Knowing adult sex ratio, such approach is promising because it requires only to count the proportion of birds of the two plumage kinds to infer the recruitment rate [@Smith2001].
+Unlike the two previous methods, this one is non-invasive and requires an acceptable time investment, which allows to collect long time series and consequently to track recruitment rate variability.
 
@@
 
-From two populations of ruddy duck, an species introduced in Europe in the 40's, this study develops a Bayesian framework to infer recruitment rate from count data.
+From two populations of ruddy duck, an species introduced in Europe in the 40's, this study introduces a simple Bayesian model to infer recruitment rate from count data.
+A large part of this study focuses on assessing the viability of such approach.
 A comparison to the results of the common approach using samples from hunting bags aims at checking if the recruitment rate variability is well tracked and if the hypothesis of the recruitment rate overestimation by sampling methods is consistent with the results.
-Using two other approaches using average population growth rates, one testes the consistency of the order of magnitude of the recruitment rate estimates from the counting method.
+Using two other approaches using average population growth rates, one testes the consistency of the order of magnitude of the recruitment rate estimated from the counting method.
 
@@
 The first reproduction was observed in the 60's.
-The population rapidely grew to reach more than 5000 individuals spread over the entire country.
-A new population set up in France from the 90's, likely because of an arrival of few individuals from the United Kingdom.
-This population set up around a single wintering spot, the Grand Lieu Lake (47.09°N, 1.67°W).
+The population rapidly grew to reach more than 5000 individuals spread over the entire country.
+A new population set up in France from the 90's, likely because of an arrival of a few individuals from the United Kingdom (UK).
+Unlike the UK population, the French one set up around a single wintering spot, the Grand Lieu Lake (47.09°N, 1.67°W).
 Since this species is considered as a threat to the white-headed duck because of the risk of genetic introgression [@MunozFuentes2007], an European plan of eradication has been adopted and control measures were taken in both countries [@GutierrezExposito2020].
 
-Just before the reproduction period, i.e.
-in winter, one can distinguish from shore individuals with male-like plumage from female-like individuals (Figure \@ref(fig:MM1)).
-It is noticeable that the apparent proportion of males in the counts from shore is far below the proportion of males in the samples from the hunting bag (Figure \@ref(fig:MM2)).
+Just before the reproduction period, one can distinguish from shore individuals with male-like plumage from female-like individuals on wintering spots (Figure \@ref(fig:MM1)).
+It is noticeable that the apparent proportion of males in the counts from shore is far below the proportion of males in the samples from hunting bags (Figure \@ref(fig:MM2)).
 
@@
 
-Since sex identification of samples is very accurate **ref required**, the counting method misses the identification of a significant part of males in the population.
-In many waterfowl species, immature individuals of both sexes display cover-up plumage similar to adult female plumage **ref required**.
+Since sex identification of samples is very accurate, such observation demonstrates that the counting method misses the identification of a significant part of males in the population.
+In many waterfowl species, immature individuals of both sexes display cover-up plumage similar to adult female plumage [@Reeber2015].
 Consequently, the counts in winter allows adult males to be distinguished from all other individuals (adult females, immature females, and immature males).
-This property can be valuable to estimate the recruitment with a few assumptions.
-One assumes in this study that the immature individuals have a equally balanced sex ratio [@Bellrose1961; @Blums1996; @Wood2021].
+This property can be valuable to estimate the recruitment with a few conditions.
+The season when adult males can be distinguished from all other individuals is sufficiently close to the breeding period to consider the age structure picture of the population to be representative of the age structure of the breeding population.
+The adult sex ratio is known and stable over the considered time period.
 
-Ruddy duck populations from the UK and from France follow similar demographic histories (Figure \@ref(fig:MM1)).
+The demography of both ruddy duck populations has been tracked thanks to exhaustive counts on the wintering spots during the period between December 1 and January 31.
+Controlled individuals between the counting date and the beginning of the reproduction period (set up to May 1 upon field observations) were subtracted from the winter count to get a proxy of the breeding population size.
+The two populations follow similar demographic histories (Figure \@ref(fig:MM1)).
 They grew freely in both countries until 1999.
@@
 From 2005 in the UK and from 2018 in France, higher control effort led to a rapid depletion in both populations.
+This history underlines that both populations observed a large spectrum of harvest effort, from no pressure to very high pressure.
+For sake of interpretation, the time period 2001-2004 was classified in the "no harvest" category in France because the harvest pressure applied over this period was very limited with no effect on the population growth.
 
@@
 
-Such biological model is thus suited to study the response of the harvest pressure on recruitment rate.
+The counting dataset, which differentiates the male-like from the female-like individuals in winter, covers 7 years and 19 years in UK and France respectively (Figure \@ref(fig:MM3)).
+Combined together, the counting datasets cover different population dynamics corresponding to a large harvest pressure spectrum.
+In France, data before 2004 correspond to a period of population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure (Figure \@ref(fig:MM3)).
+In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period.
+In parallel, the sampling dataset, which corresponds to individuals shot in winter during control operations, covers 9 years only in UK, with 5 years in common with the corresponding counting time series.
 
-## Data time series
+## Recruitment rate inference from count data
 
-The best picture stands just before the breeding period, so generally in winter for waterfowl.
+Adult sex ratio in waterfowl is generally biased towards males (Figure \@ref(fig:MM2), @Wood2021).
+As a consequence, deducing the immature proportion from the adult male proportion is not straightforward because the adult female proportion is not just equal to the adult male proportion.
+Even if the adult sex ratio may vary over a long time range, it is relatively stable over a few years [@Wood2021].
+Sex identification in adult samples from hunting bags was available for both ruddy duck populations.
+However, the small population size in France was limiting to provide useful sex proportion estimates.
+A comparison over months in UK showed no difference in male proportion, so adult samples collected over the whole year were used to estimate the male proportion.
+A comparison of male proportion among years with more 500 samples did not show significant differences among years.
+As a consequence, all adult samples were pooled to estimate the adult male proportion, see Equation \@ref(eq:pm).
 
-The French and the UK wintering populations have been monitored since the presence of the first individuals was validated.
-
-$PR_{i, t}$ estimation by the counting method
-
-## Modelling
+From the cumulated counts of male-like individuals in the wintering population, the proportion of immature individuals can be deduced following Equation \@ref(eq:pic).
+Only years with more than 100 individuals were selected.
+From this proportion and abundance index of breeding population sizes, the recruitment rate is straightforward, see Equation \@ref(eq:r).
+If the absolute value of the breeding population size is accessible, the recruitment can be estimated, see Equation \@ref(eq:R).
 
@@
 |:--|:----|:----------------------------|
-|$SM$|Data|Total number of adult males sampled|
-|$SF$|Data|Total number of adult females sampled|
-|$PM$|Parameter|Proportion of males in the adult part of population at its equilibrium|
-|$S_{i, t}$|Data|Number of individuals sampled in population $i$ in year $t$|
-|$SJ_{i, t}$|Data|Number of juveniles sampled in population $i$ in year $t$|
+|$pm$|Parameter|Proportion of males in the adult part of population at its equilibrium|
+|$SM$|Data|Total number of adult males in samples|
+|$SF$|Data|Total number of adult females in samples|
 |$CM_{i, t}$|Data|Cumulated number of type-male individuals counted in population $i$ in year $t$|
+|$pi_{i, t}$|Parameter|Proportion of immature individuals in population $i$ in year $t$|
 |$C_{i, t}$|Data|Cumulated number of individuals counted in population $i$ in year $t$|
-|$Cmax_{i, t}$|Data|Maximum number of individuals counted in population $i$ in year $t$|
-|$PR_{i, t}$|Parameter|Proportion of recruits in population $i$ in year $t$|
 |$R_{i, t}$|Parameter|Number of recruits in population $i$ in year $t$|
-|$P_{i, t}$|Parameter|Number of recruits in population $i$ in year $t$ per breeder in year $t - 1$|
-|$N0_{j}$|Parameter|Intercept of the regression model, the $j$ index refers to a restricted time period for a single population|
-|$SD_{j}$|Parameter|Standard deviation of the regression model, the $j$ index refers to a restricted time period for a single population|
-|$\lambda_{j}$|Parameter|Population growth rate (in $year^{-1}$), the $j$ index refers to a restricted time period for a single population|
-|$S_{j}$|Parameter|Adult survival rate (in $year^{-1}$), the $j$ index refers to a restricted time period for a single population|
-
-The proportion of males in the adult part of a population at its equilibrium is inferred from the samples.
-
-Indeed, the available French samples were insufficient to get Since the . Age in samples  to 2003 to 2020 for the United Kingdom and
+|$N_{i, t}$|Data|Maximum number of individuals counted in the wintering population $i$ in year $t$ minus removals before reproduction (proxy of the breeding population size)|
+|$r_{i, t}$|Parameter|Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per breeder in year $t - 1$|
 
 \begin{equation}
-PM \sim {\sf Beta}(SM, SF)
-(\#eq:PM)
+pm \sim {\sf Beta}(SM, SF)
+(\#eq:pm)
 \end{equation}
 
-$PR_{i, t}$ estimation by the counting method
+\begin{equation}
+CM_{i, t} \sim {\sf Binom}(pm.(1 - pi_{i, t}), C_{i, t})
+(\#eq:pic)
+\end{equation}
 
 \begin{equation}
-CM_{i, t} \sim {\sf Binom}(PM.(1 - PR_{i, t}), C_{i, t})
-(\#eq:PRC)
+r_{i, t} = \frac{pi_{i, t}.N_{i, t}}{N_{i, t - 1}}
+(\#eq:r)
 \end{equation}
 
-$PR_{i, t}$ estimation by the sampling method
-
 \begin{equation}
-SJ_{i, t} \sim {\sf Binom}(PR_{i, t}, S_{i, t})
-(\#eq:PRS)
-\end{equation}
-
-From $PR_{i, t}$ estimates, $R_{i, t}$ and $P_{i, t}$ are expressed as:
-
-\begin{equation}
-R_{i, t} = PR_{i, t}.Cmax_{i, t}
+R_{i, t} = pi_{i, t}.N_{i, t}
 (\#eq:R)
@@
 
+## Comparison with the sampling method
+
+A direct estimation of the immature proportion can be done using samples from the hunting bag, see Equation \@ref(eq:pis).
+By combining estimates of immature proportion and Equation \@ref(eq:r), annual recruitment rates from samples were estimated to assess their consistency with the counting method.
+One selected only years with more than 100 individuals controlled in winter, which correspond to a 9-year period in UK.
+This sampling time series covers 5 years of the UK counting time series.
+
+|Name|Class|Description|
+|:--|:----|:----------------------------|
+|$SI_{i, t}$|Data|Number of immatures sampled in the wintering population $i$ in year $t$|
+|$S_{i, t}$|Data|Number of individuals sampled in the wintering population $i$ in year $t$|
+
 \begin{equation}
-P_{i, t} = \frac{PR_{i, t}.Cmax_{i, t}}{Cmax_{i, t - 1}}
-(\#eq:P)
+SI_{i, t} \sim {\sf Binom}(pi_{i, t}, S_{i, t})
+(\#eq:pis)
 \end{equation}
 
-The index $j$ refers to a single population $i$ and a restricted time period for $t$.
-The evolution of the population size can be related to the time to infer the population growth rate.
-It is a linear regression on the logarithm scale:
+## Result validation using population growth rates
 
-\begin{equation}
-log(Cmax_{i, t}) \sim {\sf Norm}(N0_{j} + log(\lambda_{j}).t, SD_{j})
-(\#eq:L)
-\end{equation}
-
-\begin{equation}
-S_{j} = \lambda_{j} - \overline{P_{i, t}}
-(\#eq:S)
-\end{equation}
-
-[@RCT2020] R package Nimble [@Valpine2017]
-
-Bayesian framework is efficient to propagate error through the parameters
-
-We generated three chains of length 25000, discarding the first 12500 samples as burn-
-in. Convergence of all chains was assessed using the Gelman and Rubin convergence
-diagnostic (R<1.1, Gelman and Rubin 1992). We fit the models using NIMBLE (de
-Valpine et al. 2017) run from R (R Core Team 2019).
-
-## Comparison with the sampling method
-
-# Result validation using survival analysis
-
-The adult survival is commonly assessed by capture mark recapture and published in the literature.
-Since the survival can also be estimated from both the counting and the sampling methods, the consistency of the two approaches is discussed by comparing the survival estimates with the literature values.
+The simple relationship (a bit more complex if one considers a species with delayed maturity [see @Robertson2008]) between population growth rate, adult survival, and recruitment rate can be used to test the accuracy of the order of magnitude of recruitment rates estimated from the counting and the sampling methods, see Equation \@ref(eq:GR).
+This relationship comes from a simple reasoning for a closed population : the population size in year $t$ is equal to the number of breeders that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
+the recruitment in year $t$.
+The growth rate of a population is thus the sum of the adult survival rate and the recruitment rate [@Flint2015].
 
@@
 |$R_{t}$|Number of recruits in year $t$|
-|$S_{t}$|Adult survival rate, i.e. proportion of breeder in year $t-1$ still alive in year $t$|
-|$P_{t}$|Number of recruits in year $t$ produced per breeder in year $t-1$|
-|$\lambda_{t}$|Growth rate of the population between the reproduction period of year $t-1$ and year $t$|
+|$s_{t}$|Adult survival rate, i.e. proportion of breeders in year $t-1$ still alive in year $t$|
+|$r_{t}$|Recruitment rate, i.e. number of recruits in year $t$ produced per breeder in year $t-1$|
+|$\lambda_{t}$|Growth rate of the population between the reproduction periods of year $t-1$ and year $t$|
 
@@
 N_{t} & = N_{t-1} - D_{t-1} + R_{t} \notag \\
-N_{t} & = N_{t-1} - (1 - S_{t}).N_{t-1} + P_{t}.N_{t-1} \notag \\
-N_{t} & = S_{t}.N_{t-1} + P_{t}.N_{t-1} \notag \\
-\frac{N_{t}}{N_{t-1}} & = S_{t} + P_{t} \notag \\
-\lambda_{t} & = S_{t} + P_{t} 
+N_{t} & = N_{t-1} - (1 - S_{t}).N_{t-1} + r_{t}.N_{t-1} \notag \\
+N_{t} & = S_{t}.N_{t-1} + r_{t}.N_{t-1} \notag \\
+\frac{N_{t}}{N_{t-1}} & = s_{t} + r_{t} \notag \\
+\lambda_{t} & = s_{t} + r_{t} 
 (\#eq:GR)
@@
 
-The population size in year $t$ is equal to the number of breeders that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
-the recruitment in year $t$.
-The growth rate of a population is thus the sum of the recruitment rate and the adult survival rate.
+Population growth rate can be assessed from the evolution of the breeding population size.
+Since there is observation error in population size estimates and process variability in population growth, only average population growth rates are estimated in this study following Equation \@ref(eq:L), which is a linear regression on the logarithm scale.
 
-The proportion of males in the adult part of a population at its equilibrium is inferred from the samples.
+|Name|Class|Description|
+|:--|:----|:----------------------------|
+|$\lambda_{i, J}$|Parameter|Average population growth rate over a restricted time interval $J$ for a population $i$ (in $year^{-1}$)|
+|$N0_{i, J}$|Parameter|Intercept of the regression model|
+|$t \in J$|Index|Year index within the time interval $J$| 
+|$\sigma_{i, J}$|Parameter|Standard deviation of the regression model|
 
-Indeed, the available French samples were insufficient to get Since the . Age in samples  to 2003 to 2020 for the United Kingdom and
+\begin{equation}
+log(N_{i, t}) \sim {\sf Norm}(N0_{i, J} + log(\lambda_{i, J}).t, \sigma_{i, J})
+(\#eq:L)
+\end{equation}
 
-develop the counting method :
+### Analysis of the adult survival
 
-2 ways to check the method :
-
-sampling method?
-Ok but few data as poorly harvested
-
-realistic survival? 
-if > 1, estimated recruitment rate is too low to support the maximum growth rate
-in the range of similar species in litterature -> satisfying
-
-blabla
+The adult survival is commonly assessed by capture mark recapture [@Lebreton2001].
+Combined to average population growth rate, adult survival rate allows average recruitment rate to be estimated, and then compared with the output of the counting method.
+Unfortunately, no banding campaign was performed on the two ruddy duck populations.
+However, the relationship \@ref(eq:GR) could still produce average adult survival estimates that can be discussed.
+To do so, average population growth rate were estimated over time intervals overlapping the counting time series and the main corresponding harvest category, see Figure \@ref(fig:MM4).
 
@@
 
-## Result validation using maximum recruitment rate proxy
+Inconsistency on recruitment rate output, which is defined on $[0;\infty]$, would either lead to average survival output $<0$, which points out an overestimation of the recruitment rate, or $>1$, which reflects that the recruitment rate cannot support the observed population growth and is thus underestimated.
+In between the two limit values, one still discuss the average survival outputs by comparison with survival values published in the literature.
+
+### Comparison to a proxy of maximum recruitment rate
+
+When population growth is not limited by a harvest pressure and/or by density-dependent effects, both recruitment rate and adult survival are close to their maximum potential.
+Given a conservative assumption on the maximum adult survival of ruddy ducks and maximum population growth rate estimated during no harvest time period, one can provide a proxy of the maximum recruitment rate.
+If the recruitment rate outputs of the model are significantly superior to this proxy, it means that the method overestimates the recruitment rate, which means that some adult male individuals are not detected during count surveys.
+To do so, one uses a conservative value of adult survival of 1.
+After reaching the threshold of 1000 individuals, the UK population growth observed a strong inflexion whereas no harvest pressure was applied.
+This observation led to split the analysis into two time series, see Figure \@ref(fig:MM5).
 
@@
 
+For all the sub-models in Section \@ref(MM), the Bayesian framework was used for its efficiency and simplicity to propagate error through the parameters.
+One generated three chains of length 500000, with a thinning of 10 to avoid autocorrelation in the samples, and we discarded the first 2000 samples as burn-in.
+Chain convergence was assessed using the Gelman and Rubin convergence diagnostic (R<1.1, @Gelman1992).
+We fit the models using NIMBLE [@Valpine2017] run from R [@RCT2020].
+Data and code are available here: https://github.com/adri-tab/Ruddy_duck_recruitment_rate.
+
 # Results
 
-## recruitment rate variability
+## Recruitment rate variability
+
+The adult male proportion that is necessary to infer on the recruitment rate from count data is estimated at 60% ($0.60 +/- 0.01$).
+
+The counting method provides estimations of recruitment rate with good varying from 0.05estimate2 choses: 
+Counting method observed huge variability .05 to .8 recruits per breeder (16 factor !) but still 
+comprised between 0 and 1 like adult survival by definition, meaning similar weight to build population growth.
+Far higher in France than in UK, to be related to the growth rates differences (-44% vs -3%)
+
+Higher on average in the early years in France when (almost) no harvest pressure (1999- 2004) but large interannual fluctuations without clear temporal trends,but small temporal correlation, especially in UK.
+
+quite stable when harvest rate is limited (France 2000), but is lower and noisier when harvest rate is huge and the population depleted.
+
+No correlation between UK and France (only X years), meaning no large scale factor drive the fluctuations.
+
+in UK only, recruitment rate estimated from sample data consistently higher than count data as expected.
+Correlation between the two methods very high.
+Sampling method overestimate the recruitment by 44% compared to the count method.
 
@@
 
-quite stable when harvest rate is limited (France 2000), but is lower and noisy when harvest rate is huge and the population depleted.
+## Consistency of the adult survival
 
-## Result validation using survival analysis
+Depletion is clear on the counting time series intervals (- 44% on average per year) compared to the not significant decrease in France over the time interval with harvest (-3% per year on average).
 
@@
 
+As described before, recruitement rate over the studied time period is far lower for the UK than France (2.5 factor). 
+Underlined adult survival are comprised between 0 and 1, whatever the method. So we cannot conclude if one can be rejected.
+
 ```{r R3, fig.cap = "(ref:R3cap)"}
@@
 
-## Result validation using maximum recruitment rate proxy
+whereas adult survival more consistent between the tow (.5 factor), meaning there was more variability in recruitment rate than in survival rate.
+growth rate difference is thus more explained by recruitment rate fluctuation than adult survival fluctuation, consistent with the demographic buffering, or canalization hypothesis [@Gaillard2003; @Lenzi2021]
+
+survival values compared to litterature? s
+
+## Comparison to a proxy of maximum recruitment rate
+
+Maximum growth rate are similar for both populations (+45%), which supports the consistency of the species.
+After a certain threshold, the growth rate falls to 6% per year even though no harvest pressure was applied on it, maybe density dependent effects.
 
@@
 
+Upon the conservative assumption of adult survival of 1, maximum recruitment rate reach .45 recruits per breeder.
+For both populations and methods, values remains below this maximum recruitment rate.
+This demonstrates that methods provide no inconsistent recruitment rate values, which is quite satisfying and cannot reject the methods.
+
 ```{r R5, fig.cap = "(ref:R5cap)"}
@@
 
+If counting method acceptable, moderate harvest pressure does not reduce the recruitment rate as expected, maybe due to density dependend mortality for juvenile but it affects it if strong harvest rate.
+That is very interesting bro but warning,  these interpretation holds on weak estimations and adult survival equal to 1!!!
+
 # Discussion
+
+adult sex ratio in line with other duck species [@Wood2021]
+
+This time lag might be not a issue if the survival of immature is close to the adult survival between the two periods. winter count and breeding season
 
@@
 
+No correlation between the two populations on recruitment rate, so maybe no large scale drivers of recruitment rate variability.
+
 age proportion -> [0, 1]
-recruitment rate [@, + inf] donc plus sensitive
+recruitment rate [0, +inf] donc plus sensitive
 
@@
 
-(ref:MM3cap) Evolution of the size of the two Ruddy duck populations in light of the data time series;
+(ref:MM3cap) Evolution of the size of the two ruddy duck populations in light of the harvest pressure and the data time series
 
-(ref:MM4cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods with data time series, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:MM4cap) Evolution of the size of the two ruddy duck populations in light of the harvest pressure; on time periods with data time series, population growth rates can be estimated (slopes in red for the UK and in blue for France)
 
-(ref:MM5cap) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:MM5cap) Evolution of the size of the two ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)
 
@@
 
-(ref:S1cap) (A1) Estimates of the number of recruits, i.e.
-the young adults reproducing for their first time; (B1) estimates of the proportion of recruits in the population at the reproduction period; (A2) & (B2) comparison of the values between the two estimation methods; bars define the 95% confidence intervals
+(ref:S1cap) (A1) Estimates of the oroportion of the immature individuals in the wintering population; (B1) estimates of the number of recruits, i.e.
+the young adults reproducing for their first time; (A2) & (B2) comparison of the values between the two estimation methods; bars define the 95% confidence intervals

### 2021-11-24 Code/Model_recruitment_rate.R

@@ -160,7 +160,7 @@ frag %>%
   scale_x_date_own(5e-2) +
   guides(size = guide_legend(title = "Samples")) +
   labs(y = "Male proportion adult")
-
+  
 # juvenile
 frag %>% 
   filter(age == 'no_ad', sex != "ind") %>% 
@@ -286,7 +286,7 @@ count_sex_app %>%
 
 # Lambda formatting ------------------------------------------------------------
 
-# estimation of lambda_max to test productivity reliability
+# estimation of lambda_max to test recruitment rate reliability
 # when ruddy ducks start to be permanently seen and growing?
 counts_1 %>% 
   ggplot(aes(x = year, y = n_pop, color = pop)) + 
@@ -436,7 +436,7 @@ JuvCode <- nimbleCode(
       
       p_juv[j] <- (1 - p_mal_ad[j] / p_mal) # * 1.741913
       
-      # recruitment and productivity estimation
+      # recruitment and recruitment rate estimation
       recruits[j] <- p_juv[j] * cnt_size_pop[j]
       
       productivity[j] <- recruits[j] / cnt_size_breeding_pop[j]
@@ -778,7 +778,7 @@ JuvOut4 %>%
   facet_wrap(~ pop, nrow = 1, scales = "free_x") +
   geom_point() +
   geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5, size = 1.2) +
-  geom_text(aes(x = method, y = .365, label = txt), color = "#de2d26", angle = 90) +
+  geom_text(aes(x = method, y = .365, label = txt), color = c_met[2], angle = 90) +
   scale_y_continuous(
     limits = c(0, .7),
     breaks = 0:7 / 10) +
@@ -797,7 +797,7 @@ JuvOut4 %>%
   facet_wrap(~ pop, nrow = 1, scales = "free_x") +
   geom_point() +
   geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5, size = 1.2) +
-  geom_text(aes(x = method, y = .5, label = txt), color = "#de2d26", angle = 90) +
+  geom_text(aes(x = method, y = .5, label = txt), color = c_met[2], angle = 90) +
   scale_y_continuous(
     labels = scales::percent_format(accuracy = 1L),
     limits = c(-.04, 1),
@@ -822,7 +822,7 @@ JuvOut4 %>%
   facet_wrap(~ pop, nrow = 1, scales = "free_x") +
   geom_point() +
[…]

### 2021-11-26 Code/Draft_ruddy_duck_recruitment_rate.Rmd
Update

@@
 One defines the recruitment rate of a breeding population as the average number of recruits produced per each breeder [@Flint2015].
-This parameter might sometimes be called "annual fertility" [@Koons2014], "reproductive rate" [@Cooch2014], "reproductive success" [@Etterson2011], or "productivity" [@Johnson1987].
+This parameter might sometimes be called "annual fertility" [@Koons2014], "fecundity" [@Arnold2018], "reproductive rate" [@Cooch2014], "reproductive success" [@Etterson2011], or "productivity" [@Johnson1987; @Hagen2008].
 If one considers a species reaching its sexual maturity as early as its first year, all individuals are mature at the reproduction season.
@@
 The fecundity can be defined by the product of the nesting success, the average clutch size, and the hatching success.
-The juvenile survival can be defined as the product of the post-hatching survival and the post-fledging survival.
+The juvenile survival can be defined as the product of the post-hatching survival and the post-fledging survival, both critical life periods [@Hill2008; @Blums2004].
 The recruitment rate as defined in the present study is thus a parameter integrating multiple life steps from the egg to the young adult attempting to breed for its first time [@Koons2017].
@@
 These decision-support tools are then key inputs to status on the best management rules to implement.
-The growth rate of a population depends on the adult survival and the recruitment rate, see Equation \@ref(eq:GR) in Section \@ref(MM).
+The growth rate of a population depends on the adult survival and the recruitment rate, see Equation \@ref(eq:GR) in Section \@ref(PGR).
 The adult survival is generally estimated by mark recapture/recovery methods [@Lebreton2001].
@@
 To do so, it is necessary to monitor nests in habitats representative of the managed population [@Blums2004].
-The survival of a chick to its recruitment can be done by capture mark recapture method quite tricky to set up [@Blums2004; Schmidt2008].
+The survival of a chick to its recruitment can be done by capture mark recapture method quite tricky to set up [@Blums2004; @Schmidt2008, @Arnold2018].
 Even if such approach provides valuable information on the dynamics of the recruitment and can highlight the bottlenecks, the fieldwork required to estimate both components is too time-consuming to generalize such approach for waterfowl population management on a long-term horizon.
@@
 The hunting bag analysis might then lead to overestimate the recruitment rate and consequently the maximum sustainable harvest rate, with adverse consequences on population conservation [@Fox2014].
+
+Alternative method: trapping methods but still biased @Arnold2018
+Finally, because juveniles are more naïve, they may be more vulnerable to capture by
+standard trapping methods (Rguibi- Idrissi et al., 2003), even if loca-
+tions and timing were otherwise unbiased.
 
@@
 This property can be valuable to estimate the recruitment with a few conditions.
-The season when adult males can be distinguished from all other individuals is sufficiently close to the breeding period to consider the age structure picture of the population to be representative of the age structure of the breeding population.
+The season when adult males can be distinguished from all other individuals is sufficiently close to the breeding period to consider the age structure picture of the population to be representative of the age structure of the breeding population^[immature survival equal to adult survival [@Arnold2018]].
 The adult sex ratio is known and stable over the considered time period.
@@
 
-## Result validation using population growth rates
+## Result validation using population growth rates {#PGR}
 
@@
 If the recruitment rate outputs of the model are significantly superior to this proxy, it means that the method overestimates the recruitment rate, which means that some adult male individuals are not detected during count surveys.
-To do so, one uses a conservative value of adult survival of 1.
+To do so, one uses a very conservative value of adult survival of 0.85 [highest survival rates observed in litterature for species of similar weight, @Nichols1997; @Krementz1997; @Buxton2004].
 After reaching the threshold of 1000 individuals, the UK population growth observed a strong inflexion whereas no harvest pressure was applied.
@@
 Data and code are available here: https://github.com/adri-tab/Ruddy_duck_recruitment_rate.
+The values *X [min; max]* reported in results are means and the associated 95% confidence interval.
 
@@
 
-The adult male proportion that is necessary to infer on the recruitment rate from count data is estimated at 60% ($0.60 +/- 0.01$).
+The adult male proportion, which is necessary to infer on the recruitment rate from count data, is estimated at 0.60 [0.59; 0.61].
+The counting method provides estimations of the recruitment rate, Figure \@ref(fig:R1)A.
+In UK, values vary from 0.06 [0.05; 0.07] to 0.27 [0.20; 0.34] recruits per breeder, and in France, 0.08 [0.04; 0.12] to 0.76 [0.64; 0.88] recruits per breeder.
+There are significant year-to-year variations with a maximum amplitude of 0.70 recruits per breeder (12-time variability).
+Recruitment rates are estimated in both populations for 5 years, and no correlation is observed between the two populations.
 
-The counting method provides estimations of recruitment rate with good varying from 0.05estimate2 choses: 
-Counting method observed huge variability .05 to .8 recruits per breeder (16 factor !) but still 
+Recruitment rates estimated from UK sample data are consistently higher than estimates from the corresponding count data Figure \@ref(fig:R1)A.
+The correlation between the two methods is strong and points out that the sampling method overestimates the recruitment rate by 44% compared to the counting method, Figure \@ref(fig:R1)B.
+
+```{r R1, fig.cap = "(ref:R1cap)"}
+knitr::include_graphics("../Output/plot_5.png")
+```
+
+## Consistency of the adult survival
+
+Average population growth rates were estimated over the period of available count data with a consistent harvest pressure.
+Average population growth rates ($\lambda$) are estimated to 0.56 [0.48; 0.65] and 0.97 [0.94; 1.00] for the UK and France respectively, which corresponds to an average annual growth proportion ($\lambda - 1$) of -0.44 [-0.52; -0.35] and -0.03 [-0.06; 0.00] in the UK and France respectively, Figure \@ref(fig:R2).
+
 comprised between 0 and 1 like adult survival by definition, meaning similar weight to build population growth.
@@
 quite stable when harvest rate is limited (France 2000), but is lower and noisier when harvest rate is huge and the population depleted.
-
-No correlation between UK and France (only X years), meaning no large scale factor drive the fluctuations.
-
-in UK only, recruitment rate estimated from sample data consistently higher than count data as expected.
-Correlation between the two methods very high.
-Sampling method overestimate the recruitment by 44% compared to the count method.
-
-```{r R1, fig.cap = "(ref:R1cap)"}
-knitr::include_graphics("../Output/plot_5.png")
-```
-
-## Consistency of the adult survival
 
@@
 
-Upon the conservative assumption of adult survival of 1, maximum recruitment rate reach .45 recruits per breeder.
+Upon the conservative assumption of adult survival of .85, maximum recruitment rate reach .45 recruits per breeder.
 For both populations and methods, values remains below this maximum recruitment rate.
@@
 # Discussion
+
+@Arnold2018 underlines the lack of data on recruitment rate for modelers (in discussion) and that new methods are required to track this variable that often drive the population growth rate
+
+Recruitment rate drives lambda? see @Arnold2018
+
+ (only X years), meaning no large scale factor drive the fluctuations of productivity.
 
@@
 
-(ref:S1cap) (A1) Estimates of the oroportion of the immature individuals in the wintering population; (B1) estimates of the number of recruits, i.e.
+(ref:S1cap) (A1) Estimates of the proportion of the immature individuals in the wintering population; (B1) estimates of the number of recruits, i.e.
 the young adults reproducing for their first time; (A2) & (B2) comparison of the values between the two estimation methods; bars define the 95% confidence intervals

### 2021-11-26 Code/Model_recruitment_rate.R

@@ -132,7 +132,7 @@ read_rds("./Data/Ruddy_duck_data.rds") %>% pluck(2) %>%
   mutate(across(pop, as_factor)) %>% arrange(pop) -> counts
 
 
-# male proportion -------------------------------------------------------------------
+# Male proportion -------------------------------------------------------------------
 
 # adult
 frag %>% 
@@ -440,6 +440,9 @@ JuvCode <- nimbleCode(
       recruits[j] <- p_juv[j] * cnt_size_pop[j]
       
       productivity[j] <- recruits[j] / cnt_size_breeding_pop[j]
+      
+      survival[j] <- cnt_size_pop[j] / cnt_size_breeding_pop[j] - productivity[j]
+      
     }
     
     # parameter estimation from hunting samples
@@ -457,6 +460,9 @@ JuvCode <- nimbleCode(
       recruits[C + k] <- p_juv[C + k] * spl_size_pop[k]
       
       productivity[C + k] <- recruits[C + k] / spl_size_breeding_pop[k]
+      
+      survival[C + k] <- spl_size_pop[k] / spl_size_breeding_pop[k] - productivity[C + k]
+      
     }
     
     # r estimation frow counts
@@ -482,16 +488,6 @@ JuvCode <- nimbleCode(
                                 sd = N_sd[R_id[t]])
     }
     
-    # survival lambda max growth
-    for (q in 1:W) {
-      survival[q] <- lambda[W_id[q]] - productivity[q]
-    }
-    
-    # survival true lambda 
-    for (p in 1:Z) {
-      survival_2[p] <- lambda[Z_id[p]] - productivity[p]
-    }
-    
   })
 
 # set seed
@@ -509,16 +505,7 @@ JuvConst <- list(C = dis(obs_tot, pop, id, TRUE),
                  R_id = lam(sub_pop, id),
                  R_id_max = lam(sub_pop, id) %>% max(),
                  TS_id = lam(sub_ts, id),
-                 TS_id_max = lam(sub_ts, id) %>% max(),
-                 W = dis(obs_tot, pop, id, TRUE) + dis(spl_tot, pop, id, TRUE),
-                 W_id = c(dis(obs_tot, pop, id) %>% if_else(. == 2, 3, .), 
-                          dis(spl_tot, pop, id)) + 2,
-                 Z = lambda_ds %>% 
[…]

### 2021-12-03 Code/Draft_ruddy_duck_recruitment_rate.Rmd
annual survival rate

@@
 The recruitment rate, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
-The proportion of young adult in a waterfowl population, which is a main step in estimating recruitment rate, is often estimated from game-hunting samples.
+The proportion of young adult in a waterfowl population, which is a main step in estimating the recruitment rate, is often estimated from game-hunting samples.
 However, this proportion is not accessible in years without harvest, or with a low harvest rate.
@@
 Such information can be used to infer the proportion of immatures, and consequently the recruitment rate.
-In using two populations of ruddy duck, *Oxyura oxyura*, this study introduces a Bayesian method to estimate the recruitment rate from count data.
-The results are compared to recruitment rate estimates from samples and their consistency to support the observed population growth rates is tested.
+In using two populations of ruddy duck, *Oxyura oxyura*, this study introduces a model to estimate the recruitment rate from count data.
+To test the accuracy of the method, the results are compared to recruitment rate estimates from samples, and their consistency to support the observed population growth rates is also tested.
 The results suggest that the counting method is a versatile tool to estimate the recruitment rate if the species is monitored during the appropriate time window.
-The results argue for considering a two-category count protocol in winter surveys of waterfowl species when adult males can be identified.
+Authors argue for considering a two-category count protocol in winter surveys of waterfowl species for which adult males can be differentiated from other individuals.
 \\par \\textbf{Keywords:} fecundity - productivity - reproductive success - age ratio - juvenile - duck"
@@
 A new population set up in France from the 90's, likely because of an arrival of a few individuals from the United Kingdom (UK).
-Unlike the UK population, the French one set up around a single wintering spot, the Grand Lieu Lake (47.09°N, 1.67°W).
-Since this species is considered as a threat to the white-headed duck because of the risk of genetic introgression [@MunozFuentes2007], an European plan of eradication has been adopted and control measures were taken in both countries [@GutierrezExposito2020].
+Unlike the UK population, the French one set up around a single wintering spot, the Grand Lieu Lake (47.09°N, 1.67°W), which falicitates the population monitoring.
+Since this species is considered as a threat to the white-headed duck because of the risk of a genetic introgression [@MunozFuentes2007], an European plan of eradication has been adopted and control measures were taken in both countries [@GutierrezExposito2020].
 
-Just before the reproduction period, one can distinguish from shore individuals with male-like plumage from female-like individuals on wintering spots (Figure \@ref(fig:MM1)).
-It is noticeable that the apparent proportion of males in the counts from shore is far below the proportion of males in the samples from hunting bags (Figure \@ref(fig:MM2)).
+Just before the reproduction period, one can distinguish from shore individuals with male-like plumage from female-like individuals on wintering spots (Figure \@ref(fig:flock)).
+It is noticeable that the apparent proportion of males in the counts from shore is far below the proportion of males in the samples from hunting bags (Figure \@ref(fig:pmal)).
 
-```{r MM1, fig.cap = "(ref:MM1cap)"}
+```{r flock, fig.cap = "(ref:flock)"}
 knitr::include_graphics("../Output/flock.png")
@@
 
-```{r MM2, fig.cap = "(ref:MM2cap)"}
-knitr::include_graphics("../Output/plot_1.png")
+```{r pmal, fig.cap = "(ref:pmal)"}
+knitr::include_graphics("../Output/plot_1_saved.png")
 ```
@@
 Since sex identification of samples is very accurate, such observation demonstrates that the counting method misses the identification of a significant part of males in the population.
-In many waterfowl species, immature individuals of both sexes display cover-up plumage similar to adult female plumage [@Reeber2015].
+In many waterfowl species, immature individuals of both sexes display cover-up plumage similar to adult females [@Reeber2015].
 Consequently, the counts in winter allows adult males to be distinguished from all other individuals (adult females, immature females, and immature males).
-This property can be valuable to estimate the recruitment with a few conditions.
-The season when adult males can be distinguished from all other individuals is sufficiently close to the breeding period to consider the age structure picture of the population to be representative of the age structure of the breeding population^[immature survival equal to adult survival [@Arnold2018]].
+This feature can be valuable to estimate the recruitment with two conditions.
+The season when adult males can be distinguished from all other individuals is sufficiently close to the breeding period to consider the age structure picture of the population to be representative of the age structure of the breeding population.
 The adult sex ratio is known and stable over the considered time period.
@@
 Controlled individuals between the counting date and the beginning of the reproduction period (set up to May 1 upon field observations) were subtracted from the winter count to get a proxy of the breeding population size.
-The two populations follow similar demographic histories (Figure \@ref(fig:MM1)).
+Time series show that the two populations follow similar demographic histories (Figure \@ref(fig:count)).
 They grew freely in both countries until 1999.
-Limited control measures have been then applied from 1999 to 2005 in the UK and from 1999 to 2018 in France.
+Control measures have been then applied from 1999 to 2005 in the UK and from 1999 to 2018 in France.
 In both countries, these management measures led to stop the expansion of the ruddy ducks.
@@
 This history underlines that both populations observed a large spectrum of harvest effort, from no pressure to very high pressure.
-For sake of interpretation, the time period 2001-2004 was classified in the "no harvest" category in France because the harvest pressure applied over this period was very limited with no effect on the population growth.
+For sake of interpretation, the time period 2001-2004 was classified in the *no harvest* category in France because the harvest pressure applied over this period was very limited with no effect on the population growth.
 
-```{r MM3, fig.cap = "(ref:MM3cap)"}
+```{r count, fig.cap = "(ref:count)"}
 knitr::include_graphics("../Output/plot_2.png")
@@
 
-The counting dataset, which differentiates the male-like from the female-like individuals in winter, covers 7 years and 19 years in UK and France respectively (Figure \@ref(fig:MM3)).
-Combined together, the counting datasets cover different population dynamics corresponding to a large harvest pressure spectrum.
-In France, data before 2004 correspond to a period of population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure (Figure \@ref(fig:MM3)).
+The counting dataset, which differentiates the male-like from the female-like individuals in winter, covers 7 years and 19 years in UK and France respectively (Figure \@ref(fig:count)).
+Combined together, the counting datasets cover different population dynamics corresponding to a gradient of harvest pressure.
+In France, data before 2004 correspond to a period of population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure (Figure \@ref(fig:count)).
 In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period.
@@
 
-Adult sex ratio in waterfowl is generally biased towards males (Figure \@ref(fig:MM2), @Wood2021).
+Adult sex ratio in waterfowl is generally biased towards males (Figure \@ref(fig:pmal), @Wood2021).
 As a consequence, deducing the immature proportion from the adult male proportion is not straightforward because the adult female proportion is not just equal to the adult male proportion.
@@
 A comparison over months in UK showed no difference in male proportion, so adult samples collected over the whole year were used to estimate the male proportion.
-A comparison of male proportion among years with more 500 samples did not show significant differences among years.
+A comparison of male proportion among years with more 500 samples did not show significant interannual differences.
 As a consequence, all adult samples were pooled to estimate the adult male proportion, see Equation \@ref(eq:pm).
 
-From the cumulated counts of male-like individuals in the wintering population, the proportion of immature individuals can be deduced following Equation \@ref(eq:pic).
+Since the additive property of the binomiale distribution, the proportion of immature individuals can be deduced from the cumulated counts of male-like individuals in the wintering population, see Equation \@ref(eq:pic).
 Only years with more than 100 individuals were selected.
@@
 |$R_{i, t}$|Parameter|Number of recruits in population $i$ in year $t$|
-|$N_{i, t}$|Data|Maximum number of individuals counted in the wintering population $i$ in year $t$ minus removals before reproduction (proxy of the breeding population size)|
+|$N_{i, t}$|Data|Proxy of the size of the breeding population $i$ in year $t$ (maximum number of individuals counted in the wintering population minus removals before reproduction)|
 |$r_{i, t}$|Parameter|Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per breeder in year $t - 1$|
@@
 The adult survival is commonly assessed by capture mark recapture [@Lebreton2001].
-Combined to average population growth rate, adult survival rate allows average recruitment rate to be estimated, and then compared with the output of the counting method.
+Combined to population growth rate, adult survival rate allows recruitment rate to be estimated, and then compared with the output of the counting method.
 Unfortunately, no banding campaign was performed on the two ruddy duck populations.
-However, the relationship \@ref(eq:GR) could still produce average adult survival estimates that can be discussed.
-To do so, average population growth rate were estimated over time intervals overlapping the counting time series and the main corresponding harvest category, see Figure \@ref(fig:MM4).
+However, the relationship \@ref(eq:GR) could still produce adult survival estimates that can be discussed.
+To do so, population growth rate overlapping the counting time series were estimated.
 
-```{r MM4, fig.cap = "(ref:MM4cap)"}
-knitr::include_graphics("../Output/plot_3.png")
-```
-
-Inconsistency on recruitment rate output, which is defined on $[0;\infty]$, would either lead to average survival output $<0$, which points out an overestimation of the recruitment rate, or $>1$, which reflects that the recruitment rate cannot support the observed population growth and is thus underestimated.
-In between the two limit values, one still discuss the average survival outputs by comparison with survival values published in the literature.
+Inconsistency on recruitment rate output, which is defined on $[0;\infty]$, would either lead to survival output $<0$, which points out an overestimation of the recruitment rate, or $>1$, which reflects that the recruitment rate cannot support the observed population growth and is thus underestimated.
+In between the two limit values, one still discuss the survival outputs by comparison with survival values published in the literature.
 
@@
 If the recruitment rate outputs of the model are significantly superior to this proxy, it means that the method overestimates the recruitment rate, which means that some adult male individuals are not detected during count surveys.
-To do so, one uses a very conservative value of adult survival of 0.85 [highest survival rates observed in litterature for species of similar weight, @Nichols1997; @Krementz1997; @Buxton2004].
+To do so, one uses a very conservative prior of adult survival by setting a uniform distribution on $[0.7;1]$ [the lower limit corresponds to the highest range of survival rates observed in litterature for species of similar weight, @Nichols1997; @Krementz1997; @Buxton2004].
 After reaching the threshold of 1000 individuals, the UK population growth observed a strong inflexion whereas no harvest pressure was applied.
-This observation led to split the analysis into two time series, see Figure \@ref(fig:MM5).
+This observation led to split the analysis into two time series, see Figure \@ref(fig:grts).
 
-```{r MM5, fig.cap = "(ref:MM5cap)"}
-knitr::include_graphics("../Output/plot_4.png")
+```{r grts, fig.cap = "(ref:grts)"}
+knitr::include_graphics("../Output/plot_3.png")
 ```
@@
 The adult male proportion, which is necessary to infer on the recruitment rate from count data, is estimated at 0.60 [0.59; 0.61].
-The counting method provides estimations of the recruitment rate, Figure \@ref(fig:R1)A.
+The counting method provides estimations of the recruitment rate, Figure \@ref(fig:rrts)A.
 In UK, values vary from 0.06 [0.05; 0.07] to 0.27 [0.20; 0.34] recruits per breeder, and in France, 0.08 [0.04; 0.12] to 0.76 [0.64; 0.88] recruits per breeder.
@@
 
-Recruitment rates estimated from UK sample data are consistently higher than estimates from the corresponding count data Figure \@ref(fig:R1)A.
-The correlation between the two methods is strong and points out that the sampling method overestimates the recruitment rate by 44% compared to the counting method, Figure \@ref(fig:R1)B.
+Recruitment rates estimated from UK sample data are consistently higher than estimates from the corresponding count data Figure \@ref(fig:rrts)A.
+The correlation between the two methods is strong and points out that the sampling method overestimates the recruitment rate by 44% compared to the counting method, Figure \@ref(fig:rrts)B.
 
-```{r R1, fig.cap = "(ref:R1cap)"}
-knitr::include_graphics("../Output/plot_5.png")
+```{r rrts, fig.cap = "(ref:rrts)"}
+knitr::include_graphics("../Output/plot_4.png")
 ```
@@
 Average population growth rates were estimated over the period of available count data with a consistent harvest pressure.
-Average population growth rates ($\lambda$) are estimated to 0.56 [0.48; 0.65] and 0.97 [0.94; 1.00] for the UK and France respectively, which corresponds to an average annual growth proportion ($\lambda - 1$) of -0.44 [-0.52; -0.35] and -0.03 [-0.06; 0.00] in the UK and France respectively, Figure \@ref(fig:R2).
+Average population growth rates ($\lambda$) are estimated to 0.56 [0.48; 0.65] and 0.97 [0.94; 1.00] for the UK and France respectively, which corresponds to an average annual growth proportion ($\lambda - 1$) of -0.44 [-0.52; -0.35] and -0.03 [-0.06; 0.00] in the UK and France respectively
 
@@
 
-```{r R2, fig.cap = "(ref:R2cap)"}
-knitr::include_graphics("../Output/plot_6.png")
+```{r srts, fig.cap = "(ref:srts)"}
+knitr::include_graphics("../Output/plot_5.png")
 ```
@@
 Underlined adult survival are comprised between 0 and 1, whatever the method. So we cannot conclude if one can be rejected.
-
-```{r R3, fig.cap = "(ref:R3cap)"}
-knitr::include_graphics("../Output/plot_7.png")
-```
 
@@
 
-```{r R4, fig.cap = "(ref:R4cap)"}
-knitr::include_graphics("../Output/plot_8.png")
+```{r mgr, fig.cap = "(ref:mgr)"}
+knitr::include_graphics("../Output/plot_6.png")
 ```
@@
 
-```{r R5, fig.cap = "(ref:R5cap)"}
-knitr::include_graphics("../Output/plot_9.png")
+```{r grcomp, fig.cap = "(ref:grcomp)"}
+knitr::include_graphics("../Output/plot_7.png")
 ```
@@
 
-```{r S1, fig.cap = "(ref:S1cap)"}
-knitr::include_graphics("../Output/plot_10.png")
+```{r sup, fig.cap = "(ref:sup)"}
+knitr::include_graphics("../Output/plot_8.png")
 ```
 
-(ref:MM1cap) Typical observation of a ruddy duck flock in winter: 4 male-like individuals with a white cheek and a black cap, 10 female-like individuals with a whitish cheek and a dark stripe across it, 3 unidentified individuals © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
+(ref:flock) Typical observation of a ruddy duck flock in winter: 4 male-like individuals with a white cheek and a black cap, 10 female-like individuals with a whitish cheek and a dark stripe across it, 3 unidentified individuals © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
 
-(ref:MM2cap) Apparent male proportion in winter counts vs male proportion in samples; for the first category, one point corresponds to the proportion of male-like plumage in a population counted in winter; for the second category, the male proportion is estimated from all ruddy ducks sampled over one year
+(ref:pmal) Apparent male proportion in winter counts vs male proportion in samples; for the first category, one point corresponds to the proportion of male-like plumage in a population counted in winter; for the second category, the male proportion is estimated from all ruddy ducks sampled over one year
 
-(ref:MM3cap) Evolution of the size of the two ruddy duck populations in light of the harvest pressure and the data time series
+(ref:count) Evolution of the size of the two ruddy duck populations in light of the harvest pressure and the data time series
 
-(ref:MM4cap) Evolution of the size of the two ruddy duck populations in light of the harvest pressure; on time periods with data time series, population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:grts) Evolution of the size of the two ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)
 
-(ref:MM5cap) Evolution of the size of the two ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)
-
-(ref:R1cap) (A) Estimates of recruitment rate, i.e.
+(ref:rrts) (A) Estimates of recruitment rate, i.e.
 average number of recruits produced per breeder; (B) Comparison of the recruitment rate values between the two estimation methods; bars define the 95% confidence intervals
 
-(ref:R2cap) (A) Estimates of the population growth rate on time intervals with data availability.
-This aims at producing survival estimates to discuss the reliability of the methods to estimate recruitment rate; (B) Growth rate estimates; bars define the 95% confidence intervals
+(ref:srts) (A) Estimates of adult survival rate using population growth rates; (B) Comparison of the survival rate values between the two estimation methods; bars define the 95% confidence intervals
 
-(ref:R3cap) (A) Average recruitment rate estimates on the time intervals defined for survival analysis; (B) Average survival estimates deduced from growth rate and recruitment rate estimates on the time intervals; bars define the 95% confidence intervals
+(ref:mgr) (A) Estimates of the maximum population growth rate without harvest; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates; bars define the 95% confidence intervals
 
-(ref:R4cap) (A) Estimates of the population growth rate without harvest; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates; bars define the 95% confidence intervals
-
-(ref:R5cap) Comparison between a proxy of maximum recruitment rate and the average recrutiment rate estimated following *counting* and *sampling* methods; the proxy is estimated during the maximum growth periods by combining the growth rate to the hypothesis of a 100% adult survival (*max growth*); this hypothesis implies that the true maximum of recruitment rate is likely higher; bars define the 95% confidence intervals
+(ref:grcomp) Comparison between proxies of maximum recruitment rate and average recruitment rates estimated with and without harvest (*counting* method only); proxies are estimated by using the maximum growth rates; bars define the 95% confidence intervals
 

### 2021-12-03 Code/Model_recruitment_rate.R

@@ -266,7 +266,8 @@ count_sex_app %>%
   select(method, `male proportion`, tot) %>% 
   bind_rows(frag %>%  
               filter(sex != "ind") %>% 
-              filter(!(pop == "FR" & year(year) < 2011), !(pop == "UK" & year(year) > 2009)) %>%
+              filter(!(pop == "FR" & year(year) < 2011), 
+                     !(pop == "UK" & year(year) > 2009)) %>%
               group_by(year, pop) %>% 
               mutate(tot = sum(shot)) %>%  
               filter(sex == "mal") %>% 
@@ -417,7 +418,7 @@ JuvCode <- nimbleCode(
     
     # parameter estimation from apparent population structure
     
-    #hierarchical prior
+    # hierarchical prior
     for (i in 1:(C_id_max + S_id_max)) {
       shape1[i] ~ dunif(1, 1e3)
       shape2[i] ~ dunif(1, 1e3)
@@ -427,7 +428,7 @@ JuvCode <- nimbleCode(
     
     for (j in 1:C) {
       
-      #p_mal_ad prior
+      # p_mal_ad prior
       # p_mal_ad[j] ~  dbeta(shape1[C_id[j]], shape2[C_id[j]])
       p_mal_ad[j] ~  dbeta(1, 1)
       
@@ -468,12 +469,14 @@ JuvCode <- nimbleCode(
     # r estimation frow counts
     
     # priors
-     
+     max_surv ~ dunif(.7, 1)
+    
     for (r in 1:R_id_max) {
       
       r_avg[r] ~ dunif(log(0.01), log(5))
       N_sd[r] ~ dunif(0, 1e3)
       log(lambda[r]) <- r_avg[r]
+      max_prod[r] <- lambda[r] - max_surv
     }
     
     for (ts in 1:TS_id_max) {
@@ -531,6 +534,7 @@ JuvInit <- list(p_mal = 0.6,
                 prior = rep(.1, 3),
                 p_juv = rep(.1, JuvConst$C + JuvConst$S),
                 log_N_int = rep(0, 6),
+                max_surv = .9,
                 r_avg = rep(log(.5), 5),
                 N_sd = rep(1, 5))
 
@@ -540,8 +544,9 @@ JuvMon <- c("p_mal",
             "p_juv",
             "recruits",
             "productivity",
-            "lambda",
             "survival",
+            "lambda",
+            "max_prod",
             "log_N_int")
 
 # mcmc parameters
[…]

### 2021-12-03 Code/Draft_ruddy_duck_recruitment_rate.Rmd
pull issue fixed

@@
 author: 
-- Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr - contribution in data analyses & writing]
-- Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr - contribution in writing]
-- Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk - contribution in data collection]
-- Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr - contribution in data collection]
-- Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthi(UK).guillemain@ofb.gouv.fr - contribution in writing]
+- Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr]
+# - Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr - contribution in writing]
+# - Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk - contribution in data collection]
+# - Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr - contribution in data collection]
+# - Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthi(UK).guillemain@ofb.gouv.fr - contribution in writing]
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-The recruitment rate parameter is complex because it is composed of two main sub-parameters : the fecundity followed by the juvenile survival [@Etterson2011].
+The recruitment rate parameter is complex because it is composed of two main sub-parameters: the fecundity followed by the juvenile survival [@Etterson2011].
 Both sub-parameters can be split into other sub-parameters.
@@
 For such question, the carrying capacity of the habitat is theoretically not reached.
-So a manager should mostly focus on collecting knowledge on the maximum recruitment rate and its variability by removing the potential density-dependent effects occurring only for the highest theoretical densities [@Peron2013, @Eraud2021].
+So a manager should mostly focus on collecting knowledge on the maximum recruitment rate and its variability by removing the potential density-dependent effects occurring only for the highest theoretical densities [@Peron2013; @Eraud2021].
 
@@
 To do so, it is necessary to monitor nests in habitats representative of the managed population [@Blums2004].
-The survival of a chick to its recruitment can be done by capture mark recapture method quite tricky to set up [@Blums2004; @Schmidt2008, @Arnold2018].
+The survival of a chick to its recruitment can be done by capture mark recapture method quite tricky to set up [@Blums2004; @Schmidt2008; @Arnold2018].
 Even if such approach provides valuable information on the dynamics of the recruitment and can highlight the bottlenecks, the fieldwork required to estimate both components is too time-consuming to generalize such approach for waterfowl population management on a long-term horizon.
@@
 The hunting bag analysis might then lead to overestimate the recruitment rate and consequently the maximum sustainable harvest rate, with adverse consequences on population conservation [@Fox2014].
-
-Alternative method: trapping methods but still biased @Arnold2018
-Finally, because juveniles are more naïve, they may be more vulnerable to capture by
-standard trapping methods (Rguibi- Idrissi et al., 2003), even if loca-
-tions and timing were otherwise unbiased.
+Inferring recruitment rate from the proportion of immature birds trapped during tagging studies is also promising [@Arnold2018], but the author admits that juveniles might be more naïve and are then more likely caught in traps [@RguibiIdrissi2003].
 
@@
 A new population set up in France from the 90's, likely because of an arrival of a few individuals from the United Kingdom (UK).
-Unlike the UK population, the French one set up around a single wintering spot, the Grand Lieu Lake (47.09°N, 1.67°W), which falicitates the population monitoring.
+Unlike the UK population, the French one set up around a single wintering spot, the Grand Lieu Lake (47.09°N, 1.67°W), which facilitates the population monitoring.
 Since this species is considered as a threat to the white-headed duck because of the risk of a genetic introgression [@MunozFuentes2007], an European plan of eradication has been adopted and control measures were taken in both countries [@GutierrezExposito2020].
@@
 
-Since sex identification of samples is very accurate, such observation demonstrates that the counting method misses the identification of a significant part of males in the population.
+Assuming sex identification of samples is very accurate, such observation demonstrates that the counting method misses the identification of a significant part of males in the population.
 In many waterfowl species, immature individuals of both sexes display cover-up plumage similar to adult females [@Reeber2015].
@@
 Adult sex ratio in waterfowl is generally biased towards males (Figure \@ref(fig:pmal), @Wood2021).
-As a consequence, deducing the immature proportion from the adult male proportion is not straightforward because the adult female proportion is not just equal to the adult male proportion.
+As a consequence, deducing the immature proportion from the adult male proportion is not straightforward because the adult proportion twice as much as the male proportion.
 Even if the adult sex ratio may vary over a long time range, it is relatively stable over a few years [@Wood2021].
 Sex identification in adult samples from hunting bags was available for both ruddy duck populations.
-However, the small population size in France was limiting to provide useful sex proportion estimates.
+However, the small population size in France prevents from providing precise sex proportion estimates.
 A comparison over months in UK showed no difference in male proportion, so adult samples collected over the whole year were used to estimate the male proportion.
@@
 
-Since the additive property of the binomiale distribution, the proportion of immature individuals can be deduced from the cumulated counts of male-like individuals in the wintering population, see Equation \@ref(eq:pic).
+Assuming the additive property of the binomial distribution, the proportion of immature individuals can be deduced from the cumulated counts of male-like individuals in the wintering population, see Equation \@ref(eq:pic).
 Only years with more than 100 individuals were selected.
@@
 |Name|Class|Description|
-|:--|:----|:----------------------------|
-|$pm$|Parameter|Proportion of males in the adult part of population at its equilibrium|
-|$SM$|Data|Total number of adult males in samples|
-|$SF$|Data|Total number of adult females in samples|
-|$CM_{i, t}$|Data|Cumulated number of type-male individuals counted in population $i$ in year $t$|
-|$pi_{i, t}$|Parameter|Proportion of immature individuals in population $i$ in year $t$|
+|:---|:----|:----------------------------|
+|$p(m|a)$|Parameter|Proportion of males in adults (or probability to be an adult male)|
+|$SAM$|Data|Total number of adult males in samples|
+|$SAF$|Data|Total number of adult females in samples|
+|$CAM_{i, t}$|Data|Cumulated number of type-male individuals counted in population $i$ in year $t$, which are assumed to be adult males|
+|$p(m \cap a)_{i, t}$|Parameter|Proportion of adult males in population $i$ in year $t$|
+|$p(a)_{i, t}$|Parameter|Proportion adults in population $i$ in year $t$|
+|$p(i)_{i, t}$|Parameter|Proportion of immatures in population $i$ in year $t$|
 |$C_{i, t}$|Data|Cumulated number of individuals counted in population $i$ in year $t$|
@@
 \begin{equation}
-pm \sim {\sf Beta}(SM, SF)
+p(m|a) \sim {\sf Beta}(SAM, SAF)
 (\#eq:pm)
@@
 
-\begin{equation}
-CM_{i, t} \sim {\sf Binom}(pm.(1 - pi_{i, t}), C_{i, t})
+\begin{align}
+CAM_{i, t} & \sim {\sf Binom}(p(m \cap a)_{i, t}, C_{i, t}) \notag \\
+& \sim {\sf Binom}(p(m|a).p(a)_{i, t}, C_{i, t}) \notag \\
+& \sim {\sf Binom}(p(m|a).(1 - p(i)_{i, t}), C_{i, t})
 (\#eq:pic)
-\end{equation}
+\end{align}
 
 \begin{equation}
-r_{i, t} = \frac{pi_{i, t}.N_{i, t}}{N_{i, t - 1}}
+r_{i, t} = \frac{p(i)_{i, t}.N_{i, t}}{N_{i, t - 1}}
 (\#eq:r)
@@
 \begin{equation}
-R_{i, t} = pi_{i, t}.N_{i, t}
+R_{i, t} = p(i)_{i, t}.N_{i, t}
 (\#eq:R)
@@
 
-A direct estimation of the immature proportion can be done using samples from the hunting bag, see Equation \@ref(eq:pis).
-By combining estimates of immature proportion and Equation \@ref(eq:r), annual recruitment rates from samples were estimated to assess their consistency with the counting method.
+A direct estimation of the immature proportion can be made using samples from hunting bags, see Equation \@ref(eq:pis).
+By combining estimates of immature proportion and Equation \@ref(eq:r), recruitment rates from samples were estimated to check the variability consistency of the counting method.
 One selected only years with more than 100 individuals controlled in winter, which correspond to a 9-year period in UK.
@@
 |Name|Class|Description|
-|:--|:----|:----------------------------|
+|:---|:----|:----------------------------|
 |$SI_{i, t}$|Data|Number of immatures sampled in the wintering population $i$ in year $t$|
@@
 \begin{equation}
-SI_{i, t} \sim {\sf Binom}(pi_{i, t}, S_{i, t})
+SI_{i, t} \sim {\sf Binom}(p(i)_{i, t}, S_{i, t})
 (\#eq:pis)
@@
 
-The simple relationship (a bit more complex if one considers a species with delayed maturity [see @Robertson2008]) between population growth rate, adult survival, and recruitment rate can be used to test the accuracy of the order of magnitude of recruitment rates estimated from the counting and the sampling methods, see Equation \@ref(eq:GR).
-This relationship comes from a simple reasoning for a closed population : the population size in year $t$ is equal to the number of breeders that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
+The simple relationship (a bit more complex if a species with delayed maturity is considered [see @Robertson2008]) between population growth rate, adult survival, and recruitment rate can be used to test the accuracy of the order of magnitude of recruitment rates estimated from the counting and the sampling methods, see Equation \@ref(eq:GR).
+This relationship comes from a simple reasoning for a closed population: the population size in year $t$ is equal to the number of breeders that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
 the recruitment in year $t$.
@@
 
-Population growth rate can be assessed from the evolution of the breeding population size.
-Since there is observation error in population size estimates and process variability in population growth, only average population growth rates are estimated in this study following Equation \@ref(eq:L), which is a linear regression on the logarithm scale.
+From the evolution of the breeding population size, population growth rate can be assessed every year to catch its variability range \@ref(eq:GR), or can be smoothed over a consistent time period to provide average population growth rate, see Equation \@ref(eq:L), which is a linear regression on the logarithm scale.
 
@@
 |:--|:----|:----------------------------|
+|$N0_{i, J}$|Parameter|Intercept of the regression model|
 |$\lambda_{i, J}$|Parameter|Average population growth rate over a restricted time interval $J$ for a population $i$ (in $year^{-1}$)|
-|$N0_{i, J}$|Parameter|Intercept of the regression model|
 |$t \in J$|Index|Year index within the time interval $J$| 
@@
 The adult survival is commonly assessed by capture mark recapture [@Lebreton2001].
-Combined to population growth rate, adult survival rate allows recruitment rate to be estimated, and then compared with the output of the counting method.
-Unfortunately, no banding campaign was performed on the two ruddy duck populations.
-However, the relationship \@ref(eq:GR) could still produce adult survival estimates that can be discussed.
-To do so, population growth rate overlapping the counting time series were estimated.
+Combined to annual population growth rate, adult survival rate allows recruitment rate to be estimated each year, and then compared with the outputs of the counting method.
+Unfortunately, no tagging study was performed on the two ruddy duck populations.
+However, the relationship \@ref(eq:GR) could still produce annual adult survival estimates that can be discussed.
+To do so, annual population growth rates overlapping the counting time series were estimated.
 
@@
 Given a conservative assumption on the maximum adult survival of ruddy ducks and maximum population growth rate estimated during no harvest time period, one can provide a proxy of the maximum recruitment rate.
-If the recruitment rate outputs of the model are significantly superior to this proxy, it means that the method overestimates the recruitment rate, which means that some adult male individuals are not detected during count surveys.
-To do so, one uses a very conservative prior of adult survival by setting a uniform distribution on $[0.7;1]$ [the lower limit corresponds to the highest range of survival rates observed in litterature for species of similar weight, @Nichols1997; @Krementz1997; @Buxton2004].
+If the recruitment rate outputs of the model are significantly superior to this proxy, it means that the method overestimates the recruitment rate, which means that some adult males are not detected during count surveys.
+By spliting the recruitment rate outputs upon harvest category and country, one can discuss of the impact of the harvest on the recruitment rate.
+
+To do so, one estimates maximum population growth rates over time periods observing no harvest using Equation \@ref(eq:L).
 After reaching the threshold of 1000 individuals, the UK population growth observed a strong inflexion whereas no harvest pressure was applied.
-This observation led to split the analysis into two time series, see Figure \@ref(fig:grts).
+This observation led to split the analysis into two time series, see Figure \@ref(fig:grts) for the time periods used in the maximum growth rate inference.
+To get proxies of maximum recruitment rate, one uses a very conservative prior on the adult survival by setting a uniform distribution on $[0.7;0.9]$ [the lower limit corresponds to the upper range of survival rates observed in literature for species of similar weight, @Nichols1997; @Krementz1997; @Buxton2004].
 
@@
 For all the sub-models in Section \@ref(MM), the Bayesian framework was used for its efficiency and simplicity to propagate error through the parameters.
-One generated three chains of length 500000, with a thinning of 10 to avoid autocorrelation in the samples, and we discarded the first 2000 samples as burn-in.
+One generated three chains of length 500000, with a thinning of 10 to avoid autocorrelation in the samples, and one discarded the first 2000 samples as burn-in.
 Chain convergence was assessed using the Gelman and Rubin convergence diagnostic (R<1.1, @Gelman1992).
-We fit the models using NIMBLE [@Valpine2017] run from R [@RCT2020].
+One fit the models using NIMBLE [@Valpine2017] run from R [@RCT2020].
 Data and code are available here: https://github.com/adri-tab/Ruddy_duck_recruitment_rate.
@@
 There are significant year-to-year variations with a maximum amplitude of 0.70 recruits per breeder (12-time variability).
-Recruitment rates are estimated in both populations for 5 years, and no correlation is observed between the two populations.
+Recruitment rates are estimated in both populations over 5 years, and no correlation is observed between the two populations.
 
-Recruitment rates estimated from UK sample data are consistently higher than estimates from the corresponding count data Figure \@ref(fig:rrts)A.
+Recruitment rates estimated from UK sample data are consistently higher than estimates from the corresponding count data, Figure \@ref(fig:rrts)A.
 The correlation between the two methods is strong and points out that the sampling method overestimates the recruitment rate by 44% compared to the counting method, Figure \@ref(fig:rrts)B.
+The precision of the two methods is in the same range.
 
@@
 ## Consistency of the adult survival
+
+Adult surival rates were deduced from the recruitment rate estimates and the observed population growth rates, Figure \@ref(fig:srts).
+In UK, values that are obtained from the counting method ranged from 0.21 [0.20; 0.22] to 0.68 [0.65; 0.72].
+In France, survival rates vary from 0.39 [0.34; 0.44] to 1.08 [0.98; 1.17].
+Like recruitment rates, there is is significant year-to-year variability in both populations with a maximum amplitude of 0.69 (2.8-time vairability).
+
+No trend in France
 
@@
 
-As described before, recruitement rate over the studied time period is far lower for the UK than France (2.5 factor). 
+As described before, recruitment rate over the studied time period is far lower for the UK than France (2.5 factor). 
 Underlined adult survival are comprised between 0 and 1, whatever the method. So we cannot conclude if one can be rejected.
@@
 
-survival values compared to litterature? s
+survival values compared to literature? s
 
@@
 
-If counting method acceptable, moderate harvest pressure does not reduce the recruitment rate as expected, maybe due to density dependend mortality for juvenile but it affects it if strong harvest rate.
+If counting method acceptable, moderate harvest pressure does not reduce the recruitment rate as expected, maybe due to density dependent mortality for juvenile but it affects it if strong harvest rate.
 That is very interesting bro but warning,  these interpretation holds on weak estimations and adult survival equal to 1!!!
@@
 
-@Arnold2018 underlines the lack of data on recruitment rate for modelers (in discussion) and that new methods are required to track this variable that often drive the population growth rate
+@Arnold2018 underlines the lack of data on recruitment rate for modellers (in discussion) and that new methods are required to track this variable that often drive the population growth rate
 

### 2021-12-03 Code/Model_recruitment_rate.R

@@ -469,7 +469,7 @@ JuvCode <- nimbleCode(
     # r estimation frow counts
     
     # priors
-     max_surv ~ dunif(.7, 1)
+     max_surv ~ dunif(.7, .9)
     
     for (r in 1:R_id_max) {
       

### 2021-12-06 Code/Draft_ruddy_duck_recruitment_rate.Rmd
result writing

@@
 |:---|:----|:----------------------------|
-|$p(m|a)$|Parameter|Proportion of males in adults (or probability to be an adult male)|
 |$SAM$|Data|Total number of adult males in samples|
 |$SAF$|Data|Total number of adult females in samples|
+|$p(m|a)$|Parameter|Proportion of males in adults (or probability to be an adult male)|
 |$CAM_{i, t}$|Data|Cumulated number of type-male individuals counted in population $i$ in year $t$, which are assumed to be adult males|
+|$C_{i, t}$|Data|Cumulated number of individuals counted in population $i$ in year $t$|
 |$p(m \cap a)_{i, t}$|Parameter|Proportion of adult males in population $i$ in year $t$|
@@
 |$p(i)_{i, t}$|Parameter|Proportion of immatures in population $i$ in year $t$|
-|$C_{i, t}$|Data|Cumulated number of individuals counted in population $i$ in year $t$|
-|$R_{i, t}$|Parameter|Number of recruits in population $i$ in year $t$|
 |$N_{i, t}$|Data|Proxy of the size of the breeding population $i$ in year $t$ (maximum number of individuals counted in the wintering population minus removals before reproduction)|
 |$r_{i, t}$|Parameter|Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per breeder in year $t - 1$|
+|$R_{i, t}$|Parameter|Number of recruits in population $i$ in year $t$|
 
@@
 Given a conservative assumption on the maximum adult survival of ruddy ducks and maximum population growth rate estimated during no harvest time period, one can provide a proxy of the maximum recruitment rate.
-If the recruitment rate outputs of the model are significantly superior to this proxy, it means that the method overestimates the recruitment rate, which means that some adult males are not detected during count surveys.
-By spliting the recruitment rate outputs upon harvest category and country, one can discuss of the impact of the harvest on the recruitment rate.
+If the recruitment rate outputs of the model are significantly superior to this proxy, it means that the method overestimates the recruitment rate, which may reflect that some adult males are not detected during count surveys.
+By splitting the recruitment rate outputs upon population and harvest category, one can discuss of the impact of the harvest on the recruitment rate.
 
 To do so, one estimates maximum population growth rates over time periods observing no harvest using Equation \@ref(eq:L).
+Since the estimate of the UK population size is noisy below 500 individuals, one rejected data before 1972 (Figure \@ref(fig:grts)).
 After reaching the threshold of 1000 individuals, the UK population growth observed a strong inflexion whereas no harvest pressure was applied.
-This observation led to split the analysis into two time series, see Figure \@ref(fig:grts) for the time periods used in the maximum growth rate inference.
-To get proxies of maximum recruitment rate, one uses a very conservative prior on the adult survival by setting a uniform distribution on $[0.7;0.9]$ [the lower limit corresponds to the upper range of survival rates observed in literature for species of similar weight, @Nichols1997; @Krementz1997; @Buxton2004].
+This observation led to split the analysis into two time series for the time periods used in the maximum growth rate inference in UK (Figure \@ref(fig:grts)).
 
@@
 ```
+
+To get proxies of maximum recruitment rate from Equation \@ref(eq:GR), one uses a very conservative prior on the adult survival by setting a uniform distribution on $[0.7;0.9]$ [the lower limit corresponds to the upper range of survival rates observed in literature for species of similar weight, the upper limit corresponds to the range of survival rates of long-life species, @Nichols1997; @Krementz1997; @Buxton2004].
+The average recruitment rates from the count data are estimated for each population and for each harvest category and compared to the maximum recruitment rates.
 
@@
 Data and code are available here: https://github.com/adri-tab/Ruddy_duck_recruitment_rate.
-The values *X [min; max]* reported in results are means and the associated 95% confidence interval.
+The values *X[Y; Z]* reported in results are the medians and the associated boundaries of the 95% confidence interval of posterior distributions.
+The median was preferred to the mean because of its robustness to skewed distribution.
 
@@
 The adult male proportion, which is necessary to infer on the recruitment rate from count data, is estimated at 0.60 [0.59; 0.61].
-The counting method provides estimations of the recruitment rate, Figure \@ref(fig:rrts)A.
+The counting method provides estimations of the recruitment rate, see Figure \@ref(fig:rrts)A.
 In UK, values vary from 0.06 [0.05; 0.07] to 0.27 [0.20; 0.34] recruits per breeder, and in France, 0.08 [0.04; 0.12] to 0.76 [0.64; 0.88] recruits per breeder.
-There are significant year-to-year variations with a maximum amplitude of 0.70 recruits per breeder (12-time variability).
+There are significant year-to-year variations with a maximum amplitude of 0.68 recruits per breeder in France (9.5-time variability).
+On the available period, recruitment rate in UK is smaller than in France.
 Recruitment rates are estimated in both populations over 5 years, and no correlation is observed between the two populations.
+The lowest recruitment rate observed in France, which is in 2019, corresponds to the start of the depletion of the population.
+It is noticeable that it is in the same range than the UK values, also estimated during a strong population depletion.
 
@@
 The correlation between the two methods is strong and points out that the sampling method overestimates the recruitment rate by 44% compared to the counting method, Figure \@ref(fig:rrts)B.
-The precision of the two methods is in the same range.
+Estimation precision is similar for the two methods.
+The UK population started to deplete from 2006, but no significant shift is observed between the two periods.
 
@@
 
-## Consistency of the adult survival
+## Consistency of the corresponding adult survival
 
-Adult surival rates were deduced from the recruitment rate estimates and the observed population growth rates, Figure \@ref(fig:srts).
+Adult survival rates were deduced from the recruitment rate estimates and the observed population growth rates, see Figure \@ref(fig:srts)A.
 In UK, values that are obtained from the counting method ranged from 0.21 [0.20; 0.22] to 0.68 [0.65; 0.72].
 In France, survival rates vary from 0.39 [0.34; 0.44] to 1.08 [0.98; 1.17].
-Like recruitment rates, there is is significant year-to-year variability in both populations with a maximum amplitude of 0.69 (2.8-time vairability).
+Like recruitment rates, there is significant year-to-year variability without clear temporal structure in both populations with a maximum amplitude of 0.69 in France (2.8-time variability).
+Unlike the recruitment rate, the survival rate in 2019 in France is not a low outlier.
 
-No trend in France
-
-Average population growth rates were estimated over the period of available count data with a consistent harvest pressure.
-Average population growth rates ($\lambda$) are estimated to 0.56 [0.48; 0.65] and 0.97 [0.94; 1.00] for the UK and France respectively, which corresponds to an average annual growth proportion ($\lambda - 1$) of -0.44 [-0.52; -0.35] and -0.03 [-0.06; 0.00] in the UK and France respectively
-
-comprised between 0 and 1 like adult survival by definition, meaning similar weight to build population growth.
-Far higher in France than in UK, to be related to the growth rates differences (-44% vs -3%)
-
-Higher on average in the early years in France when (almost) no harvest pressure (1999- 2004) but large interannual fluctuations without clear temporal trends,but small temporal correlation, especially in UK.
-
-quite stable when harvest rate is limited (France 2000), but is lower and noisier when harvest rate is huge and the population depleted.
-
-Depletion is clear on the counting time series intervals (- 44% on average per year) compared to the not significant decrease in France over the time interval with harvest (-3% per year on average).
+As expected, survival rates estimated from UK sample data are consistently lower than estimates from the corresponding count data, Figure \@ref(fig:srts)A.
+The correlation between the two methods points out that the sampling method underestimates the survival rate by 32% compared to the counting method, Figure \@ref(fig:srts)B.
+Estimation precision is similar for the two methods.
+The sample time series shows that the adult survival is significantly lower after 2006, which corresponds to the start of the depletion of the UK population.
 
@@
 
+## Comparison to a proxy of maximum recruitment rate
+
+Maximum population growth rates are similar for the two populations, 1.45[1.36; 1.55] and 1.47[1.29; 1.68] for UK and France respectively, which corresponds to an increase of about 45% per year (Figure \@ref(fig:mgr)).
+After reaching a certain size (~ 1000 individuals), the UK population growth rate falls to 1.06[1.04; 1.07], which corresponds to an increase of 6% per year, even though no harvest pressure was applied on it.
+
+```{r mgr, fig.cap = "(ref:mgr)"}
+knitr::include_graphics("../Output/plot_6.png")
+```
+
+Upon the conservative assumption of a maximum adult survival uniformly distributed over 0.70 to 0.90, the maximum recruitment rate reaches 0.65[0.51; 0.75] and 0.66[0.45; 0.90] recruits per breeder for UK and France respectively (Figure \@ref(fig:grcomp)).
+When there is no harvest, the estimated average recruitment rate of the French population is in the range of the proxy of the maximum recruitment rate, 0.51[0.46; 0.57] recruits per breeder.
+When there is harvest, the average recruitment rates are 0.14[0.12; 0.16] and 0.33[0.31; 0.36] recruits per breeder for the UK and France respectively.
+Both values are significantly lower than the maximum recruitment rate, and it is noticeable that average recruitment rate is lower for the UK population, which is in depletion over the count time series, than for the French population, which of the dynamic is mostly stable over the monitored period.
+
+```{r grcomp, fig.cap = "(ref:grcomp)"}
+knitr::include_graphics("../Output/plot_7.png")
+```
+
+# Discussion
+
+Count data differentiating adult males from other individuals in waterfowl are successfully used in this study to provide an estimation of recruitment rates.
+
+*facts*
+
+This demonstrates that methods provide no inconsistent recruitment rate values, which is quite satisfying and cannot reject the methods.
+
+# reparler des notions en introduction
+rajouter en intro cette notion.
+a same value of age ratio can correspond to different fecundity levels, so warning. 
+a same value of age ratio for a population in depletion and a population stabilized are not similary productive, the first one is less productive than the second one ! 
+so warning, focus on recruitment rate and not on age ratio if the dynamics of the population varies!
+recruitment proportion is commonly used in to discuss the fecundity of a population, but unlike the recruitment rate, this indicator does not consider the dynamics of the population and does reflect fecundity only on stable growth period
+
+If counting method acceptable, moderate harvest pressure does not reduce the recruitment rate as expected, maybe due to density dependent mortality for juvenile but it affects it if strong harvest rate.
+That is very interesting bro but warning,  these interpretation holds on weak estimations and adult survival equal to 1!!!
+
+the consistency between the two max growth rate strenghtens that seems to be a biological limit for this species, at least in the western Palearctic.
+
 As described before, recruitment rate over the studied time period is far lower for the UK than France (2.5 factor). 
 Underlined adult survival are comprised between 0 and 1, whatever the method. So we cannot conclude if one can be rejected.
+
+France affected the recruitment, UK affected adult survival for a similar result -> depletion
 
@@
 
-## Comparison to a proxy of maximum recruitment rate
-
-Maximum growth rate are similar for both populations (+45%), which supports the consistency of the species.
-After a certain threshold, the growth rate falls to 6% per year even though no harvest pressure was applied on it, maybe density dependent effects.
-
-```{r mgr, fig.cap = "(ref:mgr)"}
-knitr::include_graphics("../Output/plot_6.png")
-```
-
-Upon the conservative assumption of adult survival of .85, maximum recruitment rate reach .45 recruits per breeder.
-For both populations and methods, values remains below this maximum recruitment rate.
-This demonstrates that methods provide no inconsistent recruitment rate values, which is quite satisfying and cannot reject the methods.
-
-```{r grcomp, fig.cap = "(ref:grcomp)"}
-knitr::include_graphics("../Output/plot_7.png")
-```
-
-If counting method acceptable, moderate harvest pressure does not reduce the recruitment rate as expected, maybe due to density dependent mortality for juvenile but it affects it if strong harvest rate.
-That is very interesting bro but warning,  these interpretation holds on weak estimations and adult survival equal to 1!!!
-
-# Discussion
+Abundance index are considered exact, which should not be true. and lead to structural error in the estimates of annual growth rate.
 
@@
 
+Message de l'article
+
 # Aknowledgements {-}
@@
 
-(ref:mgr) (A) Estimates of the maximum population growth rate without harvest; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates; bars define the 95% confidence intervals
+(ref:mgr) (A) Estimates of the maximum population growth without harvest $(\lambda - 1)$; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates; bars define the 95% confidence intervals
 

### 2021-12-07 Code/Draft_ruddy_duck_recruitment_rate.Rmd
solving issue of abstract in yaml

@@
 # - Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr - contribution in data collection]
-# - Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthi(UK).guillemain@ofb.gouv.fr - contribution in writing]
+# - Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthieu.guillemain@ofb.gouv.fr - contribution in writing]
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-abstract: "The response of a waterfowl population to a harvest pressure depends on its capacity to renew.
+# abstract: "The response of a waterfowl population to a harvest pressure depends on its capacity to renew.
 The recruitment, i.e.
@@
                       fig.align = "center",
-                      fig.pos = 'H') # hold position
+                      fig.pos = 'H') # hold position / require package "float"
 
@@
 ```
+
+# Abstract {-}
+
+The response of a waterfowl population to a harvest pressure depends on its capacity to renew.
+The recruitment, i.e.
+the number of young adults reproducing for the first time, is a key indicator to describe the renewal of a population, and therefore an essential tool in population management.
+The recruitment rate, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
+The proportion of young adult in a waterfowl population, which is a main step in estimating the recruitment rate, is often estimated from game-hunting samples.
+However, this proportion is not accessible in years without harvest, or with a low harvest rate.
+Moreover, the age-structure in the harvest samples does not necessarily reflect the underlying age-structure of the population.
+It is often skewed towards immatures and can lead to an overestimation of the recruitment rate.
+In waterfowl, adult males usually display brighter colours than females and immatures from both sexes.
+This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
+Such information can be used to infer the proportion of immatures, and consequently the recruitment rate.
+In using two populations of ruddy duck, *Oxyura oxyura*, this study introduces a model to estimate the recruitment rate from count data.
+To test the accuracy of the method, the results are compared to recruitment rate estimates from samples, and their consistency to support the observed population growth rates is also tested.
+The results suggest that the counting method is a versatile tool to estimate the recruitment rate if the species is monitored during the appropriate time window.
+Authors argue for considering a two-category count protocol in winter surveys of waterfowl species for which adult males can be differentiated from other individuals.
+
+**Keywords**: fecundity - productivity - reproductive success - age ratio - juvenile - duck
 
@@
 the recruits, to the number of all individuals at the reproduction season of the previous year, $t - 1$.
-The declination for species with age at first reproduction higher than 1 follows the same process [see @Robertson2008].
+The declination for species with delayed maturity follows the same process [see @Robertson2008].
 
@@
 
-Population conservation, harvesting, and control require tools in population management to reach targets such as a population size for example [@Shea1998].
+Population conservation, harvesting, and control require tools in population management in order to reach quantitative targets, such as a population size for example [@Shea1998].
 To do so, a first "blind" strategy is to test iteratively management rules to reach progressively an optimum.
@@
 The difficulty to track juvenile survival [@Schmidt2008] demonstrates the interest to focus on the recruitment rate when accessible.
-The key to predict the growth rate of a population is then to estimate the average value and the variability of the recruitment rate, potentially influenced by the harvest pressure [@Anders2005].
+The key to predict the growth rate of a population is then to estimate the value of the recruitment rate and its variability, potentially influenced by the harvest pressure [@Anders2005].
 
@@
 The population renewal is slowed down because the recruitment rate parameter is affected by density-dependent effects [@Gunnarsson2013].
-For an endemic species evolving in stable environmental conditions, a balance between the recruitment rate and the adult mortality rate explains the stabilization of the number of individuals [@Flint2015].
+For an endemic species evolving in stable environmental conditions, the recruitment rate compensates on average the adult mortality rate, which explains the stabilization of the number of individuals [@Flint2015].
 
@@
 The heterogeneity of quality among immature individuals might explain this by inducing a compensation process if the harvest selects mostly individuals with a low reproductive potential [@Lindberg2013; @Gimenez2017].
-The recruitment rate can even be enhanced because the harvest pressure reduces the densities and potentially the competition for space and food [@Nummi2015; @Peron2012].
-In conclusion, a limited exploitation of a population observing heterogeneity and/or density-dependent effect might increase or at least not affect the recruitment rate.
-Under moderate exploitation, a new balance between adult mortality and recruitment rate can be theoretically reached [@Tsikliras2018].
+The recruitment rate can even increase because the harvest pressure reduces the densities and potentially the competition for space and food [@Nummi2015; @Peron2012].
+In conclusion, a limited exploitation of a population observing heterogeneity and/or density-dependent process might paradoxically increase or at least not affect the recruitment rate.
+Under moderate exploitation, a new balance between recruitment rate and adult mortality can be theoretically reached [@Tsikliras2018].
 The size of a newly exploited population in stable environmental conditions should thus reach a different equilibrium, which is expected to be lower than the size without exploitation.
@@
 The recruitment rate parameter of a population can be assessed first by evaluating the fecundity of the breeding population and then by monitoring the survival from the egg life stage to the first-breeding event.
-To do so, it is necessary to monitor nests in habitats representative of the managed population [@Blums2004].
+To do so, it is necessary to monitor nests in habitats representative of the considered population [@Blums2004].
 The survival of a chick to its recruitment can be done by capture mark recapture method quite tricky to set up [@Blums2004; @Schmidt2008; @Arnold2018].
@@
 Using two other approaches using average population growth rates, one testes the consistency of the order of magnitude of the recruitment rate estimated from the counting method.
+The results also show insights on the response of the recruitment rate to a gradient of the harvest pressure.
 
@@
 
-To get proxies of maximum recruitment rate from Equation \@ref(eq:GR), one uses a very conservative prior on the adult survival by setting a uniform distribution on $[0.7;0.9]$ [the lower limit corresponds to the upper range of survival rates observed in literature for species of similar weight, the upper limit corresponds to the range of survival rates of long-life species, @Nichols1997; @Krementz1997; @Buxton2004].
+To get proxies of maximum recruitment rate from Equation \@ref(eq:GR), one uses a very conservative prior on the adult survival by setting a uniform distribution on $[0.7; 0.9]$ [the lower limit corresponds to the upper range of survival rates observed in literature for species of similar weight, the upper limit corresponds to the range of survival rates of long-life species, @Nichols1997; @Krementz1997; @Buxton2004].
 The average recruitment rates from the count data are estimated for each population and for each harvest category and compared to the maximum recruitment rates.
@@
 
-Maximum population growth rates are similar for the two populations, 1.45[1.36; 1.55] and 1.47[1.29; 1.68] for UK and France respectively, which corresponds to an increase of about 45% per year (Figure \@ref(fig:mgr)).
-After reaching a certain size (~ 1000 individuals), the UK population growth rate falls to 1.06[1.04; 1.07], which corresponds to an increase of 6% per year, even though no harvest pressure was applied on it.
+Maximum population growth rates are similar for the two populations, 1.45 [1.36; 1.55] and 1.47 [1.29; 1.68] for UK and France respectively, which corresponds to an increase of about 45% per year (Figure \@ref(fig:mgr)).
+After reaching a certain size (~ 1000 individuals), the UK population growth rate falls to 1.06 [1.04; 1.07], which corresponds to an increase of 6% per year, even though no harvest pressure was applied on it.
 
@@
 
-Upon the conservative assumption of a maximum adult survival uniformly distributed over 0.70 to 0.90, the maximum recruitment rate reaches 0.65[0.51; 0.75] and 0.66[0.45; 0.90] recruits per breeder for UK and France respectively (Figure \@ref(fig:grcomp)).
-When there is no harvest, the estimated average recruitment rate of the French population is in the range of the proxy of the maximum recruitment rate, 0.51[0.46; 0.57] recruits per breeder.
-When there is harvest, the average recruitment rates are 0.14[0.12; 0.16] and 0.33[0.31; 0.36] recruits per breeder for the UK and France respectively.
+Upon the conservative assumption of a maximum adult survival uniformly distributed over 0.70 to 0.90, the maximum recruitment rate reaches 0.65 [0.51; 0.75] and 0.66 [0.45; 0.90] recruits per breeder for UK and France respectively (Figure \@ref(fig:grcomp)).
+When there is no harvest, the estimated average recruitment rate of the French population is in the range of the proxy of the maximum recruitment rate, 0.51 [0.46; 0.57] recruits per breeder.
+When there is harvest, the average recruitment rates are 0.14 [0.12; 0.16] and 0.33 [0.31; 0.36] recruits per breeder for the UK and France respectively.
 Both values are significantly lower than the maximum recruitment rate, and it is noticeable that average recruitment rate is lower for the UK population, which is in depletion over the count time series, than for the French population, which of the dynamic is mostly stable over the monitored period.
@@
 
-*facts*
+## facts
 
@@
 
-# reparler des notions en introduction
+## reparler des notions en introduction
+
+figure 8 -> One said that recruitment rate can increase if harvest because of heterogenity and density dependence, but it is not what we observe in France, likely because the exploitation is high inhere and consequently impact.
+
+obviously, harvest impact the pre
+
 rajouter en intro cette notion.
@@
 
-(ref:S1cap) (A1) Estimates of the proportion of the immature individuals in the wintering population; (B1) estimates of the number of recruits, i.e.
+(ref:sup) (A1) Estimates of the proportion of the immature individuals in the wintering population; (B1) estimates of the number of recruits, i.e.
 the young adults reproducing for their first time; (A2) & (B2) comparison of the values between the two estimation methods; bars define the 95% confidence intervals

### 2021-12-07 Code/Model_recruitment_rate.R

@@ -21,7 +21,8 @@ theme_set(theme_bw())
 
 tibble(id = 1:20) %>% 
   ggplot(aes(x = 1, y = id, color = factor(id))) +
-  geom_point(size = 10) + scale_y_continuous(aes(breaks = id)) -> p; p
+  geom_point(size = 10) + 
+  scale_y_continuous(aes(breaks = id)) -> p; p
 
 ggplot_build(p)$data
 

### 2021-12-07 Code/Draft_ruddy_duck_recruitment_rate.Rmd
update

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-# abstract: "The response of a waterfowl population to a harvest pressure depends on its capacity to renew.
-The recruitment, i.e.
-the number of young adults reproducing for the first time, is a key indicator to describe the renewal of a population, and therefore an essential tool in population management.
-The recruitment rate, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
-The proportion of young adult in a waterfowl population, which is a main step in estimating the recruitment rate, is often estimated from game-hunting samples.
-However, this proportion is not accessible in years without harvest, or with a low harvest rate.
-Moreover, the age-structure in the harvest samples does not necessarily reflect the underlying age-structure of the population.
-It is often skewed towards immatures and can lead to an overestimation of the recruitment rate.
-In waterfowl, adult males usually display brighter colours than females and immatures from both sexes.
-This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
-Such information can be used to infer the proportion of immatures, and consequently the recruitment rate.
-In using two populations of ruddy duck, *Oxyura oxyura*, this study introduces a model to estimate the recruitment rate from count data.
-To test the accuracy of the method, the results are compared to recruitment rate estimates from samples, and their consistency to support the observed population growth rates is also tested.
-The results suggest that the counting method is a versatile tool to estimate the recruitment rate if the species is monitored during the appropriate time window.
-Authors argue for considering a two-category count protocol in winter surveys of waterfowl species for which adult males can be differentiated from other individuals.
-\\par \\textbf{Keywords:} fecundity - productivity - reproductive success - age ratio - juvenile - duck"
-output: 
-  bookdown::pdf_document2:
-    latex_engine: xelatex
-    toc: false
-header-includes:
-  \usepackage{float}
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, 
-                      warning = FALSE, 
-                      message = FALSE, 
-                      out.width = "100%", 
-                      fig.align = "center",
-                      fig.pos = 'H') # hold position / require package "float"
-
-Sys.setlocale("LC_ALL", "English")
-
-```
-
-# Abstract {-}
-
+abstract: | 
   The response of a waterfowl population to a harvest pressure depends on its capacity to renew.
@@
   **Keywords**: fecundity - productivity - reproductive success - age ratio - juvenile - duck
+output: 
+    bookdown::pdf_document2:
+    toc: false
+    latex_engine: xelatex
+header-includes:
+  \usepackage{float}
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, 
+                      warning = FALSE, 
+                      message = FALSE, 
+                      out.width = "100%", 
+                      fig.align = "center",
+                      fig.pos = 'H') # hold position / require package "float"
+
+Sys.setlocale("LC_ALL", "English")
+
+```
 
@@
 
-Count data differentiating adult males from other individuals in waterfowl are successfully used in this study to provide an estimation of recruitment rates.
+Count data differentiating adult males from other individuals in waterfowl are successfully used in this study to provide estimations of recruitment rates.
 
-## facts
+# Accuracy of the recruitment rate
+
+As expected, always lower than samples. Variability consistent to the samples as well, but slightly enhanced compared to the raw immature proportion \@ref(fig:sup)
 
@@
 
-figure 8 -> One said that recruitment rate can increase if harvest because of heterogenity and density dependence, but it is not what we observe in France, likely because the exploitation is high inhere and consequently impact.
+figure 8 -> One said that recruitment rate can increase if harvest because of heterogeneity and density dependence, but it is not what we observe in France, likely because the exploitation is high inhere and consequently impact.
 

### 2021-12-09 Code/Draft_ruddy_duck_recruitment_rate.Rmd
update

@@
 lang: en-EN
-bibliography: C:/Users/adrien.tableau/Dropbox/JabRef/Library_OFB.bib
+bibliography: C:/Users/Adrien/Dropbox/JabRef/Library_OFB.bib
 csl: https://www.zotero.org/styles/apa
@@
 - Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr]
-# - Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr - contribution in writing]
-# - Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk - contribution in data collection]
-# - Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr - contribution in data collection]
-# - Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthieu.guillemain@ofb.gouv.fr - contribution in writing]
+# - Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
+- Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk]
+- Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr]
+# - Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthieu.guillemain@ofb.gouv.fr]
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
     bookdown::pdf_document2:
+      latex_engine: xelatex
       toc: false
-    latex_engine: xelatex
+    bookdown::word_document2:
+      toc: false
 header-includes:
   \usepackage{float}
+
 ---
@@
 
-Population conservation, harvesting, and control require tools in population management in order to reach quantitative targets, such as a population size for example [@Shea1998].
+Population conservation, harvesting, and control require tools in population management in order to reach quantitative targets, such as a population size for example [@Shea1998; @Johnson2021].
 To do so, a first "blind" strategy is to test iteratively management rules to reach progressively an optimum.
@@
 When there is harvest, the average recruitment rates are 0.14 [0.12; 0.16] and 0.33 [0.31; 0.36] recruits per breeder for the UK and France respectively.
-Both values are significantly lower than the maximum recruitment rate, and it is noticeable that average recruitment rate is lower for the UK population, which is in depletion over the count time series, than for the French population, which of the dynamic is mostly stable over the monitored period.
+Both values are significantly lower than the maximum recruitment rate, and it is noticeable that the average recruitment rate is lower for the UK population, which is in depletion over the count time series, than for the French population, which of the dynamic is mostly stable over the monitored period.
 
@@
 
-# Accuracy of the recruitment rate
+# Accuracy of the counting method
 
-As expected, always lower than samples. Variability consistent to the samples as well, but slightly enhanced compared to the raw immature proportion \@ref(fig:sup)
+As expected, countalways lower than samples. Variability consistent to the samples as well, but slightly enhanced compared to the raw immature proportion \@ref(fig:sup)
 
 This demonstrates that methods provide no inconsistent recruitment rate values, which is quite satisfying and cannot reject the methods.
+
+A significant change in adult sex ratio would bias the estimation of recruitment rate.
+Since this measure integrates many bird cohorts, especially if it is a long-living species, updates on this ratio are not necessary on a year based
+
+#
+
+# Impact of harvest on recruitment rate
 

### 2021-12-09 Code/Draft_ruddy_duck_recruitment_rate.Rmd
Discussion update

@@
 lang: en-EN
-bibliography: C:/Users/Adrien/Dropbox/JabRef/Library_OFB.bib
+bibliography: C:/Users/adrien.tableau/Dropbox/JabRef/Library_OFB.bib
 csl: https://www.zotero.org/styles/apa
@@
 
-# Accuracy of the counting method
+## Accuracy of the counting method
 
-As expected, countalways lower than samples. Variability consistent to the samples as well, but slightly enhanced compared to the raw immature proportion \@ref(fig:sup)
+*comparison to sample estimation*
+
+As expected, count always lower than samples. Variability consistent to the samples as well, but slightly enhanced compared to the raw immature proportion \@ref(fig:sup)
 
@@
 A significant change in adult sex ratio would bias the estimation of recruitment rate.
-Since this measure integrates many bird cohorts, especially if it is a long-living species, updates on this ratio are not necessary on a year based
+However, since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong.
+Significant variations can then occur only every multiple years, especially if it is a long-living species.
+Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.
 
-#
+*survival consistency*
 
-# Impact of harvest on recruitment rate
+survival in the range of the expected values -> so no apparent issue of scaling issue, 
+so we the hypothesis that male-like individuals corresponds only to adult male is validated.
 
-## reparler des notions en introduction
+*average recruitment rate compared to a proxy of maximum recruitment rate*
 
-figure 8 -> One said that recruitment rate can increase if harvest because of heterogeneity and density dependence, but it is not what we observe in France, likely because the exploitation is high inhere and consequently impact.
+always below the maximum recruitment rate so satisfying.
 
-obviously, harvest impact the pre
+overall, even if we don't have access to the true value of the recruitment rate to test the accuracy of the counting method, a beam of arguments support that the method is not heavily biased, unlike the sampling method, as demonstrated in @Fox2014, even if one considers a picture limited to the end of the hunting season [@Fox2016].
 
-rajouter en intro cette notion.
+## Impact of harvest
+
+Survival higher than one but warning, observation error, so annual population growth rate might be inaccurate on some years.
+Also, we assumed a closed population, but it is possible that years with adult survival higher than one corresponds to an arrival of birds from the other population.
+As a consequence, the recruitment rate on these years may also be overestimated.
+
+No impact on the adult survival in France, but it seems that there is an impact in UK.
+Why, because harvest occurs mostly in summer in France, and in winter in UK.
+
+Impact on the recruitment rate in France, and no impact in UK.
+
+Overall, the harvest seems to impact differently France and UK, likely because the harvest mortality occurs mostly in summer in France, and in winter in UK.
+
+But caution, a population model linking the harvest rate to state variables describing the population is necessary to check precisely the impact of the harvest pressure, and test hypothesis on the functioning of harvest mortality.
+
+Recruitment rate decrease with the harvest for the French population, and the UK population recruitment rate is also below the French one. 
+it seems quite clear that the harvest affects the recruitment rate.
+
+*reference to introduction*
+
+It is quite consistent with the assumption in introduction.
+Since studied populations are in colonization of their habitat, maximum densities are not reached, and then no or limited density dependent effect occur.
+
+If counting method acceptable, moderate harvest pressure does not reduce the recruitment rate as expected, maybe due to density dependent mortality for juvenile but it affects it if strong harvest rate.
+That is very interesting bro but warning, these interpretation holds on weak estimations and adult survival equal to 1!!!
+
+it may be possible that first sight of density dependence occur in UK from 1000 individuals, but was not big enough to fully curb the population growth.
+Since there is a strong correlation between the number of birds observed and the number of sites occupied by the birds, this inflexion may be simply due to a limited number of new favourable sites to be colonized.
+
+What about heterogeneity, figure 8 -> One said that recruitment rate can increase if harvest because of heterogeneity and density dependence, but it is not what we observe in France, likely because the exploitation is high inhere and consequently impact.
+
+## Use of recruitment rate
+
+*better than sex ratio*
 a same value of age ratio can correspond to different fecundity levels, so warning. 
@@
 recruitment proportion is commonly used in to discuss the fecundity of a population, but unlike the recruitment rate, this indicator does not consider the dynamics of the population and does reflect fecundity only on stable growth period
-
-If counting method acceptable, moderate harvest pressure does not reduce the recruitment rate as expected, maybe due to density dependent mortality for juvenile but it affects it if strong harvest rate.
-That is very interesting bro but warning, these interpretation holds on weak estimations and adult survival equal to 1!!!
 
@@
 
-# Aknowledgements {-}
+# Acknowledgments {-}
 
@@
 This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the *Office Français de la Biodiversité*, and Jean-Marc Gillier from the *Société Nationale pour la Protection de la Nature*.
-The authors aknowledge all the contributors of the data collection in the UK and in France, especially: **UK names**, Vincent Fontaine, Denis Lacourpaille, Jules Joly, Justin Potier, Valentin Boniface, Alexis Laroche, and Médéric Lortion, **other names**.
+The authors acknowledge all the contributors of the data collection in the UK and in France, especially: **UK names**, Vincent Fontaine, Denis Lacourpaille, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface.
 

### 2021-12-13 Code/Draft_ruddy_duck_recruitment_rate.Rmd
start of the discussion

@@
 One defines the recruitment rate of a breeding population as the average number of recruits produced per each breeder [@Flint2015].
-This parameter might sometimes be called "annual fertility" [@Koons2014], "fecundity" [@Arnold2018], "reproductive rate" [@Cooch2014], "reproductive success" [@Etterson2011], or "productivity" [@Johnson1987; @Hagen2008].
+This parameter might sometimes be called *annual fertility* [@Koons2014], *fecundity* [@Arnold2018], *reproductive rate* [@Cooch2014], *reproductive success* [@Etterson2011], or *productivity* [@Johnson1987; @Hagen2008].
 If one considers a species reaching its sexual maturity as early as its first year, all individuals are mature at the reproduction season.
@@
 Population conservation, harvesting, and control require tools in population management in order to reach quantitative targets, such as a population size for example [@Shea1998; @Johnson2021].
-To do so, a first "blind" strategy is to test iteratively management rules to reach progressively an optimum.
+To do so, a first blind strategy is to test iteratively management rules to reach progressively an optimum.
 Such approach can be time-consuming given the intrinsic variability of the population response to its changing environment.
@@
 
-From two populations of ruddy duck, an species introduced in Europe in the 40's, this study introduces a simple Bayesian model to infer recruitment rate from count data.
+From two populations of ruddy duck, a species introduced in Europe in the 40's, this study introduces a simple Bayesian model to infer recruitment rate from count data.
 A large part of this study focuses on assessing the viability of such approach.
-A comparison to the results of the common approach using samples from hunting bags aims at checking if the recruitment rate variability is well tracked and if the hypothesis of the recruitment rate overestimation by sampling methods is consistent with the results.
-Using two other approaches using average population growth rates, one testes the consistency of the order of magnitude of the recruitment rate estimated from the counting method.
-The results also show insights on the response of the recruitment rate to a gradient of the harvest pressure.
+The comparison between two populations is a first step in testing the consistency of the method.
+A comparison to the common approach using samples from hunting bags aims at checking if the recruitment rate variability is well tracked and if the hypothesis stating that the recruitment rate is overestimated when using sample data is consistent with the results.
+Using two other approaches using population growth rates, one testes the consistency of the order of magnitude of the recruitment rate estimated from the counting method.
+Studying two populations is also an opportunity to discuss the response of the recruitment rate to a gradient of harvest pressure.
 
@@
 
-Count data differentiating adult males from other individuals in waterfowl are successfully used in this study to provide estimations of recruitment rates.
+Count data in waterfowl that differentiate adult males from other individuals are successfully used in this study to infer estimations of recruitment rates.
+In the following section, one checks the accuracy of the method from indicators developed for this purpose.
+From the outputs on the two populations of ruddy duck, one discusses about the effects of the harvest on the recruitment rate and its implication on waterfowl management.
 
@@
 
-*comparison to sample estimation*
+The recruitment rates from the counting method are in the same range for the two populations.
+The higher variability observed in France is not unexpected because the French time series is longer than in UK and corresponds to a larger harvest spectrum.
+There is no temporal correlation between the two populations, so it suggests that the recruitment rate is more likely influenced by regional-scale factors than large-scale ones.
+This is consistent with the literature because **[ref required]** shows that nesting success and hatching success are conditional to water levels, and that juvenile survival in waterfowl depends mostly on weather conditions [@Blums2004].
 
-As expected, count always lower than samples. Variability consistent to the samples as well, but slightly enhanced compared to the raw immature proportion \@ref(fig:sup)
-
-This demonstrates that methods provide no inconsistent recruitment rate values, which is quite satisfying and cannot reject the methods.
+Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation to the outputs of the counting method demonstrates its ability to track the recruitment rate variability.
+This result is quite robust because the correlation is obvious with only five years available.
+However, this conclusion might be overrated because two successive population size values are part of the recruitment rate, which mechanically scale the outputs.
+It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit less obvious than for the recruitment rate (Figure \@ref(fig:sup)A).
 
@@
 
+The counting method always provides recruitment rates significantly lower than the sampling method, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl.
+This is a robust result because the approaches used to control ruddy ducks differ strongly from usual hunting practices and are apriori less selective.
+It is especially clear in the UK, for which control operations occur in winter using long-range rifles that prevents the bird from escaping from the threat.
+The assumption on the accuracy of the counting method is then not excluded from the consistency of this result.
+
 *survival consistency*
+
+Annual adult survival rates deduced from observed population growth rates and recruitment estimates are on average in line with the literature **ref required**.
 
@@
 
-## Impact of harvest
+## Influence of harvest on the recruitment rate
+
+## Implication in waterfowl management
 

### 2021-12-14 Code/Draft_ruddy_duck_recruitment_rate.Rmd
update before conflict text

@@
 The recruitment rates from the counting method are in the same range for the two populations.
-The higher variability observed in France is not unexpected because the French time series is longer than in UK and corresponds to a larger harvest spectrum.
+The higher variability observed in France is not unexpected because the French time series is longer and covers a larger harvest spectrum than in UK.
 There is no temporal correlation between the two populations, so it suggests that the recruitment rate is more likely influenced by regional-scale factors than large-scale ones.
-This is consistent with the literature because **[ref required]** shows that nesting success and hatching success are conditional to water levels, and that juvenile survival in waterfowl depends mostly on weather conditions [@Blums2004].
+This is consistent with the literature because both nesting/hatching success and juvenile survival are conditional to the onset of lay, which is tightly related to local weather parameters, i.e.
+the spring temperature, the cumulative rainfall, and the water levels [@Folliot2017; @Blums2004; @Dzus1998].
 
 Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation to the outputs of the counting method demonstrates its ability to track the recruitment rate variability.
-This result is quite robust because the correlation is obvious with only five years available.
+This result is quite robust because the correlation is obvious with only a five-year time series.
 However, this conclusion might be overrated because two successive population size values are part of the recruitment rate, which mechanically scale the outputs.
-It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit less obvious than for the recruitment rate (Figure \@ref(fig:sup)A).
+It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit more noisy than for the recruitment rate (Figure \@ref(fig:sup)A).
 
@@
 ## Influence of harvest on the recruitment rate
+
+Variability within harvest category represents
 

### 2021-12-14 Code/Model_recruitment_rate.R

@@ -467,7 +467,7 @@ JuvCode <- nimbleCode(
       
     }
     
-    # r estimation frow counts
+    # r estimation from counts
     
     # priors
      max_surv ~ dunif(.7, .9)

### 2021-12-14 Code/Draft_ruddy_duck_recruitment_rate.Rmd
end of conflict

@@
 |$C_{i, t}$|Data|Cumulated number of individuals counted in population $i$ in year $t$|
-|$p(m \cap a)_{i, t}$|Parameter|Proportion of adult males in population $i$ in year $t$|
+|$p(a \cap m)_{i, t}$|Parameter|Proportion of adult males in population $i$ in year $t$|
 |$p(a)_{i, t}$|Parameter|Proportion adults in population $i$ in year $t$|
@@
 \begin{align}
-CAM_{i, t} & \sim {\sf Binom}(p(m \cap a)_{i, t}, C_{i, t}) \notag \\
+CAM_{i, t} & \sim {\sf Binom}(p(a \cap m)_{i, t}, C_{i, t}) \notag \\
 & \sim {\sf Binom}(p(m|a).p(a)_{i, t}, C_{i, t}) \notag \\

### 2021-12-15 Code/Draft_ruddy_duck_recruitment_rate.tex
readme edit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
]{article}
\title{Recruitment rate estimation in waterfowl using a non-invasive method}
\author{Adrien Tableau\footnote{Office Français de la Biodiversité, France - \href{mailto:adrien.tableau@ofb.gouv.fr}{\nolinkurl{adrien.tableau@ofb.gouv.fr}}} \and Iain Henderson\footnote{Animal and Plant Health Agency, United Kingdom - \href{mailto:iain.henderson@apha.gov.uk}{\nolinkurl{iain.henderson@apha.gov.uk}}} \and Sébastien Reeber\footnote{Société Nationale pour la Protection de la Nature, France - \href{mailto:sebastien.reeber@snpn.fr}{\nolinkurl{sebastien.reeber@snpn.fr}}}}
\date{15 décembre, 2021}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Recruitment rate estimation in waterfowl using a non-invasive method},
  pdfauthor={Adrien Tableau; Iain Henderson; Sébastien Reeber},
  pdflang={en-EN},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\ifXeTeX
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\begin{document}
\maketitle
\begin{abstract}
The response of a waterfowl population to a harvest pressure depends on its capacity to renew. The recruitment, i.e.~the number of young adults reproducing for the first time, is a key indicator to describe the renewal of a population, and therefore an essential tool in population management. The recruitment rate, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species. The proportion of young adult in a waterfowl population, which is a main step in estimating the recruitment rate, is often estimated from game-hunting samples. However, this proportion is not accessible in years without harvest, or with a low harvest rate. Moreover, the age-structure in the harvest samples does not necessarily reflect the underlying age-structure of the population. It is often skewed towards immatures and can lead to an overestimation of the recruitment rate. In waterfowl, adult males usually display brighter colours than females and immatures from both sexes. This dichromatism can be characterized and monitored from count surveys, a non-invasive method. Such information can be used to infer the proportion of immatures, and consequently the recruitment rate. In using two populations of ruddy duck, \emph{Oxyura oxyura}, this study introduces a model to estimate the recruitment rate from count data. To test the accuracy of the method, the results are compared to recruitment rate estimates from samples, and their consistency to support the observed population growth rates is also tested. The results suggest that the counting method is a versatile tool to estimate the recruitment rate if the species is monitored during the appropriate time window. Authors argue for considering a two-category count protocol in winter surveys of waterfowl species for which adult males can be differentiated from other individuals.

\textbf{Keywords}: fecundity - productivity - reproductive success - age ratio - juvenile - duck
\end{abstract}

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\emph{Definitions}

One defines the recruitment rate of a breeding population as the average number of recruits produced per each breeder (\protect\hyperlink{ref-Flint2015}{Flint, 2015}). This parameter might sometimes be called \emph{annual fertility} (\protect\hyperlink{ref-Koons2014}{Koons et al., 2014}), \emph{fecundity} (\protect\hyperlink{ref-Arnold2018}{Arnold, 2018}), \emph{reproductive rate} (\protect\hyperlink{ref-Cooch2014}{Cooch et al., 2014}), \emph{reproductive success} (\protect\hyperlink{ref-Etterson2011}{Etterson et al., 2011}), or \emph{productivity} (\protect\hyperlink{ref-Hagen2008}{Hagen \& Loughin, 2008}; \protect\hyperlink{ref-Johnson1987}{D. H. Johnson et al., 1987}). If one considers a species reaching its sexual maturity as early as its first year, all individuals are mature at the reproduction season. The recruitment rate is thus the ratio of the number of young adults at the reproduction season of year \(t\), i.e.~the recruits, to the number of all individuals at the reproduction season of the previous year, \(t - 1\). The declination for species with delayed maturity follows the same process (see \protect\hyperlink{ref-Robertson2008}{Robertson, 2008}).

The recruitment rate parameter is complex because it is composed of two main sub-parameters: the fecundity followed by the juvenile survival (\protect\hyperlink{ref-Etterson2011}{Etterson et al., 2011}). Both sub-parameters can be split into other sub-parameters. The fecundity can be defined by the product of the nesting success, the average clutch size, and the hatching success. The juvenile survival can be defined as the product of the post-hatching survival and the post-fledging survival, both critical life periods (\protect\hyperlink{ref-Blums2004}{Blums \& Clark, 2004}; \protect\hyperlink{ref-Hill2008}{Hill \& Ellis, 2008}). The recruitment rate as defined in the present study is thus a parameter integrating multiple life steps from the egg to the young adult attempting to breed for its first time (\protect\hyperlink{ref-Koons2017}{Koons et al., 2017}).

\emph{The use of recruitment rate in population management}

Population conservation, harvesting, and control require tools in population management in order to reach quantitative targets, such as a population size for example (\protect\hyperlink{ref-Johnson2021}{F. A. Johnson et al., 2021}; \protect\hyperlink{ref-Shea1998}{Shea \& NCEAS Working Group on Population Management, 1998}). To do so, a first blind strategy is to test iteratively management rules to reach progressively an optimum. Such approach can be time-consuming given the intrinsic variability of the population response to its changing environment. It might also suffer from a rejection of the stakeholders because of the lack of transparency of an arbitrary process (\protect\hyperlink{ref-Williams2013}{Williams \& Madsen, 2013}). Another approach is to try to predict the growth rate of the population as a function of the harvest rate and other environment inputs. Such modelling approach can provide clues on the recovering time of a population upon a variety of management rules (\protect\hyperlink{ref-Otis2006}{Otis, 2006}). These decision-support tools are then key inputs to status on the best management rules to implement. The growth rate of a population depends on the adult survival and the recruitment rate, see Equation \eqref{eq:GR} in Section \ref{PGR}. The adult survival is generally estimated by mark recapture/recovery methods (\protect\hyperlink{ref-Lebreton2001}{Lebreton, 2001}). If the harvest pressure behaves as an additive mortality to the natural mortality for adults (e.g. \protect\hyperlink{ref-Iverson2013}{Iverson et al., 2013}), it is quite straightforward to predict its impact on the adult survival.

According to \protect\hyperlink{ref-Blums2004}{Blums \& Clark} (\protect\hyperlink{ref-Blums2004}{2004}), the recruitment for diving ducks is poorly related to fecundity, but depends mostly on juvenile survival, which is driven by weather conditions. Studies on other birds concludes also on the weak correlation between fecundity and recruitment (e.g. \protect\hyperlink{ref-Murray2000}{Murray, 2000}). The difficulty to track juvenile survival (\protect\hyperlink{ref-Schmidt2008}{Schmidt et al., 2008}) demonstrates the interest to focus on the recruitment rate when accessible. The key to predict the growth rate of a population is then to estimate the value of the recruitment rate and its variability, potentially influenced by the harvest pressure (\protect\hyperlink{ref-Anders2005}{Anders \& Marshall, 2005}).

\emph{The variability of the recruitment rate}

A waterfowl species that is released in a favourable habitat will naturally expand (\protect\hyperlink{ref-Malthus1872}{Malthus, 1872}). The growth of such a population reflects that the recruitment rate parameter is higher than the adult mortality rate. This expansion is limited by the carrying capacity of the habitat (\protect\hyperlink{ref-Sayre2008}{Sayre, 2008}). When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity because the individuals compete for space or/and food (\protect\hyperlink{ref-Nummi2015}{Nummi et al., 2015}), which prevents the population from expanding. This competition commonly induces a lower fecundity or/and a higher juvenile mortality. The population renewal is slowed down because the recruitment rate parameter is affected by density-dependent effects (\protect\hyperlink{ref-Gunnarsson2013}{Gunnarsson et al., 2013}). For an endemic species evolving in stable environmental conditions, the recruitment rate compensates on average the adult mortality rate, which explains the stabilization of the number of individuals (\protect\hyperlink{ref-Flint2015}{Flint, 2015}).

Hunting waterfowl induces adult and juvenile mortality (\protect\hyperlink{ref-Bellrose1980}{Bellrose, 1980}). Intuitively, one expects that such a hunted population would observe lower adult survival and recruitment rate and then would rapidly deplete to its extinction. However, hunting waterfowl has a long history, and most of these species have persisted over time (\protect\hyperlink{ref-Cooch2014}{Cooch et al., 2014}). So paradoxically, hunting immature individuals does not always induce a decrease of recruitment rate. The heterogeneity of quality among immature individuals might explain this by inducing a compensation process if the harvest selects mostly individuals with a low reproductive potential (\protect\hyperlink{ref-Gimenez2017}{Gimenez et al., 2017}; \protect\hyperlink{ref-Lindberg2013}{Lindberg et al., 2013}). The recruitment rate can even increase because the harvest pressure reduces the densities and potentially the competition for space and food (\protect\hyperlink{ref-Nummi2015}{Nummi et al., 2015}; \protect\hyperlink{ref-Peron2012}{Péron et al., 2012}). In conclusion, a limited exploitation of a population observing heterogeneity and/or density-dependent process might paradoxically increase or at least not affect the recruitment rate. Under moderate exploitation, a new balance between recruitment rate and adult mortality can be theoretically reached (\protect\hyperlink{ref-Tsikliras2018}{Tsikliras \& Froese, 2018}). The size of a newly exploited population in stable environmental conditions should thus reach a different equilibrium, which is expected to be lower than the size without exploitation. This process explains why the exploitation of many waterfowl species is sustainable and has been lasting over a long time.

The fecundity and the juvenile survival, the two components of the recruitment rate, are very sensitive to weather conditions (\protect\hyperlink{ref-Blums2004}{Blums \& Clark, 2004}; \protect\hyperlink{ref-Folliot2017}{Folliot et al., 2017}). The environment conditions are thus an additional source of variability of the recruitment rate. The response of the recruitment rate to a harvest pressure is thus not straightforward. It is noisy because of its sensitivity to environmental factors. It also depends directly on the intensity of harvest on the immature individuals and indirectly on the mitigation of the density-dependent effects.

\emph{What part of the recruitment rate gradient is necessary for population management?}

In population management, a common question for a threaten or an invasive species is to determine the measure that will allow to reach a targeted population size/density in a limited time period (\protect\hyperlink{ref-Shea1998}{Shea \& NCEAS Working Group on Population Management, 1998}). For such question, the carrying capacity of the habitat is theoretically not reached. So a manager should mostly focus on collecting knowledge on the maximum recruitment rate and its variability by removing the potential density-dependent effects occurring only for the highest theoretical densities (\protect\hyperlink{ref-Eraud2021}{Eraud et al., 2021}; \protect\hyperlink{ref-Peron2013}{Péron, 2013}).

\emph{What is the ideal biological model to explore the gradient of the recruitment rate?}

Hunting waterfowl has a long history, but the biological effects of long-term moratorium have been poorly surveyed (\protect\hyperlink{ref-MartinezAbrain2013}{Martínez-Abraín et al., 2013}). So there are few opportunities to explore the response of the recruitment rate to a gradient of pressure for such species. Alien vertebrate species are commonly introduced during release events (\protect\hyperlink{ref-Saul2016}{Saul et al., 2016}). They usually observe a first period of colonisation on new territories without management measures. Since its presence might impact the balance of the colonized ecosystem, it happens that managers take strong control measures to restrict the growth of these populations or to eradicate them (\protect\hyperlink{ref-Oficialdegui2020}{Oficialdegui et al., 2020}). Such species are then a good biological model to overcome density-dependent effects and explore the effects of a harvest gradient on recruitment rate.

\emph{How to estimate recruitment rate?}

The recruitment rate parameter of a population can be assessed first by evaluating the fecundity of the breeding population and then by monitoring the survival from the egg life stage to the first-breeding event. To do so, it is necessary to monitor nests in habitats representative of the considered population (\protect\hyperlink{ref-Blums2004}{Blums \& Clark, 2004}). The survival of a chick to its recruitment can be done by capture mark recapture method quite tricky to set up (\protect\hyperlink{ref-Arnold2018}{Arnold, 2018}; \protect\hyperlink{ref-Blums2004}{Blums \& Clark, 2004}; \protect\hyperlink{ref-Schmidt2008}{Schmidt et al., 2008}). Even if such approach provides valuable information on the dynamics of the recruitment and can highlight the bottlenecks, the fieldwork required to estimate both components is too time-consuming to generalize such approach for waterfowl population management on a long-term horizon.

In waterfowl, it is generally impossible to differentiate immatures from adults without examining birds in hand. Aging individuals from samples coming from hunting bags is thus another opportunity to estimate the recruit proportion. Such approach suffers from two caveats. First, it is impossible to assess the recruitment rate when there is no harvest whereas it is necessary to predict what would happen if a moratorium seems required to allow a population to recover, or to calibrate a priori the harvest effort necessary to control an invasive species (\protect\hyperlink{ref-Smith2005}{G. C. Smith et al., 2005}). Secondly, like any predation action, hunting is selective and provides a biased picture of the population age structure in favour of the immatures (\protect\hyperlink{ref-Bellrose1980}{Bellrose, 1980}; \protect\hyperlink{ref-Fox2014}{A. D. Fox et al., 2014}). The hunting bag analysis might then lead to overestimate the recruitment rate and consequently the maximum sustainable harvest rate, with adverse consequences on population conservation (\protect\hyperlink{ref-Fox2014}{A. D. Fox et al., 2014}). Inferring recruitment rate from the proportion of immature birds trapped during tagging studies is also promising (\protect\hyperlink{ref-Arnold2018}{Arnold, 2018}), but the author admits that juveniles might be more naïve and are then more likely caught in traps (\protect\hyperlink{ref-RguibiIdrissi2003}{Rguibi-Idrissi et al., 2003}).

In waterfowl, it is common that adult males display brighter colours than adult females and immatures of both sexes, which display cover-up plumage similar to adult females (\protect\hyperlink{ref-Johnson1999}{K. P. Johnson, 1999}). This discrepancy is valuable because it can be observed and quantified from distance. Knowing adult sex ratio, such approach is promising because it requires only to count the proportion of birds of the two plumage kinds to infer the recruitment rate (\protect\hyperlink{ref-Smith2001}{C. M. Smith et al., 2001}). Unlike the two previous methods, this one is non-invasive and requires an acceptable time investment, which allows to collect long time series and consequently to track recruitment rate variability.

\emph{Methodology of this study}

From two populations of ruddy duck, a species introduced in Europe in the 40's, this study introduces a simple Bayesian model to infer recruitment rate from count data. A large part of this study focuses on assessing the viability of such approach. The comparison between two populations is a first step in testing the consistency of the method. A comparison to the common approach using samples from hunting bags aims at checking if the recruitment rate variability is well tracked and if the hypothesis stating that the recruitment rate is overestimated when using sample data is consistent with the results. Using two other approaches using population growth rates, one testes the consistency of the order of magnitude of the recruitment rate estimated from the counting method. Studying two populations is also an opportunity to discuss the response of the recruitment rate to a gradient of harvest pressure.

\hypertarget{MM}{%
\section{Materials \& methods}\label{MM}}

\hypertarget{biological-model}{%
\subsection{Biological model}\label{biological-model}}

The ruddy duck is a diving duck species introduced in the United Kingdom in the 40's (\protect\hyperlink{ref-GutierrezExposito2020}{Gutiérrez-Expósito et al., 2020}). The first reproduction was observed in the 60's. The population rapidly grew to reach more than 5000 individuals spread over the entire country. A new population set up in France from the 90's, likely because of an arrival of a few individuals from the United Kingdom (UK). Unlike the UK population, the French one set up around a single wintering spot, the Grand Lieu Lake (47.09°N, 1.67°W), which facilitates the population monitoring. Since this species is considered as a threat to the white-headed duck because of the risk of a genetic introgression (\protect\hyperlink{ref-MunozFuentes2007}{Muñoz-Fuentes et al., 2007}), an European plan of eradication has been adopted and control measures were taken in both countries (\protect\hyperlink{ref-GutierrezExposito2020}{Gutiérrez-Expósito et al., 2020}).

Just before the reproduction period, one can distinguish from shore individuals with male-like plumage from female-like individuals on wintering spots (Figure \ref{fig:flock}). It is noticeable that the apparent proportion of males in the counts from shore is far below the proportion of males in the samples from hunting bags (Figure \ref{fig:pmal}).

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{../Output/flock}

}

\caption{Typical observation of a ruddy duck flock in winter: 4 male-like individuals with a white cheek and a black cap, 10 female-like individuals with a whitish cheek and a dark stripe across it, 3 unidentified individuals © Jay McGowan - 3 February 2013 - Tompkins, New York, United States}\label{fig:flock}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{../Output/plot_1_saved}

}

\caption{Apparent male proportion in winter counts vs male proportion in samples; for the first category, one point corresponds to the proportion of male-like plumage in a population counted in winter; for the second category, the male proportion is estimated from all ruddy ducks sampled over one year}\label{fig:pmal}
\end{figure}

Assuming sex identification of samples is very accurate, such observation demonstrates that the counting method misses the identification of a significant part of males in the population. In many waterfowl species, immature individuals of both sexes display cover-up plumage similar to adult females (\protect\hyperlink{ref-Reeber2015}{Reeber, 2015}). Consequently, the counts in winter allows adult males to be distinguished from all other individuals (adult females, immature females, and immature males). This feature can be valuable to estimate the recruitment with two conditions. The season when adult males can be distinguished from all other individuals is sufficiently close to the breeding period to consider the age structure picture of the population to be representative of the age structure of the breeding population. The adult sex ratio is known and stable over the considered time period.

The demography of both ruddy duck populations has been tracked thanks to exhaustive counts on the wintering spots during the period between December 1 and January 31. Controlled individuals between the counting date and the beginning of the reproduction period (set up to May 1 upon field observations) were subtracted from the winter count to get a proxy of the breeding population size. Time series show that the two populations follow similar demographic histories (Figure \ref{fig:count}). They grew freely in both countries until 1999. Control measures have been then applied from 1999 to 2005 in the UK and from 1999 to 2018 in France. In both countries, these management measures led to stop the expansion of the ruddy ducks. From 2005 in the UK and from 2018 in France, higher control effort led to a rapid depletion in both populations. This history underlines that both populations observed a large spectrum of harvest effort, from no pressure to very high pressure. For sake of interpretation, the time period 2001-2004 was classified in the \emph{no harvest} category in France because the harvest pressure applied over this period was very limited with no effect on the population growth.

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{../Output/plot_2}

}

\caption{Evolution of the size of the two ruddy duck populations in light of the harvest pressure and the data time series}\label{fig:count}
\end{figure}

The counting dataset, which differentiates the male-like from the female-like individuals in winter, covers 7 years and 19 years in UK and France respectively (Figure \ref{fig:count}). Combined together, the counting datasets cover different population dynamics corresponding to a gradient of harvest pressure. In France, data before 2004 correspond to a period of population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure (Figure \ref{fig:count}). In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period. In parallel, the sampling dataset, which corresponds to individuals shot in winter during control operations, covers 9 years only in UK, with 5 years in common with the corresponding counting time series.

\hypertarget{recruitment-rate-inference-from-count-data}{%
\subsection{Recruitment rate inference from count data}\label{recruitment-rate-inference-from-count-data}}

Adult sex ratio in waterfowl is generally biased towards males (Figure \ref{fig:pmal}, \protect\hyperlink{ref-Wood2021}{Wood et al.} (\protect\hyperlink{ref-Wood2021}{2021})). As a consequence, deducing the immature proportion from the adult male proportion is not straightforward because the adult proportion twice as much as the male proportion. Even if the adult sex ratio may vary over a long time range, it is relatively stable over a few years (\protect\hyperlink{ref-Wood2021}{Wood et al., 2021}). Sex identification in adult samples from hunting bags was available for both ruddy duck populations. However, the small population size in France prevents from providing precise sex proportion estimates. A comparison over months in UK showed no difference in male proportion, so adult samples collected over the whole year were used to estimate the male proportion. A comparison of male proportion among years with more 500 samples did not show significant interannual differences. As a consequence, all adult samples were pooled to estimate the adult male proportion, see Equation \eqref{eq:pm}.

Assuming the additive property of the binomial distribution, the proportion of immature individuals can be deduced from the cumulated counts of male-like individuals in the wintering population, see Equation \eqref{eq:pic}. Only years with more than 100 individuals were selected. From this proportion and abundance index of breeding population sizes, the recruitment rate is straightforward, see Equation \eqref{eq:r}. If the absolute value of the breeding population size is accessible, the recruitment can be estimated, see Equation \eqref{eq:R}.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.11}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.13}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.76}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule
\endhead
\(SAM\) & Data & Total number of adult males in samples \\
\(SAF\) & Data & Total number of adult females in samples \\
\(p(m|a)\) & Parameter & Proportion of males in adults (or probability to be an adult male) \\
\(CAM_{i, t}\) & Data & Cumulated number of type-male individuals counted in population \(i\) in year \(t\), which are assumed to be adult males \\
\(C_{i, t}\) & Data & Cumulated number of individuals counted in population \(i\) in year \(t\) \\
\(p(a \cap m)_{i, t}\) & Parameter & Proportion of adult males in population \(i\) in year \(t\) \\
\(p(a)_{i, t}\) & Parameter & Proportion adults in population \(i\) in year \(t\) \\
\(p(i)_{i, t}\) & Parameter & Proportion of immatures in population \(i\) in year \(t\) \\
\(N_{i, t}\) & Data & Proxy of the size of the breeding population \(i\) in year \(t\) (maximum number of individuals counted in the wintering population minus removals before reproduction) \\
\(r_{i, t}\) & Parameter & Recruitment rate, i.e.~number of recruits in population \(i\) in year \(t\) per breeder in year \(t - 1\) \\
\(R_{i, t}\) & Parameter & Number of recruits in population \(i\) in year \(t\) \\
\bottomrule
\end{longtable}

\begin{equation}
p(m|a) \sim {\sf Beta}(SAM, SAF)
\label{eq:pm}
\end{equation}

\begin{align}
CAM_{i, t} & \sim {\sf Binom}(p(a \cap m)_{i, t}, C_{i, t}) \notag \\
& \sim {\sf Binom}(p(m|a).p(a)_{i, t}, C_{i, t}) \notag \\
& \sim {\sf Binom}(p(m|a).(1 - p(i)_{i, t}), C_{i, t})
\label{eq:pic}
\end{align}

\begin{equation}
r_{i, t} = \frac{p(i)_{i, t}.N_{i, t}}{N_{i, t - 1}}
\label{eq:r}
\end{equation}

\begin{equation}
R_{i, t} = p(i)_{i, t}.N_{i, t}
\label{eq:R}
\end{equation}

\hypertarget{comparison-with-the-sampling-method}{%
\subsection{Comparison with the sampling method}\label{comparison-with-the-sampling-method}}

A direct estimation of the immature proportion can be made using samples from hunting bags, see Equation \eqref{eq:pis}. By combining estimates of immature proportion and Equation \eqref{eq:r}, recruitment rates from samples were estimated to check the variability consistency of the counting method. One selected only years with more than 100 individuals controlled in winter, which correspond to a 9-year period in UK. This sampling time series covers 5 years of the UK counting time series.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.11}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.13}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.76}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule
\endhead
\(SI_{i, t}\) & Data & Number of immatures sampled in the wintering population \(i\) in year \(t\) \\
\(S_{i, t}\) & Data & Number of individuals sampled in the wintering population \(i\) in year \(t\) \\
\bottomrule
\end{longtable}

\begin{equation}
SI_{i, t} \sim {\sf Binom}(p(i)_{i, t}, S_{i, t})
\label{eq:pis}
\end{equation}

\hypertarget{PGR}{%
\subsection{Result validation using population growth rates}\label{PGR}}

The simple relationship (a bit more complex if a species with delayed maturity is considered, see \protect\hyperlink{ref-Robertson2008}{Robertson} (\protect\hyperlink{ref-Robertson2008}{2008})) between population growth rate, adult survival, and recruitment rate can be used to test the accuracy of the order of magnitude of recruitment rates estimated from the counting and the sampling methods, see Equation \eqref{eq:GR}. This relationship comes from a simple reasoning for a closed population: the population size in year \(t\) is equal to the number of breeders that survived over year \(t - 1\) plus the offspring produced in year \(t - 1\) that survived until the reproduction period of year \(t\), i.e.~the recruitment in year \(t\). The growth rate of a population is thus the sum of the adult survival rate and the recruitment rate (\protect\hyperlink{ref-Flint2015}{Flint, 2015}).

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.09}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.91}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule
\endhead
\(N_{t}\) & Number of breeders in year \(t\) \\
\(D_{t}\) & Number of breeders dead after reproduction in year \(t\) \\
\(R_{t}\) & Number of recruits in year \(t\) \\
\(s_{t}\) & Adult survival rate, i.e.~proportion of breeders in year \(t-1\) still alive in year \(t\) \\
\(r_{t}\) & Recruitment rate, i.e.~number of recruits in year \(t\) produced per breeder in year \(t-1\) \\
\(\lambda_{t}\) & Growth rate of the population between the reproduction periods of year \(t-1\) and year \(t\) \\
\bottomrule
\end{longtable}

\begin{align} 
N_{t} & = N_{t-1} - D_{t-1} + R_{t} \notag \\
N_{t} & = N_{t-1} - (1 - S_{t}).N_{t-1} + r_{t}.N_{t-1} \notag \\
N_{t} & = S_{t}.N_{t-1} + r_{t}.N_{t-1} \notag \\
\frac{N_{t}}{N_{t-1}} & = s_{t} + r_{t} \notag \\
\lambda_{t} & = s_{t} + r_{t} 
\label{eq:GR}
\end{align}

From the evolution of the breeding population size, population growth rate can be assessed every year to catch the variability range \eqref{eq:GR}, or can be smoothed over a consistent time period to provide average population growth rate, see Equation \eqref{eq:L}, which is a linear regression on the logarithm scale.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.08}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.78}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule
\endhead
\(N0_{i, J}\) & Parameter & Intercept of the regression model \\
\(\lambda_{i, J}\) & Parameter & Average population growth rate over a restricted time interval \(J\) for a population \(i\) (in \(year^{-1}\)) \\
\(t \in J\) & Index & Year index within the time interval \(J\) \\
\(\sigma_{i, J}\) & Parameter & Standard deviation of the regression model \\
\bottomrule
\end{longtable}

\begin{equation}
log(N_{i, t}) \sim {\sf Norm}(N0_{i, J} + log(\lambda_{i, J}).t, \sigma_{i, J})
\label{eq:L}
\end{equation}

\hypertarget{analysis-of-the-adult-survival}{%
\subsubsection{Analysis of the adult survival}\label{analysis-of-the-adult-survival}}

The adult survival is commonly assessed by capture mark recapture (\protect\hyperlink{ref-Lebreton2001}{Lebreton, 2001}). Combined to annual population growth rate, adult survival rate allows recruitment rate to be estimated each year, and then compared with the outputs of the counting method. Unfortunately, no tagging study was performed on the two ruddy duck populations. However, the relationship \eqref{eq:GR} could still produce annual adult survival estimates that can be discussed. To do so, annual population growth rates overlapping the counting time series were estimated.

Inconsistency on recruitment rate output, which is defined on \([0;\infty]\), would either lead to survival output \(<0\), which points out an overestimation of the recruitment rate, or \(>1\), which reflects that the recruitment rate cannot support the observed population growth and is thus underestimated. In between the two limit values, one still discuss the survival outputs by comparison with survival values published in the literature.

\hypertarget{comparison-to-a-proxy-of-maximum-recruitment-rate}{%
\subsubsection{Comparison to a proxy of maximum recruitment rate}\label{comparison-to-a-proxy-of-maximum-recruitment-rate}}

When population growth is not limited by a harvest pressure and/or by density-dependent effects, both recruitment rate and adult survival are close to their maximum potential. Given a conservative assumption on the maximum adult survival of ruddy ducks and maximum population growth rate estimated during no harvest time period, one can provide a proxy of the maximum recruitment rate. If the recruitment rate outputs of the model are significantly superior to this proxy, it means that the method overestimates the recruitment rate, which may reflect that some adult males are not detected during count surveys. By splitting the recruitment rate outputs upon population and harvest category, one can discuss of the impact of the harvest on the recruitment rate.

To do so, one estimates maximum population growth rates over time periods observing no harvest using Equation \eqref{eq:L}. Since the estimate of the UK population size is noisy below 500 individuals, one rejected data before 1972 (Figure \ref{fig:grts}). After reaching the threshold of 1000 individuals, the UK population growth observed a strong inflexion whereas no harvest pressure was applied. This observation led to split the analysis into two time series for the time periods used in the maximum growth rate inference in UK (Figure \ref{fig:grts}).

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{../Output/plot_3}

}

\caption{Evolution of the size of the two ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)}\label{fig:grts}
\end{figure}

To get proxies of maximum recruitment rate from Equation \eqref{eq:GR}, one uses a very conservative prior on the adult survival by setting a uniform distribution on \([0.7; 0.9]\) (\protect\hyperlink{ref-Buxton2004}{Buxton et al., 2004}; \protect\hyperlink{ref-Krementz1997}{Krementz et al., 1997}; the lower limit corresponds to the upper range of survival rates observed in literature for species of similar weight, the upper limit corresponds to the range of survival rates of long-life species, \protect\hyperlink{ref-Nichols1997}{Nichols et al., 1997}). The average recruitment rates from the count data are estimated for each population and for each harvest category and compared to the maximum recruitment rates.

For all the sub-models in Section \ref{MM}, the Bayesian framework was used for its efficiency and simplicity to propagate error through the parameters. One generated three chains of length 500000, with a thinning of 10 to avoid autocorrelation in the samples, and one discarded the first 2000 samples as burn-in. Chain convergence was assessed using the Gelman and Rubin convergence diagnostic (R\textless1.1, \protect\hyperlink{ref-Gelman1992}{Gelman \& Rubin} (\protect\hyperlink{ref-Gelman1992}{1992})). One fit the models using NIMBLE (\protect\hyperlink{ref-Valpine2017}{de Valpine et al., 2017}) run from R (\protect\hyperlink{ref-RCT2020}{R Core Team, 2020}). Data and code are available here: \url{https://github.com/adri-tab/Ruddy_duck_recruitment_rate}. The values \emph{X{[}Y; Z{]}} reported in results are the medians and the associated boundaries of the 95\% confidence interval of posterior distributions. The median was preferred to the mean because of its robustness to skewed distribution.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{recruitment-rate-variability}{%
\subsection{Recruitment rate variability}\label{recruitment-rate-variability}}

The adult male proportion, which is necessary to infer on the recruitment rate from count data, is estimated at 0.60 {[}0.59; 0.61{]}. The counting method provides estimations of the recruitment rate, see Figure \ref{fig:rrts}A. In UK, values vary from 0.06 {[}0.05; 0.07{]} to 0.27 {[}0.20; 0.34{]} recruits per breeder, and in France, 0.08 {[}0.04; 0.12{]} to 0.76 {[}0.64; 0.88{]} recruits per breeder. There are significant year-to-year variations with a maximum amplitude of 0.68 recruits per breeder in France (9.5-time variability). On the available period, recruitment rate in UK is smaller than in France. Recruitment rates are estimated in both populations over 5 years, and no correlation is observed between the two populations. The lowest recruitment rate observed in France, which is in 2019, corresponds to the start of the depletion of the population. It is noticeable that it is in the same range than the UK values, also estimated during a strong population depletion.

Recruitment rates estimated from UK sample data are consistently higher than estimates from the corresponding count data, Figure \ref{fig:rrts}A. The correlation between the two methods is strong and points out that the sampling method overestimates the recruitment rate by 44\% compared to the counting method, Figure \ref{fig:rrts}B. Estimation precision is similar for the two methods. The UK population started to deplete from 2006, but no significant shift is observed between the two periods.

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{../Output/plot_4}

}

\caption{(A) Estimates of recruitment rate, i.e.~average number of recruits produced per breeder; (B) Comparison of the recruitment rate values between the two estimation methods; bars define the 95\% confidence intervals}\label{fig:rrts}
\end{figure}

\hypertarget{consistency-of-the-corresponding-adult-survival}{%
\subsection{Consistency of the corresponding adult survival}\label{consistency-of-the-corresponding-adult-survival}}

Adult survival rates were deduced from the recruitment rate estimates and the observed population growth rates, see Figure \ref{fig:srts}A. In UK, values that are obtained from the counting method ranged from 0.21 {[}0.20; 0.22{]} to 0.68 {[}0.65; 0.72{]}. In France, survival rates vary from 0.39 {[}0.34; 0.44{]} to 1.08 {[}0.98; 1.17{]}. Like recruitment rates, there is significant year-to-year variability without clear temporal structure in both populations with a maximum amplitude of 0.69 in France (2.8-time variability). Unlike the recruitment rate, the survival rate in 2019 in France is not a low outlier.

As expected, survival rates estimated from UK sample data are consistently lower than estimates from the corresponding count data, Figure \ref{fig:srts}A. The correlation between the two methods points out that the sampling method underestimates the survival rate by 32\% compared to the counting method, Figure \ref{fig:srts}B. Estimation precision is similar for the two methods. The sample time series shows that the adult survival is significantly lower after 2006, which corresponds to the start of the depletion of the UK population.

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{../Output/plot_5}

}

\caption{(A) Estimates of adult survival rate using population growth rates; (B) Comparison of the survival rate values between the two estimation methods; bars define the 95\% confidence intervals}\label{fig:srts}
\end{figure}

\hypertarget{comparison-to-a-proxy-of-maximum-recruitment-rate-1}{%
\subsection{Comparison to a proxy of maximum recruitment rate}\label{comparison-to-a-proxy-of-maximum-recruitment-rate-1}}

Maximum population growth rates are similar for the two populations, 1.45 {[}1.36; 1.55{]} and 1.47 {[}1.29; 1.68{]} for UK and France respectively, which corresponds to an increase of about 45\% per year (Figure \ref{fig:mgr}). After reaching a certain size (\textasciitilde{} 1000 individuals), the UK population growth rate falls to 1.06 {[}1.04; 1.07{]}, which corresponds to an increase of 6\% per year, even though no harvest pressure was applied on it.

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{../Output/plot_6}

}

\caption{(A) Estimates of the maximum population growth without harvest \((\lambda - 1)\); two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates; bars define the 95\% confidence intervals}\label{fig:mgr}
\end{figure}

Upon the conservative assumption of a maximum adult survival uniformly distributed over 0.70 to 0.90, the maximum recruitment rate reaches 0.65 {[}0.51; 0.75{]} and 0.66 {[}0.45; 0.90{]} recruits per breeder for UK and France respectively (Figure \ref{fig:grcomp}). When there is no harvest, the estimated average recruitment rate of the French population is in the range of the proxy of the maximum recruitment rate, 0.51 {[}0.46; 0.57{]} recruits per breeder. When there is harvest, the average recruitment rates are 0.14 {[}0.12; 0.16{]} and 0.33 {[}0.31; 0.36{]} recruits per breeder for the UK and France respectively. Both values are significantly lower than the maximum recruitment rate, and it is noticeable that the average recruitment rate is lower for the UK population, which is in depletion over the count time series, than for the French population, which of the dynamic is mostly stable over the monitored period.

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{../Output/plot_7}

}

\caption{Comparison between proxies of maximum recruitment rate and average recruitment rates estimated with and without harvest (\emph{counting} method only); proxies are estimated by using the maximum growth rates; bars define the 95\% confidence intervals}\label{fig:grcomp}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Count data in waterfowl that differentiate adult males from other individuals are successfully used in this study to infer estimations of recruitment rates. In the following section, one checks the accuracy of the method from indicators developed for this purpose. From the outputs on the two populations of ruddy duck, one discusses about the effects of the harvest on the recruitment rate and its implication on waterfowl management.

\hypertarget{accuracy-of-the-counting-method}{%
\subsection{Accuracy of the counting method}\label{accuracy-of-the-counting-method}}

The recruitment rates from the counting method vary in the same range for the two populations, which demonstrate a certain robustness of the method. The higher variability observed in France is not unexpected because the French time series is longer and covers a larger harvest spectrum than in UK. There is no temporal correlation between the two populations, so it suggests that the recruitment rate fluctuations are more likely influenced by regional-scale factors than large-scale ones. This is consistent with the literature because both nesting/hatching success and juvenile survival are conditional to the onset of lay, which is tightly related to local weather parameters, e.g.~the spring temperature, the cumulative rainfall, and the water levels (\protect\hyperlink{ref-Blums2004}{Blums \& Clark, 2004}; \protect\hyperlink{ref-Dzus1998}{Dzus \& Clark, 1998}; \protect\hyperlink{ref-Folliot2017}{Folliot et al., 2017}).

Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation to the outputs of the counting method demonstrates the ability of this method to track the recruitment rate variability. This result is quite robust because the correlation is obvious with only a five-year time series. However, this conclusion might be overrated because two successive population size values are part of the recruitment rate, which mechanically scale the outputs. It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit more noisy than for the recruitment rate (Figure \ref{fig:sup}A). The uncertainty of the estimates is very close to the outputs from the sampling method. Even though the counting method is more indirect to infer on the recruitment rate, it performs to provide a good confidence of the outputs which is satisfying.

A significant change in adult sex ratio would bias the estimation of recruitment rate. However, since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong. Significant variations can then occur only every multiple years, especially if it is a long-living species. Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.

The counting method always provides recruitment rates significantly lower than the sampling method by factor close to 2 on average, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable to hunters (\protect\hyperlink{ref-Bellrose1980}{Bellrose, 1980}; \protect\hyperlink{ref-Fox2014}{A. D. Fox et al., 2014}). This is a robust result because the approaches used to control ruddy ducks differ strongly from usual hunting practices and are apriori less selective. It is especially clear in the UK, for which control operations occur in winter using long-range rifles that prevents the bird from escaping from the threat. The assumption on the accuracy of the counting method is then not excluded from the consistency of this result.

Annual adult survival rates deduced from observed population growth rates and recruitment estimates are on average in line with the literature \textbf{ref required}. which confirms that the counting method is poorly biased if there is some bias.

survival in the range of the expected values -\textgreater{} so no apparent issue of scaling issue,
so we the hypothesis that male-like individuals corresponds only to adult male is validated.

\emph{average recruitment rate compared to a proxy of maximum recruitment rate}

always below the maximum recruitment rate so satisfying.

overall, even if we don't have access to the true value of the recruitment rate to test the accuracy of the counting method, a beam of arguments support that the method is not heavily biased, unlike the sampling method, as demonstrated in \protect\hyperlink{ref-Fox2014}{A. D. Fox et al.} (\protect\hyperlink{ref-Fox2014}{2014}), even if one considers a picture limited to the end of the hunting season (\protect\hyperlink{ref-Fox2016}{A. Fox et al., 2016}).

\hypertarget{influence-of-harvest-on-the-recruitment-rate}{%
\subsection{Influence of harvest on the recruitment rate}\label{influence-of-harvest-on-the-recruitment-rate}}

Variability within harvest category represents

\hypertarget{implication-in-waterfowl-management}{%
\subsection{Implication in waterfowl management}\label{implication-in-waterfowl-management}}

Survival higher than one but warning, observation error, so annual population growth rate might be inaccurate on some years. Also, we assumed a closed population, but it is possible that years with adult survival higher than one corresponds to an arrival of birds from the other population. As a consequence, the recruitment rate on these years may also be overestimated.

No impact on the adult survival in France, but it seems that there is an impact in UK.
Why, because harvest occurs mostly in summer in France, and in winter in UK.

The harvest pressure, simplified into a two-category variable, explains already 19.3\% and 4.6\% of the variance of the recruitment rate and the adult survival rate of the French population. The harvest pressure described as a continuous variable is expected to explain even more the recruitment rate variability. The harvest pressure applied in France is thus a factor that likely induces a strong immature mortality, but a moderate one on the adults.

Impact on the recruitment rate in France, and no impact in UK.

Overall, the harvest seems to impact differently France and UK, likely because the harvest mortality occurs mostly in summer in France, and in winter in UK.

But caution, a population model linking the harvest rate to state variables describing the population is necessary to check precisely the impact of the harvest pressure, and test hypothesis on the functioning of harvest mortality.

Recruitment rate decrease with the harvest for the French population, and the UK population recruitment rate is also below the French one.
it seems quite clear that the harvest affects the recruitment rate.

\emph{reference to introduction}

It is quite consistent with the assumption in introduction. Since studied populations are in colonization of their habitat, maximum densities are not reached, and then no or limited density dependent effect occur.

If counting method acceptable, moderate harvest pressure does not reduce the recruitment rate as expected, maybe due to density dependent mortality for juvenile but it affects it if strong harvest rate. That is very interesting bro but warning, these interpretation holds on weak estimations and adult survival equal to 1!!!

it may be possible that first sight of density dependence occur in UK from 1000 individuals, but was not big enough to fully curb the population growth. Since there is a strong correlation between the number of birds observed and the number of sites occupied by the birds, this inflexion may be simply due to a limited number of new favourable sites to be colonized.

What about heterogeneity, figure 8 -\textgreater{} One said that recruitment rate can increase if harvest because of heterogeneity and density dependence, but it is not what we observe in France, likely because the exploitation is high inhere and consequently impact.

\hypertarget{use-of-recruitment-rate}{%
\subsection{Use of recruitment rate}\label{use-of-recruitment-rate}}

\emph{better than sex ratio}
a same value of age ratio can correspond to different fecundity levels, so warning.
a same value of age ratio for a population in depletion and a population stabilized are not similary productive, the first one is less productive than the second one !
so warning, focus on recruitment rate and not on age ratio if the dynamics of the population varies! recruitment proportion is commonly used in to discuss the fecundity of a population, but unlike the recruitment rate, this indicator does not consider the dynamics of the population and does reflect fecundity only on stable growth period

the consistency between the two max growth rate strenghtens that seems to be a biological limit for this species, at least in the western Palearctic.

As described before, recruitment rate over the studied time period is far lower for the UK than France (2.5 factor).
Underlined adult survival are comprised between 0 and 1, whatever the method. So we cannot conclude if one can be rejected.

France affected the recruitment, UK affected adult survival for a similar result -\textgreater{} depletion

whereas adult survival more consistent between the tow (.5 factor), meaning there was more variability in recruitment rate than in survival rate. growth rate difference is thus more explained by recruitment rate fluctuation than adult survival fluctuation, consistent with the demographic buffering, or canalization hypothesis (\protect\hyperlink{ref-Gaillard2003}{Gaillard \& Yoccoz, 2003}; \protect\hyperlink{ref-Lenzi2021}{Lenzi et al., 2021})

survival values compared to literature? s

Abundance index are considered exact, which should not be true. and lead to structural error in the estimates of annual growth rate.

\protect\hyperlink{ref-Arnold2018}{Arnold} (\protect\hyperlink{ref-Arnold2018}{2018}) underlines the lack of data on recruitment rate for modellers (in discussion) and that new methods are required to track this variable that often drive the population growth rate

Recruitment rate drives lambda? see \protect\hyperlink{ref-Arnold2018}{Arnold} (\protect\hyperlink{ref-Arnold2018}{2018})

(only X years), meaning no large scale factor drive the fluctuations of productivity.

adult sex ratio in line with other duck species (\protect\hyperlink{ref-Wood2021}{Wood et al., 2021})

This time lag might be not a issue if the survival of immature is close to the adult survival between the two periods. winter count and breeding season

According to \protect\hyperlink{ref-Blums2004}{Blums \& Clark} (\protect\hyperlink{ref-Blums2004}{2004}), the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, which is driven by weather conditions. Studies on other birds also conclude of the poor correlation between fecundity and recruitment (e.g. \protect\hyperlink{ref-Murray2000}{Murray, 2000}).

No correlation between the two populations on recruitment rate, so maybe no large scale drivers of recruitment rate variability.

age proportion -\textgreater{} {[}0, 1{]}
recruitment rate {[}0, +inf{]} donc plus sensitive

recruitment rate the most sensitive parameter to density dependent effect (\protect\hyperlink{ref-Koons2014}{Koons et al., 2014})

Hypothesis of time invariant sex ratios in immature and adult: limits

For hunted duck species in North America, the mortality rate on first-year immatures compared to adults is higher by a factor ranging from 1.4 to 2 (Bellrose 1980).

recruitment rate:
Stabilité grosse pop,
variabilité petite pop.

Au dela du fait que la croissance de la pop est la somme de la survie et de la recruitment rate,
intéressant de parler de la variability de la survie vs variabilité de la recruitment rate
pour parler de risk management.

recruitment rate: drop from 2009

one considers harvest as an impact only adult survival, not on age structure, but we should, oversimplistics

blabla

\protect\hyperlink{ref-Nichols2006}{Nichols \& Williams} (\protect\hyperlink{ref-Nichols2006}{2006}) -\textgreater{} il faut compter de la façon dont on demande

Taper des jeunes c'est pas forcément très utile car density dependence

If the maximum recruitment rate is not reached under optimal conditions (maximum population size not reached, and limited harvest pressure), it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the recruitment rate.

Condition of application to other species.

Message de l'article

\hypertarget{acknowledgments}{%
\section*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{section}{Acknowledgments}

This work was partly funded by the LIFE Oxyura project (LIFE17 NAT/FR/000942) through the LIFE program. This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the \emph{Office Français de la Biodiversité}, and Jean-Marc Gillier from the \emph{Société Nationale pour la Protection de la Nature}. The authors acknowledge all the contributors of the data collection in the UK and in France, especially: \textbf{UK names}, Vincent Fontaine, Denis Lacourpaille, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Anders2005}{}}%
Anders, A. D., \& Marshall, M. R. (2005). Increasing the accuracy of productivity and survival estimates in assessing landbird population status. \emph{Conservation Biology}, \emph{19}(1), 66--74.

\leavevmode\vadjust pre{\hypertarget{ref-Arnold2018}{}}%
Arnold, T. W. (2018). Using ring-recovery and within-season recapture data to estimate fecundity and population growth. \emph{Ecology and Evolution}, \emph{8}(20), 10298--10305. \url{https://doi.org/10.1002/ece3.4506}

\leavevmode\vadjust pre{\hypertarget{ref-Bellrose1980}{}}%
Bellrose, F. C. (1980). \emph{Ducks, geese \& swans of {N}orth {A}merica: 3rd. ed}. Stackpole Books.

\leavevmode\vadjust pre{\hypertarget{ref-Blums2004}{}}%
Blums, P., \& Clark, R. G. (2004). Correlates of lifetime reproductive success in three species of {E}uropean ducks. \emph{Oecologia}, \emph{140}(1), 61--67. \url{https://doi.org/10.1007/s00442-004-1573-8}

\leavevmode\vadjust pre{\hypertarget{ref-Buxton2004}{}}%
Buxton, N. E., Summers, R. W., \& Peach, W. J. (2004). Survival rate of female \emph{{B}oldeneye bucephala} clangula. \emph{Bird Study}, \emph{51}(3), 280--281. \url{https://doi.org/10.1080/00063650409461365}

\leavevmode\vadjust pre{\hypertarget{ref-Cooch2014}{}}%
Cooch, E. G., Guillemain, M., Boomer, G. S., Lebreton, J.-D., \& Nichols, J. D. (2014). The effects of harvest on waterfowl populations. \emph{Wildfowl}, 220--276.

\leavevmode\vadjust pre{\hypertarget{ref-Valpine2017}{}}%
de Valpine, P., Turek, D., Paciorek, C. J., Anderson-Bergman, C., Lang, D. T., \& Bodik, R. (2017). Programming with models: Writing statistical algorithms for general model structures with {NIMBLE}. \emph{Journal of Computational and Graphical Statistics}, \emph{26}(2), 403--413. \url{https://doi.org/10.1080/10618600.2016.1172487}

\leavevmode\vadjust pre{\hypertarget{ref-Dzus1998}{}}%
Dzus, E. H., \& Clark, R. G. (1998). Brood survival and recruitment of mallards in relation to wetland density and hatching date. \emph{The Auk}, \emph{115}(2), 311--318. \url{https://doi.org/10.2307/4089189}

\leavevmode\vadjust pre{\hypertarget{ref-Eraud2021}{}}%
Eraud, C., Devaux, T., Villers, A., Johnson, F. A., \& Francesiaz, C. (2021). Popharvest: An {R} package to assess the sustainability of harvesting regimes of bird populations. \emph{Ecology and Evolution}. \url{https://doi.org/10.1002/ece3.8212}

\leavevmode\vadjust pre{\hypertarget{ref-Etterson2011}{}}%
Etterson, M. A., Ellis-Felege, S. N., Evers, D., Gauthier, G., Grzybowski, J. A., Mattsson, B. J., Nagy, L. R., Olsen, B. J., Pease, C. M., Burg, M. P. van der, \& Potvien, A. (2011). Modeling fecundity in birds: Conceptual overview, current models, and considerations for future developments. \emph{Ecological Modelling}, \emph{222}(14), 2178--2190. \url{https://doi.org/10.1016/j.ecolmodel.2010.10.013}

\leavevmode\vadjust pre{\hypertarget{ref-Flint2015}{}}%
Flint, P. (2015). Population dynamics of sea ducks: Using models to understand the causes, consequences, evolution, and management of variation in life history characteristics. In J.-P. L. Savard, D. V. Derksen, D. Esler, \& J. M. Eadie (Eds.), \emph{Ecology and conservation of {N}orth {A}merican sea ducks} (Vol. 46, pp. 63--96). CRC Press.

\leavevmode\vadjust pre{\hypertarget{ref-Folliot2017}{}}%
Folliot, B., Caizergues, A., Barbotin, A., \& Guillemain, M. (2017). Environmental and individual correlates of common pochard (\emph{{A}ythya ferina}) nesting success. \emph{European Journal of Wildlife Research}, \emph{63}(4). \url{https://doi.org/10.1007/s10344-017-1126-1}

\leavevmode\vadjust pre{\hypertarget{ref-Fox2014}{}}%
Fox, A. D., Clausen, K. K., Dalby, L., Christensen, T. K., \& Sunde, P. (2014). Age-ratio bias among hunter-based surveys of {E}urasian wigeon \emph{{A}nas penelope} based on wing vs. Field samples. \emph{Ibis}, \emph{157}(2), 391--395. \url{https://doi.org/10.1111/ibi.12229}

\leavevmode\vadjust pre{\hypertarget{ref-Fox2016}{}}%
Fox, A., Clausen, K., Dalby, L., Christensen, T., \& Sunde, P. (2016). Between-year variations in sex/age ratio bias in hunter wings of eurasian wigeon (\emph{{A}nas penelope}) compared to field samples. \emph{Ornis Fennica}, \emph{93}(1), 26.

\leavevmode\vadjust pre{\hypertarget{ref-Gaillard2003}{}}%
Gaillard, J.-M., \& Yoccoz, N. G. (2003). {TEMPORAL} {VARIATION} {IN} {SURVIVAL} {OF} {MAMMALS}: A {CASE} {OF} {ENVIRONMENTAL} {CANALIZATION}? \emph{Ecology}, \emph{84}(12), 3294--3306. \url{https://doi.org/10.1890/02-0409}

\leavevmode\vadjust pre{\hypertarget{ref-Gelman1992}{}}%
Gelman, A., \& Rubin, D. B. (1992). Inference from iterative simulation using multiple sequences. \emph{Statistical Science}, \emph{7}(4). \url{https://doi.org/10.1214/ss/1177011136}

\leavevmode\vadjust pre{\hypertarget{ref-Gimenez2017}{}}%
Gimenez, O., Cam, E., \& Gaillard, J.-M. (2017). Individual heterogeneity and capture-recapture models: What, why and how? \emph{Oikos}, \emph{127}(5), 664--686. \url{https://doi.org/10.1111/oik.04532}

\leavevmode\vadjust pre{\hypertarget{ref-Gunnarsson2013}{}}%
Gunnarsson, G., Elmberg, J., Pöysä, H., Nummi, P., Sjöberg, K., Dessborn, L., \& Arzel, C. (2013). Density dependence in ducks: A review of the evidence. \emph{European Journal of Wildlife Research}, \emph{59}(3), 305--321. \url{https://doi.org/10.1007/s10344-013-0716-9}

\leavevmode\vadjust pre{\hypertarget{ref-GutierrezExposito2020}{}}%
Gutiérrez-Expósito, C., Pernollet, C., Adriaens, T., \& Henderson, I. (2020). Ruddy duck (\emph{{O}xyura jamaicensis} {G}melin, 1789). In \emph{Invasive birds: Global trends and impacts} (pp. 200--205). {CABI}.

\leavevmode\vadjust pre{\hypertarget{ref-Hagen2008}{}}%
Hagen, C. A., \& Loughin, T. M. (2008). Productivity estimates from upland bird harvests: Estimating variance and necessary sample sizes. \emph{Journal of Wildlife Management}, \emph{72}(6), 1369--1375. \url{https://doi.org/10.2193/2007-390}

\leavevmode\vadjust pre{\hypertarget{ref-Hill2008}{}}%
Hill, D. A., \& Ellis, N. (2008). Survival and age related changes in the foraging behaviour and time budget of tufted ducklings aythya fuligula. \emph{Ibis}, \emph{126}(4), 544--550. \url{https://doi.org/10.1111/j.1474-919x.1984.tb02079.x}

\leavevmode\vadjust pre{\hypertarget{ref-Iverson2013}{}}%
Iverson, S. A., Reed, E. T., Hughes, R. J., \& Forbes, M. R. (2013). Age and breeding stage-related variation in the survival and harvest of temperate-breeding {C}anada geese in {O}ntario. \emph{The Journal of Wildlife Management}, \emph{78}(1), 24--34. \url{https://doi.org/10.1002/jwmg.636}

\leavevmode\vadjust pre{\hypertarget{ref-Johnson1987}{}}%
Johnson, D. H., Sparling, D. W., \& Cowardin, L. M. (1987). A model of the productivity of the mallard duck. \emph{Ecological Modelling}, \emph{38}(3-4), 257--275. \url{https://doi.org/10.1016/0304-3800(87)90100-1}

\leavevmode\vadjust pre{\hypertarget{ref-Johnson2021}{}}%
Johnson, F. A., Heldbjerg, H., Nagy, S., \& Madsen, J. (2021). Setting population-size targets for geese causing socio-economic conflicts. \emph{Ambio}, \emph{51}(1), 209--225. \url{https://doi.org/10.1007/s13280-021-01539-5}

\leavevmode\vadjust pre{\hypertarget{ref-Johnson1999}{}}%
Johnson, K. P. (1999). The evolution of bill coloration and plumage dimorphism supports the transference hypothesis in dabbling ducks. \emph{Behavioral Ecology}, \emph{10}(1), 63--67. \url{https://doi.org/10.1093/beheco/10.1.63}

\leavevmode\vadjust pre{\hypertarget{ref-Koons2017}{}}%
Koons, D. N., Arnold, T. W., \& Schaub, M. (2017). Understanding the demographic drivers of realized population growth rates. \emph{Ecological Applications}, \emph{27}(7), 2102--2115. \url{https://doi.org/10.1002/eap.1594}

\leavevmode\vadjust pre{\hypertarget{ref-Koons2014}{}}%
Koons, D. N., Gunnarsson, G., Schmutz, J. A., \& Rotella, J. J. (2014). Drivers of waterfowl population dynamics : From teal to swans. \emph{Wildfowl}, \emph{Special Issue 4}, 169--191. \url{http://wildfowl.wwt.org.uk/index.php/wildfowl/article/view/2606}

\leavevmode\vadjust pre{\hypertarget{ref-Krementz1997}{}}%
Krementz, D. G., Barker, R. J., \& Nichols, J. D. (1997). Sources of variation in waterfowl survival rates. \emph{The Auk}, \emph{114}(1), 93--102. \url{https://doi.org/10.2307/4089068}

\leavevmode\vadjust pre{\hypertarget{ref-Lebreton2001}{}}%
Lebreton, J.-D. (2001). The use of bird rings in the study of survival. \emph{Ardea}, \emph{89}(1), 85--100.

\leavevmode\vadjust pre{\hypertarget{ref-Lenzi2021}{}}%
Lenzi, O., Ozgul, A., Salguero-Gómez, R., \& Paniw, M. (2021). Beyond demographic buffering: Context dependence in demographic strategies across animals. \emph{BioRxiv}. \url{https://doi.org/10.1101/2021.06.08.447594}

\leavevmode\vadjust pre{\hypertarget{ref-Lindberg2013}{}}%
Lindberg, M. S., Sedinger, J. S., \& Lebreton, J.-D. (2013). Individual heterogeneity in black brant survival and recruitment with implications for harvest dynamics. \emph{Ecology and Evolution}, \emph{3}(12), 4045--4056. \url{https://doi.org/10.1002/ece3.767}

\leavevmode\vadjust pre{\hypertarget{ref-Malthus1872}{}}%
Malthus, T. R. (1872). \emph{An essay on the principle of population}.

\leavevmode\vadjust pre{\hypertarget{ref-MartinezAbrain2013}{}}%
Martínez-Abraín, A., Viedma, C., Gómez, J. A., Bartolomé, M. A., Jiménez, J., Genovart, M., \& Tenan, S. (2013). Assessing the effectiveness of a hunting moratorium on target and non-target species. \emph{Biological Conservation}, \emph{165}, 171--178. \url{https://doi.org/10.1016/j.biocon.2013.06.009}

\leavevmode\vadjust pre{\hypertarget{ref-MunozFuentes2007}{}}%
Muñoz-Fuentes, V., Vilà, C., Green, A. J., Negro, J. J., \& Sorenson, M. D. (2007). Hybridization between white-headed ducks and introduced ruddy ducks in {S}pain. \emph{Molecular Ecology}, \emph{16}(3), 629--638. \url{https://doi.org/10.1111/j.1365-294x.2006.03170.x}

\leavevmode\vadjust pre{\hypertarget{ref-Murray2000}{}}%
Murray, B. G. (2000). Measuring annual reproductive success in birds. \emph{The Condor}, \emph{102}(2), 470--473. \url{https://doi.org/10.1093/condor/102.2.470}

\leavevmode\vadjust pre{\hypertarget{ref-Nichols1997}{}}%
Nichols, J. D., Hines, J. E., \& Blums, P. (1997). {TESTS} {FOR} {SENESCENT} {DECLINE} {IN} {ANNUAL} {SURVIVAL} {PROBABILITIES} {OF} {COMMON} {POCHARDS}, {AYTHYA} {FERINA}. \emph{Ecology}, \emph{78}(4), 1009--1018. \url{https://doi.org/10.1890/0012-9658(1997)078\%5B1009:tfsdia\%5D2.0.co;2}

\leavevmode\vadjust pre{\hypertarget{ref-Nichols2006}{}}%
Nichols, J. D., \& Williams, B. K. (2006). Monitoring for conservation. \emph{Trends in Ecology \& Evolution}, \emph{21}(12), 668--673. \url{https://doi.org/10.1016/j.tree.2006.08.007}

\leavevmode\vadjust pre{\hypertarget{ref-Nummi2015}{}}%
Nummi, P., Holopainen, S., Rintala, J., \& Pöysä, H. (2015). Mechanisms of density dependence in ducks: Importance of space and per capita food. \emph{Oecologia}, \emph{177}(3), 679--688. \url{https://doi.org/10.1007/s00442-014-3133-1}

\leavevmode\vadjust pre{\hypertarget{ref-Oficialdegui2020}{}}%
Oficialdegui, F. J., Delibes-Mateos, M., Green, A. J., Sánchez, M. I., Boyero, L., \& Clavero, M. (2020). Rigid laws and invasive species management. \emph{Conservation Biology}. \url{https://doi.org/10.1111/cobi.13481}

\leavevmode\vadjust pre{\hypertarget{ref-Otis2006}{}}%
Otis, D. L. (2006). Mourning dove hunting regulation strategy based on annual harvest statistics and banding data. \emph{The Journal of Wildlife Management}, \emph{70}(5), 1302--1307. \url{https://doi.org/10.2193/0022-541x(2006)70\%5B1302:mdhrsb\%5D2.0.co;2}

\leavevmode\vadjust pre{\hypertarget{ref-Peron2013}{}}%
Péron, G. (2013). Compensation and additivity of anthropogenic mortality: Life-history effects and review of methods. \emph{Journal of Animal Ecology}, \emph{82}(2), 408--417. \url{https://doi.org/10.1111/1365-2656.12014}

\leavevmode\vadjust pre{\hypertarget{ref-Peron2012}{}}%
Péron, G., Nicolai, C. A., \& Koons, D. N. (2012). Demographic response to perturbations: The role of compensatory density dependence in a north american duck under variable harvest regulations and changing habitat. \emph{Journal of Animal Ecology}, \emph{81}(5), 960--969. \url{https://doi.org/10.1111/j.1365-2656.2012.01980.x}

\leavevmode\vadjust pre{\hypertarget{ref-RCT2020}{}}%
R Core Team,. (2020). \emph{{R}: A language and environment for statistical computing}. R Foundation for Statistical Computing. \url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-Reeber2015}{}}%
Reeber, S. (2015). \emph{Wildfowl of {E}urope, {A}sia and {N}orth {A}merica} (C. Helm, Ed.).

\leavevmode\vadjust pre{\hypertarget{ref-RguibiIdrissi2003}{}}%
Rguibi-Idrissi, H., Julliard, R., \& Bairlein, F. (2003). Variation in the stopover duration of reed warblers acrocephalus scirpaceus in {M}orocco: Effects of season, age and site. \emph{Ibis}, \emph{145}(4), 650--656. \url{https://doi.org/10.1046/j.1474-919x.2003.00208.x}

\leavevmode\vadjust pre{\hypertarget{ref-Robertson2008}{}}%
Robertson, G. J. (2008). Using winter juvenile/adult ratios as indices of recruitment in population models. \emph{Waterbirds}, \emph{31}(sp2), 152--158. \url{https://doi.org/10.1675/1524-4695-31.sp2.152}

\leavevmode\vadjust pre{\hypertarget{ref-Saul2016}{}}%
Saul, W.-C., Roy, H. E., Booy, O., Carnevali, L., Chen, H.-J., Genovesi, P., Harrower, C. A., Hulme, P. E., Pagad, S., Pergl, J., \& Jeschke, J. M. (2016). Assessing patterns in introduction pathways of alien species by linking major invasion data bases. \emph{Journal of Applied Ecology}, \emph{54}(2), 657--669. \url{https://doi.org/10.1111/1365-2664.12819}

\leavevmode\vadjust pre{\hypertarget{ref-Sayre2008}{}}%
Sayre, N. F. (2008). The genesis, history, and limits of carrying capacity. \emph{Annals of the Association of American Geographers}, \emph{98}(1), 120--134. \url{https://doi.org/10.1080/00045600701734356}

\leavevmode\vadjust pre{\hypertarget{ref-Schmidt2008}{}}%
Schmidt, K. A., Rush, S. A., \& Ostfeld, R. S. (2008). Wood thrush nest success and post-fledging survival across a temporal pulse of small mammal abundance in an oak forest. \emph{Journal of Animal Ecology}, \emph{77}(4), 830--837. \url{https://doi.org/10.1111/j.1365-2656.2008.01378.x}

\leavevmode\vadjust pre{\hypertarget{ref-Shea1998}{}}%
Shea, K., \& NCEAS Working Group on Population Management, the. (1998). Management of populations in conservation, harvesting and control. \emph{Trends in Ecology \& Evolution}, \emph{13}(9), 371--375. \url{https://doi.org/10.1016/s0169-5347(98)01381-0}

\leavevmode\vadjust pre{\hypertarget{ref-Smith2001}{}}%
Smith, C. M., Goudie, R. I., \& Cooke, F. (2001). Winter age ratios and the assessment of recruitment of harlequin ducks. \emph{Waterbirds: The International Journal of Waterbird Biology}, \emph{24}(1), 39. \url{https://doi.org/10.2307/1522241}

\leavevmode\vadjust pre{\hypertarget{ref-Smith2005}{}}%
Smith, G. C., Henderson, I. S., \& Robertson, P. A. (2005). A model of ruddy duck \emph{{O}xyura jamaicensis} eradication for the {UK}. \emph{Journal of Applied Ecology}, \emph{42}(3), 546--555. \url{https://doi.org/10.1111/j.1365-2664.2005.00977.x}

\leavevmode\vadjust pre{\hypertarget{ref-Tsikliras2018}{}}%
Tsikliras, A. C., \& Froese, R. (2018). Maximum sustainable yield. In \emph{Encyclopedia of ecology (second edition)} (Vol. 1, pp. 108--115). Elsevier. \url{https://doi.org/10.1016/b978-0-12-409548-9.10601-3}

\leavevmode\vadjust pre{\hypertarget{ref-Williams2013}{}}%
Williams, J. H., \& Madsen, J. (2013). Stakeholder perspectives and values when setting waterbird population targets: Implications for flyway management planning in a european context. \emph{{PLOS} {ONE}}, \emph{8}(11), e81836. \url{https://doi.org/10.1371/journal.pone.0081836}

\leavevmode\vadjust pre{\hypertarget{ref-Wood2021}{}}%
Wood, K. A., Brides, K., Durham, M. E., \& Hearn, R. D. (2021). Adults have more male-biased sex ratios than first-winter juveniles in wintering duck populations. \emph{Avian Research}, \emph{12}(1). \url{https://doi.org/10.1186/s40657-021-00286-1}

\end{CSLReferences}

\newpage

\hypertarget{appendix-supplement}{%
\appendix}

\hypertarget{supplement}{%
\section{Supplement}\label{supplement}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{../Output/plot_8}

}

\caption{(A1) Estimates of the proportion of the immature individuals in the wintering population; (B1) estimates of the number of recruits, i.e.~the young adults reproducing for their first time; (A2) \& (B2) comparison of the values between the two estimation methods; bars define the 95\% confidence intervals}\label{fig:sup}
\end{figure}

\end{document}

### 2021-12-15 Code/Model_recruitment_rate.R

@@ -377,7 +377,7 @@ ds %>%
              aes(x = year, y = n_pop, color = sub_pop)) +
   scale_x_date_own(1e-2)
 
-# Data formatting function ----------------------------------------------------------
+# Data formatting functions ---------------------------------------------------------
 
 dis <- function(fil = NA, gp = NA, col = NA, size = FALSE) {
   fil = enquo(fil)
@@ -683,6 +683,32 @@ lambda_ds %>%
          lambda = exp(r_avg)) %>% 
   ungroup() -> lambda_ds_2
 
+JuvOut4 %>% 
+  filter(par == "productivity", method == "counting") %>% 
+  mutate(gp = if_else(year(year) %in% c(1999, 2002:2004) & pop == "FR", 
+                      "no_harvest", "harvest")) %>% 
+  group_by(pop, gp) %>% 
+  mutate(mean = mean(`50%`)) %>% 
+  ungroup() %>% 
+  mutate(res = `50%` - mean) %>% 
+  group_by(pop) %>% 
+  mutate(var_tot = var(`50%`),
+         var_res = var(res), 
+         prop_var = (var_tot - var_res) / var_tot)
+
+JuvOut4 %>% 
+  filter(par == "survival", method == "counting") %>% 
+  mutate(gp = if_else(year(year) %in% c(1999, 2002:2004) & pop == "FR", 
+                      "no_harvest", "harvest")) %>% 
+  group_by(pop, gp) %>% 
+  mutate(mean = mean(`50%`)) %>% 
+  ungroup() %>% 
+  mutate(res = `50%` - mean) %>% 
+  group_by(pop) %>% 
+  mutate(var_tot = var(`50%`),
+         var_res = var(res), 
+         prop_var = (var_tot - var_res) / var_tot)
+
 # Output plot -----------------------------------------------------------------------
 
 raw_plot(p_juv, title = "Proportion of recruits in the population", 
@@ -799,7 +825,7 @@ JuvOut4 %>%
   scale_y_continuous(
     limits = c(0, 1),
     breaks = 0:10 / 10) +
-  labs(x = NULL, y = "recruitment rate") -> prod_max_plot; prod_max_plot
+  labs(x = NULL, y = "average recruitment rate") -> prod_max_plot; prod_max_plot
 
 # Plots for article ------------------------------------------------------------------
 
@@ -1001,4 +1027,3 @@ list(plo, c(1:length(plo)), plo_dim) %>%
                  ..1, 
[…]

### 2021-12-16 Code/Draft_ruddy_duck_recruitment_rate.Rmd
readme edit

@@
 
-The simple relationship (a bit more complex if a species with delayed maturity is considered [see @Robertson2008]) between population growth rate, adult survival, and recruitment rate can be used to test the accuracy of the order of magnitude of recruitment rates estimated from the counting and the sampling methods, see Equation \@ref(eq:GR).
+The simple relationship (a bit more complex if a species with delayed maturity is considered, see @Robertson2008) between population growth rate, adult survival, and recruitment rate can be used to test the accuracy of the order of magnitude of recruitment rates estimated from the counting and the sampling methods, see Equation \@ref(eq:GR).
 This relationship comes from a simple reasoning for a closed population: the population size in year $t$ is equal to the number of breeders that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
@@
 
-From the evolution of the breeding population size, population growth rate can be assessed every year to catch its variability range \@ref(eq:GR), or can be smoothed over a consistent time period to provide average population growth rate, see Equation \@ref(eq:L), which is a linear regression on the logarithm scale.
+From the evolution of the breeding population size, population growth rate can be assessed every year to catch the variability range \@ref(eq:GR), or can be smoothed over a consistent time period to provide average population growth rate, see Equation \@ref(eq:L), which is a linear regression on the logarithm scale.
 
@@
 
-The recruitment rates from the counting method are in the same range for the two populations.
+The recruitment rates from the counting method are in the same range for the two populations, which demonstrate a certain consistency in the results obtained from the method.
 The higher variability observed in France is not unexpected because the French time series is longer and covers a larger harvest spectrum than in UK.
-There is no temporal correlation between the two populations, so it suggests that the recruitment rate is more likely influenced by regional-scale factors than large-scale ones.
-This is consistent with the literature because both nesting/hatching success and juvenile survival are conditional to the onset of lay, which is tightly related to local weather parameters, i.e.
+There is no temporal correlation between the two populations, so it suggests that the recruitment rate fluctuations are more likely influenced by regional-scale factors than large-scale ones.
+This is consistent with the literature because both nesting/hatching success and juvenile survival are conditional to the onset of lay, which is tightly related to local weather parameters, e.g.
 the spring temperature, the cumulative rainfall, and the water levels [@Folliot2017; @Blums2004; @Dzus1998].
 
-Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation to the outputs of the counting method demonstrates its ability to track the recruitment rate variability.
+Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation to the outputs of the counting method demonstrates the ability of this method to track the recruitment rate variability.
 This result is quite robust because the correlation is obvious with only a five-year time series.
@@
 It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit more noisy than for the recruitment rate (Figure \@ref(fig:sup)A).
-
-A significant change in adult sex ratio would bias the estimation of recruitment rate.
+The uncertainty of the estimates is very close to the outputs from the sampling method.
+Even though the counting method is more indirect to infer on the recruitment rate, it performs to provide a good confidence of the outputs which is satisfying.
+A significant change in adult sex ratio would bias the evolution of the recruitment rate.
 However, since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong.
@@
 Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.
-
-The counting method always provides recruitment rates significantly lower than the sampling method, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl.
+The counting method always provides recruitment rates significantly lower than the sampling method by factor close to 2 on average, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable to hunters [@Bellrose1980; @Fox2014].
 This is a robust result because the approaches used to control ruddy ducks differ strongly from usual hunting practices and are apriori less selective.
@@
 
-*survival consistency*
+Annual adult survival rates deduced from observed population growth rates and recruitment estimates are on average in line with the literature **ref required**.
+which confirms that the counting method is poorly biased if there is some bias.
 
-Annual adult survival rates deduced from observed population growth rates and recruitment estimates are on average in line with the literature **ref required**.
-
-survival in the range of the expected values -> so no apparent issue of scaling issue, 
-so we the hypothesis that male-like individuals corresponds only to adult male is validated.
+survival in the range of the expected values -> so no apparent issue of scaling issue, so we the hypothesis that male-like individuals corresponds only to adult male is validated.
 
@@
 Why, because harvest occurs mostly in summer in France, and in winter in UK.
+
+The harvest pressure, simplified into a two-category variable, explains already 19.3% and 4.6% of the variance of the recruitment rate and the adult survival rate of the French population.
+The harvest pressure described as a continuous variable is expected to explain even more the recruitment rate variability.
+The harvest pressure applied in France is thus a factor that likely induces a strong immature mortality, but a moderate one on the adults.
 

### 2021-12-16 Code/Draft_ruddy_duck_recruitment_rate.Rmd
discussion update

@@
 
+@Arnold2018 underlines the lack of data on recruitment rate for modellers and that new methods are required to track this variable that often drive the population growth rate
+
+Recruitment rate drives lambda? see @Arnold2018
+
 ## Accuracy of the counting method
@@
 the spring temperature, the cumulative rainfall, and the water levels [@Folliot2017; @Blums2004; @Dzus1998].
+
+According to @Blums2004, the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, which is driven by weather conditions.
+Studies on other birds also conclude of the poor correlation between fecundity and recruitment [e.g.
+@Murray2000].
 
@@
 Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.
-The counting method always provides recruitment rates significantly lower than the sampling method by factor close to 2 on average, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable to hunters [@Bellrose1980; @Fox2014].
-This is a robust result because the approaches used to control ruddy ducks differ strongly from usual hunting practices and are apriori less selective.
+adult sex ratio in line with other duck species [@Wood2021]
+
+The counting method always provides recruitment rates significantly lower than the sampling method by factor close to 2 on average, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable than adults to hunters [@Bellrose1980; @Fox2014].
+This is a robust result because the approach used to control ruddy ducks differ strongly from usual hunting practices and are apriori less selective.
 It is especially clear in the UK, for which control operations occur in winter using long-range rifles that prevents the bird from escaping from the threat.
@@
 
-Annual adult survival rates deduced from observed population growth rates and recruitment estimates are on average in line with the literature **ref required**.
-which confirms that the counting method is poorly biased if there is some bias.
+Annual adult survival rates deduced from observed population growth rates and recruitment rate estimates are on average in line with the literature [@Nichols1997; @Krementz1997; @Buxton2004], which reflects that there is no apparent scaling issue.
+The hypothesis that male-like individuals corresponds only to adult males is then acceptable.
+This result supports that the counting method is poorly biased if there is some bias.
+However, some adult survival rate values are outside the range of expected values even if they never significantly exceed the maximum possible value of one.
+This may be due to an observation error of the growth rate that is not possible to assess in the population size estimations, or more likely to the corruption of the closed population hypothesis.
+Indeed, if there is an arrival from another ruddy duck population, the true recruitment cannot explain by itself the apparent population growth and indirectly leads to an overestimation of the adult survival.
 
-survival in the range of the expected values -> so no apparent issue of scaling issue, so we the hypothesis that male-like individuals corresponds only to adult male is validated.
+Proxies of recruitment rate are very close for the two populations, which demonstrates that population parameters can be assessed even with small size populations. 
+The consistency between the two max growth rate strenghtens that seems to be a biological limit for this species, at least in the western Palearctic.
 
-*average recruitment rate compared to a proxy of maximum recruitment rate*
+The average recruitment rate estimated from the counting method, with or without harvest, are for both populations below the proxies of maximum recruitment rate.
+The assumption on the accuracy of the counting method is then not excluded from the consistency of this result.
 
-always below the maximum recruitment rate so satisfying.
+Overall, even if the true value of the recruitment rate is accessible to test the accuracy of the counting method, a beam of arguments support that the method is not heavily biased.
+This consists in a big difference with the sampling method as demonstrated in @Fox2014, even if one considers a picture limited to the end of the hunting season to limit the bias [@Fox2016].
 
-overall, even if we don't have access to the true value of the recruitment rate to test the accuracy of the counting method, a beam of arguments support that the method is not heavily biased, unlike the sampling method, as demonstrated in @Fox2014, even if one considers a picture limited to the end of the hunting season [@Fox2016].
+Survival higher than one but warning, observation error, so annual population growth rate might be inaccurate on some years.
+Also, we assumed a closed population, but it is possible that years with adult survival higher than one corresponds to an arrival of birds from the other population.
+As a consequence, the recruitment rate on these years may also be overestimated.
+Abundance index are considered exact, which should not be true.
+and lead to structural error in the estimates of annual growth rate.
+
+This time lag might be not a issue if the survival of immature is close to the adult survival between the two periods. winter count and breeding season
 
@@
 
-## Implication in waterfowl management
-
-Survival higher than one but warning, observation error, so annual population growth rate might be inaccurate on some years.
-Also, we assumed a closed population, but it is possible that years with adult survival higher than one corresponds to an arrival of birds from the other population.
-As a consequence, the recruitment rate on these years may also be overestimated.
-
+The harvest pressure, simplified into a two-category variable, explains already 19.3% and 4.6% of the variance of the recruitment rate and the adult survival rate of the French population.
+The harvest pressure described as a continuous variable is expected to explain even more the recruitment rate variability.
+The harvest pressure applied in France is thus a factor that likely induces a strong immature mortality, but a moderate one on the adults.
 No impact on the adult survival in France, but it seems that there is an impact in UK.
@@
 
-The harvest pressure, simplified into a two-category variable, explains already 19.3% and 4.6% of the variance of the recruitment rate and the adult survival rate of the French population.
-The harvest pressure described as a continuous variable is expected to explain even more the recruitment rate variability.
-The harvest pressure applied in France is thus a factor that likely induces a strong immature mortality, but a moderate one on the adults.
-
 Impact on the recruitment rate in France, and no impact in UK.
+Recruitment rate decrease with the harvest for the French population, and the UK population recruitment rate is also below the French one. 
+it seems quite clear that the harvest affects the recruitment rate.
 
@@
 
-But caution, a population model linking the harvest rate to state variables describing the population is necessary to check precisely the impact of the harvest pressure, and test hypothesis on the functioning of harvest mortality.
-
-Recruitment rate decrease with the harvest for the French population, and the UK population recruitment rate is also below the French one. 
-it seems quite clear that the harvest affects the recruitment rate.
-
-*reference to introduction*
-
-It is quite consistent with the assumption in introduction.
-Since studied populations are in colonization of their habitat, maximum densities are not reached, and then no or limited density dependent effect occur.
-
-If counting method acceptable, moderate harvest pressure does not reduce the recruitment rate as expected, maybe due to density dependent mortality for juvenile but it affects it if strong harvest rate.
-That is very interesting bro but warning, these interpretation holds on weak estimations and adult survival equal to 1!!!
-
-it may be possible that first sight of density dependence occur in UK from 1000 individuals, but was not big enough to fully curb the population growth.
-Since there is a strong correlation between the number of birds observed and the number of sites occupied by the birds, this inflexion may be simply due to a limited number of new favourable sites to be colonized.
-
-What about heterogeneity, figure 8 -> One said that recruitment rate can increase if harvest because of heterogeneity and density dependence, but it is not what we observe in France, likely because the exploitation is high inhere and consequently impact.
-
-## Use of recruitment rate
-
-*better than sex ratio*
-a same value of age ratio can correspond to different fecundity levels, so warning. 
-a same value of age ratio for a population in depletion and a population stabilized are not similary productive, the first one is less productive than the second one ! 
-so warning, focus on recruitment rate and not on age ratio if the dynamics of the population varies!
-recruitment proportion is commonly used in to discuss the fecundity of a population, but unlike the recruitment rate, this indicator does not consider the dynamics of the population and does reflect fecundity only on stable growth period
-
-the consistency between the two max growth rate strenghtens that seems to be a biological limit for this species, at least in the western Palearctic.
-
-As described before, recruitment rate over the studied time period is far lower for the UK than France (2.5 factor). 
-Underlined adult survival are comprised between 0 and 1, whatever the method. So we cannot conclude if one can be rejected.
-
 France affected the recruitment, UK affected adult survival for a similar result -> depletion
 
-whereas adult survival more consistent between the tow (.5 factor), meaning there was more variability in recruitment rate than in survival rate.
-growth rate difference is thus more explained by recruitment rate fluctuation than adult survival fluctuation, consistent with the demographic buffering, or canalization hypothesis [@Gaillard2003; @Lenzi2021]
+For hunted duck species in North America, the mortality rate on first-year immatures compared to adults is higher by a factor ranging from 1.4 to 2 (Bellrose 1980).
 
-survival values compared to literature? s
-
-Abundance index are considered exact, which should not be true. and lead to structural error in the estimates of annual growth rate.
-
-@Arnold2018 underlines the lack of data on recruitment rate for modellers (in discussion) and that new methods are required to track this variable that often drive the population growth rate
-
-Recruitment rate drives lambda? see @Arnold2018
-
- (only X years), meaning no large scale factor drive the fluctuations of productivity.
-
-adult sex ratio in line with other duck species [@Wood2021]
-
-This time lag might be not a issue if the survival of immature is close to the adult survival between the two periods. winter count and breeding season
-
-According to @Blums2004, the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, which is driven by weather conditions.
-Studies on other birds also conclude of the poor correlation between fecundity and recruitment [e.g.
-@Murray2000].
-
-No correlation between the two populations on recruitment rate, so maybe no large scale drivers of recruitment rate variability.
-
-age proportion -> [0, 1]
-recruitment rate [0, +inf] donc plus sensitive
-
-recruitment rate the most sensitive parameter to density dependent effect [@Koons2014]
-
-Hypothesis of time invariant sex ratios in immature and adult: limits
-
-For hunted duck species in North America, the mortality rate on first-year immatures compared to adults is higher by a factor ranging from 1.4 to 2 (Bellrose 1980).
+so more variability should be reflected in a difference of variability between the 2 parameters, but not true, how to make our result consistent with these observations?
+Huge harvest pressure, and adult are targetted, so compensation on adult survival.
 
@@
 
-Au dela du fait que la croissance de la pop est la somme de la survie et de la recruitment rate, 
-intéressant de parler de la variability de la survie vs variabilité de la recruitment rate 
-pour parler de risk management.
+## Implication in waterfowl management
+
+recruitment rate the most sensitive parameter to density dependent effect [@Koons2014]
+
+Population growth is equally influenced by adult survival and recruitment
+whereas adult survival more consistent between the tow (.5 factor), meaning there was more variability in recruitment rate than in survival rate.
+growth rate difference is thus more explained by recruitment rate fluctuation than adult survival fluctuation, consistent with the demographic buffering, or canalization hypothesis [@Gaillard2003; @Lenzi2021]
+
+Au dela du fait que la croissance de la pop est la somme de la survie et de la recruitment rate, intéressant de parler de la variability de la survie vs variabilité de la recruitment rate pour parler de risk management.
+
+But caution, a population model linking the harvest rate to state variables describing the population is necessary to check precisely the impact of the harvest pressure, and test hypothesis on the functioning of harvest mortality.
+
+It is quite consistent with the assumption in introduction.
+Since studied populations are in colonization of their habitat, maximum densities are not reached, and then no or limited density dependent effect occur.
+
+it may be possible that first sight of density dependence occur in UK from 1000 individuals, but was not big enough to fully curb the population growth.
+Since there is a strong correlation between the number of birds observed and the number of sites occupied by the birds, this inflexion may be simply due to a limited number of new favourable sites to be colonized.
+
+If counting method acceptable, moderate harvest pressure does not reduce the recruitment rate as expected, maybe due to density dependent mortality for juvenile but it affects it if strong harvest rate.
+That is very interesting bro but warning, these interpretation holds on weak estimations and adult survival equal to 1!!!
+
+What about heterogeneity, figure 8 -> One said that recruitment rate can increase if harvest because of heterogeneity and density dependence, but it is not what we observe in France, likely because the exploitation is high inhere and consequently impact.
+
+recruitment rate better than sex ratio: a same value of age ratio can correspond to different fecundity levels, so warning.
+a same value of age ratio for a population in depletion and a population stabilized are not similary productive, the first one is less productive than the second one !
+so warning, focus on recruitment rate and not on age ratio if the dynamics of the population varies!
+recruitment proportion is commonly used in to discuss the fecundity of a population, but unlike the recruitment rate, this indicator does not consider the dynamics of the population and does reflect fecundity only on stable growth period
+
+Taper des jeunes c'est pas forcément très utile car density dependence
 
@@
 
-blabla
-
-@Nichols2006 -> il faut compter de la façon dont on demande
-
-Taper des jeunes c'est pas forcément très utile car density dependence
-
 If the maximum recruitment rate is not reached under optimal conditions (maximum population size not reached, and limited harvest pressure), it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the recruitment rate.
@@
 
+# Conclusion
+
 Message de l'article
+@Nichols2006 -> il faut compter de la façon dont on demande
 

### 2021-12-17 Code/Draft_ruddy_duck_recruitment_rate.Rmd
Discussion edit

@@
 lang: en-EN
-bibliography: C:/Users/adrien.tableau/Dropbox/JabRef/Library_OFB.bib
+bibliography: "`r paste0('C:/Users/', Sys.info()[['user']], '/Dropbox/JabRef/Library_OFB.bib')`"
 csl: https://www.zotero.org/styles/apa
@@
       toc: false
+      keep_md: true
+      keep_tex: true
     bookdown::word_document2:
@@
 
-One defines the recruitment rate of a breeding population as the average number of recruits produced per each breeder [@Flint2015].
-This parameter might sometimes be called *annual fertility* [@Koons2014], *fecundity* [@Arnold2018], *reproductive rate* [@Cooch2014], *reproductive success* [@Etterson2011], or *productivity* [@Johnson1987; @Hagen2008].
+One defines the *recruitment rate* of a breeding population as the average number of recruits produced per each breeder [@Flint2015].
+This parameter might sometimes be called *annual fertility* [@Koons2014], *reproductive rate* [@Cooch2014], *reproductive success* [@Etterson2011], or *productivity* [@Johnson1987; @Hagen2008].
 If one considers a species reaching its sexual maturity as early as its first year, all individuals are mature at the reproduction season.
@@
 
-The recruitment rate parameter is complex because it is composed of two main sub-parameters: the fecundity followed by the juvenile survival [@Etterson2011].
+The recruitment rate parameter is complex because it is composed of two main sub-parameters: the fecundity followed by the juvenile survival [@Etterson2011; @Arnold2018].
 Both sub-parameters can be split into other sub-parameters.
@@
 The juvenile survival can be defined as the product of the post-hatching survival and the post-fledging survival, both critical life periods [@Hill2008; @Blums2004].
-The recruitment rate as defined in the present study is thus a parameter integrating multiple life steps from the egg to the young adult attempting to breed for its first time [@Koons2017].
+The recruitment rate as defined in the present study is thus a parameter integrating multiple life steps from the egg production to the young adult attempting to breed for its first time [@Koons2017].
 
@@
 For all the sub-models in Section \@ref(MM), the Bayesian framework was used for its efficiency and simplicity to propagate error through the parameters.
+One used only uninformative priors on all parameters.
 One generated three chains of length 500000, with a thinning of 10 to avoid autocorrelation in the samples, and one discarded the first 2000 samples as burn-in.
@@
 Data and code are available here: https://github.com/adri-tab/Ruddy_duck_recruitment_rate.
-The values *X[Y; Z]* reported in results are the medians and the associated boundaries of the 95% confidence interval of posterior distributions.
+The values **X[Y; Z]** reported in results are the medians and the associated boundaries of the 95% confidence interval of posterior distributions.
 The median was preferred to the mean because of its robustness to skewed distribution.
@@
 Count data in waterfowl that differentiate adult males from other individuals are successfully used in this study to infer estimations of recruitment rates.
-In the following section, one checks the accuracy of the method from indicators developed for this purpose.
-From the outputs on the two populations of ruddy duck, one discusses about the effects of the harvest on the recruitment rate and its implication on waterfowl management.
-
-@Arnold2018 underlines the lack of data on recruitment rate for modellers and that new methods are required to track this variable that often drive the population growth rate
-
-Recruitment rate drives lambda? see @Arnold2018
+Since this parameter usually explains mostly the population growth in waterfowl [e.g.
+@Arnold2018], the non-invasive method introduced in this study might be considered as a candidate tool to monitor at low cost waterfowl populations.
+In the following section, one checks the accuracy of the method from indicators that were developed for this purpose.
+From the outputs on the two populations of ruddy duck, one discusses about the effects of the harvest on the recruitment rate and its implications on waterfowl management.
 
@@
 
-According to @Blums2004, the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, which is driven by weather conditions.
-Studies on other birds also conclude of the poor correlation between fecundity and recruitment [e.g.
-@Murray2000].
-
 Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation to the outputs of the counting method demonstrates the ability of this method to track the recruitment rate variability.
 This result is quite robust because the correlation is obvious with only a five-year time series.
-However, this conclusion might be overrated because two successive population size values are part of the recruitment rate, which mechanically scale the outputs.
+The assumption of a constant adult sex ratio over time is then acceptable.
+A significant change in adult sex ratio would indeed bias the evolution of the recruitment rate.
+However, since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong.
+Significant variations can then occur only every multiple years, especially if the species is long-living.
+Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base, as done in this study (the average adult sex ratio found here is in line with other duck species [@Wood2021]).
+However, the robustness of the correlation between the two methods might be overrated because two successive population size values are part of the recruitment rate, which mechanically scale the outputs.
 It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit more noisy than for the recruitment rate (Figure \@ref(fig:sup)A).
-The uncertainty of the estimates is very close to the outputs from the sampling method.
-Even though the counting method is more indirect to infer on the recruitment rate, it performs to provide a good confidence of the outputs which is satisfying.
-A significant change in adult sex ratio would bias the evolution of the recruitment rate.
-However, since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong.
-Significant variations can then occur only every multiple years, especially if it is a long-living species.
-Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.
-adult sex ratio in line with other duck species [@Wood2021]
+The uncertainty around the estimates is similar for the two methods even though the counting method is more indirect to infer on the recruitment rate, which is satisfying.
 
@@
 
-Annual adult survival rates deduced from observed population growth rates and recruitment rate estimates are on average in line with the literature [@Nichols1997; @Krementz1997; @Buxton2004], which reflects that there is no apparent scaling issue.
-The hypothesis that male-like individuals corresponds only to adult males is then acceptable.
-This result supports that the counting method is poorly biased if there is some bias.
+Annual adult survival rates deduced from observed population growth rates and recruitment rate estimates are on average in line with the literature on species of similar weight [@Nichols1997; @Krementz1997; @Buxton2004], which reflects that there is no apparent scaling issue.
+The assumption that male-like individuals corresponds only to adult males is then acceptable.
+This result supports that the counting method is poorly biased if it is.
 However, some adult survival rate values are outside the range of expected values even if they never significantly exceed the maximum possible value of one.
-This may be due to an observation error of the growth rate that is not possible to assess in the population size estimations, or more likely to the corruption of the closed population hypothesis.
+A likely reason to these extreme values is a corruption of the closed population assumption on some years.
 Indeed, if there is an arrival from another ruddy duck population, the true recruitment cannot explain by itself the apparent population growth and indirectly leads to an overestimation of the adult survival.
 
-Proxies of recruitment rate are very close for the two populations, which demonstrates that population parameters can be assessed even with small size populations. 
-The consistency between the two max growth rate strenghtens that seems to be a biological limit for this species, at least in the western Palearctic.
+Proxies of maximum recruitment rate are very close for the two populations.
+This strengthens that the ruddy duck species reached its intrinsic biological reproduction limit, at least given the environmental conditions of the western Palearctic.
+The consistency of this intermediate result suggests that the population size estimation is quite good for both populations and demonstrates that their evolution is well tracked even with a few individuals.
+The average recruitment rate estimated from the counting method, with or without harvest, are not different of significantly below the proxies of maximum recruitment rate for both populations.
+The assumption on the accuracy of the counting method is then again not excluded from the consistency of this result.
 
-The average recruitment rate estimated from the counting method, with or without harvest, are for both populations below the proxies of maximum recruitment rate.
-The assumption on the accuracy of the counting method is then not excluded from the consistency of this result.
-
-Overall, even if the true value of the recruitment rate is accessible to test the accuracy of the counting method, a beam of arguments support that the method is not heavily biased.
-This consists in a big difference with the sampling method as demonstrated in @Fox2014, even if one considers a picture limited to the end of the hunting season to limit the bias [@Fox2016].
-
-Survival higher than one but warning, observation error, so annual population growth rate might be inaccurate on some years.
-Also, we assumed a closed population, but it is possible that years with adult survival higher than one corresponds to an arrival of birds from the other population.
-As a consequence, the recruitment rate on these years may also be overestimated.
-Abundance index are considered exact, which should not be true.
-and lead to structural error in the estimates of annual growth rate.
-
-This time lag might be not a issue if the survival of immature is close to the adult survival between the two periods. winter count and breeding season
+Overall, even if the true values of recruitment rates are not directly accessible to test the accuracy of the counting method, a beam of arguments support that the method is not heavily biased.
+This consists in a big difference with the sampling method as demonstrated in @Fox2014, even when one considers a picture limited to the end of the hunting season to limit the bias [@Fox2016].
 
@@
 ## Implication in waterfowl management
+
+Recruitment rate drives lambda [@Arnold2018]
+
+According to @Blums2004, the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, the other component of the recruitment that is moslty driven by the weather conditions.
+Studies on other birds also conclude of the poor correlation between fecundity and recruitment [e.g.
+@Murray2000].
 

### 2021-12-17 Code/Model_recruitment_rate.R

@@ -954,7 +954,7 @@ grid.arrange(
 
 grid.arrange(
   grobs = list(raw_surv + 
-                 labs(title = "A", y = "survival rate") + 
+                 labs(title = "A", y = "adult survival rate") + 
                  theme(legend.position = "bottom"),
                cor_surv + labs(title = "B")),
   width = c(5, 3),

### 2022-01-04 Code/Draft_ruddy_duck_recruitment_rate.Rmd
discussion edits

@@
 In using two populations of ruddy duck, *Oxyura oxyura*, this study introduces a model to estimate the recruitment rate from count data.
-To test the accuracy of the method, the results are compared to recruitment rate estimates from samples, and their consistency to support the observed population growth rates is also tested.
+To test the accuracy of the method, the results are compared to estimates of recruitment rate using samples.
+One also tests the conistency of the recruitment rates to support the observed population growth rates.
 The results suggest that the counting method is a versatile tool to estimate the recruitment rate if the species is monitored during the appropriate time window.
@@
       toc: false
-      keep_md: true
-      keep_tex: true
+      keep_md: false
+      keep_tex: false
     bookdown::word_document2:
@@
 The declination for species with delayed maturity follows the same process [see @Robertson2008].
-
 The recruitment rate parameter is complex because it is composed of two main sub-parameters: the fecundity followed by the juvenile survival [@Etterson2011; @Arnold2018].
@@
 These decision-support tools are then key inputs to status on the best management rules to implement.
-The growth rate of a population depends on the adult survival and the recruitment rate, see Equation \@ref(eq:GR) in Section \@ref(PGR).
+The growth rate of a population depends on the adult survival rate and the recruitment rate, see Equation \@ref(eq:GR) in Section \@ref(PGR).
 The adult survival is generally estimated by mark recapture/recovery methods [@Lebreton2001].
@@
 @Iverson2013], it is quite straightforward to predict its impact on the adult survival.
-
+The recruitment rate is thus a key parameter to build management scenarios.
 According to @Blums2004, the recruitment for diving ducks is poorly related to fecundity, but depends mostly on juvenile survival, which is driven by weather conditions.
@@
 This expansion is limited by the carrying capacity of the habitat [@Sayre2008].
-When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity because the individuals compete for space or/and food [@Nummi2015], which prevents the population from expanding.
+When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity because the individuals compete for space or/and food [@Nummi2015], which prevents the population from expanding indefinitely.
 This competition commonly induces a lower fecundity or/and a higher juvenile mortality.
@@
 
-Hunting waterfowl induces adult and juvenile mortality [@Bellrose1980].
+Waterfowl hunting induces adult and juvenile mortality [@Bellrose1980].
 Intuitively, one expects that such a hunted population would observe lower adult survival and recruitment rate and then would rapidly deplete to its extinction.
@@
 For such question, the carrying capacity of the habitat is theoretically not reached.
-So a manager should mostly focus on collecting knowledge on the maximum recruitment rate and its variability by removing the potential density-dependent effects occurring only for the highest theoretical densities [@Peron2013; @Eraud2021].
+Any manager should then mostly focus on collecting knowledge on the maximum recruitment rate and its variability by removing the potential density-dependent effects occurring only for the highest theoretical densities [@Peron2013; @Eraud2021].
 
@@
 
-Hunting waterfowl has a long history, but the biological effects of long-term moratorium have been poorly surveyed [@MartinezAbrain2013].
+Waterfowl hunting has a long history, but the biological effects of long-term moratorium have been poorly surveyed [@MartinezAbrain2013].
 So there are few opportunities to explore the response of the recruitment rate to a gradient of pressure for such species.
-Alien vertebrate species are commonly introduced during release events [@Saul2016].
-They usually observe a first period of colonisation on new territories without management measures.
+Alien vertebrate species are commonly introduced during release events [@Saul2016] and usually observe a first period of colonisation on new territories without management measures.
 Since its presence might impact the balance of the colonized ecosystem, it happens that managers take strong control measures to restrict the growth of these populations or to eradicate them [@Oficialdegui2020].
@@
 To do so, it is necessary to monitor nests in habitats representative of the considered population [@Blums2004].
-The survival of a chick to its recruitment can be done by capture mark recapture method quite tricky to set up [@Blums2004; @Schmidt2008; @Arnold2018].
-Even if such approach provides valuable information on the dynamics of the recruitment and can highlight the bottlenecks, the fieldwork required to estimate both components is too time-consuming to generalize such approach for waterfowl population management on a long-term horizon.
+The survival of a chick to its recruitment can be done by capture mark recapture methods that are quite tricky to set up [@Blums2004; @Schmidt2008; @Arnold2018].
+Even if such approach provides valuable information on the dynamics of the recruitment and can highlight the bottlenecks of survival, the fieldwork required to estimate both components is too time-consuming to generalize such approach for waterfowl population management on a long-term horizon.
 
@@
 The hunting bag analysis might then lead to overestimate the recruitment rate and consequently the maximum sustainable harvest rate, with adverse consequences on population conservation [@Fox2014].
-Inferring recruitment rate from the proportion of immature birds trapped during tagging studies is also promising [@Arnold2018], but the author admits that juveniles might be more naïve and are then more likely caught in traps [@RguibiIdrissi2003].
+Inferring recruitment rate from the proportion of immature birds trapped during tagging studies is also promising [@Arnold2018], but the author admits that immatures might be more naïve and are then more likely caught in traps [@RguibiIdrissi2003].
 
@@
 A new population set up in France from the 90's, likely because of an arrival of a few individuals from the United Kingdom (UK).
-Unlike the UK population, the French one set up around a single wintering spot, the Grand Lieu Lake (47.09°N, 1.67°W), which facilitates the population monitoring.
+Unlike the UK population, the French one set up around a single wintering spot, the Grand Lieu Lake (47.09°N, 1.67°W), which facilitates the monitoring of this population.
 Since this species is considered as a threat to the white-headed duck because of the risk of a genetic introgression [@MunozFuentes2007], an European plan of eradication has been adopted and control measures were taken in both countries [@GutierrezExposito2020].
 
-Just before the reproduction period, one can distinguish from shore individuals with male-like plumage from female-like individuals on wintering spots (Figure \@ref(fig:flock)).
-It is noticeable that the apparent proportion of males in the counts from shore is far below the proportion of males in the samples from hunting bags (Figure \@ref(fig:pmal)).
+Just before the reproduction period on wintering spots, one can distinguish from shore individuals with male-like plumage from female-like individuals (Figure \@ref(fig:flock)).
 
@@
 
+It is noticeable that the apparent proportion of males in the counts from shore is far below the proportion of males in the samples from hunting bags (Figure \@ref(fig:pmal)).
+Assuming sex identification of samples is very accurate, such observation demonstrates that the counting method misses the identification of a significant part of males in the population.
+In many waterfowl species, immature individuals of both sexes display cover-up plumage similar to adult females [@Reeber2015].
+Consequently, the counts in winter allows adult males to be distinguished from all other individuals (adult females, immature females, and immature males).
+This feature can be valuable to estimate the recruitment with two conditions.
+The season when adult males can be distinguished from all other individuals is sufficiently close to the breeding period to consider this age structure picture to be representative of the age structure of the breeding population.
+The adult sex ratio is known and stable over the considered time period.
+
 ```{r pmal, fig.cap = "(ref:pmal)"}
@@
 
-Assuming sex identification of samples is very accurate, such observation demonstrates that the counting method misses the identification of a significant part of males in the population.
-In many waterfowl species, immature individuals of both sexes display cover-up plumage similar to adult females [@Reeber2015].
-Consequently, the counts in winter allows adult males to be distinguished from all other individuals (adult females, immature females, and immature males).
-This feature can be valuable to estimate the recruitment with two conditions.
-The season when adult males can be distinguished from all other individuals is sufficiently close to the breeding period to consider the age structure picture of the population to be representative of the age structure of the breeding population.
-The adult sex ratio is known and stable over the considered time period.
-
 The demography of both ruddy duck populations has been tracked thanks to exhaustive counts on the wintering spots during the period between December 1 and January 31.
-Controlled individuals between the counting date and the beginning of the reproduction period (set up to May 1 upon field observations) were subtracted from the winter count to get a proxy of the breeding population size.
+Controlled individuals between the counting date and the beginning of the reproduction period (set up to May 1 upon field observations of nests) were subtracted from the winter count to get a proxy of the breeding population size.
 Time series show that the two populations follow similar demographic histories (Figure \@ref(fig:count)).
@@
 Adult sex ratio in waterfowl is generally biased towards males (Figure \@ref(fig:pmal), @Wood2021).
-As a consequence, deducing the immature proportion from the adult male proportion is not straightforward because the adult proportion twice as much as the male proportion.
+As a consequence, deducing the immature proportion from the adult male proportion is not straightforward because the adult proportion is not just twice as much as the male-like proportion.
 Even if the adult sex ratio may vary over a long time range, it is relatively stable over a few years [@Wood2021].
 Sex identification in adult samples from hunting bags was available for both ruddy duck populations.
-However, the small population size in France prevents from providing precise sex proportion estimates.
+However, the small population size in France prevents from getting precise adult sex ratios.
 A comparison over months in UK showed no difference in male proportion, so adult samples collected over the whole year were used to estimate the male proportion.
 A comparison of male proportion among years with more 500 samples did not show significant interannual differences.
-As a consequence, all adult samples were pooled to estimate the adult male proportion, see Equation \@ref(eq:pm).
+As a consequence, all adult samples were pooled to estimate the male proportion in adults, see Equation \@ref(eq:pm).
 
@@
 |$SAF$|Data|Total number of adult females in samples|
-|$p(m|a)$|Parameter|Proportion of males in adults (or probability to be an adult male)|
+|$p(m|a)$|Parameter|Proportion of males in adults (or probability to be a male knowing it is an adult)|
 |$CAM_{i, t}$|Data|Cumulated number of type-male individuals counted in population $i$ in year $t$, which are assumed to be adult males|
@@
 A direct estimation of the immature proportion can be made using samples from hunting bags, see Equation \@ref(eq:pis).
-By combining estimates of immature proportion and Equation \@ref(eq:r), recruitment rates from samples were estimated to check the variability consistency of the counting method.
+By combining estimates of immature proportion and Equation \@ref(eq:r), recruitment rates from samples were estimated to check the consistency of the counting method to track the recruitment variability.
 One selected only years with more than 100 individuals controlled in winter, which correspond to a 9-year period in UK.
@@
 If the recruitment rate outputs of the model are significantly superior to this proxy, it means that the method overestimates the recruitment rate, which may reflect that some adult males are not detected during count surveys.
-By splitting the recruitment rate outputs upon population and harvest category, one can discuss of the impact of the harvest on the recruitment rate.
+By splitting the recruitment rate outputs upon populations and harvest categories, one can discuss of the impact of the harvest on the recruitment rate.
 
@@
 For all the sub-models in Section \@ref(MM), the Bayesian framework was used for its efficiency and simplicity to propagate error through the parameters.
-One used only uninformative priors on all parameters.
+One used uninformative priors on all parameters.
 One generated three chains of length 500000, with a thinning of 10 to avoid autocorrelation in the samples, and one discarded the first 2000 samples as burn-in.
@@
 
-The adult male proportion, which is necessary to infer on the recruitment rate from count data, is estimated at 0.60 [0.59; 0.61].
+The male proportion in adults, which is necessary to infer on the recruitment rate from count data, is estimated at 0.60 [0.59; 0.61].
 The counting method provides estimations of the recruitment rate, see Figure \@ref(fig:rrts)A.
@@
 The recruitment rates from the counting method are in the same range for the two populations, which demonstrate a certain consistency in the results obtained from the method.
-The higher variability observed in France is not unexpected because the French time series is longer and covers a larger harvest spectrum than in UK.
+The higher variability observed in France is not unexpected because the French time series is longer and covers a larger spectrum of harvest pressure than in UK.
 There is no temporal correlation between the two populations, so it suggests that the recruitment rate fluctuations are more likely influenced by regional-scale factors than large-scale ones.
@@
 
-Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation to the outputs of the counting method demonstrates the ability of this method to track the recruitment rate variability.
+Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation with the outputs of the counting method demonstrates the ability of this method to track the recruitment rate variability.
 This result is quite robust because the correlation is obvious with only a five-year time series.
-The assumption of a constant adult sex ratio over time is then acceptable.
-A significant change in adult sex ratio would indeed bias the evolution of the recruitment rate.
-However, since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong.
+The assumption of a constant adult sex ratio over time is then acceptable because a significant change in adult sex ratio among years would have biased the evolution of the recruitment rate.
+Since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong.
 Significant variations can then occur only every multiple years, especially if the species is long-living.
-Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base, as done in this study (the average adult sex ratio found here is in line with other duck species [@Wood2021]).
-However, the robustness of the correlation between the two methods might be overrated because two successive population size values are part of the recruitment rate, which mechanically scale the outputs.
+Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.
+The adult sex ratio found here on a restricted time period is in line with other duck species [@Wood2021].
+The robustness of the correlation between the two methods might be overrated because two successive population size values are parts of the recruitment rate, which mechanically scale the outputs.
 It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit more noisy than for the recruitment rate (Figure \@ref(fig:sup)A).
@@
 
-The counting method always provides recruitment rates significantly lower than the sampling method by factor close to 2 on average, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable than adults to hunters [@Bellrose1980; @Fox2014].
-This is a robust result because the approach used to control ruddy ducks differ strongly from usual hunting practices and are apriori less selective.
-It is especially clear in the UK, for which control operations occur in winter using long-range rifles that prevents the bird from escaping from the threat.
+The counting method always provides recruitment rates significantly lower than the sampling method by a factor close to 2, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable than adults to hunters [@Bellrose1980; @Fox2014].
+This is a robust result because the approach used to control ruddy ducks differ strongly from usual hunting practices and is apriori less selective.
+It is especially clear in the UK, for which control operations occurred in winter using long-range rifles that prevents the bird from escaping from the threat.
 The assumption on the accuracy of the counting method is then not excluded from the consistency of this result.
@@
 The consistency of this intermediate result suggests that the population size estimation is quite good for both populations and demonstrates that their evolution is well tracked even with a few individuals.
-The average recruitment rate estimated from the counting method, with or without harvest, are not different of significantly below the proxies of maximum recruitment rate for both populations.
+The average recruitment rates estimated from the counting method, with and without harvest, are never higher to the proxies of maximum recruitment rate for both populations.
 The assumption on the accuracy of the counting method is then again not excluded from the consistency of this result.
@@
 
-Variability within harvest category represents
+Only the French population was monitored without and with harvest.
+The harvest pressure, that is defined into a binary variable (no harvest / harvest), affects on average negatively both the recruitment rate and the adult survival for this population.
+If it explains only 4.6% of the variance of the adult survival, this values raises to 19.3% for the recruitment rate of the French population.
+The harvest pressure applied in France is thus a factor that likely induces a strong immature mortality impacting the recruitment rate, but a negligible mortality on the adults.
+This conclusion is obviously opposed to the hypothesis that the heterogeneity in immature individuals would dampen the effect of the harvest pressure on the resulting recruitment rate.
+Two points can however explain this apparent discrepancy.
+Since the management targeted the eradication of this ruddy duck population, the harvest on immatures might be simply far higher than what the population could support and finally affect the recruitment rate.
+Also, the population started to be controlled when it was still growing at a high rate, so immature individuals were not suffering from density-dependent effects and were consequently sensitive to any other mortality sources, like a new harvest pressure for instance.
+The low impact on adult survival compared to the recruitment rate might reflect the timing of the management strategy adopted in France.
+Since the culling efficiency is poor on the wintering spot, most of control operations occur in summer just after the reproduction, when the proportion of immature birds is the largest.
 
-The harvest pressure, simplified into a two-category variable, explains already 19.3% and 4.6% of the variance of the recruitment rate and the adult survival rate of the French population.
-The harvest pressure described as a continuous variable is expected to explain even more the recruitment rate variability.
-The harvest pressure applied in France is thus a factor that likely induces a strong immature mortality, but a moderate one on the adults.
-No impact on the adult survival in France, but it seems that there is an impact in UK.
-Why, because harvest occurs mostly in summer in France, and in winter in UK.
+The UK population depleted from 2005, which also corresponds to a significant increase of the harvest effort.
+Even if only a few data from the sampling method are available before and after this year, it seems the recruitment rate is not affected by this increase of harvest whereas the adult survival dropped significantly.
+This differs strongly from the response of the French population to the harvest, likely because a significant part of the control operations in UK occur in winter, i.e.
+just before the reproduction period.
+Indeed, winter corresponds to the season with the lowest proportion of immatures in the population because their survival is far lower than adults, so the adult survival is more affected by this type of harvest pressure.
+
+The strong divergence of the response to the harvest pressure of the two ruddy duck populations demonstrates that it is necessary to account for the time window over which the harvest occurs to produce proper predictions.
+Controlling a waterfowl population is then not only a question about how big should be the harvest effort, but also when this effort should focus within a year to be efficient.
+This requires further investigation to calibrate the response of the population to different harvest regimes.
+
+The ruddy duck population in UK starts to deplete from 2005, a year that
+
+No impact on the adult survival in France, but it seems that there is an impact in UK. Why, because harvest occurs mostly in summer in France, and in winter in UK.
 
@@
 
-# (APPENDIX) Supplement {-}
+# (APPENDIX) sup {-}
 

### 2022-01-05 Code/Draft_ruddy_duck_recruitment_rate.Rmd
discussion edits

@@
                       fig.align = "center",
-                      fig.pos = 'H') # hold position / require package "float"
+                      fig.pos = 'H') # hold position / requires package "float"
 
@@
 The demography of both ruddy duck populations has been tracked thanks to exhaustive counts on the wintering spots during the period between December 1 and January 31.
-Controlled individuals between the counting date and the beginning of the reproduction period (set up to May 1 upon field observations of nests) were subtracted from the winter count to get a proxy of the breeding population size.
+Controlled individuals between the counting date and the beginning of the reproduction period (set up to May 1 upon field observations of nests and @Bellrose1980) were subtracted from the winter count to get a proxy of the breeding population size.
 Time series show that the two populations follow similar demographic histories (Figure \@ref(fig:count)).
@@
 
-## Comparison to a proxy of maximum recruitment rate
+## Comparison to maximum recruitment rate
 
@@
 Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.
-The adult sex ratio found here on a restricted time period is in line with other duck species [@Wood2021].
+The adult sex ratio found here on a restricted time period is in line with the ruddy duck in its native area [0.62 in @Bellrose1980] and with other duck species [@Wood2021].
 The robustness of the correlation between the two methods might be overrated because two successive population size values are parts of the recruitment rate, which mechanically scale the outputs.
@@
 The counting method always provides recruitment rates significantly lower than the sampling method by a factor close to 2, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable than adults to hunters [@Bellrose1980; @Fox2014].
+For hunted duck species in North America, the harvest-induced mortality on first-year immatures compared to adults is higher by a factor ranging from 1.3 to 2.6 [@Bellrose1980].
 This is a robust result because the approach used to control ruddy ducks differ strongly from usual hunting practices and is apriori less selective.
@@
 
-Overall, even if the true values of recruitment rates are not directly accessible to test the accuracy of the counting method, a beam of arguments support that the method is not heavily biased.
-This consists in a big difference with the sampling method as demonstrated in @Fox2014, even when one considers a picture limited to the end of the hunting season to limit the bias [@Fox2016].
+Overall, even if the true values of recruitment rates are not known to test the accuracy of the counting method, a beam of arguments support that the method is not heavily biased.
+This consists in a big difference with the sampling method as demonstrated in @Fox2014, even when one considers a age structure picture that is limited to the end of the hunting season to limit the bias [@Fox2016].
 
@@
 
-Only the French population was monitored without and with harvest.
+Only the French population was monitored over a period covering years without and with harvest.
 The harvest pressure, that is defined into a binary variable (no harvest / harvest), affects on average negatively both the recruitment rate and the adult survival for this population.
-If it explains only 4.6% of the variance of the adult survival, this values raises to 19.3% for the recruitment rate of the French population.
-The harvest pressure applied in France is thus a factor that likely induces a strong immature mortality impacting the recruitment rate, but a negligible mortality on the adults.
+If it explains only 4.6% of the variance of the adult survival, this value raises to 19.3% for the recruitment rate of the French population.
+The harvest pressure applied in France is thus a factor that likely induces a strong juvenile mortality that impacts the recruitment rate, but a negligible mortality on the adults.
 This conclusion is obviously opposed to the hypothesis that the heterogeneity in immature individuals would dampen the effect of the harvest pressure on the resulting recruitment rate.
 Two points can however explain this apparent discrepancy.
-Since the management targeted the eradication of this ruddy duck population, the harvest on immatures might be simply far higher than what the population could support and finally affect the recruitment rate.
-Also, the population started to be controlled when it was still growing at a high rate, so immature individuals were not suffering from density-dependent effects and were consequently sensitive to any other mortality sources, like a new harvest pressure for instance.
-The low impact on adult survival compared to the recruitment rate might reflect the timing of the management strategy adopted in France.
+Since the management targeted the eradication of this ruddy duck population, the harvest on immatures might be simply far higher than what the population could compensate and finally affects the recruitment rate.
+Also, the population started to be controlled when it was still growing at a high rate, so immature individuals were not suffering from density-dependent effects and were consequently sensitive to any additional mortality sources, like a new harvest pressure for instance.
+The low impact of the harvest on adult survival compared to the recruitment rate might reflect the timing of the management strategy adopted in France.
 Since the culling efficiency is poor on the wintering spot, most of control operations occur in summer just after the reproduction, when the proportion of immature birds is the largest.
@@
 The UK population depleted from 2005, which also corresponds to a significant increase of the harvest effort.
-Even if only a few data from the sampling method are available before and after this year, it seems the recruitment rate is not affected by this increase of harvest whereas the adult survival dropped significantly.
-This differs strongly from the response of the French population to the harvest, likely because a significant part of the control operations in UK occur in winter, i.e.
-just before the reproduction period.
-Indeed, winter corresponds to the season with the lowest proportion of immatures in the population because their survival is far lower than adults, so the adult survival is more affected by this type of harvest pressure.
+Even if only a few data from the sampling method are available before and after this year, it seems that the recruitment rate is not affected by this increase of harvest whereas the adult survival dropped significantly.
+This differs strongly from the response of the French population to the harvest.
+Unlike in France, a significant part of the control operations in UK occur in winter, which could explain this difference.
+Indeed, winter corresponds to the season just before the reproduction period, when the proportion of immatures in the population is the lowest over the year because their survival is far lower than adults.
+The adult survival seems then mostly affected by the harvest pressure applied in winter.
 
@@
 Controlling a waterfowl population is then not only a question about how big should be the harvest effort, but also when this effort should focus within a year to be efficient.
-This requires further investigation to calibrate the response of the population to different harvest regimes.
+These first insights suggest that further investigation are required to calibrate the response of the population to different harvest regimes.
 
-The ruddy duck population in UK starts to deplete from 2005, a year that
+## Implication in waterfowl management
 
-No impact on the adult survival in France, but it seems that there is an impact in UK. Why, because harvest occurs mostly in summer in France, and in winter in UK.
+Age ratio, or indifferently recruit proportion, is commonly used to describe the renewal capacity of a waterfowl population and to track the evolution of its productivity [@Smith2001; @Bellrose1980; @Robertson2008; @Rodway2015; @Zimmerman2010].
+The recruitment rate is however a better indicator than the age ratio, the latter being likely preferred for its simple estimation.
+Indeed, the age ratio/recruit proportion does not account for the dynamics of the population and reflects the population productivity only if the population growth is steady.
+For instance, for a breeding population of 100 individuals reaching the next year 100 individuals and then 60 individuals the second year, a 50% recruit proportion in both cases reflect actually a strong decrease of the recruitment rate, respectively 0.5 recruits per breeder and 0.3 recruits per breeder.
+This demonstrates that the age ratio/recruit proportion suffers of caveats that can be misleading for a manager.
+According to @Blums2004, the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, the other component of the recruitment that is mostly driven by the weather conditions.
+Studies on other birds also conclude of the poor correlation between fecundity and recruitment [e.g.
+@Murray2000].
+Fecundity alone does not provide all knowledge required for a manager to understand the dynamics of a population.
 
-Impact on the recruitment rate in France, and no impact in UK.
-Recruitment rate decrease with the harvest for the French population, and the UK population recruitment rate is also below the French one. 
-it seems quite clear that the harvest affects the recruitment rate.
+Immatures are more prone to die naturally than adults and are submitted to density dependent processes, which explains why the recruitment rate is the most sensitive parameter to density dependent effect [@Koons2014].
+This means that culling mostly immatures can have limited effect of the population dynamics if the harvest pressure remains moderate.
+Considering that the harvest effort occurring in winter impacted mostly adult survival and led to the strong depletion of the UK ruddy duck population, harvesting early in the season should be favoured to maximize the catch in preserving the resources.
+If the goal is to eradicate a species, a strong effort should be done especially in the late season as done in the UK, just before the reproduction.
 
-Overall, the harvest seems to impact differently France and UK, likely because the harvest mortality occurs mostly in summer in France, and in winter in UK.
+Tracking recruitment rate: If the maximum recruitment rate is not reached under optimal conditions (maximum population size not reached, and limited harvest pressure), it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the recruitment rate.
 
-France affected the recruitment, UK affected adult survival for a similar result -> depletion
-
-For hunted duck species in North America, the mortality rate on first-year immatures compared to adults is higher by a factor ranging from 1.4 to 2 (Bellrose 1980).
-
-so more variability should be reflected in a difference of variability between the 2 parameters, but not true, how to make our result consistent with these observations?
-Huge harvest pressure, and adult are targetted, so compensation on adult survival.
+Condition of application to other species.
 
@@
 
-## Implication in waterfowl management
-
 Recruitment rate drives lambda [@Arnold2018]
 
-According to @Blums2004, the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, the other component of the recruitment that is moslty driven by the weather conditions.
-Studies on other birds also conclude of the poor correlation between fecundity and recruitment [e.g.
-@Murray2000].
-
-recruitment rate the most sensitive parameter to density dependent effect [@Koons2014]
-
 Population growth is equally influenced by adult survival and recruitment
-whereas adult survival more consistent between the tow (.5 factor), meaning there was more variability in recruitment rate than in survival rate.
+whereas adult survival more consistent between the two (.5 factor), meaning there was more variability in recruitment rate than in survival rate.
 growth rate difference is thus more explained by recruitment rate fluctuation than adult survival fluctuation, consistent with the demographic buffering, or canalization hypothesis [@Gaillard2003; @Lenzi2021]
-
 Au dela du fait que la croissance de la pop est la somme de la survie et de la recruitment rate, intéressant de parler de la variability de la survie vs variabilité de la recruitment rate pour parler de risk management.
-
 But caution, a population model linking the harvest rate to state variables describing the population is necessary to check precisely the impact of the harvest pressure, and test hypothesis on the functioning of harvest mortality.
@@
 
-recruitment rate better than sex ratio: a same value of age ratio can correspond to different fecundity levels, so warning.
-a same value of age ratio for a population in depletion and a population stabilized are not similary productive, the first one is less productive than the second one !
-so warning, focus on recruitment rate and not on age ratio if the dynamics of the population varies!
-recruitment proportion is commonly used in to discuss the fecundity of a population, but unlike the recruitment rate, this indicator does not consider the dynamics of the population and does reflect fecundity only on stable growth period
-
-Taper des jeunes c'est pas forcément très utile car density dependence
-
 recruitment rate: drop from 2009
 
-one considers harvest as an impact only adult survival, not on age structure, but we should, oversimplistics
+one considers harvest as an impact only adult survival, not on age structure, but we should, oversimplification
 
-If the maximum recruitment rate is not reached under optimal conditions (maximum population size not reached, and limited harvest pressure), it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the recruitment rate.
-
-Condition of application to other species.
-
-# Conclusion
+## Conclusion
 

### 2022-01-06 Code/Draft_ruddy_duck_recruitment_rate.Rmd
Discussion edits

@@
 - Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr]
-# - Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
 - Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk]
 - Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr]
+# - Jean-François Maillard^[Office Français de la Biodiversité, France - jean-francois.maillard@ofb.gouv.fr]
 # - Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthieu.guillemain@ofb.gouv.fr]
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+# - Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
+date: "`r Sys.setlocale('LC_TIME', 'English'); format(Sys.time(), '%d %B, %Y')`"
 abstract: |
@@
 header-includes:
-  \usepackage{float}
+  - \usepackage{float} # allows to hold figure position
+  - \usepackage[left]{lineno} # line numbering
 ---
@@
 \newpage
+
+\linenumbers
 
@@
 Only the French population was monitored over a period covering years without and with harvest.
-The harvest pressure, that is defined into a binary variable (no harvest / harvest), affects on average negatively both the recruitment rate and the adult survival for this population.
+The harvest pressure, that is defined into a binary variable (no harvest / harvest), affects on average negatively both the recruitment rate and the adult survival for this population, as expected.
 If it explains only 4.6% of the variance of the adult survival, this value raises to 19.3% for the recruitment rate of the French population.
@@
 Since the management targeted the eradication of this ruddy duck population, the harvest on immatures might be simply far higher than what the population could compensate and finally affects the recruitment rate.
-Also, the population started to be controlled when it was still growing at a high rate, so immature individuals were not suffering from density-dependent effects and were consequently sensitive to any additional mortality sources, like a new harvest pressure for instance.
-The low impact of the harvest on adult survival compared to the recruitment rate might reflect the timing of the management strategy adopted in France.
+Also, the population started to be controlled when it was still growing at a high rate, so immature individuals were not suffering from density-dependent effects and were consequently sensitive to any additional mortality sources, like the new harvest pressure for instance.
+The low impact of the harvest on adult survival compared to the recruitment rate might be due to the timing of the management strategy adopted in France.
 Since the culling efficiency is poor on the wintering spot, most of control operations occur in summer just after the reproduction, when the proportion of immature birds is the largest.
@@
 Unlike in France, a significant part of the control operations in UK occur in winter, which could explain this difference.
-Indeed, winter corresponds to the season just before the reproduction period, when the proportion of immatures in the population is the lowest over the year because their survival is far lower than adults.
+Indeed, winter corresponds to the season just before the reproduction period, when the proportion of immatures in the population is the lowest because their survival is far lower than adults.
 The adult survival seems then mostly affected by the harvest pressure applied in winter.
@@
 The strong divergence of the response to the harvest pressure of the two ruddy duck populations demonstrates that it is necessary to account for the time window over which the harvest occurs to produce proper predictions.
-Controlling a waterfowl population is then not only a question about how big should be the harvest effort, but also when this effort should focus within a year to be efficient.
-These first insights suggest that further investigation are required to calibrate the response of the population to different harvest regimes.
+Controlling a waterfowl population is then not only a question about how big should be the harvest effort, but also mostly when this effort should focus within a year to be efficient.
+These first insights suggest that further investigation on this topic would be valuable to calibrate the response of the population to different harvest regimes.
 
@@
 
+Adult survival and recruitment rate equally influence the growth rate variability of the studied ruddy duck populations, but the relative variability of the adult survival is more than three times smaller than the recruitment rate.
+This observation is in line with the demographic buffering, or canalization hypothesis, on the survival parameter [@Gaillard2003; @Lenzi2021].
+@Arnold2018 even demonstrates that the recruitment rate can be the most influencing factor on the population growth in some waterfowl species.
+Population management in waterfowl cannot neglect a careful monitoring of the recruitment rate, even if this parameter might be more difficult to track compared to the adult survival.
+Tracking the recruitment rate provides however only indications to a manager that a population renewal is affected.
+If the maximum recruitment rate is not reached under optimal conditions, it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the recruitment rate.
+
 Immatures are more prone to die naturally than adults and are submitted to density dependent processes, which explains why the recruitment rate is the most sensitive parameter to density dependent effect [@Koons2014].
-This means that culling mostly immatures can have limited effect of the population dynamics if the harvest pressure remains moderate.
-Considering that the harvest effort occurring in winter impacted mostly adult survival and led to the strong depletion of the UK ruddy duck population, harvesting early in the season should be favoured to maximize the catch in preserving the resources.
-If the goal is to eradicate a species, a strong effort should be done especially in the late season as done in the UK, just before the reproduction.
+This means that culling mostly immatures could have only a limited effect on the population dynamics if the harvest pressure remains moderate.
+Considering that the harvest effort occurring in winter impacted mostly adult survival and led to the strong depletion of the UK ruddy duck population, harvesting early in the season should be favoured if the objective is to maximize the catch in preserving the resources.
+If the goal is to eradicate a population, a strong effort should be done especially in the late season as done in the UK, just before the reproduction.
+These hypotheses still requires to be tested thanks to a population model linking the harvest rate to state variables describing the population.
+Such tool is valuable to assess precisely the impact of the harvest pressure and to test hypotheses on the functioning of the harvest mortality [@Plard2019].
 
-Tracking recruitment rate: If the maximum recruitment rate is not reached under optimal conditions (maximum population size not reached, and limited harvest pressure), it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the recruitment rate.
-
-Condition of application to other species.
+Applying the method implemented in this study on any other waterfowl species implies that the adult males can be distinguished from the other individuals at distance.
+The latter in the season the dichromatism is observable, the better it is to assure a relevant picture of the age structure of the population and then good estimates of the recruitment rate.
 
@@
 variabilité petite pop.
-
-Recruitment rate drives lambda [@Arnold2018]
-
-Population growth is equally influenced by adult survival and recruitment
-whereas adult survival more consistent between the two (.5 factor), meaning there was more variability in recruitment rate than in survival rate.
-growth rate difference is thus more explained by recruitment rate fluctuation than adult survival fluctuation, consistent with the demographic buffering, or canalization hypothesis [@Gaillard2003; @Lenzi2021]
-Au dela du fait que la croissance de la pop est la somme de la survie et de la recruitment rate, intéressant de parler de la variability de la survie vs variabilité de la recruitment rate pour parler de risk management.
-But caution, a population model linking the harvest rate to state variables describing the population is necessary to check precisely the impact of the harvest pressure, and test hypothesis on the functioning of harvest mortality.
 
@@
 The authors acknowledge all the contributors of the data collection in the UK and in France, especially: **UK names**, Vincent Fontaine, Denis Lacourpaille, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface.
+
+\nolinenumbers
 

### 2022-01-07 Code/Draft_ruddy_duck_recruitment_rate.Rmd
discussion edits

@@
 - Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr]
-# - Jean-François Maillard^[Office Français de la Biodiversité, France - jean-francois.maillard@ofb.gouv.fr]
-# - Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthieu.guillemain@ofb.gouv.fr]
-# - Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
 date: "`r Sys.setlocale('LC_TIME', 'English'); format(Sys.time(), '%d %B, %Y')`"
@@
 output: 
+    bookdown::word_document2:
+      toc: false
     bookdown::pdf_document2:
@@
       keep_tex: false
-    bookdown::word_document2:
-      toc: false
 header-includes:
@@
 
-In waterfowl, it is generally impossible to differentiate immatures from adults without examining birds in hand.
+In waterfowl, it is generally impossible to differentiate immatures from adults without examining birds at hand.
 Aging individuals from samples coming from hunting bags is thus another opportunity to estimate the recruit proportion.
@@
 
-The counting method always provides recruitment rates significantly lower than the sampling method by a factor close to 2, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable than adults to hunters [@Bellrose1980; @Fox2014].
+The counting method always provides recruitment rates significantly lower than the sampling method by a factor close to two, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable than adults to hunters [@Bellrose1980; @Fox2014].
 For hunted duck species in North America, the harvest-induced mortality on first-year immatures compared to adults is higher by a factor ranging from 1.3 to 2.6 [@Bellrose1980].
@@
 The low impact of the harvest on adult survival compared to the recruitment rate might be due to the timing of the management strategy adopted in France.
-Since the culling efficiency is poor on the wintering spot, most of control operations occur in summer just after the reproduction, when the proportion of immature birds is the largest.
+Since the culling efficiency is poor on the wintering spot, most of control operations occured in summer just after the reproduction, when the proportion of immature birds is the largest.
 
 The UK population depleted from 2005, which also corresponds to a significant increase of the harvest effort.
-Even if only a few data from the sampling method are available before and after this year, it seems that the recruitment rate is not affected by this increase of harvest whereas the adult survival dropped significantly.
+Even if only a few data from the sampling method are available before and after this year, it seems that the recruitment rate is not affected by this increase of harvest, but was already in the lower range of the recruitment rate of the French population when it was exploited.
+On the other hand, the adult survival dropped significantly from 2005.
 This differs strongly from the response of the French population to the harvest.
-Unlike in France, a significant part of the control operations in UK occur in winter, which could explain this difference.
-Indeed, winter corresponds to the season just before the reproduction period, when the proportion of immatures in the population is the lowest because their survival is far lower than adults.
+Unlike in France, a significant part of the control operations in UK occured in winter, which could explain this difference.
+Indeed, winter corresponds to the season just before the reproduction period, when the proportion of immatures in the population is the lowest, the juvenile survival being far lower than adults [e.g.
+@Arnold2018].
 The adult survival seems then mostly affected by the harvest pressure applied in winter.
 
-The strong divergence of the response to the harvest pressure of the two ruddy duck populations demonstrates that it is necessary to account for the time window over which the harvest occurs to produce proper predictions.
+The different responses to the harvest pressure of the two ruddy duck populations demonstrate that it is necessary to account for the time window over which the harvest occurs to produce proper predictions.
 Controlling a waterfowl population is then not only a question about how big should be the harvest effort, but also mostly when this effort should focus within a year to be efficient.
@@
 Age ratio, or indifferently recruit proportion, is commonly used to describe the renewal capacity of a waterfowl population and to track the evolution of its productivity [@Smith2001; @Bellrose1980; @Robertson2008; @Rodway2015; @Zimmerman2010].
-The recruitment rate is however a better indicator than the age ratio, the latter being likely preferred for its simple estimation.
+The recruitment rate is however a better indicator than the age ratio, the latter having only the advantage to be more directly accessible.
 Indeed, the age ratio/recruit proportion does not account for the dynamics of the population and reflects the population productivity only if the population growth is steady.
-For instance, for a breeding population of 100 individuals reaching the next year 100 individuals and then 60 individuals the second year, a 50% recruit proportion in both cases reflect actually a strong decrease of the recruitment rate, respectively 0.5 recruits per breeder and 0.3 recruits per breeder.
-This demonstrates that the age ratio/recruit proportion suffers of caveats that can be misleading for a manager.
+For instance, for a breeding population of 100 individuals reaching the next year 100 individuals and then 60 individuals the second year, a 1:1 age ratio, or a 50% recruit proportion, for both years reflects actually a strong decrease of the recruitment rate, respectively 0.5 recruits per breeder and 0.3 recruits per breeder.
+This demonstrates that the age ratio/recruit proportion suffers from caveats that can be misleading for a manager.
 According to @Blums2004, the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, the other component of the recruitment that is mostly driven by the weather conditions.
@@
 @Murray2000].
-Fecundity alone does not provide all knowledge required for a manager to understand the dynamics of a population.
+Fecundity alone does not provide all knowledge that is required for a manager to understand the dynamics of a population.
 
 Adult survival and recruitment rate equally influence the growth rate variability of the studied ruddy duck populations, but the relative variability of the adult survival is more than three times smaller than the recruitment rate.
-This observation is in line with the demographic buffering, or canalization hypothesis, on the survival parameter [@Gaillard2003; @Lenzi2021].
+This observation is in line with the demographic buffering, or canalization hypothesis, on the adult survival parameter [@Gaillard2003; @Lenzi2021].
 @Arnold2018 even demonstrates that the recruitment rate can be the most influencing factor on the population growth in some waterfowl species.
 Population management in waterfowl cannot neglect a careful monitoring of the recruitment rate, even if this parameter might be more difficult to track compared to the adult survival.
-Tracking the recruitment rate provides however only indications to a manager that a population renewal is affected.
-If the maximum recruitment rate is not reached under optimal conditions, it is a signal that one should explore the intermediate life stages before recruitment to provide a diagnostic on the changes of the environment conditions that have affected the recruitment rate.
+A drop in the recruitment rate points out that a population renewal is affected but does show the causation.
+It is however a signal that intermediate life stages before recruitment must be investigated to provide a diagnostic on the changes of the environment conditions that have affected the recruitment rate.
 
 Immatures are more prone to die naturally than adults and are submitted to density dependent processes, which explains why the recruitment rate is the most sensitive parameter to density dependent effect [@Koons2014].
-This means that culling mostly immatures could have only a limited effect on the population dynamics if the harvest pressure remains moderate.
-Considering that the harvest effort occurring in winter impacted mostly adult survival and led to the strong depletion of the UK ruddy duck population, harvesting early in the season should be favoured if the objective is to maximize the catch in preserving the resources.
-If the goal is to eradicate a population, a strong effort should be done especially in the late season as done in the UK, just before the reproduction.
+The evolution of the UK population size seems to show first sights of density dependence from 1000 individuals, but the process was obviously not big enough to fully curb the population growth.
+Since there is a strong correlation between the number of birds observed and the number of sites occupied by the birds, this inflexion may be simply due to a limited number of new favourable sites to be colonized.
+Such density dependent process implies that culling mostly immatures could have only a limited effect on the population dynamics if the harvest pressure remains moderate.
+Considering that the harvest effort occurring in winter impacted mostly the adult survival and led to the strong depletion of the UK ruddy duck population, harvesting early in the season should be favoured if the objective is to maximize the catch in preserving the resources.
+If the goal is to eradicate a population, a strong effort should be done especially in the late season just before the reproduction, as done in the UK.
 These hypotheses still requires to be tested thanks to a population model linking the harvest rate to state variables describing the population.
@@
 
-Applying the method implemented in this study on any other waterfowl species implies that the adult males can be distinguished from the other individuals at distance.
-The latter in the season the dichromatism is observable, the better it is to assure a relevant picture of the age structure of the population and then good estimates of the recruitment rate.
-
-recruitment rate: 
-Stabilité grosse pop, 
-variabilité petite pop.
-
-It is quite consistent with the assumption in introduction.
-Since studied populations are in colonization of their habitat, maximum densities are not reached, and then no or limited density dependent effect occur.
-
-it may be possible that first sight of density dependence occur in UK from 1000 individuals, but was not big enough to fully curb the population growth.
-Since there is a strong correlation between the number of birds observed and the number of sites occupied by the birds, this inflexion may be simply due to a limited number of new favourable sites to be colonized.
-
-If counting method acceptable, moderate harvest pressure does not reduce the recruitment rate as expected, maybe due to density dependent mortality for juvenile but it affects it if strong harvest rate.
-That is very interesting bro but warning, these interpretation holds on weak estimations and adult survival equal to 1!!!
-
-What about heterogeneity, figure 8 -> One said that recruitment rate can increase if harvest because of heterogeneity and density dependence, but it is not what we observe in France, likely because the exploitation is high inhere and consequently impact.
-
-recruitment rate: drop from 2009
-
-one considers harvest as an impact only adult survival, not on age structure, but we should, oversimplification
-
-## Conclusion
-
-Message de l'article
-@Nichols2006 -> il faut compter de la façon dont on demande
+Applying the method implemented in this study to any other waterfowl species implies that the adult males can be distinguished from the other individuals at distance.
+The latter in the season the dichromatism is observable, the better it is to ensure a relevant picture of the age structure of the population and then good estimates of the recruitment rate.
+Modifying standard monitoring protocols to distinguish male-like from female-like individuals is almost costless but provides substantial increases in the efficiency and usefulness of monitoring results in conservation [@Nichols2006].
 

### 2022-01-11 Code/Draft_ruddy_duck_recruitment_rate.Rmd
equation edit

@@
 To test the accuracy of the method, the results are compared to estimates of recruitment rate using samples.
-One also tests the conistency of the recruitment rates to support the observed population growth rates.
+One also tests the consistency of the recruitment rates to support the observed population growth rates.
 The results suggest that the counting method is a versatile tool to estimate the recruitment rate if the species is monitored during the appropriate time window.
@@
 output: 
-    bookdown::word_document2:
-      toc: false
     bookdown::pdf_document2:
@@
       keep_tex: false
+    bookdown::word_document2:
+      toc: false
 header-includes:
@@
 In waterfowl, it is generally impossible to differentiate immatures from adults without examining birds at hand.
-Aging individuals from samples coming from hunting bags is thus another opportunity to estimate the recruit proportion.
+Ageing individuals from samples coming from hunting bags is thus another opportunity to estimate the recruit proportion.
 Such approach suffers from two caveats.
@@
 N_{t} & = N_{t-1} - D_{t-1} + R_{t} \notag \\
-N_{t} & = N_{t-1} - (1 - S_{t}).N_{t-1} + r_{t}.N_{t-1} \notag \\
-N_{t} & = S_{t}.N_{t-1} + r_{t}.N_{t-1} \notag \\
+N_{t} & = N_{t-1} - (1 - s_{t}).N_{t-1} + r_{t}.N_{t-1} \notag \\
+N_{t} & = s_{t}.N_{t-1} + r_{t}.N_{t-1} \notag \\
 \frac{N_{t}}{N_{t-1}} & = s_{t} + r_{t} \notag \\
@@
 The low impact of the harvest on adult survival compared to the recruitment rate might be due to the timing of the management strategy adopted in France.
-Since the culling efficiency is poor on the wintering spot, most of control operations occured in summer just after the reproduction, when the proportion of immature birds is the largest.
+Since the culling efficiency is poor on the wintering spot, most of control operations occurred in summer just after the reproduction, when the proportion of immature birds is the largest.
 
@@
 This differs strongly from the response of the French population to the harvest.
-Unlike in France, a significant part of the control operations in UK occured in winter, which could explain this difference.
+Unlike in France, a significant part of the control operations in UK occurred in winter, which could explain this difference.
 Indeed, winter corresponds to the season just before the reproduction period, when the proportion of immatures in the population is the lowest, the juvenile survival being far lower than adults [e.g.

### 2022-10-12 Code/Draft_ruddy_duck_vital_rates.Rmd
Figures update and revision Intro + M&M

--- 2022-01-11 Code/Draft_ruddy_duck_recruitment_rate.Rmd
@@
 bibliography: "`r paste0('C:/Users/', Sys.info()[['user']], '/Dropbox/JabRef/Library_OFB.bib')`"
-csl: https://www.zotero.org/styles/apa
+csl: https://www.zotero.org/styles/apa # https://osf.io/download/6hw3q/ 
 link-citations: true
-title: "Recruitment rate estimation in waterfowl using a non-invasive method"
+title: "Exploiting delayed dichromatism to disentangle the effects of adult survival and recruitment on the population dynamics of waterfowl"
 author: 
 - Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr]
+- Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
 - Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk]
 - Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr]
+- Matthieu Guillemain^[Office Français de la Biodiversité, France - matthieu.guillemain@ofb.gouv.fr]
+
 date: "`r Sys.setlocale('LC_TIME', 'English'); format(Sys.time(), '%d %B, %Y')`"
 abstract: |
-  The response of a waterfowl population to a harvest pressure depends on its capacity to renew.
-The recruitment, i.e.
-the number of young adults reproducing for the first time, is a key indicator to describe the renewal of a population, and therefore an essential tool in population management.
-The recruitment rate, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species.
-The proportion of young adult in a waterfowl population, which is a main step in estimating the recruitment rate, is often estimated from game-hunting samples.
-However, this proportion is not accessible in years without harvest, or with a low harvest rate.
-Moreover, the age-structure in the harvest samples does not necessarily reflect the underlying age-structure of the population.
-It is often skewed towards immatures and can lead to an overestimation of the recruitment rate.
-In waterfowl, adult males usually display brighter colours than females and immatures from both sexes.
-This dichromatism can be characterized and monitored from count surveys, a non-invasive method.
-Such information can be used to infer the proportion of immatures, and consequently the recruitment rate.
-In using two populations of ruddy duck, *Oxyura oxyura*, this study introduces a model to estimate the recruitment rate from count data.
-To test the accuracy of the method, the results are compared to estimates of recruitment rate using samples.
-One also tests the consistency of the recruitment rates to support the observed population growth rates.
-The results suggest that the counting method is a versatile tool to estimate the recruitment rate if the species is monitored during the appropriate time window.
-Authors argue for considering a two-category count protocol in winter surveys of waterfowl species for which adult males can be differentiated from other individuals.
+  Monitoring the number of individuals is by far the most popular strategy for investigating the environmental factors ruling the population dynamics and for measuring the efficacy of management actions targeting population recovery, control or eradication.
+Unfortunately, it is insufficient for the basic understanding of the demographic mechanisms and more specifically to assess to what extent population growth rate is affected by changes in adult survival rather than to variations in reproductive parameters.
+Usual methods to inform survival and recruitment (capture-mark-recapture, game-hunting bag) rely on catching birds, which suffers from two main drawbacks.
+The selectivity of the catch methods can lead to a biased representation of the underlying population structure.
+In practice, catching and releasing birds is also an invasive approach implying additional disturbances for endangered species or is simply forbidden for invasive alien species.
+In several waterfowl species, a sexual dichromatism is observed in adults whereas immatures of both sexes display a cover-up plumage similar to adult females.
+From populations of Ruddy duck, a species displaying this pattern, this study introduces a non-invasive method based on count data to estimate the respective contribution of the recruitment and the adult survival in the population growth rate.
+Performing count survey during the appropriate time window to differentiate both plumage types is sufficient to provide long-term time series of the two main demographic components of the population growth rate.
 
-  **Keywords**: fecundity - productivity - reproductive success - age ratio - juvenile - duck
+  **Keywords**: fecundity - productivity - reproductive success - age ratio - duck
 output: 
@@
 
-Sys.setlocale("LC_ALL", "English")
+Sys.setlocale("LC_ALL", "en_US.UTF-8")
 
@@
 
-*Definitions*
+Assessing population growth-rate is a key step towards a better understanding of factors underlying the dynamics of natural populations [e.g.
+@Niel2005].
+It is also, crucial for measuring the efficacy of management actions eventually undertaken to help population recovery, control or eradication [@Shea1998].
+Among the approaches available to managers for reaching these goals, those relying on the monitoring of the numbers of individuals are by far the most popular [@Rintala2022].
+In many instances, these methods even enable environmental factors underlying changes in population size to be investigated and hence some mitigating actions to be implemented [@Faillettaz2019].
+Unfortunately, relying on the monitoring of numbers of individuals alone hinder basic understanding of the demographic mechanisms underlying changes in population growth rates.
+More specifically, based on counts only, it is nearly impossible to assess to what extent population growth rate is affected by change survival rather than to variations in reproductive parameters [@Austin2000].
+To assess the relative sensitivity of population growth rate to factors affecting survival or reproductive parameters, demographers most often rely on the monitoring of individuals by capture-mark-recapture [@Lebreton1992].
+In practice however, one cannot always rely on the capture and release of individuals owing to the legal status of the species that preclude any additional disturbance (critically endangered species), or forbid the release of alive individuals (invasive species or pests).
+Indeed, although highly efficient for assessing demographic parameters, CMR methods have potential drawbacks such as being invasive and hardly affordable when time and money are limiting.
+Genetic monitoring is a non-invasive alternative to CMR approaches based on physical captures, but it suffers from being costly and time-consuming and requires quite rigorous sampling schemes in the field.
 
-One defines the *recruitment rate* of a breeding population as the average number of recruits produced per each breeder [@Flint2015].
-This parameter might sometimes be called *annual fertility* [@Koons2014], *reproductive rate* [@Cooch2014], *reproductive success* [@Etterson2011], or *productivity* [@Johnson1987; @Hagen2008].
-If one considers a species reaching its sexual maturity as early as its first year, all individuals are mature at the reproduction season.
-The recruitment rate is thus the ratio of the number of young adults at the reproduction season of year $t$, i.e.
-the recruits, to the number of all individuals at the reproduction season of the previous year, $t - 1$.
-The declination for species with delayed maturity follows the same process [see @Robertson2008].
-The recruitment rate parameter is complex because it is composed of two main sub-parameters: the fecundity followed by the juvenile survival [@Etterson2011; @Arnold2018].
-Both sub-parameters can be split into other sub-parameters.
-The fecundity can be defined by the product of the nesting success, the average clutch size, and the hatching success.
-The juvenile survival can be defined as the product of the post-hatching survival and the post-fledging survival, both critical life periods [@Hill2008; @Blums2004].
-The recruitment rate as defined in the present study is thus a parameter integrating multiple life steps from the egg production to the young adult attempting to breed for its first time [@Koons2017].
-
-*The use of recruitment rate in population management*
-
-Population conservation, harvesting, and control require tools in population management in order to reach quantitative targets, such as a population size for example [@Shea1998; @Johnson2021].
-To do so, a first blind strategy is to test iteratively management rules to reach progressively an optimum.
-Such approach can be time-consuming given the intrinsic variability of the population response to its changing environment.
-It might also suffer from a rejection of the stakeholders because of the lack of transparency of an arbitrary process [@Williams2013].
-Another approach is to try to predict the growth rate of the population as a function of the harvest rate and other environment inputs.
-Such modelling approach can provide clues on the recovering time of a population upon a variety of management rules [@Otis2006].
-These decision-support tools are then key inputs to status on the best management rules to implement.
-The growth rate of a population depends on the adult survival rate and the recruitment rate, see Equation \@ref(eq:GR) in Section \@ref(PGR).
-The adult survival is generally estimated by mark recapture/recovery methods [@Lebreton2001].
-If the harvest pressure behaves as an additive mortality to the natural mortality for adults [e.g.
-@Iverson2013], it is quite straightforward to predict its impact on the adult survival.
-The recruitment rate is thus a key parameter to build management scenarios.
-According to @Blums2004, the recruitment for diving ducks is poorly related to fecundity, but depends mostly on juvenile survival, which is driven by weather conditions.
-Studies on other birds concludes also on the weak correlation between fecundity and recruitment [e.g.
-@Murray2000].
-The difficulty to track juvenile survival [@Schmidt2008] demonstrates the interest to focus on the recruitment rate when accessible.
-The key to predict the growth rate of a population is then to estimate the value of the recruitment rate and its variability, potentially influenced by the harvest pressure [@Anders2005].
-
-*The variability of the recruitment rate*
-
-A waterfowl species that is released in a favourable habitat will naturally expand [@Malthus1872].
-The growth of such a population reflects that the recruitment rate parameter is higher than the adult mortality rate.
-This expansion is limited by the carrying capacity of the habitat [@Sayre2008].
-When a waterfowl species is endemic of an ecosystem and evolves in stable environmental conditions, the population size varies around the carrying capacity because the individuals compete for space or/and food [@Nummi2015], which prevents the population from expanding indefinitely.
-This competition commonly induces a lower fecundity or/and a higher juvenile mortality.
-The population renewal is slowed down because the recruitment rate parameter is affected by density-dependent effects [@Gunnarsson2013].
-For an endemic species evolving in stable environmental conditions, the recruitment rate compensates on average the adult mortality rate, which explains the stabilization of the number of individuals [@Flint2015].
-
-Waterfowl hunting induces adult and juvenile mortality [@Bellrose1980].
-Intuitively, one expects that such a hunted population would observe lower adult survival and recruitment rate and then would rapidly deplete to its extinction.
-However, hunting waterfowl has a long history, and most of these species have persisted over time [@Cooch2014].
-So paradoxically, hunting immature individuals does not always induce a decrease of recruitment rate.
-The heterogeneity of quality among immature individuals might explain this by inducing a compensation process if the harvest selects mostly individuals with a low reproductive potential [@Lindberg2013; @Gimenez2017].
-The recruitment rate can even increase because the harvest pressure reduces the densities and potentially the competition for space and food [@Nummi2015; @Peron2012].
-In conclusion, a limited exploitation of a population observing heterogeneity and/or density-dependent process might paradoxically increase or at least not affect the recruitment rate.
-Under moderate exploitation, a new balance between recruitment rate and adult mortality can be theoretically reached [@Tsikliras2018].
-The size of a newly exploited population in stable environmental conditions should thus reach a different equilibrium, which is expected to be lower than the size without exploitation.
-This process explains why the exploitation of many waterfowl species is sustainable and has been lasting over a long time.
-
-The fecundity and the juvenile survival, the two components of the recruitment rate, are very sensitive to weather conditions [@Blums2004; @Folliot2017].
-The environment conditions are thus an additional source of variability of the recruitment rate.
-The response of the recruitment rate to a harvest pressure is thus not straightforward.
-It is noisy because of its sensitivity to environmental factors.
-It also depends directly on the intensity of harvest on the immature individuals and indirectly on the mitigation of the density-dependent effects.
-
-*What part of the recruitment rate gradient is necessary for population management?*
-
-In population management, a common question for a threaten or an invasive species is to determine the measure that will allow to reach a targeted population size/density in a limited time period [@Shea1998].
-For such question, the carrying capacity of the habitat is theoretically not reached.
-Any manager should then mostly focus on collecting knowledge on the maximum recruitment rate and its variability by removing the potential density-dependent effects occurring only for the highest theoretical densities [@Peron2013; @Eraud2021].
-
-*What is the ideal biological model to explore the gradient of the recruitment rate?*
-
-Waterfowl hunting has a long history, but the biological effects of long-term moratorium have been poorly surveyed [@MartinezAbrain2013].
-So there are few opportunities to explore the response of the recruitment rate to a gradient of pressure for such species.
-Alien vertebrate species are commonly introduced during release events [@Saul2016] and usually observe a first period of colonisation on new territories without management measures.
-Since its presence might impact the balance of the colonized ecosystem, it happens that managers take strong control measures to restrict the growth of these populations or to eradicate them [@Oficialdegui2020].
-Such species are then a good biological model to overcome density-dependent effects and explore the effects of a harvest gradient on recruitment rate.
-
-*How to estimate recruitment rate?*
-
-The recruitment rate parameter of a population can be assessed first by evaluating the fecundity of the breeding population and then by monitoring the survival from the egg life stage to the first-breeding event.
-To do so, it is necessary to monitor nests in habitats representative of the considered population [@Blums2004].
-The survival of a chick to its recruitment can be done by capture mark recapture methods that are quite tricky to set up [@Blums2004; @Schmidt2008; @Arnold2018].
-Even if such approach provides valuable information on the dynamics of the recruitment and can highlight the bottlenecks of survival, the fieldwork required to estimate both components is too time-consuming to generalize such approach for waterfowl population management on a long-term horizon.
-
-In waterfowl, it is generally impossible to differentiate immatures from adults without examining birds at hand.
-Ageing individuals from samples coming from hunting bags is thus another opportunity to estimate the recruit proportion.
-Such approach suffers from two caveats.
-First, it is impossible to assess the recruitment rate when there is no harvest whereas it is necessary to predict what would happen if a moratorium seems required to allow a population to recover, or to calibrate a priori the harvest effort necessary to control an invasive species [@Smith2005].
-Secondly, like any predation action, hunting is selective and provides a biased picture of the population age structure in favour of the immatures [@Bellrose1980; @Fox2014].
-The hunting bag analysis might then lead to overestimate the recruitment rate and consequently the maximum sustainable harvest rate, with adverse consequences on population conservation [@Fox2014].
-Inferring recruitment rate from the proportion of immature birds trapped during tagging studies is also promising [@Arnold2018], but the author admits that immatures might be more naïve and are then more likely caught in traps [@RguibiIdrissi2003].
-
-In waterfowl, it is common that adult males display brighter colours than adult females and immatures of both sexes, which display cover-up plumage similar to adult females [@Johnson1999].
-This discrepancy is valuable because it can be observed and quantified from distance.
-Knowing adult sex ratio, such approach is promising because it requires only to count the proportion of birds of the two plumage kinds to infer the recruitment rate [@Smith2001].
-Unlike the two previous methods, this one is non-invasive and requires an acceptable time investment, which allows to collect long time series and consequently to track recruitment rate variability.
-
-*Methodology of this study*
-
-From two populations of ruddy duck, a species introduced in Europe in the 40's, this study introduces a simple Bayesian model to infer recruitment rate from count data.
-A large part of this study focuses on assessing the viability of such approach.
-The comparison between two populations is a first step in testing the consistency of the method.
-A comparison to the common approach using samples from hunting bags aims at checking if the recruitment rate variability is well tracked and if the hypothesis stating that the recruitment rate is overestimated when using sample data is consistent with the results.
-Using two other approaches using population growth rates, one testes the consistency of the order of magnitude of the recruitment rate estimated from the counting method.
-Studying two populations is also an opportunity to discuss the response of the recruitment rate to a gradient of harvest pressure.
+Most of the time therefore, managers do the best of a bad job using counts as the only viable option for assessing population growth rates and the efficacy of management actions.
+Counting the number of immatures produced in addition to alive adults may enable adult survival and recruitment rate to be estimated, and hence provides knowledge on which of adult survival and productivity has the most influence on population growth rate.
+Unfortunately, this is only possible for a limited number of species in which broodless adults and those with young display the same detectability.
+Alternatively, assessing the sex and age structure in hunting bags has been used to infer the role of decreasing reproductive success in population declines in a number of games species including ducks and geese [@Fox2018].
+However, this approach is not suited for protected/endangered species.
+Here we exemplify how, in dimorphic species, delayed maturity of males can be exploited for estimating productivity/recruitment and adult survival by differentiating male-like to female-like individuals in repeated winter counts.
+We use two non-native Ruddy duck populations introduced in Europe as a study model.
+As numerous duck species, the Ruddy duck is dimorphic, with newly born males looking like females until the pre-nuptial moult, which occurs in late winter.
+This means that the apparent proportion of males increases over the course of the wintering season.
+These changes in apparent proportions of males during this period are therefore directly related to the proportion of immatures into the populations and hence to the reproductive success and the recruitment rate.
+This study details the method to infer adult survival and recruitment rate from count data and an application is developed to assess the impact of two different eradication strategies deployed in Great Britain and in France respectively.
 
@@
 
-## Biological model
+The ruddy duck is a stiff-tailed duck native from the Americas.
+From seven individuals initially acclimated in the Slimbridge Wetland Centre in the Great Britain in the 40’s [@GutierrezExposito2020], a feral population began to establish with the first observed reproductive attempts in the wild in the 60s.
+This feral population rapidly grew and spread to the entire country to reach more than 5000 individuals in the early 2000s.
+By the end of the 80s, a feral population began to establish also on the continent in France, supposedly owing to the arrival of individuals born in Great Britain (GB).
+Contrary to what was observed in GB, the French Ruddy duck population did not spread much, with the vast majority of observations and breeding attempts clustering in Western France.
+In France, during winter, almost none Ruddy duck is observed outside the lake of Grand Lieu (47.09°N, 1.67°W), which is thus the exclusive wintering site and greatly facilitates the monitoring of this population.
 
-The ruddy duck is a diving duck species introduced in the United Kingdom in the 40's [@GutierrezExposito2020].
-The first reproduction was observed in the 60's.
-The population rapidly grew to reach more than 5000 individuals spread over the entire country.
-A new population set up in France from the 90's, likely because of an arrival of a few individuals from the United Kingdom (UK).
-Unlike the UK population, the French one set up around a single wintering spot, the Grand Lieu Lake (47.09°N, 1.67°W), which facilitates the monitoring of this population.
-Since this species is considered as a threat to the white-headed duck because of the risk of a genetic introgression [@MunozFuentes2007], an European plan of eradication has been adopted and control measures were taken in both countries [@GutierrezExposito2020].
+The feral European populations of Ruddy duck are considered as a major threat to the endangered native White-headed duck population of the South-Western Mediterranean because of hybridization risks and thus an elevated potential for genetic pollution and eventually genetic assimilation of the latter by the former [@MunozFuentes2007].
+In order to mitigate the risks of genetic pollution of the White-headed duck by Ruddy ducks a European plan of eradication of Ruddy ducks has been adopted by the European Commission.
+As result, eradication measures were taken both in France and Great Britain in the late 90s [@GutierrezExposito2020].
 
-Just before the reproduction period on wintering spots, one can distinguish from shore individuals with male-like plumage from female-like individuals (Figure \@ref(fig:flock)).
+The release non-native species in the wild, including Ruddy ducks, is forbidden in countries of the European Union, and the status of White-headed duck populations is highly unfavourable.
+The use of capture-mark-recapture to monitor populations of these species was thus impossible.
+As a result, the effectiveness of the eradication of Ruddy ducks and management actions intended to favour the  recovery of White-headed ducks were essentially checked through censuses.
+Censuses, however, prevented from assessing the relative effects or variations in survival and reproductive success/recruitment on population growth rate, which is a key towards a proper understanding of the efficacy of management actions.
+
+Like many other ducks, Ruddy ducks and White-headed ducks both display delayed dichromatism.
+Thereby immature males acquire the typical colourful plumage of their kind over the course of the interbreeding season, which is typically in late winter.
+Immature males are thus confounded to immature and adult females before their moult (Figure \@ref(fig:flock)).
+The delayed dichromatism explains why the apparent proportion of females in the winter counts is always higher than the true proportion of females that can be observed in winter hunting bags (Figure \@ref(fig:pfem)).
+Thus, by monitoring the seasonal evolution of apparent sex-ratio, it is possible to assess the proportion of immature males within a population, which, combined with reliable counts and under some reasonable assumptions (e.g., even sex-ratio at birth, comparable mortality rates between females and males among juveniles, and unchanging adult sex-ratio over the monitoring period), enables both adult survival rate and recruitment rate to be estimated.
 
@@
 
-It is noticeable that the apparent proportion of males in the counts from shore is far below the proportion of males in the samples from hunting bags (Figure \@ref(fig:pmal)).
-Assuming sex identification of samples is very accurate, such observation demonstrates that the counting method misses the identification of a significant part of males in the population.
-In many waterfowl species, immature individuals of both sexes display cover-up plumage similar to adult females [@Reeber2015].
-Consequently, the counts in winter allows adult males to be distinguished from all other individuals (adult females, immature females, and immature males).
-This feature can be valuable to estimate the recruitment with two conditions.
-The season when adult males can be distinguished from all other individuals is sufficiently close to the breeding period to consider this age structure picture to be representative of the age structure of the breeding population.
-The adult sex ratio is known and stable over the considered time period.
-
-```{r pmal, fig.cap = "(ref:pmal)"}
-knitr::include_graphics("../Output/plot_1_saved.png")
+```{r pfem, fig.cap = "(ref:pfem)"}
+knitr::include_graphics("../Output/plot_pres_1.png")
 ```
 
-The demography of both ruddy duck populations has been tracked thanks to exhaustive counts on the wintering spots during the period between December 1 and January 31.
-Controlled individuals between the counting date and the beginning of the reproduction period (set up to May 1 upon field observations of nests and @Bellrose1980) were subtracted from the winter count to get a proxy of the breeding population size.
-Time series show that the two populations follow similar demographic histories (Figure \@ref(fig:count)).
-They grew freely in both countries until 1999.
-Control measures have been then applied from 1999 to 2005 in the UK and from 1999 to 2018 in France.
-In both countries, these management measures led to stop the expansion of the ruddy ducks.
-From 2005 in the UK and from 2018 in France, higher control effort led to a rapid depletion in both populations.
-This history underlines that both populations observed a large spectrum of harvest effort, from no pressure to very high pressure.
-For sake of interpretation, the time period 2001-2004 was classified in the *no harvest* category in France because the harvest pressure applied over this period was very limited with no effect on the population growth.
+We monitored the demography of Ruddy duck populations in both Great-Britain and France thanks to exhaustive counts on the wintering grounds during the period between December 1 and January 31.
+In addition to these exhaustive counts, censuses that distinguished female-like from male-like individuals were performed on the same period from 2007 to 2012 in UK, and in 1999, 2001-2009, 2012-2019 in France.
+During these years, the eradication effort, under control of the authorities in both countries, was also scrupulously recorded.
 
-```{r count, fig.cap = "(ref:count)"}
-knitr::include_graphics("../Output/plot_2.png")
-```
+## Vital rates inference from count data
 
-The counting dataset, which differentiates the male-like from the female-like individuals in winter, covers 7 years and 19 years in UK and France respectively (Figure \@ref(fig:count)).
-Combined together, the counting datasets cover different population dynamics corresponding to a gradient of harvest pressure.
-In France, data before 2004 correspond to a period of population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure (Figure \@ref(fig:count)).
-In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period.
-In parallel, the sampling dataset, which corresponds to individuals shot in winter during control operations, covers 9 years only in UK, with 5 years in common with the corresponding counting time series.
+In waterfowl, the survival is commonly higher in adult males than in adult females.
+Adult sex ratio is then generally biased towards males (Figure \@ref(fig:pfem), @Wood2021).
+As a consequence, deducing the immature proportion from the observed proportion of adult males is not straightforward because the adult proportion is not just twice as much as the observed proportion of adult males.
+There was no counts differentiating male-like from female-like during the breeding period to estimate properly adult sex ratio.
+But sex identification in adults from hunting bags was available for both Ruddy duck populations.
+However, the small population size in France prevented from getting precise adult sex ratios.
+A comparison over months in UK showed no difference in male proportion, so data from adults collected over the whole year were used to estimate the male proportion.
+A comparison of male proportion among years with more than 500 samples did not show significant interannual differences.
+Even if the adult sex ratio may vary over a long time range, it is relatively stable over a few years [@Wood2021].
+As a consequence, hunting bag data from all adults were pooled to estimate the male proportion in adults, see Equation \@ref(eq:pm).
 
-## Recruitment rate inference from count data
-
-Adult sex ratio in waterfowl is generally biased towards males (Figure \@ref(fig:pmal), @Wood2021).
-As a consequence, deducing the immature proportion from the adult male proportion is not straightforward because the adult proportion is not just twice as much as the male-like proportion.
-Even if the adult sex ratio may vary over a long time range, it is relatively stable over a few years [@Wood2021].
-Sex identification in adult samples from hunting bags was available for both ruddy duck populations.
-However, the small population size in France prevents from getting precise adult sex ratios.
-A comparison over months in UK showed no difference in male proportion, so adult samples collected over the whole year were used to estimate the male proportion.
-A comparison of male proportion among years with more 500 samples did not show significant interannual differences.
-As a consequence, all adult samples were pooled to estimate the male proportion in adults, see Equation \@ref(eq:pm).
-
-Assuming the additive property of the binomial distribution, the proportion of immature individuals can be deduced from the cumulated counts of male-like individuals in the wintering population, see Equation \@ref(eq:pic).
-Only years with more than 100 individuals were selected.
-From this proportion and abundance index of breeding population sizes, the recruitment rate is straightforward, see Equation \@ref(eq:r).
-If the absolute value of the breeding population size is accessible, the recruitment can be estimated, see Equation \@ref(eq:R).
+Assuming the additive property of the binomial distribution, the proportion of immatures can be deduced from the cumulated counts of male-like individuals in the wintering population, see Equation \@ref(eq:pic).
+From this proportion and the interannual variation of an abundance index of the population size, the adult survival and the recruitment rate are straightforward, see Equation \@ref(eq:vr).
+If the absolute value of the population size is accessible, numbers of adults and recruits can be estimated, see Equation \@ref(eq:VR).
 
@@
 |:---|:----|:----------------------------|
-|$SAM$|Data|Total number of adult males in samples|
-|$SAF$|Data|Total number of adult females in samples|
+|$AM$|Data|Total number of adult males in hunting bags|
+|$AF$|Data|Total number of adult females in hunting bags|
 |$p(m|a)$|Parameter|Proportion of males in adults (or probability to be a male knowing it is an adult)|
-|$CAM_{i, t}$|Data|Cumulated number of type-male individuals counted in population $i$ in year $t$, which are assumed to be adult males|
+|$CML_{i, t}$|Data|Cumulated number of male-like individuals counted in population $i$ in year $t$, which are assumed to be adult males|
 |$C_{i, t}$|Data|Cumulated number of individuals counted in population $i$ in year $t$|
 |$p(a \cap m)_{i, t}$|Parameter|Proportion of adult males in population $i$ in year $t$|
-|$p(a)_{i, t}$|Parameter|Proportion adults in population $i$ in year $t$|
+|$p(a)_{i, t}$|Parameter|Proportion of adults in population $i$ in year $t$|
 |$p(i)_{i, t}$|Parameter|Proportion of immatures in population $i$ in year $t$|
-|$N_{i, t}$|Data|Proxy of the size of the breeding population $i$ in year $t$ (maximum number of individuals counted in the wintering population minus removals before reproduction)|
-|$r_{i, t}$|Parameter|Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per breeder in year $t - 1$|
+|$N_{i, t}$|Data|Size of the population $i$ in year $t$ (maximum number of individuals counted in the wintering population)|
+|$s_{i, t}$|Parameter|Adult survival rate, i.e. proportion of adults in year $t-1$ still alive in year $t$|
+|$r_{i, t}$|Parameter|Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per adult in year $t - 1$|
+|$S_{i, t}$|Parameter|Number of adults in population $i$ in year $t$|
 |$R_{i, t}$|Parameter|Number of recruits in population $i$ in year $t$|
@@
 \begin{equation}
-p(m|a) \sim {\sf Beta}(SAM, SAF)
+p(m|a) \sim {\sf Beta}(AM, AF)
 (\#eq:pm)
@@
 \begin{align}
-CAM_{i, t} & \sim {\sf Binom}(p(a \cap m)_{i, t}, C_{i, t}) \notag \\
+CML_{i, t} & \sim {\sf Binom}(p(a \cap m)_{i, t}, C_{i, t}) \notag \\
 & \sim {\sf Binom}(p(m|a).p(a)_{i, t}, C_{i, t}) \notag \\
@@
 \begin{equation}
+s_{i, t} = \frac{(1 - p(i)_{i, t}).N_{i, t}}{N_{i, t - 1}}
 r_{i, t} = \frac{p(i)_{i, t}.N_{i, t}}{N_{i, t - 1}}
-(\#eq:r)
+(\#eq:vr)
 \end{equation}
@@
 \begin{equation}
+S_{i, t} = (1 - p(i)_{i, t}).N_{i, t}
 R_{i, t} = p(i)_{i, t}.N_{i, t}
-(\#eq:R)
+(\#eq:VR)
 \end{equation}
 
-## Comparison with the sampling method
+## Validation of the count method {#VAL}
 
-A direct estimation of the immature proportion can be made using samples from hunting bags, see Equation \@ref(eq:pis).
-By combining estimates of immature proportion and Equation \@ref(eq:r), recruitment rates from samples were estimated to check the consistency of the counting method to track the recruitment variability.
-One selected only years with more than 100 individuals controlled in winter, which correspond to a 9-year period in UK.
-This sampling time series covers 5 years of the UK counting time series.
+The relevance of the method was assessed by testing the likelihood of the estimates of vital rates.
+For the adult survival rate, we checked if the values belonged to the expected interval $[0;1]$, and we compared the average values to literature data, since the adult survival is commonly assessed by capture mark recapture [@Lebreton2001].
+For the recruitment rate, which is rarely explicitly informed in the literature, we checked if the values were equal or below the maximum recruitment rate expected without exploitation and without negative density-dependence processes.
+If the recruitment rate outputs of the model are significantly superior to the maximum recruitment rate, it means that the method overestimates the recruitment rate, which reflects that a part of adult males are not detected during count surveys.
+A null or negative recruitment rate would mean that immature males moulted before the survey or/and the assumption on the stability of the adult sex ratio is too strong.
 
-|Name|Class|Description|
-|:---|:----|:----------------------------|
-|$SI_{i, t}$|Data|Number of immatures sampled in the wintering population $i$ in year $t$|
-|$S_{i, t}$|Data|Number of individuals sampled in the wintering population $i$ in year $t$|
-
-\begin{equation}
-SI_{i, t} \sim {\sf Binom}(p(i)_{i, t}, S_{i, t})
-(\#eq:pis)
-\end{equation}
-
-## Result validation using population growth rates {#PGR}
-
-The simple relationship (a bit more complex if a species with delayed maturity is considered, see @Robertson2008) between population growth rate, adult survival, and recruitment rate can be used to test the accuracy of the order of magnitude of recruitment rates estimated from the counting and the sampling methods, see Equation \@ref(eq:GR).
-This relationship comes from a simple reasoning for a closed population: the population size in year $t$ is equal to the number of breeders that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
+The maximum recruitment rate is deduced by using its relationship to the maximum growth rate and the maximum survival, see Equation \@ref(eq:GR).
+This relationship becomes more complex if a species with delayed maturity is considered, see @Robertson2008.
+This relationship comes from a simple reasoning for a closed population: the population size in year $t$ is equal to the number of adults that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
 the recruitment in year $t$.
 The growth rate of a population is thus the sum of the adult survival rate and the recruitment rate [@Flint2015].
+For an open population, adult survival and recruitment rate are confounded to adult and recruit migration respectively.
 
@@
 |:--|:----------------------------|
-|$N_{t}$|Number of breeders in year $t$|
-|$D_{t}$|Number of breeders dead after reproduction in year $t$|
+|$N_{t}$|Number of adults in year $t$|
+|$D_{t}$|Number of adults dead during year $t$|
 |$R_{t}$|Number of recruits in year $t$|
-|$s_{t}$|Adult survival rate, i.e. proportion of breeders in year $t-1$ still alive in year $t$|
-|$r_{t}$|Recruitment rate, i.e. number of recruits in year $t$ produced per breeder in year $t-1$|
-|$\lambda_{t}$|Growth rate of the population between the reproduction periods of year $t-1$ and year $t$|
+|$s_{t}$|Adult survival rate, i.e. proportion of adults in year $t-1$ still alive in year $t$|
+|$r_{t}$|Recruitment rate, i.e. number of recruits in year $t$ produced per adult in year $t-1$|
+|$\lambda_{t}$|Growth rate of the population between year $t-1$ and year $t$|
 
@@
 
-From the evolution of the breeding population size, population growth rate can be assessed every year to catch the variability range \@ref(eq:GR), or can be smoothed over a consistent time period to provide average population growth rate, see Equation \@ref(eq:L), which is a linear regression on the logarithm scale.
+The maximum growth rate occurred for both populations during their expansion phase, when the evolution of the population size is the steepest.
+To estimate robust maximum population growth rates for both populations, we smoothed annual population growth rate over a consistent time period by using a linear regression on the logarithm scale, see Equation \@ref(eq:L).
 
@@
 
-### Analysis of the adult survival
-
-The adult survival is commonly assessed by capture mark recapture [@Lebreton2001].
-Combined to annual population growth rate, adult survival rate allows recruitment rate to be estimated each year, and then compared with the outputs of the counting method.
-Unfortunately, no tagging study was performed on the two ruddy duck populations.
-However, the relationship \@ref(eq:GR) could still produce annual adult survival estimates that can be discussed.
-To do so, annual population growth rates overlapping the counting time series were estimated.
-
-Inconsistency on recruitment rate output, which is defined on $[0;\infty]$, would either lead to survival output $<0$, which points out an overestimation of the recruitment rate, or $>1$, which reflects that the recruitment rate cannot support the observed population growth and is thus underestimated.
-In between the two limit values, one still discuss the survival outputs by comparison with survival values published in the literature.
-
-### Comparison to a proxy of maximum recruitment rate
-
-When population growth is not limited by a harvest pressure and/or by density-dependent effects, both recruitment rate and adult survival are close to their maximum potential.
-Given a conservative assumption on the maximum adult survival of ruddy ducks and maximum population growth rate estimated during no harvest time period, one can provide a proxy of the maximum recruitment rate.
-If the recruitment rate outputs of the model are significantly superior to this proxy, it means that the method overestimates the recruitment rate, which may reflect that some adult males are not detected during count surveys.
-By splitting the recruitment rate outputs upon populations and harvest categories, one can discuss of the impact of the harvest on the recruitment rate.
-
-To do so, one estimates maximum population growth rates over time periods observing no harvest using Equation \@ref(eq:L).
-Since the estimate of the UK population size is noisy below 500 individuals, one rejected data before 1972 (Figure \@ref(fig:grts)).
-After reaching the threshold of 1000 individuals, the UK population growth observed a strong inflexion whereas no harvest pressure was applied.
-This observation led to split the analysis into two time series for the time periods used in the maximum growth rate inference in UK (Figure \@ref(fig:grts)).
+Practically, we used a very conservative prior on the maximum adult survival by setting a uniform distribution on $[0.7; 0.9]$ [the lower limit corresponds to the upper range of survival rates observed in literature for species of similar weight, the upper limit corresponds to the range of survival rates of long-life species, @Nichols1997; @Krementz1997; @Buxton2004].
+For the maximum growth rate, we rejected GB data before 1972 since the size estimate of this population is noisy below 50 individuals (Figure \@ref(fig:grts)).
+After reaching the threshold of 1000 individuals, the GB population growth observed a strong inflexion whereas no harvest pressure was applied.
+This observation suggests that a negative density-dependence process might occur beyond 1000 individuals and led to consider only the first sequence of the time series to infer on the maximum growth rate in GB.
+For the French population, the two sequences with birds and without harvest were used, i.e.
+1994-1999 and 2001-2004 (Figure \@ref(fig:grts)).
 
 ```{r grts, fig.cap = "(ref:grts)"}
-knitr::include_graphics("../Output/plot_3.png")
+knitr::include_graphics("../Output/plot_pres_1.png")
 ```
 
-To get proxies of maximum recruitment rate from Equation \@ref(eq:GR), one uses a very conservative prior on the adult survival by setting a uniform distribution on $[0.7; 0.9]$ [the lower limit corresponds to the upper range of survival rates observed in literature for species of similar weight, the upper limit corresponds to the range of survival rates of long-life species, @Nichols1997; @Krementz1997; @Buxton2004].
-The average recruitment rates from the count data are estimated for each population and for each harvest category and compared to the maximum recruitment rates.
+Even if both vital rates vary in a likely range following the validation methodology previously described, it does not prove that the interannual variability is well tracked.
+In order to validate this aspect, we compared the outputs of the count-based method, i.e.
+the immature proportion and both vital rates, to outputs based on data from hunting bags.
+Only data from years with more than 100 individuals harvested in winter were selected, which corresponded to a time series covering 5 years of the GB count time series.
+The presence of the bursa of Fabricius enables immature to be identified [@Hochbaum1942].
+The immature proportion in hunting bags is then straightforward, see Equation \@ref(eq:pis).
+The adult survival and the recruitment rate are then estimated by combining this proportion and Equation \@ref(eq:vr).
+We discuss the ability of the count method to track the interannual variability by comparing the outputs of the two methods.
 
-For all the sub-models in Section \@ref(MM), the Bayesian framework was used for its efficiency and simplicity to propagate error through the parameters.
-One used uninformative priors on all parameters.
-One generated three chains of length 500000, with a thinning of 10 to avoid autocorrelation in the samples, and one discarded the first 2000 samples as burn-in.
+|Name|Class|Description|
+|:---|:----|:----------------------------|
+|$SI_{i, t}$|Data|Number of immatures sampled in the wintering population $i$ in year $t$|
+|$S_{i, t}$|Data|Number of individuals sampled in the wintering population $i$ in year $t$|
+
+\begin{equation}
+SI_{i, t} \sim {\sf Binom}(p(i)_{i, t}, S_{i, t})
+(\#eq:pis)
+\end{equation}
+
+## Impact of harvest strategies on population dynamics
+
+Harvest strategies differ from GB to France.
+In GB, harvest occurred in winter, i.e.
+before the breeding period, whereas it is done mostly in summer in France, i.e.
+during and after the breeding period.
+To assess the impact of the two strategies on the population dynamics and conclude on which strategy is the most suited to reach the management objectives, we compared vital rates with and without harvest in light of the underlying population growth rates.
+
+For all the sub-models in Section \@ref(VAL), the Bayesian framework was used for its efficiency and simplicity to propagate error through the parameters.
+We used uninformative priors on all parameters.
+We generated three chains of length 500000, with a thinning of 10 to avoid autocorrelation in the samples, and we discarded the first 2000 samples as burn-in.
 Chain convergence was assessed using the Gelman and Rubin convergence diagnostic (R<1.1, @Gelman1992).
-One fit the models using NIMBLE [@Valpine2017] run from R [@RCT2020].
-Data and code are available here: https://github.com/adri-tab/Ruddy_duck_recruitment_rate.
+We fit the models using NIMBLE [@Valpine2017] run from R [@RCT2022].
+Data and code are available here: https://github.com/adri-tab/Ruddy_duck_vital_rates.
 The values **X[Y; Z]** reported in results are the medians and the associated boundaries of the 95% confidence interval of posterior distributions.
@@
 # Results
+
+Time series derived from these counts show that the two populations display comparable growth (Figure \@ref(fig:count)).
+They grew freely in both countries until 1999.
+Although shooting was initiated in 1999 in both countries, both populations begun declining only after much higher harvest efforts had been achieved in 2005 and 2018 respectively (figure 3).
+This history underlines that both populations observed a large spectrum of harvest effort, from no pressure to very high pressure.
+For sake of interpretation, we ascribed the time-period 2001-2004 to the no harvest category in France because the harvest pressure was negligible hand hence had no effect on the population growth.
+
+```{r count, fig.cap = "(ref:count)"}
+knitr::include_graphics("../Output/plot_pres_2.png")
+```
+
+Also, covering contrasted populations dynamics corresponding to different levels of harvesting pressure (Figure 2.3).
+In France, data before 2004 correspond to a period of steady population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure (Figure 2.3).
+In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period.
+In parallel, the sampling dataset, which corresponds to individuals shot in winter during control operations, covers 9 years only in UK, with 5 years in common with the corresponding counting time series.
+Estimating recruitment rate from the evolution of apparent sex-ratio in winter
+
+Time series show that the two populations follow similar demographic histories (Figure \@ref(fig:count)).
+They grew freely in both countries until 1999.
+Control measures have been then applied from 1999 to 2005 in the UK and from 1999 to 2018 in France.
+In both countries, these management measures led to stop the expansion of the ruddy ducks.
+From 2005 in the UK and from 2018 in France, higher control effort led to a rapid depletion in both populations.
+For sake of interpretation, the time period 2001-2004 was classified in the *no harvest* category in France because the harvest pressure applied over this period was very limited with no effect on the population growth.
+
+```{r count, fig.cap = "(ref:count)"}
+knitr::include_graphics("../Output/plot_2.png")
+```
+
+The counting dataset, which differentiates the male-like from the female-like individuals in winter, covers 7 years and 19 years in UK and France respectively (Figure \@ref(fig:count)).
+Combined together, the counting datasets cover different population dynamics corresponding to a gradient of harvest pressure.
+In France, data before 2004 correspond to a period of population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure (Figure \@ref(fig:count)).
+In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period.
+In parallel, the sampling dataset, which corresponds to individuals shot in winter during control operations, covers 9 years only in UK, with 5 years in common with the corresponding counting time series.
 
@@
 # Discussion
+
+Parler que le model peut etre amélioré en déclarant une proportion de mal à la repro dynamique qui prend en compte les années de fort recrutement vs les faibles recrutements.
+
+Dire qu'il semble y avoir une meilleure detection en hiver au vu de la mortalité adulte entre UK & France
 
@@
 
-(ref:flock) Typical observation of a ruddy duck flock in winter: 4 male-like individuals with a white cheek and a black cap, 10 female-like individuals with a whitish cheek and a dark stripe across it, 3 unidentified individuals © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
+(ref:flock) Typical observation of a ruddy duck flock in winter: 10 female-like individuals with a whitish cheek and a dark stripe across it, 4 male-like individuals with a white cheek and a black cap, 3 unidentified individuals © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
 
-(ref:pmal) Apparent male proportion in winter counts vs male proportion in samples; for the first category, one point corresponds to the proportion of male-like plumage in a population counted in winter; for the second category, the male proportion is estimated from all ruddy ducks sampled over one year
+(ref:pfem) Apparent female proportion in winter counts vs female proportion in hunting bags; for the first category, one point corresponds to the proportion of female-like plumage in a population counted in winter; for the second category, the female proportion is estimated from all ruddy ducks controlled over one year in a single population
 

### 2022-10-12 PCItemplate-main/manuscript.Rmd
Figures update and revision Intro + M&M

---
title: "An RMarkdown template for a preprint recommended by one of the Peer Communities In"
which_PCI: ecology
# which_PCI MUST BE one of:
# ecology
# evolbiol: Evolutionary Biology
# paleo: Palaeontology
# animsci: Animal Science
# genomics
# archaeo: Archaeology
# zool: Zoology
# mcb: Mathematical & Computational Biology
# fws: Forest & Wood Science
open_data: true #set to false if no open data in article
open_code: true #set to false if no open code in article
author:
  - firstname: Adrien 
    lastname: Tableau
    affiliation: 1
  - firstname: Alain
    lastname: Caizergues
    affiliation: 1,2,a
lastauthor: 
  - firstname: Carol #add last author here if more than one author for proper styling
    lastname: Senior #can be kept empty/deleted otherwise
    affiliation: 1,a
address:
  - code: 1
    address: Some Institute of Technology, Department, Street, City, State, Zip
  - code: 2
    address: Another University Department, Street, City, State, Zip
authorship-notes:
  # - code: a  #delete or keep empty if not needed
    # note: These authors contributed equally to this work. #delete or keep empty if not needed
email: email\@example.com #the "\" before the "@" is needed to escape and compile properly!
abstract: |
  Write your abstract here in a single paragraph, no line breaks. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract.
keywords:
  - one
  - two
  - optional
  - optional
  - optional

recommender: Their name here

reviewers: Their name here if known, and X anonymous reviewers

date: 01^st^ October 2020
pubyear: "`r format(Sys.time(), '%Y')`"
preprintserver: bioRxiv
preprintversion: X
preprintnumber: xxxx
preprintDOI: xxxxxxxxxxx #important: "naked" doi, do NOT include the https://doi.org/ prefix here
recommendationDOI: yyyyyyy #important: "naked" doi, do NOT include the https://doi.org/ prefix here

bibliography: bibliography/sample.bib #replace with path (relative or absolute) to your own
#needs to be a .bib file, BibTeX/BibLaTeX and their Better BibTeX equivalents all work,
#but see README for details
#inline R code that outputs an unquoted path can be used (see pubyear above for how to include r code in YAML header)
caption-side: bottom #where are figure captions relative to figure? can be top or bottom
###############################################################
#### DO NOT CHANGE these settings unless you know what you're doing
#### results not guaranteed if altered
biblio-style: authoryear
output:  
  bookdown::pdf_document2:
#can be pdf_document instead of bookdown::pdf_document2
#but that would disable figure and table cross-referencing
#see https://bookdown.org/yihui/rmarkdown-cookbook/cross-ref.html
    fig_caption: true
    template: template/main.tex
    citation_package: biblatex

---
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)
options(knitr.table.format = "latex")
#document-wide settings: 
# load required packages
# unless explicitly indicated, do not print code alongside with code chunks in document
#### end of "DO NOT CHANGE" block
################################
```

# Introduction {-}

 Note the use of {-} in headers in the code to suppress automatic header numbering. Do not forget it when you add your own custom headers and subheaders!

This is an example of text. This is an example of citation [@levins1971regional]. If you want to format your citations with authors outside parentheses, you can too: "@levins1971regional once wrote something important".

And now, some filler text. Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

In their study, @preprint combine mathematical analysis, Individual-Based simulations and experiments to study expansions. At vero eos et accusamus et iusto odio dignissimos ducimus qui blanditiis praesentium voluptatum deleniti atque corrupti quos dolores et quas molestias excepturi sint occaecati cupiditate non provident, similique sunt in culpa qui officia deserunt mollitia animi, id est laborum et dolorum fuga. Et harum quidem rerum facilis est et expedita distinctio. Nam libero tempore, cum soluta nobis est eligendi optio cumque nihil impedit quo minus id quod maxime placeat facere possimus, omnis voluptas assumenda est, omnis dolor repellendus. Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae. Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat

Lists can be included following standard Markdown syntax:

- this is a thing

- this is another thing

- one last thing

List elements don't need to be separated by empty lines:

- a thing
- another thing

Lists can be numbered:

1. a thing

2. a second thing

# Methods {-}

This is filler text. At vero eos et accusamus et iusto odio dignissimos ducimus qui blanditiis praesentium voluptatum deleniti atque corrupti quos dolores et quas molestias excepturi sint occaecati cupiditate non provident, similique sunt in culpa qui officia deserunt mollitia animi, id est laborum et dolorum fuga. Et harum quidem rerum facilis est et expedita distinctio. Nam libero tempore, cum soluta nobis est eligendi optio cumque nihil impedit quo minus id quod maxime placeat facere possimus, omnis voluptas assumenda est, omnis dolor repellendus. Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae. Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat. [@shigesada1997biological].

## Statistical analyses {-}

See how to create headers and various levels of subheaders with `#`.

Markdown documents can use LaTeX syntax to display equations, either inline: $y_{i} = x$, or as their own equation block:

$$y_{i} = x$$

# Results {-}

```{r fig1, fig.cap="A good figure, generated directly from code within the RMarkdown document", out.width="10cm"}
x <- 1:10
y <- 11:20
plot(x, y)
```

This is your Results part. Results are sometimes displayed in tables. Tables can be included in several ways. **The preferred method for this template** is through the use of the `kable` package and its extensions, like `kableExtra` for restyling. Other methods, like inputting a Markdown table directly in the text, have been known to lead to pagination bugs in some cases, that have not been ironed out yet.

```{r table-kable}
knitr::kable(head(mtcars[, 1:4]), caption= "This is a table", 
             booktabs = T)
#the format="latex" argument is REQUIRED for proper rendering in final pdf
# but it is set globally at the start of the document,
# so no need to give it in each table
```

Note in the code where the caption is, and that tables are automatically numbered in the order of their introduction (independently of the chunk label).

Tables and figures can be cross-referenced using the `\@ref(a:b)` syntax, where *a* is the object type ("tab" for tables, "fig" for figures, "eqn" for equations) and *b* the label of the corresponding chunk. For instance, see Table \@ref(tab:table-kable).

```{r table-kable2}
knitr::kable(head(mtcars[, 5:8]), caption= "This is also a table, but with vertical lines kept")
```

One can also easily add figures, either directly generated with code in-file (like Fig.\@ref(fig:fig1)), or imported (like Figs.\@ref(fig:fig2)-\@ref(fig:fig3)). There are several ways to import figures in RMarkdown; the preferred way in this template is to use `include_figures()` within a chunk. This makes in particular figure scaling much easier, and allows you to import pdf files as figures.

Figures will place themselves in order of their introduction in the code, but where there is room, so not necessarily where you placed them *relative to text*. Play with structure to obtain the effect you want.

```{r fig2, fig.cap="A great figure", out.width="5cm"}
include_graphics("resources/Fig2.png")
```

```{r fig3, out.width="10cm", fig.cap="A figure, but imported for a pdf source. Advantages: never pixelated, even when zooming in; also, text is selectable"}
knitr::include_graphics("resources/Fig3.pdf")
```

If you want to display the actual code for a chunk (may be useful for some methods paper), you can:

```{r test, echo = TRUE, eval = FALSE}
a <- 2
```

(set the `echo` argument of the chunk to `TRUE` to override the default and make the code visible, set the `eval` argument to `FALSE` if this is code you want to show, not run. See general Rmarkdown help pages for more on chunk arguments)

# Discussion {-}

Some filler text again. At vero eos et accusamus et iusto odio dignissimos ducimus qui blanditiis praesentium voluptatum deleniti atque corrupti quos dolores et quas molestias excepturi sint occaecati cupiditate non provident, similique sunt in culpa qui officia deserunt mollitia animi, id est laborum et dolorum fuga. Et harum quidem rerum facilis est et expedita distinctio. Nam libero tempore, cum soluta nobis est eligendi optio cumque nihil impedit quo minus id quod maxime placeat facere possimus, omnis voluptas assumenda est, omnis dolor repellendus. Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae. Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat. [@crooks2004spatial]

Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo. Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt. Neque porro quisquam est, qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat voluptatem. Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur? Quis autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur, vel illum qui dolorem eum fugiat quo voluptas nulla pariatur?[@shigesada1997biological].

# Data accessibility {-}

Data are available online: link and/or DOI of the webpage hosting the data

# Supplementary material {-}

Script and codes are available online: link and/or DOI of the webpage hosting the script and codes

# Acknowledgements {-}

This is where you place your Acknowledgements (may include funding sources; or that can be in its separate section).

# Conflicts of interest disclosure {-}

The authors of this preprint declare that they have no financial conflict of interest with the content of this article. XXX and XXX are recommenders for PCI XXX

<!--a little bit of LaTeX verbatim to place the references here, before the (potential) appendices:-->
\printbibliography[notcategory=ignore]

<!--delete or comment out everything after this if you do not use appendices-->
# Appendix {-}

This is your appendix 1 or the link to appendix 1  
This is your appendix 2 or the link to appendix 2

### 2022-10-12 Code/Model_vital_rates.R

@@ -264,7 +264,7 @@ frag %>%
   rowid_to_column() -> ds
 
 count_sex_app %>% 
-  mutate(method = "in visual detection",
+  mutate(method = "in counts",
          tot = obs_type_mal + obs_type_fem,
          `female proportion` = obs_type_fem / tot) %>% 
   select(method, `female proportion`, tot) %>% 
@@ -290,11 +290,6 @@ count_sex_app %>%
   xlab("") + 
   labs(title = "Female proportion", caption = "data from UK & French Ruddy duck populations") +
   coord_flip() -> intro_plot; intro_plot
-
-# ggsave(filename = "P:\\LIFE_Presentations\\2022-09-15_duck symposium\\prop_fem.png", 
-#        plot = intro_plot,
-#        width = 8, 
-#        height = 2, dpi = 600)
 
 # Lambda formatting ------------------------------------------------------------
 
@@ -1284,7 +1279,7 @@ raw_ds2 %>%
 
 raw_ds2 %>% 
   filter(var == "survival") %>% 
-  ggplot(aes(x = year, y = `50%`, color = pop, shape = label)) +
+  ggplot(aes(x = year, y = `50%`, color = pop)) +
   geom_pointrange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5, size = 1.1) +
   geom_hline(yintercept = 1, alpha = 0.1, size = 1.5) +
   scale_color_manual(values = c_pop) +
@@ -1297,9 +1292,7 @@ raw_ds2 %>%
     limits = c(0, 1.15),
     breaks = 0:11 / 10,
     minor_breaks = NULL) +
-  guides(x = guide_axis(angle = 45), 
-         color = guide_legend(title = "", order = 1),
-         shape = guide_legend(title = "", order = 2)) +
+  guides(x = guide_axis(angle = 45)) +
   labs(title = "Apparent survival rate", 
        subtitle = "adult survival & potential adult migration", 
        x = NULL, y = NULL) +
@@ -1307,7 +1300,7 @@ raw_ds2 %>%
 
 raw_ds2 %>% 
   filter(var == "productivity") %>% 
-  ggplot(aes(x = year, y = `50%`, color = pop, shape = label)) +
+  ggplot(aes(x = year, y = `50%`, color = pop)) +
   geom_pointrange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5, size = 1.1) +
[…]

### 2022-10-19 Code/Draft_ruddy_duck_vital_rates.Rmd
graphs for article

@@
 @Niel2005].
-It is also, crucial for measuring the efficacy of management actions eventually undertaken to help population recovery, control or eradication [@Shea1998].
-Among the approaches available to managers for reaching these goals, those relying on the monitoring of the numbers of individuals are by far the most popular [@Rintala2022].
+It is also crucial for measuring the efficacy of management actions eventually undertaken to help population recovery, control, or eradication [@Shea1998].
+Among the approaches available to managers for reaching these goals, those relying on the monitoring of the number of individuals are by far the most popular [@Rintala2022].
 In many instances, these methods even enable environmental factors underlying changes in population size to be investigated and hence some mitigating actions to be implemented [@Faillettaz2019].
 Unfortunately, relying on the monitoring of numbers of individuals alone hinder basic understanding of the demographic mechanisms underlying changes in population growth rates.
-More specifically, based on counts only, it is nearly impossible to assess to what extent population growth rate is affected by change survival rather than to variations in reproductive parameters [@Austin2000].
+More specifically, based on counts only, it is nearly impossible to assess to what extent population growth rate is affected by change in survival rather than to variations in reproductive parameters [@Austin2000].
 To assess the relative sensitivity of population growth rate to factors affecting survival or reproductive parameters, demographers most often rely on the monitoring of individuals by capture-mark-recapture [@Lebreton1992].
-In practice however, one cannot always rely on the capture and release of individuals owing to the legal status of the species that preclude any additional disturbance (critically endangered species), or forbid the release of alive individuals (invasive species or pests).
+In practice however, one cannot always rely on the capture and release of individuals owing to the legal status of the species that precludes any additional disturbance, e.g.
+critically endangered species, or forbid the release of alive individuals, e.g.
+invasive species or pests.
 Indeed, although highly efficient for assessing demographic parameters, CMR methods have potential drawbacks such as being invasive and hardly affordable when time and money are limiting.
@@
 
-Most of the time therefore, managers do the best of a bad job using counts as the only viable option for assessing population growth rates and the efficacy of management actions.
+Most of the time therefore, managers do the best of a bad job using counts as the only viable option for tracking population growth rates and assessing the efficiency of management actions.
 Counting the number of immatures produced in addition to alive adults may enable adult survival and recruitment rate to be estimated, and hence provides knowledge on which of adult survival and productivity has the most influence on population growth rate.
@@
 The ruddy duck is a stiff-tailed duck native from the Americas.
-From seven individuals initially acclimated in the Slimbridge Wetland Centre in the Great Britain in the 40’s [@GutierrezExposito2020], a feral population began to establish with the first observed reproductive attempts in the wild in the 60s.
+From seven individuals initially acclimated in the Slimbridge Wetland Centre in the Great Britain in the 40’s [@GutierrezExposito2020], a feral population began to establish with the first observed reproductive attempts in the wild in the 60s (Figure \@ref(fig:count)).
 This feral population rapidly grew and spread to the entire country to reach more than 5000 individuals in the early 2000s.
-By the end of the 80s, a feral population began to establish also on the continent in France, supposedly owing to the arrival of individuals born in Great Britain (GB).
-Contrary to what was observed in GB, the French Ruddy duck population did not spread much, with the vast majority of observations and breeding attempts clustering in Western France.
+By the end of the 80s, a feral population began to establish also on the continent in France, supposedly owing to the arrival of individuals born in Great Britain.
+Contrary to what was observed in Great Britain, the French Ruddy duck population did not spread much, with the vast majority of observations and breeding attempts clustering in Western France.
 In France, during winter, almost none Ruddy duck is observed outside the lake of Grand Lieu (47.09°N, 1.67°W), which is thus the exclusive wintering site and greatly facilitates the monitoring of this population.
+The feral European populations of Ruddy duck are considered as a major threat to the endangered native White-headed duck population of the South-Western Mediterranean because of hybridization risks and thus an elevated potential for genetic pollution and eventually genetic assimilation of the latter by the former [@MunozFuentes2007].
+In order to mitigate the risks of genetic pollution of the White-headed duck by Ruddy ducks, a European plan of eradication of Ruddy ducks has been adopted by the European Commission.
+As a result, eradication measures were taken both in Great Britain and in France from the late 90s [@GutierrezExposito2020], (Figure \@ref(fig:count)).
+The release of non-native species in the wild, including Ruddy ducks, is forbidden in countries of the European Union, and the status of White-headed duck populations is highly unfavourable.
+The use of capture-mark-recapture to monitor populations of these species is thus impossible.
+As a result, the effectiveness of the eradication of Ruddy ducks and management actions intended to favour the recovery of White-headed ducks were essentially checked through censuses.
+Censuses, however, prevented from assessing the relative effects or variations in survival and reproductive success/recruitment on population growth rate, which is a key towards a proper understanding of the efficiency of management actions.
 
-The feral European populations of Ruddy duck are considered as a major threat to the endangered native White-headed duck population of the South-Western Mediterranean because of hybridization risks and thus an elevated potential for genetic pollution and eventually genetic assimilation of the latter by the former [@MunozFuentes2007].
-In order to mitigate the risks of genetic pollution of the White-headed duck by Ruddy ducks a European plan of eradication of Ruddy ducks has been adopted by the European Commission.
-As result, eradication measures were taken both in France and Great Britain in the late 90s [@GutierrezExposito2020].
-
-The release non-native species in the wild, including Ruddy ducks, is forbidden in countries of the European Union, and the status of White-headed duck populations is highly unfavourable.
-The use of capture-mark-recapture to monitor populations of these species was thus impossible.
-As a result, the effectiveness of the eradication of Ruddy ducks and management actions intended to favour the  recovery of White-headed ducks were essentially checked through censuses.
-Censuses, however, prevented from assessing the relative effects or variations in survival and reproductive success/recruitment on population growth rate, which is a key towards a proper understanding of the efficacy of management actions.
+```{r count, fig.cap = "(ref:count)"}
+knitr::include_graphics("../Output/plot_pres_2.png")
+```
 
@@
 The delayed dichromatism explains why the apparent proportion of females in the winter counts is always higher than the true proportion of females that can be observed in winter hunting bags (Figure \@ref(fig:pfem)).
-Thus, by monitoring the seasonal evolution of apparent sex-ratio, it is possible to assess the proportion of immature males within a population, which, combined with reliable counts and under some reasonable assumptions (e.g., even sex-ratio at birth, comparable mortality rates between females and males among juveniles, and unchanging adult sex-ratio over the monitoring period), enables both adult survival rate and recruitment rate to be estimated.
+Thus, it is possible to assess the proportion of immatures within a population by monitoring the seasonal evolution of apparent sex ratio and assuming an even sex ratio at birth, comparable mortality rates between females and males among juveniles, and unchanging adult sex ratio over the monitoring period.
+Then, both adult survival rate and recruitment rate can be estimated by combining the age ratio thus obtained with reliable counts over two successive winters.
 
@@
 We monitored the demography of Ruddy duck populations in both Great-Britain and France thanks to exhaustive counts on the wintering grounds during the period between December 1 and January 31.
-In addition to these exhaustive counts, censuses that distinguished female-like from male-like individuals were performed on the same period from 2007 to 2012 in UK, and in 1999, 2001-2009, 2012-2019 in France.
+In addition to these exhaustive counts, censuses that distinguished female-like from male-like individuals were performed on the same period from 2007 to 2012 in Great Britain, which corresponded to a strong decline of the population.
+In France, the monitoring spread over 1999, 2001-2009, and 2012-2019, which corresponded to a stabilized population size excepted over 2001-2004, for which the population grew rapidly without control effort.
 During these years, the eradication effort, under control of the authorities in both countries, was also scrupulously recorded.
@@
 |$N_{i, t}$|Data|Size of the population $i$ in year $t$ (maximum number of individuals counted in the wintering population)|
-|$s_{i, t}$|Parameter|Adult survival rate, i.e. proportion of adults in year $t-1$ still alive in year $t$|
-|$r_{i, t}$|Parameter|Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per adult in year $t - 1$|
+|$s_{i, t}$|Parameter|Adult survival rate, i.e. proportion of individuals in year $t-1$ still alive in year $t$|
+|$r_{i, t}$|Parameter|Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per individual in year $t - 1$|
 |$S_{i, t}$|Parameter|Number of adults in population $i$ in year $t$|
@@
 \begin{equation}
-s_{i, t} = \frac{(1 - p(i)_{i, t}).N_{i, t}}{N_{i, t - 1}}
-r_{i, t} = \frac{p(i)_{i, t}.N_{i, t}}{N_{i, t - 1}}
+\begin{aligned}
+s_{i, t} & = \frac{N_{i, t}}{N_{i, t - 1}}.(1 - p(i)_{i, t}) \\
+r_{i, t} & = \frac{N_{i, t}}{N_{i, t - 1}}.p(i)_{i, t}
+\end{aligned}
 (\#eq:vr)
@@
 \begin{equation}
-S_{i, t} = (1 - p(i)_{i, t}).N_{i, t}
-R_{i, t} = p(i)_{i, t}.N_{i, t}
+\begin{aligned}
+S_{i, t} & = N_{i, t}.(1 - p(i)_{i, t}) \\
+R_{i, t} & = N_{i, t}.p(i)_{i, t}
+\end{aligned}
 (\#eq:VR)
@@
 
-## Validation of the count method {#VAL}
+## Validation of the count-based method {#VAL}
 
@@
 If the recruitment rate outputs of the model are significantly superior to the maximum recruitment rate, it means that the method overestimates the recruitment rate, which reflects that a part of adult males are not detected during count surveys.
-A null or negative recruitment rate would mean that immature males moulted before the survey or/and the assumption on the stability of the adult sex ratio is too strong.
+A null or negative recruitment rate would mean that immature males moulted before the survey or/and the assumption on the stability of the adult sex ratio is overrated.
 
 The maximum recruitment rate is deduced by using its relationship to the maximum growth rate and the maximum survival, see Equation \@ref(eq:GR).
-This relationship becomes more complex if a species with delayed maturity is considered, see @Robertson2008.
 This relationship comes from a simple reasoning for a closed population: the population size in year $t$ is equal to the number of adults that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
@@
 For an open population, adult survival and recruitment rate are confounded to adult and recruit migration respectively.
+This relationship becomes more complex if a species with delayed maturity is considered, see @Robertson2008.
 
@@
 |$R_{t}$|Number of recruits in year $t$|
-|$s_{t}$|Adult survival rate, i.e. proportion of adults in year $t-1$ still alive in year $t$|
-|$r_{t}$|Recruitment rate, i.e. number of recruits in year $t$ produced per adult in year $t-1$|
+|$s_{t}$|Adult survival rate, i.e. proportion of individuals in year $t-1$ still alive in year $t$|
+|$r_{t}$|Recruitment rate, i.e. number of recruits in year $t$ produced per individual in year $t-1$|
 |$\lambda_{t}$|Growth rate of the population between year $t-1$ and year $t$|
@@
 
-The maximum growth rate occurred for both populations during their expansion phase, when the evolution of the population size is the steepest.
+The maximum growth rate occurred for both populations during their expansion phase, when the evolution of the population size was the steepest.
 To estimate robust maximum population growth rates for both populations, we smoothed annual population growth rate over a consistent time period by using a linear regression on the logarithm scale, see Equation \@ref(eq:L).
@@
 
-Practically, we used a very conservative prior on the maximum adult survival by setting a uniform distribution on $[0.7; 0.9]$ [the lower limit corresponds to the upper range of survival rates observed in literature for species of similar weight, the upper limit corresponds to the range of survival rates of long-life species, @Nichols1997; @Krementz1997; @Buxton2004].
-For the maximum growth rate, we rejected GB data before 1972 since the size estimate of this population is noisy below 50 individuals (Figure \@ref(fig:grts)).
-After reaching the threshold of 1000 individuals, the GB population growth observed a strong inflexion whereas no harvest pressure was applied.
-This observation suggests that a negative density-dependence process might occur beyond 1000 individuals and led to consider only the first sequence of the time series to infer on the maximum growth rate in GB.
-For the French population, the two sequences with birds and without harvest were used, i.e.
-1994-1999 and 2001-2004 (Figure \@ref(fig:grts)).
+Practically, we used a very conservative prior on the maximum adult survival by setting a uniform distribution on $[0.7; 0.9]$.
+The lower limit corresponds to the upper range of survival rates observed in literature for species of similar weight, the upper limit corresponds to the range of survival rates of long-life species [@Nichols1997; @Krementz1997; @Buxton2004].
+For the maximum growth rate, we rejected Great Britain data before 1972 since the size estimate of this population is noisy below 50 individuals (Figure \@ref(fig:count)).
+After reaching the threshold of 1000 individuals, the Great Britain population growth observed a strong inflexion whereas no harvest pressure was applied.
+This observation suggests that a negative density-dependence process might occur beyond 1000 individuals and led to consider only the first sequence of the time series to infer on the maximum growth rate in Great Britain.
+For the French population, the two sequences without harvest were used, i.e.
+1994-1999 and 2001-2004.
 
-```{r grts, fig.cap = "(ref:grts)"}
-knitr::include_graphics("../Output/plot_pres_1.png")
-```
-
-Even if both vital rates vary in a likely range following the validation methodology previously described, it does not prove that the interannual variability is well tracked.
+Even if both vital rates vary in a likely range following the validation methodology previously described, it does not prove that the interannual variability is properly tracked.
 In order to validate this aspect, we compared the outputs of the count-based method, i.e.
 the immature proportion and both vital rates, to outputs based on data from hunting bags.
-Only data from years with more than 100 individuals harvested in winter were selected, which corresponded to a time series covering 5 years of the GB count time series.
+Only data from years with more than 100 individuals harvested in winter were selected.
+It covers 5 years of the count time series of the Great Britain population.
 The presence of the bursa of Fabricius enables immature to be identified [@Hochbaum1942].
-The immature proportion in hunting bags is then straightforward, see Equation \@ref(eq:pis).
-The adult survival and the recruitment rate are then estimated by combining this proportion and Equation \@ref(eq:vr).
-We discuss the ability of the count method to track the interannual variability by comparing the outputs of the two methods.
+The estimation of the immature proportion in hunting bags is then straightforward, see Equation \@ref(eq:pis).
+The adult survival and the recruitment rate were then estimated by combining this proportion and Equation \@ref(eq:vr).
+We discussed the ability of the count-based method to accurately track the interannual variability by comparing the outputs of the two methods.
 
@@
 
-Harvest strategies differ from GB to France.
-In GB, harvest occurred in winter, i.e.
-before the breeding period, whereas it is done mostly in summer in France, i.e.
+Harvest strategies differed from Great Britain to France.
+In Great Britain, harvest occurred in winter, i.e.
+before the breeding period, whereas it was done mostly in summer in France, i.e.
 during and after the breeding period.
-To assess the impact of the two strategies on the population dynamics and conclude on which strategy is the most suited to reach the management objectives, we compared vital rates with and without harvest in light of the underlying population growth rates.
+To identify the impact of the two strategies on the population dynamics, we assessed how vital rates responded to the harvest by considering both average drops and variability changes.
+To complete our understanding on why the Great Britain strategy was more performing on controlling Ruddy ducks, we compared vital rates with and without harvest in light of the underlying average population growth rates.
 
-For all the sub-models in Section \@ref(VAL), the Bayesian framework was used for its efficiency and simplicity to propagate error through the parameters.
+For all the sub-models in Section \@ref(MM), the Bayesian framework was used for its efficiency and simplicity to propagate error through the parameters.
 We used uninformative priors on all parameters.
@@
 
-Time series derived from these counts show that the two populations display comparable growth (Figure \@ref(fig:count)).
+Higher mortality on adult in GB -> better detection in winter;
+
+Time series derived from count data show that the two populations display comparable initial growth (Figure \@ref(fig:count2)).
 They grew freely in both countries until 1999.
@@
 For sake of interpretation, we ascribed the time-period 2001-2004 to the no harvest category in France because the harvest pressure was negligible hand hence had no effect on the population growth.
-
-```{r count, fig.cap = "(ref:count)"}
-knitr::include_graphics("../Output/plot_pres_2.png")
-```
 
@@
 For sake of interpretation, the time period 2001-2004 was classified in the *no harvest* category in France because the harvest pressure applied over this period was very limited with no effect on the population growth.
-
-```{r count, fig.cap = "(ref:count)"}
-knitr::include_graphics("../Output/plot_2.png")
-```
 
@@
 
-(ref:flock) Typical observation of a ruddy duck flock in winter: 10 female-like individuals with a whitish cheek and a dark stripe across it, 4 male-like individuals with a white cheek and a black cap, 3 unidentified individuals © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
+(ref:count) Evolution of the size of the two Ruddy duck populations; y-axis is on a log-scale
 
-(ref:pfem) Apparent female proportion in winter counts vs female proportion in hunting bags; for the first category, one point corresponds to the proportion of female-like plumage in a population counted in winter; for the second category, the female proportion is estimated from all ruddy ducks controlled over one year in a single population
+(ref:flock) Typical observation of a Ruddy duck flock in winter: 10 female-like individuals with a whitish cheek and a dark stripe across it, 4 male-like individuals with a white cheek and a black cap, 3 unidentified individuals © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
 
-(ref:count) Evolution of the size of the two ruddy duck populations in light of the harvest pressure and the data time series
+(ref:pfem) Female proportion in hunting bags vs in counts; For the first category, one point corresponds to the female proportion estimated from direct sex identification of the Ruddy ducks controlled over one year in a population; for the second category, one point is the proportion of female-like plumage in a population counted in winter.
+The difference is due to immature males that look like females before moulting
 
-(ref:grts) Evolution of the size of the two ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:grts) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)
 

### 2022-10-19 Code/Model_vital_rates.R

@@ -27,16 +27,18 @@ tibble(id = 1:20) %>%
 ggplot_build(p)$data
 
 c_pop <- c("#F8766D", "#619CFF")
-c_met <- c("#00BF7D", "#D89000")
+c_met <- c("#a1d99b", "#41ab5d")
 
 # Plot functions --------------------------------------------------------------------
 
-scale_y_percent <- list(
+scale_y_percent <- function() {
+  list(
   scale_y_continuous(
   labels = scales::percent_format(accuracy = 1L),
   limits = c(0, 1),
   breaks = 0:10 / 10, 
-  minor_breaks = NULL))
+  minor_breaks = NULL)
+  )}
 
 scale_x_date_own <- function(coef = 5e-2) {
   list(
@@ -69,7 +71,7 @@ raw_plot <- function(para,
     scale_color_manual(values = c_met) +
     labs(title = title, x = NULL, y = NULL) -> tmp
   
-  if (percent == TRUE) {tmp + scale_y_percent} else {tmp}
+  if (percent == TRUE) {tmp + scale_y_percent()} else {tmp}
 }
 
 cor_plot <- function(col, title = NULL, limits = NA, breaks = NA, one = 1) {
@@ -108,7 +110,7 @@ cor_plot <- function(col, title = NULL, limits = NA, breaks = NA, one = 1) {
     geom_errorbar(aes(ymin = !!sym(col_c[1]), ymax = !!sym(col_c[3]), color = year), 
                   width = 0, alpha = .5, size = 1.2) +
     labs(title = title, 
-         y = "counting method", x = "sampling method") +
+         y = "count-based method", x = "hunting-bag-based method") +
     scale_x_continuous(expand = expansion(mult = c(0, 0)), 
                        limits = if (any(is.na(limits))) {
                          c(0, 1.1 * max(unlist(tmp[-1])))
@@ -128,10 +130,12 @@ cor_plot <- function(col, title = NULL, limits = NA, breaks = NA, one = 1) {
 # Data import -----------------------------------------------------------------------
 
 read_rds("./Data/Ruddy_duck_data.rds") %>% pluck(1) %>% 
-  mutate(across(pop, as_factor)) %>% arrange(pop) -> frag
+  mutate(pop = if_else(pop == "UK", "GB", pop),
+         across(pop, as_factor)) %>% arrange(pop) -> frag
 
 read_rds("./Data/Ruddy_duck_data.rds") %>% pluck(2) %>% 
[…]

### 2022-10-19 Code/Model_vital_rates.R
average graph comparison update

@@ -964,16 +964,16 @@ JuvOut4 %>%
 ########### j'en suis la ##############
 #######################################
 
-
-
 JuvOut4 %>% 
+  mutate(par = if_else(par == "lambda" & str_detect(method, "small"), "lambda_avg", par),
+         pop = if_else(par == "lambda_avg", method %>% str_sub(1, 2) %>% as_factor(), pop)) %>% 
   filter(par %>% str_ends("_avg")) %>% 
   mutate(box = case_when(
-    method == "proxy" ~ "maximum",
+    method %>% str_detect("proxy|small") ~ "maximum",
     method %>% str_detect(".no|samp") ~ "no harvest",
     TRUE ~ "harvest") %>% 
       as_factor() %>% 
-      fct_relevel(c("maximum", "no harvest", "harvest")), 
+      fct_relevel(c("no harvest", "harvest", "maximum")), 
     across(contains("%"), ~ if_else(method %>% str_detect("sampl"), NA_real_, .x)),
     txt = if_else(str_detect(method, "sampling"), "NO DATA", NA_character_)) %>% 
   select(-year, -method) %>% 
@@ -983,7 +983,8 @@ JuvOut4 %>%
               ggplot(aes(x = pop, y = `50%`, color = pop)) +
   facet_wrap(~ box, nrow = 1) +
   theme(strip.text.x = element_text(size = 14)) +
-  geom_pointrange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5, size = 1.2) +
+  geom_pointrange(aes(ymin = `2.5%`, ymax = `97.5%`), 
+                  alpha = 0.5, size = 1.5, fatten = 1.2) +
   geom_text(aes(x = pop, y = .43, label = txt), color = c_pop[1], angle = 90) +
   scale_color_manual(values = c_pop, guide = "none") +
   scale_y_percent() +

### 2023-02-09 Code/Draft_ruddy_duck_vital_rates.Rmd
update of all figures, ready for results

@@
 invasive species or pests.
-Indeed, although highly efficient for assessing demographic parameters, CMR methods have potential drawbacks such as being invasive and hardly affordable when time and money are limiting.
-Genetic monitoring is a non-invasive alternative to CMR approaches based on physical captures, but it suffers from being costly and time-consuming and requires quite rigorous sampling schemes in the field.
+Indeed, although highly efficient for assessing demographic parameters, capture-mark-recapture methods have potential drawbacks such as being invasive and hardly affordable when time and money are limiting.
+Genetic monitoring is a non-invasive alternative to capture-mark-recapture approaches based on physical captures, but it suffers from being costly and time-consuming and requires quite rigorous sampling schemes in the field.
 
@@
 
-The ruddy duck is a stiff-tailed duck native from the Americas.
+The Ruddy duck is a stiff-tailed duck native from the Americas.
 From seven individuals initially acclimated in the Slimbridge Wetland Centre in the Great Britain in the 40’s [@GutierrezExposito2020], a feral population began to establish with the first observed reproductive attempts in the wild in the 60s (Figure \@ref(fig:count)).
@@
 ```{r count, fig.cap = "(ref:count)"}
-knitr::include_graphics("../Output/plot_pres_2.png")
+knitr::include_graphics("../Output/plot_paper_2.png")
 ```
@@
 Immature males are thus confounded to immature and adult females before their moult (Figure \@ref(fig:flock)).
-The delayed dichromatism explains why the apparent proportion of females in the winter counts is always higher than the true proportion of females that can be observed in winter hunting bags (Figure \@ref(fig:pfem)).
+The delayed dichromatism explains why the apparent proportion of females in the winter counts is always higher than the true proportion of females that can be assessed in winter hunting bags (Figure \@ref(fig:pfem)).
 Thus, it is possible to assess the proportion of immatures within a population by monitoring the seasonal evolution of apparent sex ratio and assuming an even sex ratio at birth, comparable mortality rates between females and males among juveniles, and unchanging adult sex ratio over the monitoring period.
@@
 ```{r pfem, fig.cap = "(ref:pfem)"}
-knitr::include_graphics("../Output/plot_pres_1.png")
+knitr::include_graphics("../Output/plot_paper_1.png")
 ```
@@
 We monitored the demography of Ruddy duck populations in both Great-Britain and France thanks to exhaustive counts on the wintering grounds during the period between December 1 and January 31.
-In addition to these exhaustive counts, censuses that distinguished female-like from male-like individuals were performed on the same period from 2007 to 2012 in Great Britain, which corresponded to a strong decline of the population.
-In France, the monitoring spread over 1999, 2001-2009, and 2012-2019, which corresponded to a stabilized population size excepted over 2001-2004, for which the population grew rapidly without control effort.
-During these years, the eradication effort, under control of the authorities in both countries, was also scrupulously recorded.
+In addition to these exhaustive counts, censuses that distinguished female-like from male-like individuals (Figure \@ref(fig:count)) were performed from 2006 to 2012 in Great Britain, which corresponded to a strong decline of the population.
+In France, the monitoring spread over 1999, 2001-2009, and 2012-2019.
+This population grew rapidly during the first years of the monitoring and then stabilized from 2006, which corresponds to a period with a strong control effort.
+During these years, the eradication effort, under control of the authorities in both countries, was recorded.
+Even if the age ratio in the hunting bag was unknown before 2006 in France, we can qualify the harvest pressure on adults in three categories for both countries: "no harvest" before 1999, "low harvest" between 1999 and 2005, and "high harvest" from 2006 (see Figure \@ref(fig:sup) in Supplementary Materials).
 
@@
 Adult sex ratio is then generally biased towards males (Figure \@ref(fig:pfem), @Wood2021).
-As a consequence, deducing the immature proportion from the observed proportion of adult males is not straightforward because the adult proportion is not just twice as much as the observed proportion of adult males.
-There was no counts differentiating male-like from female-like during the breeding period to estimate properly adult sex ratio.
+As a consequence, deducing the immature proportion within a population from the observed proportion of adult males is not straightforward because the adult proportion is not just twice as much as the observed proportion of adult males.
+There were no counts differentiating male-like from female-like during the breeding period to estimate properly adult sex ratio.
 But sex identification in adults from hunting bags was available for both Ruddy duck populations.
 However, the small population size in France prevented from getting precise adult sex ratios.
-A comparison over months in UK showed no difference in male proportion, so data from adults collected over the whole year were used to estimate the male proportion.
+A comparison over months in UK showed no difference in adult male proportion, so data from adults collected over the whole year were used to estimate the male proportion.
 A comparison of male proportion among years with more than 500 samples did not show significant interannual differences.
@@
 
-The relevance of the method was assessed by testing the likelihood of the estimates of vital rates.
-For the adult survival rate, we checked if the values belonged to the expected interval $[0;1]$, and we compared the average values to literature data, since the adult survival is commonly assessed by capture mark recapture [@Lebreton2001].
-For the recruitment rate, which is rarely explicitly informed in the literature, we checked if the values were equal or below the maximum recruitment rate expected without exploitation and without negative density-dependence processes.
-If the recruitment rate outputs of the model are significantly superior to the maximum recruitment rate, it means that the method overestimates the recruitment rate, which reflects that a part of adult males are not detected during count surveys.
+The relevance of the method was assessed by testing the likelihood of the estimates of both vital rates.
+For the adult survival rate, we checked if the values belonged to the expected interval $[0;1]$, and we compared the values to literature data that are accessible because the adult survival is commonly assessed by capture-mark-recapture [@Lebreton2001].
+The maximum adult survival is defined as a uniform distribution on $[0.7; 1]$.
+The lower limit corresponds to the upper range of survival rates observed in literature for waterfowl species of similar weight, the upper limit was set to 1 because the survival rates of long-life waterfowl species are very high [@Nichols1997; @Krementz1997; @Buxton2004].
+The recruitment rate is more tricky to validate because it is not upper bounded as it is defined on $[0;\infty[$ and is neither explicitly informed in the literature.
+We then developed an approach to estimate the maximum recruitment rate expected without exploitation and without negative density-dependence processes and we checked if all values were equal or below this maximum recruitment rate.
+If the recruitment rate outputs of the model are significantly higher than the maximum recruitment rate, it means that the method overestimates the recruitment rate, which reflects that a part of adult males are not detected during count surveys.
 A null or negative recruitment rate would mean that immature males moulted before the survey or/and the assumption on the stability of the adult sex ratio is overrated.
 
-The maximum recruitment rate is deduced by using its relationship to the maximum growth rate and the maximum survival, see Equation \@ref(eq:GR).
+Practically, the maximum recruitment rate is deduced by using its relationship to the maximum growth rate and the maximum survival, see Equation \@ref(eq:GR).
 This relationship comes from a simple reasoning for a closed population: the population size in year $t$ is equal to the number of adults that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
@@
 \frac{N_{t}}{N_{t-1}} & = s_{t} + r_{t} \notag \\
-\lambda_{t} & = s_{t} + r_{t} 
+\lambda_{t} & = s_{t} + r_{t} \notag \\
+r_{t} & = \lambda_{t} - s_{t}  
 (\#eq:GR)
@@
 To estimate robust maximum population growth rates for both populations, we smoothed annual population growth rate over a consistent time period by using a linear regression on the logarithm scale, see Equation \@ref(eq:L).
+To do so, we rejected Great Britain data before 1972 since the size estimate of this population is noisy below 50 individuals (Figure \@ref(fig:count)).
+After reaching the threshold of 1000 individuals, the Great Britain population growth observed a strong inflexion whereas no harvest pressure was applied.
+This observation suggests that a negative density-dependence process might occur beyond 1000 individuals and led to consider only the first sequence of the time series to infer on the maximum growth rate in Great Britain, i.e.
+1972-1981.
+For the French population, the sequence without harvest spread over the period 1994-1999.
 
@@
 \end{equation}
-
-Practically, we used a very conservative prior on the maximum adult survival by setting a uniform distribution on $[0.7; 0.9]$.
-The lower limit corresponds to the upper range of survival rates observed in literature for species of similar weight, the upper limit corresponds to the range of survival rates of long-life species [@Nichols1997; @Krementz1997; @Buxton2004].
-For the maximum growth rate, we rejected Great Britain data before 1972 since the size estimate of this population is noisy below 50 individuals (Figure \@ref(fig:count)).
-After reaching the threshold of 1000 individuals, the Great Britain population growth observed a strong inflexion whereas no harvest pressure was applied.
-This observation suggests that a negative density-dependence process might occur beyond 1000 individuals and led to consider only the first sequence of the time series to infer on the maximum growth rate in Great Britain.
-For the French population, the two sequences without harvest were used, i.e.
-1994-1999 and 2001-2004.
 
@@
 
-Harvest strategies differed from Great Britain to France.
-In Great Britain, harvest occurred in winter, i.e.
-before the breeding period, whereas it was done mostly in summer in France, i.e.
-during and after the breeding period.
-To identify the impact of the two strategies on the population dynamics, we assessed how vital rates responded to the harvest by considering both average drops and variability changes.
-To complete our understanding on why the Great Britain strategy was more performing on controlling Ruddy ducks, we compared vital rates with and without harvest in light of the underlying average population growth rates.
+Over the "high harvest" period, i.e.
+from 2006, control strategies differed from Great Britain to France.
+In Great Britain, harvest occurred mostly in winter, i.e.
+before the breeding period (53.4% of adults were shot before the 30^th May), whereas it was done mostly in summer in France, i.e.
+during and after the breeding period (81.6% of adults were shot after the 30^th May).
+We compared the population growth of the two populations in light of the difference between the two strategies and we assessed the respective response of the vital rates by comparing the average values over the "high harvest" period to proxies of maximum vital rates estimated when both populations reached their maximum growth (see Section \@ref(VAL)).
 
@@
 We used uninformative priors on all parameters.
+As the maximum growth rate is an life history trait quite stable among populations of a same species, we used a uninformative hierarchical prior for this parameter.
 We generated three chains of length 500000, with a thinning of 10 to avoid autocorrelation in the samples, and we discarded the first 2000 samples as burn-in.
@@
 
-Higher mortality on adult in GB -> better detection in winter;
+The harvest rate on adults is considered high for both populations from 2006 (Figure \@ref(fig:sup)).
+There is no significant difference between the two populations, t-test p-value : 0.23, and the average harvest rate is 0.49.
 
-Time series derived from count data show that the two populations display comparable initial growth (Figure \@ref(fig:count2)).
+Time series derived from count data show that the two populations display comparable initial growth.
 They grew freely in both countries until 1999.
@@
 
-Also, covering contrasted populations dynamics corresponding to different levels of harvesting pressure (Figure 2.3).
-In France, data before 2004 correspond to a period of steady population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure (Figure 2.3).
+Also, covering contrasted populations dynamics corresponding to different levels of harvesting pressure.
+In France, data before 2004 correspond to a period of steady population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure .
 In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period.
@@
 
-The counting dataset, which differentiates the male-like from the female-like individuals in winter, covers 7 years and 19 years in UK and France respectively (Figure \@ref(fig:count)).
+The counting dataset, which differentiates the male-like from the female-like individuals in winter, covers 7 years and 19 years in UK and France respectively.
 Combined together, the counting datasets cover different population dynamics corresponding to a gradient of harvest pressure.
-In France, data before 2004 correspond to a period of population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure (Figure \@ref(fig:count)).
+In France, data before 2004 correspond to a period of population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure.
 In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period.
@@
 
-## Recruitment rate variability
+## Likelihood of the range of vital rates
+
+########## OLD Recruitment rate variability
 
 The male proportion in adults, which is necessary to infer on the recruitment rate from count data, is estimated at 0.60 [0.59; 0.61].
-The counting method provides estimations of the recruitment rate, see Figure \@ref(fig:rrts)A.
+The counting method provides estimations of the recruitment rate, see Figure \@ref(fig:vital).
 In UK, values vary from 0.06 [0.05; 0.07] to 0.27 [0.20; 0.34] recruits per breeder, and in France, 0.08 [0.04; 0.12] to 0.76 [0.64; 0.88] recruits per breeder.
@@
 
-Recruitment rates estimated from UK sample data are consistently higher than estimates from the corresponding count data, Figure \@ref(fig:rrts)A.
-The correlation between the two methods is strong and points out that the sampling method overestimates the recruitment rate by 44% compared to the counting method, Figure \@ref(fig:rrts)B.
+Recruitment rates estimated from UK sample data are consistently higher than estimates from the corresponding count data, Figure \@ref(fig:vital).
+The correlation between the two methods is strong and points out that the sampling method overestimates the recruitment rate by 44% compared to the counting method, Figure \@ref(fig:vital).
 Estimation precision is similar for the two methods.
@@
 
-```{r rrts, fig.cap = "(ref:rrts)"}
-knitr::include_graphics("../Output/plot_4.png")
+```{r vital, fig.cap = "(ref:vital)"}
+knitr::include_graphics("../Output/plot_paper_3.png")
 ```
 
-## Consistency of the corresponding adult survival
+## Likelihood of interannual variability in vital rates
 
-Adult survival rates were deduced from the recruitment rate estimates and the observed population growth rates, see Figure \@ref(fig:srts)A.
+########## Consistency of the corresponding adult survival
+
+Adult survival rates were deduced from the recruitment rate estimates and the observed population growth rates, see Figure \@ref(fig:cor).
 In UK, values that are obtained from the counting method ranged from 0.21 [0.20; 0.22] to 0.68 [0.65; 0.72].
@@
 
-As expected, survival rates estimated from UK sample data are consistently lower than estimates from the corresponding count data, Figure \@ref(fig:srts)A.
-The correlation between the two methods points out that the sampling method underestimates the survival rate by 32% compared to the counting method, Figure \@ref(fig:srts)B.
+As expected, survival rates estimated from UK sample data are consistently lower than estimates from the corresponding count data, Figure \@ref(fig:cor)A.
+The correlation between the two methods points out that the sampling method underestimates the survival rate by 32% compared to the counting method, Figure \@ref(fig:cor)B.
 Estimation precision is similar for the two methods.
@@
 
-```{r srts, fig.cap = "(ref:srts)"}
-knitr::include_graphics("../Output/plot_5.png")
+```{r cor, fig.cap = "(ref:cor)"}
+knitr::include_graphics("../Output/plot_paper_4.png")
 ```
 
-## Comparison to maximum recruitment rate
+########### Comparison to maximum recruitment rate
 
@@
 
-```{r mgr, fig.cap = "(ref:mgr)"}
-knitr::include_graphics("../Output/plot_6.png")
-```
-
-Upon the conservative assumption of a maximum adult survival uniformly distributed over 0.70 to 0.90, the maximum recruitment rate reaches 0.65 [0.51; 0.75] and 0.66 [0.45; 0.90] recruits per breeder for UK and France respectively (Figure \@ref(fig:grcomp)).
+Upon the conservative assumption of a maximum adult survival uniformly distributed over 0.70 to 0.90, the maximum recruitment rate reaches 0.65 [0.51; 0.75] and 0.66 [0.45; 0.90] recruits per breeder for UK and France respectively.
 When there is no harvest, the estimated average recruitment rate of the French population is in the range of the proxy of the maximum recruitment rate, 0.51 [0.46; 0.57] recruits per breeder.
@@
 
-```{r grcomp, fig.cap = "(ref:grcomp)"}
-knitr::include_graphics("../Output/plot_7.png")
+## Response of vital rates to two different harvest strategies
+
+```{r mgr, fig.cap = "(ref:mgr)"}
+knitr::include_graphics("../Output/plot_paper_5_mod.png")
 ```
@@
 # Discussion
+
+dire que disantangle recruitment rate to survival bring capital information to understand the process of harvest on the population dynamics and brings knowledge sur les leviers pour mieux conserver une espèce
+on mieux la controler for invasive species
+
+dire que grâce à ça, on peut fournir des variations interannuelles de survie et de recruitment, ce que ne permet pas toujours la cmr en raison des tailles d'échantillon.
+
+Higher mortality on adult in GB -> better detection in winter;
 
@@
 The robustness of the correlation between the two methods might be overrated because two successive population size values are parts of the recruitment rate, which mechanically scale the outputs.
-It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit more noisy than for the recruitment rate (Figure \@ref(fig:sup)A).
+It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit more noisy than for the recruitment rate.
 The uncertainty around the estimates is similar for the two methods even though the counting method is more indirect to infer on the recruitment rate, which is satisfying.
@@
 
-# Supplement
+# Supplementary Materials
 
 ```{r sup, fig.cap = "(ref:sup)"}
-knitr::include_graphics("../Output/plot_8.png")
+knitr::include_graphics("../Output/plot_paper_6.png")
 ```
 
-(ref:count) Evolution of the size of the two Ruddy duck populations; y-axis is on a log-scale
+(ref:count) Evolution of the size of the two Ruddy duck populations; y-axis is on a log-scale; data availability refers to censuses that distinguished female-like from male-like individuals
 
@@
 
-(ref:pfem) Female proportion in hunting bags vs in counts; For the first category, one point corresponds to the female proportion estimated from direct sex identification of the Ruddy ducks controlled over one year in a population; for the second category, one point is the proportion of female-like plumage in a population counted in winter.
-The difference is due to immature males that look like females before moulting
+(ref:pfem) Female proportion in hunting bags vs in counts; For the first category, one point corresponds to the female proportion estimated from direct sex identification of the Ruddy ducks controlled over one year in a population; for the second category, one point is the proportion of female-like plumage in a population counted in winter; the difference is due to immature males that look like females before moulting
 
-(ref:grts) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:vital) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)
 
-(ref:rrts) (A) Estimates of recruitment rate, i.e.
+(ref:cor) (A) Estimates of recruitment rate, i.e.
 average number of recruits produced per breeder; (B) Comparison of the recruitment rate values between the two estimation methods; bars define the 95% confidence intervals
 
-(ref:srts) (A) Estimates of adult survival rate using population growth rates; (B) Comparison of the survival rate values between the two estimation methods; bars define the 95% confidence intervals
+(ref:mgr) Comparison between proxies of maximum recruitment rate and average recruitment rates estimated with and without harvest (*counting* method only); proxies are estimated by using the maximum growth rates; bars define the 95% confidence intervals ;Estimates of the maximum population growth without harvest $(\lambda - 1)$; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates; bars define the 95% confidence intervals
 
-(ref:mgr) (A) Estimates of the maximum population growth without harvest $(\lambda - 1)$; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates; bars define the 95% confidence intervals
-
-(ref:grcomp) Comparison between proxies of maximum recruitment rate and average recruitment rates estimated with and without harvest (*counting* method only); proxies are estimated by using the maximum growth rates; bars define the 95% confidence intervals
-
-(ref:sup) (A1) Estimates of the proportion of the immature individuals in the wintering population; (B1) estimates of the number of recruits, i.e.
-the young adults reproducing for their first time; (A2) & (B2) comparison of the values between the two estimation methods; bars define the 95% confidence intervals
+(ref:sup) Estimated harvest rate on adults; as only half of the controlled individuals were aged in France before 2009, we applied the average age ratio over 2009-2019 on unaged individuals to provide an estimate of the number of adults in hunting bags before 2009; since the harvest increased over time, we split the time series in 3 categories : we defined a "no harvest" period before 1999 because the harvest rate for both countries was mostly null and always below 10%, we defined a "low harvest" between 1999 and 2005 because the harvest rate varies around 20% for both countries, the harvest rate then raises significantly for both countries, so we defined a "high harvest" period from 2006

### 2023-02-09 Code/Model_test_nimble_2.R

@@ -1,0 +1,93 @@
+Sys.setlocale("LC_ALL", "en_US.UTF-8")
+
+require(tidyverse)
+require(nimble)
+require(MCMCvis)
+require(mcmcplots)
+
+set.seed(0)
+
+sd_sim = .3
+
+tibble(
+  ts = c(rep(1, 11),
+         rep(2, 5))) %>% 
+  rowid_to_column("x") %>% 
+  mutate(
+    y = c(
+    (13:23 / 3) %>% map_dbl(~ rnorm(1, .x, sd_sim)),
+    (10:14 / 3) %>% map_dbl(~ rnorm(1, .x, sd_sim))
+  ), 
+  N = exp(y)) -> q
+
+JuvCode <- nimbleCode(
+  {
+   # hierarchical prior on lambda max
+    r_max_avg ~ dunif(log(0.01), log(5))
+    r_max_sd ~ dunif(0, 1e2)
+    
+    prior_r_max ~ dnorm(r_max_avg, sd = r_max_sd)
+    log(prior_lambda_max) <- prior_r_max
+    
+    for (r in 1:2) {
+      # 
+      r_avg[r] ~ dnorm(r_max_avg, sd = r_max_sd)
+      # r_avg[r] ~ dunif(log(0.01), log(5))
+      N_int[r] ~ dnorm(0, sd = 1e2)
+      N_sd[r] ~ dunif(0, 1e3)
+      
+      log(lambda[r]) <- r_avg[r]
+    }
+    
+    # LL
+    for (t in 1:R) {
+      
+     N[t] ~ dnorm(N_int[ts_id[t]] + r_avg[ts_id[t]] * year[t], 
+                                sd = N_sd[ts_id[t]])
+    }
+    
+  })
+
+# model data
+JuvConst <- list(R = nrow(q),
+                 ts_id = q$ts)
+
+JuvData <- list(N = q$y,
+                year = q$x)
+
+JuvInit <- list(r_max_avg = log(1), 
+                r_max_sd = 1, 
+                prior_r_max = log(1.5),
+                N_int = rep(0, 2),
+                r_avg = rep(log(1), 2),
+                N_sd = rep(1, 2))
+
+# node targets
+JuvMon <- c("r_max_avg", 
+            "r_max_sd", 
+            "prior_r_max",
+            "r_avg",
+            "prior_lambda_max",
+            "lambda")
+
[…]

### 2023-02-09 Code/Model_vital_rates.R

@@ -4,8 +4,6 @@ Sys.setlocale("LC_ALL", "en_US.UTF-8")
 
 require(tidyverse)
 require(rlang)
-require(openxlsx)
-require(readxl)
 require(lubridate)
 require(nimble)
 require(MCMCvis)
@@ -50,31 +48,7 @@ scale_x_date_own <- function(coef = 5e-2) {
     labs(x = NULL) 
   )}
 
-raw_plot <- function(para,
-                     title = NA, 
-                     ylimits = NA, 
-                     percent = FALSE) {
-  para = enquo(para)
-  
-  JuvOut4 %>%
-    filter(par == as_name(para)) %>% 
-    ggplot(aes(x = year, y = `50%`, color = method, fill = method)) +
-    facet_wrap(~ pop, nrow = 1, scales = "free_x") +
-    # geom_line(linetype = "dashed", alpha = .5) +
-    geom_point() +
-    geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = .5, size = 1.2) +
-    scale_y_continuous(limits = if (is.na(ylimits)) {c(0, NA)} else {c(0, ylimits)}) +
-    scale_x_date(date_breaks = "2 years", 
-                 date_minor_breaks = "1 year", 
-                 date_labels = "%Y") +
-    guides(x =  guide_axis(angle = 45)) +
-    scale_color_manual(values = c_met) +
-    labs(title = title, x = NULL, y = NULL) -> tmp
-  
-  if (percent == TRUE) {tmp + scale_y_percent()} else {tmp}
-}
-
-cor_plot <- function(col, title = NULL, limits = NA, breaks = NA, one = 1) {
+cor_plot <- function(col, title = NULL, subtitle = NULL, limits = NA, breaks = NA, one = 1) {
   
   col2 <- str_c(col, c("2.5%", "50%", "97.5%"), sep = "_")
   col_s <- str_c(col2, "sampling", sep = "_")
@@ -102,14 +76,15 @@ cor_plot <- function(col, title = NULL, limits = NA, breaks = NA, one = 1) {
     ggplot(aes(x = !!sym(col_s[2]), y = !!sym(col_c[2]))) +
     geom_abline(intercept = 0, 
                 slope = slope[1],
-                color = "gray", size = 1.3) +
+                color = "gray", linewidth = 1.3) +
     geom_abline(slope = 1, linetype = "dashed", alpha = one) +
     geom_point(aes(color = year)) +
[…]

### 2023-04-06 Code/Draft_ruddy_duck_vital_rates2.Rmd
Second version submitted to coauthors

--- 2023-02-09 Code/Draft_ruddy_duck_vital_rates.Rmd
@@
 - Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr]
-- Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
 - Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk]
@@
 - Matthieu Guillemain^[Office Français de la Biodiversité, France - matthieu.guillemain@ofb.gouv.fr]
+- Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
 
@@
 
-  **Keywords**: fecundity - productivity - reproductive success - age ratio - duck
+  **Keywords**: fecundity - productivity - reproductive success - age ratio - vital rate - duck
 output: 
@@
 
-We monitored the demography of Ruddy duck populations in both Great-Britain and France thanks to exhaustive counts on the wintering grounds during the period between December 1 and January 31.
+We monitored the demography of Ruddy duck populations in both Great Britain and France thanks to exhaustive counts on the wintering grounds during the period between December 1 and January 31.
 In addition to these exhaustive counts, censuses that distinguished female-like from male-like individuals (Figure \@ref(fig:count)) were performed from 2006 to 2012 in Great Britain, which corresponded to a strong decline of the population.
@@
 During these years, the eradication effort, under control of the authorities in both countries, was recorded.
-Even if the age ratio in the hunting bag was unknown before 2006 in France, we can qualify the harvest pressure on adults in three categories for both countries: "no harvest" before 1999, "low harvest" between 1999 and 2005, and "high harvest" from 2006 (see Figure \@ref(fig:sup) in Supplementary Materials).
+Even if the age ratio in the hunting bag was uncertain before 2009 in France, we can qualify the harvest pressure on adults in three categories for both countries: "no harvest" before 1999, "low harvest" between 1999 and 2005, and "high harvest" from 2006 (see Figure \@ref(fig:sup) in Supplementary Materials).
 
@@
 However, the small population size in France prevented from getting precise adult sex ratios.
-A comparison over months in UK showed no difference in adult male proportion, so data from adults collected over the whole year were used to estimate the male proportion.
+A comparison over months in Great Britain showed no difference in adult male proportion, so data from adults collected over the whole year were used to estimate the male proportion.
 A comparison of male proportion among years with more than 500 samples did not show significant interannual differences.
@@
 To do so, we rejected Great Britain data before 1972 since the size estimate of this population is noisy below 50 individuals (Figure \@ref(fig:count)).
-After reaching the threshold of 1000 individuals, the Great Britain population growth observed a strong inflexion whereas no harvest pressure was applied.
+After reaching the threshold of 1000 individuals, the Great Britain (GB) population growth observed a strong inflexion whereas no harvest pressure was applied.
 This observation suggests that a negative density-dependence process might occur beyond 1000 individuals and led to consider only the first sequence of the time series to infer on the maximum growth rate in Great Britain, i.e.
 1972-1981.
-For the French population, the sequence without harvest spread over the period 1994-1999.
+For the French (FR) population, the sequence without harvest spread over the period 1994-1999.
 
@@
 Only data from years with more than 100 individuals harvested in winter were selected.
-It covers 5 years of the count time series of the Great Britain population.
+It covers 5 years of the count time series of the GB population.
 The presence of the bursa of Fabricius enables immature to be identified [@Hochbaum1942].
@@
 
-## Impact of harvest strategies on population dynamics
+## Impact of harvest strategies on vital rates {#HARV}
 
-Over the "high harvest" period, i.e.
+Over the high harvest period, i.e.
 from 2006, control strategies differed from Great Britain to France.
 In Great Britain, harvest occurred mostly in winter, i.e.
-before the breeding period (53.4% of adults were shot before the 30^th May), whereas it was done mostly in summer in France, i.e.
-during and after the breeding period (81.6% of adults were shot after the 30^th May).
-We compared the population growth of the two populations in light of the difference between the two strategies and we assessed the respective response of the vital rates by comparing the average values over the "high harvest" period to proxies of maximum vital rates estimated when both populations reached their maximum growth (see Section \@ref(VAL)).
+before the breeding period (53.4% of adults were shot before the 30^th^ May), whereas it was done mostly in summer in France, i.e.
+during and after the breeding period (81.6% of adults were shot after the 30^th^ May).
+We compared the population growth of the two populations in light of the difference between the two strategies and we assessed the respective response of the vital rates by comparing the average values over the high harvest period to proxies of maximum vital rates estimated when both populations reached their maximum growth (see Section \@ref(VAL)).
+In France, a LIFE project was implemented from early 2019 to intensify the harvest pressure, especially in winter and spring.
+We then excluded the FR time series from 2019 in order to provide a comparison of homogeneous harvest strategies.
 
@@
 We used uninformative priors on all parameters.
-As the maximum growth rate is an life history trait quite stable among populations of a same species, we used a uninformative hierarchical prior for this parameter.
+As the maximum growth rate is an life history trait expected to be stable among populations of a same species, we used a uninformative hierarchical prior for this parameter.
 We generated three chains of length 500000, with a thinning of 10 to avoid autocorrelation in the samples, and we discarded the first 2000 samples as burn-in.
@@
 
-The harvest rate on adults is considered high for both populations from 2006 (Figure \@ref(fig:sup)).
-There is no significant difference between the two populations, t-test p-value : 0.23, and the average harvest rate is 0.49.
+## Estimates of vital rates
 
-Time series derived from count data show that the two populations display comparable initial growth.
-They grew freely in both countries until 1999.
-Although shooting was initiated in 1999 in both countries, both populations begun declining only after much higher harvest efforts had been achieved in 2005 and 2018 respectively (figure 3).
-This history underlines that both populations observed a large spectrum of harvest effort, from no pressure to very high pressure.
-For sake of interpretation, we ascribed the time-period 2001-2004 to the no harvest category in France because the harvest pressure was negligible hand hence had no effect on the population growth.
-
-Also, covering contrasted populations dynamics corresponding to different levels of harvesting pressure.
-In France, data before 2004 correspond to a period of steady population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure .
-In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period.
-In parallel, the sampling dataset, which corresponds to individuals shot in winter during control operations, covers 9 years only in UK, with 5 years in common with the corresponding counting time series.
-Estimating recruitment rate from the evolution of apparent sex-ratio in winter
-
-Time series show that the two populations follow similar demographic histories (Figure \@ref(fig:count)).
-They grew freely in both countries until 1999.
-Control measures have been then applied from 1999 to 2005 in the UK and from 1999 to 2018 in France.
-In both countries, these management measures led to stop the expansion of the ruddy ducks.
-From 2005 in the UK and from 2018 in France, higher control effort led to a rapid depletion in both populations.
-For sake of interpretation, the time period 2001-2004 was classified in the *no harvest* category in France because the harvest pressure applied over this period was very limited with no effect on the population growth.
-
-The counting dataset, which differentiates the male-like from the female-like individuals in winter, covers 7 years and 19 years in UK and France respectively.
-Combined together, the counting datasets cover different population dynamics corresponding to a gradient of harvest pressure.
-In France, data before 2004 correspond to a period of population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure.
-In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period.
-In parallel, the sampling dataset, which corresponds to individuals shot in winter during control operations, covers 9 years only in UK, with 5 years in common with the corresponding counting time series.
-
-## Likelihood of the range of vital rates
-
-########## OLD Recruitment rate variability
-
-The male proportion in adults, which is necessary to infer on the recruitment rate from count data, is estimated at 0.60 [0.59; 0.61].
-The counting method provides estimations of the recruitment rate, see Figure \@ref(fig:vital).
-In UK, values vary from 0.06 [0.05; 0.07] to 0.27 [0.20; 0.34] recruits per breeder, and in France, 0.08 [0.04; 0.12] to 0.76 [0.64; 0.88] recruits per breeder.
-There are significant year-to-year variations with a maximum amplitude of 0.68 recruits per breeder in France (9.5-time variability).
-On the available period, recruitment rate in UK is smaller than in France.
-Recruitment rates are estimated in both populations over 5 years, and no correlation is observed between the two populations.
-The lowest recruitment rate observed in France, which is in 2019, corresponds to the start of the depletion of the population.
-It is noticeable that it is in the same range than the UK values, also estimated during a strong population depletion.
-
-Recruitment rates estimated from UK sample data are consistently higher than estimates from the corresponding count data, Figure \@ref(fig:vital).
-The correlation between the two methods is strong and points out that the sampling method overestimates the recruitment rate by 44% compared to the counting method, Figure \@ref(fig:vital).
-Estimation precision is similar for the two methods.
-The UK population started to deplete from 2006, but no significant shift is observed between the two periods.
+The method successfully provides estimates and associated uncertainty of immature proportion in both wintering populations.
+The adult survival rate and the recruitment rate, i.e.
+the two components of the population growth rate, are also successfully estimated by the method (Figure \@ref(fig:vital)).
 
@@
 
-## Likelihood of interannual variability in vital rates
+The male proportion in adults, which is a prerequisite to infer on the immature proportion, is estimated at 0.60 [0.59; 0.61].
+The immature proportion varies from 0.16 [0.07; 0.24] to 0.54 [0.44; 0.62].
+The lower values are similar for the two populations, but the range of the immature proportions in the GB population is far more limited (upper value : 0.30 [0.14; 0.45]).
+Conversely from the FR population, there is no significant difference from a year to another because all estimates overlap each other.
+We can notice that the immature proportion for the GB population is quite stable over the time series available, but a slight decrease is observed for the FR population.
 
-########## Consistency of the corresponding adult survival
+The adult survival rate varies from 0.33 [0.29; 0.37] to 1.01 [0.89; 1.14].
+All estimates are not significantly outside the range of a survival rate defined without immigration [0; 1].
+There is no trend in adult survival for both populations, but we can notice two main differences.
+The GB population has the lowest values of adult survival rate, and even if there are significant difference among years for both populations, the variability for the GB population is a bit lower.
 
-Adult survival rates were deduced from the recruitment rate estimates and the observed population growth rates, see Figure \@ref(fig:cor).
-In UK, values that are obtained from the counting method ranged from 0.21 [0.20; 0.22] to 0.68 [0.65; 0.72].
-In France, survival rates vary from 0.39 [0.34; 0.44] to 1.08 [0.98; 1.17].
-Like recruitment rates, there is significant year-to-year variability without clear temporal structure in both populations with a maximum amplitude of 0.69 in France (2.8-time variability).
-Unlike the recruitment rate, the survival rate in 2019 in France is not a low outlier.
+The recruitment rate varies from 0.09 [0.04; 0.13] to 0.75 [0.63; 0.87].
+All estimates are above 0, meaning there is no senseless estimation.
+The maximum recruitment rate being 0.68 [0.36; 0.78] (see Section \@ref(MAX)), all estimates are not significantly outside the range of a recruitment rate defined without immigration [0; 0.78].
+Similarly to the immature proportion, there is no trend for the GB population, but a decrease is observed for the FR population even if it is more noisy than for the immature proportion.
+Similarly to the survival rate, the GB population has low values of recruitment rate, and even if there are significant difference among years for both populations, the variability for the GB population is far lower than for the FR population.
 
-As expected, survival rates estimated from UK sample data are consistently lower than estimates from the corresponding count data, Figure \@ref(fig:cor)A.
-The correlation between the two methods points out that the sampling method underestimates the survival rate by 32% compared to the counting method, Figure \@ref(fig:cor)B.
-Estimation precision is similar for the two methods.
-The sample time series shows that the adult survival is significantly lower after 2006, which corresponds to the start of the depletion of the UK population.
+The lower variability of both vital rates for the GB population explains its steady population growth over the considered time series compared to the FR one (Figure \@ref(fig:count)).
+In the FR population, the range of the recruitment rate (min/max difference: 0.66) is higher than the range of the survival rate (min/max difference: 0.58).
+Conversely, the range of the survival rate (min/max difference: 0.46) in the GB population is far higher than the range of the recruitment rate (min/max difference: 0.08).
+This reflects that the variability of the population growth rate for the FR population is mostly driven by the recruitment rate whereas the main source of variability is the adult survival for the GB population.
+
+## Likelihood of year-to-year variability in vital rates
+
+From the 5 years available in the GB population, immature proportions estimated from the count-based method is positively correlated to estimates from hunting-bag data (Figure \@ref(fig:cor)).
+The hypothesis that the proportion of female-like individuals in the wintering populations reflects the age structure of these populations is then supported by this result.
+This correlation is even stronger when considering vital rates (Figure \@ref(fig:cor)).
+This result is expected since the two methods estimating these vital rates have a component in common, the growth rate of the population (see Equation \@ref(eq:vr)).
+
+It is noticeable that a 1:1 correlation is never obtained.
+Immature proportion is lower in counts than in hunting bags.
+This result is not surprising because immature individuals are expected to be more vulnerable to hunting than adults and might then be over represented in hunting bag data.
+As a consequence, the correlation coefficients for both vital rates differ also from 1.
+This difference between the two approaches does not question the previous validation of the ability of the count method to track the interannual variability of the vital rates.
 
@@
 
-########### Comparison to maximum recruitment rate
+## Response of vital rates to two different harvest strategies {#MAX}
 
-Maximum population growth rates are similar for the two populations, 1.45 [1.36; 1.55] and 1.47 [1.29; 1.68] for UK and France respectively, which corresponds to an increase of about 45% per year (Figure \@ref(fig:mgr)).
-After reaching a certain size (~ 1000 individuals), the UK population growth rate falls to 1.06 [1.04; 1.07], which corresponds to an increase of 6% per year, even though no harvest pressure was applied on it.
+Under no harvest pressure, the maximum growth rates of the two populations are very similar (Figure \@ref(fig:mgr)), respectively 1.45 [1.36; 1.55] and 1.52 [1.25; 1.88] for the GB and the FR population.
+It corresponds to an increase of 45% per year and 52% per year respectively.
+Based on literature, we assumed a maximum survival rate similar for both populations 0.85 [0.7; 1].
+We deduced a maximum recruitment rate of 0.60 [0.42; 0.78] for the GB population and of 0.68 [0.36; 1.06] for the FR population.
 
-Upon the conservative assumption of a maximum adult survival uniformly distributed over 0.70 to 0.90, the maximum recruitment rate reaches 0.65 [0.51; 0.75] and 0.66 [0.45; 0.90] recruits per breeder for UK and France respectively.
-When there is no harvest, the estimated average recruitment rate of the French population is in the range of the proxy of the maximum recruitment rate, 0.51 [0.46; 0.57] recruits per breeder.
-When there is harvest, the average recruitment rates are 0.14 [0.12; 0.16] and 0.33 [0.31; 0.36] recruits per breeder for the UK and France respectively.
-Both values are significantly lower than the maximum recruitment rate, and it is noticeable that the average recruitment rate is lower for the UK population, which is in depletion over the count time series, than for the French population, which of the dynamic is mostly stable over the monitored period.
+Even if a significant part of culled individuals were not aged for some years, the proxies of the harvest pressure show significant shifts over time (Figure \@ref(fig:sup)).
+This led to consider a high harvest pressure period from 2006 with no clear difference between the two populations because the signal is very noisy.
+Even if the harvest pressure seems comparable, the harvest strategies differ significantly between the two populations, see Section \@ref(HARV).
+Under high harvest pressure, the growth rate drops to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for the GB and the FR population respectively, which corresponds to a decrease of 44% per year and 3% per year.
+This result shows that the GB population significantly decreases during the high harvest pressure whereas the FR population is stabilized.
 
-## Response of vital rates to two different harvest strategies
+The adult survival rate under high harvest pressure is 0.44 [0.42; 0.46] and 0.62 [0.59; 0.64] for the GB and FR populations respectively, when the recruitment rate decreases to 0.13 [0.11; 0.15] and 0.32 [0.29; 0.34] respectively.
+The drop of the growth rate of the GB population corresponds to a similar average drop in adult survival rate (-0.41) and in recruitment rate (-0.47).
+The stabilisation of the FR population is more due to a decrease of the recruitment rate (-0.36) than from a drop of the adult survival (-0.23).
+For both populations, the response to the harvest pressure is significant, meaning that the compensation of the population to harvest mortality is limited if not null.
+Vital rates of the GB population are more affected than the FR population.
+Unsurprisingly, the FR harvest strategy focusing on the postbreeding period affects mostly the recruits.
+However, the prebreeding harvest strategy applied on the GB population that targets adults performs to deplete strongly the recruitment to a very low level (0.13 recruits produced per breeder).
 
@@
 
-dire que disantangle recruitment rate to survival bring capital information to understand the process of harvest on the population dynamics and brings knowledge sur les leviers pour mieux conserver une espèce
-on mieux la controler for invasive species
+The delayed sexual dichromatism in a waterfowl species is used here to develop and test the validity of a new non-invasive method aiming to break down the apparent growth of a population into its two structural components: the adult survival rate and the recruitment rate.
+Disentangling the contribution of these two vital rates is a key to understand the response of a population to management measures.
+The development of the method holds on a unique dataset of two comparable populations of a same species, the Ruddy duck.
+Both populations were tracked from their introduction in a similar environment to a period of intense harvest pressure.
+This common history trajectory enables to evaluate the maximum growth of this species, but also its response to harvest pressure.
+As the harvest strategy differs between the two populations, we detail the similarities and the discrepancies of the responses of the vital rates, and open the discussion to the generalization of this method and its implications for waterfowl management.
 
-dire que grâce à ça, on peut fournir des variations interannuelles de survie et de recruitment, ce que ne permet pas toujours la cmr en raison des tailles d'échantillon.
+## Accuracy of the method to estimate vital rates
 
-Higher mortality on adult in GB -> better detection in winter;
+The vital rate estimates of the two populations belongs to the same range, which demonstrate a certain consistency in the results obtained from the method.
+The higher variability observed in France is not unexpected because the FR time series covers a larger spectrum of harvest pressure than the GB time series.
+There is no temporal correlation between the two populations, suggesting that fluctuations of the vital rates are more likely influenced by regional factors than large-scale ones.
+This is consistent with the literature because both nesting/hatching success and juvenile survival are conditional to the onset of laying, which is tightly related to local weather parameters, e.g.
+spring temperature, cumulative rainfall, and water levels [@Folliot2017; @Blums2004; @Dzus1998].
 
-Parler que le model peut etre amélioré en déclarant une proportion de mal à la repro dynamique qui prend en compte les années de fort recrutement vs les faibles recrutements.
+The estimates of adult survival rates correspond to values observed in the literature on species of similar weight [@Nichols1997; @Krementz1997; @Buxton2004].
+This reflects that there is no apparent scaling issue of this method based on count data.
+The assumption that male-like individuals correspond only to adult males is then acceptable.
+This result supports that the count-based method is poorly biased if it is.
+However, some adult survival rate values are outside the range of expected values even if they are never significantly over 1.
+A likely reason to these extreme values is a corruption of the closed population assumption on some years.
+Indeed, if there is an arrival from another Ruddy duck population, the true recruitment cannot explain by itself the population growth and leads to an adult survival combining true adult survival and adult immigration.
+For the FR population, upper outliers of recruitment rates and survival rates are observed in 2002 and 2012, suggesting immigration events on these years.
 
-Dire qu'il semble y avoir une meilleure detection en hiver au vu de la mortalité adulte entre UK & France
+The theoretical maximum of the population growth rate is a potential which is determined by intrinsic life history traits of a species [@Dmitriew2010].
+This potential is expected to be similar to all Ruddy duck populations anywhere in the world.
+Realised maximum growth rates that are observed in this study are very close for the two populations.
+It suggests that the Ruddy duck species reached its intrinsic biological reproduction limit, at least given the environmental conditions of the western Palearctic, i.e.
+about 0.6 recruits per breeder.
+The consistency of this intermediate result suggests that the population size estimation is quite accurate for both populations and demonstrates that their changes are well tracked even with a few individuals.
+The average recruitment rates estimated from the count-based method, with and without harvest, are never higher to the proxies of maximum recruitment rate for both populations.
+The assumption on the accuracy of the count-based method is then again not excluded from the consistency of this result.
 
-Count data in waterfowl that differentiate adult males from other individuals are successfully used in this study to infer estimations of recruitment rates.
-Since this parameter usually explains mostly the population growth in waterfowl [e.g.
-@Arnold2018], the non-invasive method introduced in this study might be considered as a candidate tool to monitor at low cost waterfowl populations.
-In the following section, one checks the accuracy of the method from indicators that were developed for this purpose.
-From the outputs on the two populations of ruddy duck, one discusses about the effects of the harvest on the recruitment rate and its implications on waterfowl management.
+Assuming that hunting-bag data provide a good picture of the year-to-year variability of age structure of a population, the strong correlation with the outputs of our method demonstrates its ability to track interannual fluctuations.
+This result is robust because such correlation is obvious despite being based on a short time series.
+This strong relationship is also satisfying because it demonstrates that the assumption of a constant adult sex ratio is not corrupted.
+The temporal autocorrelation of the adult sex ratio might be explained because it integrates many age-cohorts, which makes it structurally strong.
+Even if significant variations of the adult sex ratio can still occur over long time periods, this result demonstrates that it is not necessary to monitor and update it on a yearly base.
+The adult sex ratio found here on a restricted time period is in line with the Ruddy duck in its native area [0.62 in @Bellrose1980] and with other duck species [@Wood2021].
+The count-based method always provides recruitment rates significantly lower than the method based on hunting-bag data by a factor close to two, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable than adults to hunters [@Bellrose1980; @Fox2014].
+The gap of vulnerability among ages observed here is realistic because the harvest-induced mortality on first-year immatures compared to adults is higher by a factor ranging from 1.3 to 2.6 for hunted duck species in North America [@Bellrose1980].
 
-## Accuracy of the counting method
+Overall, even if the true values of vital rates are not known to test properly the accuracy of the count-based method, a beam of arguments supports that the method is not heavily biased.
+This consists in a big difference with the method based on hunting-bag data as demonstrated in @Fox2014, even when one considers an age structure picture that is limited to the end of the hunting season to limit the bias [@Fox2016].
 
-The recruitment rates from the counting method are in the same range for the two populations, which demonstrate a certain consistency in the results obtained from the method.
-The higher variability observed in France is not unexpected because the French time series is longer and covers a larger spectrum of harvest pressure than in UK.
-There is no temporal correlation between the two populations, so it suggests that the recruitment rate fluctuations are more likely influenced by regional-scale factors than large-scale ones.
-This is consistent with the literature because both nesting/hatching success and juvenile survival are conditional to the onset of lay, which is tightly related to local weather parameters, e.g.
-the spring temperature, the cumulative rainfall, and the water levels [@Folliot2017; @Blums2004; @Dzus1998].
+## Population response to managment measures
 
-Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation with the outputs of the counting method demonstrates the ability of this method to track the recruitment rate variability.
-This result is quite robust because the correlation is obvious with only a five-year time series.
-The assumption of a constant adult sex ratio over time is then acceptable because a significant change in adult sex ratio among years would have biased the evolution of the recruitment rate.
-Since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong.
-Significant variations can then occur only every multiple years, especially if the species is long-living.
-Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.
-The adult sex ratio found here on a restricted time period is in line with the ruddy duck in its native area [0.62 in @Bellrose1980] and with other duck species [@Wood2021].
-The robustness of the correlation between the two methods might be overrated because two successive population size values are parts of the recruitment rate, which mechanically scale the outputs.
-It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit more noisy than for the recruitment rate.
-The uncertainty around the estimates is similar for the two methods even though the counting method is more indirect to infer on the recruitment rate, which is satisfying.
+Both populations respond to the harvest pressure by a large drop of their growth, meaning that the the harvest effort is efficient to control this species.
+The response of the GB population differs from the FR one because it displays a strong decrease when the FR population is stabilized.
+The analysis of the growth alone does not provide clues to understand the mechanisms explaining the difference of the responses between the two populations.
+Indeed, the bigger drop of growth of the GB population might be either explained either by a decrease of the adult survival or of the reproduction success, or even of both components.
 
-The counting method always provides recruitment rates significantly lower than the sampling method by a factor close to two, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable than adults to hunters [@Bellrose1980; @Fox2014].
-For hunted duck species in North America, the harvest-induced mortality on first-year immatures compared to adults is higher by a factor ranging from 1.3 to 2.6 [@Bellrose1980].
-This is a robust result because the approach used to control ruddy ducks differ strongly from usual hunting practices and is apriori less selective.
-It is especially clear in the UK, for which control operations occurred in winter using long-range rifles that prevents the bird from escaping from the threat.
-The assumption on the accuracy of the counting method is then not excluded from the consistency of this result.
+The vital rates that are estimated thanks to the count-based method developed in here demonstrates that the harvest pressure affects on average negatively both the adult survival and the recruitment rate for the two populations.
+The harvest pressure occurring mostly before the breeding period for the GB population targets by definition future breeders, and consequently induces mortality on these individuals and prevent them from reproducing.
+We observe that adult survival and recruitment rate decrease equally, which is a result in line with the expectation of this harvest strategy.
+For the FR population, the harvest pressure occurs mostly during the postbreeding period.
+This strategy should induce a higher mortality on immatures as they are expected to be more vulnerable to hunting [@Bellrose1980; @Fox2014].
+However, this anthropogenic mortality on immatures might be compensated by a higher survival of the future recruits because the competition for resources decreased [@Cooch2014].
+The recruitment rate of the FR population being more affected by 50% than the adult survival, the results support the hypothesis that no or only a limited compensation occurs.
+This conclusion is even strengthened by the result on the GB population.
+Indeed, we should expect a decrease of the recruitment rate lower than the adult survival rate if immatures would take advantage of additional resources that unexploited by culled adult individuals.
+The equal decrease of the two vital rates supports then that there is no compensation    at all.
+This might be explained because the population size is far lower than the carrying capacity of the available habitats, so there is no competition for resources that potentially enhances the survival of the remaining immature individuals [@Peron2013].
 
-Annual adult survival rates deduced from observed population growth rates and recruitment rate estimates are on average in line with the literature on species of similar weight [@Nichols1997; @Krementz1997; @Buxton2004], which reflects that there is no apparent scaling issue.
-The assumption that male-like individuals corresponds only to adult males is then acceptable.
-This result supports that the counting method is poorly biased if it is.
-However, some adult survival rate values are outside the range of expected values even if they never significantly exceed the maximum possible value of one.
-A likely reason to these extreme values is a corruption of the closed population assumption on some years.
-Indeed, if there is an arrival from another ruddy duck population, the true recruitment cannot explain by itself the apparent population growth and indirectly leads to an overestimation of the adult survival.
+Considering a population following a compensatory functioning, the FR strategy is counter-productive.
+As studied Ruddy duck populations are expanding, it seems that the postbreeding harvest strategy still provides some results to control such populations.
+However, the analysis of the year-to-year variability of the recruitment rate demonstrates that the prebreeding strategy buffers strongly the recruitment and maintains it to a low level every single year whereas the recruitment rate of the postbreeding strategy is highly variable.
+This might be because harvesting future breeders ensures to avoid the production of potential recruits whereas the postbreeding harvest strategy cannot prevent and dampen exceptional reproduction events.
+Another source of variability is the detectability of the immature individuals because they spread over many reproduction habitats with a dense vegetation when the future breeders are gathered on a limited number of wintering spots in open water [@Johnsgard1996].
 
-Proxies of maximum recruitment rate are very close for the two populations.
-This strengthens that the ruddy duck species reached its intrinsic biological reproduction limit, at least given the environmental conditions of the western Palearctic.
-The consistency of this intermediate result suggests that the population size estimation is quite good for both populations and demonstrates that their evolution is well tracked even with a few individuals.
-The average recruitment rates estimated from the counting method, with and without harvest, are never higher to the proxies of maximum recruitment rate for both populations.
-The assumption on the accuracy of the counting method is then again not excluded from the consistency of this result.
+The different responses to the harvest pressure of the two Ruddy duck populations demonstrate that it is necessary to account for the time window over which the harvest occurs to produce proper predictions.
+Predicting the impact of a harvest pressure on a waterfowl population is then not only a question about how big is the harvest effort, but also mostly when this effort occur within a year.
+More generally, disentangling the adult survival to the recruitment rate brings capital information to understand the process of harvest on population dynamics and brings key knowledge to improve the conservation of hunted species or the control of invasive species.
 
-Overall, even if the true values of recruitment rates are not known to test the accuracy of the counting method, a beam of arguments support that the method is not heavily biased.
-This consists in a big difference with the sampling method as demonstrated in @Fox2014, even when one considers a age structure picture that is limited to the end of the hunting season to limit the bias [@Fox2016].
+## Implication for waterfowl management
 
-## Influence of harvest on the recruitment rate
-
-Only the French population was monitored over a period covering years without and with harvest.
-The harvest pressure, that is defined into a binary variable (no harvest / harvest), affects on average negatively both the recruitment rate and the adult survival for this population, as expected.
-If it explains only 4.6% of the variance of the adult survival, this value raises to 19.3% for the recruitment rate of the French population.
-The harvest pressure applied in France is thus a factor that likely induces a strong juvenile mortality that impacts the recruitment rate, but a negligible mortality on the adults.
-This conclusion is obviously opposed to the hypothesis that the heterogeneity in immature individuals would dampen the effect of the harvest pressure on the resulting recruitment rate.
-Two points can however explain this apparent discrepancy.
-Since the management targeted the eradication of this ruddy duck population, the harvest on immatures might be simply far higher than what the population could compensate and finally affects the recruitment rate.
-Also, the population started to be controlled when it was still growing at a high rate, so immature individuals were not suffering from density-dependent effects and were consequently sensitive to any additional mortality sources, like the new harvest pressure for instance.
-The low impact of the harvest on adult survival compared to the recruitment rate might be due to the timing of the management strategy adopted in France.
-Since the culling efficiency is poor on the wintering spot, most of control operations occurred in summer just after the reproduction, when the proportion of immature birds is the largest.
-
-The UK population depleted from 2005, which also corresponds to a significant increase of the harvest effort.
-Even if only a few data from the sampling method are available before and after this year, it seems that the recruitment rate is not affected by this increase of harvest, but was already in the lower range of the recruitment rate of the French population when it was exploited.
-On the other hand, the adult survival dropped significantly from 2005.
-This differs strongly from the response of the French population to the harvest.
-Unlike in France, a significant part of the control operations in UK occurred in winter, which could explain this difference.
-Indeed, winter corresponds to the season just before the reproduction period, when the proportion of immatures in the population is the lowest, the juvenile survival being far lower than adults [e.g.
-@Arnold2018].
-The adult survival seems then mostly affected by the harvest pressure applied in winter.
-
-The different responses to the harvest pressure of the two ruddy duck populations demonstrate that it is necessary to account for the time window over which the harvest occurs to produce proper predictions.
-Controlling a waterfowl population is then not only a question about how big should be the harvest effort, but also mostly when this effort should focus within a year to be efficient.
-These first insights suggest that further investigation on this topic would be valuable to calibrate the response of the population to different harvest regimes.
-
-## Implication in waterfowl management
-
-Age ratio, or indifferently recruit proportion, is commonly used to describe the renewal capacity of a waterfowl population and to track the evolution of its productivity [@Smith2001; @Bellrose1980; @Robertson2008; @Rodway2015; @Zimmerman2010].
+The immature proportion, or age ratio, is commonly used to describe the renewal capacity of a waterfowl population and to track the changes of its productivity [@Smith2001; @Bellrose1980; @Robertson2008; @Rodway2015; @Zimmerman2010].
 The recruitment rate is however a better indicator than the age ratio, the latter having only the advantage to be more directly accessible.
-Indeed, the age ratio/recruit proportion does not account for the dynamics of the population and reflects the population productivity only if the population growth is steady.
-For instance, for a breeding population of 100 individuals reaching the next year 100 individuals and then 60 individuals the second year, a 1:1 age ratio, or a 50% recruit proportion, for both years reflects actually a strong decrease of the recruitment rate, respectively 0.5 recruits per breeder and 0.3 recruits per breeder.
-This demonstrates that the age ratio/recruit proportion suffers from caveats that can be misleading for a manager.
+Indeed, the immature proportion/age ratio does not account for the dynamics of the population and reflects the population productivity only if the population growth is steady.
+For instance, let's consider a breeding population of 100 individuals reaching the next year 100 individuals, which then drops to 60 individuals the second year.
+and a constant immature proportion over the 3 years of 50% immature proportion, corresponding to a 1:1 age ratio, for both years reflects actually a strong decrease of the recruitment rate, respectively 0.5 recruits per breeder and 0.3 recruits per breeder.
+This demonstrates that the immature proportion/age ratio suffers from caveats that can be misleading for a manager.
 According to @Blums2004, the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, the other component of the recruitment that is mostly driven by the weather conditions.
@@
 
-Adult survival and recruitment rate equally influence the growth rate variability of the studied ruddy duck populations, but the relative variability of the adult survival is more than three times smaller than the recruitment rate.
+Adult survival and recruitment rate equally influence the growth rate variability of the studied Ruddy duck populations, but the relative variability of the adult survival is more than three times smaller than the recruitment rate.
 This observation is in line with the demographic buffering, or canalization hypothesis, on the adult survival parameter [@Gaillard2003; @Lenzi2021].
@@
 Immatures are more prone to die naturally than adults and are submitted to density dependent processes, which explains why the recruitment rate is the most sensitive parameter to density dependent effect [@Koons2014].
-The evolution of the UK population size seems to show first sights of density dependence from 1000 individuals, but the process was obviously not big enough to fully curb the population growth.
+The evolution of the GB population size seems to show first sights of density dependence from 1000 individuals, but the process was obviously not big enough to fully curb the population growth.
 Since there is a strong correlation between the number of birds observed and the number of sites occupied by the birds, this inflexion may be simply due to a limited number of new favourable sites to be colonized.
 Such density dependent process implies that culling mostly immatures could have only a limited effect on the population dynamics if the harvest pressure remains moderate.
-Considering that the harvest effort occurring in winter impacted mostly the adult survival and led to the strong depletion of the UK ruddy duck population, harvesting early in the season should be favoured if the objective is to maximize the catch in preserving the resources.
-If the goal is to eradicate a population, a strong effort should be done especially in the late season just before the reproduction, as done in the UK.
+Considering that the harvest effort occurring in winter impacted mostly the adult survival and led to the strong depletion of the GB Ruddy duck population, harvesting early in the season should be favoured if the objective is to maximize the catch in preserving the resources.
+If the goal is to eradicate a population, a strong effort should be done especially in the late season just before the reproduction, as done in the Great Britain.
 These hypotheses still requires to be tested thanks to a population model linking the harvest rate to state variables describing the population.
@@
 Applying the method implemented in this study to any other waterfowl species implies that the adult males can be distinguished from the other individuals at distance.
-The latter in the season the dichromatism is observable, the better it is to ensure a relevant picture of the age structure of the population and then good estimates of the recruitment rate.
+The latter in the season the dichromatism is observable, the better it is to ensure a relevant picture of the age structure of the population and then good estimates of the adult survival and the recruitment rate.
 Modifying standard monitoring protocols to distinguish male-like from female-like individuals is almost costless but provides substantial increases in the efficiency and usefulness of monitoring results in conservation [@Nichols2006].
@@
 This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the *Office Français de la Biodiversité*, and Jean-Marc Gillier from the *Société Nationale pour la Protection de la Nature*.
-The authors acknowledge all the contributors of the data collection in the UK and in France, especially: **UK names**, Vincent Fontaine, Denis Lacourpaille, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface.
+The authors acknowledge all the contributors of the data collection; in the Great Britain, Wildfowl & Wetlands Trust staff past and present, particularly Baz Hughes, Peter Cranswick and Colette Hall, and all the project workers at the Animal and Plant Health Agency and its predecessors; in France, Office Français de la Biodiversité staff past and present, especially Vincent Fontaine, Denis Lacourpaille, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface, and the whole team of the Société Nationale pour la Protection de la Nature of the Grand Lieu lake.
 
@@
 
-(ref:vital) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:vital) Evolution of the proportion of immatures in the population and the vital rates in light of the harvest pressure; bars define the 95% confidence intervals
 
-(ref:cor) (A) Estimates of recruitment rate, i.e.
-average number of recruits produced per breeder; (B) Comparison of the recruitment rate values between the two estimation methods; bars define the 95% confidence intervals
+(ref:cor) Comparison of the parameter estimates between the count-based method and hunting-bag-based method.
+Only 5 years for the GB population were available; bars define the 95% confidence intervals; the red arrow defined the average difference in proportion between the two methods
 
-(ref:mgr) Comparison between proxies of maximum recruitment rate and average recruitment rates estimated with and without harvest (*counting* method only); proxies are estimated by using the maximum growth rates; bars define the 95% confidence intervals ;Estimates of the maximum population growth without harvest $(\lambda - 1)$; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates; bars define the 95% confidence intervals
+(ref:mgr) Comparison between average population growth rate under no pressure vs under high harvest pressure; The harvest effort for the GB population occurs moslty during the prebreeding period, whereas it focuses mostly during and after the breeding period for the FR population (see \@ref(HARV) for details); the average vital rates during the maximum growth time period are proxies deduced literature and estimates of maximum population growth estimates (see \@ref(VAL) for details); the vital rates under high harvest pressure are average values over a time period observing a consistent harvest pressure; bars define the 95% confidence intervals
 
-(ref:sup) Estimated harvest rate on adults; as only half of the controlled individuals were aged in France before 2009, we applied the average age ratio over 2009-2019 on unaged individuals to provide an estimate of the number of adults in hunting bags before 2009; since the harvest increased over time, we split the time series in 3 categories : we defined a "no harvest" period before 1999 because the harvest rate for both countries was mostly null and always below 10%, we defined a "low harvest" between 1999 and 2005 because the harvest rate varies around 20% for both countries, the harvest rate then raises significantly for both countries, so we defined a "high harvest" period from 2006
+(ref:sup) Definition of 3 periods of different harvest pressure from an index of harvest rate; as only half of the controlled individuals were aged in France before 2009, we applied the average age ratio over 2009-2019 on unaged individuals to provide an estimate of the number of adults in hunting bags before 2009; since the harvest increased over time, we split the time series in 3 categories : we defined a "no harvest" period before 1999 because the harvest rate for both countries was mostly null and always below 10%, we defined a "low harvest" between 1999 and 2005 because the harvest rate varies around 20% for both countries, the harvest rate then raises significantly for both countries, so we defined a "high harvest" period from 2006

### 2023-04-06 Code/Model_vital_rates.R

@@ -4,7 +4,6 @@ Sys.setlocale("LC_ALL", "en_US.UTF-8")
 
 require(tidyverse)
 require(rlang)
-require(lubridate)
 require(nimble)
 require(MCMCvis)
 require(mcmcplots)
@@ -251,15 +250,13 @@ count_sex_app %>%
   select(method, `female proportion`, tot) %>% 
   bind_rows(frag %>%  
               filter(sex != "ind") %>% 
-              filter(!(pop == "FR" & year(year) < 2011), 
-                     !(pop == "GB" & year(year) > 2009)) %>%
               group_by(year, pop) %>% 
               mutate(tot = sum(shot)) %>%  
               filter(sex == "fem") %>% 
               mutate(fem = sum(shot)) %>% 
               distinct(year, pop, fem, tot) %>%
               ungroup() %>% 
-              filter(tot > 50) %>%
+              filter(tot > 100) %>%
               mutate(`female proportion` = fem / tot, 
                      method = "Female %\nin hunting bags") %>%
               select(method, `female proportion`, tot)) %>% 
@@ -319,7 +316,26 @@ frag %>%
   geom_line(aes(y = n_win)) +
   facet_wrap( ~ pop, ncol = 1, scales = "free_y") +
   scale_x_date_own(0)
-  
+
+frag %>%
+  filter(age != "ind") %>% 
+  group_by(pop, age) %>% 
+  summarise(across(shot, sum)) %>% 
+  group_by(pop) %>% 
+  mutate(tot = sum(shot)) %>% 
+  ungroup() %>% 
+  filter(age == "ad") %>% 
+  mutate(prop = shot / tot)
+
+frag %>%
+  filter(age == "ad") %>% 
+  group_by(pop, repro) %>% 
+  summarise(across(shot, sum)) %>% 
+  group_by(pop) %>% 
+  mutate(tot = sum(shot)) %>% 
+  ungroup() %>% 
+  filter(repro == "before_rep") %>% 
+  mutate(prop = shot / tot)
 
 frag %>%  
   mutate(shot = shot * if_else(age == "ind", 1e-2 * p_ad, 1)) %>% 

### 2023-05-09 Code/Draft_ruddy_duck_vital_rates.Rmd
Accounting for M. Guillemain comments

@@
 - Matthieu Guillemain^[Office Français de la Biodiversité, France - matthieu.guillemain@ofb.gouv.fr]
-
+- Jean-François Maillard^[Office Français de la Biodiversité, France - jean-francois.maillard@ofb.gouv.fr]
+- Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
 date: "`r Sys.setlocale('LC_TIME', 'English'); format(Sys.time(), '%d %B, %Y')`"
@@
   Monitoring the number of individuals is by far the most popular strategy for investigating the environmental factors ruling the population dynamics and for measuring the efficacy of management actions targeting population recovery, control or eradication.
-Unfortunately, it is insufficient for the basic understanding of the demographic mechanisms and more specifically to assess to what extent population growth rate is affected by changes in adult survival rather than to variations in reproductive parameters.
-Usual methods to inform survival and recruitment (capture-mark-recapture, game-hunting bag) rely on catching birds, which suffers from two main drawbacks.
-The selectivity of the catch methods can lead to a biased representation of the underlying population structure.
-In practice, catching and releasing birds is also an invasive approach implying additional disturbances for endangered species or is simply forbidden for invasive alien species.
-In several waterfowl species, a sexual dichromatism is observed in adults whereas immatures of both sexes display a cover-up plumage similar to adult females.
-From populations of Ruddy duck, a species displaying this pattern, this study introduces a non-invasive method based on count data to estimate the respective contribution of the recruitment and the adult survival in the population growth rate.
-Performing count survey during the appropriate time window to differentiate both plumage types is sufficient to provide long-term time series of the two main demographic components of the population growth rate.
+Unfortunately, it is insufficient for the basic understanding of the demographic mechanisms and more specifically to assess the extend to which population growth rate is affected by changes in adult survival rather than to variations in reproductive parameters.
+Usual methods to inform survival and recruitment (capture-mark-recapture, game-hunting bag) rely on catching animals in a way or another, which suffers from three main drawbacks.
+First, the selectivity of the catching methods can lead to a biased representation of the underlying population structure.
+In practice, catching and releasing animals is also an invasive approach implying additional stress to the individuals.
+Furthermore releasing may simply be forbidden for invasive alien species.
+In many waterfowl species, a sexual dichromatism is observed in adults whereas immatures of both sexes display a cover-up plumage similar to adult females.
+From populations of Ruddy duck, a species displaying this pattern, this study introduces a non-invasive method based on count data to estimate the respective contribution of recruitment and adult survival on population growth rate.
+Relying on matricial models shows that performing count survey during the appropriate time window to differentiate both plumage types is sufficient to provide long-term time series of the two main demographic components of the population growth rate.
 
-  **Keywords**: fecundity - productivity - reproductive success - age ratio - duck
+  **Keywords**: reproductive success - fecundity - productivity - age ratio - vital rate - duck - Oxyura jamaicensis
+
 output: 
@@
 
-Assessing population growth-rate is a key step towards a better understanding of factors underlying the dynamics of natural populations [e.g.
+Assessing population growth rate is a key step towards a better understanding of factors underlying the dynamics of natural populations [e.g.
 @Niel2005].
-It is also crucial for measuring the efficacy of management actions eventually undertaken to help population recovery, control, or eradication [@Shea1998].
+It is also crucial for measuring the efficacy of management actions potentially undertaken to help population recovery, control, or eradication [@Shea1998].
 Among the approaches available to managers for reaching these goals, those relying on the monitoring of the number of individuals are by far the most popular [@Rintala2022].
-In many instances, these methods even enable environmental factors underlying changes in population size to be investigated and hence some mitigating actions to be implemented [@Faillettaz2019].
-Unfortunately, relying on the monitoring of numbers of individuals alone hinder basic understanding of the demographic mechanisms underlying changes in population growth rates.
-More specifically, based on counts only, it is nearly impossible to assess to what extent population growth rate is affected by change in survival rather than to variations in reproductive parameters [@Austin2000].
+In many instances, these methods even enable environmental factors underlying changes in population size to be investigated, and hence some mitigating actions to be implemented [@Faillettaz2019].
+Unfortunately, relying on the monitoring of numbers of individuals alone generally hinders basic understanding of the demographic mechanisms underlying changes in population growth rates.
+More specifically, based on counts only, it is nearly impossible to assess the extent to which population growth rate is affected by change in survival rather than to variations in reproductive parameters [@Austin2000].
 To assess the relative sensitivity of population growth rate to factors affecting survival or reproductive parameters, demographers most often rely on the monitoring of individuals by capture-mark-recapture [@Lebreton1992].
-In practice however, one cannot always rely on the capture and release of individuals owing to the legal status of the species that precludes any additional disturbance, e.g.
+In practice, however, one cannot always rely on the capture and release of individuals owing to the legal status of the species that precludes any additional disturbance, e.g.
 critically endangered species, or forbid the release of alive individuals, e.g.
 invasive species or pests.
+Furthermore, capture and handling may also be so stressful to individuals that their behaviour and/or the very parameters under study may be altered.
 Indeed, although highly efficient for assessing demographic parameters, capture-mark-recapture methods have potential drawbacks such as being invasive and hardly affordable when time and money are limiting.
-Genetic monitoring is a non-invasive alternative to capture-mark-recapture approaches based on physical captures, but it suffers from being costly and time-consuming and requires quite rigorous sampling schemes in the field.
+Genetic monitoring is a non-invasive alternative to capture-mark-recapture approaches based on physical captures, but it suffers from being costly and time-consuming, and requires quite rigorous sampling schemes in the field [@Caniglia2011].
 
 Most of the time therefore, managers do the best of a bad job using counts as the only viable option for tracking population growth rates and assessing the efficiency of management actions.
-Counting the number of immatures produced in addition to alive adults may enable adult survival and recruitment rate to be estimated, and hence provides knowledge on which of adult survival and productivity has the most influence on population growth rate.
-Unfortunately, this is only possible for a limited number of species in which broodless adults and those with young display the same detectability.
-Alternatively, assessing the sex and age structure in hunting bags has been used to infer the role of decreasing reproductive success in population declines in a number of games species including ducks and geese [@Fox2018].
-However, this approach is not suited for protected/endangered species.
-Here we exemplify how, in dimorphic species, delayed maturity of males can be exploited for estimating productivity/recruitment and adult survival by differentiating male-like to female-like individuals in repeated winter counts.
-We use two non-native Ruddy duck populations introduced in Europe as a study model.
+Counting the number of immatures produced in addition to live adults may enable adult survival and recruitment rate to be estimated, and hence provides knowledge on which of these components has the most influence on population growth rate.
+The recruitment rate, defined as the average number of immatures per breeder reaching the sexual maturity, is the product of two main sub-parameters: the fecundity followed by the juvenile survival [@Arnold2018; @Etterson2011].
+The recruitment rate is sometimes also called annual fertility [e.g.
+@Koons2014], reproductive rate [e.g.
+@Cooch2014], reproductive success [e.g.
+@Etterson2011], or productivity [e.g.
+@Hagen2008; @Johnson1987].
+Unfortunately, assessing the respective contribution of the adult survival and the recruitment rate is only possible for a limited number of species in which broodless adults and those with young display the same detectability.
+Alternatively, assessing the sex and age structure in hunting bags has been used to infer the role of decreasing reproductive success in population declines in a number of game species including ducks [e.g.
+@Fox2018].
+However, this approach is not suited for protected/endangered species with no harvest.
+Here we exemplify how, in dimorphic species, delayed maturity of males can be exploited for estimating adult survival and recruitment rate by differentiating male-like to female-like individuals in repeated winter counts.
+We used two non-native Ruddy duck, *Oxyura jamaicensis*, populations introduced in Europe as a study model.
 As numerous duck species, the Ruddy duck is dimorphic, with newly born males looking like females until the pre-nuptial moult, which occurs in late winter.
 This means that the apparent proportion of males increases over the course of the wintering season.
-These changes in apparent proportions of males during this period are therefore directly related to the proportion of immatures into the populations and hence to the reproductive success and the recruitment rate.
-This study details the method to infer adult survival and recruitment rate from count data and an application is developed to assess the impact of two different eradication strategies deployed in Great Britain and in France respectively.
+These changes in apparent proportions of males during this period are therefore directly related to the proportion of immatures into the populations and hence to reproductive success and recruitment rate of the previous spring/summer.
+This study details the method to infer adult survival and recruitment rate from count data, and an application is developed to assess the impact of two different eradication strategies deployed in Great Britain and in France, respectively.
 
@@
 The Ruddy duck is a stiff-tailed duck native from the Americas.
-From seven individuals initially acclimated in the Slimbridge Wetland Centre in the Great Britain in the 40’s [@GutierrezExposito2020], a feral population began to establish with the first observed reproductive attempts in the wild in the 60s (Figure \@ref(fig:count)).
-This feral population rapidly grew and spread to the entire country to reach more than 5000 individuals in the early 2000s.
+From seven individuals initially acclimated in the Slimbridge Wetland Centre in Great Britain in the 40’s [@GutierrezExposito2020], a feral population began to establish with the first observed reproductive attempts in the wild in the 60s (Figure \@ref(fig:count)).
+This feral population rapidly grew and spread to the entire country to reach more than 5,000 individuals in the early 2000s.
 By the end of the 80s, a feral population began to establish also on the continent in France, supposedly owing to the arrival of individuals born in Great Britain.
 Contrary to what was observed in Great Britain, the French Ruddy duck population did not spread much, with the vast majority of observations and breeding attempts clustering in Western France.
-In France, during winter, almost none Ruddy duck is observed outside the lake of Grand Lieu (47.09°N, 1.67°W), which is thus the exclusive wintering site and greatly facilitates the monitoring of this population.
-The feral European populations of Ruddy duck are considered as a major threat to the endangered native White-headed duck population of the South-Western Mediterranean because of hybridization risks and thus an elevated potential for genetic pollution and eventually genetic assimilation of the latter by the former [@MunozFuentes2007].
-In order to mitigate the risks of genetic pollution of the White-headed duck by Ruddy ducks, a European plan of eradication of Ruddy ducks has been adopted by the European Commission.
-As a result, eradication measures were taken both in Great Britain and in France from the late 90s [@GutierrezExposito2020], (Figure \@ref(fig:count)).
+In France, during winter, almost no Ruddy duck is observed outside the lake of Grand Lieu (47.09°N, 1.67°W), which is thus the exclusive wintering site and greatly facilitates the monitoring of this population.
+The feral European populations of Ruddy duck are considered as a major threat to the endangered native White-headed duck, *Oxyura leucocephala*, population of the South-Western Mediterranean, because of hybridization risks and thus an elevated potential for genetic pollution and eventually genetic assimilation of the latter by the former [@MunozFuentes2007].
+In order to mitigate the risks of genetic pollution of the White-headed duck by Ruddy ducks, eradication measures were taken both in Great Britain and in France from the late 90s [@GutierrezExposito2020], (Figure \@ref(fig:count)), and a European Ruddy duck eradication plan has been initiated by the European Commission in 1999 [@Hughes1999].
 The release of non-native species in the wild, including Ruddy ducks, is forbidden in countries of the European Union, and the status of White-headed duck populations is highly unfavourable.
 The use of capture-mark-recapture to monitor populations of these species is thus impossible.
-As a result, the effectiveness of the eradication of Ruddy ducks and management actions intended to favour the recovery of White-headed ducks were essentially checked through censuses.
-Censuses, however, prevented from assessing the relative effects or variations in survival and reproductive success/recruitment on population growth rate, which is a key towards a proper understanding of the efficiency of management actions.
+As a result, the effectiveness of the eradication of Ruddy ducks and management actions intended to favour the recovery of White-headed ducks were essentially assessed through censuses.
+Censuses, however, prevented assessment of the relative effects or changes in survival and recruitment rate on population growth rate, which is a key towards a proper understanding of the efficiency of management actions.
 
 ```{r count, fig.cap = "(ref:count)"}
-knitr::include_graphics("../Output/plot_paper_2.png")
+knitr::include_graphics("../Output/plot_paper_2_mod.png")
 ```
@@
 Immature males are thus confounded to immature and adult females before their moult (Figure \@ref(fig:flock)).
-The delayed dichromatism explains why the apparent proportion of females in the winter counts is always higher than the true proportion of females that can be assessed in winter hunting bags (Figure \@ref(fig:pfem)).
-Thus, it is possible to assess the proportion of immatures within a population by monitoring the seasonal evolution of apparent sex ratio and assuming an even sex ratio at birth, comparable mortality rates between females and males among juveniles, and unchanging adult sex ratio over the monitoring period.
+Delayed dichromatism explains why the apparent proportion of females in the winter counts is always greater than the true proportion of females that can be assessed in removals (Figure \@ref(fig:pfem)).
+Thus, it is possible to assess the proportion of immatures within a population by monitoring the seasonal evolution of apparent sex ratio and assuming an even sex ratio at birth [@Bellrose1961; @Blums1996], comparable mortality rates between females and males among juveniles [@Ramula2018; @Wood2021], and unchanging adult sex ratio over the monitoring period [@Devineau2010].
 Then, both adult survival rate and recruitment rate can be estimated by combining the age ratio thus obtained with reliable counts over two successive winters.
@@
 ```{r pfem, fig.cap = "(ref:pfem)"}
-knitr::include_graphics("../Output/plot_paper_1.png")
+knitr::include_graphics("../Output/plot_paper_1_saved.png")
 ```
 
-We monitored the demography of Ruddy duck populations in both Great-Britain and France thanks to exhaustive counts on the wintering grounds during the period between December 1 and January 31.
-In addition to these exhaustive counts, censuses that distinguished female-like from male-like individuals (Figure \@ref(fig:count)) were performed from 2006 to 2012 in Great Britain, which corresponded to a strong decline of the population.
+We monitored the demography of Ruddy duck populations in both Great Britain and France using exhaustive counts on the wintering grounds during the period between December 1 and January 31.
+In addition, counts that distinguished female-like from male-like individuals were performed from 2006 to 2012 in Great Britain, which corresponded to a period of strong decline of the population (Figure \@ref(fig:count)).
 In France, the monitoring spread over 1999, 2001-2009, and 2012-2019.
-This population grew rapidly during the first years of the monitoring and then stabilized from 2006, which corresponds to a period with a strong control effort.
-During these years, the eradication effort, under control of the authorities in both countries, was recorded.
-Even if the age ratio in the hunting bag was unknown before 2006 in France, we can qualify the harvest pressure on adults in three categories for both countries: "no harvest" before 1999, "low harvest" between 1999 and 2005, and "high harvest" from 2006 (see Figure \@ref(fig:sup) in Supplementary Materials).
+This population grew rapidly during the first years of the monitoring and then stabilized from 2006 onwards, which corresponds to a period with a high culling effort, which was performed under control of the authorities in both countries.
+Even if the age ratio in the removals was uncertain before 2009 in France, it is possible to grossly qualify the culling pressure in three categories for both countries: "no culling" before 1999, "low culling" between 1999 and 2005, and "high culling" from 2006 (see Figure \@ref(fig:sup) in Supplementary Materials).
 
@@
 
-In waterfowl, the survival is commonly higher in adult males than in adult females.
-Adult sex ratio is then generally biased towards males (Figure \@ref(fig:pfem), @Wood2021).
-As a consequence, deducing the immature proportion within a population from the observed proportion of adult males is not straightforward because the adult proportion is not just twice as much as the observed proportion of adult males.
-There were no counts differentiating male-like from female-like during the breeding period to estimate properly adult sex ratio.
-But sex identification in adults from hunting bags was available for both Ruddy duck populations.
+In waterfowl, survival rate is commonly higher in adult males than in adult females [@Baldassarre2014].
+Adult sex ratio is then generally biased towards males (Figure \@ref(fig:pfem), @Szekely2014).
+As a consequence, deducing the proportion of immatures within a population from the observed proportion of adult males is not straightforward because the proportion of adults is not just twice as much as the observed proportion of adult males.
+There were no counts differentiating male-like from female-like individuals during the breeding period to properly estimate adult sex ratio.
+But sex identification in adults from removals was available for both Ruddy duck populations.
 However, the small population size in France prevented from getting precise adult sex ratios.
-A comparison over months in UK showed no difference in adult male proportion, so data from adults collected over the whole year were used to estimate the male proportion.
-A comparison of male proportion among years with more than 500 samples did not show significant interannual differences.
+A comparison over months in Great Britain showed no difference in proportion of males in adults, so data from adults collected over the whole year were used to estimate the proportion of males in adults.
+A comparison of proportion of males in adults among years with more than 500 samples did not show significant interannual differences.
 Even if the adult sex ratio may vary over a long time range, it is relatively stable over a few years [@Wood2021].
-As a consequence, hunting bag data from all adults were pooled to estimate the male proportion in adults, see Equation \@ref(eq:pm).
+As a consequence, removal data from all adults were pooled to estimate the proportion of males in adults, see Equation \@ref(eq:pm).
 
 Assuming the additive property of the binomial distribution, the proportion of immatures can be deduced from the cumulated counts of male-like individuals in the wintering population, see Equation \@ref(eq:pic).
-From this proportion and the interannual variation of an abundance index of the population size, the adult survival and the recruitment rate are straightforward, see Equation \@ref(eq:vr).
-If the absolute value of the population size is accessible, numbers of adults and recruits can be estimated, see Equation \@ref(eq:VR).
+From this proportion and the interannual variation of an abundance index of population size, adult survival and recruitment rate are straightforward, see Equation \@ref(eq:vr).
+If the absolute value of population size is accessible, numbers of adults and recruits can be estimated, see Equation \@ref(eq:VR).
 
@@
 |$AF$|Data|Total number of adult females in hunting bags|
-|$p(m|a)$|Parameter|Proportion of males in adults (or probability to be a male knowing it is an adult)|
-|$CML_{i, t}$|Data|Cumulated number of male-like individuals counted in population $i$ in year $t$, which are assumed to be adult males|
+|$p(m|a)$|Parameter|Proportion of males in adults (or probability of being a male knowing it is an adult)|
+|$CML_{i, t}$|Data|Cumulated number of male-like individuals counted in population $i$ in year $t$, which are assumed to all be adult males|
 |$C_{i, t}$|Data|Cumulated number of individuals counted in population $i$ in year $t$|
@@
 The relevance of the method was assessed by testing the likelihood of the estimates of both vital rates.
-For the adult survival rate, we checked if the values belonged to the expected interval $[0;1]$, and we compared the values to literature data that are accessible because the adult survival is commonly assessed by capture-mark-recapture [@Lebreton2001].
-The maximum adult survival is defined as a uniform distribution on $[0.7; 1]$.
-The lower limit corresponds to the upper range of survival rates observed in literature for waterfowl species of similar weight, the upper limit was set to 1 because the survival rates of long-life waterfowl species are very high [@Nichols1997; @Krementz1997; @Buxton2004].
-The recruitment rate is more tricky to validate because it is not upper bounded as it is defined on $[0;\infty[$ and is neither explicitly informed in the literature.
-We then developed an approach to estimate the maximum recruitment rate expected without exploitation and without negative density-dependence processes and we checked if all values were equal or below this maximum recruitment rate.
-If the recruitment rate outputs of the model are significantly higher than the maximum recruitment rate, it means that the method overestimates the recruitment rate, which reflects that a part of adult males are not detected during count surveys.
+For adult survival rate, we checked if the values belonged to the expected interval $[0;1]$, and we compared the values to literature data on ducks that are accessible because adult survival is commonly assessed by capture-mark-recapture [@Lebreton2001].
+Maximum adult survival is defined as a uniform distribution on $[0.7;1]$.
+The lower limit corresponds to the upper range of survival rates in literature for waterfowl species of similar weight, the upper limit was set to 1 because survival rates of long-life waterfowl species are very high [@Nichols1997; @Krementz1997; @Buxton2004; @Nichols1992].
+Recruitment rate is more tricky to validate because it is not upper bounded as it is defined on $[0;\infty[$ and is neither explicitly informed in the literature.
+Only some components of the recruitment rate are generally described, such as nesting rate, nesting success, clutch size, hatching success, early stage survival, but rarely survival between the juvenile stage and the first breeding season, which is necessary to inform recruitment rate [e.g.
+@Baldassarre2014].
+We then developed an approach to estimate the maximum expected recruitment rate without exploitation and without negative density-dependence processes, and we checked if all values were equal or below this maximum recruitment rate.
+If the recruitment rate outputs of the model are significantly higher than maximum recruitment rate, it means that the method overestimates such rate, which reflects that a part of adult males are not detected during count surveys.
 A null or negative recruitment rate would mean that immature males moulted before the survey or/and the assumption on the stability of the adult sex ratio is overrated.
 
-Practically, the maximum recruitment rate is deduced by using its relationship to the maximum growth rate and the maximum survival, see Equation \@ref(eq:GR).
-This relationship comes from a simple reasoning for a closed population: the population size in year $t$ is equal to the number of adults that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
+Practically, maximum recruitment rate is deduced by using its relationship to maximum growth rate and maximum adult survival rate, see Equation \@ref(eq:GR).
+This relationship comes from a simple reasoning for a closed population: population size in year $t$ is equal to the number of adults that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
 the recruitment in year $t$.
-The growth rate of a population is thus the sum of the adult survival rate and the recruitment rate [@Flint2015].
-For an open population, adult survival and recruitment rate are confounded to adult and recruit migration respectively.
+The growth rate of a population is thus the sum of adult survival rate plus recruitment rate [@Flint2015].
+For an open population, adult survival and recruitment rates are confounded to adult and recruit migrations, respectively.
 This relationship becomes more complex if a species with delayed maturity is considered, see @Robertson2008.
@@
 
-The maximum growth rate occurred for both populations during their expansion phase, when the evolution of the population size was the steepest.
+The maximum growth rate occurred for both populations during their expansion phase, when change in population size was the steepest.
 To estimate robust maximum population growth rates for both populations, we smoothed annual population growth rate over a consistent time period by using a linear regression on the logarithm scale, see Equation \@ref(eq:L).
-To do so, we rejected Great Britain data before 1972 since the size estimate of this population is noisy below 50 individuals (Figure \@ref(fig:count)).
-After reaching the threshold of 1000 individuals, the Great Britain population growth observed a strong inflexion whereas no harvest pressure was applied.
-This observation suggests that a negative density-dependence process might occur beyond 1000 individuals and led to consider only the first sequence of the time series to infer on the maximum growth rate in Great Britain, i.e.
+To do so, we discarded Great Britain data before 1972 since the size estimate of this population was noisy below 50 individuals (Figure \@ref(fig:count)).
+After reaching the threshold of 1,000 individuals, the Great Britain (GB) population growth showed a strong inflexion whereas no culling pressure was applied.
+This observation suggests that a negative density-dependence process might occur beyond 1,000 individuals and led to consider only the first sequence of the time series to infer on the maximum growth rate in Great Britain, i.e.
 1972-1981.
-For the French population, the sequence without harvest spread over the period 1994-1999.
+For the French (FR) population, the sequence without culling effort spread over the period 1994-1999.
 
@@
 
-Even if both vital rates vary in a likely range following the validation methodology previously described, it does not prove that the interannual variability is properly tracked.
+Even if both vital rates vary in a likely range following the validation methodology previously described, this does not prove that the interannual variability is properly tracked.
 In order to validate this aspect, we compared the outputs of the count-based method, i.e.
-the immature proportion and both vital rates, to outputs based on data from hunting bags.
-Only data from years with more than 100 individuals harvested in winter were selected.
-It covers 5 years of the count time series of the Great Britain population.
-The presence of the bursa of Fabricius enables immature to be identified [@Hochbaum1942].
-The estimation of the immature proportion in hunting bags is then straightforward, see Equation \@ref(eq:pis).
-The adult survival and the recruitment rate were then estimated by combining this proportion and Equation \@ref(eq:vr).
+the proportion of immatures and both vital rates, to outputs based on data from removals.
+Only data from years with more than 100 individuals culled in winter were selected.
+It covers 5 years of the count time series of the GB population.
+The presence of the bursa of Fabricius enables immature to be identified with certainty [@Hochbaum1942].
+The estimation of the proportion of immatures in removals is then straightforward, see Equation \@ref(eq:pis).
+Adult survival and recruitment rates were then estimated by combining this proportion and Equation \@ref(eq:vr).
 We discussed the ability of the count-based method to accurately track the interannual variability by comparing the outputs of the two methods.
@@
 
-## Impact of harvest strategies on population dynamics
+## Impact of culling strategies on vital rates {#HARV}
 
-Over the "high harvest" period, i.e.
-from 2006, control strategies differed from Great Britain to France.
-In Great Britain, harvest occurred mostly in winter, i.e.
-before the breeding period (53.4% of adults were shot before the 30^th May), whereas it was done mostly in summer in France, i.e.
-during and after the breeding period (81.6% of adults were shot after the 30^th May).
-We compared the population growth of the two populations in light of the difference between the two strategies and we assessed the respective response of the vital rates by comparing the average values over the "high harvest" period to proxies of maximum vital rates estimated when both populations reached their maximum growth (see Section \@ref(VAL)).
+Over the high culling period, i.e.
+from 2006 onwards, strategies differed between Great Britain to France.
+In Great Britain, culling mostly occurred in winter, i.e.
+before the breeding period (53.4% of adults were shot before the 30^th^ May), whereas it was mostly done in summer in France, i.e.
+during and after the breeding period (81.6% of adults were shot after the 30^th^ May).
+We compared the population growth of the two populations in light of the difference between the two strategies and we assessed the respective response of the vital rates by comparing the average values over the high culling period to proxies of maximum vital rates estimated when both populations reached their maximum growth (see Section \@ref(VAL)).
+In France, a LIFE project was implemented from early 2019 to intensify the culling pressure, especially during winter and spring.
+We therefore excluded the FR time series from 2019 in order to provide a comparison of homogeneous culling strategies.
 
@@
 We used uninformative priors on all parameters.
-As the maximum growth rate is an life history trait quite stable among populations of a same species, we used a uninformative hierarchical prior for this parameter.
-We generated three chains of length 500000, with a thinning of 10 to avoid autocorrelation in the samples, and we discarded the first 2000 samples as burn-in.
+As the maximum growth rate is a life history trait expected to be stable among populations of a given species, we used an uninformative hierarchical prior for this parameter.
+We generated three chains of length 500,000, with a thinning of 10 to avoid autocorrelation in the samples, and we discarded the first 2,000 samples as burn-in.
 Chain convergence was assessed using the Gelman and Rubin convergence diagnostic (R<1.1, @Gelman1992).
@@
 Data and code are available here: https://github.com/adri-tab/Ruddy_duck_vital_rates.
-The values **X[Y; Z]** reported in results are the medians and the associated boundaries of the 95% confidence interval of posterior distributions.
+The values **X[Y; Z]** reported in Section \@ref(RES) are the medians and the associated boundaries of the 95% confidence interval of posterior distributions.
 The median was preferred to the mean because of its robustness to skewed distribution.
 
-# Results
+# Results {#RES}
 
-The harvest rate on adults is considered high for both populations from 2006 (Figure \@ref(fig:sup)).
-There is no significant difference between the two populations, t-test p-value : 0.23, and the average harvest rate is 0.49.
+## Estimates of vital rates
 
-Time series derived from count data show that the two populations display comparable initial growth.
-They grew freely in both countries until 1999.
-Although shooting was initiated in 1999 in both countries, both populations begun declining only after much higher harvest efforts had been achieved in 2005 and 2018 respectively (figure 3).
-This history underlines that both populations observed a large spectrum of harvest effort, from no pressure to very high pressure.
-For sake of interpretation, we ascribed the time-period 2001-2004 to the no harvest category in France because the harvest pressure was negligible hand hence had no effect on the population growth.
-
-Also, covering contrasted populations dynamics corresponding to different levels of harvesting pressure.
-In France, data before 2004 correspond to a period of steady population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure .
-In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period.
-In parallel, the sampling dataset, which corresponds to individuals shot in winter during control operations, covers 9 years only in UK, with 5 years in common with the corresponding counting time series.
-Estimating recruitment rate from the evolution of apparent sex-ratio in winter
-
-Time series show that the two populations follow similar demographic histories (Figure \@ref(fig:count)).
-They grew freely in both countries until 1999.
-Control measures have been then applied from 1999 to 2005 in the UK and from 1999 to 2018 in France.
-In both countries, these management measures led to stop the expansion of the ruddy ducks.
-From 2005 in the UK and from 2018 in France, higher control effort led to a rapid depletion in both populations.
-For sake of interpretation, the time period 2001-2004 was classified in the *no harvest* category in France because the harvest pressure applied over this period was very limited with no effect on the population growth.
-
-The counting dataset, which differentiates the male-like from the female-like individuals in winter, covers 7 years and 19 years in UK and France respectively.
-Combined together, the counting datasets cover different population dynamics corresponding to a gradient of harvest pressure.
-In France, data before 2004 correspond to a period of population growth with (almost) no harvest pressure, whereas data in the following years correspond to a stabilized population size with a significant level of harvest pressure.
-In UK, the counting data correspond to a quick population depletion associated to a high level of harvest pressure, especially before the reproduction period.
-In parallel, the sampling dataset, which corresponds to individuals shot in winter during control operations, covers 9 years only in UK, with 5 years in common with the corresponding counting time series.
-
-## Likelihood of the range of vital rates
-
-########## OLD Recruitment rate variability
-
-The male proportion in adults, which is necessary to infer on the recruitment rate from count data, is estimated at 0.60 [0.59; 0.61].
-The counting method provides estimations of the recruitment rate, see Figure \@ref(fig:vital).
-In UK, values vary from 0.06 [0.05; 0.07] to 0.27 [0.20; 0.34] recruits per breeder, and in France, 0.08 [0.04; 0.12] to 0.76 [0.64; 0.88] recruits per breeder.
-There are significant year-to-year variations with a maximum amplitude of 0.68 recruits per breeder in France (9.5-time variability).
-On the available period, recruitment rate in UK is smaller than in France.
-Recruitment rates are estimated in both populations over 5 years, and no correlation is observed between the two populations.
-The lowest recruitment rate observed in France, which is in 2019, corresponds to the start of the depletion of the population.
-It is noticeable that it is in the same range than the UK values, also estimated during a strong population depletion.
-
-Recruitment rates estimated from UK sample data are consistently higher than estimates from the corresponding count data, Figure \@ref(fig:vital).
-The correlation between the two methods is strong and points out that the sampling method overestimates the recruitment rate by 44% compared to the counting method, Figure \@ref(fig:vital).
-Estimation precision is similar for the two methods.
-The UK population started to deplete from 2006, but no significant shift is observed between the two periods.
+The method successfully provided estimates and associated uncertainty of immature proportion in both wintering populations.
+Adult survival rate and recruitment rate, i.e.
+the two components of the population growth rate, were also estimated by the method successfully(Figure \@ref(fig:vital)).
 
@@
 
-########### Comparison to maximum recruitment rate
+## Response of vital rates to the harvest strategy {#MAX}
 
-Maximum population growth rates are similar for the two populations, 1.45 [1.36; 1.55] and 1.47 [1.29; 1.68] for UK and France respectively, which corresponds to an increase of about 45% per year (Figure \@ref(fig:mgr)).
-After reaching a certain size (~ 1000 individuals), the UK population growth rate falls to 1.06 [1.04; 1.07], which corresponds to an increase of 6% per year, even though no harvest pressure was applied on it.
+Under no harvest pressure, the maximum growth rates of the two populations are very similar (Figure \@ref(fig:mgr)), respectively 1.45 [1.36; 1.55] and 1.52 [1.25; 1.88] for the GB and the FR population.
+It corresponds to an increase of 45% per year and 52% per year respectively.
+Based on literature, we assumed a maximum survival rate similar for both populations 0.85 [0.7; 1].
+We deduced a maximum recruitment rate of 0.60 [0.42; 0.78] for the GB population and of 0.68 [0.36; 1.06] for the FR population.
 
-Upon the conservative assumption of a maximum adult survival uniformly distributed over 0.70 to 0.90, the maximum recruitment rate reaches 0.65 [0.51; 0.75] and 0.66 [0.45; 0.90] recruits per breeder for UK and France respectively.
-When there is no harvest, the estimated average recruitment rate of the French population is in the range of the proxy of the maximum recruitment rate, 0.51 [0.46; 0.57] recruits per breeder.
-When there is harvest, the average recruitment rates are 0.14 [0.12; 0.16] and 0.33 [0.31; 0.36] recruits per breeder for the UK and France respectively.
-Both values are significantly lower than the maximum recruitment rate, and it is noticeable that the average recruitment rate is lower for the UK population, which is in depletion over the count time series, than for the French population, which of the dynamic is mostly stable over the monitored period.
+Even if a significant part of culled individuals were not aged for some years, the proxies of the harvest pressure show significant shifts over time (Figure \@ref(fig:sup)).
+This led to consider a high harvest pressure period from 2006 with no clear difference between the two populations because the signal is very noisy.
+Even if the harvest pressure seems comparable, the harvest strategies differ significantly between the two populations, see Section \@ref(HARV).
+Under high harvest pressure, the growth rate drops to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for the GB and the FR population respectively, which corresponds to a decrease of 44% per year and 3% per year.
+This result shows that the GB population significantly decreases during the high harvest pressure whereas the FR population is stabilized.
 
-## Response of vital rates to two different harvest strategies
+The adult survival rate under high harvest pressure is 0.44 [0.42; 0.46] and 0.62 [0.59; 0.64] for the GB and FR populations respectively, when the recruitment rate decreases to 0.13 [0.11; 0.15] and 0.32 [0.29; 0.34] respectively.
+The drop of the growth rate of the GB population corresponds to a similar average drop in adult survival rate (-0.41) and in recruitment rate (-0.47).
+The stabilisation of the FR population is more due to a decrease of the recruitment rate (-0.36) than from a drop of the adult survival (-0.23).
+For both populations, the response to the harvest pressure is significant, meaning that the compensation of the population to harvest mortality is limited if not null.
+Vital rates of the GB population are more affected than the FR population.
+Unsurprisingly, the FR harvest strategy focusing on the postbreeding period affects mostly the recruits.
+However, the prebreeding harvest strategy applied on the GB population that targets adults performs to deplete strongly the recruitment to a very low level (0.13 recruits produced per breeder).
 
@@
 
-dire que disantangle recruitment rate to survival bring capital information to understand the process of harvest on the population dynamics and brings knowledge sur les leviers pour mieux conserver une espèce
-on mieux la controler for invasive species
+The delayed sexual dichromatism in a waterfowl species is used here to develop and test the validity of a new non-invasive method aiming to break down the apparent growth of a population into its two structural components: the adult survival rate and the recruitment rate.
+Disentangling the contribution of these two vital rates is a key to understand the response of a population to management measures.
+The development of the method holds on a unique dataset of two comparable populations of a same species, the Ruddy duck.
+Both populations were tracked from their introduction in a similar environment to a period of intense harvest pressure.
+This common history trajectory enables to evaluate the maximum growth of this species, but also its response to harvest pressure.
+As the harvest strategy differs between the two populations, we detail the similarities and the discrepancies of the responses of the vital rates, and open the discussion to the generalization of this method and its implications for waterfowl management.
 
-dire que grâce à ça, on peut fournir des variations interannuelles de survie et de recruitment, ce que ne permet pas toujours la cmr en raison des tailles d'échantillon.
+## Accuracy of the method to estimate vital rates
 
-Higher mortality on adult in GB -> better detection in winter;
+The vital rate estimates of the two populations belongs to the same range, which demonstrate a certain consistency in the results obtained from the method.
+The higher variability observed in France is not unexpected because the FR time series covers a larger spectrum of harvest pressure than the GB time series.
+There is no temporal correlation between the two populations, suggesting that fluctuations of the vital rates are more likely influenced by regional factors than large-scale ones.
+This is consistent with the literature because both nesting/hatching success and juvenile survival are conditional to the onset of laying, which is tightly related to local weather parameters, e.g.
+spring temperature, cumulative rainfall, and water levels [@Folliot2017; @Blums2004; @Dzus1998].
 
-Parler que le model peut etre amélioré en déclarant une proportion de mal à la repro dynamique qui prend en compte les années de fort recrutement vs les faibles recrutements.
+The estimates of adult survival rates correspond to values observed in the literature on species of similar weight [@Nichols1997; @Krementz1997; @Buxton2004].
+This reflects that there is no apparent scaling issue of this method based on count data.
+The assumption that male-like individuals correspond only to adult males is then acceptable.
+This result supports that the count-based method is poorly biased if it is.
+However, some adult survival rate values are outside the range of expected values even if they are never significantly over 1.
+A likely reason to these extreme values is a corruption of the closed population assumption on some years.
+Indeed, if there is an arrival from another Ruddy duck population, the true recruitment cannot explain by itself the population growth and leads to an adult survival combining true adult survival and adult immigration.
+For the FR population, upper outliers of recruitment rates and survival rates are observed in 2002 and 2012, suggesting immigration events on these years.
 
-Dire qu'il semble y avoir une meilleure detection en hiver au vu de la mortalité adulte entre UK & France
+The theoretical maximum of the population growth rate is a potential which is determined by intrinsic life history traits of a species [@Dmitriew2010].
+This potential is expected to be similar to all Ruddy duck populations anywhere in the world.
+Realised maximum growth rates that are observed in this study are very close for the two populations.
+It suggests that the Ruddy duck species reached its intrinsic biological reproduction limit, at least given the environmental conditions of the western Palearctic, i.e.
+about 0.6 recruits per breeder.
+The consistency of this intermediate result suggests that the population size estimation is quite accurate for both populations and demonstrates that their changes are well tracked even with a few individuals.
+The average recruitment rates estimated from the count-based method, with and without harvest, are never higher to the proxies of maximum recruitment rate for both populations.
+The assumption on the accuracy of the count-based method is then again not excluded from the consistency of this result.
 
-Count data in waterfowl that differentiate adult males from other individuals are successfully used in this study to infer estimations of recruitment rates.
-Since this parameter usually explains mostly the population growth in waterfowl [e.g.
-@Arnold2018], the non-invasive method introduced in this study might be considered as a candidate tool to monitor at low cost waterfowl populations.
-In the following section, one checks the accuracy of the method from indicators that were developed for this purpose.
-From the outputs on the two populations of ruddy duck, one discusses about the effects of the harvest on the recruitment rate and its implications on waterfowl management.
+Assuming that hunting-bag data provide a good picture of the year-to-year variability of age structure of a population, the strong correlation with the outputs of our method demonstrates its ability to track interannual fluctuations.
+This result is robust because such correlation is obvious despite being based on a short time series.
+This strong relationship is also satisfying because it demonstrates that the assumption of a constant adult sex ratio is not corrupted.
+The temporal autocorrelation of the adult sex ratio might be explained because it integrates many age-cohorts, which makes it structurally strong.
+Even if significant variations of the adult sex ratio can still occur over long time periods, this result demonstrates that it is not necessary to monitor and update it on a yearly base.
+The adult sex ratio found here on a restricted time period is in line with the Ruddy duck in its native area [0.62 in @Bellrose1980] and with other duck species [@Wood2021].
+The count-based method always provides recruitment rates significantly lower than the method based on hunting-bag data by a factor close to two, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable than adults to hunters [@Bellrose1980; @Fox2014].
+The gap of vulnerability among ages observed here is realistic because the harvest-induced mortality on first-year immatures compared to adults is higher by a factor ranging from 1.3 to 2.6 for hunted duck species in North America [@Bellrose1980].
 
-## Accuracy of the counting method
+Overall, even if the true values of vital rates are not known to test properly the accuracy of the count-based method, a beam of arguments supports that the method is not heavily biased.
+This consists in a big difference with the method based on hunting-bag data as demonstrated in @Fox2014, even when one considers an age structure picture that is limited to the end of the hunting season to limit the bias [@Fox2016].
 
-The recruitment rates from the counting method are in the same range for the two populations, which demonstrate a certain consistency in the results obtained from the method.
-The higher variability observed in France is not unexpected because the French time series is longer and covers a larger spectrum of harvest pressure than in UK.
-There is no temporal correlation between the two populations, so it suggests that the recruitment rate fluctuations are more likely influenced by regional-scale factors than large-scale ones.
-This is consistent with the literature because both nesting/hatching success and juvenile survival are conditional to the onset of lay, which is tightly related to local weather parameters, e.g.
-the spring temperature, the cumulative rainfall, and the water levels [@Folliot2017; @Blums2004; @Dzus1998].
+## Population response to management measures
 
-Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation with the outputs of the counting method demonstrates the ability of this method to track the recruitment rate variability.
-This result is quite robust because the correlation is obvious with only a five-year time series.
-The assumption of a constant adult sex ratio over time is then acceptable because a significant change in adult sex ratio among years would have biased the evolution of the recruitment rate.
-Since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong.
-Significant variations can then occur only every multiple years, especially if the species is long-living.
-Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.
-The adult sex ratio found here on a restricted time period is in line with the ruddy duck in its native area [0.62 in @Bellrose1980] and with other duck species [@Wood2021].
-The robustness of the correlation between the two methods might be overrated because two successive population size values are parts of the recruitment rate, which mechanically scale the outputs.
-It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit more noisy than for the recruitment rate.
-The uncertainty around the estimates is similar for the two methods even though the counting method is more indirect to infer on the recruitment rate, which is satisfying.
+Both populations respond to the harvest pressure by a large drop of their growth, meaning that the the harvest effort is efficient to control this species.
+The response of the GB population differs from the FR one because it displays a strong decrease when the FR population is stabilized.
+The analysis of the growth alone does not provide clues to understand the mechanisms explaining the difference of the responses between the two populations.
+Indeed, the bigger drop of growth of the GB population might be explained either by a decrease of the adult survival or of the reproduction success, or even of both components.
 
-The counting method always provides recruitment rates significantly lower than the sampling method by a factor close to two, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable than adults to hunters [@Bellrose1980; @Fox2014].
-For hunted duck species in North America, the harvest-induced mortality on first-year immatures compared to adults is higher by a factor ranging from 1.3 to 2.6 [@Bellrose1980].
-This is a robust result because the approach used to control ruddy ducks differ strongly from usual hunting practices and is apriori less selective.
-It is especially clear in the UK, for which control operations occurred in winter using long-range rifles that prevents the bird from escaping from the threat.
-The assumption on the accuracy of the counting method is then not excluded from the consistency of this result.
+The vital rates that are estimated thanks to the count-based method developed in here demonstrates that the harvest pressure affects on average negatively both the adult survival and the recruitment rate for the two populations.
+The harvest pressure occurring mostly before the breeding period for the GB population targets by definition future breeders, and consequently induces mortality on these individuals and prevent them from reproducing.
+We observe that adult survival and recruitment rate decrease equally, which is a result in line with the expectation of this harvest strategy.
+For the FR population, the harvest pressure occurs mostly during the postbreeding period.
+This strategy should induce a higher mortality on immatures as they are expected to be more vulnerable to hunting [@Bellrose1980; @Fox2014].
+However, this anthropogenic mortality on immatures might be compensated by a higher survival of the future recruits because the competition for resources decreased [@Cooch2014].
+The recruitment rate of the FR population being more affected by 50% than the adult survival, the results support the hypothesis that no or only a limited compensation occurs.
+This conclusion is even strengthened by the result on the GB population.
+Indeed, we should expect a decrease of the recruitment rate lower than the adult survival rate if immatures would take advantage of additional resources that unexploited by culled adult individuals.
+The equal decrease of the two vital rates supports then that there is no compensation.
+This might be explained because the population size is far lower than the carrying capacity of the available habitats, so there is no competition for resources that potentially enhances the survival of the remaining immature individuals [@Peron2013].
 
-Annual adult survival rates deduced from observed population growth rates and recruitment rate estimates are on average in line with the literature on species of similar weight [@Nichols1997; @Krementz1997; @Buxton2004], which reflects that there is no apparent scaling issue.
-The assumption that male-like individuals corresponds only to adult males is then acceptable.
-This result supports that the counting method is poorly biased if it is.
-However, some adult survival rate values are outside the range of expected values even if they never significantly exceed the maximum possible value of one.
-A likely reason to these extreme values is a corruption of the closed population assumption on some years.
-Indeed, if there is an arrival from another ruddy duck population, the true recruitment cannot explain by itself the apparent population growth and indirectly leads to an overestimation of the adult survival.
+Considering a population following a compensatory functioning, the FR strategy is counter-productive.
+As studied Ruddy duck populations are expanding, it seems that the postbreeding harvest strategy still provides some results to control such populations.
+However, the analysis of the year-to-year variability of the recruitment rate demonstrates that the prebreeding strategy buffers strongly the recruitment and maintains it to a low level every single year whereas the recruitment rate of the postbreeding strategy is highly variable.
+This might be because harvesting future breeders ensures to avoid the production of potential recruits whereas the postbreeding harvest strategy cannot prevent and dampen exceptional reproduction events.
+The postbreeding strategy relies also on a detectability of immatures highly variable because these individuals spread over many reproduction habitats more or less accessible whereas the prebreeding strategy targets individuals that are more easily localized because they usually gather in open water on a limited number of wintering spots [@Johnsgard1996].
 
-Proxies of maximum recruitment rate are very close for the two populations.
-This strengthens that the ruddy duck species reached its intrinsic biological reproduction limit, at least given the environmental conditions of the western Palearctic.
-The consistency of this intermediate result suggests that the population size estimation is quite good for both populations and demonstrates that their evolution is well tracked even with a few individuals.
-The average recruitment rates estimated from the counting method, with and without harvest, are never higher to the proxies of maximum recruitment rate for both populations.
-The assumption on the accuracy of the counting method is then again not excluded from the consistency of this result.
+The response to the harvest pressure of the two Ruddy duck populations demonstrate that it is necessary to account for the time window over which the harvest occurs to produce proper predictions.
+Predicting the impact of a harvest pressure on a waterfowl population is then not only a question about how big is the harvest effort, but also mostly when this effort occur within a year.
+More generally, disentangling the adult survival to the recruitment rate brings capital information to understand the process of harvest on population dynamics and brings key knowledge to improve the conservation of hunted species or the control of invasive species.
 
-Overall, even if the true values of recruitment rates are not known to test the accuracy of the counting method, a beam of arguments support that the method is not heavily biased.
-This consists in a big difference with the sampling method as demonstrated in @Fox2014, even when one considers a age structure picture that is limited to the end of the hunting season to limit the bias [@Fox2016].
+## Implication for waterfowl management
 
-## Influence of harvest on the recruitment rate
+Tracking variations in abundance is a common tool to determine the conservation status of a population [e.g @Folliot2022].
+However, the abundance alone does not provide information on the mechanisms behind the changes of population growth.
+Identifying the determinants requires to dig into the components structuring the population growth to provide relevant tools to managers [e.g.
+@Austin2000].
+Gaining knowledge on adult survival and the recruitment rate often involves capture mark recapture surveys [e.g.
+@Arnold2018].
+However, tracking fine time-scale variability of the vital rates requires to mark a minimum number of individuals [e.g.
+@Souchay2016] and prevents from implementing such method to small populations with a poor conservation status for instance.
+The method introduced in here bypasses these caveats to provide annual estimates of vital rates.
 
-Only the French population was monitored over a period covering years without and with harvest.
-The harvest pressure, that is defined into a binary variable (no harvest / harvest), affects on average negatively both the recruitment rate and the adult survival for this population, as expected.
-If it explains only 4.6% of the variance of the adult survival, this value raises to 19.3% for the recruitment rate of the French population.
-The harvest pressure applied in France is thus a factor that likely induces a strong juvenile mortality that impacts the recruitment rate, but a negligible mortality on the adults.
-This conclusion is obviously opposed to the hypothesis that the heterogeneity in immature individuals would dampen the effect of the harvest pressure on the resulting recruitment rate.
-Two points can however explain this apparent discrepancy.
-Since the management targeted the eradication of this ruddy duck population, the harvest on immatures might be simply far higher than what the population could compensate and finally affects the recruitment rate.
-Also, the population started to be controlled when it was still growing at a high rate, so immature individuals were not suffering from density-dependent effects and were consequently sensitive to any additional mortality sources, like the new harvest pressure for instance.
-The low impact of the harvest on adult survival compared to the recruitment rate might be due to the timing of the management strategy adopted in France.
-Since the culling efficiency is poor on the wintering spot, most of control operations occurred in summer just after the reproduction, when the proportion of immature birds is the largest.
-
-The UK population depleted from 2005, which also corresponds to a significant increase of the harvest effort.
-Even if only a few data from the sampling method are available before and after this year, it seems that the recruitment rate is not affected by this increase of harvest, but was already in the lower range of the recruitment rate of the French population when it was exploited.
-On the other hand, the adult survival dropped significantly from 2005.
-This differs strongly from the response of the French population to the harvest.
-Unlike in France, a significant part of the control operations in UK occurred in winter, which could explain this difference.
-Indeed, winter corresponds to the season just before the reproduction period, when the proportion of immatures in the population is the lowest, the juvenile survival being far lower than adults [e.g.
-@Arnold2018].
-The adult survival seems then mostly affected by the harvest pressure applied in winter.
-
-The different responses to the harvest pressure of the two ruddy duck populations demonstrate that it is necessary to account for the time window over which the harvest occurs to produce proper predictions.
-Controlling a waterfowl population is then not only a question about how big should be the harvest effort, but also mostly when this effort should focus within a year to be efficient.
-These first insights suggest that further investigation on this topic would be valuable to calibrate the response of the population to different harvest regimes.
-
-## Implication in waterfowl management
-
-Age ratio, or indifferently recruit proportion, is commonly used to describe the renewal capacity of a waterfowl population and to track the evolution of its productivity [@Smith2001; @Bellrose1980; @Robertson2008; @Rodway2015; @Zimmerman2010].
-The recruitment rate is however a better indicator than the age ratio, the latter having only the advantage to be more directly accessible.
-Indeed, the age ratio/recruit proportion does not account for the dynamics of the population and reflects the population productivity only if the population growth is steady.
-For instance, for a breeding population of 100 individuals reaching the next year 100 individuals and then 60 individuals the second year, a 1:1 age ratio, or a 50% recruit proportion, for both years reflects actually a strong decrease of the recruitment rate, respectively 0.5 recruits per breeder and 0.3 recruits per breeder.
-This demonstrates that the age ratio/recruit proportion suffers from caveats that can be misleading for a manager.
+The immature proportion, or age ratio, is commonly used to describe the renewal capacity of a waterfowl population and to track the changes of its productivity [@Smith2001; @Bellrose1980; @Robertson2008; @Rodway2015; @Zimmerman2010].
+The recruitment rate remains however a better indicator than the immature proportion/age ratio, the latter having only the advantage to be more directly accessible.
+Indeed, the immature proportion/age ratio does not account for the dynamics of the population and reflects the population productivity only if the population growth is steady.
+For instance, let's consider a breeding population of 100 individuals reaching the next year 100 individuals, which then drops to 60 individuals the second year.
+If there is 50% of immature individuals each year, corresponding to a 1:1 age ratio, we may conclude that the productivity is stable over time.
+But it actually corresponds to a strong decrease of the productivity (-40%) because the recruitment rate evolved from 0.5 recruits per breeder the first year to 0.3 recruits per breeder the second year.
+This demonstrates that the immature proportion/age ratio suffers from caveats that can be misleading for a manager.
 According to @Blums2004, the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, the other component of the recruitment that is mostly driven by the weather conditions.
@@
 
-Adult survival and recruitment rate equally influence the growth rate variability of the studied ruddy duck populations, but the relative variability of the adult survival is more than three times smaller than the recruitment rate.
-This observation is in line with the demographic buffering, or canalization hypothesis, on the adult survival parameter [@Gaillard2003; @Lenzi2021].
-@Arnold2018 even demonstrates that the recruitment rate can be the most influencing factor on the population growth in some waterfowl species.
-Population management in waterfowl cannot neglect a careful monitoring of the recruitment rate, even if this parameter might be more difficult to track compared to the adult survival.
-A drop in the recruitment rate points out that a population renewal is affected but does show the causation.
-It is however a signal that intermediate life stages before recruitment must be investigated to provide a diagnostic on the changes of the environment conditions that have affected the recruitment rate.
-
-Immatures are more prone to die naturally than adults and are submitted to density dependent processes, which explains why the recruitment rate is the most sensitive parameter to density dependent effect [@Koons2014].
-The evolution of the UK population size seems to show first sights of density dependence from 1000 individuals, but the process was obviously not big enough to fully curb the population growth.
-Since there is a strong correlation between the number of birds observed and the number of sites occupied by the birds, this inflexion may be simply due to a limited number of new favourable sites to be colonized.
-Such density dependent process implies that culling mostly immatures could have only a limited effect on the population dynamics if the harvest pressure remains moderate.
-Considering that the harvest effort occurring in winter impacted mostly the adult survival and led to the strong depletion of the UK ruddy duck population, harvesting early in the season should be favoured if the objective is to maximize the catch in preserving the resources.
-If the goal is to eradicate a population, a strong effort should be done especially in the late season just before the reproduction, as done in the UK.
-These hypotheses still requires to be tested thanks to a population model linking the harvest rate to state variables describing the population.
-Such tool is valuable to assess precisely the impact of the harvest pressure and to test hypotheses on the functioning of the harvest mortality [@Plard2019].
-
 Applying the method implemented in this study to any other waterfowl species implies that the adult males can be distinguished from the other individuals at distance.
-The latter in the season the dichromatism is observable, the better it is to ensure a relevant picture of the age structure of the population and then good estimates of the recruitment rate.
+The latter in the season the dichromatism is observable, the better it is to ensure a relevant picture of the age structure of the population and then good estimates of the adult survival and the recruitment rate.
 Modifying standard monitoring protocols to distinguish male-like from female-like individuals is almost costless but provides substantial increases in the efficiency and usefulness of monitoring results in conservation [@Nichols2006].
@@
 
-(ref:count) Evolution of the size of the two Ruddy duck populations; y-axis is on a log-scale; data availability refers to censuses that distinguished female-like from male-like individuals
+(ref:count) Changes in the size of the two Ruddy duck populations; y-axis is on a log-scale; sex ratio refers to censuses that distinguished female-like from male-like individuals
 
-(ref:flock) Typical observation of a Ruddy duck flock in winter: 10 female-like individuals with a whitish cheek and a dark stripe across it, 4 male-like individuals with a white cheek and a black cap, 3 unidentified individuals © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
+(ref:flock) Typical observation of a Ruddy duck flock during winter: 10 female-like individuals with a whitish striped cheek, 4 male-like individuals with a white cheek and a black cap, 3 unidentified individuals (1^st^, 4^th^, and 10^th^ from left) © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
 
-(ref:pfem) Female proportion in hunting bags vs in counts; For the first category, one point corresponds to the female proportion estimated from direct sex identification of the Ruddy ducks controlled over one year in a population; for the second category, one point is the proportion of female-like plumage in a population counted in winter; the difference is due to immature males that look like females before moulting
+(ref:pfem) Proportion of females in removals vs in counts; For the first category, one data point corresponds to the female proportion estimated from direct sex identification of the Ruddy ducks culled over one year in a population; for the second category, one data point correponds to the proportion of individuals wearing female-like plumage in a population counted in winter; the difference of female proportion between the two approaches is due to immature males that look like females before moulting; this difference is at the root of the method introduced inhere to disentangle survival from recruitment
 
-(ref:vital) Evolution of the size of the two Ruddy duck populations in light of the harvest pressure; on time periods without harvest, maximum population growth rates can be estimated (slopes in red for the UK and in blue for France)
+(ref:vital) Evolution of the proportion of immatures in the population and the vital rates in light of the culling pressure; bars define the 95% confidence intervals
 
-(ref:cor) (A) Estimates of recruitment rate, i.e.
-average number of recruits produced per breeder; (B) Comparison of the recruitment rate values between the two estimation methods; bars define the 95% confidence intervals
+(ref:cor) Comparison of the parameter estimates between the count-based method and removal-based method; only 5 years for the GB population were available; bars define the 95% confidence intervals; the red arrow defined the average difference in proportion between the two methods
 
-(ref:mgr) Comparison between proxies of maximum recruitment rate and average recruitment rates estimated with and without harvest (*counting* method only); proxies are estimated by using the maximum growth rates; bars define the 95% confidence intervals ;Estimates of the maximum population growth without harvest $(\lambda - 1)$; two estimates were produced for the UK population because the growth rate dropped significantly when its size reached more than 1000 individuals; (B) Growth rate estimates; bars define the 95% confidence intervals
+(ref:mgr) Comparison between average population growth rate under no pressure vs under high culling pressure; the culling effort for the GB population occurs moslty during the prebreeding period, whereas it focuses mostly during and after the breeding period for the FR population (see \@ref(HARV) for details); the average vital rates during the maximum growth time period are proxies deduced literature and estimates of maximum population growth estimates (see \@ref(VAL) for details); the vital rates under high culling pressure are average values over a time period observing a consistent culling pressure; bars define the 95% confidence intervals
 
-(ref:sup) Estimated harvest rate on adults; as only half of the controlled individuals were aged in France before 2009, we applied the average age ratio over 2009-2019 on unaged individuals to provide an estimate of the number of adults in hunting bags before 2009; since the harvest increased over time, we split the time series in 3 categories : we defined a "no harvest" period before 1999 because the harvest rate for both countries was mostly null and always below 10%, we defined a "low harvest" between 1999 and 2005 because the harvest rate varies around 20% for both countries, the harvest rate then raises significantly for both countries, so we defined a "high harvest" period from 2006
+(ref:sup) Definition of 3 periods of different culling pressures from an index based on an estimation of the adult culling rate; as only half of the culled individuals were aged in France before 2009, we applied the average age ratio over 2009-2019 on unaged individuals to provide an estimate of the number of adults in removals before 2009; the culling rate increased over time, we then split the time series in 3 categories : we defined a "no culling" period before 1999 because the culling rate for both countries was mostly null and always below 10%, we defined a "low culling" between 1999 and 2005 because the culling rate varies around 20% for both countries, the culling rate then raises significantly for both countries, so we defined a "high culling" period from 2006

### 2023-05-09 Code/Model_test_nimble.R

@@ -1,58 +1,0 @@
-library(nimble)
-library(tidyverse)
-require(mcmcplots)
-set.seed(1)
-
-modCode <- nimbleCode(
-  {
-    a ~ dnorm(0, sd = 1e4)
-    er ~ dunif(0, 1e4)
-    
-    for (i in 1:10) {
-      
-      log(y[i]) ~ dnorm(a, sd = er)
-      
-    }
-  }
-)
-
-modConsts <- list()
-
-modData <- list(y = rnorm(10, 1e2, 10) %>% exp())
-
-modInits <- list(a = 0, er = 1)
-
-Rmodel <- nimbleModel(code = modCode, 
-                      name = "Rmodel",
-                      constants = modConsts,
-                      data = modData, 
-                      inits = modInits)
-
-
-# node targets
-to_mon <- c("a", "er")
-
-## 3. build MCMC from compiled or not compiled model
-# Parameters monitored (i.e., for which estimates are saved)
-mcmc <- buildMCMC(Rmodel, monitors = to_mon) 
-
-## 4. compile MCMC, requires the Rmodel to be compiled
-Cmodel <- compileNimble(Rmodel)
-Cmcmc <- compileNimble(mcmc, project = Rmodel)
-
-# mcmc parameters
-niter <- 1e5 ; nburnin <- 1e4 ; nsample <- 1e4 ; nchain <- 1
-
-Csamples <- runMCMC(Cmcmc, niter = niter, nburnin = nburnin, 
-                    thin = max(1, round((niter - nburnin) / nsample)), nchains = nchain)
-Csamples <- (if (is_list(Csamples)) {Csamples} else {list(Csamples)}) %>% map(as_tibble)
-
-# plot for convergence check
-Csamples %>% walk(~ .x %>% mcmcplot())
-
-## 7. convert samples to an mcmc object for the coda package
-Csamples_coda <- Csamples %>% "[["(1) %>% as.mcmc()
-
-## summarize output
-Csamples_coda %>% summary()
-

### 2023-05-09 Code/Model_test_nimble_2.R

@@ -1,93 +1,0 @@
-Sys.setlocale("LC_ALL", "en_US.UTF-8")
-
-require(tidyverse)
-require(nimble)
-require(MCMCvis)
-require(mcmcplots)
-
-set.seed(0)
-
-sd_sim = .3
-
-tibble(
-  ts = c(rep(1, 11),
-         rep(2, 5))) %>% 
-  rowid_to_column("x") %>% 
-  mutate(
-    y = c(
-    (13:23 / 3) %>% map_dbl(~ rnorm(1, .x, sd_sim)),
-    (10:14 / 3) %>% map_dbl(~ rnorm(1, .x, sd_sim))
-  ), 
-  N = exp(y)) -> q
-
-JuvCode <- nimbleCode(
-  {
-   # hierarchical prior on lambda max
-    r_max_avg ~ dunif(log(0.01), log(5))
-    r_max_sd ~ dunif(0, 1e2)
-    
-    prior_r_max ~ dnorm(r_max_avg, sd = r_max_sd)
-    log(prior_lambda_max) <- prior_r_max
-    
-    for (r in 1:2) {
-      # 
-      r_avg[r] ~ dnorm(r_max_avg, sd = r_max_sd)
-      # r_avg[r] ~ dunif(log(0.01), log(5))
-      N_int[r] ~ dnorm(0, sd = 1e2)
-      N_sd[r] ~ dunif(0, 1e3)
-      
-      log(lambda[r]) <- r_avg[r]
-    }
-    
-    # LL
-    for (t in 1:R) {
-      
-     N[t] ~ dnorm(N_int[ts_id[t]] + r_avg[ts_id[t]] * year[t], 
-                                sd = N_sd[ts_id[t]])
-    }
-    
-  })
-
-# model data
-JuvConst <- list(R = nrow(q),
-                 ts_id = q$ts)
-
-JuvData <- list(N = q$y,
-                year = q$x)
-
-JuvInit <- list(r_max_avg = log(1), 
-                r_max_sd = 1, 
-                prior_r_max = log(1.5),
-                N_int = rep(0, 2),
-                r_avg = rep(log(1), 2),
-                N_sd = rep(1, 2))
-
-# node targets
-JuvMon <- c("r_max_avg", 
-            "r_max_sd", 
-            "prior_r_max",
-            "r_avg",
-            "prior_lambda_max",
-            "lambda")
-
[…]

### 2023-05-09 Code/Model_vital_rates.R

@@ -43,7 +43,7 @@ scale_x_date_own <- function(coef = 5e-2) {
                  date_minor_breaks = "1 year", 
                  date_labels = "%Y", 
                  expand = c(coef, coef)),
-    guides(x =  guide_axis(angle = 45)),
+    guides(x =  guide_axis(angle = 90)),
     labs(x = NULL) 
   )}
 
@@ -84,7 +84,7 @@ cor_plot <- function(col, title = NULL, subtitle = NULL, limits = NA, breaks = N
                   width = 0, alpha = .5, linewidth = 1.2) +
     labs(title = title,
          subtitle = subtitle,
-         y = "count-based method", x = "hunting-bag-based method") +
+         y = "count-based method", x = "removal-based method") +
     scale_x_continuous(expand = expansion(mult = c(0, 0)), 
                        limits = if (any(is.na(limits))) {
                          c(0, 1.1 * max(unlist(tmp[-1])))
@@ -258,7 +258,7 @@ count_sex_app %>%
               ungroup() %>% 
               filter(tot > 100) %>%
               mutate(`female proportion` = fem / tot, 
-                     method = "Female %\nin hunting bags") %>%
+                     method = "Female %\nin removals") %>%
               select(method, `female proportion`, tot)) %>% 
   mutate(method = as_factor(method)) %>% 
   ggplot(aes(x = method, y = `female proportion`, size = tot)) + 
@@ -380,15 +380,16 @@ counts_1 %>%
   scale_color_manual(name = "", 
                      labels = c("GB population", "FR population"), 
                      values = c_pop) +
-  scale_y_continuous(name = "adult harvest rate",
+  scale_y_continuous(name = "Culling rate index",
                      labels = scales::percent, 
                      breaks = scales::pretty_breaks(), 
                      limits = c(0, 1)) +
   annotate(geom = "text",
            x = c(ymd(19790101), ymd(20020101), ymd(20130101)),
            y = rep(.95, 3), 
-           label = c("no\nharvest", "low\nharvest", "high\nharvest"),
-           size = 4) -> harv_plot; harv_plot
+           label = c("no\nculling", "low\nculling", "high\nculling"),
+           size = 4) +
+  theme(axis.text.x = element_text(angle = 90)) -> harv_plot; harv_plot
 
 frag %>% 
   filter(age == "ad") %>% 
@@ -810,8 +811,8 @@ lambda_ds %>%
          lambda = exp(r_avg)) %>% 
[…]

### 2023-09-27 Code/Draft_ruddy_duck_vital_rates.Rmd
revison A. Caizergues

@@
 Unfortunately, it is insufficient for the basic understanding of the demographic mechanisms and more specifically to assess the extend to which population growth rate is affected by changes in adult survival rather than to variations in reproductive parameters.
-Usual methods to inform survival and recruitment (capture-mark-recapture, game-hunting bag) rely on catching animals in a way or another, which suffers from three main drawbacks.
+Usual methods to inform survival and recruitment, e.g.
+capture-mark-recapture and game-hunting bag, rely on catching animals in a way or another, which suffers from three main drawbacks.
 First, the selectivity of the catching methods can lead to a biased representation of the underlying population structure.
@@
 From populations of Ruddy duck, a species displaying this pattern, this study introduces a non-invasive method based on count data to estimate the respective contribution of recruitment and adult survival on population growth rate.
-Relying on matricial models shows that performing count survey during the appropriate time window to differentiate both plumage types is sufficient to provide long-term time series of the two main demographic components of the population growth rate.
+Relying on population models shows that performing count survey during the appropriate time window to differentiate both plumage types is sufficient to provide long-term time series of the two main demographic components of the population growth rate.
 
   **Keywords**: reproductive success - fecundity - productivity - age ratio - vital rate - duck - Oxyura jamaicensis
-
 output: 
+    bookdown::word_document2:
+      toc: false
     bookdown::pdf_document2:
@@
 As a result, the effectiveness of the eradication of Ruddy ducks and management actions intended to favour the recovery of White-headed ducks were essentially assessed through censuses.
-Censuses, however, prevented assessment of the relative effects or changes in survival and recruitment rate on population growth rate, which is a key towards a proper understanding of the efficiency of management actions.
+Censuses, however, prevented assessment of the relative effects or changes in survival and recruitment rates on population growth rate, which is a key towards a proper understanding of the efficiency of management actions.
 
@@
 Like many other ducks, Ruddy ducks and White-headed ducks both display delayed dichromatism.
-Thereby immature males acquire the typical colourful plumage of their kind over the course of the interbreeding season, which is typically in late winter.
-Immature males are thus confounded to immature and adult females before their moult (Figure \@ref(fig:flock)).
-Delayed dichromatism explains why the apparent proportion of females in the winter counts is always greater than the true proportion of females that can be assessed in removals (Figure \@ref(fig:pfem)).
+Thereby, immature males are confounded to immature and adult females until a first moult occurring in mid-winter for the early ones [@Baldassarre2014], but most likely latter for most of them.
+Immature males acquire then the typical colourful breeding plumage of their kind in spring [@Baldassarre2014].
+Delayed dichromatism explains why the apparent proportion of females in winter counts is always greater than the true proportion of females that can be assessed in counts of late winter or in removals (Figures \@ref(fig:flock) & \@ref(fig:pfem)).
 Thus, it is possible to assess the proportion of immatures within a population by monitoring the seasonal evolution of apparent sex ratio and assuming an even sex ratio at birth [@Bellrose1961; @Blums1996], comparable mortality rates between females and males among juveniles [@Ramula2018; @Wood2021], and unchanging adult sex ratio over the monitoring period [@Devineau2010].
@@
 A comparison over months in Great Britain showed no difference in proportion of males in adults, so data from adults collected over the whole year were used to estimate the proportion of males in adults.
-A comparison of proportion of males in adults among years with more than 500 samples did not show significant interannual differences.
+A comparison of proportion of males in adults among years with more than 500 samples did not exhibit significant interannual differences.
 Even if the adult sex ratio may vary over a long time range, it is relatively stable over a few years [@Wood2021].
@@
 |:---|:----|:----------------------------|
-|$AM$|Data|Total number of adult males in hunting bags|
-|$AF$|Data|Total number of adult females in hunting bags|
+|$AM$|Data|Total number of adult males in removals|
+|$AF$|Data|Total number of adult females in removals|
 |$p(m|a)$|Parameter|Proportion of males in adults (or probability of being a male knowing it is an adult)|
@@
 The lower limit corresponds to the upper range of survival rates in literature for waterfowl species of similar weight, the upper limit was set to 1 because survival rates of long-life waterfowl species are very high [@Nichols1997; @Krementz1997; @Buxton2004; @Nichols1992].
-Recruitment rate is more tricky to validate because it is not upper bounded as it is defined on $[0;\infty[$ and is neither explicitly informed in the literature.
+Recruitment rate is more tricky to validate because it is not upper bounded as it is defined on $[0;\infty[$ and is neither explicitly informed in literature.
 Only some components of the recruitment rate are generally described, such as nesting rate, nesting success, clutch size, hatching success, early stage survival, but rarely survival between the juvenile stage and the first breeding season, which is necessary to inform recruitment rate [e.g.
@@
 Adult survival rate and recruitment rate, i.e.
-the two components of the population growth rate, were also estimated by the method successfully(Figure \@ref(fig:vital)).
+the two components of the population growth rate, were also estimated by the method successfully (Figure \@ref(fig:vital)).
 
@@
 
-The male proportion in adults, which is a prerequisite to infer on the immature proportion, is estimated at 0.60 [0.59; 0.61].
-The immature proportion varies from 0.16 [0.07; 0.24] to 0.54 [0.44; 0.62].
-The lower values are similar for the two populations, but the range of the immature proportions in the GB population is far more limited (upper value : 0.30 [0.14; 0.45]).
-Conversely from the FR population, there is no significant difference from a year to another because all estimates overlap each other.
-We can notice that the immature proportion for the GB population is quite stable over the time series available, but a slight decrease is observed for the FR population.
+The proportion of males in adults, which is a prerequisite to infer the proportion of immatures, was estimated at 0.60 [0.59; 0.61].
+The immature proportion ranged from 0.16 [0.07; 0.24] to 0.54 [0.44; 0.62].
+The lower values were similar for the two populations, but the range of the proportions of immatures in the GB population was far more limited (upper value : 0.30 [0.14; 0.45]).
+As opposed to the FR population, there were no significant differences from a year to another because all estimates overlapped with each other.
+The proportion of immatures was relatively stable over time in the GB populations, while a slight decrease was observed for the FR population.
 
-The adult survival rate varies from 0.33 [0.29; 0.37] to 1.01 [0.89; 1.14].
-All estimates are not significantly outside the range of a survival rate defined without immigration [0; 1].
-There is no trend in adult survival for both populations, but we can notice two main differences.
-The GB population has the lowest values of adult survival rate, and even if there are significant difference among years for both populations, the variability for the GB population is a bit lower.
+Adult survival rate ranged from 0.33 [0.29; 0.37] to 1.01 [0.89; 1.14].
+No estimates were significantly outside the range of a survival rates defined without immigration [0; 1].
+No trend in adult survival rate was recorded in any of the two populations, although two main differences could be observed.
+The GB population had the lowest values of adult survival rate, and even if there were significant difference among years for both populations, the variability for the GB population was slightly lower.
 
-The recruitment rate varies from 0.09 [0.04; 0.13] to 0.75 [0.63; 0.87].
-All estimates are above 0, meaning there is no senseless estimation.
-The maximum recruitment rate being 0.68 [0.36; 0.78] (see Section \@ref(MAX)), all estimates are not significantly outside the range of a recruitment rate defined without immigration [0; 0.78].
-Similarly to the immature proportion, there is no trend for the GB population, but a decrease is observed for the FR population even if it is more noisy than for the immature proportion.
-Similarly to the survival rate, the GB population has low values of recruitment rate, and even if there are significant difference among years for both populations, the variability for the GB population is far lower than for the FR population.
+Recruitment rate ranges from 0.09 [0.04; 0.13] to 0.75 [0.63; 0.87].
+All estimates were above 0, meaning there was no senseless estimation.
+The maximum recruitment rate being 0.68 [0.36; 0.78] (see Section \@ref(MAX)), no estimates were significantly outside the range of recruitment rates defined without immigration [0; 0.78].
+Similarly to the immature proportion, no trend was observed for the GB population, but recruitment rate decreased for the FR population even if it was more noisy than for the immature proportion.
+As for survival rate, the GB population had low values of recruitment rate, and even if significant differences were observed among years for both populations, the variability for the GB population was far lower than for the FR population.
 
-The lower variability of both vital rates for the GB population explains its steady population growth over the considered time series compared to the FR one (Figure \@ref(fig:count)).
-In the FR population, the range of the recruitment rate (min/max difference: 0.66) is higher than the range of the survival rate (min/max difference: 0.58).
-Conversely, the range of the survival rate (min/max difference: 0.46) in the GB population is far higher than the range of the recruitment rate (min/max difference: 0.08).
-This reflects that the variability of the population growth rate for the FR population is mostly driven by the recruitment rate whereas the main source of variability is the adult survival for the GB population.
+The lower variability of both vital rates for the GB population explained its steady population growth over the considered time series compared to the FR one (Figure \@ref(fig:count)).
+In the FR population, the range of recruitment rates (min/max difference: 0.66) was higher than the range of survival rates (min/max difference: 0.58).
+Conversely, the range of survival rates (min/max difference: 0.46) in the GB population was far higher than the range of recruitment rate (min/max difference: 0.08).
+This reflected that the variability of the population growth rate for the FR population was mostly driven by recruitment rate whereas the main source of variability was the adult survival for the GB population.
 
@@
 
-From the 5 years available in the GB population, immature proportions estimated from the count-based method is positively correlated to estimates from hunting-bag data (Figure \@ref(fig:cor)).
-The hypothesis that the proportion of female-like individuals in the wintering populations reflects the age structure of these populations is then supported by this result.
-This correlation is even stronger when considering vital rates (Figure \@ref(fig:cor)).
-This result is expected since the two methods estimating these vital rates have a component in common, the growth rate of the population (see Equation \@ref(eq:vr)).
+From the five years available in the GB population, proportions of immatures that were estimated from the count-based method were positively correlated to estimates from removal data (Figure \@ref(fig:cor)).
+The hypothesis that the proportion of female-like individuals in the wintering populations reflected the age structure of these populations was then supported by this result.
+This correlation was even stronger when considering vital rates (Figure \@ref(fig:cor)).
+This stronger correlation was expected since the two methods estimating these vital rates had a component in common, the growth rate of the population (see Equation \@ref(eq:vr)).
 
-It is noticeable that a 1:1 correlation is never obtained.
-Immature proportion is lower in counts than in hunting bags.
-This result is not surprising because immature individuals are expected to be more vulnerable to hunting than adults and might then be over represented in hunting bag data.
-As a consequence, the correlation coefficients for both vital rates differ also from 1.
-This difference between the two approaches does not question the previous validation of the ability of the count method to track the interannual variability of the vital rates.
+It was noticeable that a 1:1 correlation was never obtained.
+Proportion of immatures was lower in counts than in removals.
+This result was not surprising because immature individuals were expected to be more vulnerable to culling than adults and might then be over represented in removal data.
+As a consequence, the correlation coefficients for both vital rates also differed from 1.
+This difference between the two approaches did not question the previous validation of the ability of the count method to track the interannual variability of the vital rates.
 
@@
 
-## Response of vital rates to the harvest strategy {#MAX}
+## Response of vital rates to culling strategies {#MAX}
 
-Under no harvest pressure, the maximum growth rates of the two populations are very similar (Figure \@ref(fig:mgr)), respectively 1.45 [1.36; 1.55] and 1.52 [1.25; 1.88] for the GB and the FR population.
-It corresponds to an increase of 45% per year and 52% per year respectively.
+Under no culling pressure, maximum growth rates were very similar for the two populations (Figure \@ref(fig:mgr)), respectively 1.45 [1.36; 1.55] and 1.52 [1.25; 1.88] for the GB and the FR population.
+It corresponded to an increase of 45% per year and 52% per year, respectively.
 Based on literature, we assumed a maximum survival rate similar for both populations 0.85 [0.7; 1].
@@
 
-Even if a significant part of culled individuals were not aged for some years, the proxies of the harvest pressure show significant shifts over time (Figure \@ref(fig:sup)).
-This led to consider a high harvest pressure period from 2006 with no clear difference between the two populations because the signal is very noisy.
-Even if the harvest pressure seems comparable, the harvest strategies differ significantly between the two populations, see Section \@ref(HARV).
-Under high harvest pressure, the growth rate drops to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for the GB and the FR population respectively, which corresponds to a decrease of 44% per year and 3% per year.
-This result shows that the GB population significantly decreases during the high harvest pressure whereas the FR population is stabilized.
+Even if a significant part of culled individuals were not aged for some years, the proxies of the culling pressure exhibit significant shifts over time (Figure \@ref(fig:sup)).
+This led to consider a high culling pressure period from 2006 onwards with no clear difference between the two populations because the signal was very noisy.
+Even if the culling pressure seemed comparable, the culling strategies differed significantly between the two populations, see Section \@ref(HARV).
+Under high culling pressure, the growth rate dropped to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for the GB and the FR population, respectively, which corresponded to a decrease of 44% per year and 3% per year.
+This result showed that the GB population significantly decreased during the high culling pressure whereas the FR population was stabilized.
 
-The adult survival rate under high harvest pressure is 0.44 [0.42; 0.46] and 0.62 [0.59; 0.64] for the GB and FR populations respectively, when the recruitment rate decreases to 0.13 [0.11; 0.15] and 0.32 [0.29; 0.34] respectively.
-The drop of the growth rate of the GB population corresponds to a similar average drop in adult survival rate (-0.41) and in recruitment rate (-0.47).
-The stabilisation of the FR population is more due to a decrease of the recruitment rate (-0.36) than from a drop of the adult survival (-0.23).
-For both populations, the response to the harvest pressure is significant, meaning that the compensation of the population to harvest mortality is limited if not null.
-Vital rates of the GB population are more affected than the FR population.
-Unsurprisingly, the FR harvest strategy focusing on the postbreeding period affects mostly the recruits.
-However, the prebreeding harvest strategy applied on the GB population that targets adults performs to deplete strongly the recruitment to a very low level (0.13 recruits produced per breeder).
+Adult survival rate under high culling pressure was 0.44 [0.42; 0.46] and 0.62 [0.59; 0.64] for the GB and FR populations, respectively, when the recruitment rate decreases to 0.13 [0.11; 0.15] and 0.32 [0.29; 0.34], respectively.
+The drop in growth rate of the GB population corresponded to a similar average drop in adult survival rate (-0.41) and in recruitment rate (-0.47).
+The stabilisation of the FR population was caused by a decrease of recruitment rate (-0.36) more than a drop in adult survival (-0.23).
+Overall, vital rates of the GB population were more affected than the FR population.
 
@@
 
-The delayed sexual dichromatism in a waterfowl species is used here to develop and test the validity of a new non-invasive method aiming to break down the apparent growth of a population into its two structural components: the adult survival rate and the recruitment rate.
-Disentangling the contribution of these two vital rates is a key to understand the response of a population to management measures.
-The development of the method holds on a unique dataset of two comparable populations of a same species, the Ruddy duck.
-Both populations were tracked from their introduction in a similar environment to a period of intense harvest pressure.
-This common history trajectory enables to evaluate the maximum growth of this species, but also its response to harvest pressure.
-As the harvest strategy differs between the two populations, we detail the similarities and the discrepancies of the responses of the vital rates, and open the discussion to the generalization of this method and its implications for waterfowl management.
+The delayed sexual dichromatism in a waterfowl species was used here to develop and test the validity of a new non-invasive method aiming to break down the apparent growth of a population into its two structural components: adult survival rate and recruitment rate.
+Disentangling the contribution of these two vital rates is a key to understand the response of a population to management measures and potentially adapt these accordingly [@Williams2002].
+The development of the method held on a unique dataset of two comparable populations of the same species, the Ruddy duck.
+Both populations were tracked from their introduction in a similar environment to a period of intense culling pressure.
+This common history trajectory enabled to evaluate the maximum growth rate of this species, but also its response to culling pressure.
+As the culling strategy differed between the two populations, we detailed the similarities and the discrepancies of the responses of vital rates, and opened the discussion to the generalization of this estimation method and its implications for waterfowl management.
 
@@
 
-The vital rate estimates of the two populations belongs to the same range, which demonstrate a certain consistency in the results obtained from the method.
-The higher variability observed in France is not unexpected because the FR time series covers a larger spectrum of harvest pressure than the GB time series.
-There is no temporal correlation between the two populations, suggesting that fluctuations of the vital rates are more likely influenced by regional factors than large-scale ones.
-This is consistent with the literature because both nesting/hatching success and juvenile survival are conditional to the onset of laying, which is tightly related to local weather parameters, e.g.
+The vital rate estimates of the two populations belonged to the same range, which demonstrates a certain consistency in the results obtained from the method.
+The greater variability observed in France is not unexpected because the FR time series covered a larger spectrum of culling pressure than the GB time series.
+There was no temporal correlation between the two populations, suggesting that fluctuations of the vital rates were more likely influenced by regional factors than large-scale ones.
+This is consistent with literature because both nesting/hatching success and juvenile survival in waterfowl species are conditional to the onset of laying, which is tightly related to local weather parameters, e.g.
 spring temperature, cumulative rainfall, and water levels [@Folliot2017; @Blums2004; @Dzus1998].
 
-The estimates of adult survival rates correspond to values observed in the literature on species of similar weight [@Nichols1997; @Krementz1997; @Buxton2004].
-This reflects that there is no apparent scaling issue of this method based on count data.
-The assumption that male-like individuals correspond only to adult males is then acceptable.
-This result supports that the count-based method is poorly biased if it is.
-However, some adult survival rate values are outside the range of expected values even if they are never significantly over 1.
-A likely reason to these extreme values is a corruption of the closed population assumption on some years.
+Count surveys used to test the method occurred in mid-winter, when the first immature males were suspected to moult and start to look like adult males.
+This might partially corrupt the assumption that male-like individuals corresponded only to adult males.
+Even though no data on adult survival rate were directly available [@Baldassarre2014], the values estimated from the method developed inhere correspond to what is observed in literature for species of similar weight [@Nichols1997; @Krementz1997; @Buxton2004].
+Then, the results reflect that there was no major scaling issue using count data from mid-winter for European Ruddy duck populations.
+In any case, the time window of the count surveys must be chosen with care, i.e.
+just before the moult of immature males, to limit the bias of the outputs.
+On a few years, the values of adult survival rate were outside the range of expected values even if they were never significantly over 1.
+A likely reason to these extreme values is a corruption of the closed population assumption during some years.
 Indeed, if there is an arrival from another Ruddy duck population, the true recruitment cannot explain by itself the population growth and leads to an adult survival combining true adult survival and adult immigration.
-For the FR population, upper outliers of recruitment rates and survival rates are observed in 2002 and 2012, suggesting immigration events on these years.
+For the FR population, upper outliers of recruitment rates and survival rates were observed in 2002 and 2012, suggesting immigration events during these years.
 
-The theoretical maximum of the population growth rate is a potential which is determined by intrinsic life history traits of a species [@Dmitriew2010].
+The theoretical maximum of the population growth rate is a potential which is determined by intrinsic life history traits of a species [@Dmitriew2010; @Niel2005].
 This potential is expected to be similar to all Ruddy duck populations anywhere in the world.
-Realised maximum growth rates that are observed in this study are very close for the two populations.
-It suggests that the Ruddy duck species reached its intrinsic biological reproduction limit, at least given the environmental conditions of the western Palearctic, i.e.
+Realised maximum growth rates that were observed in this study were very close for the two populations.
+This suggests that the Ruddy duck species reached its intrinsic biological reproduction limit, at least given the environmental conditions of the western Palearctic, i.e.
 about 0.6 recruits per breeder.
-The consistency of this intermediate result suggests that the population size estimation is quite accurate for both populations and demonstrates that their changes are well tracked even with a few individuals.
-The average recruitment rates estimated from the count-based method, with and without harvest, are never higher to the proxies of maximum recruitment rate for both populations.
-The assumption on the accuracy of the count-based method is then again not excluded from the consistency of this result.
+The consistency of this intermediate result suggests that the population size estimation was quite accurate for both populations and demonstrates that their changes were well tracked even with relatively few individuals.
+The average recruitment rates estimated from the count-based method, with and without culling pressure, were never higher to the proxies of maximum recruitment rate for both populations.
+The assumption on the accuracy of the count-based method is then again not excluded considering the consistency of this result.
 
-Assuming that hunting-bag data provide a good picture of the year-to-year variability of age structure of a population, the strong correlation with the outputs of our method demonstrates its ability to track interannual fluctuations.
-This result is robust because such correlation is obvious despite being based on a short time series.
-This strong relationship is also satisfying because it demonstrates that the assumption of a constant adult sex ratio is not corrupted.
-The temporal autocorrelation of the adult sex ratio might be explained because it integrates many age-cohorts, which makes it structurally strong.
-Even if significant variations of the adult sex ratio can still occur over long time periods, this result demonstrates that it is not necessary to monitor and update it on a yearly base.
+Assuming that removal data provided a good picture of the year-to-year variability of age structure of a population, that is, any bias of shooting towards a given age or sex category is relatively stable over time, the strong correlation with the outputs of our method demonstrates its ability to track interannual fluctuations.
+This result is robust because such correlation was obvious despite being based on a short time series.
+This strong relationship was also satisfying because it demonstrates that the assumption of a constant adult sex ratio is not corrupted.
+The temporal autocorrelation of the adult sex ratio might be explained by the fact that it integrates many age-cohorts, which makes it structurally strong.
+Even if significant variations of the adult sex ratio can still occur over long time periods, this result demonstrates that it is not necessary to monitor and update it on a yearly basis.
 The adult sex ratio found here on a restricted time period is in line with the Ruddy duck in its native area [0.62 in @Bellrose1980] and with other duck species [@Wood2021].
-The count-based method always provides recruitment rates significantly lower than the method based on hunting-bag data by a factor close to two, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable than adults to hunters [@Bellrose1980; @Fox2014].
-The gap of vulnerability among ages observed here is realistic because the harvest-induced mortality on first-year immatures compared to adults is higher by a factor ranging from 1.3 to 2.6 for hunted duck species in North America [@Bellrose1980].
+The count-based method always provided recruitment rates significantly lower than the method based on removal data by a factor close to two, a result that is in line with the hypothesis that the shooting is generally selective towards immature individuals in waterfowl because they are more vulnerable to hunters than adults [@Bellrose1980; @Fox2014].
+The gap of vulnerability among ages observed here is realistic because the culling-induced mortality on first-year immatures compared to adults is higher by a factor ranging from 1.3 to 2.6 for hunted duck species in North America [@Bellrose1980].
 
-Overall, even if the true values of vital rates are not known to test properly the accuracy of the count-based method, a beam of arguments supports that the method is not heavily biased.
-This consists in a big difference with the method based on hunting-bag data as demonstrated in @Fox2014, even when one considers an age structure picture that is limited to the end of the hunting season to limit the bias [@Fox2016].
+Overall, even if the true values of vital rates are not known to properly test the accuracy of the count-based method, a beam of arguments supports that the method is not heavily biased.
+This consists in a big difference with methods based on removal/hunting bag data as demonstrated in @Fox2014, even when one considers an age structure picture that is limited to the end of the hunting season to limit the bias [@Fox2016].
 
@@
 
-Both populations respond to the harvest pressure by a large drop of their growth, meaning that the the harvest effort is efficient to control this species.
-The response of the GB population differs from the FR one because it displays a strong decrease when the FR population is stabilized.
-The analysis of the growth alone does not provide clues to understand the mechanisms explaining the difference of the responses between the two populations.
-Indeed, the bigger drop of growth of the GB population might be explained either by a decrease of the adult survival or of the reproduction success, or even of both components.
+Both populations responded to culling pressure by a large drop of their growth rate, meaning that the compensation of the population to culling mortality was limited if not null.
+The eradication effort as implemented here was therefore efficient to impact this species.
+The response of the GB population differed from the FR one because it showed a strong decrease when the FR population got stabilized.
+The analysis of the growth rate alone did not provide clues to understand the mechanisms explaining the difference of the responses between the two populations.
+Indeed, the bigger drop of growth rate of the GB population might be explained either by a decrease of adult survival or reproduction success, or both components.
 
-The vital rates that are estimated thanks to the count-based method developed in here demonstrates that the harvest pressure affects on average negatively both the adult survival and the recruitment rate for the two populations.
-The harvest pressure occurring mostly before the breeding period for the GB population targets by definition future breeders, and consequently induces mortality on these individuals and prevent them from reproducing.
-We observe that adult survival and recruitment rate decrease equally, which is a result in line with the expectation of this harvest strategy.
-For the FR population, the harvest pressure occurs mostly during the postbreeding period.
+The vital rates that were estimated thanks to the count-based method supported that the eradication pressure affected negatively both adult survival and recruitment rates in the two populations on average.
+The culling pressure occurring mostly before the breeding period for the GB population targeted by definition future breeders, and consequently induced mortality on these individuals and prevented them from reproducing.
+We observed that adult survival and recruitment rates decreased equally, which is a result in line with the expectation of this culling strategy.
+For the FR population, the culling pressure mostly occurred during the postbreeding period.
 This strategy should induce a higher mortality on immatures as they are expected to be more vulnerable to hunting [@Bellrose1980; @Fox2014].
-However, this anthropogenic mortality on immatures might be compensated by a higher survival of the future recruits because the competition for resources decreased [@Cooch2014].
-The recruitment rate of the FR population being more affected by 50% than the adult survival, the results support the hypothesis that no or only a limited compensation occurs.
+However, this anthropogenic mortality on immatures might be compensated by a higher survival of the future recruits because competition for resources is decreased [@Cooch2014].
+Recruitment rate of the FR population being more affected by 50% than adult survival, the results support the hypothesis that no or only a limited compensation occurred.
 This conclusion is even strengthened by the result on the GB population.
-Indeed, we should expect a decrease of the recruitment rate lower than the adult survival rate if immatures would take advantage of additional resources that unexploited by culled adult individuals.
-The equal decrease of the two vital rates supports then that there is no compensation.
-This might be explained because the population size is far lower than the carrying capacity of the available habitats, so there is no competition for resources that potentially enhances the survival of the remaining immature individuals [@Peron2013].
+Indeed, we should expect a decrease of recruitment rate lower than adult survival rate if immatures would take advantage of additional resources getting unexploited by culled adult individuals.
+The equal decrease of the two vital rates supports that there is no such compensation.
+This might be explained by the fact that the population size was far lower than the carrying capacity of the available habitats, so there was no competition for resources that potentially enhanced the survival of the remaining immature individuals [@Peron2013].
 
-Considering a population following a compensatory functioning, the FR strategy is counter-productive.
-As studied Ruddy duck populations are expanding, it seems that the postbreeding harvest strategy still provides some results to control such populations.
-However, the analysis of the year-to-year variability of the recruitment rate demonstrates that the prebreeding strategy buffers strongly the recruitment and maintains it to a low level every single year whereas the recruitment rate of the postbreeding strategy is highly variable.
-This might be because harvesting future breeders ensures to avoid the production of potential recruits whereas the postbreeding harvest strategy cannot prevent and dampen exceptional reproduction events.
-The postbreeding strategy relies also on a detectability of immatures highly variable because these individuals spread over many reproduction habitats more or less accessible whereas the prebreeding strategy targets individuals that are more easily localized because they usually gather in open water on a limited number of wintering spots [@Johnsgard1996].
+If we were in a situation where shooting mortality was fully compensatory to natural mortality, the FR strategy would be counter-productive.
+However, since the studied Ruddy duck populations are expanding, it seemed that the postbreeding culling strategy still provided some results to control such populations.
+However, the analysis of the year-to-year variability of recruitment rate demonstrates that the prebreeding strategy strongly buffered the recruitment and maintained it to a low level every single year, whereas recruitment rate of the postbreeding strategy was highly variable.
+This might be because culling future breeders ensured to avoid the production of potential recruits whereas the postbreeding culling strategy could not prevent and dampen exceptional reproduction events.
+The postbreeding strategy is also affected by the highly variable detectability of immatures, as these individuals spread over many reproduction habitats with variable accessibility to observers/shooters, whereas the prebreeding strategy targeted individuals that were more easily localized because they usually gathered in open water on a limited number of wintering spots [@Johnsgard1996].
 
-The response to the harvest pressure of the two Ruddy duck populations demonstrate that it is necessary to account for the time window over which the harvest occurs to produce proper predictions.
-Predicting the impact of a harvest pressure on a waterfowl population is then not only a question about how big is the harvest effort, but also mostly when this effort occur within a year.
-More generally, disentangling the adult survival to the recruitment rate brings capital information to understand the process of harvest on population dynamics and brings key knowledge to improve the conservation of hunted species or the control of invasive species.
+The response of the two Ruddy duck populations to culling pressure demonstrates the necessity to account for the culling season to produce proper predictions.
+Predicting the impact of a harvest pressure on a waterfowl population is then not only a question about how big is the harvest effort, but also mostly when this effort occurs within a year [@Kokko1998].
+More generally, disentangling role of adult survival from that of recruitment rate brings capital information to understand the process of harvest on population dynamics and brings key knowledge to improve the conservation of hunted species or the control of invasive species.
 
@@
 Tracking variations in abundance is a common tool to determine the conservation status of a population [e.g @Folliot2022].
-However, the abundance alone does not provide information on the mechanisms behind the changes of population growth.
-Identifying the determinants requires to dig into the components structuring the population growth to provide relevant tools to managers [e.g.
+However, abundance alone does not provide information on the mechanisms behind the changes of population growth [@Williams2002].
+Identifying the determinants requires to dig into the components structuring population growth to provide relevant tools to managers [e.g.
 @Austin2000].
-Gaining knowledge on adult survival and the recruitment rate often involves capture mark recapture surveys [e.g.
+Gaining knowledge on adult survival and recruitment rates often involves capture mark recapture surveys [e.g.
 @Arnold2018].
@@
 @Souchay2016] and prevents from implementing such method to small populations with a poor conservation status for instance.
-The method introduced in here bypasses these caveats to provide annual estimates of vital rates.
+The method introduced here bypasses these caveats to provide annual estimates of vital rates.
 
 The immature proportion, or age ratio, is commonly used to describe the renewal capacity of a waterfowl population and to track the changes of its productivity [@Smith2001; @Bellrose1980; @Robertson2008; @Rodway2015; @Zimmerman2010].
-The recruitment rate remains however a better indicator than the immature proportion/age ratio, the latter having only the advantage to be more directly accessible.
+The recruitment rate however remains a better indicator than the immature proportion/age ratio, the latter only having the advantage of being more directly accessible.
 Indeed, the immature proportion/age ratio does not account for the dynamics of the population and reflects the population productivity only if the population growth is steady.
-For instance, let's consider a breeding population of 100 individuals reaching the next year 100 individuals, which then drops to 60 individuals the second year.
-If there is 50% of immature individuals each year, corresponding to a 1:1 age ratio, we may conclude that the productivity is stable over time.
+For instance, let us consider a breeding population of 100 individuals reaching 100 individuals the next year, which then drops to 60 individuals the second year.
+If there are 50% immatures each year, corresponding to a 1:1 age ratio, we may be tempted to conclude that the productivity is stable over time.
 But it actually corresponds to a strong decrease of the productivity (-40%) because the recruitment rate evolved from 0.5 recruits per breeder the first year to 0.3 recruits per breeder the second year.
 This demonstrates that the immature proportion/age ratio suffers from caveats that can be misleading for a manager.
-According to @Blums2004, the recruitment for diving ducks is not related to the fecundity, but depends mostly on the juvenile survival, the other component of the recruitment that is mostly driven by the weather conditions.
+According to @Blums2004, recruitment in diving ducks is not related to fecundity, but depends mostly on juvenile survival, the other component of the recruitment that is mostly driven by weather conditions.
 Studies on other birds also conclude of the poor correlation between fecundity and recruitment [e.g.
@@
 
-Applying the method implemented in this study to any other waterfowl species implies that the adult males can be distinguished from the other individuals at distance.
-The latter in the season the dichromatism is observable, the better it is to ensure a relevant picture of the age structure of the population and then good estimates of the adult survival and the recruitment rate.
-Modifying standard monitoring protocols to distinguish male-like from female-like individuals is almost costless but provides substantial increases in the efficiency and usefulness of monitoring results in conservation [@Nichols2006].
+Applying the method implemented in this study to any other waterfowl species implies that the adult males can be distinguished from the other individuals at a distance.
+The latter in the season the dichromatism is observable, the better it is to ensure a relevant picture of the age structure of the population and then good estimates of adult survival and recruitment rates.
+This pattern occurs in most of stifftail duck species [@Johnsgard1996], and in many other duck species, among with Tufted duck, *Aythya fuligula*, Black scotter, *Melanitta nigra americana*, Common Goldeneye, *Bucephala clangula americana* [@Johnsgard1978; @Bellrose1980].
+Modifying standard monitoring protocols to distinguish male-like from female-like individuals is almost costless but would provide substantial increases in the efficiency and usefulness of monitoring results in conservation [@Nichols2006].
 
@@
 This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the *Office Français de la Biodiversité*, and Jean-Marc Gillier from the *Société Nationale pour la Protection de la Nature*.
-The authors acknowledge all the contributors of the data collection; in the Great Britain, Wildfowl & Wetlands Trust staff past and present, particularly Baz Hughes, Peter Cranswick and Colette Hall, and all the project workers at the Animal and Plant Health Agency and its predecessors; in France, Office Français de la Biodiversité staff past and present, especially Vincent Fontaine, Denis Lacourpaille, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface, and the whole team of the Société Nationale pour la Protection de la Nature of the Grand Lieu lake.
+The authors acknowledge all the contributors of the data collection; in Great Britain, Wildfowl & Wetlands Trust staff past and present, particularly Baz Hughes, Peter Cranswick and Colette Hall, and all the project workers at the Animal and Plant Health Agency and its predecessors; in France, Office Français de la Biodiversité staff past and present, especially Vincent Fontaine, Denis Lacourpaille, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface, and the whole team of the Société Nationale pour la Protection de la Nature of the Grand Lieu lake.
 
@@
 
-(ref:flock) Typical observation of a Ruddy duck flock during winter: 10 female-like individuals with a whitish striped cheek, 4 male-like individuals with a white cheek and a black cap, 3 unidentified individuals (1^st^, 4^th^, and 10^th^ from left) © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
+(ref:flock) Typical observation of a Ruddy duck flock during winter: ten female-like individuals with a whitish striped cheek, four male-like individuals with a white cheek and a black cap, three unidentified individuals (1^st^, 4^th^, and 10^th^ from left) © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
 
-(ref:pfem) Proportion of females in removals vs in counts; For the first category, one data point corresponds to the female proportion estimated from direct sex identification of the Ruddy ducks culled over one year in a population; for the second category, one data point correponds to the proportion of individuals wearing female-like plumage in a population counted in winter; the difference of female proportion between the two approaches is due to immature males that look like females before moulting; this difference is at the root of the method introduced inhere to disentangle survival from recruitment
+(ref:pfem) Proportion of females in removals vs in counts; For the first category, one data point corresponds to the proportion of females estimated from direct sex identification of the Ruddy ducks culled over one year in a population; for the second category, one data point corresponds to the proportion of individuals wearing female-like plumage in a population counted in winter; the difference of proportion of females between the two approaches is due to immature males that look like females before moulting; this difference is at the root of the method introduced inhere to disentangle adult survival from recruitment
 
-(ref:vital) Evolution of the proportion of immatures in the population and the vital rates in light of the culling pressure; bars define the 95% confidence intervals
+(ref:vital) Changes in the proportion of immatures and the vital rates in light of the culling pressure for both populations; bars define the 95% confidence intervals
 
-(ref:cor) Comparison of the parameter estimates between the count-based method and removal-based method; only 5 years for the GB population were available; bars define the 95% confidence intervals; the red arrow defined the average difference in proportion between the two methods
+(ref:cor) Comparison of the parameter estimates between the count-based method and removal-based method; only five years for the GB population were available; bars define the 95% confidence intervals; the red arrow indicates the average difference between the two methods
 
-(ref:mgr) Comparison between average population growth rate under no pressure vs under high culling pressure; the culling effort for the GB population occurs moslty during the prebreeding period, whereas it focuses mostly during and after the breeding period for the FR population (see \@ref(HARV) for details); the average vital rates during the maximum growth time period are proxies deduced literature and estimates of maximum population growth estimates (see \@ref(VAL) for details); the vital rates under high culling pressure are average values over a time period observing a consistent culling pressure; bars define the 95% confidence intervals
+(ref:mgr) Comparison between average population growth rate under no pressure vs under high culling pressure; the culling effort for the GB population moslty occurred during the prebreeding period, whereas it mostly focused during and after the breeding period for the FR population (see \@ref(HARV) for details); the average vital rates during the time period of maximum growth were proxies deduced from literature and estimates of maximum population growth (see \@ref(VAL) for details); the vital rates under high culling pressure were average values over a time period observing a consistent culling pressure; bars define the 95% confidence intervals
 
-(ref:sup) Definition of 3 periods of different culling pressures from an index based on an estimation of the adult culling rate; as only half of the culled individuals were aged in France before 2009, we applied the average age ratio over 2009-2019 on unaged individuals to provide an estimate of the number of adults in removals before 2009; the culling rate increased over time, we then split the time series in 3 categories : we defined a "no culling" period before 1999 because the culling rate for both countries was mostly null and always below 10%, we defined a "low culling" between 1999 and 2005 because the culling rate varies around 20% for both countries, the culling rate then raises significantly for both countries, so we defined a "high culling" period from 2006
+(ref:sup) Definition of three periods of different culling pressures from an index based on an estimation of adult culling rate; as only half of the culled individuals were aged in France before 2009, we applied the average age ratio over 2009-2019 on unaged individuals to provide an estimate of the number of adults in removals before 2009; culling rate increased over time, we therefore split the time series in three categories : we defined a "no culling" period before 1999 because the culling rate for both countries was mostly null and always below 10%, we defined a "low culling" between 1999 and 2005 because the culling rate varied around 20% for both countries, the culling rate then raised significantly for both countries, so we defined a "high culling" period from 2006 onwards

### 2023-11-30 Code/Draft_ruddy_duck_vital_rates.Rmd
before revision AC

@@
 link-citations: true
-title: "Exploiting delayed dichromatism to disentangle the effects of adult survival and recruitment on the population dynamics of waterfowl"
+title: "Delayed dichromatism as a convenient tool to disentangle the effects of survival and productivity on population growth in waterfowl"
 author: 
@@
 abstract: |
-  Monitoring the number of individuals is by far the most popular strategy for investigating the environmental factors ruling the population dynamics and for measuring the efficacy of management actions targeting population recovery, control or eradication.
-Unfortunately, it is insufficient for the basic understanding of the demographic mechanisms and more specifically to assess the extend to which population growth rate is affected by changes in adult survival rather than to variations in reproductive parameters.
-Usual methods to inform survival and recruitment, e.g.
-capture-mark-recapture and game-hunting bag, rely on catching animals in a way or another, which suffers from three main drawbacks.
-First, the selectivity of the catching methods can lead to a biased representation of the underlying population structure.
-In practice, catching and releasing animals is also an invasive approach implying additional stress to the individuals.
-Furthermore releasing may simply be forbidden for invasive alien species.
-In many waterfowl species, a sexual dichromatism is observed in adults whereas immatures of both sexes display a cover-up plumage similar to adult females.
-From populations of Ruddy duck, a species displaying this pattern, this study introduces a non-invasive method based on count data to estimate the respective contribution of recruitment and adult survival on population growth rate.
-Relying on population models shows that performing count survey during the appropriate time window to differentiate both plumage types is sufficient to provide long-term time series of the two main demographic components of the population growth rate.
+  Monitoring the number of individuals is by far the most popular strategy for studying the environmental factors that rule population dynamics and for measuring the efficacy of management actions aimed at population recovery, control or eradication.
+Monitoring the number of individuals is unfortunately inefficient in assessing the mechanisms underlying demographic processes and, more specifically, the extent to which population growth rate is affected by changes in adult survival rather than variations in reproductive parameters.
+Common methods to inform survival and recruitment, such as capture-mark-recapture and game-hunting bag, rely on catching animals, which has three main disadvantages.
+First, the selectivity of capture methods can lead to a biased representation of the underlying population structure.
+Second, catching and releasing animals is quite invasive, and causes stress, thus posing an additional threat to often endangered species.
+Third, release may simply be impossible, especially for invasive alien species that represent a threat to native taxa.
+In many waterfowl species, sexual dichromatism is observed in adults, while immatures of both sexes display a cover-up plumage similar to that of adult females.
+Using the example of non-native populations of Ruddy duck in Western Europe, a species that exhibits this pattern, we present a non-invasive method based on count data to estimate the respective contribution of recruitment and adult survival to the population growth rate.
+Using counts, we illustrate how changes in the apparent sex ratio (due to female-like young males gradually acquiring the plumage of adult male) can be exploited to partition population growth rate into adult survival and recruitment, and how this contributes significantly to understanding the demographic impact of two different eradication strategies.
 
-  **Keywords**: reproductive success - fecundity - productivity - age ratio - vital rate - duck - Oxyura jamaicensis
+  **Keywords**: population dynamics - survival - reproductive success - fecundity - vital rate - age ratio - intrinsic growth rate - invasive species - Oxyura jamaicensis
 output: 
-    bookdown::word_document2:
-      toc: false
     bookdown::pdf_document2:
@@
 It is also crucial for measuring the efficacy of management actions potentially undertaken to help population recovery, control, or eradication [@Shea1998].
-Among the approaches available to managers for reaching these goals, those relying on the monitoring of the number of individuals are by far the most popular [@Rintala2022].
-In many instances, these methods even enable environmental factors underlying changes in population size to be investigated, and hence some mitigating actions to be implemented [@Faillettaz2019].
+Among the approaches available to managers for reaching these goals, those relying on the monitoring of the number of individuals, i.e.
+counts, are by far the most popular [@Rintala2022].
+In many cases, these methods even allow the investigation of environmental factors underlying changes in population size and thus help in the implementation of mitigation measures [@Faillettaz2019].
 Unfortunately, relying on the monitoring of numbers of individuals alone generally hinders basic understanding of the demographic mechanisms underlying changes in population growth rates.
@@
 invasive species or pests.
-Furthermore, capture and handling may also be so stressful to individuals that their behaviour and/or the very parameters under study may be altered.
-Indeed, although highly efficient for assessing demographic parameters, capture-mark-recapture methods have potential drawbacks such as being invasive and hardly affordable when time and money are limiting.
-Genetic monitoring is a non-invasive alternative to capture-mark-recapture approaches based on physical captures, but it suffers from being costly and time-consuming, and requires quite rigorous sampling schemes in the field [@Caniglia2011].
+Furthermore, although highly efficient for assessing demographic parameters, CMR methods suffer from potential drawbacks such as being invasive (and may affect behaviour and hence survival or breeding success) and hardly affordable when time and money are limiting.
+Genetic monitoring is a non-invasive alternative to capture-mark-recapture approaches based on physical captures, but it suffers from being costly, and requires large samples [@Caniglia2011].
 
-Most of the time therefore, managers do the best of a bad job using counts as the only viable option for tracking population growth rates and assessing the efficiency of management actions.
-Counting the number of immatures produced in addition to live adults may enable adult survival and recruitment rate to be estimated, and hence provides knowledge on which of these components has the most influence on population growth rate.
-The recruitment rate, defined as the average number of immatures per breeder reaching the sexual maturity, is the product of two main sub-parameters: the fecundity followed by the juvenile survival [@Arnold2018; @Etterson2011].
-The recruitment rate is sometimes also called annual fertility [e.g.
-@Koons2014], reproductive rate [e.g.
-@Cooch2014], reproductive success [e.g.
-@Etterson2011], or productivity [e.g.
-@Hagen2008; @Johnson1987].
-Unfortunately, assessing the respective contribution of the adult survival and the recruitment rate is only possible for a limited number of species in which broodless adults and those with young display the same detectability.
-Alternatively, assessing the sex and age structure in hunting bags has been used to infer the role of decreasing reproductive success in population declines in a number of game species including ducks [e.g.
-@Fox2018].
-However, this approach is not suited for protected/endangered species with no harvest.
-Here we exemplify how, in dimorphic species, delayed maturity of males can be exploited for estimating adult survival and recruitment rate by differentiating male-like to female-like individuals in repeated winter counts.
-We used two non-native Ruddy duck, *Oxyura jamaicensis*, populations introduced in Europe as a study model.
-As numerous duck species, the Ruddy duck is dimorphic, with newly born males looking like females until the pre-nuptial moult, which occurs in late winter.
-This means that the apparent proportion of males increases over the course of the wintering season.
-These changes in apparent proportions of males during this period are therefore directly related to the proportion of immatures into the populations and hence to reproductive success and recruitment rate of the previous spring/summer.
-This study details the method to infer adult survival and recruitment rate from count data, and an application is developed to assess the impact of two different eradication strategies deployed in Great Britain and in France, respectively.
+Most of the time therefore, managers do the best of a bad job using counts as a viable option for tracking population growth rates and assessing the relevance of management actions.
+Counting the number of immatures in addition to live adults may enable the estimation of adult survival and recruitment rates, and thus allow the assessment of their relative influence on population growth rate.
+The adult survival rate is the proportion of adults staying alive over one year.
+The recruitment rate, also called productivity [e.g.
+@Hagen2008; @Johnson1987], is the average number of immatures per breeder that reach the sexual maturity, and is thus the product of two main sub-parameters: the fecundity followed by the juvenile survival [@Arnold2018; @Etterson2011].
+Alternatively, assessing the sex and age structure in hunting bags has been used to infer the role of decreasing reproductive success in population declines in a number of games species including ducks and geese but suffers from intractable biases [@Fox2018].
+Of course, this approach is not suited for protected / endangered species.
+
+Here we exemplify how, in dimorphic species, delayed maturity of males can be used for estimating adult survival and recruitment rate by differentiating male-like from female-like individuals in repeated winter counts.
+We used two non-native Ruddy duck (*Oxyura jamaicensis*) populations introduced in Europe as a study model.
+As numerous duck species, Ruddy duck is dimorphic, with newly born males looking like females until the pre-nuptial moult, which occurs in late winter.
+As a result of this late prenuptial moult of immatures, the apparent proportion of males increases over the course of the wintering season.
+These changes in apparent proportions of males during this period are therefore directly related to the proportion of immatures into the populations and thus to the reproductive success of the previous breeding season, and the recruitment rate of the next one.
+We estimated adult survival and recruitment rates and assessed the impact of two different eradication strategies deployed in Great Britain and in France, respectively.
 
@@
 The Ruddy duck is a stiff-tailed duck native from the Americas.
-From seven individuals initially acclimated in the Slimbridge Wetland Centre in Great Britain in the 40’s [@GutierrezExposito2020], a feral population began to establish with the first observed reproductive attempts in the wild in the 60s (Figure \@ref(fig:count)).
+From seven individuals initially acclimated in the Slimbridge Wetland Centre in the 1940’s [@GutierrezExposito2020], a feral population began to establish in Great Britain with the first observed reproductive attempts in the wild in the 60s (Figure \@ref(fig:count)).
 This feral population rapidly grew and spread to the entire country to reach more than 5,000 individuals in the early 2000s.
-By the end of the 80s, a feral population began to establish also on the continent in France, supposedly owing to the arrival of individuals born in Great Britain.
-Contrary to what was observed in Great Britain, the French Ruddy duck population did not spread much, with the vast majority of observations and breeding attempts clustering in Western France.
-In France, during winter, almost no Ruddy duck is observed outside the lake of Grand Lieu (47.09°N, 1.67°W), which is thus the exclusive wintering site and greatly facilitates the monitoring of this population.
-The feral European populations of Ruddy duck are considered as a major threat to the endangered native White-headed duck, *Oxyura leucocephala*, population of the South-Western Mediterranean, because of hybridization risks and thus an elevated potential for genetic pollution and eventually genetic assimilation of the latter by the former [@MunozFuentes2007].
-In order to mitigate the risks of genetic pollution of the White-headed duck by Ruddy ducks, eradication measures were taken both in Great Britain and in France from the late 90s [@GutierrezExposito2020], (Figure \@ref(fig:count)), and a European Ruddy duck eradication plan has been initiated by the European Commission in 1999 [@Hughes1999].
+By the end of the 80s, a feral population began to establish also on the continent, especially in France, supposedly owing to the arrival of individuals born in Great Britain.
+But contrary to what was observed in Great Britain, the Ruddy ducks did not spread much in France, with the vast majority of observations and breeding attempts clustering in the North-West of the territory.
+In France, during winter, almost no Ruddy duck is observed outside the lake of Grand Lieu (47.09°N, 1.67°W), which greatly facilitates the monitoring of this population.
+Ruddy ducks are considered as a major threat to the endangered native White-headed duck (*Oxyura leucocephala*) of the South-Western Mediterranean, because of hybridization and thus an elevated risk for genetic pollution and genetic assimilation of the latter by the former [@MunozFuentes2007].
+In order to mitigate the risks of genetic pollution of the White-headed duck by Ruddy ducks, eradication measures were implemented in the late 1990's both in Great Britain and in France [@GutierrezExposito2020], (Figure \@ref(fig:count)), and were followed by a European Ruddy duck eradication plan in 1999 [@Hughes1999].
+
+The trends and status of populations of both White-headed duck and Ruddy duck are being assessed through population censuses only.
+It is thus impossible to assess the factors that impair recovery of the former and the efficacy of the eradication program for the latter.
+In particular, based on censuses alone, one cannot assess of the relative effects of changes in adult survival and breeding success or recruitment rates on population growth rates which are prerequisites to identifying limiting factors.
+
 The release of non-native species in the wild, including Ruddy ducks, is forbidden in countries of the European Union, and the status of White-headed duck populations is highly unfavourable.
@@
 
-Like many other ducks, Ruddy ducks and White-headed ducks both display delayed dichromatism.
-Thereby, immature males are confounded to immature and adult females until a first moult occurring in mid-winter for the early ones [@Baldassarre2014], but most likely latter for most of them.
-Immature males acquire then the typical colourful breeding plumage of their kind in spring [@Baldassarre2014].
-Delayed dichromatism explains why the apparent proportion of females in winter counts is always greater than the true proportion of females that can be assessed in counts of late winter or in removals (Figures \@ref(fig:flock) & \@ref(fig:pfem)).
+Like many other ducks, Ruddy ducks and White-headed ducks both display delayed dichromatism, thereby, young males acquire the typical colourful plumage of their kind (and thus can be distinguished from females) over the course of the interbreeding season, typically by mid-winter for the earliest individuals [@Baldassarre2014].
+Delayed dichromatism generally explains the discrepancies in the estimates of the apparent proportions of females in winter vs spring, or between winters counts and hunting bags counts (Figures \@ref(fig:flock) & \@ref(fig:pfem)).
+
+Interestingly, delayed dichromatism allow the estimation of adult survival and recruitment rates by monitoring the seasonal evolution of apparent sex ratio and by making few and reasonable assumptions (see below).
+
 Thus, it is possible to assess the proportion of immatures within a population by monitoring the seasonal evolution of apparent sex ratio and assuming an even sex ratio at birth [@Bellrose1961; @Blums1996], comparable mortality rates between females and males among juveniles [@Ramula2018; @Wood2021], and unchanging adult sex ratio over the monitoring period [@Devineau2010].
@@
 This work was partly funded by the LIFE Oxyura project (LIFE17 NAT/FR/000942) through the LIFE program.
+We warmly thank Jay McGowan for allowing us to publish his Ruddy duck photograph.
 This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the *Office Français de la Biodiversité*, and Jean-Marc Gillier from the *Société Nationale pour la Protection de la Nature*.
@@
 
-(ref:count) Changes in the size of the two Ruddy duck populations; y-axis is on a log-scale; sex ratio refers to censuses that distinguished female-like from male-like individuals
+(ref:count) Temporal evolution of the number of Ruddy ducks (log scale) in Great Britain (GB) and France (FR) from 1960 to 2019, with indications of periods when apparent sex ratios were also recorded.
 
-(ref:flock) Typical observation of a Ruddy duck flock during winter: ten female-like individuals with a whitish striped cheek, four male-like individuals with a white cheek and a black cap, three unidentified individuals (1^st^, 4^th^, and 10^th^ from left) © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
+(ref:flock) Typical observation of a Ruddy duck flock during winter including ten female-like individuals with a whitish striped cheek (some of which are immature males), four male-like individuals with a white cheek and a black cap, three probable males individuals (1^st^, 4^th^, and 10^th^ from left) © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
 

### 2023-11-30 Code/Model_vital_rates_longer_ts.R

@@ -103,11 +103,11 @@ cor_plot <- function(col, title = NULL, subtitle = NULL, limits = NA, breaks = N
 
 # Data import -----------------------------------------------------------------------
 
-read_rds("./Data/Ruddy_duck_data.rds") %>% pluck(1) %>% 
+read_rds("./Data/Ruddy_duck_data_longer_ts.rds") %>% pluck(1) %>% 
   mutate(pop = if_else(pop == "UK", "GB", pop),
          across(pop, as_factor)) %>% arrange(pop) -> frag
 
-read_rds("./Data/Ruddy_duck_data.rds") %>% pluck(2) %>% 
+read_rds("./Data/Ruddy_duck_data_longer_ts.rds") %>% pluck(2) %>% 
   mutate(pop = if_else(pop == "UK", "GB", pop),
          across(pop, as_factor)) %>% arrange(pop) -> counts
 
@@ -988,7 +988,7 @@ ds %>%
   mutate(across(sq, ~ replace_na(.x, 0)),
          label = if_else(sq != 0, "available\ndichromatic\ndata", "") %>% as_factor() %>% 
            fct_relevel("", after = Inf)) %>% 
-  filter(year(year) <= 2019) -> ts_seq2
+  filter(year(year) <= 2023) -> ts_seq2
 
 ggplot() +
   scale_y_log10(minor_breaks = NULL) +
@@ -997,15 +997,15 @@ ggplot() +
                      labels = c("GB population", "FR population")) +
   geom_line(data = ts_seq2 %>% arrange(pop, year), 
             aes(x = year, y = n_win, color = pop), linewidth = 1, alpha = .5, linetype = "dotted") +
-  geom_line(data = ts_seq2 %>% filter(sq != 0), 
-            aes(x = year, y = n_win, color = pop, group = sq), linewidth = 2, alpha = .5) + 
+  # geom_line(data = ts_seq2 %>% filter(sq != 0), 
+  #           aes(x = year, y = n_win, color = pop, group = sq), linewidth = 2, alpha = .5) + 
   geom_point(data = ts_seq2, 
              aes(x = year, y = n_win), size = 3, color = "white") +
   geom_point(data = ts_seq2, 
              aes(x = year, y = n_win, color = pop), 
              size = 2, alpha = .5) + 
-  geom_point(data = ts_seq2 %>% filter(sq != 0), 
-             aes(x = year, y = n_win, color = pop), alpha = 1) +
+  # geom_point(data = ts_seq2 %>% filter(sq != 0), 
[…]

### 2024-05-27 Code/Draft_ruddy_duck_vital_rates.Rmd
Before instatext

@@
 link-citations: true
-title: "Delayed dichromatism as a convenient tool to disentangle the effects of survival and productivity on population growth in waterfowl"
+title: "Delayed dichromatism as a convenient tool to disentangle the effects of survival and productivity on the population dynamics in waterfowl"
 author: 
@@
 abstract: |
-  Monitoring the number of individuals is by far the most popular strategy for studying the environmental factors that rule population dynamics and for measuring the efficacy of management actions aimed at population recovery, control or eradication.
-Monitoring the number of individuals is unfortunately inefficient in assessing the mechanisms underlying demographic processes and, more specifically, the extent to which population growth rate is affected by changes in adult survival rather than variations in reproductive parameters.
-Common methods to inform survival and recruitment, such as capture-mark-recapture and game-hunting bag, rely on catching animals, which has three main disadvantages.
-First, the selectivity of capture methods can lead to a biased representation of the underlying population structure.
-Second, catching and releasing animals is quite invasive, and causes stress, thus posing an additional threat to often endangered species.
-Third, release may simply be impossible, especially for invasive alien species that represent a threat to native taxa.
-In many waterfowl species, sexual dichromatism is observed in adults, while immatures of both sexes display a cover-up plumage similar to that of adult females.
-Using the example of non-native populations of Ruddy duck in Western Europe, a species that exhibits this pattern, we present a non-invasive method based on count data to estimate the respective contribution of recruitment and adult survival to the population growth rate.
-Using counts, we illustrate how changes in the apparent sex ratio (due to female-like young males gradually acquiring the plumage of adult male) can be exploited to partition population growth rate into adult survival and recruitment, and how this contributes significantly to understanding the demographic impact of two different eradication strategies.
+  Monitoring the number of individuals is by far the most popular strategy for investigating the environmental factors ruling population dynamics and for measuring the efficacy of management actions targeting population recovery, control or eradication.
+Monitoring numbers is unfortunately inefficient for assessing the mechanisms underlying demographic processes and more specifically to assess the extent to which population growth rate is affected by changes in adult survival rather than fluctuations in reproductive parameters.
+Usual methods to inform adult survival and productivity, such as capture-mark-recapture and analysis of game-hunting bag, rely on catching animals which suffers from three main drawbacks.
+Firstly, the selectivity of the catching methods can lead to a biased representation of the underlying population structure.
+Secondly, catching and releasing animals is quite invasive, which generates stress, and is thus an additional threat to often already endangered species.
+Thirdly, releasing may simply be impossible especially for invasive alien species that represent a threat to native taxa.
+In many waterfowl species, a sexual dichromatism is observed in adults whereas immatures of both sexes display a cover-up plumage similar to adult females.
+Using the case of alien Ruddy duck populations of Western Europe (a species displaying this pattern), we introduce a non-invasive method based on count data distinguishing plumage types that allows estimating the respective contribution of adult survival and productivity (also called recruitment rate) to population growth rate.
+We exemplify how, based on count surveys, changes in apparent male proportion (due to female-like immature males progressively acquiring adult male plumage) can be exploited to decompose population growth rate into adult survival and recruitment rates, greatly helping to understand the demographic impact of two different eradication strategies.
 
-  **Keywords**: population dynamics - survival - reproductive success - fecundity - vital rate - age ratio - intrinsic growth rate - invasive species - Oxyura jamaicensis
+  **Keywords**: intrinsic growth rate - vital rate - reproductive success - fecundity - sex ratio - age ratio - Oxyura jamaicensis
 output: 
@@
 
-Assessing population growth rate is a key step towards a better understanding of factors underlying the dynamics of natural populations [e.g.
+Assessing population growth rate is a first step towards a better understanding of factors underlying the dynamics of natural populations [e.g.
 @Niel2005].
 It is also crucial for measuring the efficacy of management actions potentially undertaken to help population recovery, control, or eradication [@Shea1998].
-Among the approaches available to managers for reaching these goals, those relying on the monitoring of the number of individuals, i.e.
-counts, are by far the most popular [@Rintala2022].
-In many cases, these methods even allow the investigation of environmental factors underlying changes in population size and thus help in the implementation of mitigation measures [@Faillettaz2019].
-Unfortunately, relying on the monitoring of numbers of individuals alone generally hinders basic understanding of the demographic mechanisms underlying changes in population growth rates.
-More specifically, based on counts only, it is nearly impossible to assess the extent to which population growth rate is affected by change in survival rather than to variations in reproductive parameters [@Austin2000].
-To assess the relative sensitivity of population growth rate to factors affecting survival or reproductive parameters, demographers most often rely on the monitoring of individuals by capture-mark-recapture [@Lebreton1992].
-In practice, however, one cannot always rely on the capture and release of individuals owing to the legal status of the species that precludes any additional disturbance, e.g.
-critically endangered species, or forbid the release of alive individuals, e.g.
-invasive species or pests.
-Furthermore, although highly efficient for assessing demographic parameters, CMR methods suffer from potential drawbacks such as being invasive (and may affect behaviour and hence survival or breeding success) and hardly affordable when time and money are limiting.
-Genetic monitoring is a non-invasive alternative to capture-mark-recapture approaches based on physical captures, but it suffers from being costly, and requires large samples [@Caniglia2011].
+Among the approaches available to managers for reaching these goals, those relying on the monitoring of the number of individuals (counts) are by far the most popular [@Rintala2022].
+In many instances, these methods enable investigating environmental factors underlying changes in population size, and hence help implementing mitigating actions [@Faillettaz2019].
+Unfortunately, relying on the monitoring of population abundance alone generally hinders basic understanding of the demographic mechanisms underlying changes in population growth rates.
+More specifically, based on counts only, it is nearly impossible to assess the extent to which population growth rate is affected by changes in adult survival rather than to variations in productivity [@Austin2000], which is defined as the product of all reproductive parameters.
+To assess the relative sensitivity of population growth rate to factors affecting adult survival or reproductive parameters, demographers most often rely on the monitoring of individuals by capture-mark-recapture [CMR, @Lebreton1992].
+In practice, however, one cannot always rely on the capture and release of individuals owing to the legal status of the species that precludes any additional disturbance (e.g.
+critically endangered species) or forbid the release of alive individuals (e.g.
+invasive species or pests).
+Furthermore, although highly efficient for assessing demographic parameters, capture-mark-recapture methods suffer from potential drawbacks such as being invasive (and may affect behaviour and hence survival or breeding success) and hardly affordable when time and money are limiting.
+Genetic monitoring is a non-invasive alternative to capture-mark-recapture, but it suffers from being costly and requires large samples [@Caniglia2011].
 
-Most of the time therefore, managers do the best of a bad job using counts as a viable option for tracking population growth rates and assessing the relevance of management actions.
-Counting the number of immatures in addition to live adults may enable the estimation of adult survival and recruitment rates, and thus allow the assessment of their relative influence on population growth rate.
-The adult survival rate is the proportion of adults staying alive over one year.
-The recruitment rate, also called productivity [e.g.
-@Hagen2008; @Johnson1987], is the average number of immatures per breeder that reach the sexual maturity, and is thus the product of two main sub-parameters: the fecundity followed by the juvenile survival [@Arnold2018; @Etterson2011].
-Alternatively, assessing the sex and age structure in hunting bags has been used to infer the role of decreasing reproductive success in population declines in a number of games species including ducks and geese but suffers from intractable biases [@Fox2018].
-Of course, this approach is not suited for protected / endangered species.
+Most of the time therefore, managers do the best of a bad job using counts as a viable option for tracking population size changes and then population growth rate, in order to assess the relevance of management actions.
+Distinguishing immatures from adults in counts enables estimating adult survival rate, i.e.
+the proportion of breeders that survived over one year, and productivity/recruitment rate, i.e.
+the number of immatures produced per breeder that reach sexual maturity, which are, by definition, recruited in the breeding population.
+This additional effort makes it possible to assess the relative influence of each of these components on population growth rate.
+Alternatively, assessing the age structure in hunting bags has been used to infer the role of decreasing reproductive success in population declines in a number of games species including ducks and geese but suffers from intractable biases [@Fox2018].
+Of course, the latter approach is not suited for protected/endangered species.
 
@@
 We used two non-native Ruddy duck (*Oxyura jamaicensis*) populations introduced in Europe as a study model.
-As numerous duck species, Ruddy duck is dimorphic, with newly born males looking like females until the pre-nuptial moult, which occurs in late winter.
+As numerous duck species, Ruddy duck is dimorphic, with newly born males looking like females until the prenuptial moult, which occurs in late winter.
 As a result of this late prenuptial moult of immatures, the apparent proportion of males increases over the course of the wintering season.
-These changes in apparent proportions of males during this period are therefore directly related to the proportion of immatures into the populations and thus to the reproductive success of the previous breeding season, and the recruitment rate of the next one.
+These changes in apparent proportions of males during this period are therefore directly related to the proportion of immatures into the populations and thus to the reproductive success of the previous breeding season.
 We estimated adult survival and recruitment rates and assessed the impact of two different eradication strategies deployed in Great Britain and in France, respectively.
@@
 The Ruddy duck is a stiff-tailed duck native from the Americas.
-From seven individuals initially acclimated in the Slimbridge Wetland Centre in the 1940’s [@GutierrezExposito2020], a feral population began to establish in Great Britain with the first observed reproductive attempts in the wild in the 60s (Figure \@ref(fig:count)).
+From seven individuals initially acclimated in the Slimbridge Wetland Centre in the 1940's [@GutierrezExposito2020], a feral population began to establish in Great Britain with the first observed reproductive attempts in the wild in the 1960's (Figure \@ref(fig:count)).
 This feral population rapidly grew and spread to the entire country to reach more than 5,000 individuals in the early 2000s.
-By the end of the 80s, a feral population began to establish also on the continent, especially in France, supposedly owing to the arrival of individuals born in Great Britain.
+By the end of the 1980's, a feral population began to establish also on the continent, especially in France, supposedly owing to the arrival of individuals born in Great Britain.
 But contrary to what was observed in Great Britain, the Ruddy ducks did not spread much in France, with the vast majority of observations and breeding attempts clustering in the North-West of the territory.
@@
 Ruddy ducks are considered as a major threat to the endangered native White-headed duck (*Oxyura leucocephala*) of the South-Western Mediterranean, because of hybridization and thus an elevated risk for genetic pollution and genetic assimilation of the latter by the former [@MunozFuentes2007].
-In order to mitigate the risks of genetic pollution of the White-headed duck by Ruddy ducks, eradication measures were implemented in the late 1990's both in Great Britain and in France [@GutierrezExposito2020], (Figure \@ref(fig:count)), and were followed by a European Ruddy duck eradication plan in 1999 [@Hughes1999].
+In order to mitigate the risks of genetic pollution of the White-headed duck by Ruddy ducks, eradication measures were implemented in the late 1990's both in Great Britain and in France [@GutierrezExposito2020], and were followed by a European Ruddy duck eradication plan in 1999 [@Hughes1999] (Figure \@ref(fig:count)).
 
@@
 It is thus impossible to assess the factors that impair recovery of the former and the efficacy of the eradication program for the latter.
-In particular, based on censuses alone, one cannot assess of the relative effects of changes in adult survival and breeding success or recruitment rates on population growth rates which are prerequisites to identifying limiting factors.
-
-The release of non-native species in the wild, including Ruddy ducks, is forbidden in countries of the European Union, and the status of White-headed duck populations is highly unfavourable.
-The use of capture-mark-recapture to monitor populations of these species is thus impossible.
-As a result, the effectiveness of the eradication of Ruddy ducks and management actions intended to favour the recovery of White-headed ducks were essentially assessed through censuses.
-Censuses, however, prevented assessment of the relative effects or changes in survival and recruitment rates on population growth rate, which is a key towards a proper understanding of the efficiency of management actions.
+In particular, based on censuses alone, one cannot assess the relative effects of changes in adult survival and recruitment rates on population growth rates which are prerequisites to identifying limiting factors.
 
 ```{r count, fig.cap = "(ref:count)"}
-knitr::include_graphics("../Output/plot_paper_2_mod.png")
+knitr::include_graphics("../Output/plot_paper_2_saved.png")
 ```
 
-Like many other ducks, Ruddy ducks and White-headed ducks both display delayed dichromatism, thereby, young males acquire the typical colourful plumage of their kind (and thus can be distinguished from females) over the course of the interbreeding season, typically by mid-winter for the earliest individuals [@Baldassarre2014].
-Delayed dichromatism generally explains the discrepancies in the estimates of the apparent proportions of females in winter vs spring, or between winters counts and hunting bags counts (Figures \@ref(fig:flock) & \@ref(fig:pfem)).
+Like many other ducks, Ruddy ducks and White-headed ducks both display delayed dichromatism thereby young males acquire the typical colourful plumage of their kind, and hence can be distinguished from females, over the course of the interbreeding season, typically by mid-winter for the earliest individuals [@Baldassarre2014].
+Delayed dichromatism generally explains the discrepancies in the estimates of apparent proportions of males counted in winter vs counted in spring, or vs males determined from individuals culled in winter (Figures \@ref(fig:flock) & \@ref(fig:pmal)).
 
-Interestingly, delayed dichromatism allow the estimation of adult survival and recruitment rates by monitoring the seasonal evolution of apparent sex ratio and by making few and reasonable assumptions (see below).
+Interestingly, delayed dichromatism enables estimating adult survival and recruitment/productivity by monitoring the seasonal evolution of apparent sex ratio and by making few and reasonable assumptions (see next section).
 
-Thus, it is possible to assess the proportion of immatures within a population by monitoring the seasonal evolution of apparent sex ratio and assuming an even sex ratio at birth [@Bellrose1961; @Blums1996], comparable mortality rates between females and males among juveniles [@Ramula2018; @Wood2021], and unchanging adult sex ratio over the monitoring period [@Devineau2010].
-Then, both adult survival rate and recruitment rate can be estimated by combining the age ratio thus obtained with reliable counts over two successive winters.
+<!-- Interestingly, delayed dichromatism allows estimating the proportion of immature males, which is the difference between the proportion of female-like individuals between winter and spring.
+The proportion of all immature individuals is twice this value if we assume an even sex ratio at birth [@Bellrose1961; @Blums1996] and comparable mortality rates between immature females and immature males [@Ramula2018; @Wood2021].
+The proportion of immatures can even be estimated only from winter counts by making the assumption of a known and unchanging adult sex ratio over the monitoring period [@Devineau2010], as developed inhere.
+By combining this proportion of immatures to reliable counts over two successive winters, both adult survival rate and recruitment rate can be estimated.
+-->
 
@@
 
-```{r pfem, fig.cap = "(ref:pfem)"}
+```{r pmal, fig.cap = "(ref:pmal)"}
 knitr::include_graphics("../Output/plot_paper_1_saved.png")
@@
 
-We monitored the demography of Ruddy duck populations in both Great Britain and France using exhaustive counts on the wintering grounds during the period between December 1 and January 31.
-In addition, counts that distinguished female-like from male-like individuals were performed from 2006 to 2012 in Great Britain, which corresponded to a period of strong decline of the population (Figure \@ref(fig:count)).
-In France, the monitoring spread over 1999, 2001-2009, and 2012-2019.
-This population grew rapidly during the first years of the monitoring and then stabilized from 2006 onwards, which corresponds to a period with a high culling effort, which was performed under control of the authorities in both countries.
-Even if the age ratio in the removals was uncertain before 2009 in France, it is possible to grossly qualify the culling pressure in three categories for both countries: "no culling" before 1999, "low culling" between 1999 and 2005, and "high culling" from 2006 (see Figure \@ref(fig:sup) in Supplementary Materials).
+We monitored the demography of Ruddy duck populations in both Great Britain (GB) and France (FR) using exhaustive counts on the wintering grounds during the period between December 1 and January 31.
+In addition, apparent sex ratios in winter could be accessible from counts that distinguished female-like from male-like individuals in winter.
+Such counts were performed from 2006 to 2012 in Great Britain, which corresponded to a period of strong decline of the population (Figure \@ref(fig:count)).
+In France, a monitoring accounting for apparent sex was performed in 1999, 2001-2009, and 2012-2019.
+The French population observed a rapid growth during the first years of the monitoring and then stabilized from 2006 onwards as a result of a high culling pressure.
+In both countries, shooting was performed under the strict control of the authorities, meaning that the exact number of individuals culled was known.
+Therefore, despite the age ratio in birds shot was not ascertained before 2009 in France, it is possible to grossly qualify the culling pressure in three categories for both countries: "no culling" before 1999, "low culling" between 1999 and 2005, and "high culling" from 2006 (see Figure \@ref(fig:sup) in Supplementary Material).
 
-## Vital rates inference from count data
+## Inferring adult survival and recruitment rates from changes in population size and proportion of immatures
 
-In waterfowl, survival rate is commonly higher in adult males than in adult females [@Baldassarre2014].
-Adult sex ratio is then generally biased towards males (Figure \@ref(fig:pfem), @Szekely2014).
-As a consequence, deducing the proportion of immatures within a population from the observed proportion of adult males is not straightforward because the proportion of adults is not just twice as much as the observed proportion of adult males.
-There were no counts differentiating male-like from female-like individuals during the breeding period to properly estimate adult sex ratio.
-But sex identification in adults from removals was available for both Ruddy duck populations.
-However, the small population size in France prevented from getting precise adult sex ratios.
-A comparison over months in Great Britain showed no difference in proportion of males in adults, so data from adults collected over the whole year were used to estimate the proportion of males in adults.
-A comparison of proportion of males in adults among years with more than 500 samples did not exhibit significant interannual differences.
-Even if the adult sex ratio may vary over a long time range, it is relatively stable over a few years [@Wood2021].
-As a consequence, removal data from all adults were pooled to estimate the proportion of males in adults, see Equation \@ref(eq:pm).
+The proportion of immature males is the difference between the proportion of female-like individuals between winter and spring.
+The proportion of all immature individuals is twice this value if we assume an even sex ratio at birth [@Bellrose1961; @Blums1996] and comparable mortality rates between immature females and immature males [@Ramula2018; @Wood2021].
+However, counts in spring suffer from inextricable issues, such as a differential detectability between females and males.
+Fortunately, the proportion of immatures can also be estimated only from winter counts by making the assumption of a known and unchanging adult sex ratio over the monitoring period [@Devineau2010].
 
-Assuming the additive property of the binomial distribution, the proportion of immatures can be deduced from the cumulated counts of male-like individuals in the wintering population, see Equation \@ref(eq:pic).
-From this proportion and the interannual variation of an abundance index of population size, adult survival and recruitment rate are straightforward, see Equation \@ref(eq:vr).
-If the absolute value of population size is accessible, numbers of adults and recruits can be estimated, see Equation \@ref(eq:VR).
+The adult sex ratio is a prerequiste to estimate the proportion of immatures.
+No counts differentiating males from females were available during the breeding period to estimate adult sex ratios, but we determined the sex of each adult that was shot as part of the eradication programme both in Great Britain and in France.
+Nevertheless, the small number of culled individuals in France prevented from getting robust adult sex ratios.
+In Great Britain, the adult sex ratio estimated from the analysis of birds shot did not differ among months, so data from adults collected over the whole biological cycle were pooled to estimate adult sex ratios.
+The proportion of males did not statistically differ between the different years with more than 500 adults shot.
+This is consistent with previous findings showing that adult sex ratios in ducks are usually highly stable over the short-term [@Wood2021] but may vary over the long-term.
+Therefore, we pooled data from all adults shot as part as the control programme to estimate the adult sex ratio, see Equation \@ref(eq:pm).
 
-|Name|Class|Description|
-|:---|:----|:----------------------------|
-|$AM$|Data|Total number of adult males in removals|
-|$AF$|Data|Total number of adult females in removals|
-|$p(m|a)$|Parameter|Proportion of males in adults (or probability of being a male knowing it is an adult)|
-|$CML_{i, t}$|Data|Cumulated number of male-like individuals counted in population $i$ in year $t$, which are assumed to all be adult males|
-|$C_{i, t}$|Data|Cumulated number of individuals counted in population $i$ in year $t$|
-|$p(a \cap m)_{i, t}$|Parameter|Proportion of adult males in population $i$ in year $t$|
-|$p(a)_{i, t}$|Parameter|Proportion of adults in population $i$ in year $t$|
-|$p(i)_{i, t}$|Parameter|Proportion of immatures in population $i$ in year $t$|
-|$N_{i, t}$|Data|Size of the population $i$ in year $t$ (maximum number of individuals counted in the wintering population)|
-|$s_{i, t}$|Parameter|Adult survival rate, i.e. proportion of individuals in year $t-1$ still alive in year $t$|
-|$r_{i, t}$|Parameter|Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per individual in year $t - 1$|
-|$S_{i, t}$|Parameter|Number of adults in population $i$ in year $t$|
-|$R_{i, t}$|Parameter|Number of recruits in population $i$ in year $t$|
+Assuming the additive property of the binomial distribution, the proportion of immatures in a population can be deduced from the proportion of males in adults (alternatively the adult sex ratio) and the proportion of male-like individuals in the different count surveys in winter (assuming all these individuals are adult males), see Equation \@ref(eq:pic).
+By combining the population growth rate, which is derived from the year-to-year variation of an index of abundance of population size, to the proportion of immatures, one can derive adult survival and recruitment rates, see Equation \@ref(eq:vr) and Figure \@ref(fig:outline).
+A key concept is that if the absolute value of population size is known, numbers of adults and recruits can be estimated, see Equation \@ref(eq:VR).
 
+| Variable name         | Description                                                                                                |
+|:-|:-------|
+| *Observation (input)* |                                                                                                            |
+| $AM$                  | Total number of adult males shot as part of the eradication programme                                      |
+| $AF$                  | Total number of adult females shot as part of the eradication programme                                    |
+| $CML_{i, t}$          | Number of male-like individuals counted in population $i$ in year $t$                                      |
+| $C_{i, t}$            | Number of individuals counted in population $i$ in year $t$                                                |
+| $N_{i, t}$            | Size of the population $i$ in year $t$ (maximum number of individuals counted in the wintering population) |
+| *Parameter (output)*  |                                                                                                            |
+| $p(m|a)$              | Proportion of males in adults (or probability of being a male knowing it is an adult)                      |
+| $p(a \cap m)_{i, t}$  | Proportion of adult males in population $i$ in year $t$                                                    |
+| $p(a)_{i, t}$         | Proportion of adults in population $i$ in year $t$                                                         |
+| $p(re)_{i, t}$        | Proportion of immatures/recruits in population $i$ in year $t$                                             |
+| $s_{i, t}$            | Adult survival rate, i.e. proportion of individuals in year $t-1$ still alive in year $t$                  |
+| $r_{i, t}$            | Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per individual in year $t - 1$     |
+| $S_{i, t}$            | Number of adults in population $i$ in year $t$                                                             |
+| $R_{i, t}$            | Number of recruits in population $i$ in year $t$                                                           |
+
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
+```
+```{=tex}
 \begin{align}
@@
 & \sim {\sf Binom}(p(m|a).p(a)_{i, t}, C_{i, t}) \notag \\
-& \sim {\sf Binom}(p(m|a).(1 - p(i)_{i, t}), C_{i, t})
+& \sim {\sf Binom}(p(m|a).(1 - p(re)_{i, t}), C_{i, t})
 (\#eq:pic)
 \end{align}
-
+```
+```{=tex}
 \begin{equation}
 \begin{aligned}
-s_{i, t} & = \frac{N_{i, t}}{N_{i, t - 1}}.(1 - p(i)_{i, t}) \\
-r_{i, t} & = \frac{N_{i, t}}{N_{i, t - 1}}.p(i)_{i, t}
+s_{i, t} & = \frac{N_{i, t}}{N_{i, t - 1}}.(1 - p(re)_{i, t}) \\
+r_{i, t} & = \frac{N_{i, t}}{N_{i, t - 1}}.p(re)_{i, t}
 \end{aligned}
@@
 \end{equation}
-
+```
+```{=tex}
 \begin{equation}
 \begin{aligned}
-S_{i, t} & = N_{i, t}.(1 - p(i)_{i, t}) \\
-R_{i, t} & = N_{i, t}.p(i)_{i, t}
+S_{i, t} & = N_{i, t}.(1 - p(re)_{i, t}) \\
+R_{i, t} & = N_{i, t}.p(re)_{i, t}
 \end{aligned}
@@
 \end{equation}
+```
+```{r outline, fig.cap = "(ref:outline)"}
+knitr::include_graphics("../Output/outline.png")
+```
 
-## Validation of the count-based method {#VAL}
+## Validating the method {#VAL}
 
-The relevance of the method was assessed by testing the likelihood of the estimates of both vital rates.
-For adult survival rate, we checked if the values belonged to the expected interval $[0;1]$, and we compared the values to literature data on ducks that are accessible because adult survival is commonly assessed by capture-mark-recapture [@Lebreton2001].
-Maximum adult survival is defined as a uniform distribution on $[0.7;1]$.
-The lower limit corresponds to the upper range of survival rates in literature for waterfowl species of similar weight, the upper limit was set to 1 because survival rates of long-life waterfowl species are very high [@Nichols1997; @Krementz1997; @Buxton2004; @Nichols1992].
-Recruitment rate is more tricky to validate because it is not upper bounded as it is defined on $[0;\infty[$ and is neither explicitly informed in literature.
-Only some components of the recruitment rate are generally described, such as nesting rate, nesting success, clutch size, hatching success, early stage survival, but rarely survival between the juvenile stage and the first breeding season, which is necessary to inform recruitment rate [e.g.
-@Baldassarre2014].
-We then developed an approach to estimate the maximum expected recruitment rate without exploitation and without negative density-dependence processes, and we checked if all values were equal or below this maximum recruitment rate.
-If the recruitment rate outputs of the model are significantly higher than maximum recruitment rate, it means that the method overestimates such rate, which reflects that a part of adult males are not detected during count surveys.
-A null or negative recruitment rate would mean that immature males moulted before the survey or/and the assumption on the stability of the adult sex ratio is overrated.
+Nevertheless, we developed an indirect approach to estimate the maximum expected recruitment rate without exploitation and assuming that the observed density levels did not affect components of the recruitment process.
+Estimated values above that maximum recruitment rate would reflect failure in detection of adult males.
+On the other hand, null or negative recruitment rates would reflect either that immature males had acquired the male plumage before the survey, or violation of the assumption about the stability of the adult sex ratio.
 
-Practically, maximum recruitment rate is deduced by using its relationship to maximum growth rate and maximum adult survival rate, see Equation \@ref(eq:GR).
+Practically, maximum recruitment rate is estimated from the difference between maximum growth rate and maximum adult survival rate, see Equation \@ref(eq:GR).
 This relationship comes from a simple reasoning for a closed population: population size in year $t$ is equal to the number of adults that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
-the recruitment in year $t$.
-The growth rate of a population is thus the sum of adult survival rate plus recruitment rate [@Flint2015].
-For an open population, adult survival and recruitment rates are confounded to adult and recruit migrations, respectively.
+the recruitment in year $t$ [@Flint2015].
+For an open population, adult survival and recruitment rates are confounded to adult and recruit migrations, respectively, but this does not corrupt the equation.
 This relationship becomes more complex if a species with delayed maturity is considered, see @Robertson2008.
 
-|Name|Description|
-|:--|:----------------------------|
-|$N_{t}$|Number of adults in year $t$|
-|$D_{t}$|Number of adults dead during year $t$|
-|$R_{t}$|Number of recruits in year $t$|
-|$s_{t}$|Adult survival rate, i.e. proportion of individuals in year $t-1$ still alive in year $t$|
-|$r_{t}$|Recruitment rate, i.e. number of recruits in year $t$ produced per individual in year $t-1$|
-|$\lambda_{t}$|Growth rate of the population between year $t-1$ and year $t$|
+| Variable name | Description                                                                                 |
+|:-|:---------|
+| $N_{t}$       | Number of adults in year $t$                                                                |
+| $D_{t}$       | Number of adults dead during year $t$                                                       |
+| $R_{t}$       | Number of recruits in year $t$                                                              |
+| $s_{t}$       | Adult survival rate, i.e. proportion of individuals in year $t-1$ still alive in year $t$   |
+| $r_{t}$       | Recruitment rate, i.e. number of recruits in year $t$ produced per individual in year $t-1$ |
+| $\lambda_{t}$ | Growth rate of the population between year $t-1$ and year $t$                               |
 
+```{=tex}
 \begin{align} 
@@
 \end{align}
+```
+Maximum adult survival was defined as a uniform distribution on the $[0.7;1]$ interval.
+The lower limit corresponded to the upper range of survival rates observed in waterfowl species of similar weight, and the upper limit was set to 1 because survival rates of long-lived waterfowl species are very high [@Nichols1997; @Krementz1997; @Buxton2004; @Nichols1992; @Koons2014].
 
-The maximum growth rate occurred for both populations during their expansion phase, when change in population size was the steepest.
-To estimate robust maximum population growth rates for both populations, we smoothed annual population growth rate over a consistent time period by using a linear regression on the logarithm scale, see Equation \@ref(eq:L).
-To do so, we discarded Great Britain data before 1972 since the size estimate of this population was noisy below 50 individuals (Figure \@ref(fig:count)).
-After reaching the threshold of 1,000 individuals, the Great Britain (GB) population growth showed a strong inflexion whereas no culling pressure was applied.
+In both populations, the maximum growth rates occurred during their geographical expansion phase before shooting operations (Figure \@ref(fig:count)).
+To estimate robust maximum population growth rates for both populations, we smoothed annual growth rate values over a consistent time-period by using a linear regression on the logarithm scale, see Equation \@ref(eq:L).
+We discarded Great Britain data before 1972 as the estimates were probably noisy when the population was low (below 50 indiviudals) (Figure \@ref(fig:count)).
+After reaching the threshold of 1,000 individuals, the Great Britain (GB) population growth showed a strong inflexion despite culling had not begun (Figure \@ref(fig:count)).
 This observation suggests that a negative density-dependence process might occur beyond 1,000 individuals and led to consider only the first sequence of the time series to infer on the maximum growth rate in Great Britain, i.e.
 1972-1981.
-For the French (FR) population, the sequence without culling effort spread over the period 1994-1999.
+For the French (FR) population, the sequence without culling pressure spread over the period 1994-1999.
 
-|Name|Class|Description|
-|:--|:----|:----------------------------|
-|$N0_{i, J}$|Parameter|Intercept of the regression model|
-|$\lambda_{i, J}$|Parameter|Average population growth rate over a restricted time interval $J$ for a population $i$ (in $year^{-1}$)|
-|$t \in J$|Index|Year index within the time interval $J$| 
-|$\sigma_{i, J}$|Parameter|Standard deviation of the regression model|
+| Variable name    | Class     | Description                                                                                              |
+|:-|:-|:--------|
+| $N0_{i, J}$      | Parameter | Intercept of the regression model                                                                        |
+| $\lambda_{i, J}$ | Parameter | Average population growth rate over a restricted time interval $J$ for a population $i$ (in $year^{-1}$) |
+| $t \in J$        | Index     | Year index within the time interval $J$                                                                  |
+| $\sigma_{i, J}$  | Parameter | Standard deviation of the regression model                                                               |
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
+```
+Estimating vital rates within an acceptable range, as defined above, would not prove that out modelling approach correctly tracked their year-to-year variability.
+In order to ascertain this aspect, we check the correlation between the proportions of immatures and both survival and recruitment rates derived from counts, to estimates derived from cullings, i.e.
+from individuals shot as part of the eradication programme.
+As we restricted the dataset to years with more than 100 individuals shot in winter for sake of precision, the analysis covered only five years of the time series of the GB population.
+Proportions of immatures were estimated from culled individuals by checking the presence of the bursa of Fabricius, which is present in immature individuals only [@Hochbaum1942]; see Equation \@ref(eq:pis).
+Adult survival and recruitment rates were then estimated by combining the proportion of immatures and Equation \@ref(eq:vr).
 
-Even if both vital rates vary in a likely range following the validation methodology previously described, this does not prove that the interannual variability is properly tracked.
-In order to validate this aspect, we compared the outputs of the count-based method, i.e.
-the proportion of immatures and both vital rates, to outputs based on data from removals.
-Only data from years with more than 100 individuals culled in winter were selected.
-It covers 5 years of the count time series of the GB population.
-The presence of the bursa of Fabricius enables immature to be identified with certainty [@Hochbaum1942].
-The estimation of the proportion of immatures in removals is then straightforward, see Equation \@ref(eq:pis).
-Adult survival and recruitment rates were then estimated by combining this proportion and Equation \@ref(eq:vr).
-We discussed the ability of the count-based method to accurately track the interannual variability by comparing the outputs of the two methods.
+| Variable name | Class        | Description                                                              |
+|:-|:-|:--------|
+| $SI_{i, t}$   | Observation  | Number of immatures culled in the wintering population $i$ in year $t$   |
+| $S_{i, t}$    | Observation  | Number of individuals culled in the wintering population $i$ in year $t$ |
 
-|Name|Class|Description|
-|:---|:----|:----------------------------|
-|$SI_{i, t}$|Data|Number of immatures sampled in the wintering population $i$ in year $t$|
-|$S_{i, t}$|Data|Number of individuals sampled in the wintering population $i$ in year $t$|
-
+```{=tex}
 \begin{equation}
-SI_{i, t} \sim {\sf Binom}(p(i)_{i, t}, S_{i, t})
+SI_{i, t} \sim {\sf Binom}(p(re)_{i, t}, S_{i, t})
 (\#eq:pis)
 \end{equation}
+```
+## Assessing the impact of culling strategies {#HARV}
 
-## Impact of culling strategies on vital rates {#HARV}
+During period when culling pressure was high (i.e.
+from 2006 onwards), eradication strategies differed between Great Britain and France.
+In Great Britain, culling mostly occurred in winter (53.4% of adults were shot before the 30th May), whereas it was mostly done during the breeding period in France (81.6% of adults were shot after the 30th May).
+We assessed whether or not these strategies had a different impact on populations by comparing growth rates and the relative contributions of both recruitment and survival on them (by comparing the average values over the high culling pressure period to proxies of maximum vital rates estimated when both populations reached their maximum growth, see Section \@ref(VAL)).
+In France, a LIFE project was implemented by early 2019 to intensify the culling pressure, especially during winter and spring.
+We therefore excluded the FR time series from 2019 onward in order to provide a comparison of homogeneous culling strategies.
 
-Over the high culling period, i.e.
-from 2006 onwards, strategies differed between Great Britain to France.
-In Great Britain, culling mostly occurred in winter, i.e.
-before the breeding period (53.4% of adults were shot before the 30^th^ May), whereas it was mostly done in summer in France, i.e.
-during and after the breeding period (81.6% of adults were shot after the 30^th^ May).
-We compared the population growth of the two populations in light of the difference between the two strategies and we assessed the respective response of the vital rates by comparing the average values over the high culling period to proxies of maximum vital rates estimated when both populations reached their maximum growth (see Section \@ref(VAL)).
-In France, a LIFE project was implemented from early 2019 to intensify the culling pressure, especially during winter and spring.
-We therefore excluded the FR time series from 2019 in order to provide a comparison of homogeneous culling strategies.
+## Statistical framework
 
-For all the sub-models in Section \@ref(MM), the Bayesian framework was used for its efficiency and simplicity to propagate error through the parameters.
+We used the Bayesian framework to implement all sub-models presented in Section \@ref(MM).
+The Bayesian framework is both straightforward and efficient to propagate error through the parameters.
 We used uninformative priors on all parameters.
-As the maximum growth rate is a life history trait expected to be stable among populations of a given species, we used an uninformative hierarchical prior for this parameter.
-We generated three chains of length 500,000, with a thinning of 10 to avoid autocorrelation in the samples, and we discarded the first 2,000 samples as burn-in.
-Chain convergence was assessed using the Gelman and Rubin convergence diagnostic (R<1.1, @Gelman1992).
-We fit the models using NIMBLE [@Valpine2017] run from R [@RCT2022].
-Data and code are available here: https://github.com/adri-tab/Ruddy_duck_vital_rates.
+As the maximum growth rate is a life history trait expected to be stable among populations of a same species, we used an uninformative hierarchical prior for this parameter.
+We generated three chains of length 500,000, with a thinning of 10 to avoid autocorrelation in the samples, discarding the first 2,000 samples as burn-in.
+We assessed chain convergence using the Gelman and Rubin convergence diagnostic (R\<1.1, @Gelman1992).
+Models were fitted using NIMBLE [@Valpine2017] run from R [@RCT2022].
 The values **X[Y; Z]** reported in Section \@ref(RES) are the medians and the associated boundaries of the 95% confidence interval of posterior distributions.
-The median was preferred to the mean because of its robustness to skewed distribution.
+We preferred the median to the mean because of its robustness to skewed distribution.
 
@@
 
-## Estimates of vital rates
+## Estimating vital rates
 
-The method successfully provided estimates and associated uncertainty of immature proportion in both wintering populations.
-Adult survival rate and recruitment rate, i.e.
-the two components of the population growth rate, were also estimated by the method successfully (Figure \@ref(fig:vital)).
+The method successfully provided estimates and associated uncertainty for the proportion of immatures in both wintering populations.
+Similarly, adult survival and recruitment rates, i.e.
+the two components accounting for population growth rate, were also properly estimated (Figure \@ref(fig:vital)).
 
@@
 The proportion of males in adults, which is a prerequisite to infer the proportion of immatures, was estimated at 0.60 [0.59; 0.61].
-The immature proportion ranged from 0.16 [0.07; 0.24] to 0.54 [0.44; 0.62].
-The lower values were similar for the two populations, but the range of the proportions of immatures in the GB population was far more limited (upper value : 0.30 [0.14; 0.45]).
-As opposed to the FR population, there were no significant differences from a year to another because all estimates overlapped with each other.
-The proportion of immatures was relatively stable over time in the GB populations, while a slight decrease was observed for the FR population.
+The proportion of immatures ranged from 0.16 [0.07; 0.24] to 0.54 [0.44; 0.62], depending on population and year.
+The lowest values were similar for the two populations, but the range of the proportion of immatures in the GB population was far more limited (upper value GB: 0.30 [0.14; 0.45] & FR: 0.54 [0.44; 0.62]).
+The proportion of immatures was stable over time in the GB population, while a slight but significant decrease was observed for the FR population.
 
 Adult survival rate ranged from 0.33 [0.29; 0.37] to 1.01 [0.89; 1.14].
-No estimates were significantly outside the range of a survival rates defined without immigration [0; 1].
-No trend in adult survival rate was recorded in any of the two populations, although two main differences could be observed.
-The GB population had the lowest values of adult survival rate, and even if there were significant difference among years for both populations, the variability for the GB population was slightly lower.
+None of the estimates was significantly outside the range of a survival rate defined without immigration [0; 1].
+No trend in adult survival rate could be evidenced in any of the two populations, but patterns differed: adult survival in GB was both lower and displayed a weaker inter-annual variability than in FR.
 
-Recruitment rate ranges from 0.09 [0.04; 0.13] to 0.75 [0.63; 0.87].
-All estimates were above 0, meaning there was no senseless estimation.
-The maximum recruitment rate being 0.68 [0.36; 0.78] (see Section \@ref(MAX)), no estimates were significantly outside the range of recruitment rates defined without immigration [0; 0.78].
-Similarly to the immature proportion, no trend was observed for the GB population, but recruitment rate decreased for the FR population even if it was more noisy than for the immature proportion.
-As for survival rate, the GB population had low values of recruitment rate, and even if significant differences were observed among years for both populations, the variability for the GB population was far lower than for the FR population.
+Recruitment rate ranged from 0.09 [0.04; 0.13] to 0.75 [0.63; 0.87].
+All estimates were above 0, that is not incompatible with reality.
+Moreover, the maximum recruitment rate was 0.68 [0.36; 0.78] (see Section \@ref(MAX)), and no estimates were significantly outside the range of recruitment rates defined without immigration [0; 0.78].
+No trend was observed for the GB population, but recruitment rate decreased for the FR population even if it was more noisy than for the proportion of immatures.
+Like for survival, the GB population had lower values of recruitment rate with weaker variability than the FR one.
 
-The lower variability of both vital rates for the GB population explained its steady population growth over the considered time series compared to the FR one (Figure \@ref(fig:count)).
-In the FR population, the range of recruitment rates (min/max difference: 0.66) was higher than the range of survival rates (min/max difference: 0.58).
+Lower values and weaker variability of both adult survival and recruitment rates estimated in GB than in FR likely explained the differing trajectories of the two Ruddy duck populations, a sharp and constant decline for the former vs slow and variable decline for the latter (Figure \@ref(fig:count)).In the FR population, the range in recruitment rates (min/max difference: 0.66) was higher than the range of survival rates (min/max difference: 0.58).
 Conversely, the range of survival rates (min/max difference: 0.46) in the GB population was far higher than the range of recruitment rate (min/max difference: 0.08).
-This reflected that the variability of the population growth rate for the FR population was mostly driven by recruitment rate whereas the main source of variability was the adult survival for the GB population.
+This suggests that the variability of population growth rate was mostly driven by recruitment in FR whereas it was mainly driven by changes in adult survival in GB.
 
-## Likelihood of year-to-year variability in vital rates
+## Examining the reliability of the estimates
 
-From the five years available in the GB population, proportions of immatures that were estimated from the count-based method were positively correlated to estimates from removal data (Figure \@ref(fig:cor)).
-The hypothesis that the proportion of female-like individuals in the wintering populations reflected the age structure of these populations was then supported by this result.
-This correlation was even stronger when considering vital rates (Figure \@ref(fig:cor)).
-This stronger correlation was expected since the two methods estimating these vital rates had a component in common, the growth rate of the population (see Equation \@ref(eq:vr)).
-
-It was noticeable that a 1:1 correlation was never obtained.
-Proportion of immatures was lower in counts than in removals.
-This result was not surprising because immature individuals were expected to be more vulnerable to culling than adults and might then be over represented in removal data.
-As a consequence, the correlation coefficients for both vital rates also differed from 1.
-This difference between the two approaches did not question the previous validation of the ability of the count method to track the interannual variability of the vital rates.
+In Great Britain, proportions of immatures derived from the count-based method were positively correlated to those observed from culled individuals (Figure \@ref(fig:cor)).
+Therefore, this result supports the hypothesis that the proportion of female-like individuals in the wintering populations is efficient for assessing age structure and hence reproductive output.
+The correlation was even stronger when considering vital rates (Figure \@ref(fig:cor)), but this was expected because the two methods estimating these vital rates shared a component, the growth rate of the population (see Equation \@ref(eq:vr)).
+Interestingly, a 1:1 correlation is never reached for any parameters.
+The proportion of immatures derived from counts is always lower than those observed in culled individuals.
+The discrepancy between these two estimates is due to the fact that immatures are probably more vulnerable to culling than adults and might then be over-represented in culling samples.
+As a consequence, the correlation coefficients for both vital rates also differed.
+However, the differences between the two approaches do not question the ability of the count method to track the interannual variability of the vital rates.
 
@@
 
-## Response of vital rates to culling strategies {#MAX}
+## Assessing the demographic response to culling strategies {#MAX}
 
-Under no culling pressure, maximum growth rates were very similar for the two populations (Figure \@ref(fig:mgr)), respectively 1.45 [1.36; 1.55] and 1.52 [1.25; 1.88] for the GB and the FR population.
-It corresponded to an increase of 45% per year and 52% per year, respectively.
-Based on literature, we assumed a maximum survival rate similar for both populations 0.85 [0.7; 1].
-We deduced a maximum recruitment rate of 0.60 [0.42; 0.78] for the GB population and of 0.68 [0.36; 1.06] for the FR population.
+When culling pressure was nil, maximum growth rates were very similar for the two populations (Figure \@ref(fig:mgr)), respectively 1.45 [1.36; 1.55] and 1.52 [1.25; 1.88] for GB and FR.
+These values corresponded to a population increase of 45% and 52% per year, respectively.
+Assuming a maximum adult survival rate of 0.85 [0.7; 1] identical for both populations, we deduced maximum recruitment rates of 0.60 [0.42; 0.78] and 0.68 [0.36; 1.06] respectively for GB and FR.
 
-Even if a significant part of culled individuals were not aged for some years, the proxies of the culling pressure exhibit significant shifts over time (Figure \@ref(fig:sup)).
-This led to consider a high culling pressure period from 2006 onwards with no clear difference between the two populations because the signal was very noisy.
-Even if the culling pressure seemed comparable, the culling strategies differed significantly between the two populations, see Section \@ref(HARV).
-Under high culling pressure, the growth rate dropped to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for the GB and the FR population, respectively, which corresponded to a decrease of 44% per year and 3% per year.
-This result showed that the GB population significantly decreased during the high culling pressure whereas the FR population was stabilized.
+Despite low sample sizes of sexed and aged culled individuals, proxies of the culling pressure exhibited significant changes over time (Figure \@ref(fig:sup)).
+We thus ranked the culling pressure according to this intensity.
+High culling pressure occurred from 2006 onwards in both populations (although the signal was noisy).
+However, despite similar levels of culling pressure, the culling strategies differed significantly between the two populations, see Section \@ref(HARV).
+Under high culling pressure, the growth rate dropped to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for GB and FR, respectively, which corresponded to a decrease of 44% per year and 3% per year.
+This result showed that the GB population significantly decreased during the high culling pressure, whereas the FR population was just stabilized.
 
-Adult survival rate under high culling pressure was 0.44 [0.42; 0.46] and 0.62 [0.59; 0.64] for the GB and FR populations, respectively, when the recruitment rate decreases to 0.13 [0.11; 0.15] and 0.32 [0.29; 0.34], respectively.
-The drop in growth rate of the GB population corresponded to a similar average drop in adult survival rate (-0.41) and in recruitment rate (-0.47).
-The stabilisation of the FR population was caused by a decrease of recruitment rate (-0.36) more than a drop in adult survival (-0.23).
-Overall, vital rates of the GB population were more affected than the FR population.
+Adult survival rate under high culling pressure was 0.44 [0.42; 0.46] and 0.62 [0.59; 0.64] for GB and FR, respectively, and recruitment rates decreased to 0.13 [0.11; 0.15] and 0.32 [0.29; 0.34], respectively.
+The resulting drop in growth rate of the GB population corresponded to a similar average drop in adult survival rate (-0.41) and in recruitment rate (-0.47), whereas the stabilisation of the FR population was more caused by a decrease of recruitment rate (-0.36) than by a drop in adult survival (-0.23).
+Overall, culling strategy adopted in GB was more effective in affecting both adult survival and recruitment rates than that adopted in FR.
 
 ```{r mgr, fig.cap = "(ref:mgr)"}
-knitr::include_graphics("../Output/plot_paper_5_mod.png")
+knitr::include_graphics("../Output/plot_paper_5_saved.png")
 ```
@@
 
-The delayed sexual dichromatism in a waterfowl species was used here to develop and test the validity of a new non-invasive method aiming to break down the apparent growth of a population into its two structural components: adult survival rate and recruitment rate.
-Disentangling the contribution of these two vital rates is a key to understand the response of a population to management measures and potentially adapt these accordingly [@Williams2002].
-The development of the method held on a unique dataset of two comparable populations of the same species, the Ruddy duck.
-Both populations were tracked from their introduction in a similar environment to a period of intense culling pressure.
-This common history trajectory enabled to evaluate the maximum growth rate of this species, but also its response to culling pressure.
-As the culling strategy differed between the two populations, we detailed the similarities and the discrepancies of the responses of vital rates, and opened the discussion to the generalization of this estimation method and its implications for waterfowl management.
+Disentangling the contribution of the vital rates to population growth rate is a key step towards a better understanding of the response of a population to management actions [@Williams2002].
+We exploited delayed sexual dichromatism to develop a new approach based on counts to breakdown the apparent growth rate of a population into two of its main structural components: adult survival and recruitment rates.
+The development of the "apparent sex ratio" method holds on a unique dataset of two comparable populations of the same species, the Ruddy duck.
+Both populations were monitored in a similar way from their introduction to a period of intense culling pressure, but the culling strategies differed sharply.
+The very similar demographic trajectories and properties of both populations in similar ecological setup enabled evaluating the response to both culling pressure and culling strategies.
 
-## Accuracy of the method to estimate vital rates
+## Monitoring adult survival and recruitment rates
 
-The vital rate estimates of the two populations belonged to the same range, which demonstrates a certain consistency in the results obtained from the method.
-The greater variability observed in France is not unexpected because the FR time series covered a larger spectrum of culling pressure than the GB time series.
-There was no temporal correlation between the two populations, suggesting that fluctuations of the vital rates were more likely influenced by regional factors than large-scale ones.
-This is consistent with literature because both nesting/hatching success and juvenile survival in waterfowl species are conditional to the onset of laying, which is tightly related to local weather parameters, e.g.
-spring temperature, cumulative rainfall, and water levels [@Folliot2017; @Blums2004; @Dzus1998].
+For the two populations, the vital rate estimates were within the same range, which demonstrates a certain consistency in the results obtained from the method.
+The greater variability observed for the FR population was expected because the time series covered a larger spectrum of culling pressure than the GB one.
+We did not detect temporal correlation in the demographic parameters between the two populations, suggesting that regional (e.g., weather) rather than large-scale factors (e.g., climate) prevailed.
+This is consistent with previous findings showing for example that in waterfowl, both breeding success and juvenile survival are conditional to the onset of laying, which is tightly related to local weather (e.g.
+spring temperature, cumulative rainfall, water levels) [@Folliot2017; @Blums2004; @Dzus1998], and to local changes in predation pressure [@Jaatinen2022].
 
-Count surveys used to test the method occurred in mid-winter, when the first immature males were suspected to moult and start to look like adult males.
-This might partially corrupt the assumption that male-like individuals corresponded only to adult males.
-Even though no data on adult survival rate were directly available [@Baldassarre2014], the values estimated from the method developed inhere correspond to what is observed in literature for species of similar weight [@Nichols1997; @Krementz1997; @Buxton2004].
-Then, the results reflect that there was no major scaling issue using count data from mid-winter for European Ruddy duck populations.
-In any case, the time window of the count surveys must be chosen with care, i.e.
-just before the moult of immature males, to limit the bias of the outputs.
-On a few years, the values of adult survival rate were outside the range of expected values even if they were never significantly over 1.
-A likely reason to these extreme values is a corruption of the closed population assumption during some years.
-Indeed, if there is an arrival from another Ruddy duck population, the true recruitment cannot explain by itself the population growth and leads to an adult survival combining true adult survival and adult immigration.
-For the FR population, upper outliers of recruitment rates and survival rates were observed in 2002 and 2012, suggesting immigration events during these years.
+Count surveys used to implement our model occurred in mid-winter, that is when some immature males might initiate their moult.
+This could partially violate the assumption that male-like individuals were adult males only, and hence potentially bias our estimates.
+Nevertheless, as survival estimates are within the range of those reported for similar species [@Nichols1997; @Krementz1997; @Buxton2004; @Buderman2023], the assumption seems reliable.
+Therefore, the time window of the count surveys must be chosen with care to fit the time preceding the initiation of moult of immature males.
 
-The theoretical maximum of the population growth rate is a potential which is determined by intrinsic life history traits of a species [@Dmitriew2010; @Niel2005].
-This potential is expected to be similar to all Ruddy duck populations anywhere in the world.
-Realised maximum growth rates that were observed in this study were very close for the two populations.
-This suggests that the Ruddy duck species reached its intrinsic biological reproduction limit, at least given the environmental conditions of the western Palearctic, i.e.
-about 0.6 recruits per breeder.
-The consistency of this intermediate result suggests that the population size estimation was quite accurate for both populations and demonstrates that their changes were well tracked even with relatively few individuals.
-The average recruitment rates estimated from the count-based method, with and without culling pressure, were never higher to the proxies of maximum recruitment rate for both populations.
-The assumption on the accuracy of the count-based method is then again not excluded considering the consistency of this result.
+In some years, the estimated values of adult survival were outside the range of expected values even if they never significantly exceed 1.
+One possible explanation for these discrepancies is a violation of the "closed population" assumption.
+Indeed, despite ringing data suggest that Ruddy ducks do not undertake long-distance seasonal movements in GB (Henderson, pers.
+com.), one cannot totally rule out that immigration occurred.
+The out of range outliers of both recruitment and survival rates that were observed in 2002 and 2012 in FR suggest then significant immigration events during these years.
 
-Assuming that removal data provided a good picture of the year-to-year variability of age structure of a population, that is, any bias of shooting towards a given age or sex category is relatively stable over time, the strong correlation with the outputs of our method demonstrates its ability to track interannual fluctuations.
-This result is robust because such correlation was obvious despite being based on a short time series.
-This strong relationship was also satisfying because it demonstrates that the assumption of a constant adult sex ratio is not corrupted.
+The theoretical maximum population growth rate is an inherent trait related to the demographic characteristics of a species [@Dmitriew2010; @Niel2005].
+This trait is expected to be similar in all Ruddy duck populations anywhere in the world.
+In our study, we observed maximum growth rates that were very close for the two populations, which was expected because both populations experienced similar ecological conditions with an absence of harvesting and similar breeding conditions.
+Moreover, before the launch of the eradication programmes, the two populations also expressed a population growth much closer to the maximum expected values than any other native ducks species at the same period in the same areas.
+In the absence of culling, both populations expressed survival rates in the upper range of those observed in other diving ducks species (e.g.
+\~ 0.8 for Common pochard [@Folliot2020; @Nichols1997]).
+Ruddy duck tend to breed latter than both Common pochard and Tufted duck, yet nest predation has been found to decrease over the breeding season in a French fishpond complex [@Bourdais2015].
+Therefore, both enhanced survival and nesting success might have accounted for high population growth rates of Ruddy duck populations in the absence of culling.
+Besides providing estimates of adult survival and recruitment rates that were consistent with observed increasing population sizes, our method also provided estimates consistent with both the decreasing phases occurring after the implementation of eradication programmes and the observed differences in populations growth rates under differing culling strategies.
+As expected, higher values of survival and recruitment rates were observed when the culling strategy focused on post-breeding season rather than the pre-breeding one.
+
+Assuming that culling data provide a good picture of the year-to-year variability of age structure [@Fox2014], the strong correlation between estimates derived from counts and those derived from culled individuals demonstrates the ability of the "apparent sex ratio" method introduced here to properly track inter-annual fluctuations in demographic parameters such as survival and recruitment rates.
+The fact that the observed correlation was strong despite being based on a short time series strengthens its robustness.
+Furthermore, the strength of the relationship between parameters obtained through the two approaches also supports that important assumptions of our model were met, such as constant adult sex ratio.
 The temporal autocorrelation of the adult sex ratio might be explained by the fact that it integrates many age-cohorts, which makes it structurally strong.
-Even if significant variations of the adult sex ratio can still occur over long time periods, this result demonstrates that it is not necessary to monitor and update it on a yearly basis.
-The adult sex ratio found here on a restricted time period is in line with the Ruddy duck in its native area [0.62 in @Bellrose1980] and with other duck species [@Wood2021].
-The count-based method always provided recruitment rates significantly lower than the method based on removal data by a factor close to two, a result that is in line with the hypothesis that the shooting is generally selective towards immature individuals in waterfowl because they are more vulnerable to hunters than adults [@Bellrose1980; @Fox2014].
-The gap of vulnerability among ages observed here is realistic because the culling-induced mortality on first-year immatures compared to adults is higher by a factor ranging from 1.3 to 2.6 for hunted duck species in North America [@Bellrose1980].
+Significant variations of the adult sex ratio can occur over the long term, but our results suggests that it is not necessary to monitor and update it on a yearly basis.
+Proportions of males in adults estimated in the present paper (i.e.
+0.60) are within the range of those observed in native Ruddy duck populations of North America (0.62 in @Bellrose1980) and in other duck species [@Wood2021].
+The "apparent sex ratio" method always provided recruitment rates significantly lower than the method based on culling data by a factor close to two, a result that is compatible with a higher vulnerability of the immatures to shooting [@Fox2014].
+For instance, in North America, immatures have been found to be 1.3 to 2.6 times more vulnerable to hunting than adults in waterfowl [@Bellrose1980].
 
-Overall, even if the true values of vital rates are not known to properly test the accuracy of the count-based method, a beam of arguments supports that the method is not heavily biased.
-This consists in a big difference with methods based on removal/hunting bag data as demonstrated in @Fox2014, even when one considers an age structure picture that is limited to the end of the hunting season to limit the bias [@Fox2016].
+Because estimates of survival were not available in the literature (see @Buderman2023), we could not properly assess the accuracy of the "apparent sex ratio" method for estimating this parameter.
+Nevertheless, circumstantial evidence support that the method is not heavily biased as detailed above.
+Thus the "count-changing sex ratio" method would be much more efficient for monitoring key demographic parameters than alternative methods such as those based on culling/hunting bag data that proved unreliable [@Fox2014], even when they focused on the end the hunting season to limit possible biases [@Fox2016].
 
-## Population response to management measures
+## Assessing eradication strategies
 
-Both populations responded to culling pressure by a large drop of their growth rate, meaning that the compensation of the population to culling mortality was limited if not null.
-The eradication effort as implemented here was therefore efficient to impact this species.
-The response of the GB population differed from the FR one because it showed a strong decrease when the FR population got stabilized.
-The analysis of the growth rate alone did not provide clues to understand the mechanisms explaining the difference of the responses between the two populations.
-Indeed, the bigger drop of growth rate of the GB population might be explained either by a decrease of adult survival or reproduction success, or both components.
+Both populations responded to culling pressure by a significant decrease in their growth rate, meaning that increasing mortality due to culling was neither compensated by increasing natural survival nor by improved breeding success.
+The culling effort was therefore efficient to affect the demographic trajectory of Ruddy duck in both contexts.
+However, the decrease in population growth was much stronger in GB, where the population displayed a sharp decrease, than in FR where the increase was stopped and population size remained more or less stable.
+The greater decline in population growth observed in GB compared to FR was due to greater decline in both adult survival and recruitment rates.
+Overall, in both countries, adult survival and recruitment rates seemed significantly affected by culling.
+However, the culling pressure occurring before the breeding season, which affect only future breeders (pre-breeding culling strategy applied in GB), seemed much more efficient in limiting both survival and recruitment rates than the culling pressure occurring during nesting and young raising, which targets breeders and newly born individuals (breeding culling strategy applied in FR).
 
-The vital rates that were estimated thanks to the count-based method supported that the eradication pressure affected negatively both adult survival and recruitment rates in the two populations on average.
-The culling pressure occurring mostly before the breeding period for the GB population targeted by definition future breeders, and consequently induced mortality on these individuals and prevented them from reproducing.
-We observed that adult survival and recruitment rates decreased equally, which is a result in line with the expectation of this culling strategy.
-For the FR population, the culling pressure mostly occurred during the postbreeding period.
-This strategy should induce a higher mortality on immatures as they are expected to be more vulnerable to hunting [@Bellrose1980; @Fox2014].
-However, this anthropogenic mortality on immatures might be compensated by a higher survival of the future recruits because competition for resources is decreased [@Cooch2014].
-Recruitment rate of the FR population being more affected by 50% than adult survival, the results support the hypothesis that no or only a limited compensation occurred.
-This conclusion is even strengthened by the result on the GB population.
-Indeed, we should expect a decrease of recruitment rate lower than adult survival rate if immatures would take advantage of additional resources getting unexploited by culled adult individuals.
-The equal decrease of the two vital rates supports that there is no such compensation.
-This might be explained by the fact that the population size was far lower than the carrying capacity of the available habitats, so there was no competition for resources that potentially enhanced the survival of the remaining immature individuals [@Peron2013].
+As expected, the pre-breeding culling strategy had the same effect on adult survival and recruitment rates because a bird killed before the breeding period reduces the breeding population (lower adult survival rate) and prevents this bird from reproducing (lower recruitment rate).
+When the breeding culling strategy was applied, the recruitment rate was more affected than adult survival.
+This was probably due to immatures being much more vulnerable to shooting than adults [@Bellrose1980; @Fox2014].
+The compensation hypothesis suggests that the higher harvest mortality of immatures can be compensated by increasing natural survival [@Cooch2014] but it is not supported by the observations.
+Overall, there was no evidence for any compensation mechanism in either context.
+This might result from the fact that both populations were far of reaching carrying capacity meaning that competition for resources was relaxed [@Peron2013].
 
-If we were in a situation where shooting mortality was fully compensatory to natural mortality, the FR strategy would be counter-productive.
-However, since the studied Ruddy duck populations are expanding, it seemed that the postbreeding culling strategy still provided some results to control such populations.
-However, the analysis of the year-to-year variability of recruitment rate demonstrates that the prebreeding strategy strongly buffered the recruitment and maintained it to a low level every single year, whereas recruitment rate of the postbreeding strategy was highly variable.
-This might be because culling future breeders ensured to avoid the production of potential recruits whereas the postbreeding culling strategy could not prevent and dampen exceptional reproduction events.
-The postbreeding strategy is also affected by the highly variable detectability of immatures, as these individuals spread over many reproduction habitats with variable accessibility to observers/shooters, whereas the prebreeding strategy targeted individuals that were more easily localized because they usually gathered in open water on a limited number of wintering spots [@Johnsgard1996].
+For more or less comparable culling pressure, the pre-breeding culling strategy proved much more effective.
+This does not mean however, that the breeding culling strategy was not so.
+On the contrary, it proved rather efficient for stopping both the population increase and range expansion in a first time, and in second time to trigger a significant population decline by slightly increasing culling pressure.
+Thus when winter culling is not possible, which was the case in FR, culling during the breeding season is effective for stopping or reversing population increases in such species.
+The breeding culling strategy is however affected by the highly variable detectability of immatures, as these individuals colonise new ponds.
+This results in fluctuations of the recruitment rate and then threatens the long term efficiency of such strategy.
+The winter culling strategy does not face such problem because of the habit of Ruddy ducks to gather in a limited number of sites during winter [@Johnsgard1996].
 
 The response of the two Ruddy duck populations to culling pressure demonstrates the necessity to account for the culling season to produce proper predictions.
-Predicting the impact of a harvest pressure on a waterfowl population is then not only a question about how big is the harvest effort, but also mostly when this effort occurs within a year [@Kokko1998].
-More generally, disentangling role of adult survival from that of recruitment rate brings capital information to understand the process of harvest on population dynamics and brings key knowledge to improve the conservation of hunted species or the control of invasive species.
+Predicting the impact of a harvest pressure on a waterfowl population is then not only a question about how big is the harvesting effort, but also when harvesting occurs [@Kokko1998].
+Counter-intuitively, our results suggest that harvesting waterfowl during breeding has far less impact on population growth rate than harvesting during winter.
+As a result, a target for effective management of harvested waterfowl should be earlier onsets and endings of hunting seasons.
 
-## Implication for waterfowl management
+## Waterfowl management implications
+
+One major implication of our study concerns the timing of harvesting for limiting its impact on populations growth rate.
+Counter-intuitively, our results show that the persistence of waterfowl populations should benefit from earlier openings and closings of the hunting seasons (see above).
 
 Tracking variations in abundance is a common tool to determine the conservation status of a population [e.g @Folliot2022].
-However, abundance alone does not provide information on the mechanisms behind the changes of population growth [@Williams2002].
-Identifying the determinants requires to dig into the components structuring population growth to provide relevant tools to managers [e.g.
-@Austin2000].
-Gaining knowledge on adult survival and recruitment rates often involves capture mark recapture surveys [e.g.
+But tracking abundance alone does not provide enough information for assessing the underlying mechanisms behind the changes of population growth [@Williams2002].
+In order to do that, one must “dig deeper” [@Austin2000], for example by monitoring individuals to assess parameters like survival and recruitment rates [e.g.
 @Arnold2018].
-However, tracking fine time-scale variability of the vital rates requires to mark a minimum number of individuals [e.g.
-@Souchay2016] and prevents from implementing such method to small populations with a poor conservation status for instance.
-The method introduced here bypasses these caveats to provide annual estimates of vital rates.
+However, monitoring individuals is time consuming [e.g.
+@Souchay2016] and not always possible in endangered species.
+The method proposed here allows circumventing these drawbacks in dichromatic species with delayed maturity of males.
+Among these species, one can mention the White headed duck in Southern Spain, which is highly endangered and cannot be safely monitored through capture-mark-recapture.
+Thus by improving census strategies and teasing apart the effects of variations of adult survival and productivity on population growth through the approach presented here, one could gather much more invaluable information about the pertinence of management strategies.
 
-The immature proportion, or age ratio, is commonly used to describe the renewal capacity of a waterfowl population and to track the changes of its productivity [@Smith2001; @Bellrose1980; @Robertson2008; @Rodway2015; @Zimmerman2010].
+<!-- The immature proportion, alternatively the age ratio, is commonly used to describe the renewal capacity of a waterfowl population and to track the changes of its productivity [@Smith2001; @Bellrose1980; @Robertson2008; @Rodway2015; @Zimmerman2010].
 The recruitment rate however remains a better indicator than the immature proportion/age ratio, the latter only having the advantage of being more directly accessible.
@@
 Fecundity alone does not provide all knowledge that is required for a manager to understand the dynamics of a population.
+-->
 
-Applying the method implemented in this study to any other waterfowl species implies that the adult males can be distinguished from the other individuals at a distance.
-The latter in the season the dichromatism is observable, the better it is to ensure a relevant picture of the age structure of the population and then good estimates of adult survival and recruitment rates.
-This pattern occurs in most of stifftail duck species [@Johnsgard1996], and in many other duck species, among with Tufted duck, *Aythya fuligula*, Black scotter, *Melanitta nigra americana*, Common Goldeneye, *Bucephala clangula americana* [@Johnsgard1978; @Bellrose1980].
-Modifying standard monitoring protocols to distinguish male-like from female-like individuals is almost costless but would provide substantial increases in the efficiency and usefulness of monitoring results in conservation [@Nichols2006].
+Finally, we want to stress the fact that the main aim of our approach was to decompose population growth into its two main components and not to provide unbiased estimates of adult survival or recruitment rates.
+However, if needed, censuses could be designed in such a way that would make this possible, at least in theory, and in species presenting an observable delayed dichromatism.
+This property concerns most stiff-tailed duck species [@Johnsgard1996], as well as numerous other duck species, among which Tufted duck, *Aythya fuligula*, Black scotter, *Melanitta nigra americana*, Common Goldeneye, *Bucephala clangula americana* [@Johnsgard1978; @Bellrose1980].
+Modifying standard monitoring protocols to distinguish male-like from female-like individuals is almost costless but worth the bet because it would provide substantial increases in the efficiency of conservation/management actions [@Nichols2006].
 
-# Acknowledgments {-}
+# Acknowledgments {.unnumbered}
 
@@
 This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the *Office Français de la Biodiversité*, and Jean-Marc Gillier from the *Société Nationale pour la Protection de la Nature*.
-The authors acknowledge all the contributors of the data collection; in Great Britain, Wildfowl & Wetlands Trust staff past and present, particularly Baz Hughes, Peter Cranswick and Colette Hall, and all the project workers at the Animal and Plant Health Agency and its predecessors; in France, Office Français de la Biodiversité staff past and present, especially Vincent Fontaine, Denis Lacourpaille, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface, and the whole team of the Société Nationale pour la Protection de la Nature of the Grand Lieu lake.
+The authors acknowledge all the contributors of the data collection; in Great Britain, Wildfowl & Wetlands Trust staff past and present, particularly Baz Hughes, Peter Cranswick and Colette Hall, and all the project workers at the Animal and Plant Health Agency and its predecessors; in France, Office Français de la Biodiversité staff past and present, especially Vincent Fontaine, Denis Lacourpaille, Agathe Pirog, Hugo Pichard, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface, and the whole team of the Société Nationale pour la Protection de la Nature of the Grand Lieu lake.
 
@@
 
-# References {-}
+# References {.unnumbered}
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 
-# (APPENDIX) sup {-}
+# (APPENDIX) sup {.unnumbered}
 
@@
 
-(ref:count) Temporal evolution of the number of Ruddy ducks (log scale) in Great Britain (GB) and France (FR) from 1960 to 2019, with indications of periods when apparent sex ratios were also recorded.
+(ref:count) Temporal changes in the number of Ruddy ducks (log scale) in Great Britain (GB) and France (FR) from 1960 to 2019, with indications of periods when apparent sex ratios were also recorded
 
-(ref:flock) Typical observation of a Ruddy duck flock during winter including ten female-like individuals with a whitish striped cheek (some of which are immature males), four male-like individuals with a white cheek and a black cap, three probable males individuals (1^st^, 4^th^, and 10^th^ from left) © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
+(ref:flock) A Ruddy duck flock observed in winter including ten female-like individuals with a whitish striped cheek (some of which are immature males), four male-like individuals with a white cheek and a black cap (all adult males), three unidentified individuals (1^st^, 4^th^, and 10^th^ from left) © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
 
-(ref:pfem) Proportion of females in removals vs in counts; For the first category, one data point corresponds to the proportion of females estimated from direct sex identification of the Ruddy ducks culled over one year in a population; for the second category, one data point corresponds to the proportion of individuals wearing female-like plumage in a population counted in winter; the difference of proportion of females between the two approaches is due to immature males that look like females before moulting; this difference is at the root of the method introduced inhere to disentangle adult survival from recruitment
+(ref:pmal) Proportions of males estimated through the internal examination of individuals shot as part of the eradication programme over one year in a population (Male % in cullings) and proportions of individuals displaying a male-like plumage in a population counted in winter (Male % in counts); discrepancies between the two estimates are due to delayed dichromatism (immature males looking like females before moulting); we exploited these discrepancies for estimating survival and productivity/recruitment
 
-(ref:vital) Changes in the proportion of immatures and the vital rates in light of the culling pressure for both populations; bars define the 95% confidence intervals
+(ref:outline) Outline of the method to disentangle the population growth into survival (adult survival rate) and productivity (recruitment rate)
 
-(ref:cor) Comparison of the parameter estimates between the count-based method and removal-based method; only five years for the GB population were available; bars define the 95% confidence intervals; the red arrow indicates the average difference between the two methods
+(ref:vital) Temporal changes in the proportions of immatures and the vital rates following the culling pressure on Ruddy duck populations of Great Britain (GB) and France (FR); vertical bars represents the 95% confidence intervals.
 
-(ref:mgr) Comparison between average population growth rate under no pressure vs under high culling pressure; the culling effort for the GB population moslty occurred during the prebreeding period, whereas it mostly focused during and after the breeding period for the FR population (see \@ref(HARV) for details); the average vital rates during the time period of maximum growth were proxies deduced from literature and estimates of maximum population growth (see \@ref(VAL) for details); the vital rates under high culling pressure were average values over a time period observing a consistent culling pressure; bars define the 95% confidence intervals
+(ref:cor) Relationship between parameter estimates derived from counts (count-based method) and those obtained from culled individuals (cull-based method); only five years were available for the GB population; bars represent 95% confidence intervals and the red arrow indicates the direction and average differences between the two methods
 
-(ref:sup) Definition of three periods of different culling pressures from an index based on an estimation of adult culling rate; as only half of the culled individuals were aged in France before 2009, we applied the average age ratio over 2009-2019 on unaged individuals to provide an estimate of the number of adults in removals before 2009; culling rate increased over time, we therefore split the time series in three categories : we defined a "no culling" period before 1999 because the culling rate for both countries was mostly null and always below 10%, we defined a "low culling" between 1999 and 2005 because the culling rate varied around 20% for both countries, the culling rate then raised significantly for both countries, so we defined a "high culling" period from 2006 onwards
+(ref:mgr) Average effects of culling pressure and different culling strategies (GB vs FR) on population growth rates (derived from counts only), and on adult survival and recruitment rates (derived from counts and apparent sex ratios); the culling pressure for the GB population moslty occurred during the prebreeding period, whereas it mostly focused during and after the breeding period for the FR population (see \@ref(HARV) for details); adult survival and recruitment rates under high culling pressure were averaged over a time-period observing a consistent culling pressure; vertical bars show 95% confidence intervals
+
+(ref:sup) Definition of three periods of different culling pressures from an index based on an estimation of adult culling rate; as only half of the culled individuals were aged in France before 2009, we applied the average age ratio over 2009-2019 on unaged individuals to provide an estimate of the number of adults in cull data before 2009; culling rate increased over time, we therefore split the time series in three categories : we defined a "no culling" period before 1999 because the culling rate for both countries was mostly null and always below 10%, we defined a "low culling" between 1999 and 2005 because the culling rate varied around 20% for both countries, the culling rate then raised significantly for both countries, so we defined a "high culling" period from 2006 onwards

### 2024-05-27 PCItemplate-main/manuscript.Rmd
Before instatext

@@
     affiliation: 1
-  - firstname: Alain
-    lastname: Caizergues
+  - firstname: Léonard
+    lastname: Heinry
     affiliation: 1,2,a
 lastauthor: 
-  - firstname: Carol #add last author here if more than one author for proper styling
-    lastname: Senior #can be kept empty/deleted otherwise
+  - firstname: Alain #add last author here if more than one author for proper styling
+    lastname: Caizergues #can be kept empty/deleted otherwise
     affiliation: 1,a

### 2024-05-27 Code/Identifiability diagnostic functions.R

@@ -1,73 +1,0 @@
-hessian <- function(funfcn, x, y, delta){
-  # function to calculate the hessian of funfcn at x
-  t <- length(x)
-  h <- matrix(0, t, t)
-  Dx <- delta * diag(t)
-  for (i in 1:t) {
-    100
-    for (j in 1:i) {
-      h[i,j] <- (do.call("funfcn", list(x + Dx[i,] + Dx[j, ], y)) # LL with
-                 - do.call("funfcn", list(x + Dx[i,] - Dx[j, ], y))
-                 - do.call("funfcn", list(x - Dx[i,] + Dx[j, ], y))
-                 + do.call("funfcn", list(x - Dx[i,] - Dx[j, ], y))) / (4 * delta^2)
-      h[j, i] <- h[i, j]
-    }
-  }
-  return(h)
-}
-
-hessianmethod <- function(funfcn, pars, y, delta, print = TRUE){
-  # Applies the Hessian method
-  # funfcn - function which return negative loglikelihood
-  # pars - parameter values at which hessian is evaluated
-  # y - data to be passed into likelihood
-  # delta - error used in calculating Hessian and cutt-off
-  # suggested value delta = 0.00001
-  # cut-off used delta*p, where p is no. pars
-  cutoff <- delta*length(pars)
-  # Finds hessian matrix
-  h <- do.call("hessian", list(funfcn, pars, y, delta))
-  # Calculates eigenvalues
-  E <- eigen(h)
-  # find standardised eigenvalues
-  standeigenvalues <- abs(E$values) / max(abs(E$values))
-  # find number of estimable parameters
-  # number of parameters with eigenvalues below the cutoff
-  noestpars <- 0
-  for (i in 1:length(pars)) {
-    if (standeigenvalues[i] > cutoff) {
-      noestpars <- noestpars + 1
-    }
-  }
-  if (print) {
-    # Prints whether model is parameter redundant or not
-    101
-    # Then prints smallest eigenvalue and
-    # number of estimable parameters
-    if (min(standeigenvalues) < cutoff) {
-      cat("model is non-identifiable or parameter redundant")
-    }
-    else {
-      cat("model is identifiable or not parameter redundant")
-    }
-    cat("\n")
-    cat('smallest standardized eigenvalue', min(standeigenvalues))
-    cat("\n")
[…]

### 2024-05-27 Code/Model_vital_rates.R

@@ -84,7 +84,7 @@ cor_plot <- function(col, title = NULL, subtitle = NULL, limits = NA, breaks = N
                   width = 0, alpha = .5, linewidth = 1.2) +
     labs(title = title,
          subtitle = subtitle,
-         y = "count-based method", x = "removal-based method") +
+         y = "count-based method", x = "culling-based method") +
     scale_x_continuous(expand = expansion(mult = c(0, 0)), 
                        limits = if (any(is.na(limits))) {
                          c(0, 1.1 * max(unlist(tmp[-1])))
@@ -243,29 +243,29 @@ frag %>%
 ## intro_plot ------------------------------------------------------------------------
 
 count_sex_app %>% 
-  mutate(method = "Female %\nin counts",
+  mutate(method = "Male %\nin counts",
          tot = obs_type_mal + obs_type_fem,
-         `female proportion` = obs_type_fem / tot) %>% 
-  select(method, `female proportion`, tot) %>% 
+         `male proportion` = obs_type_mal / tot) %>% 
+  select(method, `male proportion`, tot) %>% 
   bind_rows(frag %>%  
               filter(sex != "ind") %>% 
               group_by(year, pop) %>% 
               mutate(tot = sum(shot)) %>%  
-              filter(sex == "fem") %>% 
-              mutate(fem = sum(shot)) %>% 
-              distinct(year, pop, fem, tot) %>%
+              filter(sex == "mal") %>% 
+              mutate(mal = sum(shot)) %>% 
+              distinct(year, pop, mal, tot) %>%
               ungroup() %>% 
               filter(tot > 100) %>%
-              mutate(`female proportion` = fem / tot, 
-                     method = "Female %\nin removals") %>%
-              select(method, `female proportion`, tot)) %>% 
+              mutate(`male proportion` = mal / tot, 
+                     method = "Male %\nin cullings") %>%
+              select(method, `male proportion`, tot)) %>% 
   mutate(method = as_factor(method)) %>% 
-  ggplot(aes(x = method, y = `female proportion`, size = tot)) + 
+  ggplot(aes(x = method, y = `male proportion`, size = tot)) + 
   geom_jitter(width = 0.2, alpha = .5) +
   scale_y_continuous(name = "", limits = c(0, 1), breaks = 0:10/10, labels = scales::percent) +
   guides(size = guide_legend(title = "Sample size")) +
   xlab("") + 
[…]

### 2024-05-28 Code/Draft_ruddy_duck_vital_rates.Rmd
after instatext

@@
 abstract: |
-  Monitoring the number of individuals is by far the most popular strategy for investigating the environmental factors ruling population dynamics and for measuring the efficacy of management actions targeting population recovery, control or eradication.
-Monitoring numbers is unfortunately inefficient for assessing the mechanisms underlying demographic processes and more specifically to assess the extent to which population growth rate is affected by changes in adult survival rather than fluctuations in reproductive parameters.
-Usual methods to inform adult survival and productivity, such as capture-mark-recapture and analysis of game-hunting bag, rely on catching animals which suffers from three main drawbacks.
-Firstly, the selectivity of the catching methods can lead to a biased representation of the underlying population structure.
-Secondly, catching and releasing animals is quite invasive, which generates stress, and is thus an additional threat to often already endangered species.
-Thirdly, releasing may simply be impossible especially for invasive alien species that represent a threat to native taxa.
-In many waterfowl species, a sexual dichromatism is observed in adults whereas immatures of both sexes display a cover-up plumage similar to adult females.
-Using the case of alien Ruddy duck populations of Western Europe (a species displaying this pattern), we introduce a non-invasive method based on count data distinguishing plumage types that allows estimating the respective contribution of adult survival and productivity (also called recruitment rate) to population growth rate.
-We exemplify how, based on count surveys, changes in apparent male proportion (due to female-like immature males progressively acquiring adult male plumage) can be exploited to decompose population growth rate into adult survival and recruitment rates, greatly helping to understand the demographic impact of two different eradication strategies.
+  Monitoring the number of individuals is by far the most popular strategy for studying the environmental factors that determine population dynamics and for measuring the effectiveness of management actions aimed at population recovery, control, or eradication.
+unfortunately, monitoring numbers is inefficient in identifying the mechanisms underlying demographic processes and, in particular, in assessing the extent to which population growth rate is influenced by changes in adult survival rather than variations in reproductive parameters.
+The usual methods for determining adult survival and productivity, such as analysing capture-mark-recapture data and game-hunting bag data, rely on capturing animals, which has three major drawbacks.
+Firstly, the selectivity of capture methods can lead to a distorted representation of the underlying population structure.
+Secondly, the capture and release of animals is quite invasive, causing stress and posing an additional threat to the often already endangered species.
+Thirdly, release may simply be impossible, especially for invasive alien species that represent a threat to native taxa.
+In many waterfowl species, sexual dichromatism is observed in adults, while immatures of both sexes display a cover-up plumage similar to that of adult females.
+Using two non-native populations of Ruddy duck, a species that exhibits this pattern, we present a non-invasive method to estimate the respective contribution of adult survival and productivity (also called recruitment rate) to the population growth rate.
+We exemplify how changes in the apparent proportion of males in counts (due to female-like immature males gradually acquiring adult male plumage) can be exploited to split the population growth rate into adult survival and recruitment rates, which contributes significantly to understanding the demographic impact of two different eradication strategies.
 
-  **Keywords**: intrinsic growth rate - vital rate - reproductive success - fecundity - sex ratio - age ratio - Oxyura jamaicensis
+  **Keywords**: intrinsic growth rate - vital rate - reproductive success - fecundity - sex ratio - age ratio - Oxyura jamaicensis - harvesting
 output: 
@@
 
-Assessing population growth rate is a first step towards a better understanding of factors underlying the dynamics of natural populations [e.g.
+Assessing the growth rate of populations is a first step towards a better understanding of the factors underlying the dynamics of natural populations [e.g.
 @Niel2005].
-It is also crucial for measuring the efficacy of management actions potentially undertaken to help population recovery, control, or eradication [@Shea1998].
-Among the approaches available to managers for reaching these goals, those relying on the monitoring of the number of individuals (counts) are by far the most popular [@Rintala2022].
-In many instances, these methods enable investigating environmental factors underlying changes in population size, and hence help implementing mitigating actions [@Faillettaz2019].
-Unfortunately, relying on the monitoring of population abundance alone generally hinders basic understanding of the demographic mechanisms underlying changes in population growth rates.
-More specifically, based on counts only, it is nearly impossible to assess the extent to which population growth rate is affected by changes in adult survival rather than to variations in productivity [@Austin2000], which is defined as the product of all reproductive parameters.
-To assess the relative sensitivity of population growth rate to factors affecting adult survival or reproductive parameters, demographers most often rely on the monitoring of individuals by capture-mark-recapture [CMR, @Lebreton1992].
-In practice, however, one cannot always rely on the capture and release of individuals owing to the legal status of the species that precludes any additional disturbance (e.g.
-critically endangered species) or forbid the release of alive individuals (e.g.
+It is also crucial for measuring the effectiveness of management actions that are taken to recover, control, or eradicate populations [@Shea1998].
+Among the approaches available to managers to achieve these goals, those that rely on monitoring the number of individuals (counts) are by far the most popular [@Rintala2022].
+In many cases, these methods allow the investigation of environmental factors underlying changes in population size, and thus help in the implementation of mitigating actions [@Faillettaz2019].
+Unfortunately, relying solely on the monitoring of population abundance generally hinders the fundamental understanding of the demographic mechanisms underlying changes in population growth rates.
+More specifically, on the basis of counts alone, it is almost impossible to assess the extent to which population growth rate is affected by changes in adult survival rather than to variations in productivity [@Austin2000], which is defined as the product of all reproductive parameters.
+To assess the relative sensitivity of population growth rate to factors affecting adult survival or reproductive parameters, demographers usually rely on the monitoring of individuals by capture-mark-recapture [@Lebreton1992].
+In practise, however, capture and release of individuals cannot always be relied upon, as the legal status of the species precludes any additional disturbance (e.g.
+critically endangered species) or prohibits the release of live individuals (e.g.
 invasive species or pests).
-Furthermore, although highly efficient for assessing demographic parameters, capture-mark-recapture methods suffer from potential drawbacks such as being invasive (and may affect behaviour and hence survival or breeding success) and hardly affordable when time and money are limiting.
-Genetic monitoring is a non-invasive alternative to capture-mark-recapture, but it suffers from being costly and requires large samples [@Caniglia2011].
+Even though capture-mark-recapture methods are efficient for assessing demographic parameters, they have potential drawbacks, such as their invasiveness, which can affect behaviour and thus survival or reproductive success, and their affordability when time and money are scarce.
+Genetic monitoring is a non-invasive alternative to capture-mark-recapture, but it is costly and requires large samples [@Caniglia2011].
 
-Most of the time therefore, managers do the best of a bad job using counts as a viable option for tracking population size changes and then population growth rate, in order to assess the relevance of management actions.
-Distinguishing immatures from adults in counts enables estimating adult survival rate, i.e.
-the proportion of breeders that survived over one year, and productivity/recruitment rate, i.e.
+Most of the time, therefore, managers make the best of a bad job by using counts as a viable option for tracking changes in population size and then population growth rate to assess the relevance of management actions.
+Distinguishing between immatures and adults in counts allows estimation of adult survival rate, i.e.
+the proportion of breeders that have survived for one year, and productivity/recruitment rate, i.e.
 the number of immatures produced per breeder that reach sexual maturity, which are, by definition, recruited in the breeding population.
-This additional effort makes it possible to assess the relative influence of each of these components on population growth rate.
-Alternatively, assessing the age structure in hunting bags has been used to infer the role of decreasing reproductive success in population declines in a number of games species including ducks and geese but suffers from intractable biases [@Fox2018].
-Of course, the latter approach is not suited for protected/endangered species.
+This additional effort makes it possible to assess the relative influence of each of these components on the population growth rate.
+Alternatively, the assessment of age structure in hunting bags has been used to infer the role of decreasing reproductive success in population declines in a number of game species, including ducks and geese, but suffers from intractable biases [@Fox2018].
+Of course, the latter approach is not appropriate for protected/endangered species.
 
-Here we exemplify how, in dimorphic species, delayed maturity of males can be used for estimating adult survival and recruitment rate by differentiating male-like from female-like individuals in repeated winter counts.
-We used two non-native Ruddy duck (*Oxyura jamaicensis*) populations introduced in Europe as a study model.
-As numerous duck species, Ruddy duck is dimorphic, with newly born males looking like females until the prenuptial moult, which occurs in late winter.
-As a result of this late prenuptial moult of immatures, the apparent proportion of males increases over the course of the wintering season.
-These changes in apparent proportions of males during this period are therefore directly related to the proportion of immatures into the populations and thus to the reproductive success of the previous breeding season.
-We estimated adult survival and recruitment rates and assessed the impact of two different eradication strategies deployed in Great Britain and in France, respectively.
+Here we exemplify how, in dimorphic species, delayed sexual maturity of males can be used to estimate adult survival and recruitment rates by distinguishing male-like and female-like individuals in repeated winter counts.
+We used two non-native European populations of Ruddy duck (*Oxyura jamaicensis*) as a study model.
+Like many other duck species, Ruddy duck is dimorphic: newborn males look like females until the prenuptial moult, which takes place in late winter.
+As a result of the late prenuptial moult of immatures, the apparent proportion of males increases over the course of the wintering season.
+These changes in the apparent proportion of males during this period are therefore directly related to the proportion of immatures in the populations and thus to the reproductive success of the previous breeding season.
+The "apparent sex ratio" method developed in this study was used to estimate adult survival and recruitment rates and evaluate the effects of two different eradication strategies used in Great Britain and France, respectively.
 
@@
 
-The Ruddy duck is a stiff-tailed duck native from the Americas.
-From seven individuals initially acclimated in the Slimbridge Wetland Centre in the 1940's [@GutierrezExposito2020], a feral population began to establish in Great Britain with the first observed reproductive attempts in the wild in the 1960's (Figure \@ref(fig:count)).
-This feral population rapidly grew and spread to the entire country to reach more than 5,000 individuals in the early 2000s.
-By the end of the 1980's, a feral population began to establish also on the continent, especially in France, supposedly owing to the arrival of individuals born in Great Britain.
-But contrary to what was observed in Great Britain, the Ruddy ducks did not spread much in France, with the vast majority of observations and breeding attempts clustering in the North-West of the territory.
-In France, during winter, almost no Ruddy duck is observed outside the lake of Grand Lieu (47.09°N, 1.67°W), which greatly facilitates the monitoring of this population.
-Ruddy ducks are considered as a major threat to the endangered native White-headed duck (*Oxyura leucocephala*) of the South-Western Mediterranean, because of hybridization and thus an elevated risk for genetic pollution and genetic assimilation of the latter by the former [@MunozFuentes2007].
-In order to mitigate the risks of genetic pollution of the White-headed duck by Ruddy ducks, eradication measures were implemented in the late 1990's both in Great Britain and in France [@GutierrezExposito2020], and were followed by a European Ruddy duck eradication plan in 1999 [@Hughes1999] (Figure \@ref(fig:count)).
+The Ruddy duck is a stiff-tailed duck native to the American continent.
+Starting from seven individuals acclimatised at the Slimbridge Wetland Centre in the 1940s [@GutierrezExposito2020], a feral population began to establish in Great Britain and the first attempts at reproduction in the wild were observed in the 1960s (Figure \@ref(fig:count)).
+This feral population grew rapidly and spread across the country until it reached more than 5,000 individuals in the early 2000s.
+In the late 1980s, a feral population also began to establish on the continent, particularly in France, supposedly due to the arrival of British-born individuals.
+However, in contrast to observations in Great Britain, Ruddy ducks have not spread very much in France.
+Most observations and breeding attempts have been concentrated in the north-west of the country.
+In France, almost no Ruddy ducks were observed in winter outside the lake of Grand Lieu (47.09°N, 1.67°W), which greatly facilitated the monitoring of this population.
+Ruddy ducks are considered a major threat to the endangered native White-headed duck (*Oxyura leucocephala*) in the south-western Mediterranean, as they hybridise and thus pose an increased risk of genetic pollution and genetic assimilation of the latter by the former [@MunozFuentes2007].
+To reduce the risk of genetic pollution of the White-headed duck by the Ruddy duck, eradication measures were implemented in both Great Britain and France in the late 1990s [@GutierrezExposito2020], followed by a European Ruddy duck eradication plan in 1999 [@Hughes1999] (Figure \@ref(fig:count)).
 
-The trends and status of populations of both White-headed duck and Ruddy duck are being assessed through population censuses only.
-It is thus impossible to assess the factors that impair recovery of the former and the efficacy of the eradication program for the latter.
-In particular, based on censuses alone, one cannot assess the relative effects of changes in adult survival and recruitment rates on population growth rates which are prerequisites to identifying limiting factors.
+The trends and status of the White-headed duck and Ruddy duck populations used to be determined exclusively by counts.
+It was therefore not possible to assess the factors impairing the recovery of the former and the effectiveness of the eradication programme for the latter.
+In particular, counts alone cannot be used to assess the relative impact of changes in adult survival and recruitment rates on population growth rates, which is a prerequisite for identifying limiting factors.
 
@@
 
-Like many other ducks, Ruddy ducks and White-headed ducks both display delayed dichromatism thereby young males acquire the typical colourful plumage of their kind, and hence can be distinguished from females, over the course of the interbreeding season, typically by mid-winter for the earliest individuals [@Baldassarre2014].
-Delayed dichromatism generally explains the discrepancies in the estimates of apparent proportions of males counted in winter vs counted in spring, or vs males determined from individuals culled in winter (Figures \@ref(fig:flock) & \@ref(fig:pmal)).
+Like many other duck species, both White-headed duck and Ruddy duck exhibit delayed dichromatism, i.e.
+young males acquire the colourful plumage typical of their species and can therefore be distinguished from females during the pre-breeding period, typically from mid-winter in the earliest individuals [@Baldassarre2014].
+Delayed dichromatism generally explains the discrepancies in estimates of apparent proportions of males counted in winter compared to proportions of males counted in spring or males identified from individuals culled in winter (Figures \@ref(fig:flock) & \@ref(fig:pmal)).
 
-Interestingly, delayed dichromatism enables estimating adult survival and recruitment/productivity by monitoring the seasonal evolution of apparent sex ratio and by making few and reasonable assumptions (see next section).
+Interestingly, delayed dichromatism allows the estimation of adult survival and recruitment/productivity by observing the seasonal evolution of the apparent sex ratio and making few and reasonable assumptions (see next section).
 
@@
 
-We monitored the demography of Ruddy duck populations in both Great Britain (GB) and France (FR) using exhaustive counts on the wintering grounds during the period between December 1 and January 31.
-In addition, apparent sex ratios in winter could be accessible from counts that distinguished female-like from male-like individuals in winter.
-Such counts were performed from 2006 to 2012 in Great Britain, which corresponded to a period of strong decline of the population (Figure \@ref(fig:count)).
-In France, a monitoring accounting for apparent sex was performed in 1999, 2001-2009, and 2012-2019.
-The French population observed a rapid growth during the first years of the monitoring and then stabilized from 2006 onwards as a result of a high culling pressure.
-In both countries, shooting was performed under the strict control of the authorities, meaning that the exact number of individuals culled was known.
-Therefore, despite the age ratio in birds shot was not ascertained before 2009 in France, it is possible to grossly qualify the culling pressure in three categories for both countries: "no culling" before 1999, "low culling" between 1999 and 2005, and "high culling" from 2006 (see Figure \@ref(fig:sup) in Supplementary Material).
+We monitored the demography of Ruddy duck populations in both Great Britain (GB) and France (FR) by conducting extensive counts on the wintering grounds between 1 December and 31 January.
+In addition, the apparent sex ratio could be determined using winter counts that distinguished between female-like and male-like individuals.
+Such counts were carried out in Great Britain from 2006 to 2012, which corresponded to a period of sharp population decline (Figure \@ref(fig:count)).
+In France, count surveys taking into account apparent sex were conducted in 1999, 2001-2009, and 2012-2019.
+The French population grew rapidly in the first years of monitoring and then stabilised from 2006 onwards as a result of high culling pressure.
+In both countries, culling was carried out under strict official control so that the exact number of individuals culled was known.
+Although the age ratio of birds culled before 2009 could not be determined with certainty in France, it was still possible to roughly categorise the culling pressure in both countries into three categories: "no culling" before 1999, "low culling" between 1999 and 2005 and "high culling" from 2006 onwards (see Figure \@ref(fig:sup) in Supplementary Material).
 
@@
 The proportion of immature males is the difference between the proportion of female-like individuals between winter and spring.
-The proportion of all immature individuals is twice this value if we assume an even sex ratio at birth [@Bellrose1961; @Blums1996] and comparable mortality rates between immature females and immature males [@Ramula2018; @Wood2021].
-However, counts in spring suffer from inextricable issues, such as a differential detectability between females and males.
-Fortunately, the proportion of immatures can also be estimated only from winter counts by making the assumption of a known and unchanging adult sex ratio over the monitoring period [@Devineau2010].
+The proportion of all immature individuals is twice as high if we assume an even sex ratio at birth [@Bellrose1961; @Blums1996] and comparable mortality rates between immature females and immature males [@Ramula2018; @Wood2021].
+However, spring counts suffer from intractable problems, such as the different detectability of females and males during this season.
+Fortunately, the proportion of immatures can also be estimated from winter counts alone, assuming a known and unchanged adult sex ratio along the monitoring period [@Devineau2010].
 
-The adult sex ratio is a prerequiste to estimate the proportion of immatures.
-No counts differentiating males from females were available during the breeding period to estimate adult sex ratios, but we determined the sex of each adult that was shot as part of the eradication programme both in Great Britain and in France.
-Nevertheless, the small number of culled individuals in France prevented from getting robust adult sex ratios.
-In Great Britain, the adult sex ratio estimated from the analysis of birds shot did not differ among months, so data from adults collected over the whole biological cycle were pooled to estimate adult sex ratios.
-The proportion of males did not statistically differ between the different years with more than 500 adults shot.
-This is consistent with previous findings showing that adult sex ratios in ducks are usually highly stable over the short-term [@Wood2021] but may vary over the long-term.
-Therefore, we pooled data from all adults shot as part as the control programme to estimate the adult sex ratio, see Equation \@ref(eq:pm).
+The adult sex ratio is therefore a prerequisite for estimating the proportion of immatures.
+This ratio could be derived from culling data as we determined the sex of each adult culled as part of the eradication programmes in both Great Britain and France.
+Nevertheless, the low number of individuals culled in France prevented from getting robust adult sex ratio.
+In Great Britain, adult sex ratios did not differ from month to month, so data from culled adults collected throughout the whole biological cycle were pooled together each year.
+The proportion of males also did not differ statistically between years in which more than 500 adults were culled.
+This is consistent with previous findings showing that the adult sex ratio in ducks is usually very stable in the short term [@Wood2021] but can fluctuate in the long term.
+Therefore, we pooled the data from all adults culled in the control programme to estimate the adult sex ratio, see Equation \@ref(eq:pm).
 
-Assuming the additive property of the binomial distribution, the proportion of immatures in a population can be deduced from the proportion of males in adults (alternatively the adult sex ratio) and the proportion of male-like individuals in the different count surveys in winter (assuming all these individuals are adult males), see Equation \@ref(eq:pic).
-By combining the population growth rate, which is derived from the year-to-year variation of an index of abundance of population size, to the proportion of immatures, one can derive adult survival and recruitment rates, see Equation \@ref(eq:vr) and Figure \@ref(fig:outline).
-A key concept is that if the absolute value of population size is known, numbers of adults and recruits can be estimated, see Equation \@ref(eq:VR).
+Assuming the additive property of the binomial distribution, the proportion of immatures in a population can be derived from the proportion of males among adults (alternatively the adult sex ratio) and the proportion of male-like individuals in the different counts in winter (assuming that all these individuals are adult males), see Equation \@ref(eq:pic).
+By combining the population growth rate derived from the interannual variation of an index of abundance with the proportion of immatures, adult survival and recruitment rates can be derived, see Equation \@ref(eq:vr) and Figure \@ref(fig:outline).
+If the absolute value of the population size is known, the number of adults and the number of recruits can be estimated, see Equation \@ref(eq:VR).
+We called this approach the "apparent sex ratio" method.
 
@@
 | *Observation (input)* |                                                                                                            |
-| $AM$                  | Total number of adult males shot as part of the eradication programme                                      |
-| $AF$                  | Total number of adult females shot as part of the eradication programme                                    |
+| $AM$                  | Total number of adult males culled as part of the eradication programme                                      |
+| $AF$                  | Total number of adult females culled as part of the eradication programme                                    |
 | $CML_{i, t}$          | Number of male-like individuals counted in population $i$ in year $t$                                      |
@@
 
-## Validating the method {#VAL}
+## Validating the "apparent sex ratio" method {#VAL}
 
-Nevertheless, we developed an indirect approach to estimate the maximum expected recruitment rate without exploitation and assuming that the observed density levels did not affect components of the recruitment process.
-Estimated values above that maximum recruitment rate would reflect failure in detection of adult males.
-On the other hand, null or negative recruitment rates would reflect either that immature males had acquired the male plumage before the survey, or violation of the assumption about the stability of the adult sex ratio.
+The relevance of the method was assessed by testing the likelihood of the estimates of both vital rates.
+For the adult survival rate, we checked whether the values were contained in the interval $[0;1]$, and compared our estimates with those in the literature [e.g.
+@Buxton2004; @Krementz1997; @Nichols1992, @Nichols1997].
+Validation of the recruitment rate was more difficult because if the lower limit is 0, the upper limit is a product of all maximum values of its individual components, i.e.
+nesting rate, nesting success, clutch-size, hatching success, pre-fledging survival, and fledging survival.
+These reproductive parameters may be known for some duck species, but the last one is never mentioned [e.g.
+@Baldassarre2014].
 
-Practically, maximum recruitment rate is estimated from the difference between maximum growth rate and maximum adult survival rate, see Equation \@ref(eq:GR).
-This relationship comes from a simple reasoning for a closed population: population size in year $t$ is equal to the number of adults that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e.
+To overcome this problem, we developed an indirect approach to estimate the maximum expected recruitment rate without exploitation and assuming that the level of population density does not affect the components of the recruitment process.
+Estimated values above this maximum recruitment rate would reflect a partial detection of adult males.
+On the other hand, null or negative recruitment rates would indicate that immature males acquired male plumage prior to the survey, or that the assumption about the stability of the adult sex ratio was violated.
+
+In practise, the maximum recruitment rate was estimated from the difference between the maximum growth rate and the maximum survival rate of adults, see Equation \@ref(eq:GR).
+This relationship resulted from a simple consideration for a closed population: the population size in year $t$ is equal to the number of adults that survived the whole year $t - 1$ plus the offspring produced in year $t - 1$ that survived to the reproductive period of year $t$, i.e.
 the recruitment in year $t$ [@Flint2015].
-For an open population, adult survival and recruitment rates are confounded to adult and recruit migrations, respectively, but this does not corrupt the equation.
-This relationship becomes more complex if a species with delayed maturity is considered, see @Robertson2008.
+In an open population, adult survival and recruitment rates are confounded with adult and recruit migrations, respectively, but this does not change the equation.
+This relationship becomes more complex when a species with delayed maturity is considered, see @Robertson2008.
 
@@
 ```
-Maximum adult survival was defined as a uniform distribution on the $[0.7;1]$ interval.
-The lower limit corresponded to the upper range of survival rates observed in waterfowl species of similar weight, and the upper limit was set to 1 because survival rates of long-lived waterfowl species are very high [@Nichols1997; @Krementz1997; @Buxton2004; @Nichols1992; @Koons2014].
+The maximum adult survival rate was defined as a uniform distribution on the interval $[0.7;1]$.
+The lower limit corresponded to the upper range of survival rates observed in waterfowl species of similar weight, and the upper limit was set to 1, as survival rates of long-lived waterfowl species are very high [@Nichols1997; @Krementz1997; @Buxton2004; @Nichols1992; @Koons2014].
 
-In both populations, the maximum growth rates occurred during their geographical expansion phase before shooting operations (Figure \@ref(fig:count)).
-To estimate robust maximum population growth rates for both populations, we smoothed annual growth rate values over a consistent time-period by using a linear regression on the logarithm scale, see Equation \@ref(eq:L).
-We discarded Great Britain data before 1972 as the estimates were probably noisy when the population was low (below 50 indiviudals) (Figure \@ref(fig:count)).
-After reaching the threshold of 1,000 individuals, the Great Britain (GB) population growth showed a strong inflexion despite culling had not begun (Figure \@ref(fig:count)).
-This observation suggests that a negative density-dependence process might occur beyond 1,000 individuals and led to consider only the first sequence of the time series to infer on the maximum growth rate in Great Britain, i.e.
+For both populations, the maximum growth rates occurred during their geographic expansion phase, which preceded the start of the eradication programmes (Figure \@ref(fig:count)).
+To estimate robust maximum population growth rates for both populations, we smoothed the annual growth rates over a consistent time period by using a linear regression on the logarithm scale, see Equation \@ref(eq:L).
+We discarded the data for Great Britain before 1972 as the estimates were likely noisy when the population was low (below 50 individuals) (Figure \@ref(fig:count)).
+After reaching the threshold of 1,000 individuals, the population growth in Great Britain (GB) showed a strong inflexion, although culling had not yet started (Figure \@ref(fig:count)).
+This observation suggests that a negative density-dependent process could occur beyond 1,000 individuals and led us to consider only the first sequence of the time series to infer the maximum growth rate in Great Britain, i.e.
 1972-1981.
-For the French (FR) population, the sequence without culling pressure spread over the period 1994-1999.
+For the French (FR) population, the sequence without culling pressure covered the period 1994-1999.
 
@@
 ```
-Estimating vital rates within an acceptable range, as defined above, would not prove that out modelling approach correctly tracked their year-to-year variability.
-In order to ascertain this aspect, we check the correlation between the proportions of immatures and both survival and recruitment rates derived from counts, to estimates derived from cullings, i.e.
-from individuals shot as part of the eradication programme.
-As we restricted the dataset to years with more than 100 individuals shot in winter for sake of precision, the analysis covered only five years of the time series of the GB population.
-Proportions of immatures were estimated from culled individuals by checking the presence of the bursa of Fabricius, which is present in immature individuals only [@Hochbaum1942]; see Equation \@ref(eq:pis).
-Adult survival and recruitment rates were then estimated by combining the proportion of immatures and Equation \@ref(eq:vr).
+Estimating vital rates within an acceptable range, as defined above, would not prove that our modelling approach correctly reflects the actual interannual variability.
+To determine this, we compared the proportions of immatures and the vital rates derived from counts with those derived from culling, i.e.
+from individuals culled as part of the eradication programme.
+As we restricted the dataset to years with more than 100 individuals culled in winter for sake of precision, the analysis only covered five years of the GB population time series.
+The proportions of immatures were estimated from the culled individuals by checking the presence of the bursa of Fabricius, which is only present in immature individuals [@Hochbaum1942]; see Equation \@ref(eq:pis).
+Adult survival and recruitment rates were then derived from the proportion of immatures and the Equation \@ref(eq:vr).
 
@@
 ```
-## Assessing the impact of culling strategies {#HARV}
+## Assessing the effects of culling strategies {#HARV}
 
-During period when culling pressure was high (i.e.
-from 2006 onwards), eradication strategies differed between Great Britain and France.
-In Great Britain, culling mostly occurred in winter (53.4% of adults were shot before the 30th May), whereas it was mostly done during the breeding period in France (81.6% of adults were shot after the 30th May).
-We assessed whether or not these strategies had a different impact on populations by comparing growth rates and the relative contributions of both recruitment and survival on them (by comparing the average values over the high culling pressure period to proxies of maximum vital rates estimated when both populations reached their maximum growth, see Section \@ref(VAL)).
-In France, a LIFE project was implemented by early 2019 to intensify the culling pressure, especially during winter and spring.
-We therefore excluded the FR time series from 2019 onward in order to provide a comparison of homogeneous culling strategies.
+During the period when culling pressure was high (i.e.
+from 2006 onwards), eradication strategies in Great Britain and France differed.
+In Great Britain, most individuals were culled in winter (53.4% of adults were culled before 30 May), whereas in France, most individuals were culled during the breeding season (81.6% of adults were culled after 30 May).
+We investigated whether these strategies had different effects on the populations by comparing the resulting growth rates and the relative contributions of both vital rates to them (by comparing the average values during the period of high culling pressure with proxies for the maximum vital rates estimated when both populations reached their maximum growth, see Section \@ref(VAL)).
+In France, a LIFE project was implemented from 2019 to intensify the culling pressure, especially in winter and spring.
+We therefore excluded the FR time series from 2019 onwards in order to compare homogeneous culling strategies.
 
@@
 
-We used the Bayesian framework to implement all sub-models presented in Section \@ref(MM).
-The Bayesian framework is both straightforward and efficient to propagate error through the parameters.
-We used uninformative priors on all parameters.
-As the maximum growth rate is a life history trait expected to be stable among populations of a same species, we used an uninformative hierarchical prior for this parameter.
-We generated three chains of length 500,000, with a thinning of 10 to avoid autocorrelation in the samples, discarding the first 2,000 samples as burn-in.
-We assessed chain convergence using the Gelman and Rubin convergence diagnostic (R\<1.1, @Gelman1992).
-Models were fitted using NIMBLE [@Valpine2017] run from R [@RCT2022].
-The values **X[Y; Z]** reported in Section \@ref(RES) are the medians and the associated boundaries of the 95% confidence interval of posterior distributions.
-We preferred the median to the mean because of its robustness to skewed distribution.
+We used the Bayesian framework to implement all submodels presented in Section \@ref(MM).
+The Bayesian framework is both straightforward and efficient in propagating error through the parameters.
+We used uninformative priors for all parameters.
+Since the maximum growth rate is a life history trait expected to be stable between populations of the same species, we used an uninformative hierarchical prior for this parameter.
+We generated three chains of length 500,000, with a thinning of 10 to avoid autocorrelation in the samples, and discarded the first 2,000 samples as burn-in.
+We checked the chain convergence using the Gelman and Rubin convergence diagnostic (R\<1.1, @Gelman1992).
+The models were fitted using NIMBLE [@Valpine2017] run from R [@RCT2022].
+The values **X[Y; Z]** reported in Section \@ref(RES) are the medians and the corresponding limits of the 95% confidence interval of the posterior distributions.
+We preferred the median to the mean because it is more robust to skewed distributions.
 
@@
 
-The method successfully provided estimates and associated uncertainty for the proportion of immatures in both wintering populations.
-Similarly, adult survival and recruitment rates, i.e.
-the two components accounting for population growth rate, were also properly estimated (Figure \@ref(fig:vital)).
+The "apparent sex ratio" method successfully provided estimates and associated uncertainties for the proportion of immatures in both wintering populations.
+Adult survival and recruitment rates, i.e.
+the two components that determine the population growth rate, were also correctly estimated (Figure \@ref(fig:vital)).
 
@@
 
-The proportion of males in adults, which is a prerequisite to infer the proportion of immatures, was estimated at 0.60 [0.59; 0.61].
-The proportion of immatures ranged from 0.16 [0.07; 0.24] to 0.54 [0.44; 0.62], depending on population and year.
-The lowest values were similar for the two populations, but the range of the proportion of immatures in the GB population was far more limited (upper value GB: 0.30 [0.14; 0.45] & FR: 0.54 [0.44; 0.62]).
-The proportion of immatures was stable over time in the GB population, while a slight but significant decrease was observed for the FR population.
+The proportion of males among adults, which is a prerequisite for inferring the proportion of immatures, was estimated at 0.60 [0.59; 0.61].
+The proportion of immatures ranged between 0.16 [0.07; 0.24] and 0.54 [0.44; 0.62], depending on the population and year.
+The lowest values were similar in both populations, but the range of the proportion of immatures in the GB population was much smaller (upper value GB: 0.30 [0.14; 0.45] & FR: 0.54 [0.44; 0.62]).
+The proportion of immatures in the GB population was stable over time, while a slight but significant decrease was observed in the FR population.
 
-Adult survival rate ranged from 0.33 [0.29; 0.37] to 1.01 [0.89; 1.14].
-None of the estimates was significantly outside the range of a survival rate defined without immigration [0; 1].
-No trend in adult survival rate could be evidenced in any of the two populations, but patterns differed: adult survival in GB was both lower and displayed a weaker inter-annual variability than in FR.
+Adult survival rates ranged between 0.33 [0.29; 0.37] and 1.01 [0.89; 1.14].
+None of the estimates were significantly outside the range of a survival rate defined without immigration [0; 1].
+No trend in adult survival rate was observed in either population, but the patterns were different: adult survival in GB was lower and showed less interannual variability than in FR.
 
-Recruitment rate ranged from 0.09 [0.04; 0.13] to 0.75 [0.63; 0.87].
-All estimates were above 0, that is not incompatible with reality.
-Moreover, the maximum recruitment rate was 0.68 [0.36; 0.78] (see Section \@ref(MAX)), and no estimates were significantly outside the range of recruitment rates defined without immigration [0; 0.78].
-No trend was observed for the GB population, but recruitment rate decreased for the FR population even if it was more noisy than for the proportion of immatures.
-Like for survival, the GB population had lower values of recruitment rate with weaker variability than the FR one.
+Recruitment rates ranged between 0.09 [0.04; 0.13] and 0.75 [0.63; 0.87].
+All estimates were above 0, which is not inconsistent with reality.
+Furthermore, the maximum recruitment rate was 0.68 [0.36; 0.78] (see Section \@ref(MAX)), and no estimate was significantly outside the range of recruitment rates defined without immigration [0; 0.78].
+No trend was observed for the GB population, but the recruitment rate decreased for the FR population, although it was more noisy than the proportion of immatures.
+As with the survival rate, the GB population also showed lower recruitment rate values with less variability than the FR population.
 
-Lower values and weaker variability of both adult survival and recruitment rates estimated in GB than in FR likely explained the differing trajectories of the two Ruddy duck populations, a sharp and constant decline for the former vs slow and variable decline for the latter (Figure \@ref(fig:count)).In the FR population, the range in recruitment rates (min/max difference: 0.66) was higher than the range of survival rates (min/max difference: 0.58).
-Conversely, the range of survival rates (min/max difference: 0.46) in the GB population was far higher than the range of recruitment rate (min/max difference: 0.08).
-This suggests that the variability of population growth rate was mostly driven by recruitment in FR whereas it was mainly driven by changes in adult survival in GB.
+Lower values and lower variability of both adult survival and recruitment rates estimated in the GB population than in the FR population probably explain the different trends of the two Ruddy duck populations: a sharp and constant decline in the GB population versus a slow and variable decline in the FR population (Figure \@ref(fig:count)).
+In the FR population, the range of recruitment rates (min/max difference: 0.66) was larger than the range of survival rates (min/max difference: 0.58).
+Conversely, the range of survival rates (min/max difference: 0.46) in the GB population was much larger than the range of recruitment rates (min/max difference: 0.08).
+This suggests that the variability in population growth rate in FR was mainly determined by recruitment, whereas in GB it was mainly determined by changes in adult survival.
 
-## Examining the reliability of the estimates
+## Testing the reliability of the estimates
 
-In Great Britain, proportions of immatures derived from the count-based method were positively correlated to those observed from culled individuals (Figure \@ref(fig:cor)).
-Therefore, this result supports the hypothesis that the proportion of female-like individuals in the wintering populations is efficient for assessing age structure and hence reproductive output.
-The correlation was even stronger when considering vital rates (Figure \@ref(fig:cor)), but this was expected because the two methods estimating these vital rates shared a component, the growth rate of the population (see Equation \@ref(eq:vr)).
-Interestingly, a 1:1 correlation is never reached for any parameters.
-The proportion of immatures derived from counts is always lower than those observed in culled individuals.
-The discrepancy between these two estimates is due to the fact that immatures are probably more vulnerable to culling than adults and might then be over-represented in culling samples.
-As a consequence, the correlation coefficients for both vital rates also differed.
-However, the differences between the two approaches do not question the ability of the count method to track the interannual variability of the vital rates.
+In Great Britain, the proportions of immatures derived from the counts were positively correlated with those derived from the culled individuals (Figure \@ref(fig:cor)).
+This result therefore supports the hypothesis that the proportion of female-like individuals in the overwintering populations is efficient for the assessment of age structure and thus productivity.
+The correlation was even stronger when looking at vital rates (Figure \@ref(fig:cor)), but this was to be expected because the two methods for estimating these vital rates had a common component, namely the population growth rate (see Equation \@ref(eq:vr)).
+Interestingly, a 1:1 correlation was not achieved for any of the parameters.
+The proportion of immatures derived from the counts is always lower than the proportion derived from the culled individuals.
+The discrepancy between these two estimates is due to the fact that immatures are probably more vulnerable to culling than adults and may therefore be over-represented in the culling samples.
+As a result, the correlation coefficients for both vital rates also differed.
+However, the differences between the two approaches do not call into question the ability of the "apparent sex ratio" method to capture interannual variability of the vital rates.
 
@@
 
-When culling pressure was nil, maximum growth rates were very similar for the two populations (Figure \@ref(fig:mgr)), respectively 1.45 [1.36; 1.55] and 1.52 [1.25; 1.88] for GB and FR.
-These values corresponded to a population increase of 45% and 52% per year, respectively.
-Assuming a maximum adult survival rate of 0.85 [0.7; 1] identical for both populations, we deduced maximum recruitment rates of 0.60 [0.42; 0.78] and 0.68 [0.36; 1.06] respectively for GB and FR.
+When there was no culling pressure, maximum growth rates were very similar for the two populations (Figure \@ref(fig:mgr)), namely 1.45 [1.36; 1.55] and 1.52 [1.25; 1.88] for GB and FR, respectively.
+These values corresponded to an increase in the population of 45% and 52% per year, respectively.
+Assuming a maximum adult survival rate of 0.85 [0.7; 1] for both populations, we derived maximum recruitment rates of 0.60 [0.42; 0.78] and 0.68 [0.36; 1.06] for GB and FR, respectively.
 
-Despite low sample sizes of sexed and aged culled individuals, proxies of the culling pressure exhibited significant changes over time (Figure \@ref(fig:sup)).
-We thus ranked the culling pressure according to this intensity.
+Despite the low number of sex- and age-determined culled individuals, the proxies of culling pressure showed significant changes over time (Figure \@ref(fig:sup)).
+Therefore, we categorised culling pressure according to this intensity.
 High culling pressure occurred from 2006 onwards in both populations (although the signal was noisy).
-However, despite similar levels of culling pressure, the culling strategies differed significantly between the two populations, see Section \@ref(HARV).
-Under high culling pressure, the growth rate dropped to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for GB and FR, respectively, which corresponded to a decrease of 44% per year and 3% per year.
-This result showed that the GB population significantly decreased during the high culling pressure, whereas the FR population was just stabilized.
+However, despite the similar level of culling pressure, the culling strategies of the two populations differed considerably, see Section \@ref(HARV).
+Under high culling pressure, the growth rate decreased to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for GB and FR, corresponding to a decrease of 44% per year and 3% per year, respectively.
+This result shows that the GB population declined significantly during the high culling pressure, while the FR population was just stabilised.
 
-Adult survival rate under high culling pressure was 0.44 [0.42; 0.46] and 0.62 [0.59; 0.64] for GB and FR, respectively, and recruitment rates decreased to 0.13 [0.11; 0.15] and 0.32 [0.29; 0.34], respectively.
-The resulting drop in growth rate of the GB population corresponded to a similar average drop in adult survival rate (-0.41) and in recruitment rate (-0.47), whereas the stabilisation of the FR population was more caused by a decrease of recruitment rate (-0.36) than by a drop in adult survival (-0.23).
-Overall, culling strategy adopted in GB was more effective in affecting both adult survival and recruitment rates than that adopted in FR.
+Adult survival rates under high culling pressure reached 0.44 [0.42; 0.46] and 0.62 [0.59; 0.64] for GB and FR, respectively, and the recruitment rates decreased to 0.13 [0.11; 0.15] and 0.32 [0.29; 0.34], respectively.
+The resulting decrease in the growth rate of the GB population corresponded to a similar average decrease in adult survival rate (-0.41) and recruitment rate (-0.47), whereas the stabilisation of the FR population was due to a decrease in recruitment rate (-0.36) rather than a decrease in adult survival (-0.23).
+Overall, the culling strategy used in GB was more effective than that used in FR as it affected both adult survival and recruitment rates.
 
@@
 
-Disentangling the contribution of the vital rates to population growth rate is a key step towards a better understanding of the response of a population to management actions [@Williams2002].
-We exploited delayed sexual dichromatism to develop a new approach based on counts to breakdown the apparent growth rate of a population into two of its main structural components: adult survival and recruitment rates.
-The development of the "apparent sex ratio" method holds on a unique dataset of two comparable populations of the same species, the Ruddy duck.
-Both populations were monitored in a similar way from their introduction to a period of intense culling pressure, but the culling strategies differed sharply.
-The very similar demographic trajectories and properties of both populations in similar ecological setup enabled evaluating the response to both culling pressure and culling strategies.
+Disentangling the contribution of vital rates to population growth rate is a key step towards better understanding a population's response to management actions [@Williams2002].
+We exploited delayed sexual dichromatism to develop a new approach based on counts to disentangle the growth rate of a population into its two major structural components: adult survival and recruitment rates.
+The development of the "apparent sex ratio" method is based on a unique data set of two comparable populations of the same species, the Ruddy duck.
+Both populations were monitored in a similar way from their introduction to a period of intense culling pressure, but the culling strategies differed markedly.
+The very similar demographic trajectories and properties of both populations under similar ecological conditions made it possible to assess the response to both culling pressure and culling strategies.
 
@@
 
-For the two populations, the vital rate estimates were within the same range, which demonstrates a certain consistency in the results obtained from the method.
-The greater variability observed for the FR population was expected because the time series covered a larger spectrum of culling pressure than the GB one.
-We did not detect temporal correlation in the demographic parameters between the two populations, suggesting that regional (e.g., weather) rather than large-scale factors (e.g., climate) prevailed.
-This is consistent with previous findings showing for example that in waterfowl, both breeding success and juvenile survival are conditional to the onset of laying, which is tightly related to local weather (e.g.
-spring temperature, cumulative rainfall, water levels) [@Folliot2017; @Blums2004; @Dzus1998], and to local changes in predation pressure [@Jaatinen2022].
+For both populations, vital rate estimates were within the same range, showing some consistency in the results obtained using the "apparent sex ratio" method.
+The greater variability observed for the FR population was to be expected as the time series covered a wider range of culling pressure than that of the GB population.
+We found no temporal correlation of demographic parameters between the two populations, suggesting that regional (e.g.
+weather) rather than large-scale factors (e.g.
+climate) were predominant.
+This is consistent with previous findings showing, for example, that in waterfowl both breeding success and juvenile survival depend on the onset of the laying period, which is closely linked to local weather (e.g.
+spring temperature, cumulative rainfall, water levels) [@Folliot2017; @Blums2004; @Dzus1998] and local changes in predation pressure [@Jaatinen2022].
 
-Count surveys used to implement our model occurred in mid-winter, that is when some immature males might initiate their moult.
-This could partially violate the assumption that male-like individuals were adult males only, and hence potentially bias our estimates.
-Nevertheless, as survival estimates are within the range of those reported for similar species [@Nichols1997; @Krementz1997; @Buxton2004; @Buderman2023], the assumption seems reliable.
-Therefore, the time window of the count surveys must be chosen with care to fit the time preceding the initiation of moult of immature males.
+The counts used to implement our model took place in mid-winter, at a time when some immature males may be initiating their moult.
+This could partially violate the assumption that the male-like individuals were exclusively adult males, and thus potentially bias our estimates.
+Nevertheless, as the survival estimates are in the range of those reported for similar species [@Nichols1997; @Krementz1997; @Buxton2004; @Buderman2023], this assumption appears to be reliable.
+Therefore, the time window for the count surveys must be carefully chosen to take into account the time before the immature males start moulting.
 
-In some years, the estimated values of adult survival were outside the range of expected values even if they never significantly exceed 1.
+In some years, the estimated adult survival values were outside the range of expected values, although never significantly above 1.
 One possible explanation for these discrepancies is a violation of the "closed population" assumption.
-Indeed, despite ringing data suggest that Ruddy ducks do not undertake long-distance seasonal movements in GB (Henderson, pers.
-com.), one cannot totally rule out that immigration occurred.
-The out of range outliers of both recruitment and survival rates that were observed in 2002 and 2012 in FR suggest then significant immigration events during these years.
+Although the ringing data suggest that Ruddy ducks do not undertake long-distance seasonal migrations in GB (Henderson, pers.
+comm.), the possibility that immigration occurred cannot be completely ruled out.
+The outliers in  survival and recruitment rates observed in FR in 2002 and 2012 suggest that there was significant immigration in those years.
 
 The theoretical maximum population growth rate is an inherent trait related to the demographic characteristics of a species [@Dmitriew2010; @Niel2005].
-This trait is expected to be similar in all Ruddy duck populations anywhere in the world.
-In our study, we observed maximum growth rates that were very close for the two populations, which was expected because both populations experienced similar ecological conditions with an absence of harvesting and similar breeding conditions.
-Moreover, before the launch of the eradication programmes, the two populations also expressed a population growth much closer to the maximum expected values than any other native ducks species at the same period in the same areas.
-In the absence of culling, both populations expressed survival rates in the upper range of those observed in other diving ducks species (e.g.
-\~ 0.8 for Common pochard [@Folliot2020; @Nichols1997]).
-Ruddy duck tend to breed latter than both Common pochard and Tufted duck, yet nest predation has been found to decrease over the breeding season in a French fishpond complex [@Bourdais2015].
-Therefore, both enhanced survival and nesting success might have accounted for high population growth rates of Ruddy duck populations in the absence of culling.
-Besides providing estimates of adult survival and recruitment rates that were consistent with observed increasing population sizes, our method also provided estimates consistent with both the decreasing phases occurring after the implementation of eradication programmes and the observed differences in populations growth rates under differing culling strategies.
-As expected, higher values of survival and recruitment rates were observed when the culling strategy focused on post-breeding season rather than the pre-breeding one.
+This trait is expected to be similar in all Ruddy duck populations around the world.
+In our study, we observed that the maximum growth rates of the two populations were very close to each other, which was to be expected as both populations were exposed to similar ecological conditions, i.e.
+no harvest and similar breeding conditions.
+In addition, before the eradication programmes started, the two populations had population growth rates much closer to the expected maximum values than any other native duck species in the same areas during the same period.
+Prior to the eradication programmes, the survival rates of both populations were in the upper range of those observed for other diving duck species (e.g.
+\~ 0.8 for the Common pochard [@Folliot2020; @Nichols1997]).
+Predation on diving duck nests was found to decrease during the breeding season in a French fishpond complex [@Bourdais2015].
+As Ruddy ducks generally breed later than Common pochards and Tufted ducks, both a higher survival rate and greater nesting success could be responsible for the high growth rates of Ruddy duck populations in the absence of culling.
+The "apparent sex ratio" method not only provided estimates of adult survival and recruitment rates that were consistent with observed increasing population sizes, but also provided estimates that were consistent with both declining phases following the implementation of eradication programmes and observed differences in population growth rates under different culling strategies.
+As expected, higher values for survival and recruitment rates were observed when the culling strategy focussed on the post-breeding season rather than the pre-breeding season.
 
-Assuming that culling data provide a good picture of the year-to-year variability of age structure [@Fox2014], the strong correlation between estimates derived from counts and those derived from culled individuals demonstrates the ability of the "apparent sex ratio" method introduced here to properly track inter-annual fluctuations in demographic parameters such as survival and recruitment rates.
+Assuming that culling data provide a good picture of interannual variability in age structure [@Fox2014], the strong correlation between estimates derived from counts and estimates derived from culled individuals demonstrates the ability of the "apparent sex ratio" method presented here to properly capture interannual variation in demographic parameters such as survival and recruitment rates.
 The fact that the observed correlation was strong despite being based on a short time series strengthens its robustness.
-Furthermore, the strength of the relationship between parameters obtained through the two approaches also supports that important assumptions of our model were met, such as constant adult sex ratio.
-The temporal autocorrelation of the adult sex ratio might be explained by the fact that it integrates many age-cohorts, which makes it structurally strong.
-Significant variations of the adult sex ratio can occur over the long term, but our results suggests that it is not necessary to monitor and update it on a yearly basis.
-Proportions of males in adults estimated in the present paper (i.e.
-0.60) are within the range of those observed in native Ruddy duck populations of North America (0.62 in @Bellrose1980) and in other duck species [@Wood2021].
-The "apparent sex ratio" method always provided recruitment rates significantly lower than the method based on culling data by a factor close to two, a result that is compatible with a higher vulnerability of the immatures to shooting [@Fox2014].
-For instance, in North America, immatures have been found to be 1.3 to 2.6 times more vulnerable to hunting than adults in waterfowl [@Bellrose1980].
+In addition, the strength of the relationship between the parameters obtained with the two approaches also indicates that important assumptions of our model were met, such as the constant adult sex ratio.
+The temporal autocorrelation of the adult sex ratio could be explained by the fact that it includes many age cohorts, which makes it structurally strong.
+In the long term, there may be substantial fluctuations in the adult sex ratio, but our results suggest that it is not necessary to monitor and update it annually.
+The proportion of males among adults estimated in the present study (i.e., 0.60) is within the range of values observed in native Ruddy duck populations of North America (0.62 in @Bellrose1980) and in other duck species [@Wood2021].
+The "apparent sex ratio" method consistently yielded recruitment rates that were almost a factor of two lower than the method based on culling data, a result consistent with higher vulnerability of immatures to shooting [@Fox2014].
+In North America, for example, immature waterfowl were found to be 1.3 to 2.6 times more vulnerable to hunting than adults [@Bellrose1980].
 
-Because estimates of survival were not available in the literature (see @Buderman2023), we could not properly assess the accuracy of the "apparent sex ratio" method for estimating this parameter.
-Nevertheless, circumstantial evidence support that the method is not heavily biased as detailed above.
-Thus the "count-changing sex ratio" method would be much more efficient for monitoring key demographic parameters than alternative methods such as those based on culling/hunting bag data that proved unreliable [@Fox2014], even when they focused on the end the hunting season to limit possible biases [@Fox2016].
+As no estimates of adult survival were available in the literature [see @Buderman2023], we were unable to properly assess the accuracy of the "apparent sex ratio" method for estimating this parameter.
+Nevertheless, there is evidence that the method is not highly biassed, as described above.
+Thus, the "apparent sex ratio" method would be much more efficient for monitoring key demographic parameters than alternative methods such as those based on culling/hunting bag data, which have been shown to be unreliable [@Fox2014], even when focussing on the end of the hunting season to limit potential bias [@Fox2016].
 
-## Assessing eradication strategies
+## Evaluating eradication strategies
 
-Both populations responded to culling pressure by a significant decrease in their growth rate, meaning that increasing mortality due to culling was neither compensated by increasing natural survival nor by improved breeding success.
-The culling effort was therefore efficient to affect the demographic trajectory of Ruddy duck in both contexts.
-However, the decrease in population growth was much stronger in GB, where the population displayed a sharp decrease, than in FR where the increase was stopped and population size remained more or less stable.
-The greater decline in population growth observed in GB compared to FR was due to greater decline in both adult survival and recruitment rates.
-Overall, in both countries, adult survival and recruitment rates seemed significantly affected by culling.
-However, the culling pressure occurring before the breeding season, which affect only future breeders (pre-breeding culling strategy applied in GB), seemed much more efficient in limiting both survival and recruitment rates than the culling pressure occurring during nesting and young raising, which targets breeders and newly born individuals (breeding culling strategy applied in FR).
+Both populations responded to culling pressure with a significant decline in their growth rate, meaning that the increased mortality from culling was not compensated for by an increase in natural survival or improved breeding success.
+Culling was therefore effective in affecting the demographic trajectory of Ruddy duck in both cases.
+However, the decline in population growth was much greater in GB, where the population declined steadily and sharply, than in FR, where the increase was halted and the population size remained more or less stable.
+The greater decline in population growth observed in GB compared to FR, was due to a greater decline in adult survival and recruitment rates.
+Overall, adult survival and recruitment rates appeared to be significantly affected by culling in both countries.
+However, culling pressure targeting future breeders (pre-breeding culling strategy applied in GB), appeared to be much more effective in reducing both survival and recruitment rates than culling pressure targeting breeders and newly born individuals during the breeding and rearing season (breeding culling strategy applied in FR).
 
-As expected, the pre-breeding culling strategy had the same effect on adult survival and recruitment rates because a bird killed before the breeding period reduces the breeding population (lower adult survival rate) and prevents this bird from reproducing (lower recruitment rate).
-When the breeding culling strategy was applied, the recruitment rate was more affected than adult survival.
-This was probably due to immatures being much more vulnerable to shooting than adults [@Bellrose1980; @Fox2014].
-The compensation hypothesis suggests that the higher harvest mortality of immatures can be compensated by increasing natural survival [@Cooch2014] but it is not supported by the observations.
-Overall, there was no evidence for any compensation mechanism in either context.
-This might result from the fact that both populations were far of reaching carrying capacity meaning that competition for resources was relaxed [@Peron2013].
+As expected, the pre-breeding culling strategy had the same effect on adult survival and recruitment rates, as a bird killed before the breeding season reduces the breeding population (lower adult survival rate) and prevents the bird from reproducing (lower recruitment rate).
+When using the breeding culling strategy, the recruitment rate was more affected than the adult survival rate.
+This was probably due to the fact that immatures are much more susceptible to being shot than adults [@Bellrose1980; @Fox2014].
+The compensation hypothesis states that the increased mortality of immatures due to harvesting may be compensated for by a higher natural survival [@Cooch2014], but this hypothesis is not supported by the observations.
+Overall, there was no evidence of a compensatory mechanism in either context.
+This could be due to the fact that both populations were far from reaching carrying capacity, meaning that competition for resources was relaxed [@Peron2013].
 
-For more or less comparable culling pressure, the pre-breeding culling strategy proved much more effective.
-This does not mean however, that the breeding culling strategy was not so.
-On the contrary, it proved rather efficient for stopping both the population increase and range expansion in a first time, and in second time to trigger a significant population decline by slightly increasing culling pressure.
-Thus when winter culling is not possible, which was the case in FR, culling during the breeding season is effective for stopping or reversing population increases in such species.
-The breeding culling strategy is however affected by the highly variable detectability of immatures, as these individuals colonise new ponds.
-This results in fluctuations of the recruitment rate and then threatens the long term efficiency of such strategy.
-The winter culling strategy does not face such problem because of the habit of Ruddy ducks to gather in a limited number of sites during winter [@Johnsgard1996].
+With more or less comparable culling pressure, the pre-breeding culling strategy proved to be much more effective.
+However, this does not mean that the breeding culling strategy was not also effective.
+On the contrary, it proved to be quite efficient in stopping both the population increase and the expansion of the distribution area, and in triggering a significant population decline in a second attempt by slightly increasing the culling pressure.
+So if winter culling is not possible, which was the case in FR, culling during the breeding season is effective in stopping or reversing the increase in populations of this species.
+However, the breeding culling strategy is compromised by the highly variable detectability of immatures, as these individuals colonise new ponds.
+This leads to fluctuations in the recruitment rate and thus jeopardises the long term efficiency of this strategy.
+This problem does not exist with the pre-breeding strategy, as the Ruddy ducks congregate at a limited number of sites in winter [@Johnsgard1996].
 
-The response of the two Ruddy duck populations to culling pressure demonstrates the necessity to account for the culling season to produce proper predictions.
-Predicting the impact of a harvest pressure on a waterfowl population is then not only a question about how big is the harvesting effort, but also when harvesting occurs [@Kokko1998].
-Counter-intuitively, our results suggest that harvesting waterfowl during breeding has far less impact on population growth rate than harvesting during winter.
-As a result, a target for effective management of harvested waterfowl should be earlier onsets and endings of hunting seasons.
+The response of the two Ruddy duck populations to culling pressure shows that it is necessary to take the culling period into account in order to make correct predictions of population trajectory.
+Predicting the effects of harvest pressure on a waterfowl population is then not only a question of the level of harvest, but also the timing of the harvest [@Kokko1998].
+Unexpectedly, our results suggest that harvesting waterfowl during the breeding season has far less impact on population growth rate than harvesting in winter.
+Consequently, a target for effective management of harvested waterfowl could be an earlier start and end of the hunting season.
 
-## Waterfowl management implications
+## Implications for waterfowl management
 
-One major implication of our study concerns the timing of harvesting for limiting its impact on populations growth rate.
-Counter-intuitively, our results show that the persistence of waterfowl populations should benefit from earlier openings and closings of the hunting seasons (see above).
+An important implication of our study concerns the timing of harvest to limit its impact on populations growth rates.
+Unexpectedly, our results show that the persistence of waterfowl populations should benefit from earlier openings and closings of the hunting season (see above).
 
-Tracking variations in abundance is a common tool to determine the conservation status of a population [e.g @Folliot2022].
-But tracking abundance alone does not provide enough information for assessing the underlying mechanisms behind the changes of population growth [@Williams2002].
-In order to do that, one must “dig deeper” [@Austin2000], for example by monitoring individuals to assess parameters like survival and recruitment rates [e.g.
+Tracking fluctuations of population abundance is a common tool for determining the conservation status of a population [e.g.
+@Folliot2022].
+But tracking abundance alone does not provide enough information to assess the underlying mechanisms behind changes in population growth [@Williams2002].
+This requires “digging deeper” [@Austin2000], for example, by monitoring individuals to assess parameters such as survival and recruitment rates [e.g.
 @Arnold2018].
-However, monitoring individuals is time consuming [e.g.
-@Souchay2016] and not always possible in endangered species.
-The method proposed here allows circumventing these drawbacks in dichromatic species with delayed maturity of males.
-Among these species, one can mention the White headed duck in Southern Spain, which is highly endangered and cannot be safely monitored through capture-mark-recapture.
-Thus by improving census strategies and teasing apart the effects of variations of adult survival and productivity on population growth through the approach presented here, one could gather much more invaluable information about the pertinence of management strategies.
+However, monitoring individuals is time-consuming [e.g.
+@Souchay2016] and not always possible for endangered species.
+The "apparent sex ratio" method makes it possible to circumvent these disadvantages in dichromatic species with delayed sexual maturity of the males.
+These species include the White headed duck in southern Spain, which is highly endangered and cannot be disturbed through capture-mark-recapture.
+Thus, by improving census strategies and disaggregating the effects of variations in adult survival and productivity on population growth using the approach presented here, much more invaluable information on the relevance of management strategies could be obtained.
 
@@
 
-Finally, we want to stress the fact that the main aim of our approach was to decompose population growth into its two main components and not to provide unbiased estimates of adult survival or recruitment rates.
-However, if needed, censuses could be designed in such a way that would make this possible, at least in theory, and in species presenting an observable delayed dichromatism.
-This property concerns most stiff-tailed duck species [@Johnsgard1996], as well as numerous other duck species, among which Tufted duck, *Aythya fuligula*, Black scotter, *Melanitta nigra americana*, Common Goldeneye, *Bucephala clangula americana* [@Johnsgard1978; @Bellrose1980].
-Modifying standard monitoring protocols to distinguish male-like from female-like individuals is almost costless but worth the bet because it would provide substantial increases in the efficiency of conservation/management actions [@Nichols2006].
+Finally, we would like to emphasise that the main aim of our approach was to decompose population growth into its two main components, and not to provide unbiased estimates of adult survival or recruitment rates.
+However, if necessary, count surveys could be designed to do so, at least in theory, and for species that exhibit observable delayed dichromatism.
+This trait affects most species of stiff-tailed duck [@Johnsgard1996], as well as numerous other duck species, including Tufted duck, *Aythya fuligula*, Black scotter, *Melanitta nigra americana*, Common Goldeneye, *Bucephala clangula americana* [@Johnsgard1978; @Bellrose1980].
+Modifying standard monitoring protocols to distinguish between male-like and female-like individuals is almost costless, but worth the effort as it would greatly increase the efficiency of conservation/management actions [@Nichols2006].
 
@@
 
-This work was partly funded by the LIFE Oxyura project (LIFE17 NAT/FR/000942) through the LIFE program.
-We warmly thank Jay McGowan for allowing us to publish his Ruddy duck photograph.
-This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the *Office Français de la Biodiversité*, and Jean-Marc Gillier from the *Société Nationale pour la Protection de la Nature*.
-The authors acknowledge all the contributors of the data collection; in Great Britain, Wildfowl & Wetlands Trust staff past and present, particularly Baz Hughes, Peter Cranswick and Colette Hall, and all the project workers at the Animal and Plant Health Agency and its predecessors; in France, Office Français de la Biodiversité staff past and present, especially Vincent Fontaine, Denis Lacourpaille, Agathe Pirog, Hugo Pichard, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface, and the whole team of the Société Nationale pour la Protection de la Nature of the Grand Lieu lake.
+This work was partly funded by the LIFE Oxyura project (LIFE17 NAT/FR/000942) under the LIFE programme.
+We warmly thank Jay McGowan for his permission to publish his photograph of Ruddy duck.
+This work was carried out at the suggestion of Jean-François Maillard and Jean-Baptiste Mouronval of the *Office Français de la Biodiversité*, and Jean-Marc Gillier of the *Société Nationale pour la Protection de la Nature*.
+The authors acknowledge all the contributors of the data collection: in Great Britain, Wildfowl & Wetlands Trust staff past and present, in particular Baz Hughes, Peter Cranswick and Colette Hall, and all the project workers at the Animal and Plant Health Agency and its predecessors & in France, Office Français de la Biodiversité staff past and present, especially Vincent Fontaine, Denis Lacourpaille, Agathe Pirog, Hugo Pichard, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface, as well as the whole team of the Société Nationale pour la Protection de la Nature of the Grand Lieu lake.
 
@@
 
-(ref:count) Temporal changes in the number of Ruddy ducks (log scale) in Great Britain (GB) and France (FR) from 1960 to 2019, with indications of periods when apparent sex ratios were also recorded
+(ref:count) Temporal changes in the number of Ruddy ducks (logarithmic scale) in Great Britain (GB) and France (FR) from 1960 to 2019, with references to periods when apparent sex ratios were reported
 
-(ref:flock) A Ruddy duck flock observed in winter including ten female-like individuals with a whitish striped cheek (some of which are immature males), four male-like individuals with a white cheek and a black cap (all adult males), three unidentified individuals (1^st^, 4^th^, and 10^th^ from left) © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
+(ref:flock) A flock of Ruddy ducks observed in winter, including ten female-like individuals with whitish striped cheek (some of which are immature males), four male-like individuals with white cheek and black cap (all adult males), three unidentified individuals (1^st^, 4^th^, and 10^th^ from left) © Jay McGowan - 3 February 2013 - Tompkins, New York, United States
 
-(ref:pmal) Proportions of males estimated through the internal examination of individuals shot as part of the eradication programme over one year in a population (Male % in cullings) and proportions of individuals displaying a male-like plumage in a population counted in winter (Male % in counts); discrepancies between the two estimates are due to delayed dichromatism (immature males looking like females before moulting); we exploited these discrepancies for estimating survival and productivity/recruitment
+(ref:pmal) Proportions of males estimated by internal examination of individuals culled in a same year under the eradication programmes (Male % in cullings), and proportions of individuals showing male-like plumage counted in winter (Male % in counts); discrepancies between the two estimates are due to delayed dichromatism (immature males looking like females before moulting); we used these discrepancies to estimate survival and productivity/recruitment
 
-(ref:outline) Outline of the method to disentangle the population growth into survival (adult survival rate) and productivity (recruitment rate)
+(ref:outline) Overview of the "apparent sex ratio" method used to disentangle the population growth into survival (adult survival rate) and productivity (recruitment rate)
 
-(ref:vital) Temporal changes in the proportions of immatures and the vital rates following the culling pressure on Ruddy duck populations of Great Britain (GB) and France (FR); vertical bars represents the 95% confidence intervals.
+(ref:vital) Temporal changes in the proportions of immatures and vital rates following culling pressure in the populations of Great Britain (GB) and France (FR); the vertical bars represent the 95% confidence intervals
 
-(ref:cor) Relationship between parameter estimates derived from counts (count-based method) and those obtained from culled individuals (cull-based method); only five years were available for the GB population; bars represent 95% confidence intervals and the red arrow indicates the direction and average differences between the two methods
+(ref:cor) Relationship between parameter estimates obtained from counts ("apparent sex ratio" method) and those obtained from culled individuals (culling-based method); only five years were available for the GB population; the bars represent 95% confidence intervals and the red arrow indicates the direction and average differences between the two methods
 
-(ref:mgr) Average effects of culling pressure and different culling strategies (GB vs FR) on population growth rates (derived from counts only), and on adult survival and recruitment rates (derived from counts and apparent sex ratios); the culling pressure for the GB population moslty occurred during the prebreeding period, whereas it mostly focused during and after the breeding period for the FR population (see \@ref(HARV) for details); adult survival and recruitment rates under high culling pressure were averaged over a time-period observing a consistent culling pressure; vertical bars show 95% confidence intervals
+(ref:mgr) Average effects of culling pressure and different culling strategies (GB versus FR) on population growth rates (derived from counts only), and on adult survival and recruitment rates (derived from counts and reports of apparent sex ratios); the culling pressure for the GB population occurred mainly during the pre-breeding season, while for the FR population it was mainly concentrated during and after the breeding season (see \@ref(HARV) for details); adult survival and recruitment rates under high culling pressure were averaged over a period in which consistent culling pressure was observed; the vertical bars show 95% confidence intervals
 
-(ref:sup) Definition of three periods of different culling pressures from an index based on an estimation of adult culling rate; as only half of the culled individuals were aged in France before 2009, we applied the average age ratio over 2009-2019 on unaged individuals to provide an estimate of the number of adults in cull data before 2009; culling rate increased over time, we therefore split the time series in three categories : we defined a "no culling" period before 1999 because the culling rate for both countries was mostly null and always below 10%, we defined a "low culling" between 1999 and 2005 because the culling rate varied around 20% for both countries, the culling rate then raised significantly for both countries, so we defined a "high culling" period from 2006 onwards
+(ref:sup) Definition of three periods with different culling pressure using an index based on an estimate of the adult culling rate; as only half of the culled individuals in France were age-determined before 2009, we applied the average age ratio over 2009-2019 to the remaining individuals in order to obtain an estimate of the number of adults in the culling data before 2009; the culling rate increased over time, so we divided the time series into three categories: we defined a "no culling" period before 1999 as the culling rate for both countries was mostly zero and always below 10%, we defined a "low culling" period between 1999 and 2005 as the culling rate for both countries fluctuated around 20%, the culling rate then increased significantly for both countries, therefore we defined a "high culling" period from 2006 onwards

### 2024-05-28 Code/Model_vital_rates.R

@@ -84,7 +84,7 @@ cor_plot <- function(col, title = NULL, subtitle = NULL, limits = NA, breaks = N
                   width = 0, alpha = .5, linewidth = 1.2) +
     labs(title = title,
          subtitle = subtitle,
-         y = "count-based method", x = "culling-based method") +
+         y = '"apparent sex ratio" method', x = "culling-based method") +
     scale_x_continuous(expand = expansion(mult = c(0, 0)), 
                        limits = if (any(is.na(limits))) {
                          c(0, 1.1 * max(unlist(tmp[-1])))
@@ -1131,7 +1131,7 @@ JuvOut4 %>%
                         ylim = c(-0.05, 1.05),
                         expand = FALSE) + 
         guides(fill = "none",
-               alpha = guide_legend(title = "Control pressure", order = 1)) +
+               alpha = guide_legend(title = "Culling pressure", order = 1)) +
         scale_y_continuous(
           labels = scales::percent_format(accuracy = 1L),
           breaks = 0:10 / 10, 

### 2024-05-28 Code/Draft_ruddy_duck_vital_rates.Rmd
minor correction

@@
 | *Parameter (output)*  |                                                                                                            |
-| $p(m|a)$              | Proportion of males in adults (or probability of being a male knowing it is an adult)                      |
+| $p(m|a)$              | Proportion of males among adults (or probability of being a male knowing it is an adult)                      |
 | $p(a \cap m)_{i, t}$  | Proportion of adult males in population $i$ in year $t$                                                    |

### 2024-05-28 Code/Draft_ruddy_duck_vital_rates.Rmd
keywords modif

@@
 
-  **Keywords**: intrinsic growth rate - vital rate - reproductive success - fecundity - sex ratio - age ratio - Oxyura jamaicensis - harvesting
+  **Keywords**: age ratio - alien species - demography - fecundity - harvesting - intrinsic growth rate - Oxyura jamaicensis - reproductive success - sex ratio - vital rate
 output: 

### 2024-05-28 Code/Model_vital_rates.R

@@ -268,7 +268,7 @@ count_sex_app %>%
   # labs(title = "Male proportion", caption = "data from GB & French Ruddy duck populations") +
   coord_flip() -> intro_plot; intro_plot
 
-# Observed harvest rate ----------------------------------------------------------
+# Observed culling rate ----------------------------------------------------------
 
 # estimation of lambda_max to test recruitment rate reliability
 # when ruddy ducks start to be permanently seen and growing?
@@ -363,7 +363,7 @@ frag %>%
   ungroup() %>% 
   filter(shot > 0) -> frag_uncor
 
-## harv_plot ------------------------------------------------------------------
+## cull_plot ------------------------------------------------------------------
 
 counts_1 %>% 
   filter(year(year) <= 2017) %>% 

### 2024-05-29 Code/Draft_ruddy_duck_vital_rates.Rmd
keywords modif

@@
 
-# Materials & methods {#MM}
+# Materials & methods
 
@@
 
-## Validating the "apparent sex ratio" method {#VAL}
+## Validating the "apparent sex ratio" method
 
@@
 The models were fitted using NIMBLE [@Valpine2017] run from R [@RCT2022].
-The values **X[Y; Z]** reported in Section \@ref(RES) are the medians and the corresponding limits of the 95% confidence interval of the posterior distributions.
+The values **X[Y; Z]** reported in Results section are the medians and the corresponding limits of the 95% confidence interval of the posterior distributions.
 We preferred the median to the mean because it is more robust to skewed distributions.
 
-# Results {#RES}
+# Results
 
@@
 All estimates were above 0, which is not inconsistent with reality.
-Furthermore, the maximum recruitment rate was 0.68 [0.36; 0.78] (see Section \@ref(MAX)), and no estimate was significantly outside the range of recruitment rates defined without immigration [0; 0.78].
+Furthermore, the maximum recruitment rate was 0.68 [0.36; 0.78] (see 3rd Results section), and no estimate was significantly outside the range of recruitment rates defined without immigration [0; 0.78].
 No trend was observed for the GB population, but the recruitment rate decreased for the FR population, although it was more noisy than the proportion of immatures.
@@
 
-## Assessing the demographic response to culling strategies {#MAX}
+## Assessing the demographic response to culling strategies
 
@@
 High culling pressure occurred from 2006 onwards in both populations (although the signal was noisy).
-However, despite the similar level of culling pressure, the culling strategies of the two populations differed considerably, see Section \@ref(HARV).
+However, despite the similar level of culling pressure, the culling strategies of the two populations differed considerably, see 3rd Materials & Methods section.
 Under high culling pressure, the growth rate decreased to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for GB and FR, corresponding to a decrease of 44% per year and 3% per year, respectively.

### 2024-05-29 Code/Draft_ruddy_duck_vital_rates.Rmd
Section references update

@@
 ```
-## Assessing the effects of culling strategies {#HARV}
+## Assessing the effects of culling strategies
 
@@
 In Great Britain, most individuals were culled in winter (53.4% of adults were culled before 30 May), whereas in France, most individuals were culled during the breeding season (81.6% of adults were culled after 30 May).
-We investigated whether these strategies had different effects on the populations by comparing the resulting growth rates and the relative contributions of both vital rates to them (by comparing the average values during the period of high culling pressure with proxies for the maximum vital rates estimated when both populations reached their maximum growth, see Section \@ref(VAL)).
+We investigated whether these strategies had different effects on the populations by comparing the resulting growth rates and the relative contributions of both vital rates to them (by comparing the average values during the period of high culling pressure with proxies for the maximum vital rates estimated when both populations reached their maximum growth, see previous section).
 In France, a LIFE project was implemented from 2019 to intensify the culling pressure, especially in winter and spring.
@@
 
-We used the Bayesian framework to implement all submodels presented in Section \@ref(MM).
+We used the Bayesian framework to implement all submodels presented in Materials & methods section.
 The Bayesian framework is both straightforward and efficient in propagating error through the parameters.

### 2024-05-29 Code/Model_vital_rates.R
layout fix

@@ -202,12 +202,10 @@ counts %>%
   mutate(year = year + years(1),
          killed_before_rep = killed_before_rep %>% replace_na(0),
          n_breed = count) %>% 
-  #          n_breed = count - killed_before_rep) %>%  # if reference is the breeding time
   select(year, pop, n_breed) -> count_breed
 
 counts %>% 
   mutate(n_pop = count) %>% 
-  #   mutate(n_pop = count - killed_before_rep) %>% 
   select(year, n_win = count, n_pop, pop) %>% 
   left_join(count_breed) %>% 
   left_join(count_sex_app) %>% 
@@ -265,7 +263,6 @@ count_sex_app %>%
   scale_y_continuous(name = "", limits = c(0, 1), breaks = 0:10/10, labels = scales::percent) +
   guides(size = guide_legend(title = "Sample size")) +
   xlab("") + 
-  # labs(title = "Male proportion", caption = "data from GB & French Ruddy duck populations") +
   coord_flip() -> intro_plot; intro_plot
 
 # Observed culling rate ----------------------------------------------------------

### 2024-06-04 Code/Draft_ruddy_duck_vital_rates.Rmd
layout fix
Revision before submission to bioarxiv

@@
 lang: en-EN
-bibliography: "`r paste0('C:/Users/', Sys.info()[['user']], '/Dropbox/JabRef/Library_OFB.bib')`"
+bibliography: "`r here::here('Data', 'vital_rates.bib')`"
 csl: https://www.zotero.org/styles/apa # https://osf.io/download/6hw3q/ 
@@
 abstract: |
-  Monitoring the number of individuals is by far the most popular strategy for studying the environmental factors that determine population dynamics and for measuring the effectiveness of management actions aimed at population recovery, control, or eradication.
+  | Monitoring the number of individuals is by far the most popular strategy for studying the environmental factors that determine population dynamics and for measuring the effectiveness of management actions aimed at population recovery, control, or eradication.
 unfortunately, monitoring numbers is inefficient in identifying the mechanisms underlying demographic processes and, in particular, in assessing the extent to which population growth rate is influenced by changes in adult survival rather than variations in reproductive parameters.
-The usual methods for determining adult survival and productivity, such as analysing capture-mark-recapture data and game-hunting bag data, rely on capturing animals, which has three major drawbacks.
+The usual method for determining adult survival and productivity relies on capturing and marking, which has three major drawbacks.
 Firstly, the selectivity of capture methods can lead to a distorted representation of the underlying population structure.
@@
 Thirdly, release may simply be impossible, especially for invasive alien species that represent a threat to native taxa.
-In many waterfowl species, sexual dichromatism is observed in adults, while immatures of both sexes display a cover-up plumage similar to that of adult females.
-Using two non-native populations of Ruddy duck, a species that exhibits this pattern, we present a non-invasive method to estimate the respective contribution of adult survival and productivity (also called recruitment rate) to the population growth rate.
+In many waterfowl species, sexual dichromatism is observed in adults, while immatures of both sexes display a plumage pattern similar to that of adult females.
+Using two non-native populations of Ruddy duck, a species that exhibits this plumage pattern, we present a non-invasive method to estimate the respective contribution of adult survival and productivity (also called recruitment rate) to the population growth rate.
 We exemplify how changes in the apparent proportion of males in counts (due to female-like immature males gradually acquiring adult male plumage) can be exploited to split the population growth rate into adult survival and recruitment rates, which contributes significantly to understanding the demographic impact of two different eradication strategies.
-
-  **Keywords**: age ratio - alien species - demography - fecundity - harvesting - intrinsic growth rate - Oxyura jamaicensis - reproductive success - sex ratio - vital rate
+  | 
+  | **Keywords**: age ratio - alien species - demography - fecundity - harvesting - intrinsic growth rate - Oxyura jamaicensis - reproductive success - sex ratio - vital rate
+linestretch: 2
 output: 
@@
 @Niel2005].
-It is also crucial for measuring the effectiveness of management actions that are taken to recover, control, or eradicate populations [@Shea1998].
+It is also crucial for measuring the effectiveness of management actions that are taken to restore, control, or eradicate populations [@Shea1998].
 Among the approaches available to managers to achieve these goals, those that rely on monitoring the number of individuals (counts) are by far the most popular [@Rintala2022].
@@
 Unfortunately, relying solely on the monitoring of population abundance generally hinders the fundamental understanding of the demographic mechanisms underlying changes in population growth rates.
-More specifically, on the basis of counts alone, it is almost impossible to assess the extent to which population growth rate is affected by changes in adult survival rather than to variations in productivity [@Austin2000], which is defined as the product of all reproductive parameters.
+More specifically, on the basis of counts alone, it is almost impossible to assess the extent to which population growth rate is affected by changes in adult survival rather than to variations in young production [@Austin2000].
 To assess the relative sensitivity of population growth rate to factors affecting adult survival or reproductive parameters, demographers usually rely on the monitoring of individuals by capture-mark-recapture [@Lebreton1992].
@@
 the proportion of breeders that have survived for one year, and productivity/recruitment rate, i.e.
-the number of immatures produced per breeder that reach sexual maturity, which are, by definition, recruited in the breeding population.
+the number of immatures produced per breeder that reach sexual maturity, which are, by definition, recruited in the breeding population [@Smith2001].
 This additional effort makes it possible to assess the relative influence of each of these components on the population growth rate.
@@
 
-Here we exemplify how, in dimorphic species, delayed sexual maturity of males can be used to estimate adult survival and recruitment rates by distinguishing male-like and female-like individuals in repeated winter counts.
+Here we exemplify how, in dimorphic species, delayed sexual maturity of males can be used to estimate adult survival and recruitment rates by distinguishing male-like and female-like individuals in winter counts.
 We used two non-native European populations of Ruddy duck (*Oxyura jamaicensis*) as a study model.
 Like many other duck species, Ruddy duck is dimorphic: newborn males look like females until the prenuptial moult, which takes place in late winter.
-As a result of the late prenuptial moult of immatures, the apparent proportion of males increases over the course of the wintering season.
-These changes in the apparent proportion of males during this period are therefore directly related to the proportion of immatures in the populations and thus to the reproductive success of the previous breeding season.
+As a result of the late prenuptial moult of immatures, the apparent proportions of males increases over the course of the wintering season.
+These changes in the apparent proportions of males during this period are therefore directly related to the proportion of immatures in the populations and thus to the reproductive success of the previous breeding season.
 The "apparent sex ratio" method developed in this study was used to estimate adult survival and recruitment rates and evaluate the effects of two different eradication strategies used in Great Britain and France, respectively.
@@
 The trends and status of the White-headed duck and Ruddy duck populations used to be determined exclusively by counts.
-It was therefore not possible to assess the factors impairing the recovery of the former and the effectiveness of the eradication programme for the latter.
+It was therefore difficult to assess the factors impairing the recovery of the former and the effectiveness of the eradication programme for the latter.
 In particular, counts alone cannot be used to assess the relative impact of changes in adult survival and recruitment rates on population growth rates, which is a prerequisite for identifying limiting factors.
 
+\setstretch{1}
 ```{r count, fig.cap = "(ref:count)"}
@@
 ```
+\setstretch{2}
 
@@
 
-Interestingly, delayed dichromatism allows the estimation of adult survival and recruitment/productivity by observing the seasonal evolution of the apparent sex ratio and making few and reasonable assumptions (see next section).
+Interestingly, delayed dichromatism allows the estimation of adult survival and recruitment/productivity by monitoring the seasonal evolution of the apparent sex ratio and making few and reasonable assumptions (see next section).
 
@@
 
+\setstretch{1}
 ```{r flock, fig.cap = "(ref:flock)"}
@@
 In both countries, culling was carried out under strict official control so that the exact number of individuals culled was known.
-Although the age ratio of birds culled before 2009 could not be determined with certainty in France, it was still possible to roughly categorise the culling pressure in both countries into three categories: "no culling" before 1999, "low culling" between 1999 and 2005 and "high culling" from 2006 onwards (see Figure \@ref(fig:sup) in Supplementary Material).
+Although the age ratio of birds culled before 2009 could not be determined with certainty in France, it was still possible to roughly categorise the culling pressure in both countries into three categories: "no culling" before 1999, "low culling" between 1999 and 2005 and "high culling" from 2006 onwards (see Figure \@ref(fig:sup) in Appendix section).
 
@@
 
-The proportion of immature males is the difference between the proportion of female-like individuals between winter and spring.
-The proportion of all immature individuals is twice as high if we assume an even sex ratio at birth [@Bellrose1961; @Blums1996] and comparable mortality rates between immature females and immature males [@Ramula2018; @Wood2021].
-However, spring counts suffer from intractable problems, such as the different detectability of females and males during this season.
-Fortunately, the proportion of immatures can also be estimated from winter counts alone, assuming a known and unchanged adult sex ratio along the monitoring period [@Devineau2010].
-
-The adult sex ratio is therefore a prerequisite for estimating the proportion of immatures.
-This ratio could be derived from culling data as we determined the sex of each adult culled as part of the eradication programmes in both Great Britain and France.
-Nevertheless, the low number of individuals culled in France prevented from getting robust adult sex ratio.
+Our method exploits the difference between the sex ratio estimated from counts before young males acquire their male-like plumage pattern (and are therefore mistaken for females) and the sex ratio among adults (which is stable over time), which can be estimated from either spring or summer counts or from hunting bags.
+In the present study, we estimated the adult sex-ratio from culled individuals using the presence/absence of the bursa of Fabricius as a criterion for distinguishing adults from immatures (see hereafter).
+In France, it was not possible to obtain sufficiently reliable estimates of the sex ratio due to the small sample size.
 In Great Britain, adult sex ratios did not differ from month to month, so data from culled adults collected throughout the whole biological cycle were pooled together each year.
@@
 
+\setstretch{1}
 | Variable name         | Description                                                                                                |
@@
 | $s_{i, t}$            | Adult survival rate, i.e. proportion of individuals in year $t-1$ still alive in year $t$                  |
-| $r_{i, t}$            | Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per individual in year $t - 1$     |
+| $r_{i, t}$            | Recruitment rate, i.e. number of recruits in population $i$ in year $t$ per adult in year $t - 1$     |
 | $S_{i, t}$            | Number of adults in population $i$ in year $t$                                                             |
@@
 nesting rate, nesting success, clutch-size, hatching success, pre-fledging survival, and fledging survival.
-These reproductive parameters may be known for some duck species, but the last one is never mentioned [e.g.
+Although some of these reproductive parameters may be known for some duck species, the last one is never mentioned [e.g.
 @Baldassarre2014].
@@
 
+\setstretch{1}
 | Variable name | Description                                                                                 |
@@
 ```
+\setstretch{2}
+
 The maximum adult survival rate was defined as a uniform distribution on the interval $[0.7;1]$.
-The lower limit corresponded to the upper range of survival rates observed in waterfowl species of similar weight, and the upper limit was set to 1, as survival rates of long-lived waterfowl species are very high [@Nichols1997; @Krementz1997; @Buxton2004; @Nichols1992; @Koons2014].
+The lower limit corresponded to the upper range of survival rates observed in waterfowl species of similar weight, and the upper limit was set to 1, as survival rates of long-lived waterfowl species can be very high [@Nichols1997; @Krementz1997; @Buxton2004; @Nichols1992; @Koons2014].
 
@@
 After reaching the threshold of 1,000 individuals, the population growth in Great Britain (GB) showed a strong inflexion, although culling had not yet started (Figure \@ref(fig:count)).
-This observation suggests that a negative density-dependent process could occur beyond 1,000 individuals and led us to consider only the first sequence of the time series to infer the maximum growth rate in Great Britain, i.e.
+This observation suggests that beyond 1,000 individuals, a negative density-dependent process probably took place and led us to consider only the first sequence of the time series to infer the maximum growth rate in Great Britain, i.e.
 1972-1981.
@@
 
+\setstretch{1}
 | Variable name    | Class     | Description                                                                                              |
@@
 ```
+\setstretch{2}
+
 Estimating vital rates within an acceptable range, as defined above, would not prove that our modelling approach correctly reflects the actual interannual variability.
@@
 
+\setstretch{1}
 | Variable name | Class        | Description                                                              |
@@
 ```
+\setstretch{2}
+
 ## Assessing the effects of culling strategies
@@
 We checked the chain convergence using the Gelman and Rubin convergence diagnostic (R\<1.1, @Gelman1992).
-The models were fitted using NIMBLE [@Valpine2017] run from R [@RCT2022].
+The models were fitted using NIMBLE [@deValpine2017] run from R [@RCT2022].
 The values **X[Y; Z]** reported in Results section are the medians and the corresponding limits of the 95% confidence interval of the posterior distributions.
@@
 
+\setstretch{1}
 ```{r vital, fig.cap = "(ref:vital)"}
@@
 This result therefore supports the hypothesis that the proportion of female-like individuals in the overwintering populations is efficient for the assessment of age structure and thus productivity.
-The correlation was even stronger when looking at vital rates (Figure \@ref(fig:cor)), but this was to be expected because the two methods for estimating these vital rates had a common component, namely the population growth rate (see Equation \@ref(eq:vr)).
+The correlation was even stronger when looking at vital rates (Figure \@ref(fig:cor)), but this was expected because the two methods for estimating these vital rates had a common component, namely the population growth rate (see Equation \@ref(eq:vr)).
 Interestingly, a 1:1 correlation was not achieved for any of the parameters.
@@
 
+\setstretch{1}
 ```{r cor, fig.cap = "(ref:cor)"}
@@
 When there was no culling pressure, maximum growth rates were very similar for the two populations (Figure \@ref(fig:mgr)), namely 1.45 [1.36; 1.55] and 1.52 [1.25; 1.88] for GB and FR, respectively.
-These values corresponded to an increase in the population of 45% and 52% per year, respectively.
+These values corresponded to population increases of 45% and 52% per year, respectively.
 Assuming a maximum adult survival rate of 0.85 [0.7; 1] for both populations, we derived maximum recruitment rates of 0.60 [0.42; 0.78] and 0.68 [0.36; 1.06] for GB and FR, respectively.
@@
 High culling pressure occurred from 2006 onwards in both populations (although the signal was noisy).
-However, despite the similar level of culling pressure, the culling strategies of the two populations differed considerably, see 3rd Materials & Methods section.
+However, despite the similar level of culling pressure, the culling strategies of the two populations and their outcomes differed considerably, see 3rd Materials & Methods section.
 Under high culling pressure, the growth rate decreased to 0.56 [0.53; 0.59] and 0.97 [0.94; 1.00] for GB and FR, corresponding to a decrease of 44% per year and 3% per year, respectively.
-This result shows that the GB population declined significantly during the high culling pressure, while the FR population was just stabilised.
+The GB population declined significantly during the high culling pressure, while the FR population was just stabilised.
 
@@
 
+\setstretch{1}
 ```{r mgr, fig.cap = "(ref:mgr)"}
@@
 For both populations, vital rate estimates were within the same range, showing some consistency in the results obtained using the "apparent sex ratio" method.
-The greater variability observed for the FR population was to be expected as the time series covered a wider range of culling pressure than that of the GB population.
+The greater variability observed for the FR population was expected as the time series covered a wider range of culling pressure than that of the GB population.
 We found no temporal correlation of demographic parameters between the two populations, suggesting that regional (e.g.
@@
 Nevertheless, as the survival estimates are in the range of those reported for similar species [@Nichols1997; @Krementz1997; @Buxton2004; @Buderman2023], this assumption appears to be reliable.
-Therefore, the time window for the count surveys must be carefully chosen to take into account the time before the immature males start moulting.
+Nevertheless, the time window for the count surveys must be carefully chosen to take into account the time before the immature males start moulting.
 
@@
 This trait is expected to be similar in all Ruddy duck populations around the world.
-In our study, we observed that the maximum growth rates of the two populations were very close to each other, which was to be expected as both populations were exposed to similar ecological conditions, i.e.
+In our study, we observed that the maximum growth rates of the two populations were very close to each other, which was expected as both populations were exposed to similar ecological conditions, i.e.
 no harvest and similar breeding conditions.
@@
 As Ruddy ducks generally breed later than Common pochards and Tufted ducks, both a higher survival rate and greater nesting success could be responsible for the high growth rates of Ruddy duck populations in the absence of culling.
-The "apparent sex ratio" method not only provided estimates of adult survival and recruitment rates that were consistent with observed increasing population sizes, but also provided estimates that were consistent with both declining phases following the implementation of eradication programmes and observed differences in population growth rates under different culling strategies.
+The "apparent sex ratio" method provided estimates of adult survival and recruitment rates that were not only consistent with observed increasing population sizes, but also compatible with both declining phases following the implementation of eradication programmes and observed differences in population growth rates under different culling strategies.
 As expected, higher values for survival and recruitment rates were observed when the culling strategy focussed on the post-breeding season rather than the pre-breeding season.
@@
 As no estimates of adult survival were available in the literature [see @Buderman2023], we were unable to properly assess the accuracy of the "apparent sex ratio" method for estimating this parameter.
-Nevertheless, there is evidence that the method is not highly biassed, as described above.
+Nevertheless, there is evidence that the method is not highly biassed, as explained above.
 Thus, the "apparent sex ratio" method would be much more efficient for monitoring key demographic parameters than alternative methods such as those based on culling/hunting bag data, which have been shown to be unreliable [@Fox2014], even when focussing on the end of the hunting season to limit potential bias [@Fox2016].
@@
 This leads to fluctuations in the recruitment rate and thus jeopardises the long term efficiency of this strategy.
-This problem does not exist with the pre-breeding strategy, as the Ruddy ducks congregate at a limited number of sites in winter [@Johnsgard1996].
+This problem does not exist with the pre-breeding strategy, as the Ruddy ducks form large flocks that occupy a limited number of sites in winter [@Johnsgard1996].
 
@@
 Unexpectedly, our results suggest that harvesting waterfowl during the breeding season has far less impact on population growth rate than harvesting in winter.
-Consequently, a target for effective management of harvested waterfowl could be an earlier start and end of the hunting season.
+Consequently, a target for effective management of harvested waterfowl could be an earlier starting and ending of the hunting season.
 
@@
 
+\setstretch{1}
+
+# Appendix {.unnumbered}
+
+```{r sup, fig.cap = "(ref:sup)"}
+knitr::include_graphics("../Output/plot_paper_6.png")
+```
+
+\newpage
+
+\nolinenumbers
+
 # Acknowledgments {.unnumbered}
 
-This work was partly funded by the LIFE Oxyura project (LIFE17 NAT/FR/000942) under the LIFE programme.
 We warmly thank Jay McGowan for his permission to publish his photograph of Ruddy duck.
@@
 
-\nolinenumbers
+# Funding {.unnumbered}
+
+This work was partly funded by the LIFE Oxyura project (LIFE17 NAT/FR/000942) under the LIFE programme.
+
+# Conflict of interest disclosure {.unnumbered}
+
+The authors declare that they comply with the PCI rule of having no financial conflicts of interest in relation to the content of the article.
+
+# Data, scripts, code, and supplementary information availability  {.unnumbered}
+
+Data and scripts are available online: https://doi.org/10.5281/zenodo.11471723; Tableau et al., 2024
+
+\newpage
 
@@
 :::
-
-\newpage
-
-# (APPENDIX) sup {.unnumbered}
-
-# Supplementary Materials
-
-```{r sup, fig.cap = "(ref:sup)"}
-knitr::include_graphics("../Output/plot_paper_6.png")
-```
 
@@
 
-(ref:mgr) Average effects of culling pressure and different culling strategies (GB versus FR) on population growth rates (derived from counts only), and on adult survival and recruitment rates (derived from counts and reports of apparent sex ratios); the culling pressure for the GB population occurred mainly during the pre-breeding season, while for the FR population it was mainly concentrated during and after the breeding season (see \@ref(HARV) for details); adult survival and recruitment rates under high culling pressure were averaged over a period in which consistent culling pressure was observed; the vertical bars show 95% confidence intervals
+(ref:mgr) Average effects of culling pressure and different culling strategies (GB versus FR) on population growth rates (derived from counts only), and on adult survival and recruitment rates (derived from counts and reports of apparent sex ratios); the culling pressure for the GB population occurred mainly during the pre-breeding season, while for the FR population it was mainly concentrated during and after the breeding season; adult survival and recruitment rates under high culling pressure were averaged over a period in which consistent culling pressure was observed; the vertical bars show 95% confidence intervals
 

### 2024-06-04 Code/Model_vital_rates.R

@@ -915,7 +915,8 @@ xy %>%
     if(x$par == "p_juv") {
       tmp = tmp + 
         theme(
-          legend.position = c(.05, .95),
+          legend.position = "inside",
+          legend.position.inside = c(.05, .95),
           legend.justification = c("left", "top")) + 
         scale_y_continuous(
           labels = scales::percent_format(accuracy = 1L),
@@ -1134,7 +1135,8 @@ JuvOut4 %>%
           breaks = 0:10 / 10, 
           minor_breaks = NULL) +
         theme(
-          legend.position = c(.05, .90),
+          legend.position = "inside",
+          legend.position.inside = c(.05, .90),
           legend.justification = c("left", "top"),
           legend.direction = "horizontal",
           legend.key.size = unit(0.5, "cm"),

### 2024-06-04 Code/Draft_ruddy_duck_vital_rates.Rmd
Data DOI change

@@
 \setstretch{1}
-```{r vital, fig.cap = "(ref:vital)"}
+```{r vital, out.height = "90%", fig.cap = "(ref:vital)"}
 knitr::include_graphics("../Output/plot_paper_3.png")
@@
 \setstretch{1}
-```{r cor, fig.cap = "(ref:cor)"}
+```{r cor, out.height = "90%", fig.cap = "(ref:cor)"}
 knitr::include_graphics("../Output/plot_paper_4.png")
@@
 
-\setstretch{1}
+\newpage
 
@@
 
+\setstretch{1}
 ```{r sup, fig.cap = "(ref:sup)"}
@@
 ```
+\setstretch{2}
 
 \newpage
-
-\nolinenumbers
 
